### 2023-09-23 ukb_mods.Rmd
basic setup

---
title: "UKB Models"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
date: '2023-09-23'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, plyr, shinySIR, patchwork, ggstatsplot, qqplotr, arrangements, purrr, parallel, rlang)
```

## Load Data

## EDA

### 2023-09-25 ukb_mods.Rmd
basic GMV model testing, plotting func

@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, plyr, shinySIR, patchwork, ggstatsplot, qqplotr, arrangements, purrr, parallel, rlang)
+pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
+
+#pull plotting functions from github
+plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
+Sys.setenv(GITHUB_PAT = "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") #if needed to access private BGDlab GAMLSS repo
+source_url(plot.func.url)
 ```
@@
 
+Load UKB Data from Jakob
+
+```{r}
+cn_data <- fread(file="data/lifespan_dataset_dkatlas_CNonly.csv")
+#table(cn_data$study)
+```
+
+Filter
+
+```{r}
+ukb_df <- cn_data %>%
+  dplyr::filter(study == "UKB") %>%
+  mutate(age_yrs = round(age_days/365.25),
+         log_age = log(age_days, base=10),
+         TBV=(sGMV+GMV+WMV+Ventricles),
+         fs_version = as.factor(fs_version),
+         sex = as.factor(sex),
+         sex_int = as.numeric(case_when(sex == "Male" ~ 0,
+                                        sex == "Female" ~ 1)),
+         sex.log_age = sex_int * log_age) %>%
+  na.omit() #1 subj has 1 missing val
+
+skim(ukb_df)
+n_unique(ukb_df$participant) == nrow(ukb_df) #no duplicate subj
+```
+
 ## EDA
+
+```{r}
+ggplot(data=ukb_df, aes(x=age_yrs)) +
+  geom_histogram(aes(fill=sex_char), alpha=0.6, position="identity")
+
+ggplot(data=ukb_df, aes(x=log_age, y=GMV)) +
+  geom_point(aes(color=sex_char), alpha=0.6)
+```
+
+## Basic Plot Function
+
+```{r}
+mean.tbv.sim <- function(df, s, ageRange){
+  tbv_list <- list()
+  for (i in ageRange){
+      tbv.val <- df %>%
+        dplyr::filter(sex == s) %>%
+        sort(log_age) %>%
+        dplyr::filter(log_age >= (i+0.01), log_age <= (i-0.01)) %>%
+        mean(TBV)
+      tbv_list <- append(tbv_list, tbv.val)
+  }
+  return(tbv_list)
+}
+
+sim.data.ukb <- function(df){
+  minAge <- min(df$log_age)
+  maxAge <- max(df$log_age)
+  ageRange <- seq(minAge, maxAge, 0.005)  # generate an age range with increments of 0.005
+
+  #sim data
+  dataToPredictM <- data.frame(log_age=ageRange,
+                               sex=c(rep(as.factor("Male"), length(ageRange))),
+                               fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
+                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
+                               sex.log_age=c(rep(0, length(ageRange))))
+  dataToPredictM$TBV <- mean.tbv.sim(df, "Male", ageRange)
+
+  dataToPredictF <- data.frame(log_age=ageRange,
+                               sex=c(rep(as.factor("Female"), length(ageRange))),
+                               fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
+                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
+                               sex.log_age=ageRange)
+  dataToPredictF$TBV <- mean.tbv.sim(df, "Female", ageRange)
+
+  # List of centiles for the fan plot
+  desiredCentiles <- c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+  # return
+  sim <- list(ageRange, dataToPredictM, dataToPredictF, desiredCentiles)
+  names(sim) <- c("ageRange", "dataToPredictM", "dataToPredictF", "desiredCentiles")
+  return(sim)
+}
+
+plot_singlestudy_centiles <- function(gamlssModel, phenotype, df, color_var)
+{
+  #use functions in GAMLSS repo plotting_functions.R to sim and predict
+  sim <- sim.data(df) #simulate data
+  pred <- centile_predict(gamlssModel, sim$dataToPredictM, sim$dataToPredictF, sim$ageRange, sim$desiredCentiles) #predict centiles
+
+  #extract vals
+  ages <- sim$ageRange
+  age_col <- df$log_age
+  male_peak_age <- pred$peak_age_M
+  female_peak_age <- pred$peak_age_F
+  unit_text <- "(log(years))"
+
+  #plot!
+  #sometimes df[,get()] works, sometimes not found...????
+  yvar <- df[[phenotype]]
+  color_col <- df[[color_var]]
+
+  sampleCentileFan <- ggplot() +
+    geom_point(aes(x=age_col, y=yvar, color=color_col), alpha=0.3) +
+    scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) + 
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[1]]), alpha=0.1, linetype="dashed") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[3]]), alpha=0.3, linetype="dashed") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[5]]), linetype="dashed", color="#40004BFF") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[7]]), alpha=0.3, linetype="dashed") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[9]]), alpha=0.1, linetype="dashed") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[1]]), alpha=0.1, linetype="longdash") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[3]]), alpha=0.3, linetype="longdash") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[5]]), linetype="longdash", color="#00441BFF") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[7]]), alpha=0.3, linetype="longdash") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[9]]), alpha=0.1, linetype="longdash") +
+    geom_point(aes(x=male_peak_age, y=pred$peak_M), color="#40004BFF", size=3) +
+    geom_point(aes(x=female_peak_age, y=pred$peak_F), color="#00441BFF", size=3) +
+    labs(title=deparse(substitute(gamlssModel))) +
+    theme(legend.title = element_blank())+
+    xlab(paste("Age at Scan", unit_text)) +
+    ylab(deparse(substitute(phenotype)))
+
+  print(sampleCentileFan)
+}
+```
+
+## Global Mod Fitting
+
+Try using `stepGAICALL.A()`
+
+```{r}
+m1 <- gamlss(formula = GMV ~ pb(log_age) + sex_char + fs_version + pb(log_age):sex_char,
+                       sigma.formula = ~  pb(log_age) + sex_char,
+                       nu.formula = ~ 1,
+                       control = gamlss.control(n.cyc = 200), 
+                       family = BCCG, data=ukb_df, trace = FALSE)
+
+m2 <- stepGAICAll.A(m1, scope=list(lower=~1, upper=~pb(log_age) + sex_char + fs_version + pb(log_age):sex_char), k=2) #AIC
+
+m2
+summary(m2)
+```
+
+seeing if things change when i hard-code a sex*age interaction term
+
+```{r}
+m3 <- gamlss(formula = GMV ~ pb(log_age) + sex + fs_version + pb(sex.log_age),
+                       sigma.formula = ~  pb(log_age) + sex,
+                       nu.formula = ~ 1,
+                       control = gamlss.control(n.cyc = 200), 
+                       family = BCCG, data=ukb_df, trace = FALSE)
+m4 <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(log_age) + sex + fs_version + pb(sex.log_age)), k=2) #AIC
+
+summary(m4)
+```
+
+Seems like stepGAICAll handles combined sex.log_age term better than an interaction
+Try adding TBV
+```{r}
+m3.tbv <- gamlss(formula = GMV ~ pb(log_age) + sex + fs_version + pb(sex.log_age) + pb(TBV),
+                       sigma.formula = ~  pb(log_age) + sex,
+                       nu.formula = ~ 1,
+                       control = gamlss.control(n.cyc = 200), 
+                       family = BCCG, data=ukb_df, trace = FALSE)
+m4.tbv <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(log_age) + sex + fs_version + pb(sex.log_age) + pb(TBV)), k=2) #AIC
+summary(m4.tbv)
+```
+
+```{r}
+plot_singlestudy_centiles(m4, "GMV", ukb_df, sex_char)
+
+```
+
+Let's look at diagnostics
+```{r}
+plot(m4)
+wp(m4, xvar=log_age, n.inter=4) #leptokurtosis (tails fo fitted distribution too light)
+dtop(m4)
+rqres.plot(m4)
+
+plot(m4.tbv)
+wp(m4.tbv, xvar=log_age, n.inter=4) #U-shape: fitted skewness too low
+dtop(m4.tbv)
+rqres.plot(m4.tbv)
+```

### 2023-09-25 ukb_mods.Rmd
fixed plots

@@
   for (i in ageRange){
-      tbv.val <- df %>%
+      tbv.df <- df %>%
         dplyr::filter(sex == s) %>%
-        sort(log_age) %>%
-        dplyr::filter(log_age >= (i+0.01), log_age <= (i-0.01)) %>%
-        mean(TBV)
+        arrange(log_age) %>%
+        dplyr::filter(log_age >= (i-0.005) & log_age <= (i+0.005)) %>%
+        summarize(m = mean(TBV, na.rm = TRUE))
+      tbv.val <-tbv.df$m
       tbv_list <- append(tbv_list, tbv.val)
   }
-  return(tbv_list)
+  return(unlist(tbv_list))
 }
@@
   #use functions in GAMLSS repo plotting_functions.R to sim and predict
-  sim <- sim.data(df) #simulate data
+  sim <- sim.data.ukb(df) #simulate data
   pred <- centile_predict(gamlssModel, sim$dataToPredictM, sim$dataToPredictF, sim$ageRange, sim$desiredCentiles) #predict centiles
@@
     scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) + 
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[1]]), alpha=0.1, linetype="dashed") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[3]]), alpha=0.3, linetype="dashed") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[5]]), linetype="dashed", color="#40004BFF") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[7]]), alpha=0.3, linetype="dashed") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[9]]), alpha=0.1, linetype="dashed") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[1]]), alpha=0.1, linetype="longdash") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[3]]), alpha=0.3, linetype="longdash") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[5]]), linetype="longdash", color="#00441BFF") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[7]]), alpha=0.3, linetype="longdash") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[9]]), alpha=0.1, linetype="longdash") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[1]]), alpha=0.1) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[2]]), alpha=0.3) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[3]]), color="#40004BFF") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[4]]), alpha=0.3) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[5]]), alpha=0.1) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[1]]), alpha=0.1) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[2]]), alpha=0.3) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[3]]), color="#00441BFF") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[4]]), alpha=0.3) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[5]]), alpha=0.1) +
     geom_point(aes(x=male_peak_age, y=pred$peak_M), color="#40004BFF", size=3) +
@@
 ```{r}
-plot_singlestudy_centiles(m4, "GMV", ukb_df, sex_char)
+plot_singlestudy_centiles(m4, "GMV", ukb_df, "sex")
 

### 2023-09-28 ukb_mods.Rmd
cleaning up messy variables, switching to age_days

--- 2023-09-23 ukb_mods.Rmd
@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, plyr, shinySIR, patchwork, ggstatsplot, qqplotr, arrangements, purrr, parallel, rlang)
+pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
+
+# #pull plotting functions from github
+# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
+# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
+# source_url(plot.func.url) #sometimes doesn't work....?
+
+source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
 ```
@@
 
+Load UKB Data from Jakob
+
+```{r}
+cn_data <- fread(file="data/lifespan_dataset_dkatlas_CNonly.csv")
+#table(cn_data$study)
+```
+
+Filter
+
+```{r}
+ukb_df <- cn_data %>%
+  dplyr::filter(study == "UKB") %>%
+  mutate(age_yrs = round(age_days/365.25),
+         log_age = log(age_days, base=10),
+         TBV=(sGMV+GMV+WMV+Ventricles),
+         fs_version = as.factor(fs_version),
+         sex = as.factor(sex),
+         sex_int = as.numeric(case_when(sex == "Male" ~ 0,
+                                        sex == "Female" ~ 1)),
+         sex.age = sex_int * age_days) %>%
+  na.omit() #1 subj has 1 missing val
+
+skim(ukb_df)
+n_unique(ukb_df$participant) == nrow(ukb_df) #no duplicate subj
+
+#write data
+write_csv(ukb_df, file="data/ukb_CN_data.csv")
+```
+
 ## EDA
+
+```{r}
+ggplot(data=ukb_df, aes(x=age_days)) +
+  geom_histogram(aes(fill=sex), alpha=0.6, position="identity")
+ggplot(data=ukb_df, aes(x=log_age)) +
+  geom_histogram(aes(fill=sex), alpha=0.6, position="identity")
+
+ggplot(data=ukb_df, aes(x=log_age, y=GMV)) +
+  geom_point(aes(color=sex), alpha=0.6)
+ggplot(data=ukb_df, aes(x=age_days, y=GMV)) +
+  geom_point(aes(color=sex), alpha=0.6)
+```
+
+Going to stick with age_days since distribution is fairly normal, not lifespan data
+
+## Basic Plot Function
+
+```{r}
+mean.tbv.sim <- function(df, s, ageRange){
+  tbv_list <- list()
+  for (i in ageRange){
+      tbv.df <- df %>%
+        dplyr::filter(sex == s) %>%
+        arrange(age_yrs) %>%
+        dplyr::filter(age_yrs >= (i-0.005) & age_yrs <= (i+0.005)) %>%
+        summarize(m = mean(TBV, na.rm = TRUE))
+      tbv.val <-tbv.df$m
+      tbv_list <- append(tbv_list, tbv.val)
+  }
+  return(unlist(tbv_list))
+}
+
+sim.data.ukb <- function(df){
+  minAge <- min(df$age_days)
+  maxAge <- max(df$age_days)
+  ageRange <- seq(minAge, maxAge, 25)  # generate an age range with increments of 25 (since scaling in days)
+
+  #sim data
+  dataToPredictM <- data.frame(age_yrs=ageRange,
+                               sex=c(rep(as.factor("Male"), length(ageRange))),
+                               fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
+                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
+                               sex.age_yrs=c(rep(0, length(ageRange))))
+  dataToPredictM$TBV <- mean.tbv.sim(df, "Male", ageRange)
+
+  dataToPredictF <- data.frame(age_yrs=ageRange,
+                               sex=c(rep(as.factor("Female"), length(ageRange))),
+                               fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
+                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
+                               sex.age_yrs=ageRange)
+  dataToPredictF$TBV <- mean.tbv.sim(df, "Female", ageRange)
+
+  # List of centiles for the fan plot
+  desiredCentiles <- c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+  # return
+  sim <- list(ageRange, dataToPredictM, dataToPredictF, desiredCentiles)
+  names(sim) <- c("ageRange", "dataToPredictM", "dataToPredictF", "desiredCentiles")
+  return(sim)
+}
+
+plot_singlestudy_centiles <- function(gamlssModel, phenotype, df, color_var)
+{
+  #use functions in GAMLSS repo plotting_functions.R to sim and predict
+  sim <- sim.data.ukb(df) #simulate data
+  pred <- centile_predict(gamlssModel, sim$dataToPredictM, sim$dataToPredictF, sim$ageRange, sim$desiredCentiles) #predict centiles
+
+  #extract vals
+  ages <- sim$ageRange
+  age_col <- df$age_yrs
+  male_peak_age <- pred$peak_age_M
+  female_peak_age <- pred$peak_age_F
+  unit_text <- "(years)"
+
+  #plot!
+  #sometimes df[,get()] works, sometimes not found...????
+  yvar <- df[[phenotype]]
+  color_col <- df[[color_var]]
+
+  sampleCentileFan <- ggplot() +
+    geom_point(aes(x=age_col, y=yvar, color=color_col), alpha=0.3) +
+    scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) + 
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[1]]), alpha=0.1) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[2]]), alpha=0.3) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[3]]), color="#40004BFF") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[4]]), alpha=0.3) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_M[[5]]), alpha=0.1) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[1]]), alpha=0.1) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[2]]), alpha=0.3) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[3]]), color="#00441BFF") +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[4]]), alpha=0.3) +
+    geom_line(aes(x=ages, y=pred$fanCentiles_F[[5]]), alpha=0.1) +
+    geom_point(aes(x=male_peak_age, y=pred$peak_M), color="#40004BFF", size=3) +
+    geom_point(aes(x=female_peak_age, y=pred$peak_F), color="#00441BFF", size=3) +
+    labs(title=deparse(substitute(gamlssModel))) +
+    theme(legend.title = element_blank())+
+    xlab(paste("Age at Scan", unit_text)) +
+    ylab(deparse(substitute(phenotype)))
+
+  print(sampleCentileFan)
+}
+```
+
+## Global Mod Fitting
+
+Try using `stepGAICALL.A()`
+
+```{r}
+m1 <- gamlss(formula = GMV ~ pb(age_days) + sex + fs_version + pb(age_days):sex,
+                       sigma.formula = ~  pb(age_days) + sex,
+                       nu.formula = ~ 1,
+                       control = gamlss.control(n.cyc = 200), 
+                       family = BCCG, data=ukb_df, trace = FALSE)
+
+m2 <- stepGAICAll.A(m1, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(age_days):sex), k=2) #AIC
+
+m2
+summary(m2)
+```
+
+seeing if things change when i hard-code a sex*age interaction term
+
+```{r}
+m3 <- gamlss(formula = GMV ~ pb(age_days) + sex + fs_version + pb(sex.age),
+                       sigma.formula = ~  pb(age_days) + sex,
+                       nu.formula = ~ 1,
+                       control = gamlss.control(n.cyc = 200), 
+                       family = BCCG, data=ukb_df, trace = FALSE)
+m4 <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(sex.age)), k=2) #AIC
+
+summary(m4)
+```
+
+Seems like stepGAICAll handles combined sex.age term better than an interaction
+Try adding TBV
+```{r}
+m3.tbv <- gamlss(formula = GMV ~ pb(age_days) + sex + fs_version + pb(sex.age) + pb(TBV),
+                       sigma.formula = ~  pb(age_days) + sex,
+                       nu.formula = ~ 1,
+                       control = gamlss.control(n.cyc = 200), 
+                       family = BCCG, data=ukb_df, trace = FALSE)
+m4.tbv <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(sex.age) + pb(TBV)), k=2) #AIC
+summary(m4.tbv)
+```
+
+```{r}
+plot_singlestudy_centiles(m4, "GMV", ukb_df, "sex")
+```
+
+Let's look at diagnostics
+```{r}
+plot(m4)
+wp(m4, xvar=age_days, n.inter=4) #leptokurtosis (tails fo fitted distribution too light)
+dtop(m4)
+rqres.plot(m4)
+
+plot(m4.tbv)
+wp(m4.tbv, xvar=age_days, n.inter=4) #U-shape: fitted skewness too low
+dtop(m4.tbv)
+rqres.plot(m4.tbv)
+```

### 2023-09-29 ukb_mods.Rmd

@@
 write_csv(ukb_df, file="data/ukb_CN_data.csv")
+
+col_list <- colnames(ukb_df)
+pheno_list <- col_list[grepl("^lh_|^rh_|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
+writeLines(pheno_list, con = "phenotype_list.txt")
 ```
@@
                        family = BCCG, data=ukb_df, trace = FALSE)
-m4.tbv <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(sex.age) + pb(TBV)), k=2) #AIC
+m4.tbv <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(sex.age) + pb(TBV)), k=2, glim.control(cyc=100)) #AIC
 summary(m4.tbv)

### 2023-10-02 ukb_mods.Rmd
global CT and SA values for better global correction

@@
 n_unique(ukb_df$participant) == nrow(ukb_df) #no duplicate subj
+```
+
+Write lists of phenotypes
+
+```{r}
+col_list <- colnames(ukb_df)
+pheno_list <- col_list[grepl("^lh_|^rh_|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
+length(pheno_list) #total number of models to run
+
+#split by measure type (Vol, SA, CT)
+vol_list <- col_list[grepl("^lh_Vol|^rh_Vol|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
+writeLines(vol_list, con = "pheno_lists/Vol_list.txt")
+
+sa_list <- col_list[grepl("^lh_SA|^rh_SA", col_list)]
+writeLines(sa_list, con = "pheno_lists/SA_list.txt")
+
+ct_list <- col_list[grepl("^lh_CT|^rh_CT", col_list)]
+writeLines(ct_list, con = "pheno_lists/CT_list.txt")
+```
+
+Add brainwide phenotypes!
+Summing across regions for volumes (rather than summing GMV, sGMV, etc) for consistency with SA and CT
+
+```{r}
+vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
+
+ukb_df_totals <- ukb_df %>%
+  mutate(Vol_total=rowSums(select(., .dots=all_of(c(vol_list_regions)))),
+         SA_total=rowSums(select(., .dots=all_of(c(sa_list)))),
+         CT_total=rowSums(select(., .dots=all_of(c(ct_list)))))
 
 #write data
-write_csv(ukb_df, file="data/ukb_CN_data.csv")
-
-col_list <- colnames(ukb_df)
-pheno_list <- col_list[grepl("^lh_|^rh_|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
-writeLines(pheno_list, con = "phenotype_list.txt")
+write_csv(ukb_df_totals, file="data/ukb_CN_data.csv")
 ```
@@
 ```
+
+# CUBIC Models
+Playing around wiht models fit on cubic using `fit_ukb_mod.R`
+
+scp-ing random model to look at (`lh_SA_middletemporal_mod.rds`)
+
+```{r}
+lh_SA_middletemporal_mod <- readRDS("ukb/gamlss_objs/lh_SA_middletemporal_mod.rds")
+summary(lh_SA_middletemporal_mod)
+
+```

### 2023-10-03 ukb_mods.Rmd
code to find sex terms in fitted mods

@@
 #split by measure type (Vol, SA, CT)
-vol_list <- col_list[grepl("^lh_Vol|^rh_Vol|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
-writeLines(vol_list, con = "pheno_lists/Vol_list.txt")
+vol_list_global <- c("GMV", "sGMV", "WMV", "Ventricles")
+writeLines(vol_list_global, con = "pheno_lists/Vol_list_global.txt")
+
+vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
+writeLines(vol_list_regions, con = "pheno_lists/Vol_list_regions.txt")
 
@@
 Going to stick with age_days since distribution is fairly normal, not lifespan data
+
+```{r}
+ggplot(data=ukb_df_totals, aes(x=age_days, y=Vol_total)) +
+  geom_point(aes(color=sex), alpha=0.6, position="identity")
+
+ggplot(data=ukb_df_totals, aes(x=age_days, y=TBV)) +
+  geom_point(aes(color=sex), alpha=0.6, position="identity") #TBV includes ventricles, WMV, subcortical GMV
+
+ggplot(data=ukb_df_totals, aes(x=age_days, y=CT_total)) +
+  geom_point(aes(color=sex), alpha=0.6, position="identity")
+
+ggplot(data=ukb_df_totals, aes(x=age_days, y=SA_total)) +
+  geom_point(aes(color=sex), alpha=0.6, position="identity")
+```
 
@@
 
+formula(lh_SA_middletemporal_mod, what="sigma")
+coef(lh_SA_middletemporal_mod, what="sigma") %>%
+  names()
+
+sfit <- summary(lh_SA_middletemporal_mod)
+
+sfit$coef.table
+
+#to get a list of all the slots in the object
+str(sfit)
 ```

### 2023-10-03 R_scripts/gamlss_helper_funs.R

@@ -1,0 +1,20 @@
+#load libraries
+library(gamlss)
+library(dplyr)
+
+### CONTENTS ###
+# find.param()
+################
+
+################
+# find.param(gamlss.rds.file, moment, string)
+### Use: read gamlss objects from RDS file and see whether the formula for a specific moment includes vars containing specified string
+### Arguments: gamlss.rds.file = .rds containing gamlss obj;  moment = c("mu", "sigma", "nu", "tau"); string = search string
+
+find.param <- function(gamlss.rds.file, moment, string) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  gamlss.obj <- readRDS(gamlss.rds.file)
+  sigma.coeff.list <- coef(gamlss.obj, what = moment) %>%
+    names()
+  any(sapply(sigma.coeff.list, function(x) grepl(string, x, ignore.case = TRUE)))
+}

### 2023-10-03 R_scripts/variance_parser.R

@@ -1,0 +1,18 @@
+
+set.seed(12345)
+
+#LOAD PACKAGES
+library(gamlss)
+library(dplyr)
+library(data.table)
+
+source("gamlss_helper_funs.R")
+
+#GET ARGS
+args <- commandArgs(trailingOnly = TRUE)
+read_path <- args[1] #path to bfp models
+save_path <- args[2] #path to save diagnostic plots
+
+#iterate across gamlss objects
+model.files <- list.files(path = read_path, pattern = "mod.rds", full.names = TRUE)
+sigma.sex <- sapply(model.files, find.param, moment="sigma", string="sex", USE.NAMES=TRUE)

### 2023-10-05 ukb_mods.Rmd
more robust parsing of sigma moments

@@
 lh_SA_middletemporal_mod <- readRDS("ukb/gamlss_objs/lh_SA_middletemporal_mod.rds")
-summary(lh_SA_middletemporal_mod)
+a <- summary(lh_SA_middletemporal_mod)
+dim(a)
 
-formula(lh_SA_middletemporal_mod, what="sigma")
+#sigma coefficients
 coef(lh_SA_middletemporal_mod, what="sigma") %>%
@@
 
-sfit <- summary(lh_SA_middletemporal_mod)
+test <- confint(lh_SA_middletemporal_mod)
+colnames(test)
+rownames(test)
+test["sigma.sexMale", "97.5 %"]
+df.test <- as.data.frame(test)
+df.test$params <- rownames(df.test)
+# Remove the row names from the dataframe
+rownames(df.test) <- NULL
 
-sfit$coef.table
+df.test
+#using drop1 to get significance, gamlss book p 125 for df info
+#get df from OG model - not sure how this would be affected by using stepGAIC?
 
-#to get a list of all the slots in the object
-str(sfit)
+str(lh_SA_middletemporal_mod)
+lh_SA_middletemporal_mod$sigma.nl.df
+
+#start df with confidence intervals
+confint.matrix <- confint(lh_SA_middletemporal_mod)
+df <- as.data.frame(confint.matrix) %>%
+  rownames_to_column(var = "moment.param") %>%
+  mutate(
+    moment = sapply(strsplit(moment.param, "\\."), `[`, 1),
+    param = sub("^[^.]+\\.", "", moment.param)
+  )
+
+#now that we have the moments, add the rest of the summary info
+sum.table <- summary(lh_SA_middletemporal_mod)
+df2 <- as.data.frame(sum.table) %>%
+  rownames_to_column(var = "param")
+#check
+rownames(sum.table) == df$param
+
+find.param <- function(gamlss.rds.file, moment, string) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  gamlss.obj <- readRDS(gamlss.rds.file)
+  sigma.coeff.list <- coef(gamlss.rds.file, what= moment) %>%
+    names()
+  any(sapply(sigma.coeff.list, function(x) grepl(string, x, ignore.case = TRUE)))
+}
+find.param(lh_SA_middletemporal_mod, "sigma", "sex")
+
+readRDS("ukb/gamlss_objs/rh_Vol_parsopercularis_mod.rds")
+
 ```

### 2023-10-05 R_scripts/gamlss_helper_funs.R

@@ -18,3 +18,36 @@ find.param <- function(gamlss.rds.file, moment, string) {
     names()
   any(sapply(sigma.coeff.list, function(x) grepl(string, x, ignore.case = TRUE)))
 }
+
+get.beta <- function(gamlss.rds.file, moment, term) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  gamlss.obj <- readRDS(gamlss.rds.file)
+  beta <- coef(gamlss.obj, what = moment)[term] %>%
+    unname()
+  return(beta)
+}
+
+list.sigma.terms <- function(gamlss.rds.file) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  gamlss.obj <- readRDS(gamlss.rds.file)
+  terms <- attr(gamlss.obj$sigma.terms, "term.labels")
+  return(terms)
+}
+
+get.sigma.df <- function(gamlss.rds.file) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  gamlss.obj <- readRDS(gamlss.rds.file)
+  gamlss.obj$sigma.df
+}
+
+get.sigma.nl.df <- function(gamlss.rds.file) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  gamlss.obj <- readRDS(gamlss.rds.file)
+  gamlss.obj$sigma.nl.df
+}
+
+get.moment.formula <- function(gamlss.rds.file, moment) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  gamlss.obj <- readRDS(gamlss.rds.file)
+  moment.form <- formula(gamlss.obj, what = moment)
+}

### 2023-10-05 R_scripts/variance_parser.R

@@ -16,12 +16,24 @@ save_path <- args[2] #path to save outputs
 
 #iterate across gamlss objects
 model.files <- list.files(path = read_path, pattern = "mod.rds", full.names = TRUE)
-sigma.sex <- sapply(model.files, find.param, moment="sigma", string="sex", USE.NAMES=TRUE)
-#make dataframe
-sigma.sex.df <- sigma.sex %>%
-  as.data.frame() %>%
-  rename(contains_sex = ".") %>%
-  rownames_to_column(var = "file") %>%
-  mutate(mod_name = basename(file))
 
-write.csv(sigma.sex.df, file=paste0(save_path, "/sigma_sex_term.csv"))
+sigma.df <- data.frame("file_path" = model.files)
+sigma.df <- sigma.df %>%
+  mutate(region = sub("_mod\\.rds$", "", basename(file_path)))
+
+#look for sex-terms
+sigma.df$contains_sex <- lapply(sigma.df$file_path, find.param, moment="sigma", string="sex")
+
+#list all terms in sigma
+sigma.df$terms <- lapply(sigma.df$file_path, list.sigma.terms)
+sigma.df$sig_form <- lapply(sigma.df$file_path, get.moment.formula, moment="sigma")
+
+#get sexMale beta weight
+sigma.df$sexMale <- lapply(sigma.df$file_path, get.beta, moment="sigma", term="sexMale")
+
+#sigma dfs
+sigma.df$degrees <- lapply(sigma.df$file_path, get.sigma.df)
+sigma.df$nl_degrees <- lapply(sigma.df$file_path, get.sigma.nl.df)
+
+
+write.csv(sigma.df, file=paste0(save_path, "/sigma.csv"))

### 2023-10-06 ukb_mods.Rmd
region eda histograms

@@
 
-## Basic Plot Function
+Look at variances
+```{r}
+plot.hist.by.sex <- function(data, pheno_list) {
+
+  # Create an empty list to store the ggplot objects
+  plots <- list()
+
+  # Loop through the list of variables
+  for (pheno in pheno_list) {
+    # Check if the variable exists in the data frame
+    if (!pheno %in% names(data)) {
+      warning(paste("Variable '", pheno, "' not found in the data frame. Skipping..."))
+      next
+    }
+
+    # Create a ggplot histogram for the variable
+    plot <- ggplot(data, aes(x = .data[[pheno]], fill= sex, color=sex)) +
+      geom_histogram(alpha=0.5, position="identity") +
+      labs(title = paste("Histogram of", pheno), x = pheno, y = "Frequency")
+
+    # Add the plot to the list
+    plots[[pheno]] <- plot
+  }
+
+  # Return the list of ggplot objects
+  return(plots)
+}
+```
+
+center
+```{r}
+ukb_centered <- ukb_df_totals %>%
+  group_by(sex) %>%
+  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
+  ungroup()
+```
 
 ```{r}
-mean.tbv.sim <- function(df, s, ageRange){
-  tbv_list <- list()
-  for (i in ageRange){
-      tbv.df <- df %>%
-        dplyr::filter(sex == s) %>%
-        arrange(age_yrs) %>%
-        dplyr::filter(age_yrs >= (i-0.005) & age_yrs <= (i+0.005)) %>%
-        summarize(m = mean(TBV, na.rm = TRUE))
-      tbv.val <-tbv.df$m
-      tbv_list <- append(tbv_list, tbv.val)
+# Generate histograms (mean-centered w/in sex)
+global_plots <- plot.hist.by.sex(ukb_centered, vol_list_global)
+
+# Print or display the histograms
+for (pheno in vol_list_global) {
+  print(global_plots[[pheno]])
 }
-  return(unlist(tbv_list))
+```
+
+```{r}
+# Generate histograms (mean-centered w/in sex)
+region_vol_plots <- plot.hist.by.sex(ukb_centered, vol_list_regions)
+
+# Print or display the histograms
+for (pheno in vol_list_regions) {
+  print(region_vol_plots[[pheno]])
 }
+```
 
-sim.data.ukb <- function(df){
-  minAge <- min(df$age_days)
-  maxAge <- max(df$age_days)
-  ageRange <- seq(minAge, maxAge, 25)  # generate an age range with increments of 25 (since scaling in days)
+```{r}
+# Generate histograms (mean-centered w/in sex)
+region_sa_plots <- plot.hist.by.sex(ukb_centered, sa_list)
 
-  #sim data
-  dataToPredictM <- data.frame(age_yrs=ageRange,
-                               sex=c(rep(as.factor("Male"), length(ageRange))),
-                               fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
-                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
-                               sex.age_yrs=c(rep(0, length(ageRange))))
-  dataToPredictM$TBV <- mean.tbv.sim(df, "Male", ageRange)
+# Print or display the histograms
+for (pheno in sa_list) {
+  print(region_sa_plots[[pheno]])
+}
+```
 
-  dataToPredictF <- data.frame(age_yrs=ageRange,
-                               sex=c(rep(as.factor("Female"), length(ageRange))),
-                               fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
-                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
-                               sex.age_yrs=ageRange)
-  dataToPredictF$TBV <- mean.tbv.sim(df, "Female", ageRange)
+```{r}
+# Generate histograms (mean-centered w/in sex)
+region_ct_plots <- plot.hist.by.sex(ukb_centered, ct_list)
 
-  # List of centiles for the fan plot
-  desiredCentiles <- c(0.1, 0.25, 0.5, 0.75, 0.9)
-
-  # return
-  sim <- list(ageRange, dataToPredictM, dataToPredictF, desiredCentiles)
-  names(sim) <- c("ageRange", "dataToPredictM", "dataToPredictF", "desiredCentiles")
-  return(sim)
-}
-
-plot_singlestudy_centiles <- function(gamlssModel, phenotype, df, color_var)
-{
-  #use functions in GAMLSS repo plotting_functions.R to sim and predict
-  sim <- sim.data.ukb(df) #simulate data
-  pred <- centile_predict(gamlssModel, sim$dataToPredictM, sim$dataToPredictF, sim$ageRange, sim$desiredCentiles) #predict centiles
-
-  #extract vals
-  ages <- sim$ageRange
-  age_col <- df$age_yrs
-  male_peak_age <- pred$peak_age_M
-  female_peak_age <- pred$peak_age_F
-  unit_text <- "(years)"
-
-  #plot!
-  #sometimes df[,get()] works, sometimes not found...????
-  yvar <- df[[phenotype]]
-  color_col <- df[[color_var]]
-
-  sampleCentileFan <- ggplot() +
-    geom_point(aes(x=age_col, y=yvar, color=color_col), alpha=0.3) +
-    scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) + 
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[1]]), alpha=0.1) +
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[2]]), alpha=0.3) +
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[3]]), color="#40004BFF") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[4]]), alpha=0.3) +
-    geom_line(aes(x=ages, y=pred$fanCentiles_M[[5]]), alpha=0.1) +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[1]]), alpha=0.1) +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[2]]), alpha=0.3) +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[3]]), color="#00441BFF") +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[4]]), alpha=0.3) +
-    geom_line(aes(x=ages, y=pred$fanCentiles_F[[5]]), alpha=0.1) +
-    geom_point(aes(x=male_peak_age, y=pred$peak_M), color="#40004BFF", size=3) +
-    geom_point(aes(x=female_peak_age, y=pred$peak_F), color="#00441BFF", size=3) +
-    labs(title=deparse(substitute(gamlssModel))) +
-    theme(legend.title = element_blank())+
-    xlab(paste("Age at Scan", unit_text)) +
-    ylab(deparse(substitute(phenotype)))
-
-  print(sampleCentileFan)
+# Print or display the histograms
+for (pheno in ct_list) {
+  print(region_ct_plots[[pheno]])
 }

### 2023-10-06 ukb_mods.Rmd
all sigma betas

@@
                        family = BCCG, data=ukb_df, trace = FALSE)
-m4.tbv <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(sex.age) + pb(TBV)), k=2, glim.control(cyc=100)) #AIC
+m4.tbv <- stepGAICAll.A(Ventricles_mod, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(sex.age) + pb(TBV)), k=2, glim.control(cyc=100)) #AIC
 summary(m4.tbv)
@@
 ```
+
+### Sex-effects visualization
+
+load info parsed from models via  `var_parse.sh` (which calls `R_scripts/variance_parser.R`)
+
+```{r}
+sigma.df <- read.csv(file="ukb/sigma.csv", row.names=1)
+skim(sigma.df)
+```
+
+all contain sex in sigma
+
+```{r}
+sigma.df<- sigma.df %>%
+  mutate(mu_form = as.factor(sub(".*~", "", mu_form)),
+         sig_form = as.factor(sig_form),
+         nu_form = as.factor(nu_form),
+         measure_type = as.factor(case_when(grepl("Vol", region) ~ "Vol",
+                                  grepl("CT", region) ~ "CT",
+                                  grepl("SA", region) ~ "SA",
+                                  TRUE ~ "Global")))
+
+table(sigma.df$sig_form)
+table(sigma.df$mu_form)
+table(sigma.df$nu_form)
+```
+
+Weirdly only nu has variation in terms selected...?
+
+plot sex effects in sigma
+```{r}
+ggplot(sigma.df, aes(x=sexMale, fill=measure_type)) +
+  geom_histogram(alpha=0.6, position = "identity", color="white")
+```
+
+Normalize sexMale's beta weight by the overall value of all sigma terms
+
+```{r}
+#get betas
+file.list <- list.files(path="/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb/gamlss_objs", pattern="_mod.rds", full.names = TRUE)
+
+betas.mat <- sapply(file.list, get.moment.betas, moment="sigma", USE.NAMES=TRUE)
+colnames(betas.mat) <- map(colnames(betas.mat), basename)
+
+```

### 2023-10-06 R_scripts/gamlss_helper_funs.R

@@ -36,6 +36,15 @@ get.beta <- function(gamlss.rds.file, moment, term) {
   return(beta)
 }
 
+get.moment.betas <- function(gamlss.rds.file, moment) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  gamlss.rds.file <- as.character(gamlss.rds.file)
+  gamlss.obj <- readRDS(gamlss.rds.file)
+  beta.list <- coef(gamlss.obj, what = moment) %>%
+    as.list()
+  return(beta.list)
+}
+
 list.sigma.terms <- function(gamlss.rds.file) {
   #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
   gamlss.rds.file <- as.character(gamlss.rds.file)

### 2023-10-06 ukb_cubic_mods.Rmd
splitting out EDA and modeling into separate Rmds

--- 2023-10-06 ukb_mods.Rmd
@@
 ---
-title: "UKB Models"
-output: 
-  html_document:
-    code_folding: hide
-    toc: true
-    toc_float: true
-date: '2023-09-23'
+title: "UKB Cubic Mods"
+output: html_document
+date: '2023-10-06'
 ---
+
+Playing around with model-selection algorithms using data cleaned & curated in `ukb_curation-eda.Rmd`
 
@@
 
-Load UKB Data from Jakob
-
 ```{r}
-cn_data <- fread(file="data/lifespan_dataset_dkatlas_CNonly.csv")
-#table(cn_data$study)
-```
-
-Filter
-
-```{r}
-ukb_df <- cn_data %>%
-  dplyr::filter(study == "UKB") %>%
-  mutate(age_yrs = round(age_days/365.25),
-         log_age = log(age_days, base=10),
-         TBV=(sGMV+GMV+WMV+Ventricles),
-         fs_version = as.factor(fs_version),
-         sex = as.factor(sex),
-         sex_int = as.numeric(case_when(sex == "Male" ~ 0,
-                                        sex == "Female" ~ 1)),
-         sex.age = sex_int * age_days) %>%
-  na.omit() #1 subj has 1 missing val
-
-skim(ukb_df)
-n_unique(ukb_df$participant) == nrow(ukb_df) #no duplicate subj
-```
-
-Write lists of phenotypes
-
-```{r}
-col_list <- colnames(ukb_df)
-pheno_list <- col_list[grepl("^lh_|^rh_|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
-length(pheno_list) #total number of models to run
-
-#split by measure type (Vol, SA, CT)
-vol_list_global <- c("GMV", "sGMV", "WMV", "Ventricles")
-writeLines(vol_list_global, con = "pheno_lists/Vol_list_global.txt")
-
-vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
-writeLines(vol_list_regions, con = "pheno_lists/Vol_list_regions.txt")
-
-sa_list <- col_list[grepl("^lh_SA|^rh_SA", col_list)]
-writeLines(sa_list, con = "pheno_lists/SA_list.txt")
-
-ct_list <- col_list[grepl("^lh_CT|^rh_CT", col_list)]
-writeLines(ct_list, con = "pheno_lists/CT_list.txt")
-```
-
-Add brainwide phenotypes!
-Summing across regions for volumes (rather than summing GMV, sGMV, etc) for consistency with SA and CT
-
-```{r}
-vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
-
-ukb_df_totals <- ukb_df %>%
-  mutate(Vol_total=rowSums(select(., .dots=all_of(c(vol_list_regions)))),
-         SA_total=rowSums(select(., .dots=all_of(c(sa_list)))),
-         CT_total=rowSums(select(., .dots=all_of(c(ct_list)))))
-
-#write data
-write_csv(ukb_df_totals, file="data/ukb_CN_data.csv")
-```
-
-## EDA
-
-```{r}
-ggplot(data=ukb_df, aes(x=age_days)) +
-  geom_histogram(aes(fill=sex), alpha=0.6, position="identity")
-ggplot(data=ukb_df, aes(x=log_age)) +
-  geom_histogram(aes(fill=sex), alpha=0.6, position="identity")
-
-ggplot(data=ukb_df, aes(x=log_age, y=GMV)) +
-  geom_point(aes(color=sex), alpha=0.6)
-ggplot(data=ukb_df, aes(x=age_days, y=GMV)) +
-  geom_point(aes(color=sex), alpha=0.6)
-```
-
-Going to stick with age_days since distribution is fairly normal, not lifespan data
-
-```{r}
-ggplot(data=ukb_df_totals, aes(x=age_days, y=Vol_total)) +
-  geom_point(aes(color=sex), alpha=0.6, position="identity")
-
-ggplot(data=ukb_df_totals, aes(x=age_days, y=TBV)) +
-  geom_point(aes(color=sex), alpha=0.6, position="identity") #TBV includes ventricles, WMV, subcortical GMV
-
-ggplot(data=ukb_df_totals, aes(x=age_days, y=CT_total)) +
-  geom_point(aes(color=sex), alpha=0.6, position="identity")
-
-ggplot(data=ukb_df_totals, aes(x=age_days, y=SA_total)) +
-  geom_point(aes(color=sex), alpha=0.6, position="identity")
-```
-
-Look at variances
-```{r}
-plot.hist.by.sex <- function(data, pheno_list) {
-
-  # Create an empty list to store the ggplot objects
-  plots <- list()
-
-  # Loop through the list of variables
-  for (pheno in pheno_list) {
-    # Check if the variable exists in the data frame
-    if (!pheno %in% names(data)) {
-      warning(paste("Variable '", pheno, "' not found in the data frame. Skipping..."))
-      next
-    }
-
-    # Create a ggplot histogram for the variable
-    plot <- ggplot(data, aes(x = .data[[pheno]], fill= sex, color=sex)) +
-      geom_histogram(alpha=0.5, position="identity") +
-      labs(title = paste("Histogram of", pheno), x = pheno, y = "Frequency")
-
-    # Add the plot to the list
-    plots[[pheno]] <- plot
-  }
-
-  # Return the list of ggplot objects
-  return(plots)
-}
-```
-
-center
-```{r}
-ukb_centered <- ukb_df_totals %>%
-  group_by(sex) %>%
-  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
-  ungroup()
-```
-
-```{r}
-# Generate histograms (mean-centered w/in sex)
-global_plots <- plot.hist.by.sex(ukb_centered, vol_list_global)
-
-# Print or display the histograms
-for (pheno in vol_list_global) {
-  print(global_plots[[pheno]])
-}
-```
-
-```{r}
-# Generate histograms (mean-centered w/in sex)
-region_vol_plots <- plot.hist.by.sex(ukb_centered, vol_list_regions)
-
-# Print or display the histograms
-for (pheno in vol_list_regions) {
-  print(region_vol_plots[[pheno]])
-}
-```
-
-```{r}
-# Generate histograms (mean-centered w/in sex)
-region_sa_plots <- plot.hist.by.sex(ukb_centered, sa_list)
-
-# Print or display the histograms
-for (pheno in sa_list) {
-  print(region_sa_plots[[pheno]])
-}
-```
-
-```{r}
-# Generate histograms (mean-centered w/in sex)
-region_ct_plots <- plot.hist.by.sex(ukb_centered, ct_list)
-
-# Print or display the histograms
-for (pheno in ct_list) {
-  print(region_ct_plots[[pheno]])
-}
+ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
 ```

### 2023-10-06 ukb_curation-eda.Rmd
splitting out EDA and modeling into separate Rmds

--- 2023-10-06 ukb_mods.Rmd
@@
 ```
+
+## Data Org
 
@@
 ```
-
-## Global Mod Fitting
-
-Try using `stepGAICALL.A()`
-
-```{r}
-m1 <- gamlss(formula = GMV ~ pb(age_days) + sex + fs_version + pb(age_days):sex,
-                       sigma.formula = ~  pb(age_days) + sex,
-                       nu.formula = ~ 1,
-                       control = gamlss.control(n.cyc = 200), 
-                       family = BCCG, data=ukb_df, trace = FALSE)
-
-m2 <- stepGAICAll.A(m1, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(age_days):sex), k=2) #AIC
-
-m2
-summary(m2)
-```
-
-seeing if things change when i hard-code a sex*age interaction term
-
-```{r}
-m3 <- gamlss(formula = GMV ~ pb(age_days) + sex + fs_version + pb(sex.age),
-                       sigma.formula = ~  pb(age_days) + sex,
-                       nu.formula = ~ 1,
-                       control = gamlss.control(n.cyc = 200), 
-                       family = BCCG, data=ukb_df, trace = FALSE)
-m4 <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(sex.age)), k=2) #AIC
-
-summary(m4)
-```
-
-Seems like stepGAICAll handles combined sex.age term better than an interaction
-Try adding TBV
-```{r}
-m3.tbv <- gamlss(formula = GMV ~ pb(age_days) + sex + fs_version + pb(sex.age) + pb(TBV),
-                       sigma.formula = ~  pb(age_days) + sex,
-                       nu.formula = ~ 1,
-                       control = gamlss.control(n.cyc = 200), 
-                       family = BCCG, data=ukb_df, trace = FALSE)
-m4.tbv <- stepGAICAll.A(m3, scope=list(lower=~1, upper=~pb(age_days) + sex + fs_version + pb(sex.age) + pb(TBV)), k=2, glim.control(cyc=100)) #AIC
-summary(m4.tbv)
-```
-
-```{r}
-plot_singlestudy_centiles(m4, "GMV", ukb_df, "sex")
-```
-
-Let's look at diagnostics
-```{r}
-plot(m4)
-wp(m4, xvar=age_days, n.inter=4) #leptokurtosis (tails fo fitted distribution too light)
-dtop(m4)
-rqres.plot(m4)
-
-plot(m4.tbv)
-wp(m4.tbv, xvar=age_days, n.inter=4) #U-shape: fitted skewness too low
-dtop(m4.tbv)
-rqres.plot(m4.tbv)
-```
-
-# CUBIC Models
-Playing around wiht models fit on cubic using `fit_ukb_mod.R`
-
-scp-ing random model to look at (`lh_SA_middletemporal_mod.rds`)
-
-```{r}
-lh_SA_middletemporal_mod <- readRDS("ukb/gamlss_objs/lh_SA_middletemporal_mod.rds")
-a <- summary(lh_SA_middletemporal_mod)
-dim(a)
-
-#sigma coefficients
-coef(lh_SA_middletemporal_mod, what="sigma") %>%
-  names()
-
-test <- confint(lh_SA_middletemporal_mod)
-colnames(test)
-rownames(test)
-test["sigma.sexMale", "97.5 %"]
-df.test <- as.data.frame(test)
-df.test$params <- rownames(df.test)
-# Remove the row names from the dataframe
-rownames(df.test) <- NULL
-
-df.test
-#using drop1 to get significance, gamlss book p 125 for df info
-#get df from OG model - not sure how this would be affected by using stepGAIC?
-
-str(lh_SA_middletemporal_mod)
-lh_SA_middletemporal_mod$sigma.nl.df
-
-#start df with confidence intervals
-confint.matrix <- confint(lh_SA_middletemporal_mod)
-df <- as.data.frame(confint.matrix) %>%
-  rownames_to_column(var = "moment.param") %>%
-  mutate(
-    moment = sapply(strsplit(moment.param, "\\."), `[`, 1),
-    param = sub("^[^.]+\\.", "", moment.param)
-  )
-
-#now that we have the moments, add the rest of the summary info
-sum.table <- summary(lh_SA_middletemporal_mod)
-df2 <- as.data.frame(sum.table) %>%
-  rownames_to_column(var = "param")
-#check
-rownames(sum.table) == df$param
-
-find.param <- function(gamlss.rds.file, moment, string) {
-  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
-  gamlss.obj <- readRDS(gamlss.rds.file)
-  sigma.coeff.list <- coef(gamlss.rds.file, what= moment) %>%
-    names()
-  any(sapply(sigma.coeff.list, function(x) grepl(string, x, ignore.case = TRUE)))
-}
-find.param(lh_SA_middletemporal_mod, "sigma", "sex")
-
-readRDS("ukb/gamlss_objs/rh_Vol_parsopercularis_mod.rds")
-
-```

### 2023-10-06 proof-of-concept_mods.Rmd
starting with simple/theoretical mods for proof-of-concept per discussions with andrew & taki

---
title: "proof-of-concept_mods"
output: html_document
date: '2023-10-06'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-10-06 proof-of-concept_mods.Rmd
narrowing age range

@@
 pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
+
+source("./R_scripts/gamlss_helper_funs.R")
 ```
 
-## R Markdown
+## Load Data
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
+```{r}
+ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
+ukb_df_age.filt <- read_csv(file="data/ukb_CN_data_agefilt.csv")
 ```
 
-## Including Plots
+Designate *basic* model formula
 
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
+```{r}
+basic.m1 <- gamlss(formula = GMV ~ poly(age_days, 3) + sex,
+                       sigma.formula = ~ poly(age_days, 3) + sex,
+                       nu.formula = ~ 1,
+                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+plot
+
+```{r}
+plot_singlestudy_centiles(basic.m1, "GMV", ukb_df_age.filt, "sex")
+```
+
+Check sex's significance in variance
+
+```{r}
+drop1(basic.m1, what="sigma", scope = "sex")
+```

### 2023-10-06 ukb_curation-eda.Rmd
narrowing age range

@@
 
-## EDA
-
 ```{r}
 ggplot(data=ukb_df, aes(x=age_days)) +
-  geom_histogram(aes(fill=sex), alpha=0.6, position="identity")
+  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
 ggplot(data=ukb_df, aes(x=log_age)) +
@@
 ```
+
+### Condense age range
+
+Random male subjects are younger than females - going to drop because having issues estimating total brain metrics in young women where there's no data
+
+```{r}
+ukb_df_totals %>%
+  dplyr::filter(age_yrs <= 49) %>%
+ggplot(data=, aes(x=age_days)) +
+  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
+
+ukb_df_totals %>%
+  dplyr::filter(age_yrs >= 79) %>%
+ggplot(data=, aes(x=age_days)) +
+  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
+```
+
+Restricting to ages 50-80, since that's when there's roughly ~100 ppts per sex
+
+```{r}
+ukb_df_age.filt <- ukb_df_totals %>%
+  dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
+#write data
+write_csv(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
+```
+
+## EDA
 

### 2023-10-06 R_scripts/gamlss_helper_funs.R

@@ -85,8 +85,8 @@ mean.tb.sim <- function(df, sex_level, ageRange, measure){
     tb.df <- df %>% 
       dplyr::filter(sex == sex_level) %>%
       arrange(age_days) %>%
-      dplyr::filter(age_days >= (i-300) & age_days <= (i+300)) %>%
-      as.data.frame()
+      dplyr::filter(age_days >= (i-183) & age_days <= (i+183)) %>%
+      as.data.frame() # +- 1.5 yrs
     tb.val <- mean(tb.df[[measure]], na.rm = TRUE)
     tb_list <- append(tb_list, tb.val)
   }
@@ -99,21 +99,19 @@ sim.data.ukb <- function(df){
   ageRange <- seq(minAge, maxAge, 25)  # generate an age range with increments of 25 (since scaling in days)
   
   #sim data
-  dataToPredictM <- data.frame(age_yrs=ageRange,
+  dataToPredictM <- data.frame(age_days=ageRange,
                                sex=c(rep(as.factor("Male"), length(ageRange))),
                                fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
-                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
-                               sex.age_yrs=c(rep(0, length(ageRange))))
+                               sex.age=c(rep(0, length(ageRange))))
   dataToPredictM$TBV <- mean.tb.sim(df, "Male", ageRange, "TBV")
   dataToPredictM$Vol_total <- mean.tb.sim(df, "Male", ageRange, "Vol_total")
   dataToPredictM$SA_total <- mean.tb.sim(df, "Male", ageRange, "SA_total")
   dataToPredictM$CT_total <- mean.tb.sim(df, "Male", ageRange, "CT_total")
   
-  dataToPredictF <- data.frame(age_yrs=ageRange,
+  dataToPredictF <- data.frame(age_days=ageRange,
                                sex=c(rep(as.factor("Female"), length(ageRange))),
                                fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
-                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
-                               sex.age_yrs=ageRange)
+                               sex.age=ageRange)
   dataToPredictF$TBV <- mean.tb.sim(df, "Female", ageRange, "TBV")
[…]

### 2023-10-11 proof-of-concept_mods.Rmd
fixed summary parsing, visualizing model outputs

@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
+pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg)
 
@@
 ukb_df_age.filt <- read_csv(file="data/ukb_CN_data_agefilt.csv")
+
+vol_list_global <- readRDS(file="R_scripts/vol_list_global.rds")
+vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
+sa_list <- readRDS(file="R_scripts/SA_list.rds")
+ct_list <- readRDS(file="R_scripts/CT_list.rds")
 ```
@@
                        family = BCCG, data=ukb_df_age.filt, trace = FALSE)
+summary(basic.m1)
 ```
@@
 ```{r}
-drop1(basic.m1, what="sigma", scope = "sex")
+K <- drop1(basic.m1, what="sigma", scope = "sex")
+class(K)
+K["Pr(Chi)"]
+K$"Pr(Chi)"[2]
+
+test.mu <- drop1(basic.m1, what="mu", scope = "sex")
+test.sig <- drop1(basic.m1, what="sigma", scope = "sex")
+
+#write for testing
+saveRDS(basic.m1, file="/Users/megardn/Desktop/BGD_Repos/combat_biovar/gamlss_basic_test_mod.rds")
+
+get.and.drop1.p("/Users/megardn/Desktop/BGD_Repos/combat_biovar/gamlss_basic_test_mod.rds", moment="sigma", term="sex")
+
+test.model.list <- c()
+test.model.list[[1]] <- basic.m1
+lapply(test.model.list, drop1, what="mu", scope = "sex")
 ```
+
+## Significance Testing
+
+scp'd models from cubic to parse
+
+```{r}
+read_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic/gamlss_objs" #path to bfp models
+```
+
+```{r}
+#iterate across gamlss objects
+model.files <- list.files(path = read_path, pattern = "mod.rds", full.names = TRUE)
+
+sigma.df <- data.frame("file_path" = model.files)
+sigma.df <- sigma.df %>%
+  mutate(region = sub("_mod\\.rds$", "", basename(as.character(file_path))))
+
+#test significance of sex term in sigma
+sigma.df$sig.sex_drop1.p <- lapply(sigma.df$file_path, get.and.drop1.p, moment="sigma", term="sex")
+
+#extract values of sex term from summary table
+summary.list <- lapply(sigma.df$file_path, get.summary)
+summary.df <- bind_rows(summary.list)
+sigma.sex.df <- summary.df %>%
+  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA)),
+    sig = case_when(p.value < 0.05 ~ TRUE,
+                    p.value >= 0.05 ~ FALSE),
+    sig_bf.corr = case_when(pheno %in% vol_list_global & p.value < (0.05/length(vol_list_global)) ~ TRUE,
+                            pheno %in% vol_list_global & p.value >= (0.05/length(vol_list_global)) ~ FALSE,
+                            !(pheno %in% vol_list_global) & p.value < (0.05/length(ct_list)) ~ TRUE,
+                            !(pheno %in% vol_list_global) & p.value >= (0.05/length(ct_list)) ~ FALSE,
+                            TRUE ~ NA),
+    label = sub("_[^_]*_", "_", pheno)) #for plotting
+
+#add more info from parcellations
+dk.parc <- dk$data %>%
+  as.data.frame() %>%
+  na.omit() %>%
+  dplyr::select(c(hemi, region, label)) %>%
+  distinct()
+
+sigma.sex.df2 <- left_join(sigma.sex.df, dk.parc, by="label")
+```
+
+Visualize
+
+```{r}
+table(sigma.sex.df2$sig)
+table(sigma.sex.df2$sig_bf.corr)
+```
+
+```{r, fig.width=11}
+sigma.sex.df2 %>%
+  dplyr::filter(sig_bf.corr == TRUE) %>%
+  ggplot() +
+  geom_col(aes(x=reorder(pheno,t_stat), y=t_stat, fill=pheno_cat)) + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
+
+sigma.sex.df2 %>%
+  dplyr::filter(sig_bf.corr == TRUE) %>%
+  ggplot() +
+  geom_col(aes(x=label, y=t_stat, fill=pheno_cat)) + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
+
+sigma.sex.df2 %>%
+  dplyr::filter(sig_bf.corr == TRUE) %>%
+  ggplot() +
+  geom_col(aes(x=region, y=t_stat, fill=hemi)) + 
+  facet_wrap(~pheno_cat) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
+```
+
+Plot on brains!
+
+```{r}
+sigma.sex.df2 %>%
+  dplyr::filter(sig_bf.corr == TRUE) %>%
+  group_by(pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = t_stat)) +
+  scale_fill_gradient2() +
+  facet_wrap(~pheno_cat) +
+  theme_void() +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("t-statistic for effect of male sex on sigma")
+```
+
+```{r}
+
+```

### 2023-10-11 ukb_curation-eda.Rmd
fixed summary parsing, visualizing model outputs

@@
 writeLines(vol_list_global, con = "pheno_lists/Vol_list_global.txt")
+saveRDS(vol_list_global, file="R_scripts/vol_list_global.rds")
 
@@
 writeLines(vol_list_regions, con = "pheno_lists/Vol_list_regions.txt")
+saveRDS(vol_list_regions, file="R_scripts/Vol_list_regions.rds")
 
@@
 writeLines(sa_list, con = "pheno_lists/SA_list.txt")
+saveRDS(sa_list, file="R_scripts/SA_list.rds")
 
@@
 writeLines(ct_list, con = "pheno_lists/CT_list.txt")
+saveRDS(ct_list, file="R_scripts/CT_list.rds")
 ```

### 2023-10-11 R_scripts/gamlss_helper_funs.R

@@ -3,6 +3,9 @@ library(gamlss)
 library(dplyr)
 library(ggplot2)
 library(data.table)
+library(broom.mixed)
+
+#note - some of these functions do not seem to work on CUBIC, as they rely on saving environment variables
 
 ### CONTENTS ###
 # find.param()
@@ -78,25 +81,21 @@ get.and.drop1.p <- function(gamlss.rds.file, moment, term) {
   #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
   gamlss.rds.file <- as.character(gamlss.rds.file)
   gamlss.obj <- readRDS(gamlss.rds.file)
+  #summary(gamlss.obj)
   t <- drop1(gamlss.obj, what = moment, scope = term)
   pval <- t$"Pr(Chi)"[2]
   return(pval)
 }
 
-get.summary.table.outputs<- function(gamlss.rds.file, moment, select1=NA, select2=NA) {
+get.summary<- function(gamlss.rds.file) {
   #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
   gamlss.rds.file <- as.character(gamlss.rds.file)
   gamlss.obj <- readRDS(gamlss.rds.file)
-  selection <- as.character(paste0(moment, ".coef.table"))
-  sum.table <- get(selection, summary(gamlss.obj, save=TRUE))
-  if (!is.na(select1)){
-    select1 <- as.character(select1)
-    select2 <- as.character(select2)
-    v <- sum.table[select1 ,select2]
-    return(v)
-  } else {
-    return(sum.table)
-  }
+  sum.table <- broom.mixed::tidy(gamlss.obj) %>%
+    as.data.frame() %>%
+    rename(t_stat = statistic)
+  sum.table$pheno <- sub("_mod\\.rds$", "", basename(gamlss.rds.file)) #append model name
+  return(sum.table)
 }
 
 ################

### 2023-10-11 R_scripts/sex_sig_test.R

@@ -8,6 +8,8 @@ library(tibble)
 
 source("R_scripts/gamlss_helper_funs.R")
 
+#note - some of these functions do not seem to work on CUBIC, as they rely on saving environment variables
+
 #GET ARGS
 args <- commandArgs(trailingOnly = TRUE)
 read_path <- args[1] #path to bfp models

### 2023-10-11 var_sig_test.sh

@@ -1,1 +1,4 @@
-Rscript --save /cbica/home/gardnerm/combat-biovar/R_scripts/sex_sig_test.R /cbica/home/gardnerm/combat-biovar/ukb_basic/gamlss_objs /cbica/home/gardnerm/combat-biovar/ukb_basic
+#cubic
+#Rscript --save /cbica/home/gardnerm/combat-biovar/R_scripts/sex_sig_test.R /cbica/home/gardnerm/combat-biovar/ukb_basic/gamlss_objs /cbica/home/gardnerm/combat-biovar/ukb_basic
+#local
+Rscript --save /Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/sex_sig_test.R /Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic/gamlss_objs /Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic

### 2023-10-11 proof-of-concept_mods.Rmd
simulating study sites

@@
 
+## Simulate Sites
+Dividing each subject into 3 sites (A, B, and C) with M:F ratios of 1:1, 4:1, and 1:4
+
 ```{r}
+sim.site.list <- c("Site_A", "Site_B", "Site_C")
+n_female <- table(ukb_df_age.filt$sex)["Female"]
+n_male <- table(ukb_df_age.filt$sex)["Male"]
 
+# Sample probabilities
+male_prob <- c(0.5, 0.125, 0.375)
+female_prob <- c(0.5, 0.375, 0.125)
+
+# Separate samples for males and females
+sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
+sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)
+
+ukb_df_sites <- ukb_df_age.filt %>%
+  mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))
+table(ukb_df_sites$sim.site, ukb_df_sites$sex)
+sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
 ```
+
+Center and visualize distributions
+
+```{r}
+ukb_sites_centered <- ukb_df_sites %>%
+  group_by(sex) %>%
+  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
+  ungroup()
+```
+
+```{r}
+# Generate histograms (mean-centered w/in sex)
+global_site_plots <- plot.hist.by.sex(ukb_sites_centered, vol_list_global, facet_fac = "sim.site")
+
+# Print or display the histograms
+for (pheno in vol_list_global) {
+  print(global_site_plots[[pheno]])
+}
+```

### 2023-10-11 ukb_curation-eda.Rmd
updating plot.hist.by.sex() and moving to gamlss_helper_funs.R for easy access across scripts

@@
 Look at variances
-```{r}
-plot.hist.by.sex <- function(data, pheno_list) {
-
-  # Create an empty list to store the ggplot objects
-  plots <- list()
-
-  # Loop through the list of variables
-  for (pheno in pheno_list) {
-    # Check if the variable exists in the data frame
-    if (!pheno %in% names(data)) {
-      warning(paste("Variable '", pheno, "' not found in the data frame. Skipping..."))
-      next
-    }
-
-    # Create a ggplot histogram for the variable
-    plot <- ggplot(data, aes(x = .data[[pheno]], fill= sex, color=sex)) +
-      geom_histogram(alpha=0.5, position="identity") +
-      labs(title = paste("Histogram of", pheno), x = pheno, y = "Frequency")
-
-    # Add the plot to the list
-    plots[[pheno]] <- plot
-  }
-
-  # Return the list of ggplot objects
-  return(plots)
-}
-```
 

### 2023-10-11 R_scripts/gamlss_helper_funs.R

@@ -192,3 +192,33 @@ plot_singlestudy_centiles <- function(gamlssModel, phenotype, df, color_var)
   print(sampleCentileFan)
 }
 
+plot.hist.by.sex <- function(data, pheno_list, facet_fac = NA) {
+  
+  # Create an empty list to store the ggplot objects
+  plots <- list()
+  
+  # Loop through the list of variables
+  for (pheno in pheno_list) {
+    # Check if the variable exists in the data frame
+    if (!pheno %in% names(data)) {
+      warning(paste("Variable '", pheno, "' not found in the data frame. Skipping..."))
+      next
+    }
+    
+    # Create a ggplot histogram for the variable
+    plot <- ggplot(data, aes(x = .data[[pheno]], fill= sex, color=sex)) +
+      geom_histogram(alpha=0.5, position="identity") +
+      labs(title = paste("Histogram of", pheno), x = pheno, y = "Frequency")
+    
+    # add facet wrapping if needed
+    if (!is.na(facet_fac)) {
+      plot <- plot + facet_wrap(~ .data[[facet_fac]])
+    }
+    
+    # Add the plot to the list
+    plots[[pheno]] <- plot
+  }
+  
+  # Return the list of ggplot objects
+  return(plots)
+}

### 2023-10-11 combat_sandbox.Rmd
working through combat for regional vols - some warninigs/errs

---
title: "Combat Sandbox"
output: html_document
date: '2023-10-06'
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(devtools, data.table, readr, ggplot2, tidyverse, mgcv)
#install.packages("devtools")
devtools::install_github("andy1764/ComBatFamily", build_vignettes = FALSE) #per discussion with Andrew, sent vignettes separately
library(ComBatFamily)
```

#Prep data

Using additional instructions from J Fortin's [initial repo](https://github.com/Jfortin1/neuroCombat_Rpackage/tree/2ccab2b42c59163717f012c70a54d20e8757b1c7).

Just using regional volume measures for now.

```{r}
ukb_df_sites <- read_csv(file="data/ukb_CN_data_simsites.csv")
vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
```

```{r}
#regional volume data to harmonize
rvol.df <- ukb_df_sites %>%
  dplyr::select(all_of(vol_list_regions))

#fake batches
batch <- as.factor(ukb_df_sites$sim.site)

#covariates
covar.mat <- ukb_df_sites %>%
  dplyr::select(c(age_days, sex)) %>%
  mutate(age_days = as.numeric(age_days)) %>%
  as.matrix()
#weirdly converts age_days to a character, idk if that will be a problem

nrow(rvol.df) == length(batch)
nrow(rvol.df) == nrow(covar.mat)
```

## ComBat

simple combat with linear model for covariates
```{r}
cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sex)
```

Getting warning - Warning: prediction from a rank-deficient fit may be misleading

combat with age as smooth
```{r}
cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sex)
```

### 2023-10-11 proof-of-concept_mods.Rmd
working through combat for regional vols - some warninigs/errs

@@
 ```{r}
+set.seed(1010101)
 sim.site.list <- c("Site_A", "Site_B", "Site_C")
@@
 # Sample probabilities
-male_prob <- c(0.5, 0.125, 0.375)
-female_prob <- c(0.5, 0.375, 0.125)
+male_prob <- c(0.33, 0.0825, 0.5875)
+female_prob <- c(0.33, 0.5875, 0.0825)
 
@@
   mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))
+
+#write out
+write_csv(ukb_df_sites, file="data/ukb_CN_data_simsites.csv")
+```
+
+```{r}
+table(ukb_df_sites$sim.site)
 table(ukb_df_sites$sim.site, ukb_df_sites$sex)
 sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
+anova(lm(age_days ~ sim.site, ukb_df_sites))
+ukb_df_sites %>%
+  group_by(sim.site) %>%
+  summarise(min = min(age_days), max = max(age_days), mean = mean(age_days))
 ```
@@
 ```
+## ComBat
+
+And now, to combat the data across the simulated sites!

### 2023-10-11 ukb_curation-eda.Rmd
working through combat for regional vols - some warninigs/errs

@@
 source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
+source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
 ```
@@
 }
+
+ukb_centered %>%
+  dplyr::filter(sex == "Male") %>%
+  select(GMV) %>%
+  var()
+
+ukb_centered %>%
+  dplyr::filter(sex == "Female") %>%
+  select(GMV) %>%
+  var()
+
+ukb_centered %>%
+  #dplyr::filter(sex == "Male") %>%
+  select(GMV) %>%
+  var()
 ```

### 2023-10-11 combat_sandbox.Rmd
fixed combat errs by using model.matrix to create covar matrix

@@
 
-#covariates
-covar.mat <- ukb_df_sites %>%
-  dplyr::select(c(age_days, sex)) %>%
-  mutate(age_days = as.numeric(age_days)) %>%
-  as.matrix()
-#weirdly converts age_days to a character, idk if that will be a problem
+#covariates - using model.matrix to dummycode sex as numeric
+covar.mat <- model.matrix( ~ age_days + sex, data=ukb_df_sites)
 
@@
 ```{r}
-cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sex)
+cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale)
 ```
-
-Getting warning - Warning: prediction from a rank-deficient fit may be misleading
 
@@
 ```{r}
-cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sex)
+cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale)
 ```

### 2023-10-12 combat_sandbox.Rmd
playing around with different combats' variance outputs, plots

@@
 ---
+
+Playing around with different combat configs using the UKB data, which has been assigned to 3 simulated study sites (see `proof-of-concept_mods.Rmd`)
 
@@
 }
-pacman::p_load(devtools, data.table, readr, ggplot2, tidyverse, mgcv)
+pacman::p_load(devtools, data.table, readr, ggplot2, tidyverse, mgcv, gamlss)
 #install.packages("devtools")
@@
 
+simple combat with no covariate preservation
+```{r}
+cf <- comfam(rvol.df, batch)
+cf$dat.combat
+cf$dat.combat[, "sim.site"] <- batch
+```
+
 simple combat with linear model for covariates
@@
 cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale)
+cf.lm$dat.combat
+```
+plot
+```{r}
+dim(cf.lm$dat.combat)[1] == length(batch)
+#add back site
+cf.lm$dat.combat[, "sim.site"] <- batch
+
+ggplot() + 
+ #geom_histogram(aes(x=cf.lm$dat.combat$GMV, y=after_stat(density), fill=cf.lm$dat.combat$sim.site, color=cf.lm$dat.combat$sim.site), alpha = .2)+
+ geom_density(aes(x=cf.lm$dat.combat$lh_Vol_superiorparietal, fill=cf.lm$dat.combat$sim.site, color=cf.lm$dat.combat$sim.site), alpha=.4) +
+  geom_vline(aes(xintercept=cf.lm_summary$sd_below.lh_v_sp, color=cf.lm_summary$sim.site)) +
+  geom_vline(aes(xintercept=cf.lm_summary$sd_above.lh_v_sp, color=cf.lm_summary$sim.site))
 ```
 
-combat with age as smooth
+Try setting site A (which has even sex distribution) as ref
+
+```{r}
+cf.lm2 <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale, ref.batch = "Site_A")
+cf.lm2$dat.combat
+cf.lm2$dat.combat[, "sim.site"] <- batch
+```
+
+combat gam with age as smooth
 ```{r}
 cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale)
+cf.gam$dat.combat[, "sim.site"] <- batch
 ```
+
+```{r}
+cf.gam2 <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale, ref.batch = "Site_A")
+cf.gam2$dat.combat[, "sim.site"] <- batch
+```
+
+combat gamlss!
+
+```{r}
+cf.gamlss <- comfam(rvol.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
+              sigma.formula = ~ age_days + sexMale, 
+              control = gamlss.control(trace = FALSE))
+cf.gamlss$dat.combat[, "sim.site"] <- batch
+```
+
+```{r}
+cf.gamlss2 <- comfam(rvol.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
+              sigma.formula = ~ age_days + sexMale, 
+              control = gamlss.control(trace = FALSE),
+              ref.batch = "Site_A")
+cf.gamlss2$dat.combat[, "sim.site"] <- batch
+```
+
+## Compare Variances
+
+Randomly picked `lh_Vol_superiorparietal` to compare
+
+Get variances
+```{r}
+#raw data
+sum_df <- ukb_df_sites %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal),
+            var.lh_v_sp = var(lh_Vol_superiorparietal))
+#no covars
+cf_summary <- cf$dat.combat %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
+
+#simple lm
+cf.lm_summary <- cf.lm$dat.combat %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
+
+#simple lm with ref site
+cf.lm2_summary <- cf.lm2$dat.combat %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
+
+#simple gam
+cf.gam_summary <- cf.gam$dat.combat %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
+
+#simple gam with ref site
+cf.gam2_summary <- cf.gam2$dat.combat %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
+
+#simple gam
+cf.gamlss_summary <- cf.gamlss$dat.combat %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
+
+#simple gam with ref site
+cf.gamlss2_summary <- cf.gamlss2$dat.combat %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
+```
+compile
+
+```{r}
+var_comp_df <- data.frame(sim.site = sim.site.list,
+                          raw_data = sum_df$var.lh_v_sp,
+                          cf = cf_summary$var.lh_v_sp,
+                          cf.lm = cf.lm_summary$var.lh_v_sp,
+                          cf.lm2 = cf.lm2_summary$var.lh_v_sp,
+                          cf.gam = cf.gam_summary$var.lh_v_sp,
+                          cf.gam2 = cf.gam2_summary$var.lh_v_sp,
+                          cf.gamlss = cf.gamlss_summary$var.lh_v_sp,
+                          cf.gamlss2 = cf.gamlss2_summary$var.lh_v_sp)
+
+var_comp_long <- var_comp_df %>%
+  pivot_longer(!sim.site, names_to="type", values_to = "variance")
+
+ggplot(var_comp_long) +
+  geom_col(aes(x=type, y=variance, fill=sim.site, color=sim.site), position = "dodge")
+
+ggplot(var_comp_long) +
+  geom_col(aes(x=type, y=variance), position = "dodge") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~sim.site)
+```

### 2023-10-12 proof-of-concept_mods.Rmd
playing around with different combats' variance outputs, plots

--- 2023-10-11 proof-of-concept_mods.Rmd
@@
 
+## Simulate Sites
+Dividing each subject into 3 sites (A, B, and C) with M:F ratios of 1:1, 1:4, and 4:1
+
 ```{r}
+set.seed(1010101)
+sim.site.list <- c("Site_A", "Site_B", "Site_C")
+n_female <- table(ukb_df_age.filt$sex)["Female"]
+n_male <- table(ukb_df_age.filt$sex)["Male"]
 
+# Sample probabilities
+male_prob <- c(0.33, 0.0825, 0.5875)
+female_prob <- c(0.33, 0.5875, 0.0825)
+
+# Separate samples for males and females
+sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
+sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)
+
+ukb_df_sites <- ukb_df_age.filt %>%
+  mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))
+
+#write out
+write_csv(ukb_df_sites, file="data/ukb_CN_data_simsites.csv")
 ```
+
+```{r}
+table(ukb_df_sites$sim.site)
+table(ukb_df_sites$sim.site, ukb_df_sites$sex)
+sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
+anova(lm(age_days ~ sim.site, ukb_df_sites))
+```
+
+Need to come up with cleaner way to do this...
+
+```{r}
+site_sum_df <- ukb_df_sites %>%
+  group_by(sim.site) %>%
+  summarise(proportion_male = mean(sex == "Male"),
+            min_age = min(age_days), max_age = max(age_days), mean_age = mean(age_days), 
+            var_gmv = var(GMV), var_wmv=var(WMV), var_sgmv=var(sGMV), var_ven = var(Ventricles),
+            sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))
+```
+
+Center and visualize distributions
+
+```{r}
+ukb_sites_sex.centered <- ukb_df_sites %>%
+  group_by(sex) %>%
+  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
+  ungroup()
+```
+
+```{r}
+# Generate histograms (mean-centered w/in sex)
+global_site_plots <- plot.hist.by.sex(ukb_sites_sex.centered, vol_list_global, facet_fac = "sim.site")
+
+# Print or display the histograms
+for (pheno in vol_list_global) {
+  print(global_site_plots[[pheno]])
+}
+```
+
+Look at uncentered distributions across sites
+```{r}
+ggplot(ukb_df_sites, aes(x=GMV)) + 
+ geom_histogram(aes(y=after_stat(density)))+
+ geom_density(alpha=.2) +
+  geom_vline(aes(xintercept=(mean(GMV)+sd(GMV)))) +
+  geom_vline(aes(xintercept=(mean(GMV)-sd(GMV))))
+
+ggplot(ukb_df_sites, aes(x=GMV)) + 
+ geom_histogram(aes(y=after_stat(density)))+
+ geom_density(alpha=.2) +
+  geom_vline(aes(xintercept=(mean(GMV)+sd(GMV)))) +
+  geom_vline(aes(xintercept=(mean(GMV)-sd(GMV)))) +
+  facet_wrap(~sim.site)
+```
+
+```{r}
+centered.site_df <- ukb_df_sites %>%
+  group_by(sim.site) %>%
+  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
+  ungroup()
+
+centered.site_sum_df <- centered.site_df %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))
+
+ggplot() + 
+ #geom_histogram(aes(x=centered.site_df$GMV, y=after_stat(density), fill=centered.site_df$sim.site, color=centered.site_df$sim.site), alpha = .2)+
+ geom_density(aes(x=centered.site_df$GMV, fill=centered.site_df$sim.site, color=centered.site_df$sim.site), alpha=.4) +
+  geom_vline(aes(xintercept=centered.site_sum_df$sd_below.gmv, color=centered.site_sum_df$sim.site)) +
+  geom_vline(aes(xintercept=centered.site_sum_df$sd_above.gmv, color=centered.site_sum_df$sim.site))
+```
+
+## ComBat
+
+And now, to combat the data across the simulated sites!

### 2023-10-12 proof-of-concept_mods.Rmd
moving site sims to EDA rmd

--- 2023-10-11 proof-of-concept_mods.Rmd
@@
 ```
-
-## Simulate Sites
-Dividing each subject into 3 sites (A, B, and C) with M:F ratios of 1:1, 4:1, and 1:4
-
-```{r}
-sim.site.list <- c("Site_A", "Site_B", "Site_C")
-n_female <- table(ukb_df_age.filt$sex)["Female"]
-n_male <- table(ukb_df_age.filt$sex)["Male"]
-
-# Sample probabilities
-male_prob <- c(0.5, 0.125, 0.375)
-female_prob <- c(0.5, 0.375, 0.125)
-
-# Separate samples for males and females
-sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
-sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)
-
-ukb_df_sites <- ukb_df_age.filt %>%
-  mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))
-table(ukb_df_sites$sim.site, ukb_df_sites$sex)
-sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
-```
-
-Center and visualize distributions
-
-```{r}
-ukb_sites_centered <- ukb_df_sites %>%
-  group_by(sex) %>%
-  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
-  ungroup()
-```
-
-```{r}
-# Generate histograms (mean-centered w/in sex)
-global_site_plots <- plot.hist.by.sex(ukb_sites_centered, vol_list_global, facet_fac = "sim.site")
-
-# Print or display the histograms
-for (pheno in vol_list_global) {
-  print(global_site_plots[[pheno]])
-}
-```

### 2023-10-12 ukb_curation-eda.Rmd
moving site sims to EDA rmd

@@
 ```
+
+## Simulate Sites
+Dividing each subject into 3 sites (A, B, and C) with M:F ratios of 1:1, 1:4, and 4:1
+
+```{r}
+set.seed(1010101)
+sim.site.list <- c("Site_A", "Site_B", "Site_C")
+n_female <- table(ukb_df_age.filt$sex)["Female"]
+n_male <- table(ukb_df_age.filt$sex)["Male"]
+
+# Sample probabilities
+male_prob <- c(0.33, 0.0825, 0.5875)
+female_prob <- c(0.33, 0.5875, 0.0825)
+
+# Separate samples for males and females
+sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
+sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)
+
+ukb_df_sites <- ukb_df_age.filt %>%
+  mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))
+
+#write out
+write_csv(ukb_df_sites, file="data/ukb_CN_data_simsites.csv")
+```
+
+```{r}
+table(ukb_df_sites$sim.site)
+table(ukb_df_sites$sim.site, ukb_df_sites$sex)
+sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
+anova(lm(age_days ~ sim.site, ukb_df_sites))
+```
+
+Need to come up with cleaner way to do this...
+
+```{r}
+site_sum_df <- ukb_df_sites %>%
+  group_by(sim.site) %>%
+  summarise(proportion_male = mean(sex == "Male"),
+            min_age = min(age_days), max_age = max(age_days), mean_age = mean(age_days), 
+            var_gmv = var(GMV), var_wmv=var(WMV), var_sgmv=var(sGMV), var_ven = var(Ventricles),
+            sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))
+```
+
+Center and visualize distributions
+
+```{r}
+ukb_sites_sex.centered <- ukb_df_sites %>%
+  group_by(sex) %>%
+  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
+  ungroup()
+```
+
+```{r}
+# Generate histograms (mean-centered w/in sex)
+global_site_plots <- plot.hist.by.sex(ukb_sites_sex.centered, vol_list_global, facet_fac = "sim.site")
+
+# Print or display the histograms
+for (pheno in vol_list_global) {
+  print(global_site_plots[[pheno]])
+}
+```
+
+Look at uncentered distributions across sites
+```{r}
+ggplot(ukb_df_sites, aes(x=GMV)) + 
+ geom_histogram(aes(y=after_stat(density)))+
+ geom_density(alpha=.2) +
+  geom_vline(aes(xintercept=(mean(GMV)+sd(GMV)))) +
+  geom_vline(aes(xintercept=(mean(GMV)-sd(GMV))))
+
+ggplot(ukb_df_sites, aes(x=GMV)) + 
+ geom_histogram(aes(y=after_stat(density)))+
+ geom_density(alpha=.2) +
+  geom_vline(aes(xintercept=(mean(GMV)+sd(GMV)))) +
+  geom_vline(aes(xintercept=(mean(GMV)-sd(GMV)))) +
+  facet_wrap(~sim.site)
+```
+
+```{r}
+centered.site_df <- ukb_df_sites %>%
+  group_by(sim.site) %>%
+  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
+  ungroup()
+
+centered.site_sum_df <- centered.site_df %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))
+
+ggplot() + 
+ #geom_histogram(aes(x=centered.site_df$GMV, y=after_stat(density), fill=centered.site_df$sim.site, color=centered.site_df$sim.site), alpha = .2)+
+ geom_density(aes(x=centered.site_df$GMV, fill=centered.site_df$sim.site, color=centered.site_df$sim.site), alpha=.4) +
+  geom_vline(aes(xintercept=centered.site_sum_df$sd_below.gmv, color=centered.site_sum_df$sim.site)) +
+  geom_vline(aes(xintercept=centered.site_sum_df$sd_above.gmv, color=centered.site_sum_df$sim.site))
+```
+
+## ComBat
+
+And now, to combat the data across the simulated sites!

### 2023-10-12 combat_sandbox.Rmd
simplified functions for getting variances across combat methods

--- 2023-10-11 combat_sandbox.Rmd
@@
 ---
+
+Playing around with different combat configs using the UKB data, which has been assigned to 3 simulated study sites (see `ukb_curation-eda.Rmd`)
 
@@
 ukb_df_sites <- read_csv(file="data/ukb_CN_data_simsites.csv")
-vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
+pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
 ```
@@
 ```{r}
-#regional volume data to harmonize
-rvol.df <- ukb_df_sites %>%
-  dplyr::select(all_of(vol_list_regions))
+#phenotypes to harmonize
+pheno.df <- ukb_df_sites %>%
+  dplyr::select(all_of(pheno_list))
 
@@
 
-nrow(rvol.df) == length(batch)
-nrow(rvol.df) == nrow(covar.mat)
+nrow(pheno.df) == length(batch)
+nrow(pheno.df) == nrow(covar.mat)
 ```
@@
 
+simple combat with no covariate preservation
+```{r}
+cf <- comfam(pheno.df, batch)
+cf$dat.combat
+```
+
 simple combat with linear model for covariates
 ```{r}
-cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale)
+cf.lm <- comfam(pheno.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale)
+cf.lm$dat.combat
+```
+add site
+```{r}
+#add back site
+#cf.lm$dat.combat[, "sim.site"] <- batch
 ```
 
-combat with age as smooth
+Try setting site A (which has even sex distribution) as ref
+
 ```{r}
-cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale)
+cf.lm_refA <- comfam(pheno.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale, ref.batch = "Site_A")
+cf.lm_refA$dat.combat
+# cf.lm_refA$dat.combat[, "sim.site"] <- batch
 ```
+
+combat gam with age as smooth
+```{r}
+cf.gam <- comfam(pheno.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale)
+# cf.gam$dat.combat[, "sim.site"] <- batch
+```
+
+```{r}
+cf.gam_refA <- comfam(pheno.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale, ref.batch = "Site_A")
+# cf.gam_refA$dat.combat[, "sim.site"] <- batch
+```
+
+combat gamlss!
+
+```{r}
+cf.gamlss <- comfam(pheno.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
+              sigma.formula = ~ age_days + sexMale, 
+              control = gamlss.control(trace = FALSE))
+# cf.gamlss$dat.combat[, "sim.site"] <- batch
+```
+
+```{r}
+cf.gamlss_refA <- comfam(pheno.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
+              sigma.formula = ~ age_days + sexMale, 
+              control = gamlss.control(trace = FALSE),
+              ref.batch = "Site_A")
+# cf.gamlss_refA$dat.combat[, "sim.site"] <- batch
+```
+
+Add back batch and covariance
+```{r}
+cf.list <- list(cf, cf.lm, cf.lm_refA, cf.gam, cf.gam_refA, cf.gamlss, cf.gamlss_refA)
+length(cf.list)
+
+combat.df.list <- lapply(cf.list, post_combat_concat, og_df = ukb_df_sites)
+names(combat.df.list) <- c("cf", "cf.lm", "cf.lm_refA", "cf.gam", "cf.gam_refA", "cf.gamlss", "cf.gamlss_refA")
+```
+
+## Compare Variances
+
+Randomly picked `lh_Vol_superiorparietal` to compare
+
+Get variances
+```{r}
+#raw data
+sum_df <- ukb_df_sites %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal),
+            var.lh_v_sp = var(lh_Vol_superiorparietal))
+
+for (i in names(combat.df.list)){
+  df <- combat.df.list[[i]] %>%
+  group_by(sim.site) %>%
+  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
+  assign(paste0(i, "_summary"), df)
+}
+```
+
+compile
+
+```{r}
+var_comp_df <- data.frame(sim.site = c("Site_A", "Site_B", "Site_C"),
+                          raw_data = sum_df$var.lh_v_sp,
+                          cf = cf_summary$var.lh_v_sp,
+                          cf.lm = cf.lm_summary$var.lh_v_sp,
+                          cf.lm_refA = cf.lm_refA_summary$var.lh_v_sp,
+                          cf.gam = cf.gam_summary$var.lh_v_sp,
+                          cf.gam_refA = cf.gam_refA_summary$var.lh_v_sp,
+                          cf.gamlss = cf.gamlss_summary$var.lh_v_sp,
+                          cf.gamlss_refA = cf.gamlss_refA_summary$var.lh_v_sp)
+
+var_comp_long <- var_comp_df %>%
+  pivot_longer(!sim.site, names_to="type", values_to = "variance")
+
+ggplot(var_comp_long) +
+  geom_col(aes(x=type, y=variance, fill=sim.site, color=sim.site), position = "dodge")
+
+ggplot(var_comp_long) +
+  geom_col(aes(x=type, y=variance), position = "dodge") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~sim.site)
+```

### 2023-10-12 ukb_curation-eda.Rmd
cleaning up

@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
+pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
 
@@
 length(pheno_list) #total number of models to run
+saveRDS(pheno_list, file="R_scripts/pheno_list.rds")
 
@@
 ukb_df_totals <- ukb_df %>%
-  mutate(Vol_total=rowSums(select(., .dots=all_of(c(vol_list_regions)))),
-         SA_total=rowSums(select(., .dots=all_of(c(sa_list)))),
-         CT_total=rowSums(select(., .dots=all_of(c(ct_list)))))
+  mutate(Vol_total=rowSums(dplyr::select(., .dots=all_of(c(vol_list_regions)))),
+         SA_total=rowSums(dplyr::select(., .dots=all_of(c(sa_list)))),
+         CT_total=rowSums(dplyr::select(., .dots=all_of(c(ct_list)))))
 
@@
   dplyr::filter(sex == "Male") %>%
-  select(GMV) %>%
+  dplyr::select(GMV) %>%
   var()
@@
   dplyr::filter(sex == "Female") %>%
-  select(GMV) %>%
+  dplyr::select(GMV) %>%
   var()
@@
   #dplyr::filter(sex == "Male") %>%
-  select(GMV) %>%
+  dplyr::select(GMV) %>%
   var()
@@
 ```
-
-## ComBat
-
-And now, to combat the data across the simulated sites!

### 2023-10-12 ukb_curation-eda.Rmd
reorg data

@@
   dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
-#write data
-write_csv(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
+# #write data
+# write_csv(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
 ```
@@
 #write out
-write_csv(ukb_df_sites, file="data/ukb_CN_data_simsites.csv")
+write_csv(ukb_df_sites, file="data/ukb_to_model/ukb_CN_data_simsites.csv")
 ```

### 2023-10-12 ukb_combat.Rmd
write out combatted dfs

--- 2023-10-12 combat_sandbox.Rmd
@@
 cf <- comfam(pheno.df, batch)
-cf$dat.combat
+# cf$dat.combat
 ```
@@
 cf.lm <- comfam(pheno.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale)
-cf.lm$dat.combat
-```
-add site
-```{r}
-#add back site
-#cf.lm$dat.combat[, "sim.site"] <- batch
+# cf.lm$dat.combat
 ```
@@
 cf.lm_refA <- comfam(pheno.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale, ref.batch = "Site_A")
-cf.lm_refA$dat.combat
-# cf.lm_refA$dat.combat[, "sim.site"] <- batch
+# cf.lm_refA$dat.combat
 ```
@@
 cf.gam <- comfam(pheno.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale)
-# cf.gam$dat.combat[, "sim.site"] <- batch
 ```
@@
 cf.gam_refA <- comfam(pheno.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale, ref.batch = "Site_A")
-# cf.gam_refA$dat.combat[, "sim.site"] <- batch
 ```
@@
               control = gamlss.control(trace = FALSE))
-# cf.gamlss$dat.combat[, "sim.site"] <- batch
 ```
@@
               ref.batch = "Site_A")
-# cf.gamlss_refA$dat.combat[, "sim.site"] <- batch
 ```
@@
 names(combat.df.list) <- c("cf", "cf.lm", "cf.lm_refA", "cf.gam", "cf.gam_refA", "cf.gamlss", "cf.gamlss_refA")
+```
+
+write out
+```{r}
+for (i in names(combat.df.list)){
+  df <- combat.df.list[[i]] %>%
+  write_csv(df, file=paste0("data/ukb_to_model/ukb_", as.character(i), "_data.csv"))
+}
 ```

### 2023-10-12 ukb_combat.Rmd
updating to qsub models from multiple .csvs - NEED TO TEST

@@
 ```{r}
-cf.lm_refA <- comfam(pheno.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale, ref.batch = "Site_A")
-# cf.lm_refA$dat.combat
+cf.lm-refA <- comfam(pheno.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale, ref.batch = "Site_A")
+# cf.lm-refA$dat.combat
 ```
@@
 ```{r}
-cf.gam_refA <- comfam(pheno.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale, ref.batch = "Site_A")
+cf.gam-refA <- comfam(pheno.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale, ref.batch = "Site_A")
 ```
@@
 ```{r}
-cf.gamlss_refA <- comfam(pheno.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
+cf.gamlss-refA <- comfam(pheno.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
               sigma.formula = ~ age_days + sexMale, 
@@
 ```{r}
-cf.list <- list(cf, cf.lm, cf.lm_refA, cf.gam, cf.gam_refA, cf.gamlss, cf.gamlss_refA)
+cf.list <- list(cf, cf.lm, cf.lm-refA, cf.gam, cf.gam-refA, cf.gamlss, cf.gamlss-refA)
 length(cf.list)
@@
 combat.df.list <- lapply(cf.list, post_combat_concat, og_df = ukb_df_sites)
-names(combat.df.list) <- c("cf", "cf.lm", "cf.lm_refA", "cf.gam", "cf.gam_refA", "cf.gamlss", "cf.gamlss_refA")
+names(combat.df.list) <- c("cf", "cf.lm", "cf.lm-refA", "cf.gam", "cf.gam-refA", "cf.gamlss", "cf.gamlss-refA")
 ```
@@
 for (i in names(combat.df.list)){
-  df <- combat.df.list[[i]] %>%
-  write_csv(df, file=paste0("data/ukb_to_model/ukb_", as.character(i), "_data.csv"))
+  df <- combat.df.list[[i]]
+  write_csv(df, file=paste0("data/ukb_to_model/ukb_", i, "_data.csv"))
 }
@@
                           cf.lm = cf.lm_summary$var.lh_v_sp,
-                          cf.lm_refA = cf.lm_refA_summary$var.lh_v_sp,
+                          cf.lm-refA = cf.lm-refA_summary$var.lh_v_sp,
                           cf.gam = cf.gam_summary$var.lh_v_sp,
-                          cf.gam_refA = cf.gam_refA_summary$var.lh_v_sp,
+                          cf.gam-refA = cf.gam-refA_summary$var.lh_v_sp,
                           cf.gamlss = cf.gamlss_summary$var.lh_v_sp,
-                          cf.gamlss_refA = cf.gamlss_refA_summary$var.lh_v_sp)
+                          cf.gamlss-refA = cf.gamlss-refA_summary$var.lh_v_sp)
 

### 2023-10-12 R_scripts/fit_basic_mod.R

@@ -21,4 +21,6 @@ model <- gamlss(formula = as.formula(paste0(pheno,"~ poly(age_days, 3) + sex")),
                      family = BCCG, data=df, trace = FALSE)
 
 #SAVE
-saveRDS(model,paste0(save_path, "/", pheno, "_mod.rds"))
+csv_name <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(as.character(args[1])))
+
+saveRDS(model,paste0(save_path, "/", pheno, "_", csv_name, "_mod.rds"))

### 2023-10-12 qsub_basic_gamlss.sh

@@ -1,19 +1,26 @@
 #!/bin/bash
 
 #Fit GAMLSS growth chart models for each phenotype
+#pass as argument EITHER path to .csv containing data OR directory of .csvs to iterate across
 
 #######################################################################################
 # SET PATHS
 base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
 #base=/Users/megardn/Desktop/BGD_Repos/combat_biovar #base path (local)
-data_csv=$base/data/ukb_CN_data_agefilt.csv #path to csv with CLEANED DATA (no duplicates, etc)
+#data_csv=$base/data/ukb_CN_data_agefilt.csv #path to csv with CLEANED DATA (no duplicates, etc)
 pheno_path=$base/pheno_lists #path to .txt files listing phenotypes (global & regional)
 mod_script=$base/R_scripts/fit_basic_mod.R #path to .R script
 
 #cd to where .Rprofile is stored (for propper package loading)
 cd $base
 #######################################################################################
-
+# GET ARGS
+if [ $# -lt 1 ]
+then
+	echo 1>&2 "$0: provide path to data (either single .csv or dir)"
+	exit 2
+fi
+#######################################################################################
 # MAKE DIRECTORIES 
 
 #study dir
@@ -40,20 +47,51 @@ if ! [ -d $gamlss_dir ]
 #######################################################################################
 # FIT MODELS
 
-#iterate through measure types (vol, SA, CT, global vols) for correct global corrections
-for list in "$pheno_path"/*
-do
-	echo $list
-	#iterate through list of phenotypes
-	while read -r pheno
+#if input .csv
+if [ -f $1 ]
+then
+	csv_name=$(basename $1 .csv)
+	echo "Pulling data from $csv_name"
+	#iterate through measure types (vol, SA, CT, global vols) for correct global corrections
+	for list in "$pheno_path"/*
 	do
-		#write bash script
-		bash_script=$bash_dir/${pheno}_basic_fit.sh
-		touch $bash_script
-		echo "Rscript --save $mod_script $data_csv $pheno $gamlss_dir" > $bash_script
+		echo "Modelling phenotypes in $list"
+		#iterate through list of phenotypes
+		while read -r pheno
+		do
+			#write bash script
[…]

### 2023-10-14 R_scripts/fit_basic_mod.R
troubleshooting and updating filenaming

@@ -22,5 +22,6 @@ model <- gamlss(formula = as.formula(paste0(pheno,"~ poly(age_days, 3) + sex")),
 
 #SAVE
 csv_name <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(as.character(args[1])))
+csv_rename <- gsub("_", "-", csv_name)
 
-saveRDS(model,paste0(save_path, "/", pheno, "_", csv_name, "_mod.rds"))
+saveRDS(model,paste0(save_path, "/", pheno, "_", csv_rename, "_mod.rds"))

### 2023-10-14 qsub_basic_gamlss.sh

@@ -5,8 +5,8 @@
 
 #######################################################################################
 # SET PATHS
-base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
-#base=/Users/megardn/Desktop/BGD_Repos/combat_biovar #base path (local)
+#base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
+base=/Users/megardn/Desktop/BGD_Repos/combat_biovar #base path (local)
 #data_csv=$base/data/ukb_CN_data_agefilt.csv #path to csv with CLEANED DATA (no duplicates, etc)
 pheno_path=$base/pheno_lists #path to .txt files listing phenotypes (global & regional)
 mod_script=$base/R_scripts/fit_basic_mod.R #path to .R script
@@ -51,6 +51,7 @@ if ! [ -d $gamlss_dir ]
 if [ -f $1 ]
 then
 	csv_name=$(basename $1 .csv)
+	csv_name=${csv_name//_/\-}
 	echo "Pulling data from $csv_name"
 	#iterate through measure types (vol, SA, CT, global vols) for correct global corrections
 	for list in "$pheno_path"/*
@@ -65,7 +66,7 @@ then
 			echo "Rscript --save $mod_script $1 $pheno $gamlss_dir" > $bash_script
 
 			#qsub bash script
-			qsub -N ${pheno}.${csv_name} -o $bash_dir/${pheno}_${csv_name}_out.txt -e $bash_dir/${pheno}_${csv_name}_err.txt $bash_script
+			#qsub -N ${pheno}.${csv_name} -o $bash_dir/${pheno}_${csv_name}_out.txt -e $bash_dir/${pheno}_${csv_name}_err.txt $bash_script
 
 		done < $list
 	done
@@ -75,6 +76,7 @@ then
 	for csv_file in "$1"/*
 	do
 		csv_name=$(basename $csv_file .csv)
+		csv_name=${csv_name//_/\-}
 		echo "Pulling data from $csv_name"
 		#iterate through measure types (vol, SA, CT, global vols) for correct global corrections
 		for list in "$pheno_path"/*
@@ -89,7 +91,7 @@ then
 				echo "Rscript --save $mod_script $csv_file $pheno $gamlss_dir" > $bash_script
 
 				#qsub bash script
[…]

### 2023-10-17 proof-of-concept_mods.Rmd
switch to modeling numeric sexMale for drop1() functionality

@@
 ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
-ukb_df_age.filt <- read_csv(file="data/ukb_CN_data_agefilt.csv")
+ukb_df_age.filt <- read_csv(file="data/ukb_to_model/ukb_CN_data_simsites.csv")
 
@@
 ```{r}
-read_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic/gamlss_objs" #path to bfp models
+read_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic/" #path to bfp models
 ```
@@
 ```
+
+Troubleshooting drop1
+
+```{r}
+model.test <- readRDS("~/Desktop/BGD_Repos/combat_biovar/ukb_basic/lh_SA_superiortemporal_ukb-CN-data-simsites_mod.rds")
+drop1(model.test, what = "sigma")
+model.test.drop <- gamlss(formula = lh_SA_superiortemporal ~ poly(age_days, 3) + sex,
+                       sigma.formula = ~ poly(age_days, 3),
+                       nu.formula = ~ 1,
+                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
+drop1(model.test.drop, what = "sigma")
+
+ukb_df_age.filt <- ukb_df_age.filt %>%
+  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
+                                        sex == "Female" ~ 0,
+                                        TRUE ~ NA)))
+model.test.refit <- gamlss(formula = lh_SA_superiortemporal ~ poly(age_days, 3) + sexMale,
+                       sigma.formula = ~ poly(age_days, 3) + sexMale,
+                       nu.formula = ~ 1,
+                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
+drop1(model.test.refit, what = "sigma", scope = "sexMale")
+```

### 2023-10-17 ukb_combat.Rmd
troubleshooting and updating filenaming
switch to modeling numeric sexMale for drop1() functionality

--- 2023-10-12 ukb_combat.Rmd
@@
 ```{r}
-ukb_df_sites <- read_csv(file="data/ukb_CN_data_simsites.csv")
+ukb_df_sites <- read_csv(file="data/ukb_to_model/ukb_CN_data_simsites.csv")
 pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
@@
 
-#covariates - using model.matrix to dummycode sex as numeric
-covar.mat <- model.matrix( ~ age_days + sex, data=ukb_df_sites)
+# #covariates - using model.matrix to dummycode sex as numeric - 
+# covar.mat <- model.matrix( ~ age_days + sexMale, data=ukb_df_sites) 
+#testing with df now that mods are being fit with sexMale
+covar.mat <- ukb_df_sites %>%
+  dplyr::select(age_days, sexMale)
 

### 2023-10-17 ukb_curation-eda.Rmd
switch to modeling numeric sexMale for drop1() functionality

@@
          sex = as.factor(sex),
-         sex_int = as.numeric(case_when(sex == "Male" ~ 0,
-                                        sex == "Female" ~ 1)),
-         sex.age = sex_int * age_days) %>%
+         sexMale = as.numeric(case_when(sex == "Male" ~ 1,
+                                        sex == "Female" ~ 0,
+                                        TRUE ~ NA)),
+         sex.age = sexMale * age_days) %>%
   na.omit() #1 subj has 1 missing val

### 2023-10-17 R_scripts/fit_basic_mod.R

@@ -14,8 +14,8 @@ pheno <- as.character(args[2])
 save_path <- args[3]
 
 #FIT BASIC MODEL
-model <- gamlss(formula = as.formula(paste0(pheno,"~ poly(age_days, 3) + sex")),
-                     sigma.formula = ~ poly(age_days, 3) + sex,
+model <- gamlss(formula = as.formula(paste0(pheno,"~ poly(age_days, 3) + sexMale")),
+                     sigma.formula = ~ poly(age_days, 3) + sexMale,
                      nu.formula = ~ 1,
                      control = gamlss.control(n.cyc = 200), 
                      family = BCCG, data=df, trace = FALSE)

### 2023-10-20 proof-of-concept_mods.Rmd
setup to save csv with drop1 results as each model is fit

@@
 
-Visualize
+###Visualize
 
@@
 
+##Troubleshooting
+
 Troubleshooting drop1
@@
 drop1(model.test.refit, what = "sigma", scope = "sexMale")
+
+model.test.refit2 <- gamlss(formula = lh_SA_superiortemporal ~ poly(age_days, 3) + sex,
+                       sigma.formula = ~ poly(age_days, 3) + sex,
+                       nu.formula = ~ 1,
+                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
+
+drop.test.obj<-drop1(model.test.refit2, what = "nu")
+df <- drop.test.obj %>%
+  as.data.frame() %>%
+  mutate(what=attributes(drop.test.obj)$heading[2]) %>%
+  add_rownames(var = "term")
+
+f <- drop1_all(model.test.refit2, name="TEST")
 ```

### 2023-10-20 R_scripts/fit_basic_mod.R

@@ -7,6 +7,8 @@ library(gamlss)
 library(dplyr)
 library(data.table)
 
+source("R_scripts/gamlss_helper_funs.R")
+
 #GET ARTS
 args <- commandArgs(trailingOnly = TRUE)
 df <- fread(args[1], stringsAsFactors = TRUE, na.strings = "")
@@ -25,3 +27,7 @@ csv_name <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(as.charact
 csv_rename <- gsub("_", "-", csv_name)
 
 saveRDS(model,paste0(save_path, "/", pheno, "_", csv_rename, "_mod.rds"))
+
+#DROP1 SIGNIFICANCE TESTING
+drop1_df <- drop1_all(model, name=pheno)
+fwrite(drop1_df, file=paste0(save_path, "/", pheno, "_", csv_rename, "_drop.csv"))

### 2023-10-20 R_scripts/gamlss_helper_funs.R

@@ -16,8 +16,13 @@ library(broom.mixed)
 # get.moment.formula()
 ################
 
-drop1_all <- function(mod_obj, list = c("mu", "sigma")){
-  n <- deparse(substitute(mod_obj))
+drop1_all <- function(mod_obj, list = c("mu", "sigma"), name = NA){
+  if (is.na(name)){
+    n <- deparse(substitute(mod_obj))
+  } else {
+    n <- name
+  }
+  
   df <- data.frame("Model"=character(),
                    "Term"=character(),
                    "Df"=double(),
@@ -27,7 +32,7 @@ drop1_all <- function(mod_obj, list = c("mu", "sigma")){
                    "Moment"=character())
   
   for (m in list){
-    print(m)
+    print(paste("drop1 from", m))
     drop.obj<-drop1(mod_obj, what = m)
     df2 <- drop.obj %>%
       as.data.frame() %>%

### 2023-10-23 R_scripts/combat_apply.R
prepping for qsubbing combat

@@ -51,7 +51,7 @@ sa_list <- readRDS(file="R_scripts/SA_list.rds")
 ct_list <- readRDS(file="R_scripts/CT_list.rds")
 #combine
 list_of_feature_lists <- list(vol_list_global, vol_list_regions, sa_list, ct_list)
-names(list_of_feature_lists) <- c("Vol_Global", "Vol_Regional", "SA", "CT")
+names(list_of_feature_lists) <- c("VolGlob", "VolReg", "SA", "CT")
 
 #DEF BATCHES & COVARS
 batch <- as.factor(raw.df[[batch.col]])
@@ -94,7 +94,7 @@ for (l in list_of_feature_lists){
   }
   
   #save cf.obj
-  saveRDS(cf.obj, file=paste0(save_path, "/", save_name,"_", names(list_of_feature_lists[i]), "_cf_obj.rds"))
+  saveRDS(cf.obj, file=paste0(save_path, "/combat_objs/", save_name,"_", names(list_of_feature_lists[i]), "_cf_obj.rds"))
   
   #row number
   cf.obj.df <- cf.obj$dat.combat %>%

### 2023-10-23 qsub_basic_gamlss.sh

@@ -72,7 +72,7 @@ then
 #if input directory of .csvs
 elif [ -d $1 ]
 then
-	for csv_file in "$1"/*
+	for csv_file in "$1"/*.csv #select JUST csvs - need to test this
 	do
 		csv_name=$(basename $csv_file .csv)
 		csv_name=${csv_name//_/\-}

### 2023-10-23 test-combat.sh

@@ -1,6 +1,91 @@
 #!/bin/bash
-cd /Users/megardn/Desktop/BGD_Repos/combat_biovar
+# Apply ComFam() with different arguments to specified dataset
 
-Rscript --save /Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/combat_apply.R \
-/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv \
-"sim.site" /Users/megardn/Desktop/BGD_Repos/combat_biovar/data "TEST_combat_script" "age_days,sexMale" 'gam, formula = y ~ s(age_days) + sexMale, ref.batch = "Site_A"'
+#######################################################################################
+# SET PATHS
+img=/cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif #singularity image
+
+base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
+#base=/Users/megardn/Desktop/BGD_Repos/combat_biovar #base path (local)
+data_csv=$base/data/ukb_CN_data_agefilt.csv #path to csv with CLEANED DATA (no duplicates, etc)
+#pheno_path=$base/pheno_lists #path to .txt files listing phenotypes (global & regional) - may need to pass as arg. on cubic/w singularity
+cf_script=$base/R_scripts/combat_apply.R #path to .R script
+save_path=$base/data/ukb_to_model
+
+cd $base #to source functions correctly
+#######################################################################################
+# MAKE DIRECTORIES 
+#study dir
+study_dir=$base/ukb_basic
+if ! [ -d $study_dir ]
+	then
+	mkdir $study_dir
+	fi
+
+#qsub script & outputs
+bash_dir=$study_dir/combat_qsub_scripts
+if ! [ -d $bash_dir ]
+	then
+	mkdir $bash_dir
+	fi
+
+#combat output obj dir
+if ! [ -d ${save_path}/combat_objs ]
+	then
+	mkdir ${save_path}/combat_objs
+	fi
+#######################################################################################
+#SET COVAR COLS
+covar_list="age_days,sexMale"
+#SET BATCH COL
+batch="sim.site"
+
+#LIST POSSIBLE CONFIGS
+config_list="cf cf.lm cf.lm_refA cf.gam cf.gam_refA cf.gamlss cf.gamlss_refA"
[…]

### 2023-10-23 ukb_curation-eda.Rmd
testing additional permutations of sim.site groupings

@@
 ```
+
+## Simulate Sites - Permutations
+
+```{r}
+set.seed(1010101)
+sim.site.list <- c("Site_A", "Site_B", "Site_C")
+n_female <- table(ukb_df_age.filt$sex)["Female"]
+n_male <- table(ukb_df_age.filt$sex)["Male"]
+
+# Sample probabilities
+male_prob <- c(0.33, 0.0825, 0.5875)
+female_prob <- c(0.33, 0.5875, 0.0825)
+
+# Create a list to store data frames
+df_list <- list()
+
+for(i in 1:5) {
+  new_col_name <- paste0("SimSite", i)
+  # Separate samples for males and females
+  sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
+  sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)
+
+  new_df <- ukb_df_age.filt %>%
+    mutate(!!new_col_name := as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))%>%
+    dplyr::select(!!new_col_name)
+
+  # Store the data frame in the list
+  df_list[[i]] <- new_df
+}
+
+# Combine data frames into a single data frame
+ukb_df_sites_perm <- bind_cols(ukb_df_age.filt, !!!df_list)
+
+# View the final dataframe
+print(ukb_df_sites_perm)
+#names(ukb_df_sites_perm)
+```
+```{r}
+sim.perm.long <- ukb_df_sites_perm %>%
+  dplyr::select(sex, age_days, contains("SimSite")) %>%
+  pivot_longer(!c(sex, age_days), names_to = "permutation", names_prefix = "SimSite", values_to = "site")
+
+prop <- sim.perm.long %>%
+  group_by(permutation, site, sex) %>%
+  summarize(Count = n(), avg_age = mean(age_days)) %>%
+  group_by(permutation, site) %>%
+  mutate(Proportion = Count / sum(Count))
+
+summary.prop.male <- prop %>%
+  dplyr::filter(sex=="Male") %>%
+  group_by(site) %>%
+  summarize(avg_prop = mean(Proportion), min_prop = min(Proportion), max_prop = max(Proportion))
+
+summary.prop.male <- prop %>%
+  group_by(site, sex) %>%
+  summarize(avg_age_days = mean(avg_age), min_age = min(avg_age), max_age = max(avg_age))
+```
+
+```{r}
+#write out
+write_csv(ukb_df_sites_perm, file="data/ukb_CN_data_simsites_perm.csv")
+```

### 2023-10-25 proof-of-concept_mods.Rmd
fixing gam output df

@@
 
-## Significance Testing
+## Review Significance testing
 
-scp'd models from cubic to parse
+scp'd csvs from cubic
 
 ```{r}
-read_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic/" #path to bfp models
+drop1.df <- fread("ukb_basic/drop1_tests.csv", stringsAsFactors = TRUE) %>%
+  dplyr::select(-V1)
+sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE)%>%
+  dplyr::select(-V1)
+sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE)%>%
+  dplyr::select(-V1)
 ```
@@
 ```{r}
-#iterate across gamlss objects
-model.files <- list.files(path = read_path, pattern = "mod.rds", full.names = TRUE)
-
-sigma.df <- data.frame("file_path" = model.files)
-sigma.df <- sigma.df %>%
-  mutate(region = sub("_mod\\.rds$", "", basename(as.character(file_path))))
-
-#test significance of sex term in sigma
-sigma.df$sig.sex_drop1.p <- lapply(sigma.df$file_path, get.and.drop1.p, moment="sigma", term="sex")
-
-#extract values of sex term from summary table
-summary.list <- lapply(sigma.df$file_path, get.summary)
-summary.df <- bind_rows(summary.list)
-sigma.sex.df <- summary.df %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  mutate(pheno_cat = as.factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA)),
-    sig = case_when(p.value < 0.05 ~ TRUE,
-                    p.value >= 0.05 ~ FALSE),
-    sig_bf.corr = case_when(pheno %in% vol_list_global & p.value < (0.05/length(vol_list_global)) ~ TRUE,
-                            pheno %in% vol_list_global & p.value >= (0.05/length(vol_list_global)) ~ FALSE,
-                            !(pheno %in% vol_list_global) & p.value < (0.05/length(ct_list)) ~ TRUE,
-                            !(pheno %in% vol_list_global) & p.value >= (0.05/length(ct_list)) ~ FALSE,
-                            TRUE ~ NA),
-    label = sub("_[^_]*_", "_", pheno)) #for plotting
-
 #add more info from parcellations
@@
 
-sigma.sex.df2 <- left_join(sigma.sex.df, dk.parc, by="label")
+sex.df <- left_join(sex.df, dk.parc, by="label")
+sum.df <- left_join(sum.df, dk.parc, by="label")
 ```
+
+Does the number of regions w significant age effects on mu change depending on combat version?
+
+```{r}
+drop1.df %>%
+  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
+  group_by(Dataset) %>%
+  summarize(n_sig_regions = sum(sig.drop_bf.corr),
+            n_regions = n())
+```
+
+No.
+
+Does the number of regions w significant sex effects on mu change depending on combat version?
+
+```{r}
+drop1.df %>%
+  dplyr::filter(Term=="sexMale" & Moment == "mu") %>%
+  group_by(Dataset) %>%
+  summarize(n_sig_regions = sum(sig.drop_bf.corr),
+            n_regions = n())
+
+sum.df %>%
+  dplyr::filter(term=="sexMale" & parameter == "mu") %>%
+  group_by(dataset) %>%
+  summarize(n_sig_regions = sum(sig.sum_bf.corr),
+            n_regions = n())
+```
+
+Slightly. Only combat w/o covariates or lm combat without a ref site are noticeably off. Consistent across drop1 and summary outputs
 
@@
 
-```{r}
-table(sigma.sex.df2$sig)
-table(sigma.sex.df2$sig_bf.corr)
-```
+Non-combatted results
 
 ```{r, fig.width=11}
-sigma.sex.df2 %>%
-  dplyr::filter(sig_bf.corr == TRUE) %>%
+sex.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset ==	"ukb-CN-data-simsites") %>%
   ggplot() +
@@
 
-sigma.sex.df2 %>%
-  dplyr::filter(sig_bf.corr == TRUE) %>%
+sex.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
   ggplot() +
@@
 ```{r}
-sigma.sex.df2 %>%
-  dplyr::filter(sig_bf.corr == TRUE) %>%
+sex.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
   group_by(pheno_cat) %>%
@@
 ```
+
+```{r}
+# Your input string
+input_string <- "xxxx ref.batch = yyyy"
+
+# Check if the input string contains "ref.batch =" using a regular expression
+if (grepl("ref\\.batch\\s*=\\s*", input_string)) {
+  # Your code for splitting the string
+  # Split the string into two parts
+  split_string <- unlist(strsplit(input_string, "ref.batch", fixed = TRUE))
+
+  # The split_string will contain two elements
+  part1 <- split_string[1]
+  part2 <- gsub("=", "", split_string[2])
+
+  # Trim any leading/trailing white spaces in part2
+  part2 <- trimws(part2)
+
+  # Output the two parts
+  print(part1)
+  print(part2)
+} else {
+  cat("The input string does not contain 'ref.batch ='\n")
+}
+```

### 2023-10-25 R_scripts/fit_basic_gam.R

@@ -26,4 +26,8 @@ saveRDS(model,paste0(save_path, "/", pheno, "_", csv_rename, "_gam.rds"))
 
 #SIGNIFICANCE TESTING
 df <- anova.gam(model)$pTerms.table
+df <- data.frame(term = row.names(df), df) %>%
+  mutate(dataset = csv_rename,
+         phenotype = pheno)
+
 fwrite(df, file=paste0(save_path, "/", pheno, "_", csv_rename, "_gam_sig.csv"))

### 2023-10-26 proof-of-concept_mods.Rmd
visuals

@@
 ct_list <- readRDS(file="R_scripts/CT_list.rds")
+pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
 ```
@@
 
-## Review Significance testing
+## Review Results
 
@@
   dplyr::select(-V1)
-sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE)%>%
+sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE) %>%
   dplyr::select(-V1)
@@
 
-Does the number of regions w significant age effects on mu change depending on combat version?
-
-```{r}
-drop1.df %>%
-  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
-  group_by(Dataset) %>%
-  summarize(n_sig_regions = sum(sig.drop_bf.corr),
-            n_regions = n())
-```
-
-No.
-
-Does the number of regions w significant sex effects on mu change depending on combat version?
-
-```{r}
-drop1.df %>%
-  dplyr::filter(Term=="sexMale" & Moment == "mu") %>%
-  group_by(Dataset) %>%
-  summarize(n_sig_regions = sum(sig.drop_bf.corr),
-            n_regions = n())
-
-sum.df %>%
-  dplyr::filter(term=="sexMale" & parameter == "mu") %>%
-  group_by(dataset) %>%
-  summarize(n_sig_regions = sum(sig.sum_bf.corr),
-            n_regions = n())
-```
-
-Slightly. Only combat w/o covariates or lm combat without a ref site are noticeably off. Consistent across drop1 and summary outputs
-
-###Visualize
-
-Non-combatted results
+## Raw Data Results
 
@@
   facet_wrap(~pheno_cat) +
+  theme_void() +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("t-statistic for effect of male sex on sigma")
+```
+
+### Combat Effects
+
+#### Site Effects
+Loading combatted data (scp'd from cubic: `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/data/ukb_to_model/*.csv' ./data/ukb_to_model`) to see how values change with different combat methods
+```{r}
+data.csvs <- list.files(path = "data/ukb_to_model", pattern = ".csv", full.names = TRUE)
+combined_data <- data.frame()
+
+for (file in data.csvs) {
+  # Read each CSV file
+  data <- fread(file)
+
+  # Add a "Source_File" column with the file name
+  data <- data %>%
+    mutate(Source_File = as.factor(basename(file)))
+
+  # Bind the data to the combined dataframe
+  combined_data <- bind_rows(combined_data, data, .id = "File_ID")
+}
+
+table(combined_data$Source_File)
+```
+
+```{r}
+var.df <- combined_data %>%
+  group_by(Source_File, sim.site) %>%
+  summarize(across(all_of(pheno_list), var))
+
+var.long <- pivot_longer(var.df, cols=c(-Source_File, -sim.site), names_to = "feature", values_to = "variance")
+```
+
+```{r}
+var.long %>%
+  dplyr::filter(!(feature %in% vol_list_global)) %>%
+  ggplot() +
+  geom_density(aes(x=variance, fill=sim.site), alpha = 0.4, position="identity") +
+  facet_wrap(~Source_File)
+
+var.long %>%
+  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.lm_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
+  ggplot() +
+  geom_density(aes(x=variance, fill=Source_File), alpha = 0.5, position="identity") +
+  facet_wrap(~sim.site)
+```
+
+```{r}
+var.by.sex <- combined_data %>%
+  group_by(Source_File, sex) %>%
+  summarize(across(all_of(pheno_list), var))
+
+var.by.sex.long <- pivot_longer(var.by.sex, cols=c(-Source_File, -sex), names_to = "feature", values_to = "variance")
+```
+
+```{r}
+var.by.sex.long %>%
+  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.lm_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
+  ggplot() +
+  geom_density(aes(x=variance, fill=sex), alpha = 0.5, position="identity") +
+  facet_wrap(~Source_File)
+```
+
+#### Mu
+
+Does the number of regions w significant age effects on mu change depending on combat version?
+
+```{r}
+drop1.df %>%
+  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
+  group_by(Dataset) %>%
+  summarize(n_sig_regions = sum(sig.drop_bf.corr),
+            n_regions = n())
+```
+
+No.
+
+Does the number of regions w significant sex effects on mu change depending on combat version?
+
+```{r}
+drop1.df %>%
+  dplyr::filter(Term=="sexMale" & Moment == "mu") %>%
+  group_by(Dataset) %>%
+  summarize(n_sig_regions = sum(sig.drop_bf.corr),
+            n_regions = n())
+
+sum.df %>%
+  dplyr::filter(term=="sexMale" & parameter == "mu") %>%
+  group_by(dataset) %>%
+  summarize(n_sig_regions = sum(sig.sum_bf.corr),
+            n_regions = n())
+```
+
+Slightly. Only combat w/o covariates or lm combat without a ref site are noticeably off. Consistent across drop1 and summary outputs
+
+```{r}
+sex.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "mu" & pheno_cat != "Global Volume") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = t_stat)) +
+  scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_void() +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("t-statistic for effect of male sex on mu")
+```
+
+#### Sigma
+
+```{r}
+sex.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
+  ggplot() +
+  geom_col(aes(x=label, y=t_stat, fill=dataset)) +
+  facet_wrap(~pheno_cat, scales = "free_x") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  #theme_void() +
+  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("t-statistic for effect of male sex on sigma")
+
+sex.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
+  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites") %>%
+  ggplot() +
+  geom_col(aes(x=label, y=t_stat, fill=dataset), position = "dodge") +
+  facet_wrap(~pheno_cat, scales = "free_x") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  #theme_void() +
+  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("t-statistic for effect of male sex on sigma")
+
+sex.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = t_stat)) +
+  scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
   theme_void() +

### 2023-10-26 proof-of-concept_mods.Rmd

@@
   dplyr::select(-V1)
-sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE) %>%
+sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE)%>%
   dplyr::select(-V1)

### 2023-10-27 proof-of-concept_mods.Rmd
gam parser, visualize results

--- 2023-10-26 proof-of-concept_mods.Rmd
@@
 
+# Analysis Outline
 ## Load Data
@@
 
+## Basic Models
 Designate *basic* model formula
@@
 
-Check sex's significance in variance
-
-```{r}
-K <- drop1(basic.m1, what="sigma", scope = "sex")
-class(K)
-K["Pr(Chi)"]
-K$"Pr(Chi)"[2]
-
-test.mu <- drop1(basic.m1, what="mu", scope = "sex")
-test.sig <- drop1(basic.m1, what="sigma", scope = "sex")
-
-#write for testing
-saveRDS(basic.m1, file="/Users/megardn/Desktop/BGD_Repos/combat_biovar/gamlss_basic_test_mod.rds")
-
-get.and.drop1.p("/Users/megardn/Desktop/BGD_Repos/combat_biovar/gamlss_basic_test_mod.rds", moment="sigma", term="sex")
-
-test.model.list <- c()
-test.model.list[[1]] <- basic.m1
-lapply(test.model.list, drop1, what="mu", scope = "sex")
-```
-
-## Review Results
+# Review Results
 
@@
 
-No.
+No.ln
 
@@
 ```
-
-##Troubleshooting
-
-Troubleshooting drop1
-
-```{r}
-model.test <- readRDS("~/Desktop/BGD_Repos/combat_biovar/ukb_basic/lh_SA_superiortemporal_ukb-CN-data-simsites_mod.rds")
-drop1(model.test, what = "sigma")
-model.test.drop <- gamlss(formula = lh_SA_superiortemporal ~ poly(age_days, 3) + sex,
-                       sigma.formula = ~ poly(age_days, 3),
-                       nu.formula = ~ 1,
-                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
-drop1(model.test.drop, what = "sigma")
-
-ukb_df_age.filt <- ukb_df_age.filt %>%
-  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
-                                        sex == "Female" ~ 0,
-                                        TRUE ~ NA)))
-model.test.refit <- gamlss(formula = lh_SA_superiortemporal ~ poly(age_days, 3) + sexMale,
-                       sigma.formula = ~ poly(age_days, 3) + sexMale,
-                       nu.formula = ~ 1,
-                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
-drop1(model.test.refit, what = "sigma", scope = "sexMale")
-
-model.test.refit2 <- gamlss(formula = lh_SA_superiortemporal ~ poly(age_days, 3) + sex,
-                       sigma.formula = ~ poly(age_days, 3) + sex,
-                       nu.formula = ~ 1,
-                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
-
-drop.test.obj<-drop1(model.test.refit2, what = "nu")
-df <- drop.test.obj %>%
-  as.data.frame() %>%
-  mutate(what=attributes(drop.test.obj)$heading[2]) %>%
-  add_rownames(var = "term")
-
-f <- drop1_all(model.test.refit2, name="TEST")
-```
-
-```{r}
-# Your input string
-input_string <- "xxxx ref.batch = yyyy"
-
-# Check if the input string contains "ref.batch =" using a regular expression
-if (grepl("ref\\.batch\\s*=\\s*", input_string)) {
-  # Your code for splitting the string
-  # Split the string into two parts
-  split_string <- unlist(strsplit(input_string, "ref.batch", fixed = TRUE))
-
-  # The split_string will contain two elements
-  part1 <- split_string[1]
-  part2 <- gsub("=", "", split_string[2])
-
-  # Trim any leading/trailing white spaces in part2
-  part2 <- trimws(part2)
-
-  # Output the two parts
-  print(part1)
-  print(part2)
-} else {
-  cat("The input string does not contain 'ref.batch ='\n")
-}
-```

### 2023-10-27 R_scripts/combat_apply.R

@@ -5,7 +5,7 @@
 ## 1. dataframe of data to be combatted
 ## 2. name of column containing batch identifier, or path to a csv containing batch identifier
 ## 3. path to save output csv
-## 4. filename for output csv
+## 4. combat config name to append to output csv's filename
 ## 5. logical indicator of whether to pass data fit_ukb_basic.R
 ## 6. list of columns to be included as covariates (OPTIONAL)
 ## 7. Additional comfam() arguments, including model, formula, ref.batch, ... (OPTIONAL)
@@ -13,7 +13,6 @@
 set.seed(12345)
 
 #LOAD PACKAGES
-#library(devtools) #may need to install ComBatFamily
 library(data.table) 
 library(readr)
 library(ggplot2) 
@@ -41,7 +40,7 @@ if (endsWith(batch.arg, '.csv')){
 }
 
 save_path <- as.character(args[3]) #path to save outputs
-save_name <- as.character(args[4])
+config_name <- as.character(args[4])
 pass <- as.logical(args[5])
 
 # see if optional args provided for combatting
@@ -161,9 +160,14 @@ final.df <- cf.merged %>%
 # }
 
 ##########################################################################
-
 #WRITE OUT
-datafile <- paste0(save_path, "/", save_name, "_data.csv")
+
+#extract csv name from input data
+csv_basename <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(as.character(args[1])))
+csv_basename <- gsub("_", "-", csv_basename)
+
+#append config name
+datafile <- paste0(save_path, "/", csv_basename, "_", config_name, "_data.csv")
 fwrite(final.df, file=datafile)
 
 ##########################################################################

### 2023-10-27 R_scripts/gam_parser.R

@@ -1,11 +1,13 @@
 
+#compile outputs from fit_basic_gam.R
+
 set.seed(12345)
 
 #LOAD PACKAGES
-library(gamlss)
+library(mgcv)
 library(dplyr)
 library(data.table)
-library(tibble)
+library(broom)
 
 source("R_scripts/gamlss_helper_funs.R")
 
@@ -21,14 +23,14 @@ args <- commandArgs(trailingOnly = TRUE)
 read_path <- args[1] #path to bfp models
 save_path <- args[2] #path to save outputs
 
-#iterate across gamlss objects
-model.files <- list.files(path = read_path, pattern = "mod.rds", full.names = TRUE)
+#iterate across gam objects
+model.files <- list.files(path = read_path, pattern = "gam.rds", full.names = TRUE)
 
 print("First Few Modfiles")
 print(model.files[1:3])
 
 #extract values of sex term from summary table
-summary.list <- lapply(model.files, get.gamlss.summary)
+summary.list <- lapply(model.files, get.summary)
 
 #dataframe
 summary.df <- bind_rows(summary.list)
@@ -48,7 +50,7 @@ summary.df <- summary.df %>%
 
 
 #remaining cols
-sigma.sex.df <- summary.df %>%
+summary.df2 <- summary.df %>%
   mutate(
     dataset = sub(".*_(.*)", "\\1", mod_name),
     pheno_cat = as.factor(case_when(
@@ -66,28 +68,27 @@ sigma.sex.df <- summary.df %>%
                             TRUE ~ NA),
     label = sub("_[^_]*_", "_", pheno)) #for plotting
 
-write.csv(sigma.sex.df, file=paste0(save_path, "/gamlss_summary.csv"))
+write.csv(summary.df2, file=paste0(save_path, "/gam_summary.csv"))
 
-#iterate across drop1 csvs
+#iterate across anova csvs
 csv.files <- list.files(path = read_path, pattern = ".csv", full.names = TRUE)
-drop1.df <- do.call(rbind, lapply(csv.files, fread))
+anova.df <- do.call(rbind, lapply(csv.files, fread))
 
-drop1.df <- drop1.df %>%
-  rename(drop1.pval = "Pr(Chi)") %>%
-  mutate(sig.drop_bf.corr = case_when(Model %in% vol_list_global & drop1.pval < (0.05/length(vol_list_global)) ~ TRUE,
-                                      Model %in% vol_list_global & drop1.pval >= (0.05/length(vol_list_global)) ~ FALSE,
[…]

### 2023-10-27 R_scripts/gamlss_helper_funs.R

@@ -4,6 +4,7 @@ library(dplyr)
 library(ggplot2)
 library(data.table)
 library(broom.mixed)
+library(broom)
 
 #note - some of these functions do not seem to work on CUBIC, as they rely on saving environment variables
 
@@ -116,7 +117,7 @@ get.moment.formula <- function(gamlss.rds.file, moment) {
   moment.form <- formula(gamlss.obj, what = moment)
 }
 
-get.summary<- function(gamlss.rds.file) {
+get.gamlss.summary<- function(gamlss.rds.file) {
   #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
   gamlss.rds.file <- as.character(gamlss.rds.file)
   gamlss.obj <- readRDS(gamlss.rds.file)
@@ -124,6 +125,18 @@ get.summary<- function(gamlss.rds.file) {
     as.data.frame() %>%
     rename(t_stat = statistic)
   sum.table$mod_name <- sub("_mod\\.rds$", "", basename(gamlss.rds.file)) #append model name
+  return(sum.table)
+}
+
+#more generic version of get.gamlss.summary()
+get.summary<- function(rds.file) {
+  #USE WITH sapply(USE.NAMES=TRUE) to keep file names with values!
+  rds.file <- as.character(rds.file)
+  obj <- readRDS(rds.file)
+  sum.table <- broom::tidy(obj) %>%
+    as.data.frame() %>%
+    rename(t_stat = statistic)
+  sum.table$mod_name <- sub("\\.rds$", "", basename(rds.file)) #append model name (agnostic of ending str)
   return(sum.table)
 }
 

### 2023-10-27 R_scripts/perm_runs.R

@@ -1,61 +1,0 @@
-set.seed(1010101)
-
-#######################################################
-#SIMULATE SITE ASSIGNMENTS
-#######################################################
-sim.site.list <- c("Site_A", "Site_B", "Site_C")
-n_female <- table(ukb_df_age.filt$sex)["Female"]
-n_male <- table(ukb_df_age.filt$sex)["Male"]
-
-# Sample probabilities
-male_prob <- c(0.33, 0.0825, 0.5875)
-female_prob <- c(0.33, 0.5875, 0.0825)
-
-# Create a list to store data frames
-df_list <- list()
-
-for(i in 1:5) {
-  # Separate samples for males and females
-  sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
-  sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)
-  
-  new_df <- ukb_df_age.filt %>%
-    mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))
-  
-  # Store the data frame in the list
-  df_list[[i]] <- new_df
-}
-#######################################################
-#RUN COMBAT
-#######################################################
-
-cf_method_names <- c("cf", "cf.lm", "cf.lm_refA", "cf.gam", "cf.gam_refA", "cf.gamlss", "cf.gamlss_refA")
-
-for(cf.name in cf_method_names) {
-  #apply combat to list of dataframes
-  cf_data_list <- lapply() #APPLY FUNCTION THAT RUNS COMBAT WITH DIFFERENT METHODS
-  
-  #make new list to store data from each method
-  assign(cf.name, df_list, envir = .GlobalEnv)
-}
-
-#######################################################
-#FIT GAMLSS
-#######################################################
-
-all_data_dfs <- c(df_list, cf, cf.lm, cf.lm_refA, cf.gam, cf.gam_refA, cf.gamlss, cf.gamlss_refA)
-names(all_data_dfs) <- c("raw_data", cf_method_names)
-
-for (data_list in all_data_dfs){
-  print(name(data_list))
-  drop1_df <- data.frame() #empty dataframe to store drop1 outputs for each phenotype/permutation
-  sum_df <- data.frame() #empty dataframe to store summary outputs for each phenotype/permutation
-  
-  for(i in length(data_list)){
-    #lapply function to fit gamlss on each phenotype
-    
-    # return drop1 and summary outputs for mu & sigma
-  }
-  
-}
-

### 2023-10-27 R_scripts/permute_sites.R

@@ -11,7 +11,8 @@ library(stringr)
 #GET ARGUMENTS
 args <- commandArgs(trailingOnly = TRUE)
 df <- fread(args[1], stringsAsFactors = TRUE, na.strings = "")
-save_path <- fread(args[1], stringsAsFactors = TRUE, na.strings = "")
+save_path <- fread(args[2], stringsAsFactors = TRUE, na.strings = "")
+pass <- as.logical(args[3]) #whether or not to automatically pass to qsub_combat.sh
 
 #extract csv name
 csv_basename <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(as.character(args[1])))
@@ -41,6 +42,15 @@ for(i in 1:5) {
   
   #save out csv w permutation number in filename
   n_perm <- str_pad(i, 3, pad = "0")
-  fname <- paste0(csv_basename, "_perm-key_", n_perm)
-  fwrite(new_df, file=paste0(save_path, "/", fname, ".csv"))
+  fname <- paste0(save_path, "/", csv_basename, "_perm-key_", n_perm, ".csv")
+  fwrite(new_df, file=fname)
+  
+  #automatically send to combat (and on to gamlss fitting)
+  if (isTRUE(pass)) {
+    print("submitting jobs to run combat")
+    cmd <- paste('/cbica/home/gardnerm/combat-biovar/qsub_combat.sh -c', fname, '-p TRUE')
+    cat(cmd)
+    system(cmd)
+  }
+  
 }

### 2023-10-30 proof-of-concept_mods.Rmd
looking at effects on variance

@@
 
+## QUESTIONS
+- effect size for poly(age_days, 3) in gam/gamlss
+- next steps abstract v paper
+- abs values of centile differences
+- z scores instead of centiles
+
 ## Setup
@@
 ```{r setup, include=FALSE, warning = FALSE, message = FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
 if(!require('pacman')) {
@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg)
+pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr)
 
@@
 ```{r}
-basic.m1 <- gamlss(formula = GMV ~ poly(age_days, 3) + sex,
-                       sigma.formula = ~ poly(age_days, 3) + sex,
+basic.m1 <- gamlss(formula = GMV ~ poly(age_days, 3) + sexMale,
+                       sigma.formula = ~ poly(age_days, 3) + sexMale,
                        nu.formula = ~ 1,
@@
 ```{r}
-plot_singlestudy_centiles(basic.m1, "GMV", ukb_df_age.filt, "sex")
+#plot_singlestudy_centiles(basic.m1, "GMV", ukb_df_age.filt, "sexMale")
 ```
@@
 
-scp'd csvs from cubic
+scp'd csvs from cubic (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/*.csv' ./ukb_basic`)
 
@@
 
-## Raw Data Results
+### Raw Data Results
 
@@
   ggplot() +
-  geom_density(aes(x=variance, fill=sim.site), alpha = 0.4, position="identity") +
-  facet_wrap(~Source_File)
+  geom_density(aes(x=variance, fill=Source_File), alpha = 0.4, position="identity") +
+  facet_wrap(~sim.site)
 
 var.long %>%
-  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.lm_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
+  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" |Source_File == "ukb_CN_data_simsites.csv" |  Source_File == "cf.gamlss_data.csv" )) %>%
   ggplot() +
@@
   facet_wrap(~sim.site)
+
+#Source_File == "cf.lm_data.csv" | 
 ```
@@
 var.by.sex.long %>%
-  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.lm_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
+  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
   ggplot() +
@@
 ```
+
+Simple t test for sex effects across combat types
+```{r}
+sex_ttest_df <- data.frame("Source_File" = levels(combined_data$Source_File))
+
+for(pheno in pheno_list){
+  # New column name
+  new_col_name <- paste0(pheno, ".p")
+
+  #run t-tests
+  out <- lapply(split(combined_data, factor(combined_data$Source_File)), function(x)t.test(x[[pheno]] ~ x[["sex"]], paired=FALSE)$p.value)
+  stopifnot(names(out) == levels(combined_data$Source_File)) #make sure results are in correct order
+
+  #append to dataframe
+  sex_ttest_df[[new_col_name]] <- unlist(unname(out))
+}
+```
+
+```{r}
+sex_ttest_long <- sex_ttest_df %>% pivot_longer(!Source_File, values_to="p.value", names_to="pheno") %>%
+  mutate(pheno = sub(".p$", "", pheno),
+         sig_bfcor = case_when(pheno %in% vol_list_global & p.value < (0.05/length(vol_list_global)) ~ TRUE,
+                            pheno %in% vol_list_global & p.value >= (0.05/length(vol_list_global)) ~ FALSE,
+                            !(pheno %in% vol_list_global) & p.value < (0.05/length(ct_list)) ~ TRUE,
+                            !(pheno %in% vol_list_global) & p.value >= (0.05/length(ct_list)) ~ FALSE,
+                            TRUE ~ NA))
+sex_ttest_long %>%
+  group_by(Source_File) %>%
+  summarize(n_sig_regions = sum(sig_bfcor),
+            n_regions = n())
+```
+
+Basically same results gam/gamlss
 
@@
 
-No.ln
+No.
 
@@
 ```
+
+Trying with ultra-strict threshold to see if results change:
+
+```{r}
+drop1.df <- drop1.df %>%
+  mutate(sig.sum.strict = case_when(drop1.pval < (0.05/length(pheno_list)) ~ TRUE,
+                            drop1.pval >= (0.05/length(pheno_list)) ~ FALSE,
+                            TRUE ~ NA))
+
+drop1.df %>%
+  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
+  group_by(Dataset) %>%
+  summarize(n_sig_regions = sum(sig.sum.strict),
+            n_regions = n())
+```
+
+No difference
 
@@
 ```
+Check number of regions w significant sex effects in sigma
+```{r}
+sex.sigma.counts <- sex.df %>%
+  dplyr::filter(term=="sexMale" & parameter == "sigma") %>%
+  group_by(dataset, pheno_cat) %>%
+  summarize(n_sig_regions = sum(sig.drop_bf.corr),
+            n_regions = n())
+sex.sigma.counts
+```
+
+More variation
+```{r}
+ggplot(sex.sigma.counts) +
+  geom_col(aes(x=dataset, y=n_sig_regions, fill=dataset)) +
+  facet_wrap(~pheno_cat) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+```
+
+Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex)
+
+```{r}
+sigma.df <- sum.df %>%
+  dplyr::filter(parameter == "sigma") %>%
+  mutate(var_est = exp(estimate))
+
+skim(sigma.df$var_est)
+
+sigma.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & term =="sexMale") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_histogram(aes(x=var_est, fill = dataset)) +
+  facet_wrap(~pheno_cat, scales= "free")
+
+sigma.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = var_est)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_void() +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("Beta weight for effect of male sex on variance")
+```
+
+#### Centiles
+
+Do centile scores change based on combat version?
+```{r}
+pred.csvs <- list.files(path = "ukb_basic", pattern = "_predictions.csv", full.names = TRUE)
+combined_preds <- data.frame()
+
+for (file in pred.csvs) {
+  # Read each CSV file
+  data <- fread(file)
+
+  # Add a "Source_File" column with the file name
+  data <- data %>%
+    mutate(Source_File = as.factor(basename(file))) %>%
+    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
+
+  # Bind the data to the combined dataframe
+  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
+}
+
+table(combined_preds$dataset)
+names(combined_preds)
+```
+calc how much centiles change
+```{r}
+cent_diffs <- combined_preds %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  arrange(dataset) %>%
+  group_by(participant) %>%
+  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
+  ungroup()
+
+diff_vol_glob <- sapply(vol_list_global, function(x) paste0("diff_", x))
+
+mean.cent.diffs <- cent_diffs %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("diff_"), mean))
+
+max.cent.diffs <- cent_diffs %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("diff_"), max))
+```
+
+Plot average differences in centiles
+```{r}
+mean.diffs.long <- mean.cent.diffs %>%
+  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+ggplot(mean.diffs.long) +
+  geom_density(aes(x=centile_diff, fill=dataset), alpha=0.5) +
+  facet_grid(dataset~pheno_cat)
+
+mean.diffs.long %>%
+  dplyr::filter(dataset== "cf.gam" | dataset == "cf.gam_refA" | dataset == "cf.gamlss" | dataset == "cf.gamlss_refA") %>%
+  ggplot() +
+  geom_density(aes(x=centile_diff, fill=dataset), alpha=0.5) +
+  scale_x_continuous(labels = label_comma())
+
+mean.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(centile_diff = centile_diff*100) %>% #for visualization purposes
+  ggplot() +
+  geom_histogram(aes(x=centile_diff, fill=dataset), alpha=0.5) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_x_continuous(labels = label_comma()) + facet_wrap(~dataset, scales = "free_y")
+```
+
+Look at max difference per phenotype
+```{r}
+max.diffs.long <- max.cent.diffs %>%
+  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+max.diffs.long %>%
+  #dplyr::filter(dataset== "cf.lm_refA" | dataset == "cf.gam_refA" | dataset == "cf.gamlss" | dataset == "cf.gamlss_refA") %>%
+  ggplot() +
+  geom_density(aes(x=max_centile_diff, fill=dataset), alpha=0.5) +
+  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat)
+
+max.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  ggplot() +
+  geom_density(aes(x=max_centile_diff, fill=dataset), alpha=0.5) +
+  scale_x_continuous(labels = label_comma()) + facet_wrap(~dataset, scales = "free_y")
+```
+
+Test for differences in distribution
+```{r}
+summary(aov(centile_diff ~ dataset, data=mean.diffs.long))
+
+attach(mean.diffs.long)
+pairwise.t.test(x=centile_diff, g=dataset, p.adj = "bonf")
+detach()
+```
+
+Seems like gamlss methods are significantly better than the other combat methods and not significantly different from raw centile scores (bonferroni-adjusted)
+
+Do the different combat methods introduce any kind of bias in centile scores?
+
+```{r}
+mean.cent.diffs.by.sex <- cent_diffs %>%
+  group_by(dataset, sex) %>%
+  summarize(across(starts_with("diff_"), mean)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+mean.cent.diffs.by.sex %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(centile_diff = centile_diff *100) %>%
+ggplot() +
+  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
+  facet_wrap(~dataset, scales = "free_y")
+
+max.cent.diffs.by.sex <- cent_diffs %>%
+  group_by(dataset, sex) %>%
+  summarize(across(starts_with("diff_"), max)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+max.cent.diffs.by.sex %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(max_centile_diff = max_centile_diff *100) %>%
+ggplot() +
+  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
+  facet_wrap(~dataset, scales = "free_y")
+```
+
+Other methods are fairly consistently underestimating male centiles and overestimating female centiles.
+
+#### GAMs
+scp'd csvs from cubic (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/*.csv' ./ukb_basic`)
+
+```{r}
+anova.gam <- fread("ukb_basic/anova_gam_tests.csv", stringsAsFactors = TRUE) %>%
+  dplyr::select(-V1)
+gam.sex.df <- fread("ukb_basic/sex_summary_gam.csv", stringsAsFactors = TRUE) %>%
+  dplyr::select(-V1)
+gam.sum.df <- fread("ukb_basic/gam_summary.csv", stringsAsFactors = TRUE)%>%
+  dplyr::select(-V1)
+```
+
+Does the number of regions w significant age effects on mean change depending on combat version?
+
+```{r}
+anova.gam %>%
+  dplyr::filter(term=="poly(age_days, 3)") %>%
+  group_by(dataset) %>%
+  summarize(n_sig_regions = sum(sig.drop_bf.corr),
+            n_regions = n())
+```
+
+No.
+
+```{r, fig.height=10, fig.width=15}
+anova.gam %>%
+  dplyr::filter(term=="poly(age_days, 3)") %>%
+  rename(f_val = "F") %>%
+  ggplot() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  geom_col(aes(x=phenotype, y=f_val, fill=dataset), position = "dodge")
+```
+
+```{r}
+anova.gam_wide.f <- anova.gam %>%
+  dplyr::filter(term=="poly(age_days, 3)") %>%
+  rename(f_val = "F") %>%
+  pivot_wider(id_cols= phenotype, names_from=dataset, values_from=f_val)
+
+anova.gam.f_diff <- anova.gam_wide.f %>%
+  rename_all(~ gsub("-", "_", .)) %>%
+  mutate(across(starts_with("cf"), ~ . - ukb_CN_data_simsites)) %>%
+  dplyr::select(!ukb_CN_data_simsites)%>%
+  pivot_longer(!phenotype, names_to = "dataset", values_to = "f_diff") %>%
+  mutate(dataset = as.factor(dataset))
+
+summary(aov(f_diff ~ dataset, data=anova.gam.f_diff))
+
+attach(anova.gam.f_diff)
+pairwise.t.test(x=f_diff, g=dataset, p.adj = "bonf")
+detach()
+
+ggplot(anova.gam.f_diff) + 
+  geom_density(aes(x=f_diff, fill=dataset), alpha=0.5)
+
+anova.gam.f_diff %>%
+  dplyr::filter(dataset == "cf.gamlss_data" | dataset == "cf.gam_data" | dataset == "cf.gam_refA_data" | dataset == "cf.gamlss_refA_data") %>%
+ggplot() + 
+  geom_density(aes(x=f_diff, fill=dataset), alpha=0.5)
+```
+
+So the effect sizes (if this F is the correct F to use as effect size) are closer to the "ground truth" for gamlss than non-gamlss combat methods.
+However, similar to other results, only combat w/o covariates and combat with linear covariates are significantly worse.
+
+Does the number of regions w significant sex effects on mean change depending on combat version?
+
+```{r}
+anova.gam %>%
+  dplyr::filter(term=="sexMale") %>%
+  group_by(dataset) %>%
+  summarize(n_sig_regions = sum(sig.drop_bf.corr),
+            n_regions = n())
+```
+
+Not much, only for combat w/out covariates or linear covariates
+
+```{r, fig.heigt=15}
+gam.sum.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & term == "sexMale") %>%
+  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites"| dataset == "cf.lm-refA-data") %>%
+  ggplot() +
+  geom_col(aes(x=label, y=t_stat, fill=dataset), position="dodge") +
+  facet_wrap(~pheno_cat, scales = "free_x") + 
+  #theme_cowplot()+
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  #theme_void() +
+  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("t-statistic for effect of male sex on mean")
+
+gam.sum.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & term == "sexMale") %>%
+  ggplot() +
+    geom_density(aes(x=t_stat, fill=dataset), alpha=0.4) +
+  facet_wrap(~dataset)
+```

### 2023-10-31 proof-of-concept_mods.Rmd
fixing sGMV pheno search

@@
   facet_wrap(~pheno_cat, scales= "free")
+```
 
+```{r, fig.height=15, fig.widt=12}
 sigma.df %>%
@@
   facet_grid(dataset~pheno_cat) + 
-  theme_void() +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  theme_bw(base_size=20) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
   ggtitle("Beta weight for effect of male sex on variance")
+```
+
+```{r, fig.height=6, fig.widt=8}
+sigma.df %>%
+  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale") %>%
+  mutate(var_est.sig = case_when(sig.sum_bf.corr == TRUE ~ var_est,
+                                 sig.sum_bf.corr == FALSE ~ 0)) %>%
+  #group_by(dataset) %>%
+  ggplot() +
+  geom_col(aes(x=dataset, y = var_est.sig, fill=dataset), position="dodge") +
+  #scale_fill_gradient2() +
+  facet_grid(~pheno) + 
+  theme_bw(base_size=16) +
+  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  ggtitle("Beta weight for effect of male sex on global volumes' variance")
+
+sigma.df.glob <- sigma.df %>%
+  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale")
+table(sigma.df.glob$dataset, sigma.df.glob$pheno)
+
+sigma.df %>%
+  dplyr::filter(pheno == "sGMV")
 ```

### 2023-10-31 R_scripts/gamlss_parser.R

@@ -38,14 +38,13 @@ summary.df <- summary.df %>%
   mutate(pheno = sapply(mod_name, function(b) {
     pheno <- ""
     for (a in pheno_list) {
-      if (grepl(a, b)) {
+      if (grepl(paste0("^", a), b)) {
         pheno <- a
         break
       }
     }
     return(pheno)
   }))
-
 
 #remaining cols
 sigma.sex.df <- summary.df %>%

### 2023-10-31 qsub_centiles.sh

@@ -7,7 +7,7 @@ img=/cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif #singularity ima
 base=/cbica/home/gardnerm/combat-biovar #base path (cubic) - also path that output csvs will be saved to
 r_script=$base/R_scripts/fit_centiles.R #path to .R script
 csv_path=$base/data/ukb_to_model #path to data csvs
-mod_path=$vase/ukb_basic/gamlss_objs #path to gamlss .rds objs
+mod_path=$base/ukb_basic/gamlss_objs #path to gamlss .rds objs
 #######################################################################################
 cd $base #to source functions correctly
 #######################################################################################

### 2023-10-31 proof-of-concept_mods.Rmd
centile diffs by sex

@@
   dplyr::select(-V1)
-sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE)%>%
+sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
   dplyr::select(-V1)
@@
   ggplot() +
-  geom_histogram(aes(x=var_est, fill = dataset)) +
+  geom_histogram(aes(x=var_est, fill = dataset), color="black", alpha=0.5, position = "identity") +
+  theme_linedraw() +
   facet_wrap(~pheno_cat, scales= "free")
@@
 
-```{r, fig.height=6, fig.widt=8}
+```{r, fig.height=5, fig.widt=5}
 sigma.df %>%
@@
   facet_grid(~pheno) + 
-  theme_bw(base_size=16) +
+  theme_bw(base_size=12) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  theme(legend.position="none") +
   #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
@@
   dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale")
-table(sigma.df.glob$dataset, sigma.df.glob$pheno)
-
-sigma.df %>%
-  dplyr::filter(pheno == "sGMV")
+table(sigma.df.glob$dataset, sigma.df.glob$sig.sum_bf.corr)
 ```
@@
 table(combined_preds$dataset)
-names(combined_preds)
 ```
@@
   group_by(participant) %>%
-  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
+  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
   ungroup()
 
-diff_vol_glob <- sapply(vol_list_global, function(x) paste0("diff_", x))
+diff_vol_glob <- sapply(vol_list_global, function(x) paste0("abs.diff_", x))
 
@@
   group_by(dataset) %>%
-  summarize(across(starts_with("diff_"), mean))
+  summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup()
 
@@
   group_by(dataset) %>%
-  summarize(across(starts_with("diff_"), max))
+  summarize(across(starts_with("abs.diff_"), max)) %>%
+  ungroup()
+#names(max.cent.diffs)
 ```
@@
 mean.diffs.long <- mean.cent.diffs %>%
-  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
   mutate(pheno_cat = as.factor(case_when(
@@
 
-ggplot(mean.diffs.long) +
-  geom_density(aes(x=centile_diff, fill=dataset), alpha=0.5) +
+mean.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+ggplot() +
+  geom_density(aes(x=centile_abs.diff, fill=dataset), alpha=0.5) +
   facet_grid(dataset~pheno_cat)
@@
   ggplot() +
-  geom_density(aes(x=centile_diff, fill=dataset), alpha=0.5) +
+  geom_density(aes(x=centile_abs.diff, fill=dataset), alpha=0.5) +
   scale_x_continuous(labels = label_comma())
@@
   dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(centile_diff = centile_diff*100) %>% #for visualization purposes
+  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visualization purposes
   ggplot() +
-  geom_histogram(aes(x=centile_diff, fill=dataset), alpha=0.5) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  scale_x_continuous(labels = label_comma()) + facet_wrap(~dataset, scales = "free_y")
+  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", binwidth=0.1, alpha=0.5, position="identity") +
+  theme_linedraw() +
+  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_x_continuous(labels = label_comma()) +
+  ggtitle("Mean absolute differences in centile scores across ComBat \nprocessing pipelines for each phenotype")
 ```
@@
 Look at max difference per phenotype
-```{r}
+```{r, fig.height=6, fig.width=7}
 max.diffs.long <- max.cent.diffs %>%
-  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
   mutate(pheno_cat = as.factor(case_when(
@@
 max.diffs.long %>%
-  #dplyr::filter(dataset== "cf.lm_refA" | dataset == "cf.gam_refA" | dataset == "cf.gamlss" | dataset == "cf.gamlss_refA") %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(max_centile_abs.diff = max_centile_abs.diff*100) %>% #for visuals
   ggplot() +
-  geom_density(aes(x=max_centile_diff, fill=dataset), alpha=0.5) +
-  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat)
+  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity") +
+  theme_linedraw(base_size = 12) +
+  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
+  ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
+```
 
+```{r}
 max.diffs.long %>%
+  mutate(max_centile_abs.diff = max_centile_abs.diff*100) %>% #for visuals
   dplyr::filter(dataset != "ukb_CN") %>%
   ggplot() +
-  geom_density(aes(x=max_centile_diff, fill=dataset), alpha=0.5) +
-  scale_x_continuous(labels = label_comma()) + facet_wrap(~dataset, scales = "free_y")
+  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity") +
+  theme_linedraw(base_size = 12) +
+  scale_x_continuous(labels = label_comma())+
+  ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
 ```
@@
 ```{r}
-summary(aov(centile_diff ~ dataset, data=mean.diffs.long))
+summary(aov(centile_abs.diff ~ dataset, data=mean.diffs.long))
 
 attach(mean.diffs.long)
-pairwise.t.test(x=centile_diff, g=dataset, p.adj = "bonf")
+pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "fdr")
 detach()
@@
 
-Seems like gamlss methods are significantly better than the other combat methods and not significantly different from raw centile scores (bonferroni-adjusted)
+Seems like gamlss methods are significantly better than the other combat methods and not significantly different from raw centile scores (BH corrected)
 
@@
   group_by(dataset, sex) %>%
-  summarize(across(starts_with("diff_"), mean)) %>%
+  summarize(across(starts_with("abs.diff_"), mean)) %>%
   ungroup() %>%
-  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
   mutate(pheno_cat = as.factor(case_when(
@@
   group_by(dataset, sex) %>%
-  summarize(across(starts_with("diff_"), max)) %>%
+  summarize(across(starts_with("abs.diff_"), max)) %>%
   ungroup() %>%
-  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
   mutate(pheno_cat = as.factor(case_when(
@@
 
+No shift in abs. differences, looking at overall diff.
+```{r, fig.height=7, fig.width=8}
+cent.diffs2 <- combined_preds %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  arrange(dataset) %>%
+  group_by(participant) %>%
+  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
+  ungroup()
+```
+
+```{r}
+max.cent.diffs.by.sex2 <- cent.diffs2 %>%
+  group_by(dataset, sex) %>%
+  summarize_at(vars(starts_with("diff")), list(max_value = ~.[which.max(abs(.))])) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+max.cent.diffs.by.sex2 %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(max_centile_diff = max_centile_diff *100) %>%
+ggplot() +
+  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
+  facet_wrap(~dataset, scales="free")
+```
 Other methods are fairly consistently underestimating male centiles and overestimating female centiles.
+
+```{r}
+mean.cent.diffs.by.sex2 <- cent.diffs2 %>%
+  group_by(dataset, sex) %>%
+  summarize(across(starts_with("diff_"), mean)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+mean.cent.diffs.by.sex2 %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(centile_diff = centile_diff *100) %>%
+ggplot() +
+  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
+  facet_wrap(~dataset, scales="free")
+```
 

### 2023-11-02 proof-of-concept_mods.Rmd
plots

@@
 - effect size for poly(age_days, 3) in gam/gamlss
-- next steps abstract v paper
-- abs values of centile differences
-- z scores instead of centiles
+- z scores instead of centiles!!!!
 
@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr)
+pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr, paletteer)
 
@@
 sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1)
+  dplyr::select(-V1) %>%
+  mutate(dataset = gsub("-data", "", dataset)) %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb-CN-simsites"))
 ```
@@
 
-Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex)
+Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex).
 
@@
   dplyr::filter(parameter == "sigma") %>%
-  mutate(var_est = exp(estimate))
+  mutate(var_est = exp(estimate),
+         )
 
@@
 
-```{r, fig.height=15, fig.widt=12}
-sigma.df %>%
+```{r, fig.height=15, fig.width=12}
+beta.plot <- sigma.df %>%
   dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
   ggtitle("Beta weight for effect of male sex on variance")
+
+ggsave("male_variance_beta.jpeg", beta.plot, width=12, height=15)
 ```
@@
 table(combined_preds$dataset)
+#skim(combined_preds)
 ```
@@
       pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+mean.diffs.long <- left_join(mean.diffs.long, dk.parc, by="label")
+```
 
+```{r}
 mean.diffs.long %>%
@@
   ggtitle("Mean absolute differences in centile scores across ComBat \nprocessing pipelines for each phenotype")
+
+cent.diffs.plot <- mean.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.6, position="identity") +
+  theme_linedraw(base_size = 12) +
+  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
+  paletteer::scale_fill_paletteer_d("Polychrome::light") +
+  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
+ggsave("mean_centile_diffs.jpeg", cent.diffs.plot)
+```
+
+```{r, fig.height=15, fig.width=12}
+mean.diffs.long %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = centile_abs.diff)) +
+  scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_void() +
+  theme_bw(base_size=20) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("Mean difference in centiles across ComBat configurations")
 ```
@@
   ggplot() +
-  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity") +
+  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.6, position="identity") +
   theme_linedraw(base_size = 12) +
   scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
+  paletteer::scale_fill_paletteer_d("Polychrome::light") +
   ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
@@
   ggplot() +
-  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity") +
+  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity", binwidth = 0.5) +
   theme_linedraw(base_size = 12) +
@@
 ```
+T test of mean diff. between sex w/in each dataset
+```{r}
+t.test.by.sex <- mean.cent.diffs.by.sex2 %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  group_by(dataset) %>%
+  summarise(t_statistic = t.test(centile_diff ~ sex)$statistic,
+            p_value = t.test(centile_diff ~ sex)$p.value) %>%
+  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
+         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
+                             p_val_fdr.corr >= 0.05 ~ FALSE))
+
+t.test.by.sex
+```
+
+Both gam methods and lm combat with ref both end up inducing a bias in how males and females centiles are shifted.
+
+#### Z-Score
+calc how much z-scores change
+
+```{r}
+z_diffs <- combined_preds %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref = "ukb_CN")) %>%
+  arrange(dataset) %>%
+  group_by(participant) %>%
+  mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
+  ungroup()
+```
+
+```{r}
+#mean
+mean.z.diffs <- z_diffs %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup()
+mean.z.long <- mean.z.diffs %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno = sub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+mean.z.long <- left_join(mean.z.long, dk.parc, by="label")
+
+#max
+max.z.diffs <- z_diffs %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), max)) %>%
+  ungroup()
+max.z.long <- max.z.diffs %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno = sub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+```
+
+```{r}
+mean.z.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  ggplot() +
+  geom_histogram(aes(x=z_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity") +
+  theme_linedraw() +
+  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_x_continuous(labels = label_comma()) +
+  ggtitle("Mean absolute differences in Z-scores across ComBat \nprocessing pipelines for each phenotype")
+```
+
+```{r, fig.height=15, fig.width=12}
+mean.z.long %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = z_abs.diff)) +
+  scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_void() +
+  theme_bw(base_size=20) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("Mean difference in z scores across ComBat configurations")
+```
+
+```{r}
+max.z.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  ggplot() +
+  geom_histogram(aes(x=max_z_abs.diff, fill=dataset), color="black", binwidth = 0.025, alpha=0.5, position="identity") +
+  theme_linedraw() +
+  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_x_continuous(labels = label_comma()) +
+  ggtitle("Mean absolute differences in Z-scores across ComBat \nprocessing pipelines for each phenotype")
+```
+
+```{r, fig.height=15, fig.width=12}
+max.z.long %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = max_z_abs.diff)) +
+  scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_void() +
+  theme_bw(base_size=20) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
+  ggtitle("Maximum difference in z scores across ComBat configurations")
+```
 

### 2023-11-03 proof-of-concept_mods.Rmd
plots, OHBM abstract analyses

@@
 Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex).
+Also getting mean variance.
 
@@
   mutate(var_est = exp(estimate),
-         )
+         var_mean = exp(moment.mean),
+         beta_ratio = var_est/var_mean)
 
@@
   facet_wrap(~pheno_cat, scales= "free")
+```
+
+```{r, fig.height=10, fig.width=8}
+proportion_beta_plot <- sigma.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = beta_ratio)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_bw(base_size=15) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
+  ggtitle("Proportion of effect of male sex on variance")
 ```
@@
 
+##### Sex-Biases
+
 Do the different combat methods introduce any kind of bias in centile scores?
@@
 ```
+
 Other methods are fairly consistently underestimating male centiles and overestimating female centiles.
@@
 ```
+
 T test of mean diff. between sex w/in each dataset
@@
 
+Test for differences in distribution
+```{r}
+summary(aov(z_abs.diff ~ dataset, data=mean.z.long))
+
+attach(mean.z.long)
+pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr")
+detach()
+```
+
+##### Sex-Biases
+
+Test for sex-specific differences in mean z-score offset
+
+```{r}
+z_diffs2 <- combined_preds %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  arrange(dataset) %>%
+  group_by(participant) %>%
+  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
+  ungroup()
+
+mean.z.diffs.by.sex2 <- z_diffs2 %>%
+  group_by(dataset, sex) %>%
+  summarize(across(starts_with("diff_"), mean)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("diff_"), values_to="z_diff", names_to = "pheno", names_prefix = "diff_") %>%
+  mutate(pheno = gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+```
+t-test
+```{r}
+t.test.by.sex_z <- mean.z.diffs.by.sex2 %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  group_by(dataset) %>%
+  summarise(t_statistic = t.test(z_diff ~ sex)$statistic,
+            p_value = t.test(z_diff ~ sex)$p.value) %>%
+  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
+         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
+                             p_val_fdr.corr >= 0.05 ~ FALSE))
+
+t.test.by.sex_z
+```
+
+Same results as centiles.
+
 #### GAMs
@@
 ```
+
+#### OHBM Abstract
+
+rerunning stats above w/o ref-site combat versions (to account for different FDR correction) per Andrew's suggestion
+
+Test for differences in centile offset distributions
+```{r}
+mean.diffs.long_filt <- mean.diffs.long %>%
+  filter(!str_detect(dataset, "_refA"))
+
+summary(aov(centile_abs.diff ~ dataset, data=mean.diffs.long_filt))
+
+attach(mean.diffs.long_filt)
+pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "fdr")
+detach()
+```
+
+Testing for differences in Z offset distributions
+
+```{r}
+mean.z.long_filt <- mean.z.long %>%
+  filter(!str_detect(dataset, "_refA"))
+
+summary(aov(z_abs.diff ~ dataset, data=mean.z.long_filt))
+
+attach(mean.z.long_filt)
+pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr")
+detach()
+```
+
+Test for sex-biases in centile offsets
+
+Test for sex-biases
+
+```{r}
+mean.z.diffs.by.sex2 %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  filter(!str_detect(dataset, "_refA")) %>%
+  group_by(dataset) %>%
+  summarise(t_statistic = t.test(z_diff ~ sex)$statistic,
+            p_value = t.test(z_diff ~ sex)$p.value) %>%
+  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
+         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
+                             p_val_fdr.corr >= 0.05 ~ FALSE))
+
+mean.cent.diffs.by.sex2 %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  filter(!str_detect(dataset, "_refA")) %>%
+  group_by(dataset) %>%
+  summarise(t_statistic = t.test(centile_diff ~ sex)$statistic,
+            p_value = t.test(centile_diff ~ sex)$p.value) %>%
+  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
+         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
+                             p_val_fdr.corr >= 0.05 ~ FALSE))
+```
+
+##### Plots
+
+Beta weight? or proportion?
+
+```{r}
+beta.plot2 <- sigma.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
+  filter(!str_detect(dataset, "-refA")) %>%
+  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"))) %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = var_est)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_bw(base_size=20) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
+  ggtitle("Beta weight for effect of male sex on variance")
+
+beta.plot2
+```
+
+```{r}
+proportion_beta_plot2 <- sigma.df %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
+  filter(!str_detect(dataset, "-refA")) %>%
+  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"))) %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = beta_ratio)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_bw(base_size=15) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
+  ggtitle("Proportion of effect of male sex on variance")
+
+proportion_beta_plot2
+```
+
+Centile plot
+
+```{r}
+cent.diffs.plot2 <- mean.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  filter(!str_detect(dataset, "_refA")) %>%
+  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.4, position="identity") +
+  theme_linedraw(base_size = 12) +
+  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
+  #paletteer::scale_fill_paletteer_d("Polychrome::light") +
+  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
+
+cent.diffs.plot2
+```

### 2023-11-03 ukb_curation-eda.Rmd
save out age-filt df to csv for permuting sites on with qsub

@@
 #table(cn_data$study)
+names(cn_data)
 ```
@@
   dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
-# #write data
-# write_csv(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
+#write data
+fwrite(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
 ```
@@
 ## EDA
+
+```{r}
+#percent female
+r <- table(ukb_df_age.filt$sex)
+
+(r["Female"]/n_unique(ukb_df_age.filt$participant))*100
+summary(ukb_df_age.filt$age_yrs)
+(table(ukb_df_age.filt$fs_version)["FS6_T1T2"])/(n_unique(ukb_df_age.filt$participant))*100
+```
 
@@
 # View the final dataframe
-print(ukb_df_sites_perm)
 #names(ukb_df_sites_perm)
 ```
+
 ```{r}

### 2023-11-06 proof-of-concept_mods.Rmd
comparing abs centile diffs w/in pheno, troubleshooting missing cf-cf.lm t-test results

@@
 ```{r}
-summary(aov(centile_abs.diff ~ dataset, data=mean.diffs.long))
-
-attach(mean.diffs.long)
-pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "fdr")
+#get centile errors for each subject
+cent.diffs2 <- combined_preds %>%
+  dplyr::filter(!str_detect(dataset, "_refA")) %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  arrange(dataset) %>%
+  group_by(participant) %>%
+  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "diff_{.col}")) %>%
+  ungroup() %>%
+  dplyr::filter(dataset != "ukb_CN") #drop ref
+
+# cf.gmvdiff <- cent.diffs2 %>%
+#   dplyr::filter(dataset == "cf") %>%
+#   dplyr::select(diff_GMV)
+# cf.lm.gmvdiff <- cent.diffs2 %>%
+#   dplyr::filter(dataset == "cf.lm") %>%
+#   dplyr::select(diff_GMV) 
+# 
+# t.test(cf.gmvdiff, cf.lm.gmvdiff, paired=TRUE)
+```
+
+NOTE: "missing" paired comparisons between cf and cf.lm diffs are because diff values are identical between the two methods (i.e.
+there is literally no difference in accuracy of centile scores between the two)
+
+```{r, warning=FALSE, message=FALSE}
+#within each phenotype, iterate paired t test between each combo of combat dfs
+
+#initalize empty dfs to store outputs
+sum.df <- data.frame(pheno = character(),
+                     term = character(),
+                     df = double(),
+                     sumsq = double(),
+                     meansq = double(),
+                     statistic = double(),
+                     p.value = double())
+t.df <- data.frame("pheno" = character(),
+                   "group1" = character(),
+                   "group2" = character(),
+                   "p.value" = double())
+pheno_diff.list <- paste0("diff_", pheno_list)
+
+#run tests
+attach(cent.diffs2)
+for (pheno in pheno_diff.list) {
+  df.aov <- tidy(aov(cent.diffs2[[pheno]] ~ cent.diffs2[["dataset"]]))
+  df.aov$pheno <- pheno
+  sum.df <- rbind(sum.df, df.aov)
+
+  df.pairwise.t <- tidy(pairwise.t.test(x=cent.diffs2[[pheno]], g=cent.diffs2[["dataset"]], paired = TRUE, p.adj = "none"))
+  df.pairwise.t$pheno <- pheno
+  t.df <- rbind(t.df, df.pairwise.t)
+}
 detach()
@@
 
-Seems like gamlss methods are significantly better than the other combat methods and not significantly different from raw centile scores (BH corrected)
+```{r}
+#apply fdr correction
+sum.df <- sum.df %>%
+  dplyr::filter(term != "Residuals") %>%
+  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
+
+summary(sum.df$p.val_fdr)
+
+t.df <- t.df %>%
+  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
+         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                             p.val_fdr >= 0.05 ~ FALSE),
+         pheno = sub("diff_", "", pheno)) %>%
+  unite(comp, c("group1", "group2")) %>%
+  mutate(comp = as.factor(comp))
+table(t.df$comp)
+
+t.df %>%
+  group_by(comp) %>%
+   summarize(n_sig_regions = sum(sig_fdr),
+            n_compare = n())
+```
 
@@
 No shift in abs. differences, looking at overall diff.
-```{r, fig.height=7, fig.width=8}
-cent.diffs2 <- combined_preds %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
-  ungroup()
-```
 
@@
 summary(aov(z_abs.diff ~ dataset, data=mean.z.long))
+tidy(aov(z_abs.diff ~ dataset, data=mean.z.long))
 
 attach(mean.z.long)
-pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr")
+t.obj <- pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr", paired = TRUE)
 detach()
+names(tidy(t.obj))
+
+test <- tidy(t.obj) %>%
+  mutate(comp = paste(group1, group2))
+table(test$comp)
 ```
@@
 
+Test
+
 Test for differences in centile offset distributions
@@
 mean.diffs.long_filt <- mean.diffs.long %>%
-  filter(!str_detect(dataset, "_refA"))
+  filter(!str_detect(dataset, "_refA")) %>%
+  filter(dataset != "ukb_CN")
 
@@
 attach(mean.diffs.long_filt)
-pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "fdr")
+pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "bonferroni")
 detach()
@@
 beta.plot2
+
+ggsave("male_variance_beta2.jpeg", beta.plot2, width=10, height=10)
 ```
@@
 proportion_beta_plot2
+ggsave("male_variance_prop.beta2.jpeg", proportion_beta_plot2, width=10, height=10)
 ```
@@
 ```{r}
+#Define scales
+scales_y <- list(
+  `Global Volume` = scale_y_continuous(limits = c(0, 4)),
+  `Regional CT` = scale_y_continuous(limits = c(0, 70)),
+  `Regional SA` = scale_y_continuous(limits = c(0, 70)),
+  `Regional Volume` = scale_y_continuous(limits = c(0, 70))
+)
+
 cent.diffs.plot2 <- mean.diffs.long %>%
@@
   theme_linedraw(base_size = 12) +
-  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
+  scale_x_continuous(labels = label_comma()) + 
+  facet_wrap(~pheno_cat) +
+  #facetscales::facet_grid_sc(pheno_cat~.,
+                #scales = list(y = scales_y)) +
   #paletteer::scale_fill_paletteer_d("Polychrome::light") +
@@
 cent.diffs.plot2
-```
+#ggsave("mean_centile_diffs2.jpeg", cent.diffs.plot2, width=10, height=10)
+```

### 2023-11-07 proof-of-concept_mods.Rmd
centile difference sig tests

@@
 - z scores instead of centiles!!!!
+- Fix age -> diff in male and female variance/intercept
 
@@
 table(sigma.df.glob$dataset, sigma.df.glob$sig.sum_bf.corr)
+```
+
+fix age, get diff between male and female pred. variance
+```{r}
+mean(ukb_df_age.filt$age_days)
 ```
@@
    summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n())
+            n_compare = n(),
+            min_pval = min(p.val_fdr),
+            max_pval = max(p.val_fdr))
 ```
@@
 No shift in abs. differences, looking at overall diff.
-
-```{r}
-max.cent.diffs.by.sex2 <- cent.diffs2 %>%
+```{r}
+#get centile errors for each subject - not abs val
+cent.diffs3 <- combined_preds %>%
+  dplyr::filter(!str_detect(dataset, "_refA")) %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  arrange(dataset) %>%
+  group_by(participant) %>%
+  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
+  ungroup() %>%
+  dplyr::filter(dataset != "ukb_CN") #drop ref
+```
+
+```{r, eval=FALSE}
+cent.diffs3 %>%
+  pivot_longer(starts_with("diff_"), names_to="pheno", values_to="diff", names_prefix = "diff_") %>%
+  ggplot() +
+  geom_density(aes(x=diff, fill=pheno)) +
+  facet_wrap(~pheno_class) +
+  theme(legend.position="none")
+```
+
+```{r}
+max.cent.diffs.by.sex2 <- cent.diffs3 %>%
   group_by(dataset, sex) %>%
@@
 ```{r}
-mean.cent.diffs.by.sex2 <- cent.diffs2 %>%
+mean.cent.diffs.by.sex2 <- cent.diffs3 %>%
   group_by(dataset, sex) %>%
@@
 Both gam methods and lm combat with ref both end up inducing a bias in how males and females centiles are shifted.
+
+```{r, warning=FALSE, message=FALSE}
+# within each df, within each phenotype, t-test between male and female diffs
+
+#initalize empty dfs to store outputs
+t.df.sex <- data.frame("pheno" = character(),
+                   "dataset" = character(),
+                   "p_value" = double()
+                   )
+
+for (pheno in pheno_diff.list){
+    df.sex.test <- cent.diffs3 %>%
+        group_by(dataset) %>%
+        summarize(p_value = t.test(.data[[pheno]][sex == "Male"], .data[[pheno]][sex == "Female"])$p.value) %>%
+        ungroup()
+    df.sex.test$pheno <- pheno
+    t.df.sex <- rbind(t.df.sex, df.sex.test)
+}
+```
+
+```{r}
+#apply fdr correction
+t.df.sex <- t.df.sex %>%
+  mutate(p.val_fdr = p.adjust(p_value, method="fdr", n = length(p_value)),
+         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                             p.val_fdr >= 0.05 ~ FALSE),
+         pheno = sub("diff_", "", pheno))
+
+t.df.sex %>%
+  group_by(dataset) %>%
+   summarize(n_sig_regions = sum(sig_fdr),
+            n_compare = n())
+```
 
@@
 Test for differences in distribution
-```{r}
-summary(aov(z_abs.diff ~ dataset, data=mean.z.long))
-tidy(aov(z_abs.diff ~ dataset, data=mean.z.long))
-
-attach(mean.z.long)
-t.obj <- pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr", paired = TRUE)
+
+```{r, warning=FALSE, message=FALSE}
+#within each phenotype, iterate paired t test between each combo of combat dfs
+
+z_diffs.filt <- z_diffs %>%
+  dplyr::filter(!str_detect(dataset, "_refA")) %>%
+  dplyr::filter(dataset != "ukb_CN")
+
+#initalize empty dfs to store outputs
+z.sum.df <- data.frame(pheno = character(),
+                     term = character(),
+                     df = double(),
+                     sumsq = double(),
+                     meansq = double(),
+                     statistic = double(),
+                     p.value = double())
+z.t.df <- data.frame("pheno" = character(),
+                   "group1" = character(),
+                   "group2" = character(),
+                   "p.value" = double())
+
+z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")
+
+#run tests
+attach(z_diffs.filt)
+for (pheno in z.pheno_abs.diff.list) {
+  df.aov <- tidy(aov(z_diffs.filt[[pheno]] ~ z_diffs.filt[["dataset"]]))
+  df.aov$pheno <- pheno
+  z.sum.df <- rbind(z.sum.df, df.aov)
+
+  df.pairwise.t <- tidy(pairwise.t.test(x=z_diffs.filt[[pheno]], g=z_diffs.filt[["dataset"]], paired = TRUE, p.adj = "none"))
+  df.pairwise.t$pheno <- pheno
+  z.t.df <- rbind(z.t.df, df.pairwise.t)
+}
 detach()
-names(tidy(t.obj))
-
-test <- tidy(t.obj) %>%
-  mutate(comp = paste(group1, group2))
-table(test$comp)
+```
+
+```{r}
+#apply fdr correction
+z.sum.df <- z.sum.df %>%
+  dplyr::filter(term != "Residuals") %>%
+  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
+
+summary(z.sum.df$p.val_fdr)
+
+z.t.df <- z.t.df %>%
+  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
+         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                             p.val_fdr >= 0.05 ~ FALSE),
+         pheno = sub("abs.diff_", "", pheno),
+         pheno = sub(".z", "", pheno)) %>%
+  unite(comp, c("group1", "group2")) %>%
+  mutate(comp = as.factor(comp))
+
+z.t.df %>%
+  group_by(comp) %>%
+   summarize(n_sig_regions = sum(sig_fdr),
+            n_compare = n(),
+            min.pval = min(p.val_fdr),
+            max.pval = max(p.val_fdr))
 ```

### 2023-11-09 proof-of-concept_mods.Rmd
keep data in permuted csvs

@@
 
+##### Variance
 Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex).
@@
 ```{r}
-mean(ukb_df_age.filt$age_days)
+pred.csvs <- list.files(path = "ukb_basic", pattern = "_variance.csv", full.names = TRUE)
+combined_var <- data.frame()
+
+for (file in pred.csvs) {
+  # Read each CSV file
+  data <- fread(file)
+
+  # Add a "Source_File" column with the file name
+  data <- data %>%
+    mutate(Source_File = as.factor(basename(file))) %>%
+    mutate(dataset = sub("_data_variance.csv", "", Source_File))
+
+  # Bind the data to the combined dataframe
+  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
+}
+
+table(combined_var$dataset)
+#skim(combined_preds)
+```
+
+```{r}
+combined_var <- combined_var %>%
+  mutate(sex_std = sex_effect/f.var)
+
+summary(sigma.df$var_est)
+summary(combined_var$sex_effect)
+
+summary(sigma.df$beta_ratio)
+summary(combined_var$sex_std)
+```
+
+```{r}
+plot.info <- sigma.df %>%
+  dplyr::filter(term == "sexMale") %>%
+  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
+  unique() %>%
+  mutate(dataset = gsub("-", "_", dataset),
+         dataset = gsub("_simsites", "", dataset))
+combined_var <- inner_join(combined_var, plot.info)
+table(combined_var$dataset)
+```
+
+plot
+
+```{r, fig.height=5, fig.widt=5}
+combined_var %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
+  group_by(dataset, pheno_cat) %>%
+  ggplot() +
+  geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = sex_std)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_bw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
+  ggtitle("Standardized effect of male sex on variance")
 ```

### 2023-11-09 R_scripts/permute_sites.R

@@ -27,10 +27,6 @@ n_male <- table(df$sex)["Male"]
 male_prob <- c(0.33, 0.0825, 0.5875)
 female_prob <- c(0.33, 0.5875, 0.0825)
 
-# Just get cols needed for site assignment
-df.filt <- df %>%
-  dplyr::select(participant, sex)
-
 #control # permutations to complete
 for(i in 1:n_permutations) {
   # Separate samples for males and females
@@ -38,12 +34,12 @@ for(i in 1:n_permutations) {
   sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)
   
   #assign to df
-  new_df <- df.filt %>%
+  new_df <- df %>%
     mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))
   
   #save out csv w permutation number in filename
   n_perm <- str_pad(i, 3, pad = "0")
-  fname <- paste0(save_path, "/", csv_basename, "_perm-key_", n_perm, ".csv")
+  fname <- paste0(save_path, "/", csv_basename, "_perm-", n_perm, ".csv")
   fwrite(new_df, file=fname)
   
 }

### 2023-11-09 proof-of-concept_mods.Rmd
updates, script to run ohbm results

@@
 - effect size for poly(age_days, 3) in gam/gamlss
-- z scores instead of centiles!!!!
 - Fix age -> diff in male and female variance/intercept
@@
 cent.diffs3 %>%
+  #dplyr::filter(dataset == "cf.gamlss") %>%
   pivot_longer(starts_with("diff_"), names_to="pheno", values_to="diff", names_prefix = "diff_") %>%
-  ggplot() +
-  geom_density(aes(x=diff, fill=pheno)) +
-  facet_wrap(~pheno_class) +
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_))) %>%
+  ggplot() +
+  geom_density(aes(x=diff, fill=dataset), alpha=0.4) +
+  facet_wrap(~pheno_cat) +
   theme(legend.position="none")

### 2023-11-09 R_scripts/ohbm_results.R

@@ -1,0 +1,232 @@
+#single script to parse outputs & get results for OHBM 2024 abstract (condensed from proof-of-concept_mods.Rmd)
+
+### LOAD ###
+pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
+
+library(data.table) 
+library(readr)
+library(ggplot2) 
+library(tidyverse) 
+library(mgcv) 
+library(gamlss)
+library(ggseg)
+library(paletteer)
+
+### CENTILES ###
+#Do centile scores change based on combat version?
+
+pred.csvs <- list.files(path = "ukb_basic", pattern = "_predictions.csv", full.names = TRUE)
+combined_preds <- data.frame()
+
+for (file in pred.csvs) {
+  # Read each CSV file
+  data <- fread(file)
+  
+  # Add a "Source_File" column with the file name
+  data <- data %>%
+    mutate(Source_File = as.factor(basename(file))) %>%
+    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
+  
+  # Bind the data to the combined dataframe
+  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
+}
+
+#calculate abs error
+cent_abs.diffs <- combined_preds %>%
+  dplyr::filter(!str_detect(dataset, "_refA")) %>% #drop ref site results
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  arrange(dataset) %>%
+  group_by(participant) %>%
+  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
+  ungroup() %>%
+  dplyr::filter(dataset != "ukb_CN") #drop raw
+
+#within each phenotype, iterate paired t test between each combo of combat dfs 
+
+#initalize empty dfs to store outputs
+sum.df <- data.frame(pheno = character(),
+                     term = character(),
+                     df = double(),
+                     sumsq = double(),
+                     meansq = double(),
+                     statistic = double(),
+                     p.value = double())
+t.df <- data.frame("pheno" = character(),
[…]

### 2023-11-13 ohbm_results.Rmd
simplified version of proof-of-concept_mods.Rmd with results, visuals

---
title: "ohbm_results"
output: html_document
date: '2023-11-12'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon)

source("./R_scripts/gamlss_helper_funs.R")
```

single script to parse outputs & get results for OHBM 2024 abstract (condensed from proof-of-concept_mods.Rmd)

load info
```{r}
data_path <- "./ukb_basic"

### LOAD ###
pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
vol_list_global <- readRDS(file="R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="R_scripts/SA_list.rds")
ct_list <- readRDS(file="R_scripts/CT_list.rds")
```

### CENTILES ###
Do centile scores change based on combat version?

```{r}
pred.csvs <- list.files(path = data_path, pattern = "_predictions.csv", full.names = TRUE)
combined_preds <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_predictions.csv", "", Source_File))

  # Bind the data to the combined dataframe
  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
}

#calculate abs error
cent_abs.diffs <- combined_preds %>%
  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
  ungroup() %>%
  dplyr::filter(dataset != "ukb_CN") #drop raw

#within each phenotype, iterate paired t test between each combo of combat dfs

#initalize empty dfs to store outputs
sum.df <- data.frame(pheno = character(),
                     term = character(),
                     df = double(),
                     sumsq = double(),
                     meansq = double(),
                     statistic = double(),
                     p.value = double())
t.df <- data.frame("pheno" = character(),
                   "group1" = character(),
                   "group2" = character(),
                   "p.value" = double())
pheno_diff.list <- paste0("abs.diff_", pheno_list)

#run tests
attach(cent_abs.diffs)
for (pheno in pheno_diff.list) {
  df.aov <- tidy(aov(cent_abs.diffs[[pheno]] ~ cent_abs.diffs[["dataset"]]))
  df.aov$pheno <- pheno
  sum.df <- rbind(sum.df, df.aov)

  df.pairwise.t <- tidy(pairwise.t.test(x=cent_abs.diffs[[pheno]], g=cent_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
  df.pairwise.t$pheno <- pheno
  t.df <- rbind(t.df, df.pairwise.t)
}
detach()

#apply fdr correction
sum.df <- sum.df %>%
  dplyr::filter(term != "Residuals") %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))

t.df <- t.df %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                             p.val_fdr >= 0.05 ~ FALSE),
         pheno = sub("abs.diff_", "", pheno)) %>%
  unite(comp, c("group1", "group2")) %>%
  mutate(comp = as.factor(comp))

t.df.sum <- t.df %>%
  group_by(comp) %>%
  summarize(n_sig_regions = sum(sig_fdr),
            n_compare = n(),
            min_pval = min(p.val_fdr),
            max_pval = max(p.val_fdr))
```

Big plot - too big to run locally :(

```{r, eval=FALSE}
#test big plot
plot.cents <- cent_abs.diffs %>%
  #dplyr::filter(dataset == "cf.gamlss") %>%
  pivot_longer(starts_with("abs.diff_"), names_to="pheno", values_to="abs.diff", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  ggplot() +
  geom_density(aes(x=diff, fill=dataset), alpha=0.4) +
  facet_wrap(~pheno_cat) +
  theme(legend.position="none")
ggsave("figs/cent_plot_test.jpeg", var.plot, width=10, height=10)
```

### Z-SCORES ###

```{r}
z_abs.diffs <- combined_preds %>%
  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref = "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
  ungroup() %>%
  dplyr::filter(dataset != "ukb_CN") #drop raw

#initalize empty dfs to store outputs
z.sum.df <- data.frame(pheno = character(),
                       term = character(),
                       df = double(),
                       sumsq = double(),
                       meansq = double(),
                       statistic = double(),
                       p.value = double())
z.t.df <- data.frame("pheno" = character(),
                     "group1" = character(),
                     "group2" = character(),
                     "p.value" = double())

z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")

#run tests
attach(z_abs.diffs)
for (pheno in z.pheno_abs.diff.list) {
  df.aov <- tidy(aov(z_abs.diffs[[pheno]] ~ z_abs.diffs[["dataset"]]))
  df.aov$pheno <- pheno
  z.sum.df <- rbind(z.sum.df, df.aov)

  df.pairwise.t <- tidy(pairwise.t.test(x=z_abs.diffs[[pheno]], g=z_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
  df.pairwise.t$pheno <- pheno
  z.t.df <- rbind(z.t.df, df.pairwise.t)
}
detach()

#apply fdr correction
z.sum.df <- z.sum.df %>%
  dplyr::filter(term != "Residuals") %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))

z.t.df <- z.t.df %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                             p.val_fdr >= 0.05 ~ FALSE),
         pheno = sub("abs.diff_", "", pheno),
         pheno = sub(".z", "", pheno)) %>%
  unite(comp, c("group1", "group2")) %>%
  mutate(comp = as.factor(comp))

sum_z.t.df <- z.t.df %>%
  group_by(comp) %>%
  summarize(n_sig_regions = sum(sig_fdr),
            n_compare = n(),
            min.pval = min(p.val_fdr),
            max.pval = max(p.val_fdr))
```

### SCALE EFFECTS OF SEX ###
fix age, get diff between male and female pred. variance
```{r}
pred.csvs <- list.files(path = data_path, pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_variance.csv", "", Source_File))

  # Bind the data to the combined dataframe
  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
}

#get sex effect
combined_var <- combined_var %>%
  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
  mutate(sex_std = sex_effect/f.var)

#get summary df for p vals, plotting
sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(dataset = gsub("-data", "", dataset),
         dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)

#add more info from parcellations
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")

combined_var <- inner_join(combined_var, plot.info)
```

```{r}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat", cf.lm = "ComBat Linear", cf.gam ="ComBatGAM", cf.gamlss ="ComBatLS")

#var plot
var.plot.df <- combined_var %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset, pheno_cat)

tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_std <- 1

var.plot <- ggplot(var.plot.df) +
  geom_rect(data = tp, fill="lightgreen", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.2) +
  ggseg::geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_std)) +
  #scale_fill_gradient2() +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("viridis::plasma", name = "Sex Effect")
var.plot

#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
```

```{r}
#plot global
var.plot.global.df <- combined_var %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA)),
         sex_std = ifelse(sig.sum_bf.corr == TRUE, sex_std, NA)) %>% #drop NS values
  group_by(dataset)

tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_std <- 1

var.plot.global <- ggplot(var.plot.global.df) +
  geom_rect(data = tp, fill="lightgreen", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.2) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = sex_std),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("viridis::plasma", name = "Sex Effect")
```

Combine plots
```{r}
p <- ggarrange(
  var.plot.global, var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

p_annotate <- annotate_figure(p, top = text_grob("Figure 1. Standardized effect of male sex on feature variance", 
               color = "black", size = 14, hjust=.65))

ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
```

```{r}
gridExtra::grid.arrange(var.plot.global,var.plot, ncol=2, widths=c(1, 1.7))
```

```{r}
p2 <- grid_arrange_shared_legend(var.plot.global, var.plot, ncol=2, nrow=1, widths=c(1, 1.7))
annotate_figure(p2, top = text_grob("Standardized effect of male sex on feature variance", 
               color = "black", size = 16))
```

### 2023-11-13 ohbm_results.Rmd
removing failed combo plot attempts

@@
 ```
-
-```{r}
-gridExtra::grid.arrange(var.plot.global,var.plot, ncol=2, widths=c(1, 1.7))
-```
-
-```{r}
-p2 <- grid_arrange_shared_legend(var.plot.global, var.plot, ncol=2, nrow=1, widths=c(1, 1.7))
-annotate_figure(p2, top = text_grob("Standardized effect of male sex on feature variance", 
-               color = "black", size = 16))
-```

### 2023-11-13 ohbm_results.Rmd
prepping to plot on PMACS

@@
 
+saveRDS(cent_abs.diffs, "data/centile_abs_diffs.rds")
+```
+
+```{r}
 #within each phenotype, iterate paired t test between each combo of combat dfs
@@
             max_pval = max(p.val_fdr))
+t.df.sum
 ```
@@
             max.pval = max(p.val_fdr))
+
+sum_z.t.df
 ```
@@
 var.plot <- ggplot(var.plot.df) +
-  geom_rect(data = tp, fill="lightgreen", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.2) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
   ggseg::geom_brain(atlas = dk, 
@@
 var.plot.global <- ggplot(var.plot.global.df) +
-  geom_rect(data = tp, fill="lightgreen", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.2) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
   ggseg::geom_brain(atlas = aseg_cerebrum, 

### 2023-11-13 R_scripts/ohbm_results.R

@@ -1,7 +1,7 @@
 #plot centile differences
 
 #define path
-cent_abs.diffs <- fread("/cbica/home/gardnerm/combat-biovar/ukb_basic")
+cent_abs.diffs <- readRDS("data/centile_abs_diffs.rds")
 
 ### LOAD ###
 vol_list_global <- readRDS(file="R_scripts/vol_list_global.rds")
@@ -9,7 +9,7 @@ vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
 sa_list <- readRDS(file="R_scripts/SA_list.rds")
 ct_list <- readRDS(file="R_scripts/CT_list.rds")
 
-library(data.table) 
+# library(data.table) 
 library(ggplot2) 
 library(tidyverse) 
 # library(ggseg)

### 2023-11-13 ohbm_results.Rmd
fix color scale

@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("viridis::plasma", name = "Sex Effect")
+  paletteer::scale_fill_paletteer_c("viridis::plasma", name = "Sex Effect", limits = c(-0.15, 0.15), breaks = c(-0.1, 0,0.1))
 var.plot
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("viridis::plasma", name = "Sex Effect")
+  paletteer::scale_fill_paletteer_c("viridis::plasma", name = "Sex Effect", limits = c(-0.15, 0.15), breaks = c(-0.1, 0,0.1))
 ```

### 2023-11-15 ohbm_results.Rmd
more tinkering plots

--- 2023-11-13 ohbm_results.Rmd
@@
 
+Little plot
+```{r}
+mean.diffs.long <- cent_abs.diffs %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+```
+
+```{r}
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
+
+cent.diffs.plot <- mean.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
+  theme_linedraw(base_size = 11) +
+  scale_x_continuous(labels = label_comma()) + 
+  facet_wrap(~pheno_cat) +
+  scale_y_continuous(labels = percent ) + 
+  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
+  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
+cent.diffs.plot
+
+ggsave("figs/mean_centile_diffs.jpeg", cent.diffs.plot)
+```
+
 ### Z-SCORES ###
@@
 ```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat", cf.lm = "ComBat Linear", cf.gam ="ComBatGAM", cf.gamlss ="ComBatLS")
-
 #var plot
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("viridis::plasma", name = "Sex Effect")
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
 var.plot
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("viridis::plasma", name = "Sex Effect")
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
+var.plot.global
 ```

### 2023-11-15 proof-of-concept_mods.Rmd

@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr, paletteer)
+pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr, paletteer, ggsegExtra, ggseg3d)
 
 source("./R_scripts/gamlss_helper_funs.R")
+# 
+# # Enable this universe
+# options(repos = c(
+#   ggseg = 'https://ggseg.r-universe.dev',
+#   CRAN = 'https://cloud.r-project.org'))
+# 
+# # Install some packages
+# install.packages('ggsegExtra')
 ```
@@
   ggtitle("t-statistic for effect of male sex on sigma")
+```
+
+Simple bar plot of sites
+```{r}
+ggplot(ukb_df_age.filt) +
+  geom_bar(aes(x=sim.site, fill=sex), color="black", alpha=0.6) +
+  theme_linedraw(base_size=12)
 ```
@@
   ggtitle("Standardized effect of male sex on variance")
+```
+
+using aseg
+
+```{r}
+plot(aseg)
 ```
@@
 ```{r}
+sigma.aseg_all <- sigma.df %>%
+  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale") %>%
+  filter(!str_detect(dataset, "-refA")) %>%
+  dplyr::select(!label) %>%
+  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss")),
+         tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "White Matter",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA)),
+         var_est = case_when(sig.sum_bf.corr == TRUE ~ var_est,
+                             TRUE ~ NA)) #just keep significant regions
+
+sigma.aseg_all <- left_join(sigma.aseg_all, aseg_cerebrum$data, by = c("tissue_class"))
+
+sigma.aseg_all %>%
+  group_by(dataset) %>%
+  ggplot() +
+  geom_brain(atlas = aseg_cerebrum, 
+             #position = position_brain(hemi ~ side),
+             aes(fill = var_est),
+             alpha=0.8) +
+  #scale_fill_gradient2() +
+  facet_wrap(~dataset) + 
+  theme_bw(base_size=20) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
+  ggtitle("Beta weight for effect of male sex on variance")
+```
+
+```{r}
 proportion_beta_plot2 <- sigma.df %>%

### 2023-11-20 ohbm_results.Rmd
adding replications

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel)
 
@@
 ct_list <- readRDS(file="R_scripts/CT_list.rds")
+
+aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
 ```
 
-### CENTILES ###
+# Main Analyses
+
+## CENTILES
 Do centile scores change based on combat version?
@@
   theme_linedraw(base_size = 11) +
-  scale_x_continuous(labels = label_comma()) + 
   facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = percent ) + 
+  scale_y_continuous(labels = scales::percent) + 
   scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
+  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
   ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
@@
 
-ggsave("figs/mean_centile_diffs.jpeg", cent.diffs.plot)
+#ggsave("figs/mean_centile_diffs.jpeg", cent.diffs.plot)
 ```
 
-### Z-SCORES ###
+## Z-SCORES
 
@@
 
-### SCALE EFFECTS OF SEX ###
+## SCALE EFFECTS OF SEX
 fix age, get diff between male and female pred. variance
@@
     mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_variance.csv", "", Source_File))
+    mutate(dataset = sub("_data_variance.csv", "", Source_File),
+           sex_standard = sex_effect/f.var)
 
@@
 }
-
-#get sex effect
-combined_var <- combined_var %>%
-  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
-  mutate(sex_std = sex_effect/f.var)
 
@@
 
+plot.info <- sigma.sex.df %>%
+  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
+  unique() %>%
+  mutate(dataset = gsub("-", "_", dataset),
+         dataset = gsub("_simsites", "", dataset))
+
 combined_var <- inner_join(combined_var, plot.info)
@@
   dplyr::filter(dataset=="ukb_CN")
-tp$sex_std <- 1
+tp$sex_standard <- 1
 
@@
              position = position_brain(hemi ~ side),
-             aes(fill = sex_std)) +
+             aes(fill = sex_standard)) +
   #scale_fill_gradient2() +
@@
                                             TRUE ~ NA)),
-         sex_std = ifelse(sig.sum_bf.corr == TRUE, sex_std, NA)) %>% #drop NS values
+         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
   group_by(dataset)
@@
   ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = sex_std),
+             aes(fill = sex_standard),
              side = c("coronal")) +
@@
 ```
+# Replication
+
+```{r}
+permutation_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_perm_csvs"
+```
+
+## CENTILES
+Do centile scores change based on combat version?
+
+```{r}
+n_perm_list <- lapply(c(1:10), str_pad, 3, pad = "0")
+n_perm_list <- paste0("perm-", n_perm_list)
+pheno_list.z <- paste0(pheno_list, ".z")
+
+# combined_preds.p <- data.frame()
+get.predictions <- function(x){
+  df <- data.frame() #new empty dataframe
+  pred.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_predictions.csv"), full.names = TRUE)
+  for (file in pred.csvs.p) {
+    # Read each CSV file
+    data <- fread(file)
+
+    # Add a "Source_File" column with the file name
+    data <- data %>%
+      mutate(Source_File = as.factor(basename(file)),
+             perm = as.factor(x)) %>%
+      mutate(dataset = gsub("_data|_predictions.csv|perm-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df <- bind_rows(df, data, .id = "File_ID")
+  }
+  return(df)
+}
+
+perm_predictions <- mclapply(n_perm_list, get.predictions, mc.preschedule = FALSE)
+```
+
+```{r}
+get.abs.diffs <- function(x){
+  df <- x %>%
+    mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+    mutate(dataset = relevel(dataset, ref= "raw")) %>%
+    arrange(dataset) %>%
+    group_by(participant) %>%
+    dplyr::mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #centile diffs
+    dplyr::select(-any_of(pheno_list)) %>% #drop
+    dplyr::mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #z-score diffs
+    dplyr::select(-any_of(pheno_list.z)) %>% #drop
+    ungroup() %>%
+    dplyr::filter(dataset != "raw") #drop raw
+  return(df)
+}
+
+perm_diffs <- mclapply(perm_predictions, get.abs.diffs, mc.preschedule = FALSE)
+```
+
+```{r}
+#within each phenotype, iterate paired t test between each combo of combat dfs 
+pheno_diff.list <- paste0("abs.diff_", pheno_list)
+
+centile.t.tests <- function(df){
+  #initialize empty dfs to store outputs
+  t.df <- data.frame("pheno" = character(),
+                     "group1" = character(),
+                     "group2" = character(),
+                     "p.value" = double())
+  #run tests
+  attach(df)
+  for (pheno in pheno_diff.list) {
+    df.pairwise.t <- tidy(pairwise.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
+    df.pairwise.t$pheno <- pheno
+    t.df <- rbind(t.df, df.pairwise.t)
+  }
+  n_comp <- nrow(t.df)
+
+  #apply fdr correction
+  t.df <- t.df %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE),
+           pheno = sub("abs.diff_", "", pheno)) %>%
+    unite(comp, c("group1", "group2")) %>%
+   dplyr::mutate(comp = as.factor(comp))
+
+  t.df.sum <- t.df %>%
+    group_by(comp) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+  return(t.df.sum)
+  #detach()
+}
+perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
+names(perm_t.tests) <- n_perm_list
+cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
+fwrite(cent_t.tests.p, file="ukb_perm_csvs/centile_t.tests.csv")
+```
+
+```{r}
+cent_t.tests.p %>%
+  group_by(comp) %>%
+  dplyr::summarize(n_sig_min=min(n_sig_regions),
+            n_sig_max = max(n_sig_regions),
+            n_sig_mean = mean(n_sig_regions))
+```
+
+Little plots
+```{r}
+plot.cent.diffs <- function(df){
+mean.diffs.long <- df %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup() %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+cent.diffs.plot <- mean.diffs.long %>%
+  dplyr::filter(dataset != "raw") %>%
+  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
+  theme_linedraw(base_size = 11) +
+  facet_wrap(~pheno_cat) +
+  scale_y_continuous(labels = scales::percent) + 
+  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
+  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
+
+print(cent.diffs.plot)
+}
+lapply(perm_diffs, plot.cent.diffs)
+```
+
+W/in subj average plots
+```{r}
+mean.subj.centile.err <- function(df){
+mean.diffs.subj <- df %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+  rowwise() %>%
+  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))
+
+return(mean.diffs.subj)
+#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
+}
+perm_cent.err <- mclapply(perm_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
+names(perm_cent.err) <- n_perm_list
+cent_err.p <- bind_rows(perm_cent.err, .id = "permutation")
+```
+
+```{r}
+cent.diffs.v.plot <- cent_err.p %>%
+  mutate(mean = mean*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=dataset, y=mean, fill=permutation))
+ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
+```
+
+## Z-SCORES
+
+```{r}
+#within each phenotype, iterate paired t test between each combo of combat dfs 
+pheno_diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
+
+z.t.tests <- function(df){
+  #initialize empty dfs to store outputs
+  t.df <- data.frame("pheno" = character(),
+                     "group1" = character(),
+                     "group2" = character(),
+                     "p.value" = double())
+  #run tests
+  attach(df)
+  for (pheno in pheno_diff.list.z) {
+    tryCatch(df.pairwise.t <- tidy(pairwise.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
+    df.pairwise.t$pheno <- pheno
+    t.df <- rbind(t.df, df.pairwise.t)
+  }
+  n_comp <- nrow(t.df)
+
+  #apply fdr correction
+  t.df <- t.df %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE),
+           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
+    unite(comp, c("group1", "group2")) %>%
+   dplyr::mutate(comp = as.factor(comp))
+
+  t.df.sum <- t.df %>%
+    group_by(comp) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+  return(t.df.sum)
+  #detach()
+}
+
+perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
+names(perm_t.tests_z) <- n_perm_list
+z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
+```
+
+```{r}
+fwrite(z_t.tests.p, file="ukb_permute/z_t.tests.csv")
+```
+
+## SCALE EFFECTS OF SEX
+fix age, get diff between male and female pred. variance
+```{r}
+# combined_preds.p <- data.frame()
+get.var <- function(x){
+  df <- data.frame() #new empty dataframe
+  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
+  for (file in var.csvs.p) {
+    # Read each CSV file
+    data <- fread(file)
+
+    # Add a "Source_File" column with the file name
+    data <- data %>%
+      mutate(Source_File = as.factor(basename(file)),
+             perm = as.factor(x)) %>%
+      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
+             sex_standard = sex_effect/f.var)
+
+    # Bind the data to the combined dataframe
+    df <- bind_rows(df, data, .id = "File_ID")
+  }
+  return(df)
+}
+
+perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
+```
+
+maybe better to get sex_summary.csv - just need signifigance
+
+```{r}
+#get summary df for p vals, plotting
+sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
+  dplyr::select(-V1) %>%
+  mutate(dataset = gsub("-data", "", dataset),
+         dataset = gsub("-", "_", dataset),
+         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
+  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
+
+#add more info from parcellations
+dk.parc <- dk$data %>%
+  as.data.frame() %>%
+  na.omit() %>%
+  dplyr::select(c(hemi, region, label)) %>%
+  distinct()
+
+sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
+
+combined_var <- inner_join(combined_var, plot.info)
+```
+
+```{r}
+plot_male_var <- function(df) {
+permutation <- unique(df[["perm"]])
+
+sex_effect_plot <- df %>%
+  mutate(sex_standard = case_when(isTRUE(sig.drop_bf.corr)~ sex_standard,
+                                   isFALSE(sig.drop_bf.corr) ~ NA)) %>% #only significant effects
+  ggplot() +
+  geom_bar(aes(x=pheno, y=sex_standard, fill=dataset)) +
+  facet_wrap(~pheno_cat) +
+  ggtitle(paste("Standardized effect of male sex on variance:", permutation))
+
+  ggsave(paste0("figs/male_variance_ohbm", permutation, ".jpeg", sex_effect_plot)
+  print(sex_effect_plot)
+}
+
+#mclapply
+
+```

### 2023-11-28 adni_eda.Rmd
finalize ohbm plots, adding EDA for next-steps

---
title: "ADNI EDA"
output: html_document
date: '2023-11-28'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)

# #pull plotting functions from github
# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
# source_url(plot.func.url) #sometimes doesn't work....?

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
```

Load data from Jakob
```{r}
adni.df <- readRDS("data/ADNIforMargaret.RDS")
#skim(adni.df)

#Lots of empty data, dropping
adni.df.filt <- adni.df %>%
  janitor::remove_empty(which = "cols")
skim(adni.df.filt)

#check for duplicate subjects
nrow(adni.df.filt) == n_unique(adni.df.filt$participant)
```

restrict to one scan per subject for visualizing demographics

```{r}
adni.df.no_dup <- adni.df.filt %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1)

nrow(adni.df.no_dup) == n_unique(adni.df.no_dup$participant)
```

## Site-wise EDA

look for variance-effecting features that are confounded by site

### Age

```{r}
ggplot(adni.df.no_dup) +
  geom_histogram(aes(x=age, fill=site), position="identity", alpha=0.5)
```

```{r}
ggplot(adni.df.no_dup) +
  geom_violin(aes(x=site, y=age, fill=sex, color=sex), scale="width", trim = FALSE, adjust=0.75)
```

```{r}
ggplot(adni.df.no_dup, aes(x=site, y=age, color=sex, fill=sex)) + 
  geom_boxplot()
```

```{r}
summary(aov(age_days ~ site, data = adni.df.filt))
```

### Sex

```{r}
ggplot(adni.df.no_dup) +
  geom_bar(aes(x=site, fill=sex))
```

```{r}
ggplot(adni.df.no_dup) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent)
```

### Diagnosis

```{r}
ggplot(adni.df.no_dup) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~dx)
```

```{r}
ggplot(adni.df.no_dup, aes(x=site, y=age, fill=dx, color=dx)) + 
  geom_boxplot() 
```

```{r}
ggplot(adni.df.no_dup, aes(x=site, y=age, fill=dx, color=dx)) + 
  geom_boxplot() +
  facet_wrap(~sex, nrow=2)
```

### 2023-11-28 ohbm_results.Rmd
finalize ohbm plots, adding EDA for next-steps

@@
   ggplot() +
-  geom_violin(aes(x=dataset, y=mean, fill=permutation))
+  geom_violin(aes(x=dataset, y=mean, fill=permutation)) +
+  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
 ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
@@
 ```{r}
-fwrite(z_t.tests.p, file="ukb_permute/z_t.tests.csv")
+fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
+```
+
+```{r}
+z_t.tests.p %>%
+  group_by(comp) %>%
+  dplyr::summarize(n_sig_min=min(n_sig_regions),
+            n_sig_max = max(n_sig_regions),
+            n_sig_mean = mean(n_sig_regions))
 ```
@@
       mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var)
+             sex_standard = sex_effect/f.var,
+             pheno = as.character(pheno))
 
@@
 perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
+perm_var.all <- bind_rows(perm_var, .id = "column_label")
 ```
@@
 #get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
+sigma.sex.df <- fread("ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
   dplyr::select(-V1) %>%
-  mutate(dataset = gsub("-data", "", dataset),
-         dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
+         dataset = gsub("-data", "", dataset),
+         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
+         dataset = gsub("-", "", dataset),
+         #dataset = as.character(dataset),
+         pheno = as.character(pheno)) %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
+         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
+  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
   dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
+  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
 
-#add more info from parcellations
-dk.parc <- dk$data %>%
-  as.data.frame() %>%
-  na.omit() %>%
-  dplyr::select(c(hemi, region, label)) %>%
-  distinct()
-
-sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
-
-combined_var <- inner_join(combined_var, plot.info)
+perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
 ```
@@
 ```{r}
-plot_male_var <- function(df) {
-permutation <- unique(df[["perm"]])
+  sex_effect_plot <- perm_var_info %>%
+    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
+                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
+    ggplot() +
+    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
+    facet_wrap(~perm, ncol=5) +
+    ggtitle(paste("Standardized effect of male sex on variance:", permutation)) +
+    theme(axis.text.x = element_blank())
 
-sex_effect_plot <- df %>%
-  mutate(sex_standard = case_when(isTRUE(sig.drop_bf.corr)~ sex_standard,
-                                   isFALSE(sig.drop_bf.corr) ~ NA)) %>% #only significant effects
-  ggplot() +
-  geom_bar(aes(x=pheno, y=sex_standard, fill=dataset)) +
-  facet_wrap(~pheno_cat) +
-  ggtitle(paste("Standardized effect of male sex on variance:", permutation))
-
-  ggsave(paste0("figs/male_variance_ohbm", permutation, ".jpeg", sex_effect_plot)
+    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
     print(sex_effect_plot)
-}
-
-#mclapply
-
 ```

### 2023-12-03 lifespan_eda.Rmd
eda of multisite data

---
title: "lifespan_eda"
output: html_document
date: '2023-11-30'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see)

# #pull plotting functions from github
# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
# source_url(plot.func.url) #sometimes doesn't work....?

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
```

Load data from Jakob
```{r}
cn.df <- fread("data/lifespan_dataset_dkatlas_CNonly.csv", stringsAsFactors = TRUE)

#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup()

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
```

## Site-Level Data
look just at data w site info

```{r}
cn.df.site <- cn.df.unique %>% 
  dplyr::filter(site != "NA") %>%
  mutate(study = droplevels(study),
         site = gsub(" ", "_", trimws(site)))
n_unique(cn.df.site$participant)
n_unique(cn.df.site$participant) == nrow(cn.df.site)

max(cn.df.site$age_days)/365.25

grep("scan", colnames(cn.df.site), value = TRUE) #no scanner info
```

```{r}
# number studies
length(levels(cn.df.site$study))

#number sites per study
cn.df.site %>%
  group_by(study) %>% 
  summarise(Result = length(unique(site)))
```

```{r}
ggplot(cn.df.site) +
  geom_violin(aes(x=study, y=age_days), scale="width")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.site) +
  geom_violin(aes(x=site, y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

check for sex differences in sites

```{r}
ggplot(cn.df.site) +
  geom_bar(aes(x=site, fill=sex))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
ggplot(cn.df.site) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Write out

```{r}
fwrite(cn.df.site, file="data/lifespan_CN_site-level.csv")
```

## Keep Study Level
Keep studies w missing site info and just treat as single-site

```{r}
#clean up site NAs
cn.df.imp <- cn.df.unique %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, site)))

unique(cn.df.imp$site)
length(is.na(cn.df.imp$site))
max(cn.df.imp$age_days)/365.25
mean(cn.df.imp$age_days)/365.25
```

```{r}
ggplot(cn.df.imp) +
  geom_violin(aes(x=study, y=age_days), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.imp) +
  geom_violin(aes(x=site, y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

check for sex differences in sites

```{r}
ggplot(cn.df.imp) +
  geom_bar(aes(x=site, fill=sex))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
ggplot(cn.df.imp) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Write out

```{r}
fwrite(cn.df.imp, file="data/lifespan_CN_imp-sites.csv")
```

### 2023-12-03 lifespan_eda.Rmd
update combat gam and combatLS formulas for s(age)*sex interactions, add sexMale terms

@@
   slice(1) %>%
-  ungroup()
+  ungroup() %>%
+  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
+                                        sex == "Female" ~ 0,
+                                        TRUE ~ NA)))
 

### 2023-12-03 qsub_combat.sh

@@ -38,7 +38,7 @@ cd $base #to source functions correctly
 #######################################################################################
 # MAKE DIRECTORIES 
 #study dir
-study_dir=$base/ukb_permute
+study_dir=$base/lifespan
 if ! [ -d $study_dir ]
 	then
 	mkdir $study_dir
@@ -94,22 +94,22 @@ do
 	#COMBAT GAM
 	elif [ $config = "cf.gam" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gam, formula = y ~ s(age_days) + sexMale'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gam, formula = y ~ s(age_days) * sexMale'" > $bash_script
 
 	#COMBAT GAM W/ REF SITE
 	elif [ $config = "cf.gam_refA" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gam, formula = y ~ s(age_days) + sexMale, ref.batch = \"Site_A\"'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gam, formula = y ~ s(age_days) * sexMale, ref.batch = \"Site_A\"'" > $bash_script
 
 	#COMBAT GAMLSS
 	elif [ $config = "cf.gamlss" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gamlss, formula = y ~ pb(age_days) + sexMale, sigma.formula = ~ age_days + sexMale'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gamlss, formula = y ~ pb(age_days) * sexMale, sigma.formula = ~ pb(age_days) * sexMale'" > $bash_script
 
 	#COMBAT GAMLSS W/ REF SITE
 	elif [ $config = "cf.gamlss_refA" ]
 	then
[…]

### 2023-12-03 lifespan_eda.Rmd
adding sex.age hard-coded int term

--- 2023-12-03 lifespan_eda.Rmd
@@
   slice(1) %>%
-  ungroup()
+  ungroup() %>%
+  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
+                                        sex == "Female" ~ 0,
+                                        TRUE ~ NA)),
+         sex.age = sexMale*age_days)
 

### 2023-12-03 qsub_combat.sh

@@ -62,7 +62,7 @@ csv_fname=$(basename $csv .csv)
 
 #######################################################################################
 #SET COVAR COLS
-covar_list="age_days,sexMale"
+covar_list="age_days,sexMale,sex.age"
 #SET BATCH COL
 batch="sim.site"
 
@@ -94,22 +94,22 @@ do
 	#COMBAT GAM
 	elif [ $config = "cf.gam" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gam, formula = y ~ s(age_days) * sexMale'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gam, formula = y ~ s(age_days) + sexMale + s(sex.age)'" > $bash_script
 
 	#COMBAT GAM W/ REF SITE
 	elif [ $config = "cf.gam_refA" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gam, formula = y ~ s(age_days) * sexMale, ref.batch = \"Site_A\"'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gam, formula = y ~ s(age_days) + sexMale + s(sex.age), ref.batch = \"Site_A\"'" > $bash_script
 
 	#COMBAT GAMLSS
 	elif [ $config = "cf.gamlss" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gamlss, formula = y ~ pb(age_days) * sexMale, sigma.formula = ~ pb(age_days) * sexMale'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list $pass 'gamlss, formula = y ~ pb(age_days) + sexMale + pb(sex.age), sigma.formula = ~ pb(age_days) + sexMale + pb(sex.age)'" > $bash_script
 
 	#COMBAT GAMLSS W/ REF SITE
 	elif [ $config = "cf.gamlss_refA" ]
 	then
[…]

### 2023-12-03 lifespan_eda.Rmd
clean up nas

--- 2023-12-03 lifespan_eda.Rmd
@@
   slice(1) %>%
-  ungroup()
+  ungroup() %>%
+  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
+                                        sex == "Female" ~ 0,
+                                        TRUE ~ NA)),
+         sex.age = sexMale*age_days,
+         log_age = log(age_days, base=10))
 
 nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
+```
+
+```{r}
+skim(cn.df.unique)
+levels(cn.df.unique$study)
+```
+
+## Visualize Features
+plot global brain features for basic qc - check again post-combat
+
+```{r}
+ggplot(cn.df.unique) +
+  geom_point(aes(x=age_days, y=GMV, color=sex))
+
+ggplot(cn.df.unique) +
+  geom_point(aes(x=age_days, y=sGMV, color=sex))
+
+ggplot(cn.df.unique) +
+  geom_point(aes(x=age_days, y=WMV, color=sex))
+
+ggplot(cn.df.unique) +
+  geom_point(aes(x=age_days, y=Ventricles, color=sex))
+```
+
+log-scaled
+```{r}
+ggplot(cn.df.unique) +
+  geom_point(aes(x=log_age, y=GMV, color=sex))
+
+ggplot(cn.df.unique) +
+  geom_point(aes(x=log_age, y=sGMV, color=sex))
+
+ggplot(cn.df.unique) +
+  geom_point(aes(x=log_age, y=WMV, color=sex))
+
+ggplot(cn.df.unique) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex))
 ```
@@
   mutate(study = droplevels(study),
-         site = gsub(" ", "_", trimws(site)))
+         site = gsub(" ", "_", trimws(site))) %>%
+  na.omit()
 n_unique(cn.df.site$participant)
@@
          study = trimws(study)) %>%
-  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, site)))
+  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, site))) %>%
+  na.omit()
 

### 2023-12-04 ohbm_results.Rmd
trying to add more memory

@@
 cent_err.p <- bind_rows(perm_cent.err, .id = "permutation")
+
+#write out for plotting later
+saveRDS(cent_err.p, "ukb_perm_csvs/centile_errors.RDS")
 ```
@@
     facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance:", permutation)) +
+    ggtitle(paste("Standardized effect of male sex on variance")) +
     theme(axis.text.x = element_blank())

### 2023-12-04 qsub_combat.sh

@@ -111,5 +111,5 @@ do
 	fi
 
 #qsub bash script
-	qsub -N ${config}.${csv_fname} -o $bash_dir/${config}_out.txt -e $bash_dir/${config}_err.txt $bash_script
+	qsub -l h_vmem=64G,s_vmem=64G -N ${config}.${csv_fname} -o $bash_dir/${config}.${csv_fname}_out.txt -e $bash_dir/${config}.${csv_fname}_err.txt $bash_script
 done

### 2023-12-04 paper_figs.Rmd
starting figures for paper, playing around with sex-bias testing

--- 2023-11-20 ohbm_results.Rmd
@@
 
-single script to parse outputs & get results for OHBM 2024 abstract (condensed from proof-of-concept_mods.Rmd)
+Figures for paper
+
+# UKB Basic Models
 
@@
 aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
-```
 
-# Main Analyses
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
 
-## CENTILES
-Do centile scores change based on combat version?
-
-```{r}
-pred.csvs <- list.files(path = data_path, pattern = "_predictions.csv", full.names = TRUE)
-combined_preds <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
-
-  # Bind the data to the combined dataframe
-  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
-}
-
-#calculate abs error
-cent_abs.diffs <- combined_preds %>%
-  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
-  ungroup() %>%
-  dplyr::filter(dataset != "ukb_CN") #drop raw
-
-saveRDS(cent_abs.diffs, "data/centile_abs_diffs.rds")
-```
-
-```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs
-
-#initalize empty dfs to store outputs
-sum.df <- data.frame(pheno = character(),
-                     term = character(),
-                     df = double(),
-                     sumsq = double(),
-                     meansq = double(),
-                     statistic = double(),
-                     p.value = double())
-t.df <- data.frame("pheno" = character(),
-                   "group1" = character(),
-                   "group2" = character(),
-                   "p.value" = double())
-pheno_diff.list <- paste0("abs.diff_", pheno_list)
-
-#run tests
-attach(cent_abs.diffs)
-for (pheno in pheno_diff.list) {
-  df.aov <- tidy(aov(cent_abs.diffs[[pheno]] ~ cent_abs.diffs[["dataset"]]))
-  df.aov$pheno <- pheno
-  sum.df <- rbind(sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.t.test(x=cent_abs.diffs[[pheno]], g=cent_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  t.df <- rbind(t.df, df.pairwise.t)
-}
-detach()
-
-#apply fdr correction
-sum.df <- sum.df %>%
-  dplyr::filter(term != "Residuals") %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-t.df <- t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-t.df.sum <- t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min_pval = min(p.val_fdr),
-            max_pval = max(p.val_fdr))
-t.df.sum
+#ukb data from ohbm_results.Rmd
+cent_err.p <- readRDS("ukb_perm_csvs/centile_errors.RDS")
+z_diffs2 <- readRDS("ukb_basic/z_diffs.RDS")
+cent.diffs3 <- readRDS("ukb_basic/cent_diffs.RDS")
+ukb_data <- fread("data/ukb_to_model/ukb_CN_data_simsites.csv")
+#cent_abs.diffs <- readRDS("data/centile_abs_diffs.rds")
 ```
@@
 
-Little plot
-```{r}
-mean.diffs.long <- cent_abs.diffs %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-```
-
-```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
-cent.diffs.plot
-
-#ggsave("figs/mean_centile_diffs.jpeg", cent.diffs.plot)
-```
-
-## Z-SCORES
-
-```{r}
-z_abs.diffs <- combined_preds %>%
-  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref = "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
-  ungroup() %>%
-  dplyr::filter(dataset != "ukb_CN") #drop raw
-
-#initalize empty dfs to store outputs
-z.sum.df <- data.frame(pheno = character(),
-                       term = character(),
-                       df = double(),
-                       sumsq = double(),
-                       meansq = double(),
-                       statistic = double(),
-                       p.value = double())
-z.t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-
-z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")
-
-#run tests
-attach(z_abs.diffs)
-for (pheno in z.pheno_abs.diff.list) {
-  df.aov <- tidy(aov(z_abs.diffs[[pheno]] ~ z_abs.diffs[["dataset"]]))
-  df.aov$pheno <- pheno
-  z.sum.df <- rbind(z.sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.t.test(x=z_abs.diffs[[pheno]], g=z_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  z.t.df <- rbind(z.t.df, df.pairwise.t)
-}
-detach()
-
-#apply fdr correction
-z.sum.df <- z.sum.df %>%
-  dplyr::filter(term != "Residuals") %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-z.t.df <- z.t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno),
-         pheno = sub(".z", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-sum_z.t.df <- z.t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min.pval = min(p.val_fdr),
-            max.pval = max(p.val_fdr))
-
-sum_z.t.df
-```
-
-## SCALE EFFECTS OF SEX
+## Sex Effects on Scale
 fix age, get diff between male and female pred. variance
@@
   paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-var.plot
 
@@
   paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-var.plot.global
 ```
@@
                color = "black", size = 14, hjust=.65))
-
+p_annotate
 ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
 ```
-# Replication
+
+## Z Errors
 
 ```{r}
-permutation_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_perm_csvs"
+ukb_sites <- ukb_data %>%
+  dplyr::select(participant, sim.site)
+
+z_diffs <- left_join(z_diffs2, ukb_sites) %>%
+  dplyr::filter(dataset != "ukb_CN")
+
+ggplot(z_diffs, aes(x=GMV.z, y=diff_GMV.z, color=sex)) +
+  geom_point(alpha=0.1, aes(shape=sim.site)) +
+  geom_smooth() +
+  facet_wrap(~dataset)
+
+ggplot(z_diffs, aes(x=GMV.z, y=diff_GMV.z, color=sex)) +
+  geom_point(alpha=0.1, aes(shape=sim.site)) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset)
+
+ggplot(z_diffs, aes(x=WMV.z, y=diff_WMV.z, color=sex)) +
+  geom_point(alpha=0.1, aes(shape=sim.site)) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset)
+
+ggplot(z_diffs) +
+  geom_jitter(aes(x=sex, y=WMV.z, color=diff_WMV.z), width = 0.2, height = 0, alpha=0.2) +
+  #geom_smooth(method="lm") +
+  facet_grid(sim.site~dataset) +
+  paletteer::scale_color_paletteer_c("grDevices::Cork") +
+  theme_bw()
+
+ggplot(z_diffs) +
+  geom_jitter(aes(x=sex, y=diff_WMV.z, color=WMV.z), width = 0.2, height = 0, alpha=0.2) +
+  #geom_smooth(method="lm") +
+  facet_grid(sim.site~dataset) +
+  #paletteer::scale_color_paletteer_c("grDevices::Cork") +
+  theme_bw()
+
+ggplot(z_diffs, aes(x=dataset, y=diff_WMV.z, color=sex)) +
+  geom_boxplot() +
+  theme_bw() +
+  stat_compare_means(aes(group = sex))
 ```
 
-## CENTILES
-Do centile scores change based on combat version?
+## Centile Errors
 
 ```{r}
-n_perm_list <- lapply(c(1:10), str_pad, 3, pad = "0")
-n_perm_list <- paste0("perm-", n_perm_list)
-pheno_list.z <- paste0(pheno_list, ".z")
+cent_diffs <- left_join(cent.diffs3, ukb_sites) %>%
+  dplyr::filter(dataset != "ukb_CN")
 
-# combined_preds.p <- data.frame()
-get.predictions <- function(x){
-  df <- data.frame() #new empty dataframe
-  pred.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_predictions.csv"), full.names = TRUE)
-  for (file in pred.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
+ggplot(cent_diffs, aes(x=GMV, y=diff_GMV, color=sex)) +
+  geom_point(alpha=0.1, aes(shape=sim.site)) +
+  #geom_smooth(method="lm") +
+  facet_wrap(~dataset)
 
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_predictions.csv|perm-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_predictions <- mclapply(n_perm_list, get.predictions, mc.preschedule = FALSE)
+ggplot(cent_diffs, aes(x=WMV, y=diff_WMV, color=sex)) +
+  geom_point(alpha=0.1, aes(shape=sim.site)) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset)
+```
+```{r}
+ggplot(z_diffs, aes(x=dataset, y=diff_WMV.z, color=sex)) +
+  geom_boxplot() +
+  theme_bw() +
+  stat_compare_means(aes(group = sex))
 ```
 
-```{r}
-get.abs.diffs <- function(x){
-  df <- x %>%
-    mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-    mutate(dataset = relevel(dataset, ref= "raw")) %>%
-    arrange(dataset) %>%
-    group_by(participant) %>%
-    dplyr::mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #centile diffs
-    dplyr::select(-any_of(pheno_list)) %>% #drop
-    dplyr::mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #z-score diffs
-    dplyr::select(-any_of(pheno_list.z)) %>% #drop
-    ungroup() %>%
-    dplyr::filter(dataset != "raw") #drop raw
-  return(df)
-}
-
-perm_diffs <- mclapply(perm_predictions, get.abs.diffs, mc.preschedule = FALSE)
-```
-
-```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs 
-pheno_diff.list <- paste0("abs.diff_", pheno_list)
-
-centile.t.tests <- function(df){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in pheno_diff.list) {
-    df.pairwise.t <- tidy(pairwise.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = sub("abs.diff_", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
-names(perm_t.tests) <- n_perm_list
-cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-fwrite(cent_t.tests.p, file="ukb_perm_csvs/centile_t.tests.csv")
-```
-
-```{r}
-cent_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
-
-Little plots
-```{r}
-plot.cent.diffs <- function(df){
-mean.diffs.long <- df %>%
-  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "raw") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
-
-print(cent.diffs.plot)
-}
-lapply(perm_diffs, plot.cent.diffs)
-```
-
+## Replication Centiles
 W/in subj average plots
-```{r}
-mean.subj.centile.err <- function(df){
-mean.diffs.subj <- df %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  rowwise() %>%
-  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))
-
-return(mean.diffs.subj)
-#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
-}
-perm_cent.err <- mclapply(perm_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
-names(perm_cent.err) <- n_perm_list
-cent_err.p <- bind_rows(perm_cent.err, .id = "permutation")
-```
 
@@
   ggplot() +
-  geom_violin(aes(x=dataset, y=mean, fill=permutation))
+  geom_violin(aes(x=dataset, y=mean, fill=permutation)) +
+  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
+
 ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
 ```
-
-## Z-SCORES
-
-```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs 
-pheno_diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
-
-z.t.tests <- function(df){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in pheno_diff.list.z) {
-    tryCatch(df.pairwise.t <- tidy(pairwise.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
-names(perm_t.tests_z) <- n_perm_list
-z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
-```
-
-```{r}
-fwrite(z_t.tests.p, file="ukb_permute/z_t.tests.csv")
-```
-
-## SCALE EFFECTS OF SEX
-fix age, get diff between male and female pred. variance
-```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var)
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-```
-
-maybe better to get sex_summary.csv - just need signifigance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(dataset = gsub("-data", "", dataset),
-         dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
-
-#add more info from parcellations
-dk.parc <- dk$data %>%
-  as.data.frame() %>%
-  na.omit() %>%
-  dplyr::select(c(hemi, region, label)) %>%
-  distinct()
-
-sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
-
-combined_var <- inner_join(combined_var, plot.info)
-```
-
-```{r}
-plot_male_var <- function(df) {
-permutation <- unique(df[["perm"]])
-
-sex_effect_plot <- df %>%
-  mutate(sex_standard = case_when(isTRUE(sig.drop_bf.corr)~ sex_standard,
-                                   isFALSE(sig.drop_bf.corr) ~ NA)) %>% #only significant effects
-  ggplot() +
-  geom_bar(aes(x=pheno, y=sex_standard, fill=dataset)) +
-  facet_wrap(~pheno_cat) +
-  ggtitle(paste("Standardized effect of male sex on variance:", permutation))
-
-  ggsave(paste0("figs/male_variance_ohbm", permutation, ".jpeg", sex_effect_plot)
-  print(sex_effect_plot)
-}
-
-#mclapply
-
-```

### 2023-12-04 proof-of-concept_mods.Rmd
starting figures for paper, playing around with sex-bias testing

@@
   dplyr::filter(dataset != "ukb_CN") #drop ref
+saveRDS(cent.diffs3, file="ukb_basic/cent_diffs.RDS")
 ```
@@
   group_by(participant) %>%
-  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
+  mutate(across(ends_with(".z"), ~ . - first(.), .names = "diff_{.col}")) %>%
   ungroup()
+#write out
+saveRDS(z_diffs2, file = "ukb_basic/z_diffs.RDS")
 

### 2023-12-05 paper_figs.Rmd
sex bias analyses

@@
   facet_grid(sim.site~dataset) +
-  #paletteer::scale_color_paletteer_c("grDevices::Cork") +
+  paletteer::scale_color_paletteer_c("grDevices::Cork") +
   theme_bw()
@@
   stat_compare_means(aes(group = sex))
+
+ggplot(z_diffs, aes(x=dataset, y=diff_WMV.z, color=sim.site)) +
+  geom_boxplot() +
+  theme_bw() +
+  stat_compare_means(aes(group = sim.site, label = paste0("p = ", after_stat(p.format))))
+```
+
+```{r}
+ggplot(z_diffs) +
+  geom_boxplot(aes(y=diff_WMV.z, x=sim.site, color=sex)) +
+  #geom_jitter(aes(y=diff_WMV.z, x=sim.site, color=sex), alpha=0.3) +
+  #geom_smooth(method="lm", aes(linetype=sim.site)) +
+  facet_wrap(~dataset)
 ```
@@
 ```{r}
-ggplot(z_diffs, aes(x=dataset, y=diff_WMV.z, color=sex)) +
+ggplot(cent_diffs, aes(x=dataset, y=diff_WMV, color=sex)) +
   geom_boxplot() +

### 2023-12-05 eda/abide_eda.Rmd
reorg

---
title: "ABIDE EDA"
output: html_document
date: '2023-11-30'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see)

# #pull plotting functions from github
# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
# source_url(plot.func.url) #sometimes doesn't work....?

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
```

Load data from Lena
```{r}
abide.df <- fread("data/ABIDE.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-X) %>%
  janitor::remove_empty(which = "cols") #drop totally empty cols

skim(abide.df)

#check for duplicate subjects
nrow(abide.df) == n_unique(abide.df$participant)
```

restrict to one scan per subject for visualizing demographics

```{r}
abide.df.no_dup <- abide.df %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1)

nrow(abide.df.no_dup) == n_unique(abide.df.no_dup$participant)

min(abide.df.no_dup$age_days)/365
max(abide.df.no_dup$age_days)/365
```

## Site-wise EDA

look for variance-effecting features that are confounded by site

### Age

```{r}
ggplot(abide.df.no_dup) +
  geom_violin(aes(x=site, y=age_days), scale="width", trim = FALSE, adjust=0.75) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
```{r}
ggplot(abide.df.no_dup) +
  geom_boxplot(aes(x=site, y=age_days)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
ggplot(abide.df.no_dup) +
  geom_violin(aes(x=site, y=age_days, fill=sex), scale="width", trim = FALSE, adjust=0.75) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Sex

```{r}
ggplot(abide.df.no_dup) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Diagnosis

```{r}
ggplot(abide.df.no_dup) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=dx)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
ggplot(abide.df.no_dup) + 
  geom_density(aes(x=lh.CT.frontalpole, fill=dx))
```

### 2023-12-05 eda/adni_eda.Rmd
reorg

--- 2023-11-28 adni_eda.Rmd
@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
+pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see)
 
@@
   group_by(participant) %>%
-  slice(1)
+  slice(1) %>%
+  ungroup() %>%
+  group_by(site) %>%
+  filter(n() >= 2) %>%
+  ungroup() #drop sites w <2 ppl
 
@@
 ggplot(adni.df.no_dup) +
-  geom_histogram(aes(x=age, fill=site), position="identity", alpha=0.5)
+  geom_violin(aes(x=site, y=age), scale="width", trim = FALSE, adjust=0.75)
 ```
@@
 ggplot(adni.df.no_dup, aes(x=site, y=age, fill=dx, color=dx)) +
-  geom_boxplot() 
+  geom_violin(scale="width", trim = FALSE, adjust=0.75)
 ```
@@
 ```
+
+Keep just CN, AD, and LMCI (same as longComBat paper)
+
+```{r}
+adni.df.no_dup %>%
+  dplyr::filter(dx == "CN" | dx == "AD" | dx == "LMCI") %>%
+ggplot() +
+  geom_bar(aes(y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=dx)) + 
+  scale_y_continuous(labels = scales::percent)
+
+adni.df.no_dup %>%
+  dplyr::filter(dx == "CN" | dx == "AD" | dx == "LMCI") %>%
+ggplot() +
+  geom_bar(aes(x=site, fill=dx))
+```

### 2023-12-05 ukb_sim.site_analyses.Rmd
cleaning up section headers

--- 2023-12-04 proof-of-concept_mods.Rmd
@@
 ---
-title: "proof-of-concept_mods"
+title: "UKB Simulation Analyses"
 output: html_document
@@
 ---
-
-## QUESTIONS
-- effect size for poly(age_days, 3) in gam/gamlss
-- Fix age -> diff in male and female variance/intercept
 
@@
 
-source("./R_scripts/gamlss_helper_funs.R")
+source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
 # 
@@
 
-# Analysis Outline
-## Load Data
+## Load
 
@@
 
-vol_list_global <- readRDS(file="R_scripts/vol_list_global.rds")
-vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
-sa_list <- readRDS(file="R_scripts/SA_list.rds")
-ct_list <- readRDS(file="R_scripts/CT_list.rds")
-pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
-```
-
-## Basic Models
-Designate *basic* model formula
-
-```{r}
-basic.m1 <- gamlss(formula = GMV ~ poly(age_days, 3) + sexMale,
-                       sigma.formula = ~ poly(age_days, 3) + sexMale,
-                       nu.formula = ~ 1,
-                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
-summary(basic.m1)
-```
-
-plot
-
-```{r}
-#plot_singlestudy_centiles(basic.m1, "GMV", ukb_df_age.filt, "sexMale")
-```
-
-# Review Results
+vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
+vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
+sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
+ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
+pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
+
+aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+```
 
@@
 
-### Raw Data Results
+Simple bar plot of sites
+```{r}
+ggplot(ukb_df_age.filt) +
+  geom_bar(aes(x=sim.site, fill=sex), color="black", alpha=0.6) +
+  theme_linedraw(base_size=12)
+```
+
+## Raw Data Results
+Review outcomes from raw (non-ComBat-ed) data
 
@@
 
-Simple bar plot of sites
-```{r}
-ggplot(ukb_df_age.filt) +
-  geom_bar(aes(x=sim.site, fill=sex), color="black", alpha=0.6) +
-  theme_linedraw(base_size=12)
-```
-
-### Combat Effects
-
-#### Site Effects
+## ComBat Effects by Moment
+
 Loading combatted data (scp'd from cubic: `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/data/ukb_to_model/*.csv' ./data/ukb_to_model`) to see how values change with different combat methods
+
 ```{r}
@@
 
-#### Centiles
+## ComBat Effects on Centiles
 
@@
 
-##### Sex-Biases
+#### Sex-Biases
 
@@
 
-#### Z-Score
+## ComBat Effects on Z-Score
 calc how much z-scores change

### 2023-12-05 ukb_sim.site_analyses.Rmd
removing GAM analyses

@@
 
-#### GAMs
-scp'd csvs from cubic (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/*.csv' ./ukb_basic`)
-
-```{r}
-anova.gam <- fread("ukb_basic/anova_gam_tests.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1)
-gam.sex.df <- fread("ukb_basic/sex_summary_gam.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1)
-gam.sum.df <- fread("ukb_basic/gam_summary.csv", stringsAsFactors = TRUE)%>%
-  dplyr::select(-V1)
-```
-
-Does the number of regions w significant age effects on mean change depending on combat version?
-
-```{r}
-anova.gam %>%
-  dplyr::filter(term=="poly(age_days, 3)") %>%
-  group_by(dataset) %>%
-  summarize(n_sig_regions = sum(sig.drop_bf.corr),
-            n_regions = n())
-```
-
-No.
-
-```{r, fig.height=10, fig.width=15}
-anova.gam %>%
-  dplyr::filter(term=="poly(age_days, 3)") %>%
-  rename(f_val = "F") %>%
-  ggplot() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  geom_col(aes(x=phenotype, y=f_val, fill=dataset), position = "dodge")
-```
-
-```{r}
-anova.gam_wide.f <- anova.gam %>%
-  dplyr::filter(term=="poly(age_days, 3)") %>%
-  rename(f_val = "F") %>%
-  pivot_wider(id_cols= phenotype, names_from=dataset, values_from=f_val)
-
-anova.gam.f_diff <- anova.gam_wide.f %>%
-  rename_all(~ gsub("-", "_", .)) %>%
-  mutate(across(starts_with("cf"), ~ . - ukb_CN_data_simsites)) %>%
-  dplyr::select(!ukb_CN_data_simsites)%>%
-  pivot_longer(!phenotype, names_to = "dataset", values_to = "f_diff") %>%
-  mutate(dataset = as.factor(dataset))
-
-summary(aov(f_diff ~ dataset, data=anova.gam.f_diff))
-
-attach(anova.gam.f_diff)
-pairwise.t.test(x=f_diff, g=dataset, p.adj = "bonf")
-detach()
-
-ggplot(anova.gam.f_diff) + 
-  geom_density(aes(x=f_diff, fill=dataset), alpha=0.5)
-
-anova.gam.f_diff %>%
-  dplyr::filter(dataset == "cf.gamlss_data" | dataset == "cf.gam_data" | dataset == "cf.gam_refA_data" | dataset == "cf.gamlss_refA_data") %>%
-ggplot() + 
-  geom_density(aes(x=f_diff, fill=dataset), alpha=0.5)
-```
-
-So the effect sizes (if this F is the correct F to use as effect size) are closer to the "ground truth" for gamlss than non-gamlss combat methods.
-However, similar to other results, only combat w/o covariates and combat with linear covariates are significantly worse.
-
-Does the number of regions w significant sex effects on mean change depending on combat version?
-
-```{r}
-anova.gam %>%
-  dplyr::filter(term=="sexMale") %>%
-  group_by(dataset) %>%
-  summarize(n_sig_regions = sum(sig.drop_bf.corr),
-            n_regions = n())
-```
-
-Not much, only for combat w/out covariates or linear covariates
-
-```{r, fig.heigt=15}
-gam.sum.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & term == "sexMale") %>%
-  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites"| dataset == "cf.lm-refA-data") %>%
-  ggplot() +
-  geom_col(aes(x=label, y=t_stat, fill=dataset), position="dodge") +
-  facet_wrap(~pheno_cat, scales = "free_x") + 
-  #theme_cowplot()+
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  #theme_void() +
-  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("t-statistic for effect of male sex on mean")
-
-gam.sum.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & term == "sexMale") %>%
-  ggplot() +
-    geom_density(aes(x=t_stat, fill=dataset), alpha=0.4) +
-  facet_wrap(~dataset)
-```
-
 #### OHBM Abstract

### 2023-12-05 eda/ukb_raw_analyses.Rmd
split out raw ukb_CN analyses from analyses of combat effects

--- 2023-12-05 ukb_sim.site_analyses.Rmd
@@
 ---
-title: "UKB Simulation Analyses"
+title: "UKB Raw Analyses"
 output: html_document
@@
 ---
+
+Mapping sex effects in raw, un-combatted data in order to establish deviation induced by ComBat-ing on simulated sites.
 
@@
 
-source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
+base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/"
+
+source(paste0(base_path, "cubic_scripts/R_scripts/gamlss_helper_funs.R"))
 # 
@@
 ```{r}
-ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
-ukb_df_age.filt <- read_csv(file="data/ukb_to_model/ukb_CN_data_simsites.csv")
+ukb_df_totals <- read_csv(file=paste0(base_path, "data/ukb_CN_data.csv"))
+ukb_df_age.filt <- read_csv(file=paste0(base_path, "data/ukb_to_model/ukb_CN_data_simsites.csv"))
 
-vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
-vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
-sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
-ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
-pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
+vol_list_global <- readRDS(file=paste0(base_path, "cubic_scripts/R_scripts/vol_list_global.rds"))
+vol_list_regions <- readRDS(file=paste0(base_path, "cubic_scripts/R_scripts/Vol_list_regions.rds"))
+sa_list <- readRDS(file=paste0(base_path, "cubic_scripts/R_scripts/SA_list.rds"))
+ct_list <- readRDS(file=paste0(base_path, "cubic_scripts/R_scripts/CT_list.rds"))
+pheno_list <- readRDS(file=paste0(base_path, "cubic_scripts/R_scripts/pheno_list.rds"))
 
-aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+#aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
 ```
@@
 ```{r}
-drop1.df <- fread("ukb_basic/drop1_tests.csv", stringsAsFactors = TRUE) %>%
+drop1.df <- fread(paste0(base_path, "ukb_basic/drop1_tests.csv"), stringsAsFactors = TRUE) %>%
   dplyr::select(-V1)
-sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE) %>%
+sex.df <- fread(paste0(base_path, "ukb_basic/sex_summary.csv"), stringsAsFactors = TRUE) %>%
   dplyr::select(-V1)
-sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
+sum.df <- fread(paste0(base_path, "ukb_basic/gamlss_summary.csv"), stringsAsFactors = TRUE) %>%
   dplyr::select(-V1) %>%
@@
 
-Simple bar plot of sites
-```{r}
-ggplot(ukb_df_age.filt) +
-  geom_bar(aes(x=sim.site, fill=sex), color="black", alpha=0.6) +
-  theme_linedraw(base_size=12)
-```
-
 ## Raw Data Results
-Review outcomes from raw (non-ComBat-ed) data
 
@@
 ```
-
-## ComBat Effects by Moment
-
-Loading combatted data (scp'd from cubic: `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/data/ukb_to_model/*.csv' ./data/ukb_to_model`) to see how values change with different combat methods
-
-```{r}
-data.csvs <- list.files(path = "data/ukb_to_model", pattern = ".csv", full.names = TRUE)
-combined_data <- data.frame()
-
-for (file in data.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file)))
-
-  # Bind the data to the combined dataframe
-  combined_data <- bind_rows(combined_data, data, .id = "File_ID")
-}
-
-table(combined_data$Source_File)
-```
-
-```{r}
-var.df <- combined_data %>%
-  group_by(Source_File, sim.site) %>%
-  summarize(across(all_of(pheno_list), var))
-
-var.long <- pivot_longer(var.df, cols=c(-Source_File, -sim.site), names_to = "feature", values_to = "variance")
-```
-
-```{r}
-var.long %>%
-  dplyr::filter(!(feature %in% vol_list_global)) %>%
-  ggplot() +
-  geom_density(aes(x=variance, fill=Source_File), alpha = 0.4, position="identity") +
-  facet_wrap(~sim.site)
-
-var.long %>%
-  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" |Source_File == "ukb_CN_data_simsites.csv" |  Source_File == "cf.gamlss_data.csv" )) %>%
-  ggplot() +
-  geom_density(aes(x=variance, fill=Source_File), alpha = 0.5, position="identity") +
-  facet_wrap(~sim.site)
-
-#Source_File == "cf.lm_data.csv" | 
-```
-
-```{r}
-var.by.sex <- combined_data %>%
-  group_by(Source_File, sex) %>%
-  summarize(across(all_of(pheno_list), var))
-
-var.by.sex.long <- pivot_longer(var.by.sex, cols=c(-Source_File, -sex), names_to = "feature", values_to = "variance")
-```
-
-```{r}
-var.by.sex.long %>%
-  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
-  ggplot() +
-  geom_density(aes(x=variance, fill=sex), alpha = 0.5, position="identity") +
-  facet_wrap(~Source_File)
-```
-
-Simple t test for sex effects across combat types
-```{r}
-sex_ttest_df <- data.frame("Source_File" = levels(combined_data$Source_File))
-
-for(pheno in pheno_list){
-  # New column name
-  new_col_name <- paste0(pheno, ".p")
-
-  #run t-tests
-  out <- lapply(split(combined_data, factor(combined_data$Source_File)), function(x)t.test(x[[pheno]] ~ x[["sex"]], paired=FALSE)$p.value)
-  stopifnot(names(out) == levels(combined_data$Source_File)) #make sure results are in correct order
-
-  #append to dataframe
-  sex_ttest_df[[new_col_name]] <- unlist(unname(out))
-}
-```
-
-```{r}
-sex_ttest_long <- sex_ttest_df %>% pivot_longer(!Source_File, values_to="p.value", names_to="pheno") %>%
-  mutate(pheno = sub(".p$", "", pheno),
-         sig_bfcor = case_when(pheno %in% vol_list_global & p.value < (0.05/length(vol_list_global)) ~ TRUE,
-                            pheno %in% vol_list_global & p.value >= (0.05/length(vol_list_global)) ~ FALSE,
-                            !(pheno %in% vol_list_global) & p.value < (0.05/length(ct_list)) ~ TRUE,
-                            !(pheno %in% vol_list_global) & p.value >= (0.05/length(ct_list)) ~ FALSE,
-                            TRUE ~ NA))
-sex_ttest_long %>%
-  group_by(Source_File) %>%
-  summarize(n_sig_regions = sum(sig_bfcor),
-            n_regions = n())
-```
-
-Basically same results gam/gamlss
-
-#### Mu
-
-Does the number of regions w significant age effects on mu change depending on combat version?
-
-```{r}
-drop1.df %>%
-  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
-  group_by(Dataset) %>%
-  summarize(n_sig_regions = sum(sig.drop_bf.corr),
-            n_regions = n())
-```
-
-No.
-
-Does the number of regions w significant sex effects on mu change depending on combat version?
-
-```{r}
-drop1.df %>%
-  dplyr::filter(Term=="sexMale" & Moment == "mu") %>%
-  group_by(Dataset) %>%
-  summarize(n_sig_regions = sum(sig.drop_bf.corr),
-            n_regions = n())
-
-sum.df %>%
-  dplyr::filter(term=="sexMale" & parameter == "mu") %>%
-  group_by(dataset) %>%
-  summarize(n_sig_regions = sum(sig.sum_bf.corr),
-            n_regions = n())
-```
-
-Slightly. Only combat w/o covariates or lm combat without a ref site are noticeably off. Consistent across drop1 and summary outputs
-
-```{r}
-sex.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "mu" & pheno_cat != "Global Volume") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = t_stat)) +
-  scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_void() +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("t-statistic for effect of male sex on mu")
-```
-
-Trying with ultra-strict threshold to see if results change:
-
-```{r}
-drop1.df <- drop1.df %>%
-  mutate(sig.sum.strict = case_when(drop1.pval < (0.05/length(pheno_list)) ~ TRUE,
-                            drop1.pval >= (0.05/length(pheno_list)) ~ FALSE,
-                            TRUE ~ NA))
-
-drop1.df %>%
-  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
-  group_by(Dataset) %>%
-  summarize(n_sig_regions = sum(sig.sum.strict),
-            n_regions = n())
-```
-
-No difference
-
-#### Sigma
-
-```{r}
-sex.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
-  ggplot() +
-  geom_col(aes(x=label, y=t_stat, fill=dataset)) +
-  facet_wrap(~pheno_cat, scales = "free_x") + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  #theme_void() +
-  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("t-statistic for effect of male sex on sigma")
-
-sex.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
-  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites") %>%
-  ggplot() +
-  geom_col(aes(x=label, y=t_stat, fill=dataset), position = "dodge") +
-  facet_wrap(~pheno_cat, scales = "free_x") + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  #theme_void() +
-  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("t-statistic for effect of male sex on sigma")
-
-sex.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = t_stat)) +
-  scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_void() +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("t-statistic for effect of male sex on sigma")
-```
-Check number of regions w significant sex effects in sigma
-```{r}
-sex.sigma.counts <- sex.df %>%
-  dplyr::filter(term=="sexMale" & parameter == "sigma") %>%
-  group_by(dataset, pheno_cat) %>%
-  summarize(n_sig_regions = sum(sig.drop_bf.corr),
-            n_regions = n())
-sex.sigma.counts
-```
-
-More variation
-```{r}
-ggplot(sex.sigma.counts) +
-  geom_col(aes(x=dataset, y=n_sig_regions, fill=dataset)) +
-  facet_wrap(~pheno_cat) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-```
-
-##### Variance
-Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex).
-Also getting mean variance.
-
-```{r}
-sigma.df <- sum.df %>%
-  dplyr::filter(parameter == "sigma") %>%
-  mutate(var_est = exp(estimate),
-         var_mean = exp(moment.mean),
-         beta_ratio = var_est/var_mean)
-
-skim(sigma.df$var_est)
-
-sigma.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & term =="sexMale") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_histogram(aes(x=var_est, fill = dataset), color="black", alpha=0.5, position = "identity") +
-  theme_linedraw() +
-  facet_wrap(~pheno_cat, scales= "free")
-```
-
-```{r, fig.height=10, fig.width=8}
-proportion_beta_plot <- sigma.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = beta_ratio)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_bw(base_size=15) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
-  ggtitle("Proportion of effect of male sex on variance")
-```
-
-```{r, fig.height=15, fig.width=12}
-beta.plot <- sigma.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = var_est)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_bw(base_size=20) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
-  ggtitle("Beta weight for effect of male sex on variance")
-
-ggsave("male_variance_beta.jpeg", beta.plot, width=12, height=15)
-```
-
-```{r, fig.height=5, fig.widt=5}
-sigma.df %>%
-  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale") %>%
-  mutate(var_est.sig = case_when(sig.sum_bf.corr == TRUE ~ var_est,
-                                 sig.sum_bf.corr == FALSE ~ 0)) %>%
-  #group_by(dataset) %>%
-  ggplot() +
-  geom_col(aes(x=dataset, y = var_est.sig, fill=dataset), position="dodge") +
-  #scale_fill_gradient2() +
-  facet_grid(~pheno) + 
-  theme_bw(base_size=12) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  theme(legend.position="none") +
-  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  ggtitle("Beta weight for effect of male sex on global volumes' variance")
-
-sigma.df.glob <- sigma.df %>%
-  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale")
-table(sigma.df.glob$dataset, sigma.df.glob$sig.sum_bf.corr)
-```
-
-fix age, get diff between male and female pred. variance
-```{r}
-pred.csvs <- list.files(path = "ukb_basic", pattern = "_variance.csv", full.names = TRUE)
-combined_var <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_variance.csv", "", Source_File))
-
-  # Bind the data to the combined dataframe
-  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
-}
-
-table(combined_var$dataset)
-#skim(combined_preds)
-```
-
-```{r}
-combined_var <- combined_var %>%
-  mutate(sex_std = sex_effect/f.var)
-
-summary(sigma.df$var_est)
-summary(combined_var$sex_effect)
-
-summary(sigma.df$beta_ratio)
-summary(combined_var$sex_std)
-```
-
-```{r}
-plot.info <- sigma.df %>%
-  dplyr::filter(term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
-  unique() %>%
-  mutate(dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset))
-combined_var <- inner_join(combined_var, plot.info)
-table(combined_var$dataset)
-```
-
-plot
-
-```{r, fig.height=5, fig.widt=5}
-combined_var %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = sex_std)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_bw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
-  ggtitle("Standardized effect of male sex on variance")
-```
-
-using aseg
-
-```{r}
-plot(aseg)
-```
-
-## ComBat Effects on Centiles
-
-Do centile scores change based on combat version?
-```{r}
-pred.csvs <- list.files(path = "ukb_basic", pattern = "_predictions.csv", full.names = TRUE)
-combined_preds <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
-
-  # Bind the data to the combined dataframe
-  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
-}
-
-table(combined_preds$dataset)
-#skim(combined_preds)
-```
-calc how much centiles change
-```{r}
-cent_diffs <- combined_preds %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
-  ungroup()
-
-diff_vol_glob <- sapply(vol_list_global, function(x) paste0("abs.diff_", x))
-
-mean.cent.diffs <- cent_diffs %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup()
-
-max.cent.diffs <- cent_diffs %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), max)) %>%
-  ungroup()
-#names(max.cent.diffs)
-```
-
-Plot average differences in centiles
-```{r}
-mean.diffs.long <- mean.cent.diffs %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-mean.diffs.long <- left_join(mean.diffs.long, dk.parc, by="label")
-```
-
-```{r}
-mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-ggplot() +
-  geom_density(aes(x=centile_abs.diff, fill=dataset), alpha=0.5) +
-  facet_grid(dataset~pheno_cat)
-
-mean.diffs.long %>%
-  dplyr::filter(dataset== "cf.gam" | dataset == "cf.gam_refA" | dataset == "cf.gamlss" | dataset == "cf.gamlss_refA") %>%
-  ggplot() +
-  geom_density(aes(x=centile_abs.diff, fill=dataset), alpha=0.5) +
-  scale_x_continuous(labels = label_comma())
-
-mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visualization purposes
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", binwidth=0.1, alpha=0.5, position="identity") +
-  theme_linedraw() +
-  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  scale_x_continuous(labels = label_comma()) +
-  ggtitle("Mean absolute differences in centile scores across ComBat \nprocessing pipelines for each phenotype")
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 12) +
-  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
-  paletteer::scale_fill_paletteer_d("Polychrome::light") +
-  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
-ggsave("mean_centile_diffs.jpeg", cent.diffs.plot)
-```
-
-```{r, fig.height=15, fig.width=12}
-mean.diffs.long %>%
-  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = centile_abs.diff)) +
-  scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_void() +
-  theme_bw(base_size=20) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("Mean difference in centiles across ComBat configurations")
-```
-
-Look at max difference per phenotype
-```{r, fig.height=6, fig.width=7}
-max.diffs.long <- max.cent.diffs %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-
-max.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(max_centile_abs.diff = max_centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 12) +
-  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
-  paletteer::scale_fill_paletteer_d("Polychrome::light") +
-  ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
-```
-
-```{r}
-max.diffs.long %>%
-  mutate(max_centile_abs.diff = max_centile_abs.diff*100) %>% #for visuals
-  dplyr::filter(dataset != "ukb_CN") %>%
-  ggplot() +
-  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity", binwidth = 0.5) +
-  theme_linedraw(base_size = 12) +
-  scale_x_continuous(labels = label_comma())+
-  ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
-```
-
-Test for differences in distribution
-```{r}
-#get centile errors for each subject
-cent.diffs2 <- combined_preds %>%
-  dplyr::filter(!str_detect(dataset, "_refA")) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "diff_{.col}")) %>%
-  ungroup() %>%
-  dplyr::filter(dataset != "ukb_CN") #drop ref
-
-# cf.gmvdiff <- cent.diffs2 %>%
-#   dplyr::filter(dataset == "cf") %>%
-#   dplyr::select(diff_GMV)
-# cf.lm.gmvdiff <- cent.diffs2 %>%
-#   dplyr::filter(dataset == "cf.lm") %>%
-#   dplyr::select(diff_GMV) 
-# 
-# t.test(cf.gmvdiff, cf.lm.gmvdiff, paired=TRUE)
-```
-
-NOTE: "missing" paired comparisons between cf and cf.lm diffs are because diff values are identical between the two methods (i.e.
-there is literally no difference in accuracy of centile scores between the two)
-
-```{r, warning=FALSE, message=FALSE}
-#within each phenotype, iterate paired t test between each combo of combat dfs
-
-#initalize empty dfs to store outputs
-sum.df <- data.frame(pheno = character(),
-                     term = character(),
-                     df = double(),
-                     sumsq = double(),
-                     meansq = double(),
-                     statistic = double(),
-                     p.value = double())
-t.df <- data.frame("pheno" = character(),
-                   "group1" = character(),
-                   "group2" = character(),
-                   "p.value" = double())
-pheno_diff.list <- paste0("diff_", pheno_list)
-
-#run tests
-attach(cent.diffs2)
-for (pheno in pheno_diff.list) {
-  df.aov <- tidy(aov(cent.diffs2[[pheno]] ~ cent.diffs2[["dataset"]]))
-  df.aov$pheno <- pheno
-  sum.df <- rbind(sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.t.test(x=cent.diffs2[[pheno]], g=cent.diffs2[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  t.df <- rbind(t.df, df.pairwise.t)
-}
-detach()
-```
-
-```{r}
-#apply fdr correction
-sum.df <- sum.df %>%
-  dplyr::filter(term != "Residuals") %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-summary(sum.df$p.val_fdr)
-
-t.df <- t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("diff_", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-table(t.df$comp)
-
-t.df %>%
-  group_by(comp) %>%
-   summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min_pval = min(p.val_fdr),
-            max_pval = max(p.val_fdr))
-```
-
-#### Sex-Biases
-
-Do the different combat methods introduce any kind of bias in centile scores?
-
-```{r}
-mean.cent.diffs.by.sex <- cent_diffs %>%
-  group_by(dataset, sex) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-
-mean.cent.diffs.by.sex %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(centile_diff = centile_diff *100) %>%
-ggplot() +
-  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
-  facet_wrap(~dataset, scales = "free_y")
-
-max.cent.diffs.by.sex <- cent_diffs %>%
-  group_by(dataset, sex) %>%
-  summarize(across(starts_with("abs.diff_"), max)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-
-max.cent.diffs.by.sex %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(max_centile_diff = max_centile_diff *100) %>%
-ggplot() +
-  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
-  facet_wrap(~dataset, scales = "free_y")
-```
-
-No shift in abs. differences, looking at overall diff.
-```{r}
-#get centile errors for each subject - not abs val
-cent.diffs3 <- combined_preds %>%
-  dplyr::filter(!str_detect(dataset, "_refA")) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
-  ungroup() %>%
-  dplyr::filter(dataset != "ukb_CN") #drop ref
-saveRDS(cent.diffs3, file="ukb_basic/cent_diffs.RDS")
-```
-
-```{r, eval=FALSE}
-cent.diffs3 %>%
-  #dplyr::filter(dataset == "cf.gamlss") %>%
-  pivot_longer(starts_with("diff_"), names_to="pheno", values_to="diff", names_prefix = "diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_))) %>%
-  ggplot() +
-  geom_density(aes(x=diff, fill=dataset), alpha=0.4) +
-  facet_wrap(~pheno_cat) +
-  theme(legend.position="none")
-```
-
-```{r}
-max.cent.diffs.by.sex2 <- cent.diffs3 %>%
-  group_by(dataset, sex) %>%
-  summarize_at(vars(starts_with("diff")), list(max_value = ~.[which.max(abs(.))])) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-
-max.cent.diffs.by.sex2 %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(max_centile_diff = max_centile_diff *100) %>%
-ggplot() +
-  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
-  facet_wrap(~dataset, scales="free")
-```
-
-Other methods are fairly consistently underestimating male centiles and overestimating female centiles.
-
-```{r}
-mean.cent.diffs.by.sex2 <- cent.diffs3 %>%
-  group_by(dataset, sex) %>%
-  summarize(across(starts_with("diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-
-mean.cent.diffs.by.sex2 %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(centile_diff = centile_diff *100) %>%
-ggplot() +
-  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
-  facet_wrap(~dataset, scales="free")
-```
-
-T test of mean diff. between sex w/in each dataset
-```{r}
-t.test.by.sex <- mean.cent.diffs.by.sex2 %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  group_by(dataset) %>%
-  summarise(t_statistic = t.test(centile_diff ~ sex)$statistic,
-            p_value = t.test(centile_diff ~ sex)$p.value) %>%
-  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
-         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
-                             p_val_fdr.corr >= 0.05 ~ FALSE))
-
-t.test.by.sex
-```
-
-Both gam methods and lm combat with ref both end up inducing a bias in how males and females centiles are shifted.
-
-```{r, warning=FALSE, message=FALSE}
-# within each df, within each phenotype, t-test between male and female diffs
-
-#initalize empty dfs to store outputs
-t.df.sex <- data.frame("pheno" = character(),
-                   "dataset" = character(),
-                   "p_value" = double()
-                   )
-
-for (pheno in pheno_diff.list){
-    df.sex.test <- cent.diffs3 %>%
-        group_by(dataset) %>%
-        summarize(p_value = t.test(.data[[pheno]][sex == "Male"], .data[[pheno]][sex == "Female"])$p.value) %>%
-        ungroup()
-    df.sex.test$pheno <- pheno
-    t.df.sex <- rbind(t.df.sex, df.sex.test)
-}
-```
-
-```{r}
-#apply fdr correction
-t.df.sex <- t.df.sex %>%
-  mutate(p.val_fdr = p.adjust(p_value, method="fdr", n = length(p_value)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("diff_", "", pheno))
-
-t.df.sex %>%
-  group_by(dataset) %>%
-   summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n())
-```
-
-## ComBat Effects on Z-Score
-calc how much z-scores change
-
-```{r}
-z_diffs <- combined_preds %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref = "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
-  ungroup()
-```
-
-```{r}
-#mean
-mean.z.diffs <- z_diffs %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup()
-mean.z.long <- mean.z.diffs %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno = sub(".z", "", pheno)) %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-mean.z.long <- left_join(mean.z.long, dk.parc, by="label")
-
-#max
-max.z.diffs <- z_diffs %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), max)) %>%
-  ungroup()
-max.z.long <- max.z.diffs %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno = sub(".z", "", pheno)) %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-```
-
-```{r}
-mean.z.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  ggplot() +
-  geom_histogram(aes(x=z_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity") +
-  theme_linedraw() +
-  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  scale_x_continuous(labels = label_comma()) +
-  ggtitle("Mean absolute differences in Z-scores across ComBat \nprocessing pipelines for each phenotype")
-```
-
-```{r, fig.height=15, fig.width=12}
-mean.z.long %>%
-  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = z_abs.diff)) +
-  scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_void() +
-  theme_bw(base_size=20) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("Mean difference in z scores across ComBat configurations")
-```
-
-```{r}
-max.z.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  ggplot() +
-  geom_histogram(aes(x=max_z_abs.diff, fill=dataset), color="black", binwidth = 0.025, alpha=0.5, position="identity") +
-  theme_linedraw() +
-  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  scale_x_continuous(labels = label_comma()) +
-  ggtitle("Mean absolute differences in Z-scores across ComBat \nprocessing pipelines for each phenotype")
-```
-
-```{r, fig.height=15, fig.width=12}
-max.z.long %>%
-  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = max_z_abs.diff)) +
-  scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_void() +
-  theme_bw(base_size=20) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("Maximum difference in z scores across ComBat configurations")
-```
-
-Test for differences in distribution
-
-```{r, warning=FALSE, message=FALSE}
-#within each phenotype, iterate paired t test between each combo of combat dfs
-
-z_diffs.filt <- z_diffs %>%
-  dplyr::filter(!str_detect(dataset, "_refA")) %>%
-  dplyr::filter(dataset != "ukb_CN")
-
-#initalize empty dfs to store outputs
-z.sum.df <- data.frame(pheno = character(),
-                     term = character(),
-                     df = double(),
-                     sumsq = double(),
-                     meansq = double(),
-                     statistic = double(),
-                     p.value = double())
-z.t.df <- data.frame("pheno" = character(),
-                   "group1" = character(),
-                   "group2" = character(),
-                   "p.value" = double())
-
-z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")
-
-#run tests
-attach(z_diffs.filt)
-for (pheno in z.pheno_abs.diff.list) {
-  df.aov <- tidy(aov(z_diffs.filt[[pheno]] ~ z_diffs.filt[["dataset"]]))
-  df.aov$pheno <- pheno
-  z.sum.df <- rbind(z.sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.t.test(x=z_diffs.filt[[pheno]], g=z_diffs.filt[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  z.t.df <- rbind(z.t.df, df.pairwise.t)
-}
-detach()
-```
-
-```{r}
-#apply fdr correction
-z.sum.df <- z.sum.df %>%
-  dplyr::filter(term != "Residuals") %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-summary(z.sum.df$p.val_fdr)
-
-z.t.df <- z.t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno),
-         pheno = sub(".z", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-z.t.df %>%
-  group_by(comp) %>%
-   summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min.pval = min(p.val_fdr),
-            max.pval = max(p.val_fdr))
-```
-
-##### Sex-Biases
-
-Test for sex-specific differences in mean z-score offset
-
-```{r}
-z_diffs2 <- combined_preds %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(ends_with(".z"), ~ . - first(.), .names = "diff_{.col}")) %>%
-  ungroup()
-#write out
-saveRDS(z_diffs2, file = "ukb_basic/z_diffs.RDS")
-
-mean.z.diffs.by.sex2 <- z_diffs2 %>%
-  group_by(dataset, sex) %>%
-  summarize(across(starts_with("diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("diff_"), values_to="z_diff", names_to = "pheno", names_prefix = "diff_") %>%
-  mutate(pheno = gsub(".z", "", pheno)) %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-```
-t-test
-```{r}
-t.test.by.sex_z <- mean.z.diffs.by.sex2 %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  group_by(dataset) %>%
-  summarise(t_statistic = t.test(z_diff ~ sex)$statistic,
-            p_value = t.test(z_diff ~ sex)$p.value) %>%
-  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
-         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
-                             p_val_fdr.corr >= 0.05 ~ FALSE))
-
-t.test.by.sex_z
-```
-
-Same results as centiles.
-
-#### OHBM Abstract
-
-rerunning stats above w/o ref-site combat versions (to account for different FDR correction) per Andrew's suggestion
-
-Test
-
-Test for differences in centile offset distributions
-```{r}
-mean.diffs.long_filt <- mean.diffs.long %>%
-  filter(!str_detect(dataset, "_refA")) %>%
-  filter(dataset != "ukb_CN")
-
-summary(aov(centile_abs.diff ~ dataset, data=mean.diffs.long_filt))
-
-attach(mean.diffs.long_filt)
-pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "bonferroni")
-detach()
-```
-
-Testing for differences in Z offset distributions
-
-```{r}
-mean.z.long_filt <- mean.z.long %>%
-  filter(!str_detect(dataset, "_refA"))
-
-summary(aov(z_abs.diff ~ dataset, data=mean.z.long_filt))
-
-attach(mean.z.long_filt)
-pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr")
-detach()
-```
-
-Test for sex-biases in centile offsets
-
-Test for sex-biases
-
-```{r}
-mean.z.diffs.by.sex2 %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  filter(!str_detect(dataset, "_refA")) %>%
-  group_by(dataset) %>%
-  summarise(t_statistic = t.test(z_diff ~ sex)$statistic,
-            p_value = t.test(z_diff ~ sex)$p.value) %>%
-  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
-         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
-                             p_val_fdr.corr >= 0.05 ~ FALSE))
-
-mean.cent.diffs.by.sex2 %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  filter(!str_detect(dataset, "_refA")) %>%
-  group_by(dataset) %>%
-  summarise(t_statistic = t.test(centile_diff ~ sex)$statistic,
-            p_value = t.test(centile_diff ~ sex)$p.value) %>%
-  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
-         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
-                             p_val_fdr.corr >= 0.05 ~ FALSE))
-```
-
-##### Plots
-
-Beta weight? or proportion?
-
-```{r}
-beta.plot2 <- sigma.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
-  filter(!str_detect(dataset, "-refA")) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"))) %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = var_est)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_bw(base_size=20) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
-  ggtitle("Beta weight for effect of male sex on variance")
-
-beta.plot2
-
-ggsave("male_variance_beta2.jpeg", beta.plot2, width=10, height=10)
-```
-
-```{r}
-sigma.aseg_all <- sigma.df %>%
-  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale") %>%
-  filter(!str_detect(dataset, "-refA")) %>%
-  dplyr::select(!label) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss")),
-         tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "White Matter",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-         var_est = case_when(sig.sum_bf.corr == TRUE ~ var_est,
-                             TRUE ~ NA)) #just keep significant regions
-
-sigma.aseg_all <- left_join(sigma.aseg_all, aseg_cerebrum$data, by = c("tissue_class"))
-
-sigma.aseg_all %>%
-  group_by(dataset) %>%
-  ggplot() +
-  geom_brain(atlas = aseg_cerebrum, 
-             #position = position_brain(hemi ~ side),
-             aes(fill = var_est),
-             alpha=0.8) +
-  #scale_fill_gradient2() +
-  facet_wrap(~dataset) + 
-  theme_bw(base_size=20) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
-  ggtitle("Beta weight for effect of male sex on variance")
-```
-
-```{r}
-proportion_beta_plot2 <- sigma.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
-  filter(!str_detect(dataset, "-refA")) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"))) %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = beta_ratio)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_bw(base_size=15) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
-  ggtitle("Proportion of effect of male sex on variance")
-
-proportion_beta_plot2
-ggsave("male_variance_prop.beta2.jpeg", proportion_beta_plot2, width=10, height=10)
-```
-
-Centile plot
-
-```{r}
-#Define scales
-scales_y <- list(
-  `Global Volume` = scale_y_continuous(limits = c(0, 4)),
-  `Regional CT` = scale_y_continuous(limits = c(0, 70)),
-  `Regional SA` = scale_y_continuous(limits = c(0, 70)),
-  `Regional Volume` = scale_y_continuous(limits = c(0, 70))
-)
-
-cent.diffs.plot2 <- mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  filter(!str_detect(dataset, "_refA")) %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.4, position="identity") +
-  theme_linedraw(base_size = 12) +
-  scale_x_continuous(labels = label_comma()) + 
-  facet_wrap(~pheno_cat) +
-  #facetscales::facet_grid_sc(pheno_cat~.,
-                #scales = list(y = scales_y)) +
-  #paletteer::scale_fill_paletteer_d("Polychrome::light") +
-  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
-
-cent.diffs.plot2
-#ggsave("mean_centile_diffs2.jpeg", cent.diffs.plot2, width=10, height=10)
-```

### 2023-12-05 ukb_sim.site_analyses.Rmd
reorg

@@
 
-## Raw Data Results
-Review outcomes from raw (non-ComBat-ed) data
-
-```{r, fig.width=11}
-sex.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset ==	"ukb-CN-data-simsites") %>%
-  ggplot() +
-  geom_col(aes(x=reorder(pheno,t_stat), y=t_stat, fill=pheno_cat)) + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
-
-sex.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
-  ggplot() +
-  geom_col(aes(x=label, y=t_stat, fill=pheno_cat)) + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
-
-sex.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
-  ggplot() +
-  geom_col(aes(x=region, y=t_stat, fill=hemi)) + 
-  facet_wrap(~pheno_cat) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
-```
-
-Plot on brains!
-
-```{r}
-sex.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
-  group_by(pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = t_stat)) +
-  scale_fill_gradient2() +
-  facet_wrap(~pheno_cat) +
-  theme_void() +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
-  ggtitle("t-statistic for effect of male sex on sigma")
-```
-
-## ComBat Effects by Moment
+## By Moment
 
@@
 
+Plot distribution of features' variances across each site
+
 ```{r}
@@
   facet_wrap(~sim.site)
-
-#Source_File == "cf.lm_data.csv" | 
-```
+```
+How does male vs. female variance across features change by combat type
 
@@
 
-Simple t test for sex effects across combat types
+Simple t test for sex effects on brain features to see whether # of significantly sex-biased features change across combat configurations
+
 ```{r}
@@
 
-Basically same results gam/gamlss
+Basically no difference across combat types
 
@@
 Check number of regions w significant sex effects in sigma
+
 ```{r}
@@
 ```
-
-More variation
+ Plot these results
+
 ```{r}
@@
   facet_wrap(~pheno_cat) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-```
-
-##### Variance
-Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex).
-Also getting mean variance.
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("# regions with significant effects of male sex on sigma")
+```
+
+## Variance
+Estimating by 3 different methods:
+1. Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex).
+1.5. getting mean variance exp(mean sigma) (not actually plotted)
+2. actual estimated variance at mean(sample age) (andrew's rec)
+
+#### Method 1: exp(sigma)
 
@@
   theme_linedraw() +
-  facet_wrap(~pheno_cat, scales= "free")
-```
+  facet_wrap(~pheno_cat, scales= "free")+
+  ggtitle("estimated variance in regions with significant effects of male sex on sigma")
+```
+
+What proportion of estimated variance is attributable to sex?
 
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork") +
   ggtitle("Proportion of effect of male sex on variance")
 ```
+
+exp(SexMale beta weight on sigma) -> estimate of male effect on variance
 
@@
 
-ggsave("male_variance_beta.jpeg", beta.plot, width=12, height=15)
+ggsave("figs/male_variance_beta.jpeg", beta.plot, width=12, height=15)
 ```
@@
 
+#### Method 2: Var at Mean Age
 fix age, get diff between male and female pred. variance
@@
 
-table(combined_var$dataset)
+table(combined_var$dataset) #confirm all data's there
 #skim(combined_preds)
@@
 
+Calculate standardized effects of male sex w/in region as (M var - F var)/F var
+
 ```{r}
 combined_var <- combined_var %>%
-  mutate(sex_std = sex_effect/f.var)
-
-summary(sigma.df$var_est)
-summary(combined_var$sex_effect)
-
-summary(sigma.df$beta_ratio)
-summary(combined_var$sex_std)
-```
-
-```{r}
+  mutate(sex_std = sex_effect/f.var) #standardized effect of male sex on variance
+
+#compare methods
+summary(sigma.df$var_est) #method 1 var
+summary(combined_var$sex_effect) #method 2 var
+
+summary(sigma.df$beta_ratio). #method 1 male effect on var
+summary(combined_var$sex_std) #method 2 male effect on var
+```
+
+```{r}
+#join with sigma.df for significance testing
 plot.info <- sigma.df %>%
@@
 combined_var <- inner_join(combined_var, plot.info)
-table(combined_var$dataset)
+table(combined_var$dataset) #check no data's lost
 ```
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, -0.05, 0, 0.05, 0.1)) +
   ggtitle("Standardized effect of male sex on variance")
@@
 
-using aseg
-
-```{r}
-plot(aseg)
-```
-
-## ComBat Effects on Centiles
+## Centiles
 
@@
 ```
-calc how much centiles change
+
+#### Magnitude error
+
+Calculate how far centiles are from "truth" (raw data's centile vals), take absolute value for magnitude error
+
 ```{r}
@@
   ungroup()
-
-diff_vol_glob <- sapply(vol_list_global, function(x) paste0("abs.diff_", x))
-
+```
+
+Calculate within feature mean and max abs. centile errors across all subjects
+
+```{r}
 mean.cent.diffs <- cent_diffs %>%
@@
 
-Plot average differences in centiles
+Plot w/in feature average magnitude of centile errors
 ```{r}
@@
 Look at max difference per phenotype
+
 ```{r, fig.height=6, fig.width=7}
@@
 
-Test for differences in distribution
-```{r}
-#get centile errors for each subject
+Test for differences in distribution of subjects' abs(cent error)
+
+```{r}
+#get abs val of centile errors for each subject
 cent.diffs2 <- combined_preds %>%
@@
 ```
+
+plot to see if assumptions for t-tests hold ...
+```{r}
+ggplot(cent.diffs2) +
+  geom_histogram(aes(x=diff_rh_CT_posteriorcingulate, fill=dataset), alpha = 0.6)
+```
+
+looks like.... no.
 
@@
 
-  df.pairwise.t <- tidy(pairwise.t.test(x=cent.diffs2[[pheno]], g=cent.diffs2[["dataset"]], paired = TRUE, p.adj = "none"))
+  df.pairwise.t <- tidy(pairwise.wilcox.test(x=cent.diffs2[[pheno]], g=cent.diffs2[["dataset"]], paired = TRUE, p.adj = "none")) #update to wilcox since t assumptions not all met
   df.pairwise.t$pheno <- pheno
@@
 
-## ComBat Effects on Z-Score
+## Z-Score
 calc how much z-scores change
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork") +
   ggtitle("Beta weight for effect of male sex on variance")

### 2023-12-06 paper_figs.Rmd
adding gams for extremeness bias tests

@@
 
-source("./R_scripts/gamlss_helper_funs.R")
+source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
 ```
@@
 ### LOAD ###
-pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
-vol_list_global <- readRDS(file="R_scripts/vol_list_global.rds")
-vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
-sa_list <- readRDS(file="R_scripts/SA_list.rds")
-ct_list <- readRDS(file="R_scripts/CT_list.rds")
+pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
+vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
+vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
+sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
+ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
 
@@
 #ukb data from ohbm_results.Rmd
-cent_err.p <- readRDS("ukb_perm_csvs/centile_errors.RDS")
+cent_err.p <- readRDS("data/ukb_perm_csvs/centile_errors.RDS")
 z_diffs2 <- readRDS("ukb_basic/z_diffs.RDS")

### 2023-12-06 ukb_sim.site_analyses.Rmd
adding gams for extremeness bias tests

@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr, paletteer, ggsegExtra, ggseg3d)
+pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr, paletteer, ggsegExtra, ggseg3d, mgcv)
 
@@
 
-summary(sigma.df$beta_ratio). #method 1 male effect on var
+summary(sigma.df$beta_ratio) #method 1 male effect on var
 summary(combined_var$sex_std) #method 2 male effect on var
@@
 
+**check w andrew**
+
 ```{r, warning=FALSE, message=FALSE}
-#within each phenotype, iterate paired t test between each combo of combat dfs
+#within each phenotype, iterate paired wilcox rank test between each combo of combat dfs
 
@@
 No shift in abs. differences, looking at overall diff.
+
 ```{r}
@@
 
-```{r, eval=FALSE}
-cent.diffs3 %>%
-  #dplyr::filter(dataset == "cf.gamlss") %>%
-  pivot_longer(starts_with("diff_"), names_to="pheno", values_to="diff", names_prefix = "diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_))) %>%
-  ggplot() +
-  geom_density(aes(x=diff, fill=dataset), alpha=0.4) +
-  facet_wrap(~pheno_cat) +
-  theme(legend.position="none")
-```
+Look at w/in subject maximum error across all feature's centiles & compare across harmonization methods
 
@@
 
-Other methods are fairly consistently underestimating male centiles and overestimating female centiles.
+all methods are fairly consistently underestimating male centiles and overestimating female centiles.
 
@@
 
-T test of mean diff. between sex w/in each dataset
+test of mean diff. between sex w/in each dataset - again need to update w nonparametric method **ask andrew about this**
 ```{r}
@@
 Both gam methods and lm combat with ref both end up inducing a bias in how males and females centiles are shifted.
+
+t tests of sex-biases in centiles w/in each feature **ask andrew about this**
 
@@
 ```
+
+### Extremeness Bias
+
+Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)
+
+do some plotting
+
+```{r}
+ggplot(cent.diffs3, aes(x=GMV, y=diff_GMV, color=sex)) +
+  geom_point(alpha=0.1) +
+  geom_smooth() +
+  facet_wrap(~dataset)
+
+ggplot(cent.diffs3, aes(x=WMV, y=diff_WMV, color=sex)) +
+  geom_point(alpha=0.1) +
+  geom_smooth() +
+  facet_wrap(~dataset, scales="free")
+```
+
+Testing on MAGNITUDE of error, not the errors themselves!
+
+```{r}
+ggplot(cent.diffs2, aes(x=GMV, y=diff_GMV, color=sex)) +
+  geom_point(alpha=0.1) +
+  geom_smooth() +
+  facet_wrap(~dataset)
+
+ggplot(cent.diffs2, aes(x=WMV, y=diff_WMV, color=sex)) +
+  geom_point(alpha=0.1) +
+  geom_smooth() +
+  facet_wrap(~dataset, scales="free")
+```
+
+*ask andrew about how best to test - would GAM be too flexible? use polynomials?*
 
@@
 
-#### OHBM Abstract
-
-rerunning stats above w/o ref-site combat versions (to account for different FDR correction) per Andrew's suggestion
-
-Test
-
-Test for differences in centile offset distributions
-```{r}
-mean.diffs.long_filt <- mean.diffs.long %>%
-  filter(!str_detect(dataset, "_refA")) %>%
-  filter(dataset != "ukb_CN")
-
-summary(aov(centile_abs.diff ~ dataset, data=mean.diffs.long_filt))
-
-attach(mean.diffs.long_filt)
-pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "bonferroni")
-detach()
-```
-
-Testing for differences in Z offset distributions
-
-```{r}
-mean.z.long_filt <- mean.z.long %>%
-  filter(!str_detect(dataset, "_refA"))
-
-summary(aov(z_abs.diff ~ dataset, data=mean.z.long_filt))
-
-attach(mean.z.long_filt)
-pairwise.t.test(x=z_abs.diff, g=dataset, p.adj = "fdr")
-detach()
-```
-
-Test for sex-biases in centile offsets
-
-Test for sex-biases
-
-```{r}
-mean.z.diffs.by.sex2 %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  filter(!str_detect(dataset, "_refA")) %>%
-  group_by(dataset) %>%
-  summarise(t_statistic = t.test(z_diff ~ sex)$statistic,
-            p_value = t.test(z_diff ~ sex)$p.value) %>%
-  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
-         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
-                             p_val_fdr.corr >= 0.05 ~ FALSE))
-
-mean.cent.diffs.by.sex2 %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  filter(!str_detect(dataset, "_refA")) %>%
-  group_by(dataset) %>%
-  summarise(t_statistic = t.test(centile_diff ~ sex)$statistic,
-            p_value = t.test(centile_diff ~ sex)$p.value) %>%
-  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
-         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
-                             p_val_fdr.corr >= 0.05 ~ FALSE))
-```
-
-##### Plots
-
-Beta weight? or proportion?
-
-```{r}
-beta.plot2 <- sigma.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
-  filter(!str_detect(dataset, "-refA")) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"))) %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = var_est)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_bw(base_size=20) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork") +
-  ggtitle("Beta weight for effect of male sex on variance")
-
-beta.plot2
-
-ggsave("male_variance_beta2.jpeg", beta.plot2, width=10, height=10)
-```
-
-```{r}
-sigma.aseg_all <- sigma.df %>%
-  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale") %>%
-  filter(!str_detect(dataset, "-refA")) %>%
-  dplyr::select(!label) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss")),
-         tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "White Matter",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-         var_est = case_when(sig.sum_bf.corr == TRUE ~ var_est,
-                             TRUE ~ NA)) #just keep significant regions
-
-sigma.aseg_all <- left_join(sigma.aseg_all, aseg_cerebrum$data, by = c("tissue_class"))
-
-sigma.aseg_all %>%
-  group_by(dataset) %>%
-  ggplot() +
-  geom_brain(atlas = aseg_cerebrum, 
-             #position = position_brain(hemi ~ side),
-             aes(fill = var_est),
-             alpha=0.8) +
-  #scale_fill_gradient2() +
-  facet_wrap(~dataset) + 
-  theme_bw(base_size=20) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
-  ggtitle("Beta weight for effect of male sex on variance")
-```
-
-```{r}
-proportion_beta_plot2 <- sigma.df %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
-  filter(!str_detect(dataset, "-refA")) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"))) %>%
-  group_by(dataset, pheno_cat) %>%
-  ggplot() +
-  geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = beta_ratio)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_bw(base_size=15) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
-  ggtitle("Proportion of effect of male sex on variance")
-
-proportion_beta_plot2
-ggsave("male_variance_prop.beta2.jpeg", proportion_beta_plot2, width=10, height=10)
-```
-
-Centile plot
-
-```{r}
-#Define scales
-scales_y <- list(
-  `Global Volume` = scale_y_continuous(limits = c(0, 4)),
-  `Regional CT` = scale_y_continuous(limits = c(0, 70)),
-  `Regional SA` = scale_y_continuous(limits = c(0, 70)),
-  `Regional Volume` = scale_y_continuous(limits = c(0, 70))
-)
-
-cent.diffs.plot2 <- mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  filter(!str_detect(dataset, "_refA")) %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.4, position="identity") +
-  theme_linedraw(base_size = 12) +
-  scale_x_continuous(labels = label_comma()) + 
-  facet_wrap(~pheno_cat) +
-  #facetscales::facet_grid_sc(pheno_cat~.,
-                #scales = list(y = scales_y)) +
-  #paletteer::scale_fill_paletteer_d("Polychrome::light") +
-  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
-
-cent.diffs.plot2
-#ggsave("mean_centile_diffs2.jpeg", cent.diffs.plot2, width=10, height=10)
-```
+### Extremeness Bias
+
+Plot error
+
+```{r}
+ggplot(z_diffs2, aes(x=GMV.z, y=diff_GMV.z, color=sex)) +
+  geom_point(alpha=0.1) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset)
+
+ggplot(z_diffs2, aes(x=WMV.z, y=diff_WMV.z, color=sex)) +
+  geom_point(alpha=0.1) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset)
+```
+
+plot magnitude error (abs val)
+
+```{r}
+ggplot(z_diffs, aes(x=GMV.z, y=abs.diff_GMV.z, color=sex)) +
+  geom_point(alpha=0.1) +
+  geom_smooth() +
+  facet_wrap(~dataset)
+
+ggplot(z_diffs, aes(x=WMV.z, y=abs.diff_WMV.z, color=sex)) +
+  geom_point(alpha=0.1) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset)
+```
+
+Try using gam to test associations (start with just GMV)
+
+```{r}
+m1 <- gam(abs.diff_GMV.z ~ s(GMV.z, by=dataset) + dataset, data=z_diffs)
+summary(m1)
+
+#test for interactive sex effects - split df by dataset and smooth by sex
+for (dset in c("cf", "cf.gam", "cf.gamlss", "cf.lm")) {
+  print(dset)
+  df <- z_diffs %>%
+    dplyr::filter(dataset == dset)
+
+  m2 <- gam(abs.diff_GMV.z ~ s(GMV.z, by=as.factor(sex)) + as.factor(sex) + s(GMV.z), data=df)
+  print(summary(m2))
+}
+```
+
+Pivot data long and run gam to test assoc. across all features
+```{r}
+dataset_test_list <- c("cf", "cf.gam", "cf.gamlss", "cf.lm")
+pheno_list.z <- paste0(pheno_list, ".z")
+
+z_diffs.long <- z_diffs %>%
+  dplyr::filter(dataset %in% dataset_test_list) %>%
+  mutate(dataset = droplevels(dataset)) %>%
+  dplyr::select(c(participant, sex, age_days, dataset, Source_File), starts_with("abs.diff_")) %>%
+  pivot_longer(starts_with("abs.diff_"), names_to="pheno", values_to="z_abs.diff", names_prefix = "abs.diff_")
+
+z_long <- z_diffs %>%
+  dplyr::filter(dataset %in% dataset_test_list) %>%
+  mutate(dataset = droplevels(dataset)) %>%
+  dplyr::select(c(participant, sex, age_days, dataset, Source_File, pheno_list.z)) %>%
+  pivot_longer(cols=all_of(pheno_list.z), names_to="pheno", values_to="z_score")
+
+z_long.full <- inner_join(z_diffs.long, z_long)
+```
+
+```{r}
+gam.ext_m1 <- gam(z_abs.diff ~ s(z_score, by=dataset) + dataset, data=z_long.full)
+summary(gam.ext_m1)
+```
+
+```{r}
+#test for interactive sex effects - split df by dataset and smooth by sex 
+for (dset in c("cf", "cf.gam", "cf.gamlss", "cf.lm")) {
+  print(dset)
+  df <- z_long.full %>%
+    dplyr::filter(dataset == dset)
+
+  m2 <- gam(z_abs.diff ~ s(z_score, by=as.factor(sex)) + as.factor(sex) + s(z_score), data=df)
+  print(summary(m2))
+}
+```

### 2023-12-07 ukb_sim.site_analyses.Rmd
apply welch's nonparametric testing

@@
 source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
+source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
 # 
@@
 ggplot(cent.diffs2) +
-  geom_histogram(aes(x=diff_rh_CT_posteriorcingulate, fill=dataset), alpha = 0.6)
-```
-
-looks like.... no.
+  geom_histogram(aes(x=diff_GMV, fill=dataset), alpha = 0.6) +
+  theme_bw()
+```
+
+looks like....
+no.
+Using welch's t test on centile erro ranks per discussion with Andrew (see Zimmerman & Zumbo, 1993).
+Using custom functions from `ranked_welch_tests.R`.
+also updating Anova to nonparametric kruskal-wallace test.
+**confirm with andrew**
 
@@
 there is literally no difference in accuracy of centile scores between the two)
-
-**check w andrew**
 
@@
 for (pheno in pheno_diff.list) {
-  df.aov <- tidy(aov(cent.diffs2[[pheno]] ~ cent.diffs2[["dataset"]]))
+
+  #group-wise
+  df.aov <- tidy(kruskal.test(cent.diffs2[[pheno]] ~ cent.diffs2[["dataset"]]))
   df.aov$pheno <- pheno
@@
 
-  df.pairwise.t <- tidy(pairwise.wilcox.test(x=cent.diffs2[[pheno]], g=cent.diffs2[["dataset"]], paired = TRUE, p.adj = "none")) #update to wilcox since t assumptions not all met
+  #'post-hoc' pairwise t tests
+  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=cent.diffs2[[pheno]], g=cent.diffs2[["dataset"]], paired = TRUE, p.adj = "none"))
   df.pairwise.t$pheno <- pheno
@@
 
-t tests of sex-biases in centiles w/in each feature **ask andrew about this**
+t tests of sex-biases in centiles w/in each feature
 
@@
         group_by(dataset) %>%
-        summarize(p_value = t.test(.data[[pheno]][sex == "Male"], .data[[pheno]][sex == "Female"])$p.value) %>%
+        summarize(p_value = rank.welch.t.test(.data[[pheno]][sex == "Male"], .data[[pheno]][sex == "Female"])$p.value) %>%
         ungroup()
@@
 for (pheno in z.pheno_abs.diff.list) {
-  df.aov <- tidy(aov(z_diffs.filt[[pheno]] ~ z_diffs.filt[["dataset"]]))
+  df.aov <- tidy(kruskal.test(z_diffs.filt[[pheno]] ~ z_diffs.filt[["dataset"]]))
   df.aov$pheno <- pheno
@@
 
-  df.pairwise.t <- tidy(pairwise.t.test(x=z_diffs.filt[[pheno]], g=z_diffs.filt[["dataset"]], paired = TRUE, p.adj = "none"))
+  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=z_diffs.filt[[pheno]], g=z_diffs.filt[["dataset"]], paired = TRUE, p.adj = "none"))
   df.pairwise.t$pheno <- pheno
@@
 t-test
+
+```{r}
+#check for skew
+ggplot(z_diffs2) +
+  geom_histogram(aes(x=diff_sGMV.z, fill=dataset))
+```
+
 ```{r}

### 2023-12-07 cubic_scripts/R_scripts/ranked_welch_tests.R

@@ -1,6 +1,11 @@
 #load libraries
 library(stats)
 
+#contains 2 functions for running welch's t-test on data rankings (accounting for both skew and unequal variances - Zimmerman & Zumbo, 1993)
+
+# rank.welch.t.test() - analogous to t.test()
+# pairwise.rank.welch.t.test() - analogous to pairwise.t.test()
+
 rank.welch.t.test <- function (x, ...) {
   UseMethod("rank.welch.t.test", x)
 }
@@ -29,7 +34,7 @@ rank.welch.t.test.default <- function(x_list, y_list,
 }
 
 
-## S3 method for class 'formula'
+## S3 method for class 'formula' (based on t.test.formula())
 rank.welch.t.test.formula <- function(formula, data, subset, na.action, ...)
 {
   if(missing(formula)
@@ -43,12 +48,12 @@ rank.welch.t.test.formula <- function(formula, data, subset, na.action, ...)
   m[[1L]] <- quote(stats::model.frame)
   m$... <- NULL
   mf <- eval(m, parent.frame())
-  print(mf)
   DNAME <- paste(names(mf), collapse = " by ")
   names(mf) <- NULL
   response <- attr(attr(mf, "terms"), "response")
-  ## rank
+  # rank values
   mf[[response]] <- rank(mf[[response]])
+  # group
   g <- factor(mf[[-response]])
   if(nlevels(g) != 2L)
     stop("grouping factor must have exactly 2 levels")

### 2023-12-11 ukb_sim.site_analyses.Rmd
adding data dist. plots

@@
 source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
+source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
 # 
@@
   theme_linedraw(base_size=12)
+```
+
+## Plot Centiles
+
+```{r}
+#load data csvs
+ukb_df_cf.gam <- fread(file="data/ukb_to_model/cf.gam_data.csv")
+ukb_df_cf.gamlss <- fread(file="data/ukb_to_model/cf.gamlss_data.csv")
+ukb_df_cf.lm <- fread(file="data/ukb_to_model/cf.lm_data.csv")
+```
+
+plot data distributions (not loading models for now)
+```{r}
+ven_raw <- ggplot(ukb_df_age.filt) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  ggtitle("Raw Data")
+
+ven_lm <- ggplot(ukb_df_cf.lm) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  ggtitle("ComBat")
+
+ven_gam <- ggplot(ukb_df_cf.gam) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  ggtitle("ComBat-GAM")
+
+ven_gamlss <- ggplot(ukb_df_cf.gamlss) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  ggtitle("ComBatLS")
+
+ggarrange(ven_raw, ven_lm, ven_gam, ven_gamlss, ncol=2, nrow=2, common.legend=TRUE)
+```
+
+```{r}
+rh_CT_bankssts_raw <- ggplot(ukb_df_age.filt) +
+  geom_point(aes(x=log_age, y=rh_CT_bankssts, color=sex), alpha =0.3) +
+  ggtitle("Raw Data")
+
+rh_CT_bankssts_lm <- ggplot(ukb_df_cf.lm) +
+  geom_point(aes(x=log_age, y=rh_CT_bankssts, color=sex), alpha =0.3) +
+  ggtitle("ComBat")
+
+rh_CT_bankssts_gam <- ggplot(ukb_df_cf.gam) +
+  geom_point(aes(x=log_age, y=rh_CT_bankssts, color=sex), alpha =0.3) +
+  ggtitle("ComBat-GAM")
+
+rh_CT_bankssts_gamlss <- ggplot(ukb_df_cf.gamlss) +
+  geom_point(aes(x=log_age, y=rh_CT_bankssts, color=sex), alpha =0.3) +
+  ggtitle("ComBatLS")
+
+ggarrange(rh_CT_bankssts_raw, rh_CT_bankssts_lm, rh_CT_bankssts_gam, rh_CT_bankssts_gamlss, ncol=2, nrow=2, common.legend=TRUE)
 ```
@@
   #group-wise
-  df.aov <- tidy(kruskal.test(cent.diffs2[[pheno]] ~ cent.diffs2[["dataset"]]))
+  df.aov <- tidy(kruskal.test(cent.diffs2[[pheno]] ~ cent.diffs2[["dataset"]])) #nonparametric b/c distribution but also data is paired? ask andrews
   df.aov$pheno <- pheno

### 2023-12-11 eda/lifespan_eda.Rmd
adding log-transform for global volume features pre-combat (for lifespan analyses)

--- 2023-12-03 lifespan_eda.Rmd
@@
 
+vol_list_global <- readRDS(file=paste0(base_path, "/cubic_scripts/R_scripts/vol_list_global.rds"))
+
+base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar"
 source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
-source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
+source(paste0(base_path, "/cubic_scripts/R_scripts/gamlss_helper_funs.R"))
 ```
@@
 ```{r}
-cn.df <- fread("data/lifespan_dataset_dkatlas_CNonly.csv", stringsAsFactors = TRUE)
+cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly.csv"), stringsAsFactors = TRUE)
 
@@
 ```{r}
-fwrite(cn.df.site, file="data/lifespan_CN_site-level.csv")
+fwrite(cn.df.site, file=paste0(base_path, "/data/lifespan_CN_site-level.csv"))
+```
+
+looking into whether all global measures should be log-transformed to account for skew (getting neg vals for ventricles with combat-GAM)
+
+```{r}
+global_vols <- cn.df.site %>%
+  dplyr::select(sex, log_age, all_of(vol_list_global)) %>%
+  pivot_longer(cols=c(-sex, -log_age), names_to = "pheno", values_to = "volume") %>%
+  mutate(log_vol = log(volume, base=10)) #test log-transform
+
+ggplot(global_vols) +
+  geom_point(aes(x=log_age, y=volume, color=sex), alpha=0.4) +
+  facet_wrap(~pheno, scales="free_y")
+
+ggplot(global_vols) +
+  geom_histogram(aes(x=volume, fill=pheno), alpha=0.5) +
+  facet_wrap(~pheno, scales="free")
+
+ggplot(global_vols) +
+  geom_histogram(aes(x=log_vol, fill=pheno), alpha=0.5) +
+  facet_wrap(~pheno, scales="free")
 ```
@@
 ```{r}
-fwrite(cn.df.imp, file="data/lifespan_CN_imp-sites.csv")
+fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites.csv"))
 ```
+
+## Visualize Harmonized Data
+
+comparing ventricle vals across harmonization pipelines (neg vals in `lifespan-CN-imp-sites_cf.gam_data.csv`), may need to log-transform
+```{r}
+lifespan_imp_gam.df <- fread(paste0(base_path, "/data/lifespan/lifespan-CN-imp-sites_cf.gam_data.csv"), stringsAsFactors = TRUE)
+lifespan_imp_gamlss.df <- fread(paste0(base_path, "/data/lifespan/lifespan-CN-imp-sites_cf.gamlss_data.csv"), stringsAsFactors = TRUE)
+
+lifespan_gam.df <- fread(paste0(base_path, "/data/lifespan/lifespan-CN-site-level_cf.gam_data.csv"), stringsAsFactors = TRUE)
+lifespan_gamlss.df <- fread(paste0(base_path, "/data/lifespan/lifespan-CN-site-level_cf.gamlss_data.csv"), stringsAsFactors = TRUE)
+```
+
+```{r}
+#mark negative vals
+# cn.df.site <- as.data.frame(cn.df.site)
+neg1 <- cn.df.site %>%
+  dplyr::filter(Ventricles <= 0)
+ven_raw <- ggplot() +
+  geom_point(data=cn.df.site, aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  geom_point(data=neg1, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
+  ggtitle("Raw Data")
+
+neg2 <- lifespan_gam.df[lifespan_gam.df$Ventricles <=0]
+ven_gam <- ggplot() +
+  geom_point(data=lifespan_gam.df, aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  geom_point(data=neg2, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
+  ggtitle("ComBat-GAM")
+
+neg3 <- lifespan_gamlss.df[lifespan_gamlss.df$Ventricles <=0]
+ven_gamlss <- ggplot(lifespan_gamlss.df) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  geom_point(data=neg3, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
+  ggtitle("ComBatLS")
+
+# cn.df.imp <- as.data.frame(cn.df.imp)
+neg4 <- cn.df.imp %>%
+  dplyr::filter(Ventricles <= 0)
+ven_raw.imp <- ggplot(cn.df.imp) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  geom_point(data=neg4, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
+  ggtitle("Raw Data - All Studies")
+
+neg5 <- lifespan_imp_gam.df[lifespan_imp_gam.df$Ventricles <=0]
+ven_gam.imp <- ggplot(lifespan_imp_gam.df) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  geom_point(data=neg5, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
+  ggtitle("ComBat-GAM - All Studies")
+
+neg6 <- lifespan_imp_gamlss.df[lifespan_imp_gamlss.df$Ventricles <=0]
+ven_gamlss.imp <- ggplot(lifespan_imp_gamlss.df) +
+  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
+  geom_point(data=neg6, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
+  ggtitle("ComBatLS - All Studies")
+
+ggarrange(ven_raw, ven_gam, ven_gamlss, ven_raw.imp, ven_gam.imp, ven_gamlss.imp, common.legend = TRUE)
+```

### 2023-12-11 cubic_scripts/R_scripts/combat_apply.R

@@ -20,6 +20,8 @@ library(mgcv)
 library(gamlss)
 library(ComBatFamily)
 
+source("/cubic_scripts/R_scripts/gamlss_helper_funs.R")
+
 ##########################################################################
 
 #GET ARGS
@@ -99,6 +101,12 @@ for (l in list_of_feature_lists){
   pheno.df <- raw.df %>%
     dplyr::select(all_of(l))
   
+  #log-transform global vals - added for lifespan data analyses, not sure if i should rerun for ukb analyses
+  if (names(list_of_feature_lists[i]) = "VolGlob"){
+    pheno.df <- pheno.df %>%
+      mutate(across(c(l), \(x) log(x, base=10)))
+  }
+  
   #make sure batch, covars, and pheno dfs are all the same length
   stopifnot(nrow(pheno.df) == length(batch))
   
@@ -123,6 +131,12 @@ for (l in list_of_feature_lists){
     } else {
       cf.obj <- eval(parse(text = paste("comfam(pheno.df, batch, covar.df,", cf.args, ")")))
     }
+  }
+  
+  #un-log-transform global vals
+  if (names(list_of_feature_lists[i]) = "VolGlob"){
+    cf.obj$dat.combat <- cf.obj$dat.combat %>%
+      mutate(across(c(l), \(x) un_log(x)))
   }
   
   #save cf.obj

### 2023-12-11 cubic_scripts/R_scripts/gamlss_helper_funs.R

@@ -15,6 +15,7 @@ library(broom)
 # get.sigma.df()
 # get.sigma.nl.df()
 # get.moment.formula()
+# un_log()
 ################
 
 drop1_all <- function(mod_obj, list = c("mu", "sigma"), name = NA, dataset = NA){
@@ -441,3 +442,6 @@ get.og.data.centiles <- function(gamlss.rds.file, og.data, get.zscores = FALSE){
     return(df)
   }
 }
+
+### UN-LOG-SCALE - used to un-transform log_age values
+un_log <- function(x){return(10^(x))}

### 2023-12-12 eda/ukb_curation-eda.Rmd
updating to make sure that ppts aren't resampled

--- 2023-10-17 ukb_curation-eda.Rmd
@@
 
+base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/"
+
 source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
-source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
+source(paste0(base_path, "cubic_scripts/R_scripts/gamlss_helper_funs.R"))
 ```
@@
 ```{r}
-cn_data <- fread(file="data/lifespan_dataset_dkatlas_CNonly.csv")
+cn_data <- fread(file= paste0(base_path, "data/lifespan_dataset_dkatlas_CNonly.csv"))
 #table(cn_data$study)
+names(cn_data)
 ```
@@
 length(pheno_list) #total number of models to run
-saveRDS(pheno_list, file="R_scripts/pheno_list.rds")
+saveRDS(pheno_list, file=paste0(base_path, "cubic_scripts/R_scripts/pheno_list.rds"))
 
@@
 vol_list_global <- c("GMV", "sGMV", "WMV", "Ventricles")
-writeLines(vol_list_global, con = "pheno_lists/Vol_list_global.txt")
-saveRDS(vol_list_global, file="R_scripts/vol_list_global.rds")
+writeLines(vol_list_global, con = paste0(base_path,"pheno_lists/Vol_list_global.txt"))
+saveRDS(vol_list_global, file=paste0(base_path,"cubic_scripts/R_scripts/vol_list_global.rds"))
 
 vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
-writeLines(vol_list_regions, con = "pheno_lists/Vol_list_regions.txt")
-saveRDS(vol_list_regions, file="R_scripts/Vol_list_regions.rds")
+writeLines(vol_list_regions, con = paste0(base_path,"pheno_lists/Vol_list_regions.txt"))
+saveRDS(vol_list_regions, file=paste0(base_path, "cubic_scripts/R_scripts/Vol_list_regions.rds"))
 
 sa_list <- col_list[grepl("^lh_SA|^rh_SA", col_list)]
-writeLines(sa_list, con = "pheno_lists/SA_list.txt")
-saveRDS(sa_list, file="R_scripts/SA_list.rds")
+writeLines(sa_list, con = paste0(base_path,"pheno_lists/SA_list.txt"))
+saveRDS(sa_list, file=paste0(base_path,"cubic_scripts/R_scripts/SA_list.rds"))
 
 ct_list <- col_list[grepl("^lh_CT|^rh_CT", col_list)]
-writeLines(ct_list, con = "pheno_lists/CT_list.txt")
-saveRDS(ct_list, file="R_scripts/CT_list.rds")
+writeLines(ct_list, con = paste0(base_path,"pheno_lists/CT_list.txt"))
+saveRDS(ct_list, file=paste0(base_path,"cubic_scripts/R_scripts/CT_list.rds"))
 ```
@@
 #write data
-write_csv(ukb_df_totals, file="data/ukb_CN_data.csv")
+write_csv(ukb_df_totals, file=paste0(base_path,"data/ukb_CN_data.csv"))
 ```
@@
   dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
-# #write data
-# write_csv(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
+#write data
+fwrite(ukb_df_age.filt, file=paste0(base_path,"data/ukb_CN_data_agefilt.csv"))
 ```
@@
 #write out
-write_csv(ukb_df_sites, file="data/ukb_to_model/ukb_CN_data_simsites.csv")
+write_csv(ukb_df_sites, file=paste0(base_path, "data/ukb_to_model/ukb_CN_data_simsites.csv"))
 ```

### 2023-12-12 ukb_sim.site_analyses.Rmd
updating to make sure that ppts aren't resampled

@@
 ukb_df_cf.gamlss <- fread(file="data/ukb_to_model/cf.gamlss_data.csv")
+# length(ukb_df_cf.gamlss$participant) == n_unique(ukb_df_cf.gamlss$participant)
 ukb_df_cf.lm <- fread(file="data/ukb_to_model/cf.lm_data.csv")

### 2023-12-12 cubic_scripts/R_scripts/combat_apply.R

@@ -101,11 +101,11 @@ for (l in list_of_feature_lists){
   pheno.df <- raw.df %>%
     dplyr::select(all_of(l))
   
-  #log-transform global vals - added for lifespan data analyses, not sure if i should rerun for ukb analyses
-  if (names(list_of_feature_lists[i]) = "VolGlob"){
-    pheno.df <- pheno.df %>%
-      mutate(across(c(l), \(x) log(x, base=10)))
-  }
+  # #log-transform global vals - added for lifespan data analyses, not sure if i should rerun for ukb analyses
+  # if (names(list_of_feature_lists[i]) = "VolGlob"){
+  #   pheno.df <- pheno.df %>%
+  #     mutate(across(c(l), \(x) log(x, base=10)))
+  # }
   
   #make sure batch, covars, and pheno dfs are all the same length
   stopifnot(nrow(pheno.df) == length(batch))
@@ -133,11 +133,11 @@ for (l in list_of_feature_lists){
     }
   }
   
-  #un-log-transform global vals
-  if (names(list_of_feature_lists[i]) = "VolGlob"){
-    cf.obj$dat.combat <- cf.obj$dat.combat %>%
-      mutate(across(c(l), \(x) un_log(x)))
-  }
+  # #un-log-transform global vals
+  # if (names(list_of_feature_lists[i]) = "VolGlob"){
+  #   cf.obj$dat.combat <- cf.obj$dat.combat %>%
+  #     mutate(across(c(l), \(x) un_log(x)))
+  # }
   
   #save cf.obj
   saveRDS(cf.obj, file=paste0(save_path, "/combat_objs/",csv_basename,"_", config_name,"_", names(list_of_feature_lists[i]), "_cf_obj.rds"))

### 2023-12-12 eda/lifespan_eda.Rmd

@@
 
-ggarrange(ven_raw, ven_gam, ven_gamlss, ven_raw.imp, ven_gam.imp, ven_gamlss.imp, common.legend = TRUE)
+neg_ventricle_plots <- ggarrange(ven_raw, ven_gam, ven_gamlss, ven_raw.imp, ven_gam.imp, ven_gamlss.imp, common.legend = TRUE)
+#ggsave(paste0(base_path, "figs/lifespan_ventricle_neg.jpg"), plot=neg_ventricle_plots)
 ```

### 2023-12-14 ohbm_results.Rmd

@@
 # Replication 
+Permuting site assignments 3x
 
 ```{r}
-permutation_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_perm_csvs"
+#path to csvs scp'd from cubic
+permutation_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm_csvs"
 ```

### 2023-12-14 paper_figs.Rmd

@@
 #ukb data from ohbm_results.Rmd
-cent_err.p <- readRDS("data/ukb_perm_csvs/centile_errors.RDS")
+cent_err.p <- readRDS("derivatives/ukb_perm_csvs/centile_errors.RDS")
 z_diffs2 <- readRDS("ukb_basic/z_diffs.RDS")

### 2023-12-14 ukb_sim.site_analyses.Rmd

@@
 ```
+
+## Varying M:F
+
+Visualize simulated sites (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/data/ukb_ratios/*.csv' ./data/ukb_ratios/`)
+
+```{r}
+ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_ratios"
+raw_files <- list.files(path = ratio_data_path, pattern = "_raw.csv", full.names = TRUE)
+
+df <- data.frame() #new empty dataframe
+for (file in raw_files) {
+  # Read each CSV file
+    data <- fread(file)
+
+    # Add a "Source_File" column with the file name
+    data <- data %>%
+      dplyr::select(c(sim.site, participant, sex, site, age_days)) %>%
+      mutate(Source_File = as.factor(basename(file))) %>%
+      mutate(dataset = gsub("_data|_raw.csv|prop-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df <- bind_rows(df, data, .id = "File_ID")
+    assign("ratio_sites", df)
+}
+length(unique(ratio_sites$Source_File))
+```
+
+```{r}
+ggplot(ratio_sites) +
+  geom_bar(aes(x=sim.site, fill=sex), position = "fill") +
+  scale_y_continuous(name = "percent",
+                     breaks = seq(0, 1, by=0.1), 
+                     labels = scales::percent(seq(0, 1, by=0.1))) +
+  facet_wrap(~Source_File)
+```
+
+```{r}
+ggplot(ratio_sites) +
+  geom_histogram(aes(x=age_days, fill=sim.site), position = "identity", alpha=0.4, binwidth = 365) +
+  facet_wrap(~Source_File)
+
+ggplot(ratio_sites) +
+  geom_histogram(aes(x=age_days, fill=Source_File, color=Source_File), position = "identity", alpha=0.4, binwidth = 365)
+```
+
+```{r}
+#path to csvs scp'd from cubic
+ratio_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
+```
+
+### Centiles
+Do centile scores change based on combat version?
+
+```{r}
+n_ratio_list <- lapply(c(1:11), str_pad, 2, pad = "0")
+n_ratio_list <- paste0("prop-", n_ratio_list)
+pheno_list.z <- paste0(pheno_list, ".z")
+
+# combined_preds.p <- data.frame()
+get.predictions.ratio <- function(x){
+  df <- data.frame() #new empty dataframe
+  pred.csvs.p <- list.files(path = ratio_path, pattern = paste0(x, ".+_predictions.csv"), full.names = TRUE)
+  for (file in pred.csvs.p) {
+    # Read each CSV file
+    data <- fread(file)
+
+    # Add a "Source_File" column with the file name
+    data <- data %>%
+      mutate(Source_File = as.factor(basename(file)),
+             perm = as.factor(x)) %>%
+      mutate(dataset = gsub("_data|_predictions.csv|prop-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df <- bind_rows(df, data, .id = "File_ID")
+  }
+  return(df)
+}
+ratio_predictions <- mclapply(n_ratio_list, get.predictions.ratio, mc.preschedule = FALSE)
+```

### 2023-12-19 cubic_scripts/qsub_centiles.sh
fixing search so cf.gam and cf.gamlss aren't conflated

@@ -35,8 +35,8 @@ if ! [ -d $save_path ]
 	fi
 #######################################################################################
 #LIST POSSIBLE CONFIGS, INCLUDING NAME OF ORIGINAL RAW DATA
-#be sure to include "data" at the end of the string so refA options aren't confused
-config_list="site-level_cf.gam site-level_cf.gamlss imp-sites_cf.gamlss imp-sites_cf.gam" #cf_data cf.lm_data cf.gam_data cf.gamlss_data #cf.lm_refA_data cf.gam_refA_data cf.gamlss_refA_data ukb_CN_data
+#be sure to include "data" at the end of the string so gam/gamlss and refA options aren't confused
+config_list="site-level_cf.gam_data site-level_cf.gamlss_data imp-sites_cf.gamlss_data imp-sites_cf.gam_data" #cf_data cf.lm_data cf.gam_data cf.gamlss_data #cf.lm_refA_data cf.gam_refA_data cf.gamlss_refA_data ukb_CN_data
 #######################################################################################
 for config in $config_list
 do

### 2023-12-19 paper_results.Rmd

--- 2023-12-14 ohbm_results.Rmd
@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix)
+#Sys.setenv("MC_CORES"=10L)
+options("mc.cores")
 
-source("./R_scripts/gamlss_helper_funs.R")
+source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
+source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
+aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
 ```
 
-single script to parse outputs & get results for OHBM 2024 abstract (condensed from proof-of-concept_mods.Rmd)
+single script to parse outputs & get results (condensed from `ukb_sim.site_analyses.Rmd`)
 
@@
 ### LOAD ###
-pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
-vol_list_global <- readRDS(file="R_scripts/vol_list_global.rds")
-vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
-sa_list <- readRDS(file="R_scripts/SA_list.rds")
-ct_list <- readRDS(file="R_scripts/CT_list.rds")
-
-aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
+vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
+vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
+sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
+ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
 ```
@@
 }
+```
 
+```{r, eval=FALSE}
 #calculate abs error
@@
 ```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs
+cent_abs.diffs <- readRDS("data/centile_abs_diffs.rds")
+```
 
+Test for differences in abs.
+error distributions across combat configurations using nonparametric kruskal-wallace test (instead of anova.
+good that it's nonparametric b/c non-normal distributions BUT also the data is paired!).
+**confirm with andrew**
+Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
+(see Zimmerman & Zumbo, 1993).
+Using custom functions from `ranked_welch_tests.R`.
+
+**need to confirm w andrew then update output table formatting as needed - skipping for now**
+
+```{r}
 #initalize empty dfs to store outputs
@@
 for (pheno in pheno_diff.list) {
-  df.aov <- tidy(aov(cent_abs.diffs[[pheno]] ~ cent_abs.diffs[["dataset"]]))
-  df.aov$pheno <- pheno
-  sum.df <- rbind(sum.df, df.aov)
+  # df.aov <- tidy(kruskal.test(cent_abs.diffs[[pheno]] ~ cent_abs.diffs[["dataset"]]))
+  # df.aov$pheno <- pheno
+  # sum.df <- rbind(sum.df, df.aov)
 
-  df.pairwise.t <- tidy(pairwise.t.test(x=cent_abs.diffs[[pheno]], g=cent_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
+  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=cent_abs.diffs[[pheno]], g=cent_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
   df.pairwise.t$pheno <- pheno
@@
 #apply fdr correction
-sum.df <- sum.df %>%
-  dplyr::filter(term != "Residuals") %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
+# sum.df <- sum.df %>%
+#   dplyr::filter(term != "Residuals") %>%
+#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
 
@@
 
-t.df.sum <- t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min_pval = min(p.val_fdr),
-            max_pval = max(p.val_fdr))
-t.df.sum
-```
-
-Big plot - too big to run locally :(
-
-```{r, eval=FALSE}
-#test big plot
-plot.cents <- cent_abs.diffs %>%
-  #dplyr::filter(dataset == "cf.gamlss") %>%
-  pivot_longer(starts_with("abs.diff_"), names_to="pheno", values_to="abs.diff", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_))) %>%
-  ggplot() +
-  geom_density(aes(x=diff, fill=dataset), alpha=0.4) +
-  facet_wrap(~pheno_cat) +
-  theme(legend.position="none")
-ggsave("figs/cent_plot_test.jpeg", var.plot, width=10, height=10)
+# t.df.sum <- t.df %>%
+#   group_by(comp) %>%
+#   summarize(n_sig_regions = sum(sig_fdr),
+#             n_compare = n(),
+#             min_pval = min(p.val_fdr),
+#             max_pval = max(p.val_fdr))
+# t.df.sum
 ```
@@
 
+look for skew in distribution
+
 ```{r}
@@
 #initalize empty dfs to store outputs
-z.sum.df <- data.frame(pheno = character(),
-                       term = character(),
-                       df = double(),
-                       sumsq = double(),
-                       meansq = double(),
-                       statistic = double(),
-                       p.value = double())
+# z.sum.df <- data.frame(pheno = character(),
+#                        term = character(),
+#                        df = double(),
+#                        sumsq = double(),
+#                        meansq = double(),
+#                        statistic = double(),
+#                        p.value = double())
 z.t.df <- data.frame("pheno" = character(),
@@
 for (pheno in z.pheno_abs.diff.list) {
-  df.aov <- tidy(aov(z_abs.diffs[[pheno]] ~ z_abs.diffs[["dataset"]]))
-  df.aov$pheno <- pheno
-  z.sum.df <- rbind(z.sum.df, df.aov)
+  # df.aov <- tidy(kruskal.test(z_abs.diffs[[pheno]] ~ z_abs.diffs[["dataset"]]))
+  # df.aov$pheno <- pheno
+  # z.sum.df <- rbind(z.sum.df, df.aov)
 
-  df.pairwise.t <- tidy(pairwise.t.test(x=z_abs.diffs[[pheno]], g=z_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
+  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=z_abs.diffs[[pheno]], g=z_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
   df.pairwise.t$pheno <- pheno
@@
 #apply fdr correction
-z.sum.df <- z.sum.df %>%
-  dplyr::filter(term != "Residuals") %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
+# z.sum.df <- z.sum.df %>%
+#   dplyr::filter(term != "Residuals") %>%
+#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
 
@@
 
-sum_z.t.df <- z.t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min.pval = min(p.val_fdr),
-            max.pval = max(p.val_fdr))
-
-sum_z.t.df
+# sum_z.t.df <- z.t.df %>%
+#   group_by(comp) %>%
+#   summarize(n_sig_regions = sum(sig_fdr),
+#             n_compare = n(),
+#             min.pval = min(p.val_fdr),
+#             max.pval = max(p.val_fdr))
+# 
+# sum_z.t.df
 ```
@@
 ```
+
 # Replication 
@@
 
-perm_diffs <- mclapply(perm_predictions, get.abs.diffs, mc.preschedule = FALSE)
 ```
@@
 ```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs 
+perm_diffs <- mclapply(perm_predictions, get.abs.diffs, mc.preschedule = FALSE)
+saveRDS(perm_diffs)
+```
+
+```{r}
+perm_diffs <- readRDS()
+```
+
+add ANOVA
+```{r}
+
+```
+
+Pairwise T tests
+
+```{r}
+#define function
 pheno_diff.list <- paste0("abs.diff_", pheno_list)
@@
   for (pheno in pheno_diff.list) {
-    df.pairwise.t <- tidy(pairwise.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
+    df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
     df.pairwise.t$pheno <- pheno
@@
 }
+```
+
+```{r}
+#apply
 perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
@@
 cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-fwrite(cent_t.tests.p, file="ukb_perm_csvs/centile_t.tests.csv")
+fwrite(cent_t.tests.p, file="derivatives/ukb_perm_csvs/centile_t.tests.csv")
 ```
@@
 ```{r}
+#define plotting functions
 plot.cent.diffs <- function(df){
@@
 }
+```
+
+```{r}
 lapply(perm_diffs, plot.cent.diffs)
@@
 
+*need to look at distribution to check for skew!*
+
 ```{r}
@@
   for (pheno in pheno_diff.list.z) {
-    tryCatch(df.pairwise.t <- tidy(pairwise.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
+    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
+    df.pairwise.t$pheno <- pheno
+    t.df <- rbind(t.df, df.pairwise.t)
+  }
+  n_comp <- nrow(t.df)
+
+  #apply fdr correction
+  t.df <- t.df %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE),
+           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
+    unite(comp, c("group1", "group2")) %>%
+   dplyr::mutate(comp = as.factor(comp))
+
+  t.df.sum <- t.df %>%
+    group_by(comp) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+  return(t.df.sum)
+  #detach()
+}
+
+perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
+names(perm_t.tests_z) <- n_perm_list
+z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
+```
+
+```{r}
+fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
+```
+
+```{r}
+z_t.tests.p %>%
+  group_by(comp) %>%
+  dplyr::summarize(n_sig_min=min(n_sig_regions),
+            n_sig_max = max(n_sig_regions),
+            n_sig_mean = mean(n_sig_regions))
+```
+
+## SCALE EFFECTS OF SEX
+fix age, get diff between male and female pred. variance
+```{r}
+# combined_preds.p <- data.frame()
+get.var <- function(x){
+  df <- data.frame() #new empty dataframe
+  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
+  for (file in var.csvs.p) {
+    # Read each CSV file
+    data <- fread(file)
+
+    # Add a "Source_File" column with the file name
+    data <- data %>%
+      mutate(Source_File = as.factor(basename(file)),
+             perm = as.factor(x)) %>%
+      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
+             sex_standard = sex_effect/f.var,
+             pheno = as.character(pheno))
+
+    # Bind the data to the combined dataframe
+    df <- bind_rows(df, data, .id = "File_ID")
+  }
+  return(df)
+}
+
+perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
+perm_var.all <- bind_rows(perm_var, .id = "column_label")
+```
+
+maybe better to get sex_summary.csv - just need signifigance
+
+```{r}
+#get summary df for p vals, plotting
+sigma.sex.df <- fread("derivatives/ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
+  dplyr::select(-V1) %>%
+  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
+         dataset = gsub("-data", "", dataset),
+         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
+         dataset = gsub("-", "", dataset),
+         #dataset = as.character(dataset),
+         pheno = as.character(pheno)) %>%
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
+         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
+  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
+  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
+  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
+
+perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
+```
+
+```{r}
+  sex_effect_plot <- perm_var_info %>%
+    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
+                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
+    ggplot() +
+    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
+    facet_wrap(~perm, ncol=5) +
+    ggtitle(paste("Standardized effect of male sex on variance")) +
+    theme(axis.text.x = element_blank())
+
+    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
+    print(sex_effect_plot)
+```
+
+# Secondary: Vary M:F
+
+```{r}
+ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
+raw_files <- list.files(path = ratio_data_path, pattern = "_pred.csv", full.names = TRUE)
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  # Read each CSV file
+    data <- fread(file)
+
+    # Add a "Source_File" column with the file name
+    data <- data %>%
+      mutate(Source_File = as.factor(basename(file))) %>%
+      mutate(dataset = gsub("_data|_raw.csv|prop-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df_list <- c(df_list, list(data))
+    assign("ratio_sites_list", df_list)
+}
+length(ratio_sites_list)
+```
+
+## CENTILES
+Do centile scores change based on combat version?
+
+need to look and see what happens for subjects who aren't in all samples!
+
+Little plots
+```{r}
+lapply(ratio_sites_list, plot.cent.diffs)
+```
+
+Plot boxplots to compare across prop.
+```{r}
+ratio_diffs.df <- bind_rows(ratio_diffs) %>% 
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(perm, dataset) %>%
+  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup() %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+ratio_diffs.df %>%
+  mutate(centile_abs.diff = centile_abs.diff * 100) %>%
+  ggplot() +
+    geom_boxplot(aes(x=perm, y=centile_abs.diff, fill=dataset)) +
+  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
+  ggtitle("Within-feature mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
+```
+
+get w/in subj mean magnitude centile error
+
+```{r}
+ratio_subj_diffs.df <- bind_rows(ratio_diffs) %>% 
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(perm, dataset) %>%
+  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup() %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+ratio_diffs.df %>%
+  mutate(centile_abs.diff = centile_abs.diff * 100) %>%
+  ggplot() +
+    geom_boxplot(aes(x=perm, y=centile_abs.diff, fill=dataset)) +
+  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
+  ggtitle("Within-feature mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
+```
+
+W/in subj average plots
+```{r, eval=FALSE}
+mean.subj.centile.err <- function(df){
+mean.diffs.subj <- df %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+  rowwise() %>%
+  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))
+
+return(mean.diffs.subj)
+#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
+}
+ratio_diffs.err <- mclapply(ratio_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
+names(ratio_diffs.err) <- levels(ratio_diffs.df$perm)
+subj.cent_err.r <- bind_rows(ratio_diffs.err, .id = "proportion")
+
+#write out for plotting later
+saveRDS(subj.cent_err.r, "derivatives/ukb_ratio/subj_centile_errors.RDS")
+```
+
+```{r}
+subj.cent_err.r <- readRDS("derivatives/ukb_ratio/subj_centile_errors.RDS")
+```
+
+```{r}
+cent.diffs.v.plot.r <- subj.cent_err.r %>%
+  mutate(mean = mean*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=proportion, y=mean, fill=dataset)) +
+  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
+ggsave(filename="figs/ratio_centile_violin.png", cent.diffs.v.plot.r)
+```
+
+```{r}
+cent.diffs.v.boxplot.r <- subj.cent_err.r %>%
+  mutate(mean = mean*100) %>% #for visuals
+  ggplot() +
+  geom_boxplot(aes(x=proportion, y=mean, fill=dataset)) +
+  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
+ggsave(filename="figs/ratio_centile_box.png", cent.diffs.v.plot.r)
+```
+
+### Sex-Biases
+
+```{r}
+cent.diffs.v.boxplot.r_sex_free <- subj.cent_err.r %>%
+  mutate(mean = mean*100) %>% #for visuals
+  ggplot() +
+  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
+  facet_wrap(~dataset, scale="free_y") +
+  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
+ggsave(filename="figs/ratio_centile_box_sex_free.png", cent.diffs.v.boxplot.r_sex_free)
+```
+
+```{r}
+#trying to add significance testing
+cent.diffs.by.sex.stat.test <- subj.cent_err.r %>%
+  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
+  group_by(dataset, proportion) %>%
+  wilcox_test(mean ~ sex) %>%
+  adjust_pvalue(method = "fdr") %>%
+  add_significance() %>%
+  add_xy_position(x = "proportion")
+
+cent.diffs.v.boxplot.r_sex <- subj.cent_err.r %>%
+  mutate(mean = mean*100) %>% #for visuals
+  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
+  droplevels() %>%
+  ggplot() +
+  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
+  facet_wrap(~dataset) +
+  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  stat_pvalue_manual(
+    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
+    #hide.ns = TRUE
+    bracket.nudge.y = -0.5
+    ) +
+  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
+cent.diffs.v.boxplot.r_sex
+ggsave(filename="figs/ratio_abs.centile_box_sex.png", cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
+```
+
+get diffs (not abs)
+
+```{r}
+get.diffs <- function(x){
+  df <- x %>%
+    mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+    mutate(dataset = relevel(dataset, ref= "raw")) %>%
+    arrange(dataset) %>%
+    group_by(participant) %>%
+    dplyr::mutate(across(all_of(pheno_list), ~ (. - first(.)), .names = "diff_{.col}")) %>% #centile diffs
+    dplyr::select(-any_of(pheno_list)) %>% #drop
+    dplyr::mutate(across(ends_with(".z"), ~ (. - first(.)), .names = "diff_{.col}")) %>% #z-score diffs
+    dplyr::select(-any_of(pheno_list.z)) %>% #drop
+    ungroup() %>%
+    dplyr::filter(dataset != "raw") #drop raw
+  return(df)
+}
+```
+
+```{r, eval=FALSE}
+#get diff in centiles and z scores across ratios
+ratio_diffs2 <- mclapply(ratio_predictions, get.diffs, mc.preschedule = FALSE)
+length(ratio_diffs2)
+```
+
+take w/in subj mean
+```{r, eval=FALSE}
+mean.subj.centile.err2 <- function(df){
+mean.diffs.subj <- df %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+  rowwise() %>%
+  dplyr::mutate(mean = mean(c_across(starts_with("diff"))))
+
+return(mean.diffs.subj)
+#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
+}
+ratio_diffs.err2 <- mclapply(ratio_diffs2, mean.subj.centile.err2, mc.preschedule = FALSE)
+names(ratio_diffs.err2) <- levels(ratio_diffs.df$perm)
+subj.cent_err.r2 <- bind_rows(ratio_diffs.err2, .id = "proportion")
+saveRDS(subj.cent_err.r2, file="derivatives/ukb_ratio/ratio_centile_diffs.rds")
+```
+
+```{r}
+#load
+subj.cent_err.r2 <- readRDS("derivatives/ukb_ratio/ratio_centile_diffs.rds")
+```
+
+```{r}
+cent.diffs.v.boxplot.r_sex2 <- subj.cent_err.r2 %>%
+  mutate(mean = mean*100) %>% #for visuals
+  ggplot() +
+  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
+  facet_wrap(~dataset) +
+  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
+ggsave(filename="figs/ratio_centile_box_sex2.png", cent.diffs.v.boxplot.r_sex)
+```
+
+Test for sex-effects in subj's mean centile errors
+
+```{r}
+cent.diffs.by.sex.stat.test <- subj.cent_err.r2 %>%
+  group_by(dataset, proportion) %>%
+  wilcox_test(mean ~ sex) %>%
+  adjust_pvalue(method = "fdr") %>%
+  add_significance() %>%
+  add_xy_position(x = "proportion", dodge = 0.8)
+
+cent.diffs.v.boxplot.r_sex_free2 <- subj.cent_err.r2 %>%
+  mutate(mean = mean*100) %>% #for visuals
+  ggplot() +
+  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
+  facet_wrap(~dataset, scale="free_y") +
+  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  stat_pvalue_manual(
+    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0.01,
+    #hide.ns = TRUE
+    ) +  
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations while varying M:F")
+cent.diffs.v.boxplot.r_sex_free2
+ggsave(filename="figs/ratio_centile_box_sex_freescale2.png", cent.diffs.v.boxplot.r_sex_free2, width = 15, height=7)
+```
+
+### Extremeness-Biases
+Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)
+
+test w/in subj avg centile ~ w/in subj avg centile error
+
+```{r}
+#mean w/in subj error
+subj.cent_err.only <- subj.cent_err.r2 %>%
+  dplyr::select(!starts_with("diff")) %>%
+  rename(mean_diff = mean)
+```
+
+```{r, eval=FALSE}
+#calc mean centile
+subj.cent.mean <- function(df){
+mean.subj <- df %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+  rowwise() %>%
+  dplyr::mutate(mean_cent = mean(c_across(pheno_list))) %>%
+  dplyr::select(!pheno_list)
+
+return(mean.subj)
+#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
+}
+
+ratio_cent_means <- mclapply(ratio_predictions, subj.cent.mean, mc.preschedule = FALSE, mc.cores=12L)
+#bind predictions
+names(ratio_cent_means) <- n_ratio_list
+ratio_cent_means.df <- bind_rows(ratio_cent_means, .id = "column_label")
+saveRDS(ratio_cent_means.df, file="derivatives/ukb_ratios/ratio_cent_means.csv")
+```
+
+## Z-SCORES
+
+```{r}
+#within each phenotype, iterate paired t test between each combo of combat dfs 
+pheno_diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
+
+z.t.tests <- function(df){
+  #initialize empty dfs to store outputs
+  t.df <- data.frame("pheno" = character(),
+                     "group1" = character(),
+                     "group2" = character(),
+                     "p.value" = double())
+  #run tests
+  attach(df)
+  for (pheno in pheno_diff.list.z) {
+    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
     df.pairwise.t$pheno <- pheno
@@
 ```
+
+# Secondary: LBCC Data

### 2023-12-20 paper_results.Rmd
reworking and updating so that more heavy computation (ie centile error calcs) happen on cubic

@@
 ```{r}
-#define plotting functions
+#define plotting
 plot.cent.diffs <- function(df){

### 2023-12-20 ukb_sim.site_analyses.Rmd
reworking and updating so that more heavy computation (ie centile error calcs) happen on cubic

--- 2023-12-12 ukb_sim.site_analyses.Rmd
@@
 ```
+
+## Varying M:F
+
+pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_diffs.csv' .`
+
+pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' .`
+
+Visualize simulated sites
+
+```{r}
+ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
+raw_files <- list.files(path = ratio_data_path, pattern = "_pred.csv", full.names = TRUE)
+
+df <- data.frame() #new empty dataframe
+for (file in raw_files) {
+  # Read each CSV file
+    data <- fread(file)
+
+    # Add a "Source_File" column with the file name
+    data <- data %>%
+      mutate(Source_File = as.factor(basename(file))) %>%
+      mutate(dataset = gsub("_data|_raw.csv|prop-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df <- bind_rows(df, data)
+    assign("ratio_sites", df)
+}
+length(unique(ratio_sites$Source_File))
+```
+
+```{r}
+ggplot(ratio_sites) +
+  geom_bar(aes(x=sim.site, fill=sex), position = "fill") +
+  scale_y_continuous(name = "percent",
+                     breaks = seq(0, 1, by=0.1), 
+                     labels = scales::percent(seq(0, 1, by=0.1))) +
+  facet_wrap(~Source_File)
+```
+
+```{r}
+ggplot(ratio_sites) +
+  geom_histogram(aes(x=age_days, fill=sim.site), position = "identity", alpha=0.4, binwidth = 365) +
+  facet_wrap(~Source_File)
+
+ggplot(ratio_sites) +
+  geom_histogram(aes(x=age_days, fill=Source_File, color=Source_File), position = "identity", alpha=0.4, binwidth = 365)
+```
+
+### Centiles

### 2023-12-20 paper_results.Rmd
fixed ratio file read-in

@@
 ---
-title: "ohbm_results"
+title: "hbm_paper_results"
 output: html_document
@@
 
+## Load Data
+
+Read in centile/z-score errors
+
 ```{r}
 ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
-raw_files <- list.files(path = ratio_data_path, pattern = "_pred.csv", full.names = TRUE)
+raw_files <- list.files(path = ratio_data_path, pattern = "_diffs.csv", full.names = TRUE)
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  print(paste("reading file", file))
+  # Read each CSV file
+    data <- fread(file)
+
+    data <- data %>%
+     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df_list[[file]] <- as.data.frame(data)
+    assign("ratio_list", df_list)
+}
+length(ratio_list)
+```
+
+Read in subject-wise summary stats
+
+```{r}
+raw_files <- list.files(path = ratio_data_path, pattern = "_subj_pred.csv", full.names = TRUE)
 
@@
 
-    # Add a "Source_File" column with the file name
     data <- data %>%
-      mutate(Source_File = as.factor(basename(file))) %>%
-      mutate(dataset = gsub("_data|_raw.csv|prop-|[0-9]|-", "", Source_File))
+      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
 
@@
     df_list <- c(df_list, list(data))
-    assign("ratio_sites_list", df_list)
-}
-length(ratio_sites_list)
+    assign("ratio_subj_list", df_list)
+}
+length(ratio_subj_list)
+ratio_subj.df <- bind_rows(ratio_subj_list)
 ```
@@
 ```{r}
-ratio_diffs.df <- bind_rows(ratio_diffs) %>% 
-  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(perm, dataset) %>%
-  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-ratio_diffs.df %>%
-  mutate(centile_abs.diff = centile_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=perm, y=centile_abs.diff, fill=dataset)) +
+# ratio_diffs.df <- bind_rows(ratio_diffs) %>% 
+#   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+#   group_by(perm, dataset) %>%
+#   dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
+#   ungroup() %>%
+#   dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+#   pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+#   mutate(pheno_cat = as.factor(case_when(
+#       pheno %in% vol_list_global ~ "Global Volume",
+#       pheno %in% vol_list_regions ~ "Regional Volume",
+#       pheno %in% sa_list ~ "Regional SA",
+#       pheno %in% ct_list ~ "Regional CT",
+#       TRUE ~ NA_character_)),
+#       label = sub("_[^_]*_", "_", pheno))
+
+ratio_subj.df %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff * 100) %>%
+  ggplot() +
+    geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=dataset)) +
   labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +

### 2023-12-27 eda/ukb_curation-eda.Rmd
qc-ing features with centile score of 1

@@
 ```{r}
+ukb_df.ppt.qc <- ukb_df %>%
+  dplyr::filter(participant %in% ppt.list)
+
+ggplot(data=ukb_df, aes(x=log_age, y=GMV)) +
+  geom_point(aes(color=sex), alpha=0.6) +
+  geom_text(data=ukb_df.ppt.qc,
+            aes(x=log_age, y=GMV, label=participant))
+
+ggplot(data=ukb_df, aes(x=log_age, y=lh_CT_cuneus)) +
+  geom_point(aes(color=sex), alpha=0.6) +
+  geom_text(data=ukb_df.ppt.qc,
+            aes(x=log_age, y=lh_CT_cuneus, label=participant)) +
+  geom_hline(yintercept=median(ukb_df$lh_CT_cuneus) + 7*stats::mad(ukb_df$lh_CT_cuneus))
+
+ggplot(data=ukb_df, aes(x=log_age, y=lh_CT_pericalcarine)) +
+  geom_point(aes(color=sex), alpha=0.6) +
+  geom_text(data=ukb_df.ppt.qc,
+            aes(x=log_age, y=lh_CT_pericalcarine, label=participant)) +
+  geom_hline(yintercept=median(ukb_df$lh_CT_pericalcarine) + 7*stats::mad(ukb_df$lh_CT_pericalcarine))
+
+ggplot(data=ukb_df, aes(x=log_age, y=lh_Vol_parahippocampal)) +
+  geom_point(aes(color=sex), alpha=0.6) +
+  geom_text(data=ukb_df.ppt.qc,
+            aes(x=log_age, y=lh_Vol_parahippocampal, label=participant)) +
+    geom_hline(yintercept=median(ukb_df$lh_Vol_parahippocampal) + 7*stats::mad(ukb_df$lh_Vol_parahippocampal))
+
+ggplot(data=ukb_df, aes(x=log_age, y=rh_CT_medialorbitofrontal)) +
+  geom_point(aes(color=sex), alpha=0.6) +
+  geom_text(data=ukb_df.ppt.qc,
+            aes(x=log_age, y=rh_CT_medialorbitofrontal, label=participant)) +
+  geom_hline(yintercept=median(ukb_df$rh_CT_medialorbitofrontal) + 7*stats::mad(ukb_df$rh_CT_medialorbitofrontal))
+```
+
+```{r}
 ggplot(data=ukb_df, aes(x=age_days)) +
@@
 
-ggplot(data=ukb_df, aes(x=log_age, y=GMV)) +
-  geom_point(aes(color=sex), alpha=0.6)
 ggplot(data=ukb_df, aes(x=age_days, y=GMV)) +
@@
 ```
+
+looking into subj. with centiles = 1
 

### 2023-12-27 paper_results.Rmd

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions)
 #Sys.setenv("MC_CORES"=10L)
@@
 
-load info
 ```{r}
@@
 ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
-```
-
+
+pheno_list.z <- paste0(pheno_list, ".z")
+```
+
+# Define Functions
+
+Pairwise T tests: within each feature, run paired t tests (welch's test on ranks) pairwise on each possible pairing of combat configs and return as a single dataframe, with FDR correction for all feature + cf pair combos.
+Currently runs on magnitude (abs val) of centile diffs, but can be run on other metrics by redefining `feature_list` Recommended for use with mclapply across list of dataframes.
+Optional argument to correct for more comparisons (e.g.
+fdr-correct across an entire list of dataframes) using `comp_multiplier` arg.
+
+```{r}
+#define function
+pheno_diff.list <- paste0("abs.diff_", pheno_list)
+
+centile.t.tests <- function(df, feature_list=pheno_diff.list, comp_multiplier=1){
+  #initialize empty dfs to store outputs
+  t.df <- data.frame("pheno" = character(),
+                     "group1" = character(),
+                     "group2" = character(),
+                     "p.value" = double())
+  #run tests
+  attach(df)
+  for (pheno in feature_list) {
+    df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
+    df.pairwise.t$pheno <- pheno
+    t.df <- rbind(t.df, df.pairwise.t)
+  }
+  n_comp <- nrow(t.df)
+
+  #apply fdr correction
+  t.df <- t.df %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE),
+           pheno = sub("abs.diff_", "", pheno)) %>%
+    unite(comp, c("group1", "group2")) %>%
+   dplyr::mutate(comp = as.factor(comp))
+
+  t.df.sum <- t.df %>%
+    group_by(comp) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+  return(t.df.sum)
+  #detach()
+}
+
+### EXAMPLE:
+# perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
+# names(perm_t.tests) <- n_perm_list
+# cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
+```
+
+Within each combat config, test if there are significant differences in M and F subject's mean abs.
+centile error.
+Can test mean (not abs) centile errors, Z errors, etc, using `to_test` arg.
+FDR corrects for number of
+
+```{r}
+sex.bias.t.tests <- function(df, to_test = "mean_cent_abs.diff", comp_multiplier=1){
+
+  #conduct t test
+  df.sex.t <- df %>%
+    group_by(dataset) %>%
+    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
+              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
+              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
+              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
+    ungroup()
+
+  #apply fdr correction
+  n_comp=nrow(df)
+  df.sex.t <- df.sex.t %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE))
+  return(df.sex.t)
+  #detach()
+}
+
+# tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
+# rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
+```
+
+```{r}
+extremeness.lm <- function(df, f, comp_multiplier=1){
+
+  #conduct t test
+  df.lm <- df %>%
+    group_by(dataset) %>%
+    #returns values for last (highest order) predictor
+    summarise(p.value = tail(tidy(lm(as.formula(f)))$p.value, n=1),
+              estimate = tail(tidy(lm(as.formula(f)))$estimate, n=1),
+              t.stat = tail(tidy(lm(as.formula(f)))$statistic[2], n=1),
+              std.error = tail(tidy(lm(as.formula(f)))$std.error[2]), n=1) %>%
+    ungroup()
+  n_comp=nrow(df)
+
+  #apply fdr correction
+  df.lm <- df.lm %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE))
+  return(df.lm)
+  #detach()
+}
+
+# tidy(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
+# summary(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
+```
 # Main Analyses
@@
   # Bind the data to the combined dataframe
-  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
-}
-```
+  combined_preds <- bind_rows(combined_preds, data)
+}
+```
+
+**qc - find ppts with centiles at 1 or 0**
+
+```{r}
+qc_cent_df <- combined_preds %>% 
+  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0, Inf, -Inf, NA))) %>%
+  dplyr::select(File_ID, participant, sex, dataset, pheno_list) %>%
+  pivot_longer(cols=pheno_list, names_to="pheno", values_to="centile")
+skimr::skim(qc_cent_df)
+
+qc_cent_df %>%
+  dplyr::filter(centile == 0)
+
+qc_cent1_df <- qc_cent_df %>%
+  dplyr::filter(centile == 1)
+
+qc_cent1_df %>%
+  ggplot() +
+  geom_bar(aes(dataset, fill=participant)) +
+  facet_wrap(~pheno)
+
+ppt.list <- qc_cent1_df %>%
+  dplyr::select(participant) %>%
+  unique() %>%
+  as.list() %>%
+  unname() %>%
+  unlist()
+
+qc_z_df <- combined_preds %>% 
+  filter_at(vars(pheno_list.z), any_vars(. %in% c(1, 0, Inf, -Inf, NA)))
+
+unique(qc_z_df$participant)
+ppt.list
+```
+
+*MOVE TO CUBIC*
 
@@
 ```{r}
+#get means w/in feature 
 mean.diffs.long <- cent_abs.diffs %>%
@@
   labs(y="Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
+  ggtitle("Figure 2. Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
 cent.diffs.plot
 
-#ggsave("figs/mean_centile_diffs.jpeg", cent.diffs.plot)
+ggsave("figs/win_feature_mean_cent_abs.diffs.jpeg", cent.diffs.plot)
 ```
@@
 
-```{r}
+```{r, eval=FALSE}
+#calc errors
 z_abs.diffs <- combined_preds %>%
@@
   dplyr::filter(dataset != "ukb_CN") #drop raw
-
+saveRDS(z_abs.diffs, "data/z_abs.diffs.rds")
+```
+
+```{r}
+z_abs.diffs <- readRDS("data/z_abs.diffs.rds")
+```
+
+significance testing
+
+```{r}
 #initalize empty dfs to store outputs
@@
 
+Little plot
+```{r}
+#get means w/in feature 
+z.mean.diffs.long <- z_abs.diffs %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno = gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+```
+
+```{r}
+z.diffs.plot <- z.mean.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  ggplot() +
+  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
+  theme_linedraw(base_size = 11) +
+  facet_wrap(~pheno_cat) +
+  scale_y_continuous(labels = scales::percent) + 
+  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
+  ggtitle("Figure 2. Absolute errors in z-scores of brain structure features across ComBat processing pipelines")
+z.diffs.plot
+
+ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", z.diffs.plot)
+```
+
 ## SCALE EFFECTS OF SEX
@@
 # Replication
-Permuting site assignments 3x
-
-```{r}
-#path to csvs scp'd from cubic
-permutation_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm_csvs"
+
+## Load Data
+
+pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*_diffs.csv' .`
+
+pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*.csv' .`
+
+Read in centile/z-score errors
+
+```{r}
+perm_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm_csvs"
+raw_files <- list.files(path = perm_data_path, pattern = "_diffs.csv", full.names = TRUE)
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  print(paste("reading file", file))
+  # Read each CSV file
+    data <- fread(file)
+
+    data <- data %>%
+     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df_list[[file]] <- as.data.frame(data)
+    assign("perm_list", df_list)
+}
+length(perm_list)
+perm.df <- bind_rows(perm_list)
+```
+
+Read in subject-wise summary stats
+
+```{r}
+raw_files <- list.files(path = perm_data_path, pattern = "_subj_pred.csv", full.names = TRUE)
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  # Read each CSV file
+    data <- fread(file)
+
+    data <- data %>%
+      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df_list <- c(df_list, list(data))
+    assign("perm_subj_list", df_list)
+}
+length(perm_subj_list)
+perm_subj.df <- bind_rows(perm_subj_list)
 ```
@@
 
-```
-
-Pairwise T tests
-
-```{r}
-#define function
-pheno_diff.list <- paste0("abs.diff_", pheno_list)
-
-centile.t.tests <- function(df){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in pheno_diff.list) {
-    df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = sub("abs.diff_", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
 ```
@@
 
+```{r}
+n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
+```
+
 ## Load Data
+
+pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_diffs.csv' .`
+
+pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' .`
 
@@
 length(ratio_list)
+ratio.df <- bind_rows(ratio_list)
+
+# #sum(is.na(ratio.df))
+# na.df <- ratio.df[rowSums(is.na(ratio.df)) > 0,]
+# skimr::skim(na.df)
+# 
+# names(ratio.df)
+ratio.df[is.na(ratio.df$diff_lh_CT_cuneus.z),] %>%
+  dplyr::select(contains("_cuneus"), "participant", "dataset")
 ```
@@
 ratio_subj.df <- bind_rows(ratio_subj_list)
-```
-
-## CENTILES
+
+max(ratio_subj.df$mean_centile)
+min(ratio_subj.df$mean_centile)
+max(ratio_subj.df$mean_z)
+min(ratio_subj.df$mean_z)
+```
+
 Do centile scores change based on combat version?
 
-need to look and see what happens for subjects who aren't in all samples!
-
-Little plots
-```{r}
-lapply(ratio_sites_list, plot.cent.diffs)
-```
+need to look and see what happens for subjects who aren't in all samples?
 
 Plot boxplots to compare across prop.
-```{r}
-# ratio_diffs.df <- bind_rows(ratio_diffs) %>% 
-#   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-#   group_by(perm, dataset) %>%
-#   dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-#   ungroup() %>%
-#   dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-#   pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-#   mutate(pheno_cat = as.factor(case_when(
-#       pheno %in% vol_list_global ~ "Global Volume",
-#       pheno %in% vol_list_regions ~ "Regional Volume",
-#       pheno %in% sa_list ~ "Regional SA",
-#       pheno %in% ct_list ~ "Regional CT",
-#       TRUE ~ NA_character_)),
-#       label = sub("_[^_]*_", "_", pheno))
-
+
+```{r, eval=FALSE}
+#too much memory to run!
+ratio.df %>%
+  dplyr::select(sim.site, prop, dataset, starts_with("abs.diff"), !ends_with(".z")) %>%
+  pivot_longer(cols= starts_with("abs.diff"), names_to="pheno", nammes_prefix="abs.diff_", values_to="cent_abs.diff") %>%
+  mutate(cent_abs.diff = cent_abs.diff * 100) %>%
+  ggplot() +
+    geom_boxplot(aes(x=prop, y=cent_abs.diff, fill=dataset)) +
+  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
+  ggtitle("Error in centile scores across ComBat \nconfigurations as sex imbalance changes")
+```
+
+Plot w/in subj means
+
+```{r}
 ratio_subj.df %>%
@@
   labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-feature mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-get w/in subj mean magnitude centile error
-
-```{r}
-ratio_subj_diffs.df <- bind_rows(ratio_diffs) %>% 
-  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(perm, dataset) %>%
-  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-ratio_diffs.df %>%
-  mutate(centile_abs.diff = centile_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=perm, y=centile_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-feature mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
+
+ratio_subj.df %>%
+  ggplot() +
+    geom_boxplot(aes(x=prop, y=mean_z_abs.diff, fill=dataset)) +
+  labs(y= "Magnitude Z-score Error", x = "M:F Ratio in Site B") +
+  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat \nconfigurations as sex imbalance changes")
 ```
@@
 W/in subj average plots
-```{r, eval=FALSE}
-mean.subj.centile.err <- function(df){
-mean.diffs.subj <- df %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  rowwise() %>%
-  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))
-
-return(mean.diffs.subj)
-#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
-}
-ratio_diffs.err <- mclapply(ratio_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
-names(ratio_diffs.err) <- levels(ratio_diffs.df$perm)
-subj.cent_err.r <- bind_rows(ratio_diffs.err, .id = "proportion")
-
-#write out for plotting later
-saveRDS(subj.cent_err.r, "derivatives/ukb_ratio/subj_centile_errors.RDS")
-```
-
-```{r}
-subj.cent_err.r <- readRDS("derivatives/ukb_ratio/subj_centile_errors.RDS")
-```
-
-```{r}
-cent.diffs.v.plot.r <- subj.cent_err.r %>%
-  mutate(mean = mean*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=proportion, y=mean, fill=dataset)) +
+
+```{r}
+cent.diffs.v.plot.r <- ratio_subj.df %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=dataset), scale="width") +
   labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
   ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
+cent.diffs.v.plot.r
 ggsave(filename="figs/ratio_centile_violin.png", cent.diffs.v.plot.r)
@@
 
-```{r}
-cent.diffs.v.boxplot.r <- subj.cent_err.r %>%
-  mutate(mean = mean*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=proportion, y=mean, fill=dataset)) +
+**Stats Tests**
+Within each M:F proportion simulated, is the magnitude of subjects' mean centile errors differ significantly by combat configuration?
+
+```{r}
+prop_abs.cent_t.tests <- mclapply(ratio_list, centile.t.tests, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
+names(prop_abs.cent_t.tests) <- n_prop_list
+prop_abs.cent_t.tests_df <- bind_rows(prop_abs.cent_t.tests, .id = "column_label")
+```
+
+Visalize
+```{r}
+ggplot(prop_abs.cent_t.tests_df) +
+  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
+  geom_hline(yintercept = 208, linetype='longdash') +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
+  facet_wrap(~comp)
+```
+
+## Sex-Biases
+
+```{r}
+ratio_subj.df %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=sex), scale="width") +
+  facet_wrap(~dataset) +
   labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-ggsave(filename="figs/ratio_centile_box.png", cent.diffs.v.plot.r)
-```
-
-### Sex-Biases
-
-```{r}
-cent.diffs.v.boxplot.r_sex_free <- subj.cent_err.r %>%
-  mutate(mean = mean*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
+```
+
+```{r}
+cent.diffs.v.boxplot.r_sex_free <- ratio_subj.df %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
   facet_wrap(~dataset, scale="free_y") +
@@
   ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
+cent.diffs.v.boxplot.r_sex_free
 ggsave(filename="figs/ratio_centile_box_sex_free.png", cent.diffs.v.boxplot.r_sex_free)
@@
 #trying to add significance testing
-cent.diffs.by.sex.stat.test <- subj.cent_err.r %>%
+abs.cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
   dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, proportion) %>%
-  wilcox_test(mean ~ sex) %>%
+  group_by(dataset, prop) %>%
+  wilcox_test(mean_cent_abs.diff ~ sex) %>%
   adjust_pvalue(method = "fdr") %>%
   add_significance() %>%
-  add_xy_position(x = "proportion")
-
-cent.diffs.v.boxplot.r_sex <- subj.cent_err.r %>%
-  mutate(mean = mean*100) %>% #for visuals
+  add_xy_position(x = "prop")
+
+abs.cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
   dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
@@
   ggplot() +
-  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
+  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
   facet_wrap(~dataset) +
@@
 cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_abs.centile_box_sex.png", cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-get diffs (not abs)
-
-```{r}
-get.diffs <- function(x){
-  df <- x %>%
-    mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-    mutate(dataset = relevel(dataset, ref= "raw")) %>%
-    arrange(dataset) %>%
-    group_by(participant) %>%
-    dplyr::mutate(across(all_of(pheno_list), ~ (. - first(.)), .names = "diff_{.col}")) %>% #centile diffs
-    dplyr::select(-any_of(pheno_list)) %>% #drop
-    dplyr::mutate(across(ends_with(".z"), ~ (. - first(.)), .names = "diff_{.col}")) %>% #z-score diffs
-    dplyr::select(-any_of(pheno_list.z)) %>% #drop
-    ungroup() %>%
-    dplyr::filter(dataset != "raw") #drop raw
-  return(df)
-}
-```
-
-```{r, eval=FALSE}
-#get diff in centiles and z scores across ratios
-ratio_diffs2 <- mclapply(ratio_predictions, get.diffs, mc.preschedule = FALSE)
-length(ratio_diffs2)
-```
-
-take w/in subj mean
-```{r, eval=FALSE}
-mean.subj.centile.err2 <- function(df){
-mean.diffs.subj <- df %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  rowwise() %>%
-  dplyr::mutate(mean = mean(c_across(starts_with("diff"))))
-
-return(mean.diffs.subj)
-#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
-}
-ratio_diffs.err2 <- mclapply(ratio_diffs2, mean.subj.centile.err2, mc.preschedule = FALSE)
-names(ratio_diffs.err2) <- levels(ratio_diffs.df$perm)
-subj.cent_err.r2 <- bind_rows(ratio_diffs.err2, .id = "proportion")
-saveRDS(subj.cent_err.r2, file="derivatives/ukb_ratio/ratio_centile_diffs.rds")
-```
-
-```{r}
-#load
-subj.cent_err.r2 <- readRDS("derivatives/ukb_ratio/ratio_centile_diffs.rds")
-```
-
-```{r}
-cent.diffs.v.boxplot.r_sex2 <- subj.cent_err.r2 %>%
-  mutate(mean = mean*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-ggsave(filename="figs/ratio_centile_box_sex2.png", cent.diffs.v.boxplot.r_sex)
-```
-
-Test for sex-effects in subj's mean centile errors
-
-```{r}
-cent.diffs.by.sex.stat.test <- subj.cent_err.r2 %>%
-  group_by(dataset, proportion) %>%
-  wilcox_test(mean ~ sex) %>%
+ggsave(filename="figs/ratio_abs.centile_box_sex.png", abs.cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
+```
+
+eval diffs (not abs. diffs)
+
+```{r}
+#trying to add significance testing
+cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
+  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
+  group_by(dataset, prop) %>%
+  wilcox_test(mean_cent_diff ~ sex) %>%
   adjust_pvalue(method = "fdr") %>%
   add_significance() %>%
-  add_xy_position(x = "proportion", dodge = 0.8)
-
-cent.diffs.v.boxplot.r_sex_free2 <- subj.cent_err.r2 %>%
-  mutate(mean = mean*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  add_xy_position(x = "prop")
+
+cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
+  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
+  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
+  droplevels() %>%
+  ggplot() +
+  geom_boxplot(aes(x=prop, y=mean_cent_diff, fill=sex)) +
+  facet_wrap(~dataset) +
+  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0.01,
+    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
     #hide.ns = TRUE
+    bracket.nudge.y = 4.5
     ) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex_free2
-ggsave(filename="figs/ratio_centile_box_sex_freescale2.png", cent.diffs.v.boxplot.r_sex_free2, width = 15, height=7)
-```
-
-### Extremeness-Biases
+  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
+cent.diffs.v.boxplot.r_sex
+ggsave(filename="figs/ratio_centile_box_sex.png", cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
+```
+
+### Significance Testing
+
+#### CENTILES
+within each M:F prop, test male vs female subject's mean abs. centile diff
+```{r}
+sex_bias_ratio_abs.tests <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, comp_multiplier = 11)
+names(sex_bias_ratio_abs.tests) <- n_prop_list
+sex_bias_ratio_abs.tests_df <- bind_rows(sex_bias_ratio_abs.tests, .id = "prop")
+```
+
+Visualize
+```{r}
+sex_bias_ratio_abs.tests_df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
+    geom_raster() +
+    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
+    theme_classic() +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+    coord_fixed()
+```
+-log(p)
+
+test actual (not abs. val) centile errors
+
+```{r}
+sex_bias_ratio_tests <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test = "mean_cent_diff", mc.preschedule = FALSE, comp_multiplier = 11)
+names(sex_bias_ratio_tests) <- n_prop_list
+sex_bias_ratio_tests_df <- bind_rows(sex_bias_ratio_tests, .id = "prop")
+```
+
+Visualize
+```{r}
+sex_bias_ratio_tests_df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
+    geom_raster() +
+    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
+    theme_classic() +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+    coord_fixed()
+```
+
+#### Z-SCORES
+
+```{r}
+sex_bias_ratio_abs.tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_abs.diff", mc.preschedule = FALSE, comp_multiplier = 11)
+names(sex_bias_ratio_abs.tests.z) <- n_prop_list
+sex_bias_ratio_abs.tests.z_df <- bind_rows(sex_bias_ratio_abs.tests.z, .id = "prop")
+```
+
+Visualize
+```{r}
+sex_bias_ratio_abs.tests.z_df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
+    geom_raster() +
+    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
+    theme_classic() +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+    coord_fixed()
+```
+
+```{r}
+sex_bias_ratio_tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_diff", mc.preschedule = FALSE, comp_multiplier = 11)
+names(sex_bias_ratio_tests.z) <- n_prop_list
+sex_bias_ratio_tests.z_df <- bind_rows(sex_bias_ratio_tests.z, .id = "prop")
+```
+
+Visualize
+```{r}
+sex_bias_ratio_tests.z_df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
+    geom_raster() +
+    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
+    theme_classic() +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+    coord_fixed()
+```
+
+## Extremeness-Biases
 Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)
 
+#### CENTILES
+
 test w/in subj avg centile ~ w/in subj avg centile error
@@
 ```{r}
-#mean w/in subj error
-subj.cent_err.only <- subj.cent_err.r2 %>%
-  dplyr::select(!starts_with("diff")) %>%
-  rename(mean_diff = mean)
-```
-
-```{r, eval=FALSE}
-#calc mean centile
-subj.cent.mean <- function(df){
-mean.subj <- df %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  rowwise() %>%
-  dplyr::mutate(mean_cent = mean(c_across(pheno_list))) %>%
-  dplyr::select(!pheno_list)
-
-return(mean.subj)
-#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
-}
-
-ratio_cent_means <- mclapply(ratio_predictions, subj.cent.mean, mc.preschedule = FALSE, mc.cores=12L)
-#bind predictions
-names(ratio_cent_means) <- n_ratio_list
-ratio_cent_means.df <- bind_rows(ratio_cent_means, .id = "column_label")
-saveRDS(ratio_cent_means.df, file="derivatives/ukb_ratios/ratio_cent_means.csv")
-```
-
-## Z-SCORES
-
-```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs 
-pheno_diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
-
-z.t.tests <- function(df){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in pheno_diff.list.z) {
-    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
-names(perm_t.tests_z) <- n_perm_list
-z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
-```
-
-```{r}
-fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
-```
-
-```{r}
-z_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
-
-## SCALE EFFECTS OF SEX
+#testing
+test.lm <- lm(mean_cent_diff ~ mean_centile*as.factor(dataset), data=ratio_subj_list[[1]])
+tidy(test.lm)
+interactions::interact_plot(test.lm, pred = mean_centile, modx = dataset)
+
+ggplot(ratio_subj_list[[4]], aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth() +
+  facet_wrap(~dataset, scales="free_y")
+```
+
+```{r}
+ext.bias_test.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ mean_cent_diff", comp_multiplier= 33, mc.preschedule=FALSE)
+names(ext.bias_test.ln) <- n_prop_list
+linear <- as.data.frame(bind_rows(ext.bias_test.ln, .id = "prop"))
+
+ext.bias_test.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
+names(ext.bias_test.sq) <- n_prop_list
+square <- as.data.frame(bind_rows(ext.bias_test.sq, .id = "prop"))
+
+ext.bias_test.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
+names(ext.bias_test.cb) <- n_prop_list
+cubic <- as.data.frame(bind_rows(ext.bias_test.cb, .id = "prop"))
+
+#one big dataframe
+ext.bias_test.df <- bind_rows(linear, square, cubic, .id = "model")
+```
+
+summarize across modeling options
+
+```{r}
+ext.bias_test.df %>%
+  group_by(dataset, prop) %>%
+  dplyr::arrange(p.val_fdr) %>%
+  slice(1) %>% #get lowest p-val
+  ungroup()
+```
+
+#### Z-SCORES
+```{r}
+ext.bias_test.z.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ mean_z_diff", comp_multiplier=33, mc.preschedule=FALSE)
+names(ext.bias_test.z.ln) <- n_prop_list
+linear <- as.data.frame(bind_rows(ext.bias_test.z.ln, .id = "prop"))
+
+ext.bias_test.z.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
+names(ext.bias_test.z.sq) <- n_prop_list
+square <- as.data.frame(bind_rows(ext.bias_test.z.sq, .id = "prop"))
+
+extremeness.lm(ratio_subj_list[[2]], f="mean_z ~ poly(mean_z_diff, 2)")
+
+ext.bias_test.z.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
+names(ext.bias_test.z.cb) <- n_prop_list
+cubic <- as.data.frame(bind_rows(ext.bias_test.z.cb, .id = "prop"))
+
+#one big dataframe
+ext.bias_test.z.df <- bind_rows(linear, square, cubic, .id = "model")
+```
+
+summarize across modeling options
+
+```{r}
+ext.bias_test.z.df %>%
+  group_by(dataset, prop) %>%
+  dplyr::arrange(p.val_fdr) %>%
+  slice(1) %>% #get lowest p-val
+  ungroup()
+```
+
+## Scale Effects of Sex
+
+sex-diff in variance as function of M:F!
+
 fix age, get diff between male and female pred. variance

### 2023-12-27 cubic_scripts/R_scripts/gamlss_helper_funs.R
troubleshooting

@@ -399,9 +399,9 @@ centile_predict <- function(gamlssModel, dataToPredictM, dataToPredictF, ageRang
   }
   
   print(paste("length:", length(fanCentiles)))
-  print(paste("max:", max(fanCentiles)))
-  print(paste("male max:", max(fanCentiles_M)))
-  print(paste("female max:", max(fanCentiles_F)))
+  print(fanCentiles)
+  #print(paste("male max:", max(fanCentiles_M)))
+  #print(paste("female max:", max(fanCentiles_F)))
   
   # to get peaks, match median point with age ranges
   med.idx <- ceiling(length(desiredCentiles) / 2) #find median centile

### 2023-12-27 paper_results.Rmd
qc for centiles at 0 and 1

@@
 ```
+
 # Main Analyses
@@
 unique(qc_z_df$participant)
-ppt.list
+
+qc_cent_df %>%
+  dplyr::filter(centile != 1) %>%
+  summarize(sprintf("%.100f", (1 - max(centile))))
+
+qc_cent_df %>%
+  summarize(sprintf("%.100f", min(centile)))
 ```
@@
 ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
+```
+
+##CENTILE FANS
+
+pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' .`
+
+```{r}
+basic_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_basic"
+raw_files <- list.files(path = basic_data_path, pattern = "_centiles.csv", full.names = TRUE)
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  print(paste("reading file", file))
+  # Read each CSV file
+    data <- fread(file)
+
+    data <- data %>%
+     mutate(dataset = gsub("-data", "", dataset))
+
+    # Bind the data to the combined dataframe
+    df_list[[file]] <- as.data.frame(data)
+    assign("basic_centiles", df_list)
+}
+length(basic_centiles)
+centiles.df <- bind_rows(basic_centiles)
+skimr::skim(centiles.df)
+```
+
+```{r}
+#also load raw data
+ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")
+```
+
+just work with global volumes for now
+
+```{r}
+global.cent.df <- centiles.df %>%
+  dplyr::filter(pheno_cat == "Global Volume")
+
+global.cent.df %>%
+  group_by(pheno) %>%
+  summarise(max=max(centile_5))
+
+names(global.cent.df)
+
+gmv.df <- global.cent.df
+
+gmv.df <- centiles.df %>%
+  dplyr::filter(pheno =="GMV")
+
+  ggplot(gmv.df) +
+      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df$GMV), alpha=0.5) +
+      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+      geom_line(aes(x=gmv.df$age_days, y=gmv.df$M_centile_1, color=gmv.df$dataset)) +
+      geom_line(aes(x=gmv.df$age_days, y=gmv.df$M_centile_2, color=gmv.df$dataset)) +
+      geom_line(aes(x=gmv.df$age_days, y=gmv.df$M_centile_3, color=gmv.df$dataset)) +
+      geom_line(aes(x=gmv.df$age_days, y=gmv.df$M_centile_4, color=gmv.df$dataset)) +
+      geom_line(aes(x=gmv.df$age_days, y=gmv.df$M_centile_5, color=gmv.df$dataset)) +
+      facet_wrap(~ pheno, scales="free_y") + 
+      labs(title="Male GMV brain charts")
+
+      geom_line(aes(x=ages, y=pred$fanCentiles[[2]])) +
+      geom_line(aes(x=ages, y=pred$fanCentiles[[3]]), alpha=0.6) +
+      geom_line(aes(x=ages, y=pred$fanCentiles[[4]]), alpha=0.8) +
+      geom_line(aes(x=ages, y=pred$fanCentiles[[5]])) +
+      geom_line(aes(x=ages, y=pred$fanCentiles[[6]]), alpha=0.8) +
+      geom_line(aes(x=ages, y=pred$fanCentiles[[7]]), alpha=0.6) +
+      geom_line(aes(x=ages, y=pred$fanCentiles[[8]]), alpha=0.4) +
+      geom_line(aes(x=ages, y=pred$fanCentiles[[9]]), alpha=0.2) +
+      geom_point(aes(x=peak_age, y=pred$peak), size=3) +
+      scale_x_continuous(breaks=tickMarks, labels=tickLabels,
+                         limits=c(tickMarks[[1]], max(age_col))) +
+      labs(title=deparse(substitute(gamlssModel))) +
+      theme(legend.title= element_blank())+
+      xlab(paste("Age at Scan", unit_text)) +
+      ylab(deparse(substitute(phenotype)))
+    annotate("text", x = max(age_col)-0.5, y = max(yvar)-10000, label = paste0("BIC=",gamlssModel$sbc))
+    print(sampleCentileFan)
 ```

### 2023-12-27 paper_results.Rmd
calc more centiles for fan plots

--- 2023-12-27 paper_results.Rmd
@@
 ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
+```
+
+##CENTILE FANS
+
+pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' .`
+
+```{r}
+basic_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_basic"
+raw_files <- list.files(path = basic_data_path, pattern = "_centiles.csv", full.names = TRUE)
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  print(paste("reading file", file))
+  # Read each CSV file
+    data <- fread(file)
+
+    data <- data %>%
+     mutate(dataset = gsub("-data", "", dataset))
+
+    # Bind the data to the combined dataframe
+    df_list[[file]] <- as.data.frame(data)
+    assign("basic_centiles", df_list)
+}
+length(basic_centiles)
+centiles.df <- bind_rows(basic_centiles)
+#skimr::skim(centiles.df)
+```
+
+```{r}
+#also load raw data
+ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")
+
+global.df.long <- ukb_cn_df %>%
+  dplyr::select(participant, age_days, sex, vol_list_global) %>%
+  pivot_longer(cols = vol_list_global, values_to = "value", names_to="pheno")
+```
+
+just work with global volumes for now
+
+```{r}
+global.cent.df <- centiles.df %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  dplyr::filter(!grepl('-refA', dataset))
+
+for (feature in unique(global.cent.df$pheno)) {
+
+  df <- global.cent.df %>%
+    dplyr::filter(pheno == feature)
+
+ plot <- ggplot() +
+      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
+      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_1, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_2, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_3, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_4, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_5, color=df$dataset)) +
+      labs(title=paste("Male", feature, "brain chart"))
+
+ print(plot)
+
+ plot <- ggplot() +
+      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
+      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_1, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_2, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_3, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_4, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_5, color=df$dataset)) +
+      labs(title=paste("Female", feature, "brain chart"))
+
+ print(plot)
+
+ plot <- ggplot() +
+      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
+      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+      geom_line(aes(x=df$age_days, y=df$centile_1, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_2, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_3, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_4, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_5, color=df$dataset)) +
+      labs(title=paste(feature, "brain chart"))
+
+ print(plot)
+
+}
 ```

### 2023-12-27 cubic_scripts/R_scripts/gamlss_helper_funs.R

@@ -286,7 +286,7 @@ sim.data.ukb <- function(df){
   #dataToPredictF$CT_total <- mean.tb.sim(df, "Female", ageRange, "CT_total")
   
   # List of centiles for the fan plot
-  desiredCentiles <- c(0.1, 0.25, 0.5, 0.75, 0.9)
+  desiredCentiles <- c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
   
   # return
   sim <- list(ageRange, dataToPredictM, dataToPredictF, desiredCentiles)

### 2023-12-28 paper_results.Rmd
fixing so that centiles = 1 are actually replaced correctly

--- 2023-12-27 paper_results.Rmd
@@
 
+# Setup 
 ```{r setup, include=FALSE}
@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel)
 #Sys.setenv("MC_CORES"=10L)
@@
 ```{r}
-data_path <- "./ukb_basic"
-
-### LOAD ###
+data_path <- "./derivatives"
+
 pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
@@
 pheno_list.z <- paste0(pheno_list, ".z")
+#also load raw data
+ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")
 ```
@@
 ```
+
 # Main Analyses
 
-## CENTILES
-Do centile scores change based on combat version?
-
-```{r}
-pred.csvs <- list.files(path = data_path, pattern = "_predictions.csv", full.names = TRUE)
+##Load Data
+
+pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' .`
+
+pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' .`
+
+```{r}
+pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_predictions.csv", full.names = TRUE)
 combined_preds <- data.frame()
@@
 ```{r}
-qc_cent_df <- combined_preds %>% 
-  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0, Inf, -Inf, NA))) %>%
+combined_preds %>% 
+  filter_at(vars(pheno_list), any_vars(. %in% c(1))) %>%
+  dplyr::select(participant) %>%
+  unique()
   dplyr::select(File_ID, participant, sex, dataset, pheno_list) %>%
@@
 unique(qc_z_df$participant)
-ppt.list
-```
-
-*MOVE TO CUBIC*
-
-```{r, eval=FALSE}
-#calculate abs error
-cent_abs.diffs <- combined_preds %>%
-  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
-  ungroup() %>%
-  dplyr::filter(dataset != "ukb_CN") #drop raw
-
-saveRDS(cent_abs.diffs, "data/centile_abs_diffs.rds")
-```
-
-```{r}
-cent_abs.diffs <- readRDS("data/centile_abs_diffs.rds")
-```
+
+qc_cent_df %>%
+  dplyr::filter(centile != 1) %>%
+  summarize(sprintf("%.100f", (1 - max(centile))))
+
+qc_cent_df %>%
+  summarize(sprintf("%.100f", min(centile)))
+```
+## CENTILES
+
+Do centile scores change based on combat version?
 
@@
 ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
+```
+
+##CENTILE FANS
+
+pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' .`
+
+```{r}
+basic_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_basic"
+raw_files <- list.files(path = basic_data_path, pattern = "_centiles.csv", full.names = TRUE)
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  print(paste("reading file", file))
+  # Read each CSV file
+    data <- fread(file)
+
+    data <- data %>%
+     mutate(dataset = gsub("-data", "", dataset))
+
+    # Bind the data to the combined dataframe
+    df_list[[file]] <- as.data.frame(data)
+    assign("basic_centiles", df_list)
+}
+length(basic_centiles)
+centiles.df <- bind_rows(basic_centiles)
+#skimr::skim(centiles.df)
+names(centiles.df)
+```
+
+```{r}
+global.df.long <- ukb_cn_df %>%
+  dplyr::select(participant, age_days, sex, vol_list_global) %>%
+  pivot_longer(cols = vol_list_global, values_to = "value", names_to="pheno")
+```
+
+just work with global volumes for now
+
+```{r}
+global.cent.df <- centiles.df %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  dplyr::filter(!grepl('-refA', dataset))
+
+for (feature in unique(global.cent.df$pheno)) {
+
+  df <- global.cent.df %>%
+    dplyr::filter(pheno == feature)
+
+ plot <- ggplot() +
+      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
+      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_1, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_2, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_3, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_4, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_5, color=df$dataset), linewidth=1.5) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_6, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_7, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_8, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$M_centile_9, color=df$dataset)) +
+      labs(title=paste("Male", feature, "brain chart"))
+
+ print(plot)
+
+ plot <- ggplot() +
+      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
+      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_1, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_2, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_3, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_4, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_5, color=df$dataset), linewidth=1.5) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_6, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_7, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_8, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$F_centile_9, color=df$dataset)) +
+      labs(title=paste("Female", feature, "brain chart"))
+
+ print(plot)
+
+ plot <- ggplot() +
+      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
+      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+      geom_line(aes(x=df$age_days, y=df$centile_1, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_2, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_3, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_4, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_5, color=df$dataset), linewidth=1.5) +
+      geom_line(aes(x=df$age_days, y=df$centile_6, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_7, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_8, color=df$dataset)) +
+      geom_line(aes(x=df$age_days, y=df$centile_9, color=df$dataset)) +
+      labs(title=paste(feature, "brain chart"))
+
+ print(plot)
+
+}
+```
+
+Looking at WMV more specifically for visualizations
+```{r}
+#wm centile lines
+wmv.df <- centiles.df %>%
+  dplyr::filter(pheno == "WMV") %>%
+  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
+  #pivot_longer(cols=contains("centile"), name)  %>%
+  pivot_longer(
+    cols = contains("centile"),
+    names_to = c("sex", ".value"),
+    names_pattern = "([MF]?)_(centile_\\d+)",
+    values_to = "value"
+  ) %>%
+ dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
+  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
+                                   sex == "F" ~ "Female")),
+         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
+```
+
+```{r}
+ggplot(wmv.df) +
+  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$WMV), alpha=0.7, width=150, color="lightgray") +
+  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
+  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
+  facet_wrap(~sex) + 
+  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "dashed")) +
+  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  theme_minimal() +
+  geom_text_repel(aes(x = max(ukb_cn_df$age_days), y = min(centile_1)),
+                  label="test",
+                  nudge_x = 1,
+                  na.rm = TRUE) + 
+  labs(title="WMV brain charts")
+```
+
+```{r}
+#looking at other phenotypes
+ven.df <- centiles.df %>%
+  dplyr::filter(pheno == "Ventricles") %>%
+  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  #pivot_longer(cols=contains("centile"), name)  %>%
+  pivot_longer(
+    cols = contains("centile"),
+    names_to = c("sex", ".value"),
+    names_pattern = "([MF]?)_(centile_\\d+)",
+    values_to = "value"
+  ) %>%
+ dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
+  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
+                                   sex == "F" ~ "Female")),
+         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
+
+ggplot(ven.df) +
+  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$Ventricles), alpha=0.7, width=150, color="lightgray") +
+  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
+  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
+  facet_wrap(~sex) + 
+  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid")) +
+  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846")) +
+  theme_minimal() +
+  labs(title="sGMV brain charts")
+```
+```{r}
+#wm centile lines
+r.sa.pc.df <- centiles.df %>%
+  dplyr::filter(pheno == "rh_SA_postcentral") %>%
+  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  #pivot_longer(cols=contains("centile"), name)  %>%
+  pivot_longer(
+    cols = contains("centile"),
+    names_to = c("sex", ".value"),
+    names_pattern = "([MF]?)_(centile_\\d+)",
+    values_to = "value"
+  ) %>%
+ dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
+  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
+                                   sex == "F" ~ "Female")),
+         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
+
+ggplot(r.sa.pc.df) +
+  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$rh_SA_postcentral), alpha=0.7, width=150, color="lightgray") +
+  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
+  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
+  facet_wrap(~sex) + 
+  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid")) +
+  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846")) +
+  theme_minimal() +
+  labs(title="SA of right postcentral brain charts")
 ```

### 2023-12-28 cubic_scripts/R_scripts/gamlss_helper_funs.R

@@ -471,13 +471,17 @@ get.og.data.centiles <- function(gamlss.rds.file, og.data, get.zscores = FALSE){
       
       #don't let centile = 1!
       if (centiles[i] == 1) {
-        centiles[i] <- 0.9999999999999999999999999 #25 dec places, should be plenty based on next-highest non-1 centile value
+        centiles[i] <- 0.99999999999999994 #largest number i could get w/o rounding to 1 (trial & error)
       }
       #don't let centile = 0!
       if (centiles[i] == 0) {
         centiles[i] <- 0.0000000000000000000000001 #25 dec places, should be plenty based on min centile
       }
   }
+  #double check there are no impossible centiles
+  if(c(1,0) %in% centiles) 
+    stop("Error: predicted centiles include 0 or 1!")
+  
   if (get.zscores == FALSE){
   return(centiles)
   } else {

### 2024-01-02 paper_results.Rmd

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain)
 #Sys.setenv("MC_CORES"=10L)
@@
 ```{r}
-data_path <- "./derivatives"
+data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"
 
@@
 pheno_list.z <- paste0(pheno_list, ".z")
+pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)
+z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")
+
 #also load raw data
@@
 Can test mean (not abs) centile errors, Z errors, etc, using `to_test` arg.
-FDR corrects for number of
+FDR corrects for number of datasets tested
 
@@
   #apply fdr correction
-  n_comp=nrow(df)
+  n_comp=length(unique(df$dataset))
   df.sex.t <- df.sex.t %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
            sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
@@
     ungroup()
-  n_comp=nrow(df)
+  n_comp=length(unique(df$dataset))
 
@@
 
-##Load Data
-
-pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' .`
-
-pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' .`
+## Load Data
+
+pulled variance from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_variance.csv' .`
+
+Get predicted centiles
+- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' .`
 
@@
 combined_preds %>% 
-  filter_at(vars(pheno_list), any_vars(. %in% c(1))) %>%
+  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0))) %>%
   dplyr::select(participant) %>%
   unique()
-  dplyr::select(File_ID, participant, sex, dataset, pheno_list) %>%
-  pivot_longer(cols=pheno_list, names_to="pheno", values_to="centile")
-skimr::skim(qc_cent_df)
-
-qc_cent_df %>%
-  dplyr::filter(centile == 0)
-
-qc_cent1_df <- qc_cent_df %>%
-  dplyr::filter(centile == 1)
-
-qc_cent1_df %>%
-  ggplot() +
-  geom_bar(aes(dataset, fill=participant)) +
-  facet_wrap(~pheno)
-
-ppt.list <- qc_cent1_df %>%
+
+combined_preds %>% 
+  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
   dplyr::select(participant) %>%
-  unique() %>%
-  as.list() %>%
-  unname() %>%
-  unlist()
-
-qc_z_df <- combined_preds %>% 
-  filter_at(vars(pheno_list.z), any_vars(. %in% c(1, 0, Inf, -Inf, NA)))
-
-unique(qc_z_df$participant)
-
-qc_cent_df %>%
-  dplyr::filter(centile != 1) %>%
-  summarize(sprintf("%.100f", (1 - max(centile))))
-
-qc_cent_df %>%
-  summarize(sprintf("%.100f", min(centile)))
-```
+  unique()
+
+# test.df <- combined_preds %>% 
+#   dplyr::select(pheno_list)
+# 
+# test.df[test.df==1] <- 0
+# 
+# format(round(max(do.call(pmax, test.df)), 20), nsmall = 20)
+```
+
+saving out csv forces some centile scores to round up to 1, BUT it appears z scores are still being calculated correctly so just going to proceed!
+
+Get errors
+- pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' .`
+
+```{r}
+raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
+
+diffs.df <- fread(file = raw_files, sep=",") %>%
+     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+```
+
+Get subject-wise summaries
+- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' .`
+
+```{r}
+raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_subj_pred.csv", full.names = TRUE) %>% unlist()
+
+subj_sum.df <- fread(file = raw_files, sep=",") %>%
+     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+```
+
 ## CENTILES
@@
                    "p.value" = double())
-pheno_diff.list <- paste0("abs.diff_", pheno_list)
 
 #run tests
-attach(cent_abs.diffs)
+attach(diffs.df)
 for (pheno in pheno_diff.list) {
-  # df.aov <- tidy(kruskal.test(cent_abs.diffs[[pheno]] ~ cent_abs.diffs[["dataset"]]))
+  # df.aov <- tidy(kruskal.test(diffs.df[[pheno]] ~ diffs.df[["dataset"]]))
   # df.aov$pheno <- pheno
@@
 
-  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=cent_abs.diffs[[pheno]], g=cent_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
+  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=diffs.df[[pheno]], g=diffs.df[["dataset"]], paired = TRUE, p.adj = "none"))
   df.pairwise.t$pheno <- pheno
@@
 
-# t.df.sum <- t.df %>%
-#   group_by(comp) %>%
-#   summarize(n_sig_regions = sum(sig_fdr),
-#             n_compare = n(),
-#             min_pval = min(p.val_fdr),
-#             max_pval = max(p.val_fdr))
-# t.df.sum
+t.df.sum <- t.df %>%
+  group_by(comp) %>%
+  summarize(n_sig_regions = sum(sig_fdr),
+            n_compare = n(),
+            min_pval = min(p.val_fdr),
+            max_pval = max(p.val_fdr))
+t.df.sum
 ```
@@
 #get means w/in feature 
-mean.diffs.long <- cent_abs.diffs %>%
+mean.diffs.long <- diffs.df %>%
+  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
@@
 
+### Sex-Biases
+
+Visualize magnitude mean w/in subj error
+
+```{r}
+subj_sum.df %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=sex)) +
+  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations")
+
+subj_sum.df %>%
+  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
+  ggplot(aes(x=dataset, y=mean_cent_diff, fill=sex)) +
+  geom_boxplot(trim=FALSE, scale="width") +
+  labs(y="Centile Error", x = "ComBat Configurations") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Within-subject mean error in centile scores across ComBat configurations")
+
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
+  ggplot(aes(x=sim.site, y=mean_cent_abs.diff, fill=sex)) +
+  geom_violin(trim=FALSE) +
+  facet_wrap(~dataset) +
+  labs(y="Magnitude Centile Error", x = "Simulated Site") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Within-subject mean magnitude error in centile scores across ComBat configurations")
+
+subj_sum.df %>%
+  ggplot() +
+  geom_bar(aes(x=sim.site, fill=sex))
+```
+
+Significance testing (w/in sub. mean abs centile error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
+
+```{r}
+sex.bias.t.tests(subj_sum.df)
+```
+
 ## Z-SCORES
@@
 look for skew in distribution
-
-```{r, eval=FALSE}
-#calc errors
-z_abs.diffs <- combined_preds %>%
-  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref = "ukb_CN")) %>%
-  arrange(dataset) %>%
-  group_by(participant) %>%
-  mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
-  ungroup() %>%
-  dplyr::filter(dataset != "ukb_CN") #drop raw
-saveRDS(z_abs.diffs, "data/z_abs.diffs.rds")
-```
-
-```{r}
-z_abs.diffs <- readRDS("data/z_abs.diffs.rds")
-```
 
@@
 
-z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")
-
 #run tests
-attach(z_abs.diffs)
+attach(diffs.df)
 for (pheno in z.pheno_abs.diff.list) {
-  # df.aov <- tidy(kruskal.test(z_abs.diffs[[pheno]] ~ z_abs.diffs[["dataset"]]))
+  # df.aov <- tidy(kruskal.test(diffs.df[[pheno]] ~ diffs.df[["dataset"]]))
   # df.aov$pheno <- pheno
@@
 
-  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=z_abs.diffs[[pheno]], g=z_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
+  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=diffs.df[[pheno]], g=diffs.df[["dataset"]], paired = TRUE, p.adj = "none"))
   df.pairwise.t$pheno <- pheno
@@
 
-# sum_z.t.df <- z.t.df %>%
-#   group_by(comp) %>%
-#   summarize(n_sig_regions = sum(sig_fdr),
-#             n_compare = n(),
-#             min.pval = min(p.val_fdr),
-#             max.pval = max(p.val_fdr))
-# 
-# sum_z.t.df
+sum_z.t.df <- z.t.df %>%
+  group_by(comp) %>%
+  summarize(n_sig_regions = sum(sig_fdr),
+            n_compare = n(),
+            min.pval = min(p.val_fdr),
+            max.pval = max(p.val_fdr))
+
+sum_z.t.df
 ```
@@
 #get means w/in feature 
-z.mean.diffs.long <- z_abs.diffs %>%
+z.mean.diffs.long <- diffs.df %>%
+  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
@@
 
+### Sex-Biases
+
+Visualize magnitude mean w/in subj error
+
+```{r}
+subj_sum.df %>%
+  ggplot() +
+  geom_violin(aes(x=dataset, y=mean_z_abs.diff, fill=sex)) +
+  labs(y= "Magnitude Z-Score Error", x = "ComBat Configurations") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
+
+subj_sum.df %>%
+  ggplot(aes(x=sex, y=mean_z_diff, fill=sex)) +
+  geom_boxplot(trim=FALSE, scale="width") +
+  facet_wrap(~dataset, scale="free_y") +
+  labs(y="Z-Score Error", x = "ComBat Configurations") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Within-subject mean error in z-scores across ComBat configurations")
+
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=sim.site, y=mean_z_abs.diff, fill=sex)) +
+  geom_violin(trim=FALSE) +
+  facet_wrap(~dataset) +
+  labs(y="Magnitude Z Error", x = "Simulated Site") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
+
+subj_sum.df %>%
+  ggplot() +
+  geom_bar(aes(x=sim.site, fill=sex))
+```
+
+Significance testing (w/in sub. mean abs z error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
+
+```{r}
+sex.bias.t.tests(subj_sum.df, to_test="mean_z_abs.diff")
+sex.bias.t.tests(subj_sum.df, to_test="mean_z_diff")
+```
+
 ## SCALE EFFECTS OF SEX
+
 fix age, get diff between male and female pred. variance
+
 ```{r}
@@
 
-##CENTILE FANS
+## CENTILE FANS
 
@@
 
-maybe better to get sex_summary.csv - just need signifigance
+maybe better to get sex_summary.csv - just need significance
 
@@
 ```
--log(p)
 

### 2024-01-04 paper_results.Rmd
plots, easier pulls

@@
 sex.bias.t.tests(subj_sum.df)
+sex.bias.t.tests(subj_sum.df, to_test="mean_cent_diff")
+```
+
+### Extremeness-Biases
+
+Group at 0-0.25, >0.25 - .75 <, 0.75-1
+
+```{r}
+subj_sum.df <- subj_sum.df %>%
+  mutate(cent_cat = as.factor(case_when(mean_centile <= 0.25 ~ "low",
+                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
+                              mean_centile >= 0.75 ~ "high")),
+         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
+                              mean_z > -1 & mean_z < 1 ~ "mid",
+                              mean_z >= 1 ~ "high")))
+```
+
+```{r}
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=mean_z, y=mean_z_diff, color=sex)) +
+  geom_point(alpha=0.5) +
+  geom_smooth() +
+  facet_wrap(~dataset, scales="free_y")
 ```

### 2024-01-04 cubic_scripts/R_scripts/combat_apply_w_shift.R

@@ -172,7 +172,6 @@ final.df <- cf.merged %>%
          SA_total=rowSums(dplyr::select(., .dots=all_of(c(sa_list)))),
          CT_total=rowSums(dplyr::select(., .dots=all_of(c(ct_list)))))
 
-
 ##########################################################################
 #WRITE OUT
 

### 2024-01-04 cubic_scripts/pull_derivatives.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+#PULL ALL OUTPUTS FROM CUBIC  TO LOCAL
+base=/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives
+
+# Main Analyses #
+if ! [ -d $base/ukb_basic ]
+	then
+	mkdir $base/ukb_basic
+	fi
+
+scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*.csv' $base/ukb_basic
+scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' $base/ukb_basic
+
+# Replication (Permutations) #
+if ! [ -d $base/ukb_perm ]
+	then
+	mkdir $base/ukb_perm
+	fi
+
+scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*.csv' $base/ukb_perm
+scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*.csv' $base/ukb_perm
+
+# Varying M:F Ratios (Proportions) #
+if ! [ -d $base/ukb_ratio ]
+	then
+	mkdir $base/ukb_ratio
+	fi
+
+scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*.csv' $base/ukb_ratio
+scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' $base/ukb_ratio
+
+# LBCC Lifespan #

### 2024-01-04 paper_results.Rmd
cleaning up stat tests for main analyses

--- 2024-01-02 paper_results.Rmd
@@
 
-Test for differences in abs.
-error distributions across combat configurations using nonparametric kruskal-wallace test (instead of anova.
-good that it's nonparametric b/c non-normal distributions BUT also the data is paired!).
-**confirm with andrew**
 Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
@@
 Using custom functions from `ranked_welch_tests.R`.
-
-**need to confirm w andrew then update output table formatting as needed - skipping for now**
 
@@
 sex.bias.t.tests(subj_sum.df)
+sex.bias.t.tests(subj_sum.df, to_test="mean_cent_diff")
+```
+
+### Extremeness-Biases
+
+Group at 0-0.25, >0.25 - .75 <, 0.75-1
+
+```{r}
+subj_sum.df <- subj_sum.df %>%
+  mutate(cent_cat = factor(case_when(mean_centile <= 0.25 ~ "low",
+                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
+                              mean_centile >= 0.75 ~ "high"), levels= c("low", "mid", "high"), ordered=TRUE),
+         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
+                              mean_z > -1 & mean_z < 1 ~ "mid",
+                              mean_z >= 1 ~ "high")))
+```
+
+```{r}
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth() +
+  facet_grid(dataset ~ sim.site, scales="free_y")
+```
+
+```{r}
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=cent_cat, y=mean_cent_diff, color=sex)) +
+  geom_violin() +
+  facet_wrap(~dataset, scales="free_y")
+
+subj_sum.df %>%
+  #dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=cent_cat, y=mean_cent_diff)) +
+  geom_violin() +
+  facet_wrap(~dataset, scales="free_y")
 ```
@@
 ```{r}
-pred.csvs <- list.files(path = data_path, pattern = "_variance.csv", full.names = TRUE)
+pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
 combined_var <- data.frame()
@@
   dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
@@
 ```{r}
-basic_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_basic"
-raw_files <- list.files(path = basic_data_path, pattern = "_centiles.csv", full.names = TRUE)
+raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_centiles.csv", full.names = TRUE)
 
@@
 ```{r}
-perm_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm_csvs"
+perm_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm"
 raw_files <- list.files(path = perm_data_path, pattern = "_diffs.csv", full.names = TRUE)
@@
 
+```{r}
+for (l in unique(ratio_subj.df$dataset)){
+plot <- ratio_subj.df %>%
+  dplyr::filter(dataset == l) %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth() +
+  facet_grid(sim.site ~ prop) +
+  ggtitle(l)
+print(plot)
+}
+```
+
 Do centile scores change based on combat version?

### 2024-01-04 paper_results.Rmd
assumption testing for w/in feat sex-diffs in error t testing

@@
 ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")
+
+#plotting info from dk atlas
+dk.parc <- dk$data %>%
+  as.data.frame() %>%
+  na.omit() %>%
+  dplyr::select(c(hemi, region, label)) %>%
+  distinct()
 ```
@@
 #define function
-pheno_diff.list <- paste0("abs.diff_", pheno_list)
-
-centile.t.tests <- function(df, feature_list=pheno_diff.list, comp_multiplier=1){
+pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)
+
+centile.t.tests <- function(df, feature_list=pheno_abs.diff.list, comp_multiplier=1){
   #initialize empty dfs to store outputs
@@
 
+modifying to test w/in feature rather than mean centiles
+
+```{r}
+sex.bias.feat.t.tests <- function(df, feature_list=pheno_list, comp_multiplier=1){
+
+  #initialize empty df to store outputs
+  t.df <- data.frame("dataset" = character(),
+                     "pheno" = character(),
+                     "p.value" = double(),
+                     "sex_diff" = double(),
+                     "t.stat" = double(),
+                     "df" = double())
+  attach(df)
+  for (pheno in feature_list) {
+  #conduct t test
+  df.sex.t <- df %>%
+    group_by(dataset) %>%
+    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$p.value,
+              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$estimate,
+              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$statistic,
+              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
+    ungroup()
+  df.sex.t$pheno <- pheno
+  t.df <- rbind(t.df, df.sex.t)
+  }
+
+  #apply fdr correction
+  n_comp=length(unique(df$dataset))*length(feature_list)
+  df.sex.t <- t.df %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE),
+           dataset=as.factor(dataset))
+
+  #summarize
+  df.sex.t.sum <- df.sex.t %>%
+    group_by(dataset) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+
+  return(df.sex.t.sum)
+  #detach()
+}
+
+# tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
+# rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
+```
+
+trying to get interpretable effect sizes w t.test (rather than ranked t test)
+
+```{r}
+sex.bias.feat.t.tests_2 <- function(df, feature_list=pheno_list, comp_multiplier=1){
+
+  #initialize empty df to store outputs
+  t.df <- data.frame("dataset" = character(),
+                     "pheno" = character(),
+                     "p.value" = double(),
+                     "sex_diff" = double(),
+                     "t.stat" = double(),
+                     "df" = double())
+  attach(df)
+  for (pheno in feature_list) {
+  #conduct t test
+  df.sex.t <- df %>%
+    group_by(dataset) %>%
+    summarise(p.value = tidy(t.test(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$p.value,
+              sex_diff = tidy(t.test(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$estimate,
+              t.stat = tidy(t.test(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$statistic,
+              df = tidy(t.test(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
+    ungroup()
+  df.sex.t$pheno <- pheno
+  t.df <- rbind(t.df, df.sex.t)
+  }
+
+  #apply fdr correction
+  n_comp=length(unique(df$dataset))*length(feature_list)
+  df.sex.t <- t.df %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE),
+           dataset=as.factor(dataset))
+
+  return(df.sex.t)
+  #detach()
+}
+
+# tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
+# rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
+```
+
 ```{r}
@@
 Do centile scores change based on combat version?
-
+Foregoing anova based on emails with andrew.
 Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
@@
 ```{r}
-#initalize empty dfs to store outputs
-sum.df <- data.frame(pheno = character(),
-                     term = character(),
-                     df = double(),
-                     sumsq = double(),
-                     meansq = double(),
-                     statistic = double(),
-                     p.value = double())
-t.df <- data.frame("pheno" = character(),
-                   "group1" = character(),
-                   "group2" = character(),
-                   "p.value" = double())
-
-#run tests
-attach(diffs.df)
-for (pheno in pheno_diff.list) {
-  # df.aov <- tidy(kruskal.test(diffs.df[[pheno]] ~ diffs.df[["dataset"]]))
-  # df.aov$pheno <- pheno
-  # sum.df <- rbind(sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=diffs.df[[pheno]], g=diffs.df[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  t.df <- rbind(t.df, df.pairwise.t)
-}
-detach()
-
-#apply fdr correction
-# sum.df <- sum.df %>%
-#   dplyr::filter(term != "Residuals") %>%
-#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-t.df <- t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-t.df.sum <- t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min_pval = min(p.val_fdr),
-            max_pval = max(p.val_fdr))
-t.df.sum
+centile.t.tests(diffs.df)
 ```
@@
 subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
   mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
   ggplot() +
-  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=sex)) +
+  geom_boxplot(aes(x=sex, y=mean_cent_abs.diff, fill=sex)) +
   labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~dataset) +
+  theme(legend.position="none") +
   ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations")
@@
   ggplot(aes(x=dataset, y=mean_cent_diff, fill=sex)) +
-  geom_boxplot(trim=FALSE, scale="width") +
+  geom_violin(trim=FALSE, scale="width") +
   labs(y="Centile Error", x = "ComBat Configurations") +
@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in centile scores across ComBat configurations")
+  ggtitle("Within-subject mean magnitude error in centile scores across \nComBat configurations")
 
@@
 sex.bias.t.tests(subj_sum.df, to_test="mean_cent_diff")
+```
+
+Sig test w/in feature diffs
+
+check for skew (welch t.test vs ranked welch t test)
+```{r}
+for (pheno in pheno_list){
+  val <- paste0("diff_", pheno)
+  p <- ggplot(diffs.df, aes(x = diffs.df[[val]], fill= sex, color=sex)) +
+      geom_histogram(alpha=0.5, position="identity") +
+    ggtitle(val)
+print(p)
+}
+```
+
+```{r}
+sex.bias.feat.t.tests(diffs.df)
+```
+
+Sig test w/in feature diffs
+```{r, warning=FALSE}
+sex_t_tests_in_feat.df <- sex.bias.feat.t.tests_2(diffs.df)
+```
+
+```{r}
+sex_t_tests_in_feat.df.sum <- sex_t_tests_in_feat.df %>%
+    group_by(dataset) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+```
+
+```{r}
+sex_t_tests_in_feat.df2 <- sex_t_tests_in_feat.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    label = sub("_[^_]*_", "_", pheno)) #for plotting
+sex_t_tests_to_plot <- left_join(sex_t_tests_in_feat.df2, dk.parc, by="label")
+```
+
+plot w/in feature t tests
+
+```{r}
+#var plot
+sex.var.plot.df <- sex_t_tests_to_plot %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ sex_diff)) %>%
+  group_by(dataset, pheno_cat)
+
+tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="ukb_CN")
+tp$sex_diff <- 1
+
+sex.var.plot <- ggplot(sex.var.plot.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = sex_diff)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.003, 0.003))
+sex.var.plot
+
+#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
+```
+
+```{r}
+#plot global
+sex.var.plot.global.df <- sex_t_tests_to_plot %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "WM",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA)),
+        sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ sex_diff)) %>% #drop NS values
+  group_by(dataset)
+
+tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="ukb_CN")
+tp$sex_standard <- 1
+
+sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = sex_diff),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.003, 0.003))
+sex.var.plot.global
+```
+
+Combine plots
+```{r}
+p <- ggarrange(
+  var.plot.global, var.plot,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.1),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+
+p_annotate <- annotate_figure(p, top = text_grob("Figure 1. Standardized effect of male sex on feature variance", 
+               color = "black", size = 14, hjust=.65))
+
+ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
 ```
@@
   dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
-
-#add more info from parcellations
-dk.parc <- dk$data %>%
-  as.data.frame() %>%
-  na.omit() %>%
-  dplyr::select(c(hemi, region, label)) %>%
-  distinct()
 
@@
 #within each phenotype, iterate paired t test between each combo of combat dfs 
-pheno_diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
+pheno_abs.diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
 
@@
   attach(df)
-  for (pheno in pheno_diff.list.z) {
+  for (pheno in pheno_abs.diff.list.z) {
     tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #

### 2024-01-06 paper_results.Rmd
cleaning up main analyses

@@
 
+Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.
+
 # Setup
+
 ```{r setup, include=FALSE}
@@
 
-modifying to test w/in feature rather than mean centiles
-
-```{r}
-sex.bias.feat.t.tests <- function(df, feature_list=pheno_list, comp_multiplier=1){
+modifying to test for sex-biases w/in feature rather than mean centiles
+
+```{r}
+pheno_diff_list <- paste0("diff_", pheno_list)
+sex.bias.feat.t.tests <- function(df, feature_list=pheno_diff_list, comp_multiplier=1){
 
@@
   for (pheno in feature_list) {
+  #print(paste("t-test:", pheno))
+
   #conduct t test
@@
     group_by(dataset) %>%
-    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
+    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
+              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
+              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
+              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
     ungroup()
-  df.sex.t$pheno <- pheno
+
+  df.sex.t$pheno <- sub("diff_", "", pheno)
+
+  #append to full df
   t.df <- rbind(t.df, df.sex.t)
   }
+  detach(df)
+
+  #also get medians for m and f
+  med_df <- df %>%
+  dplyr::group_by(dataset, sex) %>%
+    summarise_at(feature_list, median) %>%
+    ungroup() %>%
+    pivot_longer(cols = starts_with("diff_"), names_to="pheno", names_prefix="diff_") %>%
+    pivot_wider(names_from = sex, values_from=value, names_glue = "med_{sex}") %>%
+    mutate(median_sex_diff = (med_Male - med_Female)) %>%
+    ungroup()
+
+  #combine feature dfs
+  result_df <- left_join(t.df, med_df)
 
@@
   n_comp=length(unique(df$dataset))*length(feature_list)
-  df.sex.t <- t.df %>%
+  result_df <- result_df %>%
     dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
@@
 
-  #summarize
-  df.sex.t.sum <- df.sex.t %>%
-    group_by(dataset) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-
-  return(df.sex.t.sum)
-  #detach()
-}
-
-# tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
-# rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
-```
-
-trying to get interpretable effect sizes w t.test (rather than ranked t test)
-
-```{r}
-sex.bias.feat.t.tests_2 <- function(df, feature_list=pheno_list, comp_multiplier=1){
-
-  #initialize empty df to store outputs
-  t.df <- data.frame("dataset" = character(),
-                     "pheno" = character(),
-                     "p.value" = double(),
-                     "sex_diff" = double(),
-                     "t.stat" = double(),
-                     "df" = double())
-  attach(df)
-  for (pheno in feature_list) {
-  #conduct t test
-  df.sex.t <- df %>%
-    group_by(dataset) %>%
-    summarise(p.value = tidy(t.test(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(t.test(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(t.test(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(t.test(formula=as.formula(paste0("diff_",pheno, " ~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
-    ungroup()
-  df.sex.t$pheno <- pheno
-  t.df <- rbind(t.df, df.sex.t)
-  }
-
-  #apply fdr correction
-  n_comp=length(unique(df$dataset))*length(feature_list)
-  df.sex.t <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           dataset=as.factor(dataset))
-
-  return(df.sex.t)
-  #detach()
-}
-
-# tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
-# rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
-```
+  # #summarize
+  # df.sex.t.sum <- df.sex.t %>%
+  #   group_by(dataset) %>%
+  #   dplyr::summarize(n_sig_regions = sum(sig_fdr),
+  #             min_pval = min(p.val_fdr),
+  #             max_pval = max(p.val_fdr)) %>%
+  #   ungroup() %>%
+  #   as.data.frame()
+
+  return(result_df)
+
+}
+# sex.bias.feat.t.tests(diffs.df, feature_list = global.diff)
+```
+
+Test for linear relationships between subject's actual centile score (calculated from raw data) and how much that score is affected by different combat methods.
 
@@
 
+Get differences in male vs female predicted variances at mean sample age
+
+```{r}
+pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
+combined_var <- data.frame()
+
+for (file in pred.csvs) {
+  # Read each CSV file
+  data <- fread(file)
+
+  # Add a "Source_File" column with the file name
+  data <- data %>%
+    mutate(Source_File = as.factor(basename(file))) %>%
+    mutate(dataset = sub("_data_variance.csv", "", Source_File),
+           sex_standard = sex_effect/f.var)
+
+  # Bind the data to the combined dataframe
+  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
+}
+
+#get summary df for p vals, plotting
+sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
+  dplyr::select(-V1) %>%
+  mutate(dataset = gsub("-data", "", dataset),
+         dataset = gsub("-", "_", dataset),
+         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
+  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
+  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
+
+sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
+
+plot.info <- sigma.sex.df %>%
+  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
+  unique() %>%
+  mutate(dataset = gsub("-", "_", dataset),
+         dataset = gsub("_simsites", "", dataset))
+
+combined_var <- inner_join(combined_var, plot.info)
+```
+
+## SCALE EFFECTS OF SEX
+
+```{r}
+#var plot
+var.plot.df <- combined_var %>%
+  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset, pheno_cat)
+
+tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="ukb_CN")
+tp$sex_standard <- 1
+
+var.plot <- ggplot(var.plot.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = sex_standard)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
+#var.plot
+
+#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
+```
+
+```{r}
+#plot global
+var.plot.global.df <- combined_var %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "WM",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA)),
+         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
+  group_by(dataset)
+
+tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="ukb_CN")
+tp$sex_standard <- 1
+
+var.plot.global <- ggplot(var.plot.global.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = sex_standard),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
+#var.plot.global
+```
+
+Combine plots
+```{r}
+p <- ggarrange(
+  var.plot.global, var.plot,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.1),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+
+p_annotate <- annotate_figure(p, top = text_grob("Figure 1B. Standardized effect of male sex on feature variance", 
+               color = "black", size = 14, hjust=.65))
+p_annotate
+ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
+```
+
 ## CENTILES
@@
 Do centile scores change based on combat version?
-Foregoing anova based on emails with andrew.
+Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)
+
 Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
@@
 Little plot
-```{r}
+```{r, eval=FALSE}
 #get means w/in feature 
@@
       label = sub("_[^_]*_", "_", pheno))
-```
-
-```{r}
+
 dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
@@
   labs(y="Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Figure 2. Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
+  ggtitle("Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
 cent.diffs.plot
@@
 
-Visualize magnitude mean w/in subj error
+Playing with different ways to visualize each subject's average centile error (i.e., raw data centile - combatted data centile) and magnitude centile error (absolute value)
 
@@
 
+#look at sex distributions of ppts across sites
 subj_sum.df %>%
   ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))
+  geom_bar(aes(x=sim.site, fill=sex))+
+  ggtitle("Distributions of male and female subjects across simulated sites")
 ```
@@
 
-Sig test w/in feature diffs
+Keeping these analyses in case I want to use one of the above violin/boxplots that show w/in subject averages (across features).
+However, while these results I think are good to have from a "be aware of how biological variables are distributed across your study sites" perspective, it doesnt really say anything interesting froma "combat methods" perspective.
+
+Testing for significant M:F differences in centile erros WITHIN each feature!
 
 check for skew (welch t.test vs ranked welch t test)
-```{r}
+```{r, eval=FALSE}
 for (pheno in pheno_list){
@@
       geom_histogram(alpha=0.5, position="identity") +
-    ggtitle(val)
+    ggtitle(val) +
+    facet_wrap(~dataset, scales="free")
 print(p)
@@
 
-```{r}
-sex.bias.feat.t.tests(diffs.df)
-```
-
-Sig test w/in feature diffs
+not normally distributed, use nonparametric t test of ranks with welch correction for different variances.
+Also calculating the median centile errors of males and females, since this seems like a more intuitive thing to plot than effect size (since the effect size is in units of rank, not actually centiles)
+
 ```{r, warning=FALSE}
-sex_t_tests_in_feat.df <- sex.bias.feat.t.tests_2(diffs.df)
-```
-
-```{r}
-sex_t_tests_in_feat.df.sum <- sex_t_tests_in_feat.df %>%
+sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df)
+
+#also get summary
+sex_t_tests_in_feat.df %>%
     group_by(dataset) %>%
@@
 sex_t_tests_to_plot <- left_join(sex_t_tests_in_feat.df2, dk.parc, by="label")
+
+# max(sex_t_tests_to_plot$median_sex_diff)*100
+# min(sex_t_tests_to_plot$median_sex_diff)*100
 ```
@@
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ sex_diff)) %>%
+         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
   group_by(dataset, pheno_cat)
@@
 sex.var.plot <- ggplot(sex.var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
             ymin = -Inf, ymax = Inf, alpha = 0.4) +
@@
              position = position_brain(hemi ~ side),
-             aes(fill = sex_diff)) +
+             aes(fill = median_sex_diff)) +
   #scale_fill_gradient2() +
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.003, 0.003))
-sex.var.plot
-
-#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
+# sex.var.plot
 ```
@@
                                             TRUE ~ NA)),
-        sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ sex_diff)) %>% #drop NS values
+        median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>% #drop NS values
   group_by(dataset)
@@
   ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = sex_diff),
+             aes(fill = median_sex_diff),
              side = c("coronal")) +
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.003, 0.003))
-sex.var.plot.global
-```
-
-Combine plots
-```{r}
-p <- ggarrange(
-  var.plot.global, var.plot,
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
+# sex.var.plot.global
+```
+
+Combine plots - need to keep playing around to make the spacing pretty when it saves out
+```{r}
+p_sex <- ggarrange(
+  sex.var.plot.global, sex.var.plot,
   ncol=2, nrow=1, 
@@
 
-p_annotate <- annotate_figure(p, top = text_grob("Figure 1. Standardized effect of male sex on feature variance", 
+p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
                color = "black", size = 14, hjust=.65))
-
-ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
-```
+p_sexannotate
+ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
+```
+
+Sex Effect is the median male centile error - median female centile error (i.e.
+units are in centiles).
+Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)?
+Honestly, not the most straightforward interpretation, i know - may make more sense to just plot the median errors separately in males and females than to plot the differences between the two.
 
@@
 
+The relationship between subjects' centile score and how much their centile score is altered by combat is complex and depends on sex and site.
+I think that rather than trying to investigate this directly, it'll be more interesting/effective to show visually with the secondary analyses where I vary M:F ratios.
+
 ## Z-SCORES
+
+*come back to this*
 
@@
 
-## SCALE EFFECTS OF SEX
-
-fix age, get diff between male and female pred. variance
-
-```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
-combined_var <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_variance.csv", "", Source_File),
-           sex_standard = sex_effect/f.var)
-
-  # Bind the data to the combined dataframe
-  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
-}
-
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(dataset = gsub("-data", "", dataset),
-         dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
-
-sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
-
-plot.info <- sigma.sex.df %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
-  unique() %>%
-  mutate(dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset))
-
-combined_var <- inner_join(combined_var, plot.info)
-```
-
-```{r}
-#var plot
-var.plot.df <- combined_var %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot <- ggplot(var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = sex_standard)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-var.plot
-
-#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
-```
-
-```{r}
-#plot global
-var.plot.global.df <- combined_var %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
-  group_by(dataset)
-
-tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot.global <- ggplot(var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = sex_standard),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-var.plot.global
-```
-
-Combine plots
-```{r}
-p <- ggarrange(
-  var.plot.global, var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_annotate <- annotate_figure(p, top = text_grob("Figure 1. Standardized effect of male sex on feature variance", 
-               color = "black", size = 14, hjust=.65))
-
-ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
-```
-
 ## CENTILE FANS
@@
 }
-length(basic_centiles)
+# length(basic_centiles)
 centiles.df <- bind_rows(basic_centiles)
 #skimr::skim(centiles.df)
-names(centiles.df)
-```
-
-```{r}
+# names(centiles.df)
+```
+
+just work with global volumes to visualize for now
+
+```{r}
+#get raw data points for visualization
 global.df.long <- ukb_cn_df %>%
@@
 ```
-
-just work with global volumes for now
 
@@
 
-```{r}
+adding jitter to age (x-axis) because otherwise frankly it's hard to look at
+
+```{r}
+dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+
 ggplot(wmv.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$WMV), alpha=0.7, width=150, color="lightgray") +
+  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$WMV), alpha=0.4, width=150, color="lightgray") +
   #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
@@
   facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "dashed")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
   theme_minimal() +
+  xlab("Age (days)") + ylab("Volume (mm3)") +
   geom_text_repel(aes(x = max(ukb_cn_df$age_days), y = min(centile_1)),
@@
                   na.rm = TRUE) + 
-  labs(title="WMV brain charts")
-```
+  labs(title="Fig 2A. WM volume brain charts")
+```
+
+*need to add labeling of % lines - ask Ayan how he did it*
+also not sure how much i should worry about the super big female outlier
+
+Look at other phenotypes
 
@@
   facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846")) +
+  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
+  theme_minimal() +
+  xlab("Age (days)") + ylab("Volume (mm3)") +
   theme_minimal() +
@@
 ```
-```{r}
-#wm centile lines
+Random cortical feature
+
+```{r}
 r.sa.pc.df <- centiles.df %>%
@@
 ```{r}
-perm_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm"
-raw_files <- list.files(path = perm_data_path, pattern = "_diffs.csv", full.names = TRUE)
+raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_diffs.csv", full.names = TRUE)
 
@@
 
-Read in subject-wise summary stats
-
-```{r}
-raw_files <- list.files(path = perm_data_path, pattern = "_subj_pred.csv", full.names = TRUE)
+Get errors
+
+```{r}
+raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
 
@@
     data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df_list <- c(df_list, list(data))
+    assign("diffs_subj.list", df_list)
+}
+length(diffs_subj.list)
+diffs_subj.df <- bind_rows(diffs_subj.list)
+```
+
+Read in subject-wise summary stats
+
+```{r}
+raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_subj_pred.csv", full.names = TRUE)
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  # Read each CSV file
+    data <- fread(file)
+
+    data <- data %>%
+      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
 
@@
 
-## CENTILES
-Do centile scores change based on combat version?
-
 ```{r}
@@
 pheno_list.z <- paste0(pheno_list, ".z")
-
-# combined_preds.p <- data.frame()
-get.predictions <- function(x){
-  df <- data.frame() #new empty dataframe
-  pred.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_predictions.csv"), full.names = TRUE)
-  for (file in pred.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_predictions.csv|perm-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_predictions <- mclapply(n_perm_list, get.predictions, mc.preschedule = FALSE)
-```
-
-```{r}
-get.abs.diffs <- function(x){
-  df <- x %>%
-    mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-    mutate(dataset = relevel(dataset, ref= "raw")) %>%
-    arrange(dataset) %>%
-    group_by(participant) %>%
-    dplyr::mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #centile diffs
-    dplyr::select(-any_of(pheno_list)) %>% #drop
-    dplyr::mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #z-score diffs
-    dplyr::select(-any_of(pheno_list.z)) %>% #drop
-    ungroup() %>%
-    dplyr::filter(dataset != "raw") #drop raw
-  return(df)
-}
-
-```
-
-```{r}
-perm_diffs <- mclapply(perm_predictions, get.abs.diffs, mc.preschedule = FALSE)
-saveRDS(perm_diffs)
-```
-
-```{r}
-perm_diffs <- readRDS()
-```
-
-add ANOVA
-```{r}
-
-```
+```
+
+## CENTILES
+
+Do each feature's centile scores change based on combat version?
 
@@
 #apply
-perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
+perm_t.tests <- mclapply(diffs_subj.list, centile.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_subj.list))
 names(perm_t.tests) <- n_perm_list
@@
 fwrite(cent_t.tests.p, file="derivatives/ukb_perm_csvs/centile_t.tests.csv")
+
+centile.t.tests(perm_subj_list[[1]])
 ```

### 2024-01-06 paper_main_results.Rmd
splitting out big rmd file

--- 2024-01-06 paper_results.Rmd
@@
 # summary(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
+```
+
+plot w/in feature average centile errors
+
+```{r}
+#define plotting
+plot.cent.diffs <- function(df){
+mean.diffs.long <- df %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup() %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+cent.diffs.plot <- mean.diffs.long %>%
+  dplyr::filter(dataset != "raw") %>%
+  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
+  theme_linedraw(base_size = 11) +
+  facet_wrap(~pheno_cat) +
+  scale_y_continuous(labels = scales::percent) + 
+  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
+  ggtitle("Absolute errors in centile scores across ComBat processing pipelines")
+
+print(cent.diffs.plot)
+}
 ```
@@
   geom_smooth() +
+  facet_grid(~dataset, scales="free_y")
+
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth() +
   facet_grid(dataset ~ sim.site, scales="free_y")
 ```
+
+test to see if slope of non-dominant group is nonzero!!!
 
@@
 cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-fwrite(cent_t.tests.p, file="derivatives/ukb_perm_csvs/centile_t.tests.csv")
-
-centile.t.tests(perm_subj_list[[1]])
+fwrite(cent_t.tests.p, file="derivatives/ukb_perm/centile_t.tests.csv")
 ```
@@
 Little plots
-```{r}
-#define plotting
-plot.cent.diffs <- function(df){
-mean.diffs.long <- df %>%
-  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "raw") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
+
+```{r}
+lapply(diffs_subj.list, plot.cent.diffs)
+```
+
+W/in subj average plots
+
+```{r}
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+cent.diffs.v.plot <- perm_subj.df %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
   ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
-
-print(cent.diffs.plot)
-}
-```
-
-```{r}
-lapply(perm_diffs, plot.cent.diffs)
-```
-
-W/in subj average plots
-```{r}
-mean.subj.centile.err <- function(df){
-mean.diffs.subj <- df %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  rowwise() %>%
-  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))
-
-return(mean.diffs.subj)
-#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
-}
-perm_cent.err <- mclapply(perm_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
-names(perm_cent.err) <- n_perm_list
-cent_err.p <- bind_rows(perm_cent.err, .id = "permutation")
-
-#write out for plotting later
-saveRDS(cent_err.p, "ukb_perm_csvs/centile_errors.RDS")
-```
-
-```{r}
-cent.diffs.v.plot <- cent_err.p %>%
-  mutate(mean = mean*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=dataset, y=mean, fill=permutation)) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
+  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=perm), scale="count") +
+  labs(y= "Mean Magnitude Centile Error", x = "ComBat Configurations") +
+  scale_x_discrete(labels=c(dataset_names)) +
+  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
+  ggtitle("Fig 1B. Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
+cent.diffs.v.plot
 ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)

### 2024-01-07 paper_lbcc_results.Rmd
cleaning up rmds, prepping to run sex-bias t tests on cubic

--- 2024-01-06 paper_main_results.Rmd
@@
 
-# Main Analyses
-
-## Load Data
-
-pulled variance from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_variance.csv' .`
-
-Get predicted centiles
-- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' .`
-
-```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_predictions.csv", full.names = TRUE)
-combined_preds <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
-
-  # Bind the data to the combined dataframe
-  combined_preds <- bind_rows(combined_preds, data)
-}
-```
-
-**qc - find ppts with centiles at 1 or 0**
-
-```{r}
-combined_preds %>% 
-  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0))) %>%
-  dplyr::select(participant) %>%
-  unique()
-
-combined_preds %>% 
-  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
-  dplyr::select(participant) %>%
-  unique()
-
-# test.df <- combined_preds %>% 
-#   dplyr::select(pheno_list)
-# 
-# test.df[test.df==1] <- 0
-# 
-# format(round(max(do.call(pmax, test.df)), 20), nsmall = 20)
-```
-
-saving out csv forces some centile scores to round up to 1, BUT it appears z scores are still being calculated correctly so just going to proceed!
-
-Get errors
-- pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
-
-diffs.df <- fread(file = raw_files, sep=",") %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-```
-
-Get subject-wise summaries
-- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_subj_pred.csv", full.names = TRUE) %>% unlist()
-
-subj_sum.df <- fread(file = raw_files, sep=",") %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-```
-
-Get differences in male vs female predicted variances at mean sample age
-
-```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
-combined_var <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_variance.csv", "", Source_File),
-           sex_standard = sex_effect/f.var)
-
-  # Bind the data to the combined dataframe
-  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
-}
-
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(dataset = gsub("-data", "", dataset),
-         dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
-
-sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
-
-plot.info <- sigma.sex.df %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
-  unique() %>%
-  mutate(dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset))
-
-combined_var <- inner_join(combined_var, plot.info)
-```
-
-## SCALE EFFECTS OF SEX
-
-```{r}
-#var plot
-var.plot.df <- combined_var %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot <- ggplot(var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = sex_standard)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-#var.plot
-
-#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
-```
-
-```{r}
-#plot global
-var.plot.global.df <- combined_var %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
-  group_by(dataset)
-
-tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot.global <- ggplot(var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = sex_standard),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-#var.plot.global
-```
-
-Combine plots
-```{r}
-p <- ggarrange(
-  var.plot.global, var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_annotate <- annotate_figure(p, top = text_grob("Figure 1B. Standardized effect of male sex on feature variance", 
-               color = "black", size = 14, hjust=.65))
-p_annotate
-ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
-```
-
-## CENTILES
-
-Do centile scores change based on combat version?
-Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)
-
-Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
-(see Zimmerman & Zumbo, 1993).
-Using custom functions from `ranked_welch_tests.R`.
-
-```{r}
-centile.t.tests(diffs.df)
-```
-
-Little plot
-```{r, eval=FALSE}
-#get means w/in feature 
-mean.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
-cent.diffs.plot
-
-ggsave("figs/win_feature_mean_cent_abs.diffs.jpeg", cent.diffs.plot)
-```
-
-### Sex-Biases
-
-Playing with different ways to visualize each subject's average centile error (i.e., raw data centile - combatted data centile) and magnitude centile error (absolute value)
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=sex, y=mean_cent_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  facet_wrap(~dataset) +
-  theme(legend.position="none") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
-  ggplot(aes(x=dataset, y=mean_cent_diff, fill=sex)) +
-  geom_violin(trim=FALSE, scale="width") +
-  labs(y="Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot(aes(x=sim.site, y=mean_cent_abs.diff, fill=sex)) +
-  geom_violin(trim=FALSE) +
-  facet_wrap(~dataset) +
-  labs(y="Magnitude Centile Error", x = "Simulated Site") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in centile scores across \nComBat configurations")
-
-#look at sex distributions of ppts across sites
-subj_sum.df %>%
-  ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))+
-  ggtitle("Distributions of male and female subjects across simulated sites")
-```
-
-Significance testing (w/in sub. mean abs centile error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
-
-```{r}
-sex.bias.t.tests(subj_sum.df)
-sex.bias.t.tests(subj_sum.df, to_test="mean_cent_diff")
-```
-
-Keeping these analyses in case I want to use one of the above violin/boxplots that show w/in subject averages (across features).
-However, while these results I think are good to have from a "be aware of how biological variables are distributed across your study sites" perspective, it doesnt really say anything interesting froma "combat methods" perspective.
-
-Testing for significant M:F differences in centile erros WITHIN each feature!
-
-check for skew (welch t.test vs ranked welch t test)
-```{r, eval=FALSE}
-for (pheno in pheno_list){
-  val <- paste0("diff_", pheno)
-  p <- ggplot(diffs.df, aes(x = diffs.df[[val]], fill= sex, color=sex)) +
-      geom_histogram(alpha=0.5, position="identity") +
-    ggtitle(val) +
-    facet_wrap(~dataset, scales="free")
-print(p)
-}
-```
-
-not normally distributed, use nonparametric t test of ranks with welch correction for different variances.
-Also calculating the median centile errors of males and females, since this seems like a more intuitive thing to plot than effect size (since the effect size is in units of rank, not actually centiles)
-
-```{r, warning=FALSE}
-sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df)
-
-#also get summary
-sex_t_tests_in_feat.df %>%
-    group_by(dataset) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-```
-
-```{r}
-sex_t_tests_in_feat.df2 <- sex_t_tests_in_feat.df %>%
-  mutate(pheno_cat = as.factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_)),
-    label = sub("_[^_]*_", "_", pheno)) #for plotting
-sex_t_tests_to_plot <- left_join(sex_t_tests_in_feat.df2, dk.parc, by="label")
-
-# max(sex_t_tests_to_plot$median_sex_diff)*100
-# min(sex_t_tests_to_plot$median_sex_diff)*100
-```
-
-plot w/in feature t tests
-
-```{r}
-#var plot
-sex.var.plot.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_diff <- 1
-
-sex.var.plot <- ggplot(sex.var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = median_sex_diff)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
-# sex.var.plot
-```
-
-```{r}
-#plot global
-sex.var.plot.global.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-        median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>% #drop NS values
-  group_by(dataset)
-
-tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = median_sex_diff),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
-# sex.var.plot.global
-```
-
-Combine plots - need to keep playing around to make the spacing pretty when it saves out
-```{r}
-p_sex <- ggarrange(
-  sex.var.plot.global, sex.var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
-               color = "black", size = 14, hjust=.65))
-p_sexannotate
-ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
-```
-
-Sex Effect is the median male centile error - median female centile error (i.e.
-units are in centiles).
-Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)?
-Honestly, not the most straightforward interpretation, i know - may make more sense to just plot the median errors separately in males and females than to plot the differences between the two.
-
-### Extremeness-Biases
-
-Group at 0-0.25, >0.25 - .75 <, 0.75-1
-
-```{r}
-subj_sum.df <- subj_sum.df %>%
-  mutate(cent_cat = factor(case_when(mean_centile <= 0.25 ~ "low",
-                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
-                              mean_centile >= 0.75 ~ "high"), levels= c("low", "mid", "high"), ordered=TRUE),
-         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
-                              mean_z > -1 & mean_z < 1 ~ "mid",
-                              mean_z >= 1 ~ "high")))
-```
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(dataset ~ sim.site, scales="free_y")
-```
-
-test to see if slope of non-dominant group is nonzero!!!
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff, color=sex)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  #dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-The relationship between subjects' centile score and how much their centile score is altered by combat is complex and depends on sex and site.
-I think that rather than trying to investigate this directly, it'll be more interesting/effective to show visually with the secondary analyses where I vary M:F ratios.
-
-## Z-SCORES
-
-*come back to this*
-
-look for skew in distribution
-
-significance testing
-
-```{r}
-#initalize empty dfs to store outputs
-# z.sum.df <- data.frame(pheno = character(),
-#                        term = character(),
-#                        df = double(),
-#                        sumsq = double(),
-#                        meansq = double(),
-#                        statistic = double(),
-#                        p.value = double())
-z.t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-
-#run tests
-attach(diffs.df)
-for (pheno in z.pheno_abs.diff.list) {
-  # df.aov <- tidy(kruskal.test(diffs.df[[pheno]] ~ diffs.df[["dataset"]]))
-  # df.aov$pheno <- pheno
-  # z.sum.df <- rbind(z.sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=diffs.df[[pheno]], g=diffs.df[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  z.t.df <- rbind(z.t.df, df.pairwise.t)
-}
-detach()
-
-#apply fdr correction
-# z.sum.df <- z.sum.df %>%
-#   dplyr::filter(term != "Residuals") %>%
-#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-z.t.df <- z.t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno),
-         pheno = sub(".z", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-sum_z.t.df <- z.t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min.pval = min(p.val_fdr),
-            max.pval = max(p.val_fdr))
-
-sum_z.t.df
-```
-
-Little plot
-```{r}
-#get means w/in feature 
-z.mean.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno = gsub(".z", "", pheno)) %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-```
-
-```{r}
-z.diffs.plot <- z.mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  ggplot() +
-  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
-  ggtitle("Figure 2. Absolute errors in z-scores of brain structure features across ComBat processing pipelines")
-z.diffs.plot
-
-ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", z.diffs.plot)
-```
-
-### Sex-Biases
-
-Visualize magnitude mean w/in subj error
-
-```{r}
-subj_sum.df %>%
-  ggplot() +
-  geom_violin(aes(x=dataset, y=mean_z_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Z-Score Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot(aes(x=sex, y=mean_z_diff, fill=sex)) +
-  geom_boxplot(trim=FALSE, scale="width") +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y="Z-Score Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=sim.site, y=mean_z_abs.diff, fill=sex)) +
-  geom_violin(trim=FALSE) +
-  facet_wrap(~dataset) +
-  labs(y="Magnitude Z Error", x = "Simulated Site") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))
-```
-
-Significance testing (w/in sub. mean abs z error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
-
-```{r}
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_abs.diff")
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_diff")
-```
-
-## CENTILE FANS
-
-pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_centiles.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("-data", "", dataset))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("basic_centiles", df_list)
-}
-# length(basic_centiles)
-centiles.df <- bind_rows(basic_centiles)
-#skimr::skim(centiles.df)
-# names(centiles.df)
-```
-
-just work with global volumes to visualize for now
-
-```{r}
-#get raw data points for visualization
-global.df.long <- ukb_cn_df %>%
-  dplyr::select(participant, age_days, sex, vol_list_global) %>%
-  pivot_longer(cols = vol_list_global, values_to = "value", names_to="pheno")
-```
-
-```{r}
-global.cent.df <- centiles.df %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('-refA', dataset))
-
-for (feature in unique(global.cent.df$pheno)) {
-
-  df <- global.cent.df %>%
-    dplyr::filter(pheno == feature)
-
- plot <- ggplot() +
-      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_9, color=df$dataset)) +
-      labs(title=paste("Male", feature, "brain chart"))
-
- print(plot)
-
- plot <- ggplot() +
-      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_9, color=df$dataset)) +
-      labs(title=paste("Female", feature, "brain chart"))
-
- print(plot)
-
- plot <- ggplot() +
-      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_9, color=df$dataset)) +
-      labs(title=paste(feature, "brain chart"))
-
- print(plot)
-
-}
-```
-
-Looking at WMV more specifically for visualizations
-```{r}
-#wm centile lines
-wmv.df <- centiles.df %>%
-  dplyr::filter(pheno == "WMV") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-```
-
-adding jitter to age (x-axis) because otherwise frankly it's hard to look at
-
-```{r}
-dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-
-ggplot(wmv.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$WMV), alpha=0.4, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  theme_minimal() +
-  xlab("Age (days)") + ylab("Volume (mm3)") +
-  geom_text_repel(aes(x = max(ukb_cn_df$age_days), y = min(centile_1)),
-                  label="test",
-                  nudge_x = 1,
-                  na.rm = TRUE) + 
-  labs(title="Fig 2A. WM volume brain charts")
-```
-
-*need to add labeling of % lines - ask Ayan how he did it*
-also not sure how much i should worry about the super big female outlier
-
-Look at other phenotypes
-
-```{r}
-#looking at other phenotypes
-ven.df <- centiles.df %>%
-  dplyr::filter(pheno == "Ventricles") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-
-ggplot(ven.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$Ventricles), alpha=0.7, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  theme_minimal() +
-  xlab("Age (days)") + ylab("Volume (mm3)") +
-  theme_minimal() +
-  labs(title="sGMV brain charts")
-```
-Random cortical feature
-
-```{r}
-r.sa.pc.df <- centiles.df %>%
-  dplyr::filter(pheno == "rh_SA_postcentral") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-
-ggplot(r.sa.pc.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$rh_SA_postcentral), alpha=0.7, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  theme_minimal() +
-  xlab("Age (days)") + ylab("Volume (mm3)") +
-  theme_minimal() +
-  labs(title="SA of right postcentral brain charts")
-```
-
-# Replication
-
-## Load Data
-
-pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*_diffs.csv' .`
-
-pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*.csv' .`
-
-Read in centile/z-score errors
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_diffs.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("perm_list", df_list)
-}
-length(perm_list)
-perm.df <- bind_rows(perm_list)
-```
-
-Get errors
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("diffs_subj.list", df_list)
-}
-length(diffs_subj.list)
-diffs_subj.df <- bind_rows(diffs_subj.list)
-```
-
-Read in subject-wise summary stats
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_subj_pred.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("perm_subj_list", df_list)
-}
-length(perm_subj_list)
-perm_subj.df <- bind_rows(perm_subj_list)
-```
-
-```{r}
-n_perm_list <- lapply(c(1:10), str_pad, 3, pad = "0")
-n_perm_list <- paste0("perm-", n_perm_list)
-pheno_list.z <- paste0(pheno_list, ".z")
-```
-
-## CENTILES
-
-Do each feature's centile scores change based on combat version?
-
-```{r}
-#apply
-perm_t.tests <- mclapply(diffs_subj.list, centile.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_subj.list))
-names(perm_t.tests) <- n_perm_list
-cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-fwrite(cent_t.tests.p, file="derivatives/ukb_perm/centile_t.tests.csv")
-```
-
-```{r}
-cent_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
-
-Little plots
-
-```{r}
-lapply(diffs_subj.list, plot.cent.diffs)
-```
-
-W/in subj average plots
-
-```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-cent.diffs.v.plot <- perm_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=perm), scale="count") +
-  labs(y= "Mean Magnitude Centile Error", x = "ComBat Configurations") +
-  scale_x_discrete(labels=c(dataset_names)) +
-  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
-  ggtitle("Fig 1B. Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
-cent.diffs.v.plot
-ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
-```
-
-## Z-SCORES
-
-*need to look at distribution to check for skew!*
-
-```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs 
-pheno_abs.diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
-
-z.t.tests <- function(df){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in pheno_abs.diff.list.z) {
-    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
-names(perm_t.tests_z) <- n_perm_list
-z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
-```
-
-```{r}
-fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
-```
-
-```{r}
-z_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
-
-## SCALE EFFECTS OF SEX
-fix age, get diff between male and female pred. variance
-```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var,
-             pheno = as.character(pheno))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-perm_var.all <- bind_rows(perm_var, .id = "column_label")
-```
-
-maybe better to get sex_summary.csv - just need significance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("derivatives/ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
-         dataset = gsub("-data", "", dataset),
-         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
-         dataset = gsub("-", "", dataset),
-         #dataset = as.character(dataset),
-         pheno = as.character(pheno)) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
-         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
-  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
-
-perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
-```
-
-```{r}
-  sex_effect_plot <- perm_var_info %>%
-    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
-                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
-    ggplot() +
-    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
-    facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance")) +
-    theme(axis.text.x = element_blank())
-
-    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
-    print(sex_effect_plot)
-```
-
-# Secondary: Vary M:F
-
-```{r}
-n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
-```
-
-## Load Data
-
-pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_diffs.csv' .`
-
-pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' .`
-
-Read in centile/z-score errors
-
-```{r}
-ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
-raw_files <- list.files(path = ratio_data_path, pattern = "_diffs.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("ratio_list", df_list)
-}
-length(ratio_list)
-ratio.df <- bind_rows(ratio_list)
-
-# #sum(is.na(ratio.df))
-# na.df <- ratio.df[rowSums(is.na(ratio.df)) > 0,]
-# skimr::skim(na.df)
-# 
-# names(ratio.df)
-ratio.df[is.na(ratio.df$diff_lh_CT_cuneus.z),] %>%
-  dplyr::select(contains("_cuneus"), "participant", "dataset")
-```
-
-Read in subject-wise summary stats
-
-```{r}
-raw_files <- list.files(path = ratio_data_path, pattern = "_subj_pred.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("ratio_subj_list", df_list)
-}
-length(ratio_subj_list)
-ratio_subj.df <- bind_rows(ratio_subj_list)
-
-max(ratio_subj.df$mean_centile)
-min(ratio_subj.df$mean_centile)
-max(ratio_subj.df$mean_z)
-min(ratio_subj.df$mean_z)
-```
-
-```{r}
-for (l in unique(ratio_subj.df$dataset)){
-plot <- ratio_subj.df %>%
-  dplyr::filter(dataset == l) %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(sim.site ~ prop) +
-  ggtitle(l)
-print(plot)
-}
-```
-
-Do centile scores change based on combat version?
-
-need to look and see what happens for subjects who aren't in all samples?
-
-Plot boxplots to compare across prop.
-
-```{r, eval=FALSE}
-#too much memory to run!
-ratio.df %>%
-  dplyr::select(sim.site, prop, dataset, starts_with("abs.diff"), !ends_with(".z")) %>%
-  pivot_longer(cols= starts_with("abs.diff"), names_to="pheno", nammes_prefix="abs.diff_", values_to="cent_abs.diff") %>%
-  mutate(cent_abs.diff = cent_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=cent_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-Plot w/in subj means
-
-```{r}
-ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-
-ratio_subj.df %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=mean_z_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Z-score Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-W/in subj average plots
-
-```{r}
-cent.diffs.v.plot.r <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=dataset), scale="width") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-cent.diffs.v.plot.r
-ggsave(filename="figs/ratio_centile_violin.png", cent.diffs.v.plot.r)
-```
-
-**Stats Tests**
-Within each M:F proportion simulated, is the magnitude of subjects' mean centile errors differ significantly by combat configuration?
-
-```{r}
-prop_abs.cent_t.tests <- mclapply(ratio_list, centile.t.tests, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
-names(prop_abs.cent_t.tests) <- n_prop_list
-prop_abs.cent_t.tests_df <- bind_rows(prop_abs.cent_t.tests, .id = "column_label")
-```
-
-Visalize
-```{r}
-ggplot(prop_abs.cent_t.tests_df) +
-  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
-  geom_hline(yintercept = 208, linetype='longdash') +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
-  facet_wrap(~comp)
-```
-
-## Sex-Biases
-
-```{r}
-ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=sex), scale="width") +
-  facet_wrap(~dataset) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-```
-
-```{r}
-cent.diffs.v.boxplot.r_sex_free <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-cent.diffs.v.boxplot.r_sex_free
-ggsave(filename="figs/ratio_centile_box_sex_free.png", cent.diffs.v.boxplot.r_sex_free)
-```
-
-```{r}
-#trying to add significance testing
-abs.cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, prop) %>%
-  wilcox_test(mean_cent_abs.diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance() %>%
-  add_xy_position(x = "prop")
-
-abs.cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  droplevels() %>%
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
-    #hide.ns = TRUE
-    bracket.nudge.y = -0.5
-    ) +
-  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_abs.centile_box_sex.png", abs.cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-eval diffs (not abs. diffs)
-
-```{r}
-#trying to add significance testing
-cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, prop) %>%
-  wilcox_test(mean_cent_diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance() %>%
-  add_xy_position(x = "prop")
-
-cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
-  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  droplevels() %>%
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_diff, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
-    #hide.ns = TRUE
-    bracket.nudge.y = 4.5
-    ) +
-  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_centile_box_sex.png", cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-### Significance Testing
-
-#### CENTILES
-within each M:F prop, test male vs female subject's mean abs. centile diff
-```{r}
-sex_bias_ratio_abs.tests <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_abs.tests) <- n_prop_list
-sex_bias_ratio_abs.tests_df <- bind_rows(sex_bias_ratio_abs.tests, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_abs.tests_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-test actual (not abs. val) centile errors
-
-```{r}
-sex_bias_ratio_tests <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test = "mean_cent_diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_tests) <- n_prop_list
-sex_bias_ratio_tests_df <- bind_rows(sex_bias_ratio_tests, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_tests_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-#### Z-SCORES
-
-```{r}
-sex_bias_ratio_abs.tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_abs.diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_abs.tests.z) <- n_prop_list
-sex_bias_ratio_abs.tests.z_df <- bind_rows(sex_bias_ratio_abs.tests.z, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_abs.tests.z_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-```{r}
-sex_bias_ratio_tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_tests.z) <- n_prop_list
-sex_bias_ratio_tests.z_df <- bind_rows(sex_bias_ratio_tests.z, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_tests.z_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-## Extremeness-Biases
-Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)
-
-#### CENTILES
-
-test w/in subj avg centile ~ w/in subj avg centile error
-
-```{r}
-#testing
-test.lm <- lm(mean_cent_diff ~ mean_centile*as.factor(dataset), data=ratio_subj_list[[1]])
-tidy(test.lm)
-interactions::interact_plot(test.lm, pred = mean_centile, modx = dataset)
-
-ggplot(ratio_subj_list[[4]], aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-```{r}
-ext.bias_test.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ mean_cent_diff", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.ln) <- n_prop_list
-linear <- as.data.frame(bind_rows(ext.bias_test.ln, .id = "prop"))
-
-ext.bias_test.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.sq) <- n_prop_list
-square <- as.data.frame(bind_rows(ext.bias_test.sq, .id = "prop"))
-
-ext.bias_test.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.cb) <- n_prop_list
-cubic <- as.data.frame(bind_rows(ext.bias_test.cb, .id = "prop"))
-
-#one big dataframe
-ext.bias_test.df <- bind_rows(linear, square, cubic, .id = "model")
-```
-
-summarize across modeling options
-
-```{r}
-ext.bias_test.df %>%
-  group_by(dataset, prop) %>%
-  dplyr::arrange(p.val_fdr) %>%
-  slice(1) %>% #get lowest p-val
-  ungroup()
-```
-
-#### Z-SCORES
-```{r}
-ext.bias_test.z.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ mean_z_diff", comp_multiplier=33, mc.preschedule=FALSE)
-names(ext.bias_test.z.ln) <- n_prop_list
-linear <- as.data.frame(bind_rows(ext.bias_test.z.ln, .id = "prop"))
-
-ext.bias_test.z.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.z.sq) <- n_prop_list
-square <- as.data.frame(bind_rows(ext.bias_test.z.sq, .id = "prop"))
-
-extremeness.lm(ratio_subj_list[[2]], f="mean_z ~ poly(mean_z_diff, 2)")
-
-ext.bias_test.z.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.z.cb) <- n_prop_list
-cubic <- as.data.frame(bind_rows(ext.bias_test.z.cb, .id = "prop"))
-
-#one big dataframe
-ext.bias_test.z.df <- bind_rows(linear, square, cubic, .id = "model")
-```
-
-summarize across modeling options
-
-```{r}
-ext.bias_test.z.df %>%
-  group_by(dataset, prop) %>%
-  dplyr::arrange(p.val_fdr) %>%
-  slice(1) %>% #get lowest p-val
-  ungroup()
-```
-
-## Scale Effects of Sex
-
-sex-diff in variance as function of M:F!
-
-fix age, get diff between male and female pred. variance
-```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var,
-             pheno = as.character(pheno))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-perm_var.all <- bind_rows(perm_var, .id = "column_label")
-```
-
-maybe better to get sex_summary.csv - just need signifigance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
-         dataset = gsub("-data", "", dataset),
-         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
-         dataset = gsub("-", "", dataset),
-         #dataset = as.character(dataset),
-         pheno = as.character(pheno)) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
-         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
-  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
-
-perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
-```
-
-```{r}
-  sex_effect_plot <- perm_var_info %>%
-    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
-                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
-    ggplot() +
-    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
-    facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance")) +
-    theme(axis.text.x = element_blank())
-
-    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
-    print(sex_effect_plot)
-```
-
 # Secondary: LBCC Data

### 2024-01-07 cubic_scripts/R_scripts/featurewise_sex_bias_test.R

@@ -1,0 +1,47 @@
+#Within each combat config, test if there are significant differences in M and F centile errors in each feature 
+
+set.seed(12345)
+
+#LOAD PACKAGES
+library(dplyr)
+library(data.table)
+
+source("R_scripts/gamlss_helper_funs.R")
+
+#pheno lists
+pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
+
+#get args
+args <- commandArgs(trailingOnly = TRUE)
+data_path <- args[1] #path to csvs
+
+
+#Read in centile/z-score errors
+
+raw_files <- list.files(path = data_path, pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
+
+df_list <- list() #new empty list
+for (file in raw_files) {
+  # Read each CSV file
+  data <- fread(file)
+  
+  data <- data %>%
+    mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
+  
+  # Bind the data to the combined dataframe
+  df_list <- c(df_list, list(data))
+  assign("diffs_perm.list", df_list)
+}
+print(paste("length=", length(diffs_perm.list)))
+
+#centile t-tests
+cent.sex_t_tests_in_feat.df <- mclapply(diffs_perm.list, sex.bias.feat.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_perm.list), feature_list=pheno_list)
+
+#save rds
+saveRDS(cent.sex_t_tests_in_feat.df, file=paste0(data_path, "/featurewise_cent_sex_bias_tests.RDS"))
+
+#z-score t-tests
+z.sex_t_tests_in_feat.df <- mclapply(diffs_perm.list, sex.bias.feat.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_perm.list), feature_list=paste0(pheno_list,".z"))
+
+#save rds
+saveRDS(z.sex_t_tests_in_feat.df, file=paste0(data_path, "/featurewise_z_sex_bias_tests.RDS"))

### 2024-01-07 cubic_scripts/pull_derivatives.sh

@@ -11,6 +11,7 @@ if ! [ -d $base/ukb_basic ]
 	fi
 echo "pulling from ukb_basic"
 scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*.csv' $base/ukb_basic
+echo "pulling from ukb_basic - subject-wise"
 scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' $base/ukb_basic
 
 # Replication (Permutations) #
@@ -20,6 +21,7 @@ if ! [ -d $base/ukb_perm ]
 	fi
 echo "pulling from ukb_permute"
 scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*.csv' $base/ukb_perm
+echo "pulling from ukb_permute - subject-wise"
 scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*.csv' $base/ukb_perm
 
 # Varying M:F Ratios (Proportions) #
@@ -29,6 +31,7 @@ if ! [ -d $base/ukb_ratio ]
 	fi
 echo "pulling from ukb_ratios"
 scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*.csv' $base/ukb_ratio
+echo "pulling from ukb_ratios - subject-wise"
 scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' $base/ukb_ratio
 
 # LBCC Lifespan #

### 2024-01-07 cubic_scripts/qsub_combat_test.sh
troubleshooting

@@ -117,7 +117,7 @@ do
 	#COMBAT GAMLSS
 	elif [ $config = "cf.gamlss" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gamlss, formula = y ~ pb(log_age) + sexMale + pb(sex.age), sigma.formula = ~ pb(age_days) + sexMale + pb(sex.age)'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gamlss, formula = y ~ pb(log_age) + sexMale + pb(sex.age), sigma.formula = ~ pb(log_age) + sexMale + pb(sex.age)'" > $bash_script
 
 	#COMBAT GAMLSS W/ REF SITE
 	elif [ $config = "cf.gamlss_refA" ]

### 2024-01-07 paper_main_results.Rmd

@@
 ```
-
-# Replication
-
-## Load Data
-
-pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*_diffs.csv' .`
-
-pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*.csv' .`
-
-Read in centile/z-score errors
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_diffs.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("perm_list", df_list)
-}
-length(perm_list)
-perm.df <- bind_rows(perm_list)
-```
-
-Get errors
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("diffs_subj.list", df_list)
-}
-length(diffs_subj.list)
-diffs_subj.df <- bind_rows(diffs_subj.list)
-```
-
-Read in subject-wise summary stats
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_subj_pred.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("perm_subj_list", df_list)
-}
-length(perm_subj_list)
-perm_subj.df <- bind_rows(perm_subj_list)
-```
-
-```{r}
-n_perm_list <- lapply(c(1:10), str_pad, 3, pad = "0")
-n_perm_list <- paste0("perm-", n_perm_list)
-pheno_list.z <- paste0(pheno_list, ".z")
-```
-
-## CENTILES
-
-Do each feature's centile scores change based on combat version?
-
-```{r}
-#apply
-perm_t.tests <- mclapply(diffs_subj.list, centile.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_subj.list))
-names(perm_t.tests) <- n_perm_list
-cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-fwrite(cent_t.tests.p, file="derivatives/ukb_perm/centile_t.tests.csv")
-```
-
-```{r}
-cent_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
-
-Little plots
-
-```{r}
-lapply(diffs_subj.list, plot.cent.diffs)
-```
-
-W/in subj average plots
-
-```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-cent.diffs.v.plot <- perm_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=perm), scale="count") +
-  labs(y= "Mean Magnitude Centile Error", x = "ComBat Configurations") +
-  scale_x_discrete(labels=c(dataset_names)) +
-  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
-  ggtitle("Fig 1B. Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
-cent.diffs.v.plot
-ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
-```
-
-## Z-SCORES
-
-*need to look at distribution to check for skew!*
-
-```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs 
-pheno_abs.diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
-
-z.t.tests <- function(df){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in pheno_abs.diff.list.z) {
-    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
-names(perm_t.tests_z) <- n_perm_list
-z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
-```
-
-```{r}
-fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
-```
-
-```{r}
-z_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
-
-## SCALE EFFECTS OF SEX
-fix age, get diff between male and female pred. variance
-```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var,
-             pheno = as.character(pheno))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-perm_var.all <- bind_rows(perm_var, .id = "column_label")
-```
-
-maybe better to get sex_summary.csv - just need significance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("derivatives/ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
-         dataset = gsub("-data", "", dataset),
-         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
-         dataset = gsub("-", "", dataset),
-         #dataset = as.character(dataset),
-         pheno = as.character(pheno)) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
-         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
-  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
-
-perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
-```
-
-```{r}
-  sex_effect_plot <- perm_var_info %>%
-    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
-                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
-    ggplot() +
-    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
-    facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance")) +
-    theme(axis.text.x = element_blank())
-
-    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
-    print(sex_effect_plot)
-```
-
-# Secondary: Vary M:F
-
-```{r}
-n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
-```
-
-## Load Data
-
-pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_diffs.csv' .`
-
-pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' .`
-
-Read in centile/z-score errors
-
-```{r}
-ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
-raw_files <- list.files(path = ratio_data_path, pattern = "_diffs.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("ratio_list", df_list)
-}
-length(ratio_list)
-ratio.df <- bind_rows(ratio_list)
-
-# #sum(is.na(ratio.df))
-# na.df <- ratio.df[rowSums(is.na(ratio.df)) > 0,]
-# skimr::skim(na.df)
-# 
-# names(ratio.df)
-ratio.df[is.na(ratio.df$diff_lh_CT_cuneus.z),] %>%
-  dplyr::select(contains("_cuneus"), "participant", "dataset")
-```
-
-Read in subject-wise summary stats
-
-```{r}
-raw_files <- list.files(path = ratio_data_path, pattern = "_subj_pred.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("ratio_subj_list", df_list)
-}
-length(ratio_subj_list)
-ratio_subj.df <- bind_rows(ratio_subj_list)
-
-max(ratio_subj.df$mean_centile)
-min(ratio_subj.df$mean_centile)
-max(ratio_subj.df$mean_z)
-min(ratio_subj.df$mean_z)
-```
-
-```{r}
-for (l in unique(ratio_subj.df$dataset)){
-plot <- ratio_subj.df %>%
-  dplyr::filter(dataset == l) %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(sim.site ~ prop) +
-  ggtitle(l)
-print(plot)
-}
-```
-
-Do centile scores change based on combat version?
-
-need to look and see what happens for subjects who aren't in all samples?
-
-Plot boxplots to compare across prop.
-
-```{r, eval=FALSE}
-#too much memory to run!
-ratio.df %>%
-  dplyr::select(sim.site, prop, dataset, starts_with("abs.diff"), !ends_with(".z")) %>%
-  pivot_longer(cols= starts_with("abs.diff"), names_to="pheno", nammes_prefix="abs.diff_", values_to="cent_abs.diff") %>%
-  mutate(cent_abs.diff = cent_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=cent_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-Plot w/in subj means
-
-```{r}
-ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-
-ratio_subj.df %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=mean_z_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Z-score Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-W/in subj average plots
-
-```{r}
-cent.diffs.v.plot.r <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=dataset), scale="width") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-cent.diffs.v.plot.r
-ggsave(filename="figs/ratio_centile_violin.png", cent.diffs.v.plot.r)
-```
-
-**Stats Tests**
-Within each M:F proportion simulated, is the magnitude of subjects' mean centile errors differ significantly by combat configuration?
-
-```{r}
-prop_abs.cent_t.tests <- mclapply(ratio_list, centile.t.tests, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
-names(prop_abs.cent_t.tests) <- n_prop_list
-prop_abs.cent_t.tests_df <- bind_rows(prop_abs.cent_t.tests, .id = "column_label")
-```
-
-Visalize
-```{r}
-ggplot(prop_abs.cent_t.tests_df) +
-  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
-  geom_hline(yintercept = 208, linetype='longdash') +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
-  facet_wrap(~comp)
-```
-
-## Sex-Biases
-
-```{r}
-ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=sex), scale="width") +
-  facet_wrap(~dataset) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-```
-
-```{r}
-cent.diffs.v.boxplot.r_sex_free <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-cent.diffs.v.boxplot.r_sex_free
-ggsave(filename="figs/ratio_centile_box_sex_free.png", cent.diffs.v.boxplot.r_sex_free)
-```
-
-```{r}
-#trying to add significance testing
-abs.cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, prop) %>%
-  wilcox_test(mean_cent_abs.diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance() %>%
-  add_xy_position(x = "prop")
-
-abs.cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  droplevels() %>%
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
-    #hide.ns = TRUE
-    bracket.nudge.y = -0.5
-    ) +
-  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_abs.centile_box_sex.png", abs.cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-eval diffs (not abs. diffs)
-
-```{r}
-#trying to add significance testing
-cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, prop) %>%
-  wilcox_test(mean_cent_diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance() %>%
-  add_xy_position(x = "prop")
-
-cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
-  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  droplevels() %>%
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_diff, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
-    #hide.ns = TRUE
-    bracket.nudge.y = 4.5
-    ) +
-  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_centile_box_sex.png", cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-### Significance Testing
-
-#### CENTILES
-within each M:F prop, test male vs female subject's mean abs. centile diff
-```{r}
-sex_bias_ratio_abs.tests <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_abs.tests) <- n_prop_list
-sex_bias_ratio_abs.tests_df <- bind_rows(sex_bias_ratio_abs.tests, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_abs.tests_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-test actual (not abs. val) centile errors
-
-```{r}
-sex_bias_ratio_tests <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test = "mean_cent_diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_tests) <- n_prop_list
-sex_bias_ratio_tests_df <- bind_rows(sex_bias_ratio_tests, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_tests_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-#### Z-SCORES
-
-```{r}
-sex_bias_ratio_abs.tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_abs.diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_abs.tests.z) <- n_prop_list
-sex_bias_ratio_abs.tests.z_df <- bind_rows(sex_bias_ratio_abs.tests.z, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_abs.tests.z_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-```{r}
-sex_bias_ratio_tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_tests.z) <- n_prop_list
-sex_bias_ratio_tests.z_df <- bind_rows(sex_bias_ratio_tests.z, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_tests.z_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-## Extremeness-Biases
-Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)
-
-#### CENTILES
-
-test w/in subj avg centile ~ w/in subj avg centile error
-
-```{r}
-#testing
-test.lm <- lm(mean_cent_diff ~ mean_centile*as.factor(dataset), data=ratio_subj_list[[1]])
-tidy(test.lm)
-interactions::interact_plot(test.lm, pred = mean_centile, modx = dataset)
-
-ggplot(ratio_subj_list[[4]], aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-```{r}
-ext.bias_test.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ mean_cent_diff", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.ln) <- n_prop_list
-linear <- as.data.frame(bind_rows(ext.bias_test.ln, .id = "prop"))
-
-ext.bias_test.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.sq) <- n_prop_list
-square <- as.data.frame(bind_rows(ext.bias_test.sq, .id = "prop"))
-
-ext.bias_test.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.cb) <- n_prop_list
-cubic <- as.data.frame(bind_rows(ext.bias_test.cb, .id = "prop"))
-
-#one big dataframe
-ext.bias_test.df <- bind_rows(linear, square, cubic, .id = "model")
-```
-
-summarize across modeling options
-
-```{r}
-ext.bias_test.df %>%
-  group_by(dataset, prop) %>%
-  dplyr::arrange(p.val_fdr) %>%
-  slice(1) %>% #get lowest p-val
-  ungroup()
-```
-
-#### Z-SCORES
-```{r}
-ext.bias_test.z.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ mean_z_diff", comp_multiplier=33, mc.preschedule=FALSE)
-names(ext.bias_test.z.ln) <- n_prop_list
-linear <- as.data.frame(bind_rows(ext.bias_test.z.ln, .id = "prop"))
-
-ext.bias_test.z.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.z.sq) <- n_prop_list
-square <- as.data.frame(bind_rows(ext.bias_test.z.sq, .id = "prop"))
-
-extremeness.lm(ratio_subj_list[[2]], f="mean_z ~ poly(mean_z_diff, 2)")
-
-ext.bias_test.z.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.z.cb) <- n_prop_list
-cubic <- as.data.frame(bind_rows(ext.bias_test.z.cb, .id = "prop"))
-
-#one big dataframe
-ext.bias_test.z.df <- bind_rows(linear, square, cubic, .id = "model")
-```
-
-summarize across modeling options
-
-```{r}
-ext.bias_test.z.df %>%
-  group_by(dataset, prop) %>%
-  dplyr::arrange(p.val_fdr) %>%
-  slice(1) %>% #get lowest p-val
-  ungroup()
-```
-
-## Scale Effects of Sex
-
-sex-diff in variance as function of M:F!
-
-fix age, get diff between male and female pred. variance
-```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var,
-             pheno = as.character(pheno))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-perm_var.all <- bind_rows(perm_var, .id = "column_label")
-```
-
-maybe better to get sex_summary.csv - just need signifigance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
-         dataset = gsub("-data", "", dataset),
-         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
-         dataset = gsub("-", "", dataset),
-         #dataset = as.character(dataset),
-         pheno = as.character(pheno)) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
-         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
-  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
-
-perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
-```
-
-```{r}
-  sex_effect_plot <- perm_var_info %>%
-    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
-                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
-    ggplot() +
-    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
-    facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance")) +
-    theme(axis.text.x = element_blank())
-
-    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
-    print(sex_effect_plot)
-```
-
-# Secondary: LBCC Data

### 2024-01-07 paper_prop_results.Rmd

--- 2024-01-04 paper_results.Rmd
@@
 
+modifying to test for sex-biases w/in feature rather than mean centiles
+
+```{r}
+pheno_diff_list <- paste0("diff_", pheno_list)
+sex.bias.feat.t.tests <- function(df, feature_list=pheno_diff_list, comp_multiplier=1){
+
+  #initialize empty df to store outputs
+  t.df <- data.frame("dataset" = character(),
+                     "pheno" = character(),
+                     "p.value" = double(),
+                     "sex_diff" = double(),
+                     "t.stat" = double(),
+                     "df" = double())
+  attach(df)
+  for (pheno in feature_list) {
+  #print(paste("t-test:", pheno))
+
+  #conduct t test
+  df.sex.t <- df %>%
+    group_by(dataset) %>%
+    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
+              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
+              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
+              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
+    ungroup()
+
+  df.sex.t$pheno <- sub("diff_", "", pheno)
+
+  #append to full df
+  t.df <- rbind(t.df, df.sex.t)
+  }
+  detach(df)
+
+  #also get medians for m and f
+  med_df <- df %>%
+  dplyr::group_by(dataset, sex) %>%
+    summarise_at(feature_list, median) %>%
+    ungroup() %>%
+    pivot_longer(cols = starts_with("diff_"), names_to="pheno", names_prefix="diff_") %>%
+    pivot_wider(names_from = sex, values_from=value, names_glue = "med_{sex}") %>%
+    mutate(median_sex_diff = (med_Male - med_Female)) %>%
+    ungroup()
+
+  #combine feature dfs
+  result_df <- left_join(t.df, med_df)
+
+  #apply fdr correction
+  n_comp=length(unique(df$dataset))*length(feature_list)
+  result_df <- result_df %>%
+    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
+           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                               p.val_fdr >= 0.05 ~ FALSE),
+           dataset=as.factor(dataset))
+
+  # #summarize
+  # df.sex.t.sum <- df.sex.t %>%
+  #   group_by(dataset) %>%
+  #   dplyr::summarize(n_sig_regions = sum(sig_fdr),
+  #             min_pval = min(p.val_fdr),
+  #             max_pval = max(p.val_fdr)) %>%
+  #   ungroup() %>%
+  #   as.data.frame()
+
+  return(result_df)
+
+}
+# sex.bias.feat.t.tests(diffs.df, feature_list = global.diff)
+```
+
+Test for linear relationships between subject's actual centile score (calculated from raw data) and how much that score is affected by different combat methods.
+
 ```{r}
@@
 
-# Main Analyses
-
-## Load Data
-
-pulled variance from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_variance.csv' .`
-
-Get predicted centiles
-- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' .`
-
-```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_predictions.csv", full.names = TRUE)
-combined_preds <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
-
-  # Bind the data to the combined dataframe
-  combined_preds <- bind_rows(combined_preds, data)
-}
-```
-
-**qc - find ppts with centiles at 1 or 0**
-
-```{r}
-combined_preds %>% 
-  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0))) %>%
-  dplyr::select(participant) %>%
-  unique()
-
-combined_preds %>% 
-  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
-  dplyr::select(participant) %>%
-  unique()
-
-# test.df <- combined_preds %>% 
-#   dplyr::select(pheno_list)
-# 
-# test.df[test.df==1] <- 0
-# 
-# format(round(max(do.call(pmax, test.df)), 20), nsmall = 20)
-```
-
-saving out csv forces some centile scores to round up to 1, BUT it appears z scores are still being calculated correctly so just going to proceed!
-
-Get errors
-- pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
-
-diffs.df <- fread(file = raw_files, sep=",") %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-```
-
-Get subject-wise summaries
-- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_subj_pred.csv", full.names = TRUE) %>% unlist()
-
-subj_sum.df <- fread(file = raw_files, sep=",") %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-```
-
-## CENTILES
-
-Do centile scores change based on combat version?
-
-Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
-(see Zimmerman & Zumbo, 1993).
-Using custom functions from `ranked_welch_tests.R`.
-
-```{r}
-#initalize empty dfs to store outputs
-sum.df <- data.frame(pheno = character(),
-                     term = character(),
-                     df = double(),
-                     sumsq = double(),
-                     meansq = double(),
-                     statistic = double(),
-                     p.value = double())
-t.df <- data.frame("pheno" = character(),
-                   "group1" = character(),
-                   "group2" = character(),
-                   "p.value" = double())
-
-#run tests
-attach(diffs.df)
-for (pheno in pheno_diff.list) {
-  # df.aov <- tidy(kruskal.test(diffs.df[[pheno]] ~ diffs.df[["dataset"]]))
-  # df.aov$pheno <- pheno
-  # sum.df <- rbind(sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=diffs.df[[pheno]], g=diffs.df[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  t.df <- rbind(t.df, df.pairwise.t)
-}
-detach()
-
-#apply fdr correction
-# sum.df <- sum.df %>%
-#   dplyr::filter(term != "Residuals") %>%
-#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-t.df <- t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-t.df.sum <- t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min_pval = min(p.val_fdr),
-            max_pval = max(p.val_fdr))
-t.df.sum
-```
-
-Little plot
-```{r}
-#get means w/in feature 
-mean.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-```
-
-```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Figure 2. Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
-cent.diffs.plot
-
-ggsave("figs/win_feature_mean_cent_abs.diffs.jpeg", cent.diffs.plot)
-```
-
-### Sex-Biases
-
-Visualize magnitude mean w/in subj error
-
-```{r}
-subj_sum.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
-  ggplot(aes(x=dataset, y=mean_cent_diff, fill=sex)) +
-  geom_boxplot(trim=FALSE, scale="width") +
-  labs(y="Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot(aes(x=sim.site, y=mean_cent_abs.diff, fill=sex)) +
-  geom_violin(trim=FALSE) +
-  facet_wrap(~dataset) +
-  labs(y="Magnitude Centile Error", x = "Simulated Site") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))
-```
-
-Significance testing (w/in sub. mean abs centile error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
-
-```{r}
-sex.bias.t.tests(subj_sum.df)
-sex.bias.t.tests(subj_sum.df, to_test="mean_cent_diff")
-```
-
-### Extremeness-Biases
-
-Group at 0-0.25, >0.25 - .75 <, 0.75-1
-
-```{r}
-subj_sum.df <- subj_sum.df %>%
-  mutate(cent_cat = factor(case_when(mean_centile <= 0.25 ~ "low",
-                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
-                              mean_centile >= 0.75 ~ "high"), levels= c("low", "mid", "high"), ordered=TRUE),
-         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
-                              mean_z > -1 & mean_z < 1 ~ "mid",
-                              mean_z >= 1 ~ "high")))
-```
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(dataset ~ sim.site, scales="free_y")
-```
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff, color=sex)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  #dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-## Z-SCORES
-
-look for skew in distribution
-
-significance testing
-
-```{r}
-#initalize empty dfs to store outputs
-# z.sum.df <- data.frame(pheno = character(),
-#                        term = character(),
-#                        df = double(),
-#                        sumsq = double(),
-#                        meansq = double(),
-#                        statistic = double(),
-#                        p.value = double())
-z.t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-
-#run tests
-attach(diffs.df)
-for (pheno in z.pheno_abs.diff.list) {
-  # df.aov <- tidy(kruskal.test(diffs.df[[pheno]] ~ diffs.df[["dataset"]]))
-  # df.aov$pheno <- pheno
-  # z.sum.df <- rbind(z.sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=diffs.df[[pheno]], g=diffs.df[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  z.t.df <- rbind(z.t.df, df.pairwise.t)
-}
-detach()
-
-#apply fdr correction
-# z.sum.df <- z.sum.df %>%
-#   dplyr::filter(term != "Residuals") %>%
-#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-z.t.df <- z.t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno),
-         pheno = sub(".z", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-sum_z.t.df <- z.t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min.pval = min(p.val_fdr),
-            max.pval = max(p.val_fdr))
-
-sum_z.t.df
-```
-
-Little plot
-```{r}
-#get means w/in feature 
-z.mean.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno = gsub(".z", "", pheno)) %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-```
-
-```{r}
-z.diffs.plot <- z.mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  ggplot() +
-  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
-  ggtitle("Figure 2. Absolute errors in z-scores of brain structure features across ComBat processing pipelines")
-z.diffs.plot
-
-ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", z.diffs.plot)
-```
-
-### Sex-Biases
-
-Visualize magnitude mean w/in subj error
-
-```{r}
-subj_sum.df %>%
-  ggplot() +
-  geom_violin(aes(x=dataset, y=mean_z_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Z-Score Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot(aes(x=sex, y=mean_z_diff, fill=sex)) +
-  geom_boxplot(trim=FALSE, scale="width") +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y="Z-Score Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=sim.site, y=mean_z_abs.diff, fill=sex)) +
-  geom_violin(trim=FALSE) +
-  facet_wrap(~dataset) +
-  labs(y="Magnitude Z Error", x = "Simulated Site") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))
-```
-
-Significance testing (w/in sub. mean abs z error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
-
-```{r}
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_abs.diff")
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_diff")
-```
-
-## SCALE EFFECTS OF SEX
-
-fix age, get diff between male and female pred. variance
-
-```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
-combined_var <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_variance.csv", "", Source_File),
-           sex_standard = sex_effect/f.var)
-
-  # Bind the data to the combined dataframe
-  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
-}
-
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(dataset = gsub("-data", "", dataset),
-         dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
-
-#add more info from parcellations
-dk.parc <- dk$data %>%
-  as.data.frame() %>%
-  na.omit() %>%
-  dplyr::select(c(hemi, region, label)) %>%
-  distinct()
-
-sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
-
-plot.info <- sigma.sex.df %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
-  unique() %>%
-  mutate(dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset))
-
-combined_var <- inner_join(combined_var, plot.info)
-```
-
-```{r}
-#var plot
-var.plot.df <- combined_var %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot <- ggplot(var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = sex_standard)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-var.plot
-
-#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
-```
-
-```{r}
-#plot global
-var.plot.global.df <- combined_var %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
-  group_by(dataset)
-
-tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot.global <- ggplot(var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = sex_standard),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-var.plot.global
-```
-
-Combine plots
-```{r}
-p <- ggarrange(
-  var.plot.global, var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_annotate <- annotate_figure(p, top = text_grob("Figure 1. Standardized effect of male sex on feature variance", 
-               color = "black", size = 14, hjust=.65))
-
-ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
-```
-
-## CENTILE FANS
-
-pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_centiles.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("-data", "", dataset))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("basic_centiles", df_list)
-}
-length(basic_centiles)
-centiles.df <- bind_rows(basic_centiles)
-#skimr::skim(centiles.df)
-names(centiles.df)
-```
-
-```{r}
-global.df.long <- ukb_cn_df %>%
-  dplyr::select(participant, age_days, sex, vol_list_global) %>%
-  pivot_longer(cols = vol_list_global, values_to = "value", names_to="pheno")
-```
-
-just work with global volumes for now
-
-```{r}
-global.cent.df <- centiles.df %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('-refA', dataset))
-
-for (feature in unique(global.cent.df$pheno)) {
-
-  df <- global.cent.df %>%
-    dplyr::filter(pheno == feature)
-
- plot <- ggplot() +
-      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_9, color=df$dataset)) +
-      labs(title=paste("Male", feature, "brain chart"))
-
- print(plot)
-
- plot <- ggplot() +
-      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_9, color=df$dataset)) +
-      labs(title=paste("Female", feature, "brain chart"))
-
- print(plot)
-
- plot <- ggplot() +
-      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_9, color=df$dataset)) +
-      labs(title=paste(feature, "brain chart"))
-
- print(plot)
-
-}
-```
-
-Looking at WMV more specifically for visualizations
-```{r}
-#wm centile lines
-wmv.df <- centiles.df %>%
-  dplyr::filter(pheno == "WMV") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-```
-
-```{r}
-ggplot(wmv.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$WMV), alpha=0.7, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "dashed")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  theme_minimal() +
-  geom_text_repel(aes(x = max(ukb_cn_df$age_days), y = min(centile_1)),
-                  label="test",
-                  nudge_x = 1,
-                  na.rm = TRUE) + 
-  labs(title="WMV brain charts")
-```
-
-```{r}
-#looking at other phenotypes
-ven.df <- centiles.df %>%
-  dplyr::filter(pheno == "Ventricles") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-
-ggplot(ven.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$Ventricles), alpha=0.7, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846")) +
-  theme_minimal() +
-  labs(title="sGMV brain charts")
-```
-```{r}
-#wm centile lines
-r.sa.pc.df <- centiles.df %>%
-  dplyr::filter(pheno == "rh_SA_postcentral") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-
-ggplot(r.sa.pc.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$rh_SA_postcentral), alpha=0.7, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846")) +
-  theme_minimal() +
-  labs(title="SA of right postcentral brain charts")
-```
-
-# Replication
-
-## Load Data
-
-pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*_diffs.csv' .`
-
-pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*.csv' .`
-
-Read in centile/z-score errors
-
-```{r}
-perm_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm"
-raw_files <- list.files(path = perm_data_path, pattern = "_diffs.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("perm_list", df_list)
-}
-length(perm_list)
-perm.df <- bind_rows(perm_list)
-```
-
-Read in subject-wise summary stats
-
-```{r}
-raw_files <- list.files(path = perm_data_path, pattern = "_subj_pred.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("perm_subj_list", df_list)
-}
-length(perm_subj_list)
-perm_subj.df <- bind_rows(perm_subj_list)
-```
-
-## CENTILES
-Do centile scores change based on combat version?
-
-```{r}
-n_perm_list <- lapply(c(1:10), str_pad, 3, pad = "0")
-n_perm_list <- paste0("perm-", n_perm_list)
-pheno_list.z <- paste0(pheno_list, ".z")
-
-# combined_preds.p <- data.frame()
-get.predictions <- function(x){
-  df <- data.frame() #new empty dataframe
-  pred.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_predictions.csv"), full.names = TRUE)
-  for (file in pred.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_predictions.csv|perm-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_predictions <- mclapply(n_perm_list, get.predictions, mc.preschedule = FALSE)
-```
-
-```{r}
-get.abs.diffs <- function(x){
-  df <- x %>%
-    mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-    mutate(dataset = relevel(dataset, ref= "raw")) %>%
-    arrange(dataset) %>%
-    group_by(participant) %>%
-    dplyr::mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #centile diffs
-    dplyr::select(-any_of(pheno_list)) %>% #drop
-    dplyr::mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #z-score diffs
-    dplyr::select(-any_of(pheno_list.z)) %>% #drop
-    ungroup() %>%
-    dplyr::filter(dataset != "raw") #drop raw
-  return(df)
-}
-
-```
-
-```{r}
-perm_diffs <- mclapply(perm_predictions, get.abs.diffs, mc.preschedule = FALSE)
-saveRDS(perm_diffs)
-```
-
-```{r}
-perm_diffs <- readRDS()
-```
-
-add ANOVA
-```{r}
-
-```
-
-```{r}
-#apply
-perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
-names(perm_t.tests) <- n_perm_list
-cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-fwrite(cent_t.tests.p, file="derivatives/ukb_perm_csvs/centile_t.tests.csv")
-```
-
-```{r}
-cent_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
-
-Little plots
+plot w/in feature average centile errors
+
 ```{r}
@@
   labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
+  ggtitle("Absolute errors in centile scores across ComBat processing pipelines")
 
@@
 }
-```
-
-```{r}
-lapply(perm_diffs, plot.cent.diffs)
-```
-
-W/in subj average plots
-```{r}
-mean.subj.centile.err <- function(df){
-mean.diffs.subj <- df %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  rowwise() %>%
-  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))
-
-return(mean.diffs.subj)
-#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
-}
-perm_cent.err <- mclapply(perm_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
-names(perm_cent.err) <- n_perm_list
-cent_err.p <- bind_rows(perm_cent.err, .id = "permutation")
-
-#write out for plotting later
-saveRDS(cent_err.p, "ukb_perm_csvs/centile_errors.RDS")
-```
-
-```{r}
-cent.diffs.v.plot <- cent_err.p %>%
-  mutate(mean = mean*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=dataset, y=mean, fill=permutation)) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
-ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
-```
-
-## Z-SCORES
-
-*need to look at distribution to check for skew!*
-
-```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs 
-pheno_diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
-
-z.t.tests <- function(df){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in pheno_diff.list.z) {
-    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
-names(perm_t.tests_z) <- n_perm_list
-z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
-```
-
-```{r}
-fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
-```
-
-```{r}
-z_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
-
-## SCALE EFFECTS OF SEX
-fix age, get diff between male and female pred. variance
-```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var,
-             pheno = as.character(pheno))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-perm_var.all <- bind_rows(perm_var, .id = "column_label")
-```
-
-maybe better to get sex_summary.csv - just need significance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("derivatives/ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
-         dataset = gsub("-data", "", dataset),
-         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
-         dataset = gsub("-", "", dataset),
-         #dataset = as.character(dataset),
-         pheno = as.character(pheno)) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
-         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
-  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
-
-perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
-```
-
-```{r}
-  sex_effect_plot <- perm_var_info %>%
-    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
-                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
-    ggplot() +
-    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
-    facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance")) +
-    theme(axis.text.x = element_blank())
-
-    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
-    print(sex_effect_plot)
 ```
@@
 ```
-
-# Secondary: LBCC Data

### 2024-01-07 paper_replication_results.Rmd

--- 2024-01-06 paper_results.Rmd
@@
 
-# Main Analyses
-
-## Load Data
-
-pulled variance from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_variance.csv' .`
-
-Get predicted centiles
-- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' .`
+plot w/in feature average centile errors
 
 ```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_predictions.csv", full.names = TRUE)
-combined_preds <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
-
-  # Bind the data to the combined dataframe
-  combined_preds <- bind_rows(combined_preds, data)
-}
-```
-
-**qc - find ppts with centiles at 1 or 0**
-
-```{r}
-combined_preds %>% 
-  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0))) %>%
-  dplyr::select(participant) %>%
-  unique()
-
-combined_preds %>% 
-  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
-  dplyr::select(participant) %>%
-  unique()
-
-# test.df <- combined_preds %>% 
-#   dplyr::select(pheno_list)
-# 
-# test.df[test.df==1] <- 0
-# 
-# format(round(max(do.call(pmax, test.df)), 20), nsmall = 20)
-```
-
-saving out csv forces some centile scores to round up to 1, BUT it appears z scores are still being calculated correctly so just going to proceed!
-
-Get errors
-- pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
-
-diffs.df <- fread(file = raw_files, sep=",") %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-```
-
-Get subject-wise summaries
-- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_subj_pred.csv", full.names = TRUE) %>% unlist()
-
-subj_sum.df <- fread(file = raw_files, sep=",") %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-```
-
-Get differences in male vs female predicted variances at mean sample age
-
-```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
-combined_var <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_variance.csv", "", Source_File),
-           sex_standard = sex_effect/f.var)
-
-  # Bind the data to the combined dataframe
-  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
-}
-
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(dataset = gsub("-data", "", dataset),
-         dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
-
-sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
-
-plot.info <- sigma.sex.df %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
-  unique() %>%
-  mutate(dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset))
-
-combined_var <- inner_join(combined_var, plot.info)
-```
-
-## SCALE EFFECTS OF SEX
-
-```{r}
-#var plot
-var.plot.df <- combined_var %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot <- ggplot(var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = sex_standard)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-#var.plot
-
-#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
-```
-
-```{r}
-#plot global
-var.plot.global.df <- combined_var %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
-  group_by(dataset)
-
-tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot.global <- ggplot(var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = sex_standard),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-#var.plot.global
-```
-
-Combine plots
-```{r}
-p <- ggarrange(
-  var.plot.global, var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_annotate <- annotate_figure(p, top = text_grob("Figure 1B. Standardized effect of male sex on feature variance", 
-               color = "black", size = 14, hjust=.65))
-p_annotate
-ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
-```
-
-## CENTILES
-
-Do centile scores change based on combat version?
-Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)
-
-Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
-(see Zimmerman & Zumbo, 1993).
-Using custom functions from `ranked_welch_tests.R`.
-
-```{r}
-centile.t.tests(diffs.df)
-```
-
-Little plot
-```{r, eval=FALSE}
-#get means w/in feature 
-mean.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+#define plotting
+plot.cent.diffs <- function(df){
+mean.diffs.long <- df %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
   group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
+  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
   ungroup() %>%
+  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
   pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
@@
 
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
-
 cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
+  dplyr::filter(dataset != "raw") %>%
   mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
@@
   scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
-cent.diffs.plot
+  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
+  ggtitle("Absolute errors in centile scores across ComBat processing pipelines")
 
-ggsave("figs/win_feature_mean_cent_abs.diffs.jpeg", cent.diffs.plot)
-```
-
-### Sex-Biases
-
-Playing with different ways to visualize each subject's average centile error (i.e., raw data centile - combatted data centile) and magnitude centile error (absolute value)
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=sex, y=mean_cent_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  facet_wrap(~dataset) +
-  theme(legend.position="none") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
-  ggplot(aes(x=dataset, y=mean_cent_diff, fill=sex)) +
-  geom_violin(trim=FALSE, scale="width") +
-  labs(y="Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot(aes(x=sim.site, y=mean_cent_abs.diff, fill=sex)) +
-  geom_violin(trim=FALSE) +
-  facet_wrap(~dataset) +
-  labs(y="Magnitude Centile Error", x = "Simulated Site") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in centile scores across \nComBat configurations")
-
-#look at sex distributions of ppts across sites
-subj_sum.df %>%
-  ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))+
-  ggtitle("Distributions of male and female subjects across simulated sites")
-```
-
-Significance testing (w/in sub. mean abs centile error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
-
-```{r}
-sex.bias.t.tests(subj_sum.df)
-sex.bias.t.tests(subj_sum.df, to_test="mean_cent_diff")
-```
-
-Keeping these analyses in case I want to use one of the above violin/boxplots that show w/in subject averages (across features).
-However, while these results I think are good to have from a "be aware of how biological variables are distributed across your study sites" perspective, it doesnt really say anything interesting froma "combat methods" perspective.
-
-Testing for significant M:F differences in centile erros WITHIN each feature!
-
-check for skew (welch t.test vs ranked welch t test)
-```{r, eval=FALSE}
-for (pheno in pheno_list){
-  val <- paste0("diff_", pheno)
-  p <- ggplot(diffs.df, aes(x = diffs.df[[val]], fill= sex, color=sex)) +
-      geom_histogram(alpha=0.5, position="identity") +
-    ggtitle(val) +
-    facet_wrap(~dataset, scales="free")
-print(p)
+print(cent.diffs.plot)
 }
-```
-
-not normally distributed, use nonparametric t test of ranks with welch correction for different variances.
-Also calculating the median centile errors of males and females, since this seems like a more intuitive thing to plot than effect size (since the effect size is in units of rank, not actually centiles)
-
-```{r, warning=FALSE}
-sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df)
-
-#also get summary
-sex_t_tests_in_feat.df %>%
-    group_by(dataset) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-```
-
-```{r}
-sex_t_tests_in_feat.df2 <- sex_t_tests_in_feat.df %>%
-  mutate(pheno_cat = as.factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_)),
-    label = sub("_[^_]*_", "_", pheno)) #for plotting
-sex_t_tests_to_plot <- left_join(sex_t_tests_in_feat.df2, dk.parc, by="label")
-
-# max(sex_t_tests_to_plot$median_sex_diff)*100
-# min(sex_t_tests_to_plot$median_sex_diff)*100
-```
-
-plot w/in feature t tests
-
-```{r}
-#var plot
-sex.var.plot.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_diff <- 1
-
-sex.var.plot <- ggplot(sex.var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = median_sex_diff)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
-# sex.var.plot
-```
-
-```{r}
-#plot global
-sex.var.plot.global.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-        median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>% #drop NS values
-  group_by(dataset)
-
-tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = median_sex_diff),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
-# sex.var.plot.global
-```
-
-Combine plots - need to keep playing around to make the spacing pretty when it saves out
-```{r}
-p_sex <- ggarrange(
-  sex.var.plot.global, sex.var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
-               color = "black", size = 14, hjust=.65))
-p_sexannotate
-ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
-```
-
-Sex Effect is the median male centile error - median female centile error (i.e.
-units are in centiles).
-Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)?
-Honestly, not the most straightforward interpretation, i know - may make more sense to just plot the median errors separately in males and females than to plot the differences between the two.
-
-### Extremeness-Biases
-
-Group at 0-0.25, >0.25 - .75 <, 0.75-1
-
-```{r}
-subj_sum.df <- subj_sum.df %>%
-  mutate(cent_cat = factor(case_when(mean_centile <= 0.25 ~ "low",
-                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
-                              mean_centile >= 0.75 ~ "high"), levels= c("low", "mid", "high"), ordered=TRUE),
-         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
-                              mean_z > -1 & mean_z < 1 ~ "mid",
-                              mean_z >= 1 ~ "high")))
-```
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(dataset ~ sim.site, scales="free_y")
-```
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff, color=sex)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  #dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-The relationship between subjects' centile score and how much their centile score is altered by combat is complex and depends on sex and site.
-I think that rather than trying to investigate this directly, it'll be more interesting/effective to show visually with the secondary analyses where I vary M:F ratios.
-
-## Z-SCORES
-
-*come back to this*
-
-look for skew in distribution
-
-significance testing
-
-```{r}
-#initalize empty dfs to store outputs
-# z.sum.df <- data.frame(pheno = character(),
-#                        term = character(),
-#                        df = double(),
-#                        sumsq = double(),
-#                        meansq = double(),
-#                        statistic = double(),
-#                        p.value = double())
-z.t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-
-#run tests
-attach(diffs.df)
-for (pheno in z.pheno_abs.diff.list) {
-  # df.aov <- tidy(kruskal.test(diffs.df[[pheno]] ~ diffs.df[["dataset"]]))
-  # df.aov$pheno <- pheno
-  # z.sum.df <- rbind(z.sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=diffs.df[[pheno]], g=diffs.df[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  z.t.df <- rbind(z.t.df, df.pairwise.t)
-}
-detach()
-
-#apply fdr correction
-# z.sum.df <- z.sum.df %>%
-#   dplyr::filter(term != "Residuals") %>%
-#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-z.t.df <- z.t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno),
-         pheno = sub(".z", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-sum_z.t.df <- z.t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min.pval = min(p.val_fdr),
-            max.pval = max(p.val_fdr))
-
-sum_z.t.df
-```
-
-Little plot
-```{r}
-#get means w/in feature 
-z.mean.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno = gsub(".z", "", pheno)) %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-```
-
-```{r}
-z.diffs.plot <- z.mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  ggplot() +
-  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
-  ggtitle("Figure 2. Absolute errors in z-scores of brain structure features across ComBat processing pipelines")
-z.diffs.plot
-
-ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", z.diffs.plot)
-```
-
-### Sex-Biases
-
-Visualize magnitude mean w/in subj error
-
-```{r}
-subj_sum.df %>%
-  ggplot() +
-  geom_violin(aes(x=dataset, y=mean_z_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Z-Score Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot(aes(x=sex, y=mean_z_diff, fill=sex)) +
-  geom_boxplot(trim=FALSE, scale="width") +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y="Z-Score Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=sim.site, y=mean_z_abs.diff, fill=sex)) +
-  geom_violin(trim=FALSE) +
-  facet_wrap(~dataset) +
-  labs(y="Magnitude Z Error", x = "Simulated Site") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))
-```
-
-Significance testing (w/in sub. mean abs z error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
-
-```{r}
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_abs.diff")
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_diff")
-```
-
-## CENTILE FANS
-
-pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' .`
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_centiles.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("-data", "", dataset))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("basic_centiles", df_list)
-}
-# length(basic_centiles)
-centiles.df <- bind_rows(basic_centiles)
-#skimr::skim(centiles.df)
-# names(centiles.df)
-```
-
-just work with global volumes to visualize for now
-
-```{r}
-#get raw data points for visualization
-global.df.long <- ukb_cn_df %>%
-  dplyr::select(participant, age_days, sex, vol_list_global) %>%
-  pivot_longer(cols = vol_list_global, values_to = "value", names_to="pheno")
-```
-
-```{r}
-global.cent.df <- centiles.df %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('-refA', dataset))
-
-for (feature in unique(global.cent.df$pheno)) {
-
-  df <- global.cent.df %>%
-    dplyr::filter(pheno == feature)
-
- plot <- ggplot() +
-      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_9, color=df$dataset)) +
-      labs(title=paste("Male", feature, "brain chart"))
-
- print(plot)
-
- plot <- ggplot() +
-      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_9, color=df$dataset)) +
-      labs(title=paste("Female", feature, "brain chart"))
-
- print(plot)
-
- plot <- ggplot() +
-      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_9, color=df$dataset)) +
-      labs(title=paste(feature, "brain chart"))
-
- print(plot)
-
-}
-```
-
-Looking at WMV more specifically for visualizations
-```{r}
-#wm centile lines
-wmv.df <- centiles.df %>%
-  dplyr::filter(pheno == "WMV") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-```
-
-adding jitter to age (x-axis) because otherwise frankly it's hard to look at
-
-```{r}
-dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-
-ggplot(wmv.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$WMV), alpha=0.4, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  theme_minimal() +
-  xlab("Age (days)") + ylab("Volume (mm3)") +
-  geom_text_repel(aes(x = max(ukb_cn_df$age_days), y = min(centile_1)),
-                  label="test",
-                  nudge_x = 1,
-                  na.rm = TRUE) + 
-  labs(title="Fig 2A. WM volume brain charts")
-```
-
-*need to add labeling of % lines - ask Ayan how he did it*
-also not sure how much i should worry about the super big female outlier
-
-Look at other phenotypes
-
-```{r}
-#looking at other phenotypes
-ven.df <- centiles.df %>%
-  dplyr::filter(pheno == "Ventricles") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-
-ggplot(ven.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$Ventricles), alpha=0.7, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  theme_minimal() +
-  xlab("Age (days)") + ylab("Volume (mm3)") +
-  theme_minimal() +
-  labs(title="sGMV brain charts")
-```
-Random cortical feature
-
-```{r}
-r.sa.pc.df <- centiles.df %>%
-  dplyr::filter(pheno == "rh_SA_postcentral") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  #pivot_longer(cols=contains("centile"), name)  %>%
-  pivot_longer(
-    cols = contains("centile"),
-    names_to = c("sex", ".value"),
-    names_pattern = "([MF]?)_(centile_\\d+)",
-    values_to = "value"
-  ) %>%
- dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
-  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
-                                   sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-
-ggplot(r.sa.pc.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$rh_SA_postcentral), alpha=0.7, width=150, color="lightgray") +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  theme_minimal() +
-  xlab("Age (days)") + ylab("Volume (mm3)") +
-  theme_minimal() +
-  labs(title="SA of right postcentral brain charts")
 ```
@@
 
-pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*_diffs.csv' .`
-
-pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*.csv' .`
-
-Read in centile/z-score errors
-
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_diffs.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("perm_list", df_list)
-}
-length(perm_list)
-perm.df <- bind_rows(perm_list)
-```
-
-Get errors
+Read in centile/z-score errors (pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*_diffs.csv' .`)
 
@@
     df_list <- c(df_list, list(data))
-    assign("diffs_subj.list", df_list)
+    assign("diffs_perm.list", df_list)
 }
-length(diffs_subj.list)
-diffs_subj.df <- bind_rows(diffs_subj.list)
+length(diffs_perm.list)
+diffs_perm.df <- bind_rows(diffs_perm.list)
 ```
 
-Read in subject-wise summary stats
+Read in subject-wise summary stats (pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*.csv' .`)
 
@@
 
-```{r}
+```{r, warning=FALSE}
 #apply
-perm_t.tests <- mclapply(diffs_subj.list, centile.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_subj.list))
+perm_t.tests <- mclapply(diffs_perm.list, centile.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_perm.list))
 names(perm_t.tests) <- n_perm_list
 cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-fwrite(cent_t.tests.p, file="derivatives/ukb_perm_csvs/centile_t.tests.csv")
-
-centile.t.tests(perm_subj_list[[1]])
+fwrite(cent_t.tests.p, file="derivatives/ukb_perm/centile_t.tests.csv")
 ```
@@
 Little plots
-```{r}
-#define plotting
-plot.cent.diffs <- function(df){
-mean.diffs.long <- df %>%
-  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "raw") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
-
-print(cent.diffs.plot)
-}
-```
 
 ```{r}
-lapply(perm_diffs, plot.cent.diffs)
+lapply(diffs_perm.list, plot.cent.diffs)
 ```
@@
 W/in subj average plots
-```{r}
-mean.subj.centile.err <- function(df){
-mean.diffs.subj <- df %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  rowwise() %>%
-  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))
-
-return(mean.diffs.subj)
-#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
-}
-perm_cent.err <- mclapply(perm_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
-names(perm_cent.err) <- n_perm_list
-cent_err.p <- bind_rows(perm_cent.err, .id = "permutation")
-
-#write out for plotting later
-saveRDS(cent_err.p, "ukb_perm_csvs/centile_errors.RDS")
-```
 
 ```{r}
-cent.diffs.v.plot <- cent_err.p %>%
-  mutate(mean = mean*100) %>% #for visuals
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+cent.diffs.v.plot <- perm_subj.df %>%
+  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
   ggplot() +
-  geom_violin(aes(x=dataset, y=mean, fill=permutation)) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
+  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=perm), scale="count") +
+  labs(y= "Mean Magnitude Centile Error", x = "ComBat Configurations") +
+  scale_x_discrete(labels=c(dataset_names)) +
+  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
+  ggtitle("Fig 1B. Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
+cent.diffs.v.plot
 ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
@@
 
-## Z-SCORES
+### Sex-Biases
 
-*need to look at distribution to check for skew!*
+```{r, warning=FALSE}
+sex_t_tests_in_feat.df <- mclapply(diffs_perm.list, sex.bias.feat.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_perm.list))
 
-```{r}
-#within each phenotype, iterate paired t test between each combo of combat dfs 
-pheno_abs.diff.list.z <- paste0("abs.diff_", pheno_list, ".z")
-
-z.t.tests <- function(df){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in pheno_abs.diff.list.z) {
-    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
+#also get summary
+sex_t_tests_in_feat.df %>%
+    group_by(dataset) %>%
     dplyr::summarize(n_sig_regions = sum(sig_fdr),
@@
     as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
-names(perm_t.tests_z) <- n_perm_list
-z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
 ```
 
-```{r}
-fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
-```
+## Z-SCORES
 
-```{r}
-z_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
-```
+*come back to this*
+
+*need to look at distribution to check for skew!*
 
 ## SCALE EFFECTS OF SEX
+
 fix age, get diff between male and female pred. variance
+
 ```{r}
@@
 ```
-
-# Secondary: Vary M:F
-
-```{r}
-n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
-```
-
-## Load Data
-
-pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_diffs.csv' .`
-
-pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' .`
-
-Read in centile/z-score errors
-
-```{r}
-ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
-raw_files <- list.files(path = ratio_data_path, pattern = "_diffs.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  print(paste("reading file", file))
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list[[file]] <- as.data.frame(data)
-    assign("ratio_list", df_list)
-}
-length(ratio_list)
-ratio.df <- bind_rows(ratio_list)
-
-# #sum(is.na(ratio.df))
-# na.df <- ratio.df[rowSums(is.na(ratio.df)) > 0,]
-# skimr::skim(na.df)
-# 
-# names(ratio.df)
-ratio.df[is.na(ratio.df$diff_lh_CT_cuneus.z),] %>%
-  dplyr::select(contains("_cuneus"), "participant", "dataset")
-```
-
-Read in subject-wise summary stats
-
-```{r}
-raw_files <- list.files(path = ratio_data_path, pattern = "_subj_pred.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("ratio_subj_list", df_list)
-}
-length(ratio_subj_list)
-ratio_subj.df <- bind_rows(ratio_subj_list)
-
-max(ratio_subj.df$mean_centile)
-min(ratio_subj.df$mean_centile)
-max(ratio_subj.df$mean_z)
-min(ratio_subj.df$mean_z)
-```
-
-```{r}
-for (l in unique(ratio_subj.df$dataset)){
-plot <- ratio_subj.df %>%
-  dplyr::filter(dataset == l) %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(sim.site ~ prop) +
-  ggtitle(l)
-print(plot)
-}
-```
-
-Do centile scores change based on combat version?
-
-need to look and see what happens for subjects who aren't in all samples?
-
-Plot boxplots to compare across prop.
-
-```{r, eval=FALSE}
-#too much memory to run!
-ratio.df %>%
-  dplyr::select(sim.site, prop, dataset, starts_with("abs.diff"), !ends_with(".z")) %>%
-  pivot_longer(cols= starts_with("abs.diff"), names_to="pheno", nammes_prefix="abs.diff_", values_to="cent_abs.diff") %>%
-  mutate(cent_abs.diff = cent_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=cent_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-Plot w/in subj means
-
-```{r}
-ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-
-ratio_subj.df %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=mean_z_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Z-score Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-W/in subj average plots
-
-```{r}
-cent.diffs.v.plot.r <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=dataset), scale="width") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-cent.diffs.v.plot.r
-ggsave(filename="figs/ratio_centile_violin.png", cent.diffs.v.plot.r)
-```
-
-**Stats Tests**
-Within each M:F proportion simulated, is the magnitude of subjects' mean centile errors differ significantly by combat configuration?
-
-```{r}
-prop_abs.cent_t.tests <- mclapply(ratio_list, centile.t.tests, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
-names(prop_abs.cent_t.tests) <- n_prop_list
-prop_abs.cent_t.tests_df <- bind_rows(prop_abs.cent_t.tests, .id = "column_label")
-```
-
-Visalize
-```{r}
-ggplot(prop_abs.cent_t.tests_df) +
-  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
-  geom_hline(yintercept = 208, linetype='longdash') +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
-  facet_wrap(~comp)
-```
-
-## Sex-Biases
-
-```{r}
-ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=sex), scale="width") +
-  facet_wrap(~dataset) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-```
-
-```{r}
-cent.diffs.v.boxplot.r_sex_free <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-cent.diffs.v.boxplot.r_sex_free
-ggsave(filename="figs/ratio_centile_box_sex_free.png", cent.diffs.v.boxplot.r_sex_free)
-```
-
-```{r}
-#trying to add significance testing
-abs.cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, prop) %>%
-  wilcox_test(mean_cent_abs.diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance() %>%
-  add_xy_position(x = "prop")
-
-abs.cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  droplevels() %>%
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
-    #hide.ns = TRUE
-    bracket.nudge.y = -0.5
-    ) +
-  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_abs.centile_box_sex.png", abs.cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-eval diffs (not abs. diffs)
-
-```{r}
-#trying to add significance testing
-cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, prop) %>%
-  wilcox_test(mean_cent_diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance() %>%
-  add_xy_position(x = "prop")
-
-cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
-  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  droplevels() %>%
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_diff, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
-    #hide.ns = TRUE
-    bracket.nudge.y = 4.5
-    ) +
-  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_centile_box_sex.png", cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-### Significance Testing
-
-#### CENTILES
-within each M:F prop, test male vs female subject's mean abs. centile diff
-```{r}
-sex_bias_ratio_abs.tests <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_abs.tests) <- n_prop_list
-sex_bias_ratio_abs.tests_df <- bind_rows(sex_bias_ratio_abs.tests, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_abs.tests_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-test actual (not abs. val) centile errors
-
-```{r}
-sex_bias_ratio_tests <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test = "mean_cent_diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_tests) <- n_prop_list
-sex_bias_ratio_tests_df <- bind_rows(sex_bias_ratio_tests, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_tests_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-#### Z-SCORES
-
-```{r}
-sex_bias_ratio_abs.tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_abs.diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_abs.tests.z) <- n_prop_list
-sex_bias_ratio_abs.tests.z_df <- bind_rows(sex_bias_ratio_abs.tests.z, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_abs.tests.z_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-```{r}
-sex_bias_ratio_tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_tests.z) <- n_prop_list
-sex_bias_ratio_tests.z_df <- bind_rows(sex_bias_ratio_tests.z, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_tests.z_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-## Extremeness-Biases
-Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)
-
-#### CENTILES
-
-test w/in subj avg centile ~ w/in subj avg centile error
-
-```{r}
-#testing
-test.lm <- lm(mean_cent_diff ~ mean_centile*as.factor(dataset), data=ratio_subj_list[[1]])
-tidy(test.lm)
-interactions::interact_plot(test.lm, pred = mean_centile, modx = dataset)
-
-ggplot(ratio_subj_list[[4]], aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-```{r}
-ext.bias_test.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ mean_cent_diff", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.ln) <- n_prop_list
-linear <- as.data.frame(bind_rows(ext.bias_test.ln, .id = "prop"))
-
-ext.bias_test.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.sq) <- n_prop_list
-square <- as.data.frame(bind_rows(ext.bias_test.sq, .id = "prop"))
-
-ext.bias_test.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.cb) <- n_prop_list
-cubic <- as.data.frame(bind_rows(ext.bias_test.cb, .id = "prop"))
-
-#one big dataframe
-ext.bias_test.df <- bind_rows(linear, square, cubic, .id = "model")
-```
-
-summarize across modeling options
-
-```{r}
-ext.bias_test.df %>%
-  group_by(dataset, prop) %>%
-  dplyr::arrange(p.val_fdr) %>%
-  slice(1) %>% #get lowest p-val
-  ungroup()
-```
-
-#### Z-SCORES
-```{r}
-ext.bias_test.z.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ mean_z_diff", comp_multiplier=33, mc.preschedule=FALSE)
-names(ext.bias_test.z.ln) <- n_prop_list
-linear <- as.data.frame(bind_rows(ext.bias_test.z.ln, .id = "prop"))
-
-ext.bias_test.z.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.z.sq) <- n_prop_list
-square <- as.data.frame(bind_rows(ext.bias_test.z.sq, .id = "prop"))
-
-extremeness.lm(ratio_subj_list[[2]], f="mean_z ~ poly(mean_z_diff, 2)")
-
-ext.bias_test.z.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.z.cb) <- n_prop_list
-cubic <- as.data.frame(bind_rows(ext.bias_test.z.cb, .id = "prop"))
-
-#one big dataframe
-ext.bias_test.z.df <- bind_rows(linear, square, cubic, .id = "model")
-```
-
-summarize across modeling options
-
-```{r}
-ext.bias_test.z.df %>%
-  group_by(dataset, prop) %>%
-  dplyr::arrange(p.val_fdr) %>%
-  slice(1) %>% #get lowest p-val
-  ungroup()
-```
-
-## Scale Effects of Sex
-
-sex-diff in variance as function of M:F!
-
-fix age, get diff between male and female pred. variance
-```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var,
-             pheno = as.character(pheno))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-perm_var.all <- bind_rows(perm_var, .id = "column_label")
-```
-
-maybe better to get sex_summary.csv - just need signifigance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
-         dataset = gsub("-data", "", dataset),
-         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
-         dataset = gsub("-", "", dataset),
-         #dataset = as.character(dataset),
-         pheno = as.character(pheno)) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
-         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
-  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
-
-perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
-```
-
-```{r}
-  sex_effect_plot <- perm_var_info %>%
-    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
-                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
-    ggplot() +
-    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
-    facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance")) +
-    theme(axis.text.x = element_blank())
-
-    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
-    print(sex_effect_plot)
-```
-
-# Secondary: LBCC Data

### 2024-01-07 paper_main_results.Rmd
commit before removing extremeness bias tests

@@
 
-source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
-source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
+source("./cubic_scripts/R_scripts/stats_tests.R")
 aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
@@
   distinct()
-```
-
-# Define Functions
-
-Pairwise T tests: within each feature, run paired t tests (welch's test on ranks) pairwise on each possible pairing of combat configs and return as a single dataframe, with FDR correction for all feature + cf pair combos.
-Currently runs on magnitude (abs val) of centile diffs, but can be run on other metrics by redefining `feature_list` Recommended for use with mclapply across list of dataframes.
-Optional argument to correct for more comparisons (e.g.
-fdr-correct across an entire list of dataframes) using `comp_multiplier` arg.
-
-```{r}
-#define function
-pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)
-
-centile.t.tests <- function(df, feature_list=pheno_abs.diff.list, comp_multiplier=1){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in feature_list) {
-    df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = sub("abs.diff_", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-### EXAMPLE:
-# perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
-# names(perm_t.tests) <- n_perm_list
-# cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-```
-
-Within each combat config, test if there are significant differences in M and F subject's mean abs.
-centile error.
-Can test mean (not abs) centile errors, Z errors, etc, using `to_test` arg.
-FDR corrects for number of datasets tested
-
-```{r}
-sex.bias.t.tests <- function(df, to_test = "mean_cent_abs.diff", comp_multiplier=1){
-
-  #conduct t test
-  df.sex.t <- df %>%
-    group_by(dataset) %>%
-    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
-    ungroup()
-
-  #apply fdr correction
-  n_comp=length(unique(df$dataset))
-  df.sex.t <- df.sex.t %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE))
-  return(df.sex.t)
-  #detach()
-}
-
-# tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
-# rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
-```
-
-modifying to test for sex-biases w/in feature rather than mean centiles
-
-```{r}
-pheno_diff_list <- paste0("diff_", pheno_list)
-sex.bias.feat.t.tests <- function(df, feature_list=pheno_diff_list, comp_multiplier=1){
-
-  #initialize empty df to store outputs
-  t.df <- data.frame("dataset" = character(),
-                     "pheno" = character(),
-                     "p.value" = double(),
-                     "sex_diff" = double(),
-                     "t.stat" = double(),
-                     "df" = double())
-  attach(df)
-  for (pheno in feature_list) {
-  #print(paste("t-test:", pheno))
-
-  #conduct t test
-  df.sex.t <- df %>%
-    group_by(dataset) %>%
-    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
-    ungroup()
-
-  df.sex.t$pheno <- sub("diff_", "", pheno)
-
-  #append to full df
-  t.df <- rbind(t.df, df.sex.t)
-  }
-  detach(df)
-
-  #also get medians for m and f
-  med_df <- df %>%
-  dplyr::group_by(dataset, sex) %>%
-    summarise_at(feature_list, median) %>%
-    ungroup() %>%
-    pivot_longer(cols = starts_with("diff_"), names_to="pheno", names_prefix="diff_") %>%
-    pivot_wider(names_from = sex, values_from=value, names_glue = "med_{sex}") %>%
-    mutate(median_sex_diff = (med_Male - med_Female)) %>%
-    ungroup()
-
-  #combine feature dfs
-  result_df <- left_join(t.df, med_df)
-
-  #apply fdr correction
-  n_comp=length(unique(df$dataset))*length(feature_list)
-  result_df <- result_df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           dataset=as.factor(dataset))
-
-  # #summarize
-  # df.sex.t.sum <- df.sex.t %>%
-  #   group_by(dataset) %>%
-  #   dplyr::summarize(n_sig_regions = sum(sig_fdr),
-  #             min_pval = min(p.val_fdr),
-  #             max_pval = max(p.val_fdr)) %>%
-  #   ungroup() %>%
-  #   as.data.frame()
-
-  return(result_df)
-
-}
-# sex.bias.feat.t.tests(diffs.df, feature_list = global.diff)
-```
-
-Test for linear relationships between subject's actual centile score (calculated from raw data) and how much that score is affected by different combat methods.
-
-```{r}
-extremeness.lm <- function(df, f, comp_multiplier=1){
-
-  #conduct t test
-  df.lm <- df %>%
-    group_by(dataset) %>%
-    #returns values for last (highest order) predictor
-    summarise(p.value = tail(tidy(lm(as.formula(f)))$p.value, n=1),
-              estimate = tail(tidy(lm(as.formula(f)))$estimate, n=1),
-              t.stat = tail(tidy(lm(as.formula(f)))$statistic[2], n=1),
-              std.error = tail(tidy(lm(as.formula(f)))$std.error[2]), n=1) %>%
-    ungroup()
-  n_comp=length(unique(df$dataset))
-
-  #apply fdr correction
-  df.lm <- df.lm %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE))
-  return(df.lm)
-  #detach()
-}
-
-# tidy(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
-# summary(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
-```
-
-plot w/in feature average centile errors
-
-```{r}
-#define plotting
-plot.cent.diffs <- function(df){
-mean.diffs.long <- df %>%
-  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "raw") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Absolute errors in centile scores across ComBat processing pipelines")
-
-print(cent.diffs.plot)
-}
 ```
@@
 
-Significance testing (w/in sub. mean abs centile error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
-
-```{r}
-sex.bias.t.tests(subj_sum.df)
-sex.bias.t.tests(subj_sum.df, to_test="mean_cent_diff")
+Significance testing (w/in sub. mean abs centile error)  - using wilcox (mann whitney u) since distributions are skewed but variances are ~= across sexes
+
+```{r}
+sex.bias.wilcox.tests(subj_sum.df)
+sex.bias.wilcox.tests(subj_sum.df, to_test="mean_cent_diff")
 ```
@@
 
-```{r}
+1) is the slope nonzero in the full sample and within each site?
+
+```{r}
+diffs.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=rh_Vol_superiortemporal, y=diff_rh_Vol_superiortemporal)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset, scales="free_y")
+
+diffs.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=rh_Vol_superiortemporal, y=diff_rh_Vol_superiortemporal)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_grid(dataset ~ sim.site, scales="free_y")
+
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset, scales="free_y")
+
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff)) +
+  geom_point(alpha=0.2) +
+  geom_smooth(method="lm") +
+  facet_grid(dataset ~ sim.site, scales="free_y")
+```
+
+2) is the slope of the nondominant sex in each site nonzero?
+
+```{r}
+diffs.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=rh_Vol_superiortemporal, y=diff_rh_Vol_superiortemporal, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset, scales="free_y")
+
+diffs.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=rh_Vol_superiortemporal, y=diff_rh_Vol_superiortemporal, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_grid(dataset ~ sim.site, scales="free_y")
+
 subj_sum.df %>%
@@
   geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(~dataset, scales="free_y")
+  geom_smooth(method="lm") +
+  facet_wrap(~dataset, scales="free_y")
 
@@
   ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
+  geom_point(alpha=0.2) +
+  geom_smooth(method="lm") +
   facet_grid(dataset ~ sim.site, scales="free_y")
 ```
-
-test to see if slope of non-dominant group is nonzero!!!
 
@@
 
-*come back to this*
-
-look for skew in distribution
-
-significance testing
-
-```{r}
-#initalize empty dfs to store outputs
-# z.sum.df <- data.frame(pheno = character(),
-#                        term = character(),
-#                        df = double(),
-#                        sumsq = double(),
-#                        meansq = double(),
-#                        statistic = double(),
-#                        p.value = double())
-z.t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-
-#run tests
-attach(diffs.df)
-for (pheno in z.pheno_abs.diff.list) {
-  # df.aov <- tidy(kruskal.test(diffs.df[[pheno]] ~ diffs.df[["dataset"]]))
-  # df.aov$pheno <- pheno
-  # z.sum.df <- rbind(z.sum.df, df.aov)
-
-  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=diffs.df[[pheno]], g=diffs.df[["dataset"]], paired = TRUE, p.adj = "none"))
-  df.pairwise.t$pheno <- pheno
-  z.t.df <- rbind(z.t.df, df.pairwise.t)
+Do z scores change based on combat version?
+
+check for skew (welch t.test vs ranked welch t test)
+```{r, eval=FALSE}
+for (val in z.pheno_abs.diff.list){
+  p <- ggplot(diffs.df, aes(x = diffs.df[[val]])) +
+      geom_histogram(alpha=0.5, position="identity") +
+    ggtitle(val) +
+    facet_wrap(~dataset, scales="free")
+print(p)
 }
-detach()
-
-#apply fdr correction
-# z.sum.df <- z.sum.df %>%
-#   dplyr::filter(term != "Residuals") %>%
-#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))
-
-z.t.df <- z.t.df %>%
-  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                             p.val_fdr >= 0.05 ~ FALSE),
-         pheno = sub("abs.diff_", "", pheno),
-         pheno = sub(".z", "", pheno)) %>%
-  unite(comp, c("group1", "group2")) %>%
-  mutate(comp = as.factor(comp))
-
-sum_z.t.df <- z.t.df %>%
-  group_by(comp) %>%
-  summarize(n_sig_regions = sum(sig_fdr),
-            n_compare = n(),
-            min.pval = min(p.val_fdr),
-            max.pval = max(p.val_fdr))
-
-sum_z.t.df
+```
+
+Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)
+
+Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
+(see Zimmerman & Zumbo, 1993).
+Using custom functions from `ranked_welch_tests.R`.
+
+```{r}
+centile.t.tests(diffs.df, feature_list=z.pheno_abs.diff.list)
 ```
@@
 Little plot
-```{r}
+```{r, eval=FALSE}
 #get means w/in feature 
-z.mean.diffs.long <- diffs.df %>%
+mean.z.diffs.long <- diffs.df %>%
   dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
@@
   pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno = gsub(".z", "", pheno)) %>%
+  mutate(pheno=gsub(".z", "", pheno)) %>%
   mutate(pheno_cat = as.factor(case_when(
@@
       label = sub("_[^_]*_", "_", pheno))
-```
-
-```{r}
-z.diffs.plot <- z.mean.diffs.long %>%
+
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
+
+z.diffs.plot <- mean.z.diffs.long %>%
   dplyr::filter(dataset != "ukb_CN") %>%
@@
   labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
-  ggtitle("Figure 2. Absolute errors in z-scores of brain structure features across ComBat processing pipelines")
+  ggtitle("Absolute errors in zscores of brain structure features across ComBat processing pipelines")
 z.diffs.plot
 
-ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", z.diffs.plot)
+ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", cent.diffs.plot)
 ```
@@
 
-Visualize magnitude mean w/in subj error
-
-```{r}
-subj_sum.df %>%
+Playing with different ways to visualize each subject's average centile error (i.e., raw data centile - combatted data centile) and magnitude centile error (absolute value)
+
+```{r}
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
   ggplot() +
-  geom_violin(aes(x=dataset, y=mean_z_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Z-Score Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot(aes(x=sex, y=mean_z_diff, fill=sex)) +
-  geom_boxplot(trim=FALSE, scale="width") +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y="Z-Score Error", x = "ComBat Configurations") +
+  geom_boxplot(aes(x=sex, y=mean_z_abs.diff, fill=sex)) +
+  labs(y= "Magnitude Z Error", x = "ComBat Configurations") +
+  facet_wrap(~dataset, scales="free_y") +
+  theme(legend.position="none") +
+  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat configurations")
+
+subj_sum.df %>%
+  ggplot(aes(x=dataset, y=mean_z_diff, fill=sex)) +
+  geom_violin(trim=FALSE, scale="width") +
+  labs(y="Z Error", x = "ComBat Configurations") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across ComBat configurations")
-
+  ggtitle("Within-subject mean magnitude error in z-scores across \nComBat configurations")
+
+#look at sex distributions of ppts across sites
 subj_sum.df %>%
   ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))
-```
-
-Significance testing (w/in sub. mean abs z error) *ask about this - might be better to use wilcox since variances are ~= across sexes*
-
-```{r}
+  geom_bar(aes(x=sim.site, fill=sex))+
+  ggtitle("Distributions of male and female subjects across simulated sites")
+```
+
+Significance testing (w/in sub. mean abs centile error)
+
+```{r}
+sex.bias.t.tests(subj_sum.df, to_test="mean_z_diff")
 sex.bias.t.tests(subj_sum.df, to_test="mean_z_abs.diff")
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_diff")
-```
+```
+
+Keeping these analyses in case I want to use one of the above violin/boxplots that show w/in subject averages (across features).
+However, while these results I think are good to have from a "be aware of how biological variables are distributed across your study sites" perspective, it doesnt really say anything interesting from a "combat methods" perspective.
+
+Testing for significant M:F differences in centile erros WITHIN each feature!
+
+check for skew (welch t.test vs ranked welch t test)
+```{r, eval=FALSE}
+for (val in z.pheno_abs.diff.list){
+  p <- ggplot(diffs.df, aes(x = diffs.df[[val]], fill= sex, color=sex)) +
+      geom_histogram(alpha=0.5, position="identity") +
+    ggtitle(val) +
+    facet_wrap(~dataset, scales="free")
+print(p)
+}
+```
+
+skewed but ~ variances and not paired, using wilcox (mann-whitney u) test. Also calculating the median z-score errors of males and females
+
+```{r, warning=FALSE}
+sex_t_tests_in_feat.df <- sex.bias.feat.wilcox.tests(diffs.df)
+
+#also get summary
+sex_t_tests_in_feat.df %>%
+    group_by(dataset) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+```
+
+```{r}
+sex_t_tests_in_feat.df2 <- sex_t_tests_in_feat.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    label = sub("_[^_]*_", "_", pheno)) #for plotting
+sex_t_tests_to_plot <- left_join(sex_t_tests_in_feat.df2, dk.parc, by="label")
+
+# max(sex_t_tests_to_plot$median_sex_diff)*100
+# min(sex_t_tests_to_plot$median_sex_diff)*100
+```
+
+plot w/in feature t tests
+
+```{r}
+#var plot
+sex.var.plot.df <- sex_t_tests_to_plot %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  group_by(dataset, pheno_cat)
+
+tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="ukb_CN")
+tp$sex_diff <- 1
+
+sex.var.plot <- ggplot(sex.var.plot.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = median_sex_diff)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
+# sex.var.plot
+```
+
+```{r}
+#plot global
+sex.var.plot.global.df <- sex_t_tests_to_plot %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "WM",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA)),
+        median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>% #drop NS values
+  group_by(dataset)
+
+tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="ukb_CN")
+tp$sex_standard <- 1
+
+sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = median_sex_diff),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
+# sex.var.plot.global
+```
+
+Combine plots - need to keep playing around to make the spacing pretty when it saves out
+```{r}
+p_sex <- ggarrange(
+  sex.var.plot.global, sex.var.plot,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.1),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+
+p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
+               color = "black", size = 14, hjust=.65))
+p_sexannotate
+ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
+```
+
+Sex Effect is the median male centile error - median female centile error (i.e.
+units are in centiles).
+Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)?
+Honestly, not the most straightforward interpretation, i know - may make more sense to just plot the median errors separately in males and females than to plot the differences between the two.
+
+### Extremeness-Biases
+
+Group at 0-0.25, >0.25 - .75 <, 0.75-1
+
+```{r}
+subj_sum.df <- subj_sum.df %>%
+  mutate(cent_cat = factor(case_when(mean_centile <= 0.25 ~ "low",
+                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
+                              mean_centile >= 0.75 ~ "high"), levels= c("low", "mid", "high"), ordered=TRUE),
+         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
+                              mean_z > -1 & mean_z < 1 ~ "mid",
+                              mean_z >= 1 ~ "high")))
+```
+
+```{r}
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth() +
+  facet_grid(~dataset, scales="free_y")
+
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth() +
+  facet_grid(dataset ~ sim.site, scales="free_y")
+```
+
+test to see if slope of non-dominant group is nonzero!!!
+
+```{r}
+subj_sum.df %>%
+  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=cent_cat, y=mean_cent_diff, color=sex)) +
+  geom_violin() +
+  facet_wrap(~dataset, scales="free_y")
+
+subj_sum.df %>%
+  #dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
+  ggplot(aes(x=cent_cat, y=mean_cent_diff)) +
+  geom_violin() +
+  facet_wrap(~dataset, scales="free_y")
+```
+
+The relationship between subjects' centile score and how much their centile score is altered by combat is complex and depends on sex and site.
+I think that rather than trying to investigate this directly, it'll be more interesting/effective to show visually with the secondary analyses where I vary M:F ratios.
 

### 2024-01-08 paper_main_results.Rmd

@@
 
-Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
+Within each feature (phenotype), iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
 (see Zimmerman & Zumbo, 1993).
@@
   dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
@@
 
-```{r, warning=FALSE}
+```{r, message=FALSE}
 sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df)
@@
 
-### Extremeness-Biases
-
-Group at 0-0.25, >0.25 - .75 <, 0.75-1
-
-```{r}
-subj_sum.df <- subj_sum.df %>%
-  mutate(cent_cat = factor(case_when(mean_centile <= 0.25 ~ "low",
-                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
-                              mean_centile >= 0.75 ~ "high"), levels= c("low", "mid", "high"), ordered=TRUE),
-         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
-                              mean_z > -1 & mean_z < 1 ~ "mid",
-                              mean_z >= 1 ~ "high")))
-```
-
-1) is the slope nonzero in the full sample and within each site?
-
-```{r}
-diffs.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=rh_Vol_superiortemporal, y=diff_rh_Vol_superiortemporal)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
+### W/o extremes
+Are extreme centiles (<5% and >95%) driving observed results?
+
+Get raw centiles and merge with centile differences
+
+```{r, warning=FALSE}
+raw_centiles <- combined_preds %>%
+  dplyr::filter(dataset == "ukb_CN")
+
+merge.df <- full_join(raw_centiles, diffs.df)
+unique(merge.df$dataset)
+dim(raw_centiles)
+dim(merge.df)
+
+merge.df <- merge.df %>%
+  dplyr::select(!File_ID)
+```
+
+```{r}
+df_list <- c()
+count <- 1
+for(pheno in pheno_list) {
+  list <- c(pheno, paste0("diff_", pheno), paste0("abs.diff_", pheno))
+
+  df <- merge.df %>%
+    dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, list) %>% #get correct cols
+    group_by(participant) %>% 
+    dplyr::filter(!any((get(pheno) < 0.05 | get(pheno) > 0.95) & dataset == "ukb_CN")) %>% #drop ppts if raw centile is <0.05 or >0.95
+    ungroup() %>%
+    dplyr::filter(dataset != "ukb_CN") #now drop raw vals
+  df_list[count] <- list(df)
+  count <- count + 1
+}
+length(df_list) == length(pheno_list)
+```
+merge into single dataframe
+```{r, message=FALSE}
+no_ext <- Reduce(full_join, df_list)
+
+#make sure no cols are lost
+ncol(no_ext) == ncol(merge.df) #lost File_ID col, not going to bother rerunning rn
+nrow(no_ext)
+```
+
+Rerun stats tests
+
+centile error diffs
+```{r}
+centile.t.tests(no_ext)
+```
+
+Little plot
+```{r, eval=FALSE}
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
+
+#get means w/in feature 
+no.ext_mean.diffs.long <- no_ext %>%
+  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), mean, na.rm=TRUE)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+no.ext_mean.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
+  theme_linedraw(base_size = 11) +
+  facet_wrap(~pheno_cat) +
+  scale_y_continuous(labels = scales::percent) + 
+  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
+  ggtitle("No extremes: Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
+cent.diffs.plot
+```
+
+```{r}
+no.ext_mean.diffs.long$type <- "no_ext"
+mean.diffs.long$type <- "full"
+
+merge.df <- full_join(no.ext_mean.diffs.long, mean.diffs.long)
+
+merge.df %>%
+  ggplot() +
+  geom_boxplot(aes(x=type, y=centile_abs.diff, fill=dataset)) +
+  ggtitle("W/in feature average magnitude of centile error, w/ and w/o extreme subjects")
+```
+
+w/in feature sex-biases
+```{r, message=FALSE}
+noext.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(no_ext, feature_list=pheno_diff_list)
+
+#compare summaries
+noext.sex_t_tests_in_feat.df %>%
+    group_by(dataset) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+
+sex_t_tests_in_feat.df %>%
+    group_by(dataset) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+```
+
+plot comparative differences in medians
+```{r}
+sex.var.plot.full.df <- sex_t_tests_to_plot %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  group_by(dataset, pheno_cat) %>%
+  dplyr::select(dataset, pheno_cat, pheno, median_sex_diff)
+
+sex.var.plot.full.df$type <- "full"
+
+noext.sex_t_tests_in_feat.df2 <- noext.sex_t_tests_in_feat.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    label = sub("_[^_]*_", "_", pheno)) %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  dplyr::select(dataset, pheno_cat, pheno, median_sex_diff) %>%
+  group_by(dataset, pheno_cat)
+
+noext.sex_t_tests_in_feat.df2$type <- "no_ext"
+
+#merge
+noext_plot.df <- full_join(sex.var.plot.full.df, noext.sex_t_tests_in_feat.df2)
+```
+
+```{r}
+noext_plot.df %>%
+  ggplot() +
+  geom_col(aes(x=pheno, y=median_sex_diff, fill=type), alpha=0.7, position="identity") +
+  theme(axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  facet_wrap(~dataset, scale="free_y")
+
+noext_plot.df %>%
+  ggplot() +
+  geom_violin(aes(x=type, y=median_sex_diff, fill=type), trim=FALSE) +
   facet_wrap(~dataset, scales="free_y")
-
-diffs.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=rh_Vol_superiortemporal, y=diff_rh_Vol_superiortemporal)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_grid(dataset ~ sim.site, scales="free_y")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_wrap(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff)) +
-  geom_point(alpha=0.2) +
-  geom_smooth(method="lm") +
-  facet_grid(dataset ~ sim.site, scales="free_y")
-```
-
-2) is the slope of the nondominant sex in each site nonzero?
-
-```{r}
-diffs.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=rh_Vol_superiortemporal, y=diff_rh_Vol_superiortemporal, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_wrap(~dataset, scales="free_y")
-
-diffs.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=rh_Vol_superiortemporal, y=diff_rh_Vol_superiortemporal, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_grid(dataset ~ sim.site, scales="free_y")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_wrap(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.2) +
-  geom_smooth(method="lm") +
-  facet_grid(dataset ~ sim.site, scales="free_y")
-```
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff, color=sex)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  #dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-The relationship between subjects' centile score and how much their centile score is altered by combat is complex and depends on sex and site.
-I think that rather than trying to investigate this directly, it'll be more interesting/effective to show visually with the secondary analyses where I vary M:F ratios.
+```
 

### 2024-01-09 paper_main_results.Rmd
running with all results

@@
 
-single script to parse outputs & get results (condensed from `ukb_sim.site_analyses.Rmd`)
-
 ```{r}
@@
 
+#pheno lists
 pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
@@
 
+#diff lists
+pheno_diff_list <- paste0("diff_", pheno_list)
+pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)
+
+#z lists
 pheno_list.z <- paste0(pheno_list, ".z")
-pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)
-z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")
+z.pheno_diff_list <- paste0(pheno_diff_list, ".z")
+z.pheno_abs.diff.list <- paste0(pheno_abs.diff.list, ".z")
 
@@
 ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")
+
+#years for plotting
+ukb_cn_df <- ukb_cn_df %>%
+  mutate(age_years = age_days/365.25)
 
@@
 ```{r}
-centile.t.tests(diffs.df)
+centile.t.tests(diffs.df, feature_list=pheno_abs.diff.list) #testing abs(error)
 ```
@@
 
+Plot errors for global phenotypes
+```{r}
+global_abs.diffs <- diffs.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", vol_list_global)) %>%
+  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")
+
+global_cent_errs_plot <- global_abs.diffs %>%
+  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
+         abs.diff = abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=pheno, y=abs.diff, fill=dataset, color=dataset), scale="count") +
+  theme_linedraw(base_size = 11) +
+  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Magnitude Centile Error", x = "Phenotype") +
+  theme(plot.title = element_text(size=12)) +
+  ggtitle("Magnitude errors in centile scores of global features across ComBat pipelines")
+ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
+```
+
+*TO-DO: SIGNIFICANCE IN PLOT*
+
 ### Sex-Biases
@@
 ```{r, message=FALSE}
-sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df)
+sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list)
 
@@
 ```{r}
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 #var plot
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5), name = "Sex Effect") #
 # sex.var.plot
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
+  paletteer::scale_fill_paletteer_c("grDevices::Cork",  limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5), name = "Sex Effect") #,
 # sex.var.plot.global
@@
 #make sure no cols are lost
-ncol(no_ext) == ncol(merge.df) #lost File_ID col, not going to bother rerunning rn
-nrow(no_ext)
+dim(no_ext)
+# #make sure no cols are lost
+# ncol(no_ext) == ncol(merge.df) #lost File_ID col, not going to bother rerunning rn
+# nrow(no_ext)
 ```
@@
 ```{r}
-centile.t.tests(no_ext)
+centile.t.tests(no_ext, feature_list=pheno_abs.diff.list)
 ```
@@
 
-merge.df <- full_join(no.ext_mean.diffs.long, mean.diffs.long)
-
-merge.df %>%
+no.ext_merge <- full_join(no.ext_mean.diffs.long, mean.diffs.long)
+
+no.ext_merge %>%
   ggplot() +
@@
 Little plot
-```{r, eval=FALSE}
+```{r}
 #get means w/in feature 
 mean.z.diffs.long <- diffs.df %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
   dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
@@
 
-Playing with different ways to visualize each subject's average centile error (i.e., raw data centile - combatted data centile) and magnitude centile error (absolute value)
+Visualize
 
@@
 
-Significance testing (w/in sub. mean abs centile error)
+Significance testing (w/in sub. mean abs z error)
 
@@
 
-Keeping these analyses in case I want to use one of the above violin/boxplots that show w/in subject averages (across features).
-However, while these results I think are good to have from a "be aware of how biological variables are distributed across your study sites" perspective, it doesnt really say anything interesting from a "combat methods" perspective.
-
-Testing for significant M:F differences in centile erros WITHIN each feature!
+Testing for significant M-F differences in z score errors WITHIN each feature!
 
@@
 ```{r, eval=FALSE}
-for (val in z.pheno_abs.diff.list){
+for (val in z.pheno_diff_list){
   p <- ggplot(diffs.df, aes(x = diffs.df[[val]], fill= sex, color=sex)) +
@@
 
-skewed but ~ variances and not paired, using wilcox (mann-whitney u) test. Also calculating the median z-score errors of males and females
+skewed but ~ variances? sticking with t test of ranks. Also calculating the median z-score errors of males and females
 
 ```{r, warning=FALSE}
-sex_t_tests_in_feat.df <- sex.bias.feat.wilcox.tests(diffs.df)
+z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=z.pheno_diff_list)
 
 #also get summary
-sex_t_tests_in_feat.df %>%
+z.sex_t_tests_in_feat.df %>%
     group_by(dataset) %>%
@@
 
-```{r}
-sex_t_tests_in_feat.df2 <- sex_t_tests_in_feat.df %>%
+### W/o extremes
+Are extreme z-scores (<-2 and >2) driving observed results?
+
+Get raw centiles and merge with centile differences
+
+```{r}
+z.df_list <- c()
+count <- 1
+for(pheno in pheno_list.z) {
+  list <- c(pheno, paste0("diff_", pheno), paste0("abs.diff_", pheno))
+
+  df <- merge.df %>%
+    dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, list) %>% #get correct cols
+    group_by(participant) %>% 
+    dplyr::filter(!any((get(pheno) < -2 | get(pheno) > 2) & dataset == "ukb_CN")) %>% #drop ppts if raw z score is < -2 or > 2
+    ungroup() %>%
+    dplyr::filter(dataset != "ukb_CN") #now drop raw vals
+  z.df_list[count] <- list(df)
+  count <- count + 1
+}
+length(z.df_list) == length(pheno_list.z)
+```
+merge into single dataframe
+```{r, message=FALSE}
+z.no_ext <- Reduce(full_join, z.df_list)
+
+#make sure no cols are lost
+dim(z.no_ext)
+```
+
+Rerun stats tests
+
+z error diffs
+```{r}
+centile.t.tests(z.no_ext, feature_list=z.pheno_abs.diff.list)
+```
+
+Little plot
+```{r}
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
+
+#get means w/in feature 
+z.no.ext_mean.diffs.long <- diffs.df %>%
+  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), mean, na.rm=TRUE)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
   mutate(pheno_cat = as.factor(case_when(
@@
       TRUE ~ NA_character_)),
-    label = sub("_[^_]*_", "_", pheno)) #for plotting
-sex_t_tests_to_plot <- left_join(sex_t_tests_in_feat.df2, dk.parc, by="label")
-
-# max(sex_t_tests_to_plot$median_sex_diff)*100
-# min(sex_t_tests_to_plot$median_sex_diff)*100
-```
-
-plot w/in feature t tests
-
-```{r}
-#var plot
-sex.var.plot.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_diff <- 1
-
-sex.var.plot <- ggplot(sex.var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = median_sex_diff)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
-# sex.var.plot
-```
-
-```{r}
-#plot global
-sex.var.plot.global.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-        median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>% #drop NS values
-  group_by(dataset)
-
-tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = median_sex_diff),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5))
-# sex.var.plot.global
-```
-
-Combine plots - need to keep playing around to make the spacing pretty when it saves out
-```{r}
-p_sex <- ggarrange(
-  sex.var.plot.global, sex.var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
-               color = "black", size = 14, hjust=.65))
-p_sexannotate
-ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
-```
-
-Sex Effect is the median male centile error - median female centile error (i.e.
-units are in centiles).
-Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)?
-Honestly, not the most straightforward interpretation, i know - may make more sense to just plot the median errors separately in males and females than to plot the differences between the two.
-
-### Extremeness-Biases
-
-Group at 0-0.25, >0.25 - .75 <, 0.75-1
-
-```{r}
-subj_sum.df <- subj_sum.df %>%
-  mutate(cent_cat = factor(case_when(mean_centile <= 0.25 ~ "low",
-                              mean_centile > 0.25 & mean_centile < 0.75 ~ "mid",
-                              mean_centile >= 0.75 ~ "high"), levels= c("low", "mid", "high"), ordered=TRUE),
-         z_cat = as.factor(case_when(mean_z <= -1 ~ "low",
-                              mean_z > -1 & mean_z < 1 ~ "mid",
-                              mean_z >= 1 ~ "high")))
-```
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(dataset ~ sim.site, scales="free_y")
-```
-
-test to see if slope of non-dominant group is nonzero!!!
-
-```{r}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff, color=sex)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-
-subj_sum.df %>%
-  #dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=cent_cat, y=mean_cent_diff)) +
-  geom_violin() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-The relationship between subjects' centile score and how much their centile score is altered by combat is complex and depends on sex and site.
-I think that rather than trying to investigate this directly, it'll be more interesting/effective to show visually with the secondary analyses where I vary M:F ratios.
+      label = sub("_[^_]*_", "_", pheno))
+```
+
+```{r}
+z.no.ext_mean.diffs.long$type <- "no_ext"
+mean.z.diffs.long$type <- "full"
+
+z.no.ext_merge <- full_join(z.no.ext_mean.diffs.long, mean.z.diffs.long)
+
+z.no.ext_merge %>%
+  ggplot() +
+  geom_boxplot(aes(x=type, y=z_abs.diff, fill=dataset)) +
+  ggtitle("W/in feature average magnitude of centile error, w/ and w/o extreme subjects")
+```
+
+w/in feature sex-biases
+```{r, message=FALSE}
+noext.z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(z.no_ext, feature_list=z.pheno_diff_list)
+
+#compare summaries
+noext.z.sex_t_tests_in_feat.df %>%
+    group_by(dataset) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+
+z.sex_t_tests_in_feat.df %>%
+    group_by(dataset) %>%
+    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+              min_pval = min(p.val_fdr),
+              max_pval = max(p.val_fdr)) %>%
+    ungroup() %>%
+    as.data.frame()
+```
 
@@
                                    sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
+         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ),
+         age_years = age_days/365.25)
 ```
@@
 
-ggplot(wmv.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$WMV), alpha=0.4, width=150, color="lightgray") +
+wmv_centile_fan <- ggplot(wmv.df) +
+  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_years, y=ukb_cn_df$WMV), alpha=0.4, width=.2, color="lightgray") +
   #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_3, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
+  geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_7, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
   facet_wrap(~sex) + 
@@
   theme_minimal() +
-  xlab("Age (days)") + ylab("Volume (mm3)") +
-  geom_text_repel(aes(x = max(ukb_cn_df$age_days), y = min(centile_1)),
-                  label="test",
-                  nudge_x = 1,
-                  na.rm = TRUE) + 
-  labs(title="Fig 2A. WM volume brain charts")
+  xlab("Age (yrs)") + ylab("Volume (mm3)") +
+  theme_linedraw(base_size = 11) +
+  theme(plot.title = element_text(size=12)) +
+  labs(title="WMV brain charts")
+wmv_centile_fan
 ```
@@
   dplyr::filter(pheno == "Ventricles") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss"| dataset == "cf.lm") %>%
   #pivot_longer(cols=contains("centile"), name)  %>%
@@
   theme_minimal() +
-  labs(title="sGMV brain charts")
+  labs(title="Ventricle brain charts")
 ```
@@
 ```
+
+## FIGURE 1
+
+Combine plots
+```{r}
+fig_1 <- ggarrange(
+  wmv_centile_fan, global_cent_errs_plot,
+  ncol=2, nrow=1, 
+  widths = c(1.3, 1),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom",
+  labels="auto"
+  )
+
+fig_1_annotate <- annotate_figure(fig_1, top = text_grob("Figure 1. Choice of ComBat configuration can induce errors in brain feature centile scores",
+               color = "black", hjust = 0, x = 0))
+fig_1_annotate
+ggsave("figs/figure_1.jpeg", fig_1_annotate, width=10, height=5)
+```

### 2024-01-10 paper_prop_results.Rmd
working through ratio results

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x)
 #Sys.setenv("MC_CORES"=10L)
@@
 
-source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
-source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
+source("./cubic_scripts/R_scripts/stats_tests.R")
 aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
 ```
-
-single script to parse outputs & get results (condensed from `ukb_sim.site_analyses.Rmd`)
 
@@
 ```{r}
-ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
-raw_files <- list.files(path = ratio_data_path, pattern = "_diffs.csv", full.names = TRUE)
+raw_files <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "_diffs.csv", full.names = TRUE)
 
@@
 # names(ratio.df)
-ratio.df[is.na(ratio.df$diff_lh_CT_cuneus.z),] %>%
-  dplyr::select(contains("_cuneus"), "participant", "dataset")
 ```
@@
 ```{r}
-raw_files <- list.files(path = ratio_data_path, pattern = "_subj_pred.csv", full.names = TRUE)
+raw_files <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "_subj_pred.csv", full.names = TRUE)
 
@@
 
+Read in predicted centiles & z scores
 ```{r}
-for (l in unique(ratio_subj.df$dataset)){
-plot <- ratio_subj.df %>%
-  dplyr::filter(dataset == l) %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_grid(sim.site ~ prop) +
-  ggtitle(l)
-print(plot)
+pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "_predictions.csv", full.names = TRUE)
+combined_preds <- data.frame()
+
+df_list <- list() #new empty list
+for (file in pred.csvs) {
+  # Read each CSV file
+  data <- fread(file)
+
+  # Add a "Source_File" column with the file name
+  data <- data %>%
+    mutate(Source_File = as.factor(basename(file))) %>%
+    mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+
+    # Bind the data to the combined dataframe
+    df_list <- c(df_list, list(data))
+    assign("ratio_pred_list", df_list)
 }
+dim(ratio_pred_list)
+
+#merge into single df
+ratio_pred.df <- bind_rows(ratio_pred_list)
 ```
 
-Do centile scores change based on combat version?
+Visualize sex-distribution across sites/proportions
 
-need to look and see what happens for subjects who aren't in all samples?
+```{r}
+ratio.df %>%
+  dplyr::filter(dataset=="cf") %>%
+ggplot() +
+  geom_bar(aes(x=prop, fill=sex)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~ sim.site)
+```
+
+## CENTILES
+
+Do centile score errors change based on combat version?
+*need to look and see what happens for subjects who aren't in all samples?*
 
 Plot boxplots to compare across prop.
-
-```{r, eval=FALSE}
-#too much memory to run!
-ratio.df %>%
-  dplyr::select(sim.site, prop, dataset, starts_with("abs.diff"), !ends_with(".z")) %>%
-  pivot_longer(cols= starts_with("abs.diff"), names_to="pheno", nammes_prefix="abs.diff_", values_to="cent_abs.diff") %>%
-  mutate(cent_abs.diff = cent_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=cent_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-Plot w/in subj means
-
-```{r}
-ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff * 100) %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
-
-ratio_subj.df %>%
-  ggplot() +
-    geom_boxplot(aes(x=prop, y=mean_z_abs.diff, fill=dataset)) +
-  labs(y= "Magnitude Z-score Error", x = "M:F Ratio in Site B") +
-  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat \nconfigurations as sex imbalance changes")
-```
-
-W/in subj average plots
-
-```{r}
-cent.diffs.v.plot.r <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=dataset), scale="width") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-cent.diffs.v.plot.r
-ggsave(filename="figs/ratio_centile_violin.png", cent.diffs.v.plot.r)
-```
 
@@
 ```{r}
-prop_abs.cent_t.tests <- mclapply(ratio_list, centile.t.tests, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
+prop_abs.cent_t.tests <- mclapply(ratio_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
 names(prop_abs.cent_t.tests) <- n_prop_list
@@
 
-Visalize
+Visualize 
 ```{r}
@@
 
-## Sex-Biases
+Plot errors for global phenotypes
+```{r}
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio_subj.df$prop))
+
+prop.global_abs.diffs <- ratio.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global)) %>%
+  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")
+
+prop.global_cent_errs_plot <- prop.global_abs.diffs %>%
+  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
+         abs.diff = abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
+  theme_linedraw(base_size = 11) +
+  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Study Site") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_x_discrete(labels=percent_f) +
+  facet_wrap(~ pheno) +
+  ggtitle("Magnitude errors in centile scores of global features across ComBat pipelines")
+prop.global_cent_errs_plot
+#ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
+```
+
+### Sex-Biases
+
+Test subj. mean averages
+```{r}
+prop.sex.bias.test <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, to_test = "mean_cent_abs.diff", comp_multiplier=11)
+names(prop.sex.bias.test) <- n_prop_list
+sex_bias_ratio_tests_df <- bind_rows(prop.sex.bias.test, .id = "prop")
+```
+
+```{r}
+for (l in unique(ratio_subj.df$dataset)){
+plot <- ratio_subj.df %>%
+  dplyr::filter(dataset == l) %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_grid(sim.site ~ prop) +
+  ggtitle(l)
+print(plot)
+}
+```
 
@@
 ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=mean_cent_abs.diff, fill=sex), scale="width") +
-  facet_wrap(~dataset) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+  dplyr::filter(dataset == "cf.lm" |dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=mean_centile*100, y=mean_cent_diff*100, color=sex)) +
+  geom_point(alpha=0.2) +
+  #geom_smooth(method="lm") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
+  facet_nested(dataset + sim.site ~ prop)
+
+ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=sim.site, y=mean_cent_diff*100)) +
+  geom_violin() +
+  #geom_smooth(method="lm") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_nested(dataset ~ prop)
 ```
@@
 ```{r}
-cent.diffs.v.boxplot.r_sex_free <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
-  facet_wrap(~dataset, scale="free_y") +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
+dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+
+prop_sex_mean_cent.plot <- ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=sim.site, y=mean_cent_diff*100, fill=sex)) +
+  geom_violin() +
+  #geom_smooth(method="lm") +
+  theme_linedraw(base_size=12) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
-cent.diffs.v.boxplot.r_sex_free
-ggsave(filename="figs/ratio_centile_box_sex_free.png", cent.diffs.v.boxplot.r_sex_free)
+  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
+  xlab("Site") + ylab("Mean Centile Error")
+prop_sex_mean_cent.plot
 ```
@@
 ```{r}
-#trying to add significance testing
-abs.cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, prop) %>%
-  wilcox_test(mean_cent_abs.diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance() %>%
-  add_xy_position(x = "prop")
+dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio_subj.df$prop))
 
-abs.cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  droplevels() %>%
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_abs.diff, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
-    #hide.ns = TRUE
-    bracket.nudge.y = -0.5
-    ) +
-  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_abs.centile_box_sex.png", abs.cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-eval diffs (not abs. diffs)
-
-```{r}
-#trying to add significance testing
-cent.diffs.by.sex.stat.test <- ratio_subj.df %>%
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  group_by(dataset, prop) %>%
-  wilcox_test(mean_cent_diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance() %>%
-  add_xy_position(x = "prop")
-
-cent.diffs.v.boxplot.r_sex <- ratio_subj.df %>%
-  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
-  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
-  droplevels() %>%
-  ggplot() +
-  geom_boxplot(aes(x=prop, y=mean_cent_diff, fill=sex)) +
-  facet_wrap(~dataset) +
-  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_pvalue_manual(
-    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
-    #hide.ns = TRUE
-    bracket.nudge.y = 4.5
-    ) +
-  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
-cent.diffs.v.boxplot.r_sex
-ggsave(filename="figs/ratio_centile_box_sex.png", cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
-```
-
-### Significance Testing
-
-#### CENTILES
-within each M:F prop, test male vs female subject's mean abs. centile diff
-```{r}
-sex_bias_ratio_abs.tests <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_abs.tests) <- n_prop_list
-sex_bias_ratio_abs.tests_df <- bind_rows(sex_bias_ratio_abs.tests, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_abs.tests_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-test actual (not abs. val) centile errors
-
-```{r}
-sex_bias_ratio_tests <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test = "mean_cent_diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_tests) <- n_prop_list
-sex_bias_ratio_tests_df <- bind_rows(sex_bias_ratio_tests, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_tests_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
-
-#### Z-SCORES
-
-```{r}
-sex_bias_ratio_abs.tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_abs.diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_abs.tests.z) <- n_prop_list
-sex_bias_ratio_abs.tests.z_df <- bind_rows(sex_bias_ratio_abs.tests.z, .id = "prop")
-```
-
-Visualize
-```{r}
-sex_bias_ratio_abs.tests.z_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
+prop_sex_mean_cent.plot2 <- ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=sex, y=mean_cent_diff*100, fill=sex)) +
+  geom_violin() +
+  #geom_smooth(method="lm") +
+  theme_linedraw(base_size=12) +
+  theme(axis.text.x = element_blank()) +
+  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
+  xlab("Site") + ylab("Mean Centile Error")
+prop_sex_mean_cent.plot2
 ```
@@
 ```{r}
-sex_bias_ratio_tests.z <- mclapply(ratio_subj_list, sex.bias.t.tests, to_test ="mean_z_diff", mc.preschedule = FALSE, comp_multiplier = 11)
-names(sex_bias_ratio_tests.z) <- n_prop_list
-sex_bias_ratio_tests.z_df <- bind_rows(sex_bias_ratio_tests.z, .id = "prop")
+prop_sex_mean_cent.plot.abs <- ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=sex, y=mean_cent_abs.diff*100, fill=sex)) +
+  geom_violin() +
+  #geom_smooth(method="lm") +
+  theme_linedraw(base_size=12) +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), legend.position="bottom") +
+  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
+  ylab("Mean Absolute Centile Error")
+prop_sex_mean_cent.plot.abs
+saveRDS(prop_sex_mean_cent.plot.abs, file="figs/mean_subj_cent_err.RDS")
 ```
 
-Visualize
-```{r}
-sex_bias_ratio_tests.z_df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         prop = factor(prop, levels=n_prop_list, ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=prop, fill=sex_diff)) + 
-    geom_raster() +
-    paletteer::scale_fill_paletteer_c("ggthemes::Classic Green-Blue") +
-    theme_classic() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-    coord_fixed()
-```
+*need to add w/in feature sex-bias testing, having issues with cubic*
 
-## Extremeness-Biases
-Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)
+### W/o extremes
+Are extreme centiles (<5% and >95%) driving observed results?
 
-#### CENTILES
+Get raw centiles and merge with centile differences
 
-test w/in subj avg centile ~ w/in subj avg centile error
+```{r, warning=FALSE}
+raw_centiles <- combined_preds %>%
+  dplyr::filter(dataset == "ukb_CN")
 
-```{r}
-#testing
-test.lm <- lm(mean_cent_diff ~ mean_centile*as.factor(dataset), data=ratio_subj_list[[1]])
-tidy(test.lm)
-interactions::interact_plot(test.lm, pred = mean_centile, modx = dataset)
+merge.df <- full_join(raw_centiles, diffs.df)
+unique(merge.df$dataset)
+dim(raw_centiles)
+dim(merge.df)
 
-ggplot(ratio_subj_list[[4]], aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth() +
-  facet_wrap(~dataset, scales="free_y")
-```
-
-```{r}
-ext.bias_test.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ mean_cent_diff", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.ln) <- n_prop_list
-linear <- as.data.frame(bind_rows(ext.bias_test.ln, .id = "prop"))
-
-ext.bias_test.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.sq) <- n_prop_list
-square <- as.data.frame(bind_rows(ext.bias_test.sq, .id = "prop"))
-
-ext.bias_test.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_centile ~ poly(mean_cent_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.cb) <- n_prop_list
-cubic <- as.data.frame(bind_rows(ext.bias_test.cb, .id = "prop"))
-
-#one big dataframe
-ext.bias_test.df <- bind_rows(linear, square, cubic, .id = "model")
-```
-
-summarize across modeling options
-
-```{r}
-ext.bias_test.df %>%
-  group_by(dataset, prop) %>%
-  dplyr::arrange(p.val_fdr) %>%
-  slice(1) %>% #get lowest p-val
-  ungroup()
-```
-
-#### Z-SCORES
-```{r}
-ext.bias_test.z.ln <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ mean_z_diff", comp_multiplier=33, mc.preschedule=FALSE)
-names(ext.bias_test.z.ln) <- n_prop_list
-linear <- as.data.frame(bind_rows(ext.bias_test.z.ln, .id = "prop"))
-
-ext.bias_test.z.sq <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 2)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.z.sq) <- n_prop_list
-square <- as.data.frame(bind_rows(ext.bias_test.z.sq, .id = "prop"))
-
-extremeness.lm(ratio_subj_list[[2]], f="mean_z ~ poly(mean_z_diff, 2)")
-
-ext.bias_test.z.cb <- mclapply(ratio_subj_list, extremeness.lm, f="mean_z ~ poly(mean_z_diff, 3)", comp_multiplier= 33, mc.preschedule=FALSE)
-names(ext.bias_test.z.cb) <- n_prop_list
-cubic <- as.data.frame(bind_rows(ext.bias_test.z.cb, .id = "prop"))
-
-#one big dataframe
-ext.bias_test.z.df <- bind_rows(linear, square, cubic, .id = "model")
-```
-
-summarize across modeling options
-
-```{r}
-ext.bias_test.z.df %>%
-  group_by(dataset, prop) %>%
-  dplyr::arrange(p.val_fdr) %>%
-  slice(1) %>% #get lowest p-val
-  ungroup()
+merge.df <- merge.df %>%
+  dplyr::select(!File_ID)
 ```

### 2024-01-11 paper_main_results.Rmd

@@
 
-**qc - find ppts with centiles at 1 or 0**
+**QC: find ppts with centiles at 1 or 0**
 
@@
   unique()
-
-# test.df <- combined_preds %>% 
-#   dplyr::select(pheno_list)
-# 
-# test.df[test.df==1] <- 0
-# 
-# format(round(max(do.call(pmax, test.df)), 20), nsmall = 20)
 ```
@@
 ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
+saveRDS(p_sexannotate, file="figs/sex_diff_win_feat.RDS")
 ```
@@
   scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  theme_minimal() +
   xlab("Age (yrs)") + ylab("Volume (mm3)") +
@@
 
+not sure whether to include
+```{r}
+#look at sex distributions of ppts across sites
+data.p <- subj_sum.df %>%
+  count(sim.site, sex) %>% 
+  group_by(sim.site) %>%
+  mutate(prop = n / sum(n)) %>%
+  ggplot(aes(x = sim.site, y = prop, fill = sex)) +
+  geom_col() +
+  theme_linedraw(base_size = 11) +
+  ggtitle("Distributions of male and female subjects across simulated sites")
+```
+
 Combine plots
+
 ```{r}
@@
 ```
+
+## FIGURE 2
+need to run `paper_prop_results.Rmd` to get panel B
+
+```{r}
+prop_sex_mean_cent.plot.abs <- readRDS(file="figs/mean_subj_cent_err.RDS")
+```
+
+```{r}
+fig_2 <- ggarrange(
+  p_sex, prop_sex_mean_cent.plot.abs,
+  ncol=2, nrow=1, 
+  widths = c(1, 1.2),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  # align = "hv",
+  common.legend = FALSE,
+  labels="auto"
+  )
+# fig_2
+
+fig_2_annotate <- annotate_figure(fig_2, top = text_grob("Figure 2. Sex-biases in centile score errors across combat configurations", color = "black", hjust = 0, x = 0))
+fig_2_annotate
+ggsave("figs/figure_2.jpeg", fig_2_annotate, width=14, height=6)
+```

### 2024-01-11 paper_prop_results.Rmd

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x, purrr)
 #Sys.setenv("MC_CORES"=10L)
@@
 source("./cubic_scripts/R_scripts/stats_tests.R")
-aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+#aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
 ```
@@
 ```{r}
+dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
@@
 
-Read in predicted centiles & z scores
+Read in predicted centiles & z scores FROM RAW DATA ONLY (to save memory)
 ```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "_predictions.csv", full.names = TRUE)
-combined_preds <- data.frame()
+pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "raw_predictions.csv", full.names = TRUE)
 
@@
     mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+    mutate(dataset = gsub("_data|_predictions.csv|prop-|[0-9]|-", "", Source_File),
+           prop=gsub("-raw_predictions.csv", "", Source_File))
 
@@
     df_list <- c(df_list, list(data))
-    assign("ratio_pred_list", df_list)
+    assign("ratio_raw_pred_list", df_list)
 }
-dim(ratio_pred_list)
+length(ratio_raw_pred_list)
 
 #merge into single df
-ratio_pred.df <- bind_rows(ratio_pred_list)
+ratio_raw_pred.df <- bind_rows(ratio_raw_pred_list)
+dim(ratio_raw_pred.df)
 ```
@@
 ```{r}
-dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-
 prop_sex_mean_cent.plot <- ratio_subj.df %>%
@@
 
-```{r, warning=FALSE}
-raw_centiles <- combined_preds %>%
-  dplyr::filter(dataset == "ukb_CN")
+```{r, message=FALSE}
+ratio.merge.df <- full_join(ratio_raw_pred.df, ratio.df)
+unique(ratio.merge.df$dataset)
+dim(ratio_raw_pred.df)
+dim(ratio.merge.df)
 
-merge.df <- full_join(raw_centiles, diffs.df)
-unique(merge.df$dataset)
-dim(raw_centiles)
-dim(merge.df)
+ratio.merge.df <- ratio.merge.df %>%
+  dplyr::select(!File_ID) %>%
+  mutate(dataset=gsub("_predictions.csv","", dataset))
+```
 
-merge.df <- merge.df %>%
-  dplyr::select(!File_ID)
+split back into separate dataframes for each proportion
+
+```{r}
+# df_list <- c()
+# for (p in unique(ratio.merge.df$prop)) {
+#   df <- ratio.merge.df %>%
+#     dplyr::filter(prop == p)
+#   df_list[p] <- list(df)
+# }
+# length(df_list) == length(unique(ratio.merge.df$prop))
+```
+
+define function to remove extremes
+```{r}
+rm_ext_cents <- function(og.data, feat_list){
+  df_ext_list <- c()
+  count <- 1
+  for(pheno in feat_list) {
+    list <- c(pheno, paste0("diff_", pheno), paste0("abs.diff_", pheno))
+
+    df <- og.data %>%
+      dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, prop, all_of(list)) %>% #get correct cols
+      group_by(participant) %>% 
+      dplyr::filter(!any((get(pheno) < 0.05 | get(pheno) > 0.95) & dataset == "raw")) %>% #drop ppts if raw centile is <0.05 or >0.95
+      ungroup() %>%
+      dplyr::filter(dataset != "raw") #now drop raw vals
+    df_ext_list[count] <- list(df)
+    count <- count + 1
+  }
+  stopifnot(length(df_ext_list) == length(pheno_list))
+  df.no_ext <- Reduce(full_join, df_ext_list)
+  return(df.no_ext)
+}
+```
+
+apply fun
+
+```{r}
+ratio_no_ext_list <- mclapply(df_list, rm_ext_cents, feat_list=pheno_list, mc.preschedule = FALSE)
 ```

### 2024-01-16 paper_prop_results.Rmd
troubleshooting

@@
 
-*need to add w/in feature sex-bias testing, having issues with cubic*
+*need to add w/in feature sex-bias testing*
 
@@
 
-```{r, message=FALSE}
-ratio.merge.df <- full_join(ratio_raw_pred.df, ratio.df)
-unique(ratio.merge.df$dataset)
-dim(ratio_raw_pred.df)
-dim(ratio.merge.df)
-
-ratio.merge.df <- ratio.merge.df %>%
-  dplyr::select(!File_ID) %>%
-  mutate(dataset=gsub("_predictions.csv","", dataset))
+```{r}
+ratio.no_ext_cent <- fread(file=paste0(data_path, "/ukb_ratio/no_ext_cent.csv"))
 ```
@@
 ```{r}
-# df_list <- c()
-# for (p in unique(ratio.merge.df$prop)) {
-#   df <- ratio.merge.df %>%
-#     dplyr::filter(prop == p)
-#   df_list[p] <- list(df)
-# }
-# length(df_list) == length(unique(ratio.merge.df$prop))
+ratio.no_ext_cent_list <- c()
+for (p in unique(ratio.no_ext_cent$prop)){
+  df <- ratio.no_ext_cent %>%
+    dplyr::filter(prop == p)
+  ratio.no_ext_cent_list <- c(ratio.no_ext_cent_list, list(df))
+}
+rm(ratio.no_ext_cent)
+length(ratio.no_ext_cent_list)
+dim(ratio.no_ext_cent_list[[1]]) == dim(ratio.no_ext_cent_list[[9]]) #expect TRUE
+sum(unique(ratio.no_ext_cent_list[[1]]$participant) == unique(ratio.no_ext_cent_list[[9]]$participant)) #expect low # matches
 ```
 
-define function to remove extremes
+Test between-combat abs. centile diffs - probably faster if i just run this back on cubic
+
 ```{r}
-rm_ext_cents <- function(og.data, feat_list){
-  df_ext_list <- c()
-  count <- 1
-  for(pheno in feat_list) {
-    list <- c(pheno, paste0("diff_", pheno), paste0("abs.diff_", pheno))
+noext.prop_abs.cent_t.tests <- mclapply(ratio.no_ext_cent_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
+names(noext.prop_abs.cent_t.tests) <- n_prop_list
+noext.prop_abs.cent_t.tests_df <- bind_rows(noext.prop_abs.cent_t.tests, .id = "column_label")
+noext.prop_abs.cent_t.tests_df
+```
 
-    df <- og.data %>%
-      dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, prop, all_of(list)) %>% #get correct cols
-      group_by(participant) %>% 
-      dplyr::filter(!any((get(pheno) < 0.05 | get(pheno) > 0.95) & dataset == "raw")) %>% #drop ppts if raw centile is <0.05 or >0.95
-      ungroup() %>%
-      dplyr::filter(dataset != "raw") #now drop raw vals
-    df_ext_list[count] <- list(df)
-    count <- count + 1
-  }
-  stopifnot(length(df_ext_list) == length(pheno_list))
-  df.no_ext <- Reduce(full_join, df_ext_list)
-  return(df.no_ext)
+*need to add w/in feature sex-bias testing*
+
+## Z-SCORES
+
+Do z score errors change based on combat version?
+
+Plot boxplots to compare across prop.
+
+**Stats Tests**
+Within each M:F proportion simulated, is the magnitude of subjects' mean centile errors differ significantly by combat configuration?
+
+```{r}
+prop_abs.z_t.tests <- mclapply(ratio_list, centile.t.tests, feature_list=z.pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
+names(prop_abs.z_t.tests) <- n_prop_list
+prop_abs.z_t.tests_df <- bind_rows(prop_abs.z_t.tests, .id = "column_label")
+```
+
+Visualize 
+```{r}
+ggplot(prop_abs.z_t.tests_df) +
+  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
+  geom_hline(yintercept = 208, linetype='longdash') +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
+  facet_wrap(~comp)
+```
+
+Plot errors for global phenotypes
+```{r}
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio_subj.df$prop))
+
+prop.global_z.abs.diffs <- ratio.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global, ".z")) %>%
+  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")
+
+prop.global_z_errs_plot <- prop.global_z.abs.diffs %>%
+  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
+         abs.diff = abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
+  theme_linedraw(base_size = 11) +
+  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Magnitude Z Error", x = "% Females in Imbalanced Study Site") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_x_discrete(labels=percent_f) +
+  facet_wrap(~ pheno) +
+  ggtitle("Magnitude errors in z scores of global features across ComBat pipelines")
+prop.global_z_errs_plot
+#ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
+```
+
+### Sex-Biases
+
+Test subj. mean averages
+```{r}
+prop.sex.bias.test <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, to_test = "mean_cent_abs.diff", comp_multiplier=11)
+names(prop.sex.bias.test) <- n_prop_list
+sex_bias_ratio_tests_df <- bind_rows(prop.sex.bias.test, .id = "prop")
+```
+
+```{r}
+for (l in unique(ratio_subj.df$dataset)){
+plot <- ratio_subj.df %>%
+  dplyr::filter(dataset == l) %>%
+  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_grid(sim.site ~ prop) +
+  ggtitle(l)
+print(plot)
 }
@@
 
-apply fun
+```{r}
+ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.lm" |dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=mean_centile*100, y=mean_cent_diff*100, color=sex)) +
+  geom_point(alpha=0.2) +
+  #geom_smooth(method="lm") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_nested(dataset + sim.site ~ prop)
+
+ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=sim.site, y=mean_cent_diff*100)) +
+  geom_violin() +
+  #geom_smooth(method="lm") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_nested(dataset ~ prop)
+```
 
 ```{r}
-ratio_no_ext_list <- mclapply(df_list, rm_ext_cents, feat_list=pheno_list, mc.preschedule = FALSE)
+prop_sex_mean_cent.plot <- ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=sim.site, y=mean_cent_diff*100, fill=sex)) +
+  geom_violin() +
+  #geom_smooth(method="lm") +
+  theme_linedraw(base_size=12) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
+  xlab("Site") + ylab("Mean Centile Error")
+prop_sex_mean_cent.plot
 ```
+
+```{r}
+dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio_subj.df$prop))
+
+prop_sex_mean_cent.plot2 <- ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=sex, y=mean_cent_diff*100, fill=sex)) +
+  geom_violin() +
+  #geom_smooth(method="lm") +
+  theme_linedraw(base_size=12) +
+  theme(axis.text.x = element_blank()) +
+  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
+  xlab("Site") + ylab("Mean Centile Error")
+prop_sex_mean_cent.plot2
+```
+
+```{r}
+prop_sex_mean_cent.plot.abs <- ratio_subj.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
+  ggplot(aes(x=sex, y=mean_cent_abs.diff*100, fill=sex)) +
+  geom_violin() +
+  #geom_smooth(method="lm") +
+  theme_linedraw(base_size=12) +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), legend.position="bottom") +
+  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
+  ylab("Mean Absolute Centile Error")
+prop_sex_mean_cent.plot.abs
+saveRDS(prop_sex_mean_cent.plot.abs, file="figs/mean_subj_cent_err.RDS")
+```
+
+*need to add w/in feature sex-bias testing*
+
+### W/o extremes
+Are extreme centiles (<5% and >95%) driving observed results?
+
+Get raw centiles and merge with centile differences
+
+```{r}
+ratio.no_ext_cent <- fread(file=paste0(data_path, "/ukb_ratio/no_ext_cent.csv"))
+```
+
+split back into separate dataframes for each proportion
+
+```{r}
+ratio.no_ext_cent_list <- c()
+for (p in unique(ratio.no_ext_cent$prop)){
+  df <- ratio.no_ext_cent %>%
+    dplyr::filter(prop == p)
+  ratio.no_ext_cent_list <- c(ratio.no_ext_cent_list, list(df))
+}
+rm(ratio.no_ext_cent)
+length(ratio.no_ext_cent_list)
+dim(ratio.no_ext_cent_list[[1]]) == dim(ratio.no_ext_cent_list[[9]]) #expect TRUE
+sum(unique(ratio.no_ext_cent_list[[1]]$participant) == unique(ratio.no_ext_cent_list[[9]]$participant)) #expect low # matches
+```
+
+Test between-combat abs. centile diffs - probably faster if i just run this back on cubic
+
+```{r}
+noext.prop_abs.cent_t.tests <- mclapply(ratio.no_ext_cent_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
+names(noext.prop_abs.cent_t.tests) <- n_prop_list
+noext.prop_abs.cent_t.tests_df <- bind_rows(noext.prop_abs.cent_t.tests, .id = "column_label")
+noext.prop_abs.cent_t.tests_df
+```
+
+*need to add w/in feature sex-bias testing*
 

### 2024-01-16 cubic_scripts/R_scripts/UntitledR.R

@@ -1,13 +1,0 @@
-essai <- function(x){
-  result <- tryCatch({
-    lm(x ~ sex + age_days, data=raw.dfs)
-  } , warning = function(w) {
-    message("warning")
-    lm(x ~ sex + age_days, data=raw.df)
-  } , error = function(e) {
-    message("error")
-    lm(x ~ sex + age_days, data=raw.df)
-  } , finally = {
-    message("done")
-  } )
-}

### 2024-01-16 cubic_scripts/R_scripts/combat_apply_w_transform.R

@@ -157,7 +157,7 @@ for (l in list_of_feature_lists){
         } )
       }
       
-      cf.obj <- cf_try(cf.arg1)
+      cf.obj <- cf_try(cf.args)
     }
   }
   

### 2024-01-16 cubic_scripts/R_scripts/stats_tests.R

@@ -7,8 +7,8 @@ library(broom)
 library(purrr)
 
 # devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/ranked_welch_tests.R?token=GHSAT0AAAAAACGAYP4GAPOTD3FZTMHHQOGCZM5SE5Q")
-devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/ranked_welch_tests.R?token=GHSAT0AAAAAACGAYP4GBRV6DV6ZXXAUMDXMZM5SWKQ")
-#source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/cubic_scripts/R_scripts/ranked_welch_tests.R")
+#devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/ranked_welch_tests.R?token=GHSAT0AAAAAACGAYP4GBRV6DV6ZXXAUMDXMZM5SWKQ")
+source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/cubic_scripts/R_scripts/ranked_welch_tests.R")
 
 pheno_list <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/combat_biovar/cubic_scripts/R_scripts/pheno_list.rds")
 

### 2024-01-19 paper_main_results.Rmd
prepping for lifespan centiles

@@
 source("./cubic_scripts/R_scripts/stats_tests.R")
+```
+
+```{r}
 aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
-```
-
-```{r}
 data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"
@@
 Little plot
-```{r, eval=FALSE}
+```{r, eval=TRUE}
 dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
@@
       label = sub("_[^_]*_", "_", pheno))
-
+```
+```{r, eval=FALSE}
 no.ext_mean.diffs.long %>%

### 2024-01-19 cubic_scripts/R_scripts/fit_lifespan_centiles.R

@@ -65,13 +65,3 @@ for (name in names(cent.list)){
 
 #write out
 fwrite(cent.df, paste0(save_path, "/", fname_str, "_predictions.csv"))
-
-#get variance for males and females at mean age
-var.list <- lapply(model.files, get.var.across.ages.lbcc, og_df = df)
-var_df <- do.call(rbind, var.list)
-var_df <- var_df %>%
-  mutate(sex_effect = (m.var - f.var))
-
-#write out
-fwrite(var_df, paste0(save_path, "/", fname_str, "_variance.csv"))
-

### 2024-01-19 paper_lbcc_results.Rmd

@@
 aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+
+data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/lifespan"
+```
+
+# QC ComBat Outputs
+
+csvs scp'd from cubic
+
+Log-Scaled
+
+```{r}
+imp.sites_log.cf_cf.gam <- fread(file=paste0(data_path, "/lifespan-CN-imp-sites_log-cf_cf.gam_data.csv"))
+sum(is.na(imp.sites_log.cf_cf.gam))
+sum(imp.sites_log.cf_cf.gam < 0)
+skimr::skim(imp.sites_log.cf_cf.gam$lh_Vol_posteriorcingulate)
+min(imp.sites_log.cf_cf.gam$lh_Vol_posteriorcingulate)
+```
+
+```{r}
+imp.sites_log.cf_cf.gamlss <- fread(file=paste0(data_path, "/lifespan-CN-imp-sites_log-cf_cf.gamlss_data.csv"))
+sum(is.na(imp.sites_log.cf_cf.gamlss))
+skimr::skim(imp.sites_log.cf_cf.gam$log_age)
+skimr::skim(imp.sites_log.cf_cf.gam$sex)
+skimr::skim(imp.sites_log.cf_cf.gam$sex.age)
+skimr::skim(imp.sites_log.cf_cf.gam$fs_version)
+skimr::skim(imp.sites_log.cf_cf.gam$Vol_total)
+```
+
+Min Val Shift
+
+```{r}
+
 ```

### 2024-01-22 eda/lifespan_eda.Rmd
fixing sites with the same name that combine across studies

@@
 
+base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar"
 vol_list_global <- readRDS(file=paste0(base_path, "/cubic_scripts/R_scripts/vol_list_global.rds"))
 
-base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar"
 source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
@@
 ## Site-Level Data
+
 look just at data w site info
@@
          site = gsub(" ", "_", trimws(site))) %>%
+  unite(site, study, site, sep = "_", remove = FALSE) %>% #so sites with the same name don't get combined across studies!
   na.omit()
@@
 ggplot(cn.df.site) +
-  geom_violin(aes(x=study, y=age_days), scale="width")+
+  geom_boxplot(aes(x=reorder(study, age_days), y=age_days))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
@@
 ggplot(cn.df.site) +
-  geom_violin(aes(x=site, y=age_days, fill=study), scale="width") +
+  geom_violin(aes(x=reorder(site, age_days), y=age_days, fill=study), scale="width") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
@@
          study = trimws(study)) %>%
-  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, site))) %>%
+  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_")))) %>%
   na.omit()
@@
 unique(cn.df.imp$site)
-length(is.na(cn.df.imp$site))
+table(cn.df.imp$site, cn.df.imp$study)
+sum(is.na(cn.df.imp$site))
 max(cn.df.imp$age_days)/365.25
@@
 ggplot(cn.df.imp) +
-  geom_violin(aes(x=study, y=age_days), scale="width") +
+  geom_violin(aes(x=reorder(study, age_days), y=age_days), scale="width") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
@@
 ggplot(cn.df.imp) +
-  geom_violin(aes(x=site, y=age_days, fill=study), scale="width") +
+  geom_violin(aes(x=reorder(site, age_days), y=age_days, fill=study), scale="width") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")

### 2024-01-22 paper_lbcc_results.Rmd
prep to rerun featurewise sex tests

@@
 
-Min Val Shift
-
-```{r}
-
-```
-
-single script to parse outputs & get results (condensed from `ukb_sim.site_analyses.Rmd`)
-
 ```{r}
@@
 ```
-
-# Define Functions
-
-Pairwise T tests: within each feature, run paired t tests (welch's test on ranks) pairwise on each possible pairing of combat configs and return as a single dataframe, with FDR correction for all feature + cf pair combos.
-Currently runs on magnitude (abs val) of centile diffs, but can be run on other metrics by redefining `feature_list` Recommended for use with mclapply across list of dataframes.
-Optional argument to correct for more comparisons (e.g.
-fdr-correct across an entire list of dataframes) using `comp_multiplier` arg.
-
-```{r}
-#define function
-pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)
-
-centile.t.tests <- function(df, feature_list=pheno_abs.diff.list, comp_multiplier=1){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in feature_list) {
-    df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = sub("abs.diff_", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-### EXAMPLE:
-# perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
-# names(perm_t.tests) <- n_perm_list
-# cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-```
-
-Within each combat config, test if there are significant differences in M and F subject's mean abs.
-centile error.
-Can test mean (not abs) centile errors, Z errors, etc, using `to_test` arg.
-FDR corrects for number of datasets tested
-
-```{r}
-sex.bias.t.tests <- function(df, to_test = "mean_cent_abs.diff", comp_multiplier=1){
-
-  #conduct t test
-  df.sex.t <- df %>%
-    group_by(dataset) %>%
-    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
-    ungroup()
-
-  #apply fdr correction
-  n_comp=length(unique(df$dataset))
-  df.sex.t <- df.sex.t %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE))
-  return(df.sex.t)
-  #detach()
-}
-
-# tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
-# rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
-```
-
-modifying to test for sex-biases w/in feature rather than mean centiles
-
-```{r}
-pheno_diff_list <- paste0("diff_", pheno_list)
-sex.bias.feat.t.tests <- function(df, feature_list=pheno_diff_list, comp_multiplier=1){
-
-  #initialize empty df to store outputs
-  t.df <- data.frame("dataset" = character(),
-                     "pheno" = character(),
-                     "p.value" = double(),
-                     "sex_diff" = double(),
-                     "t.stat" = double(),
-                     "df" = double())
-  attach(df)
-  for (pheno in feature_list) {
-  #print(paste("t-test:", pheno))
-
-  #conduct t test
-  df.sex.t <- df %>%
-    group_by(dataset) %>%
-    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
-    ungroup()
-
-  df.sex.t$pheno <- sub("diff_", "", pheno)
-
-  #append to full df
-  t.df <- rbind(t.df, df.sex.t)
-  }
-  detach(df)
-
-  #also get medians for m and f
-  med_df <- df %>%
-  dplyr::group_by(dataset, sex) %>%
-    summarise_at(feature_list, median) %>%
-    ungroup() %>%
-    pivot_longer(cols = starts_with("diff_"), names_to="pheno", names_prefix="diff_") %>%
-    pivot_wider(names_from = sex, values_from=value, names_glue = "med_{sex}") %>%
-    mutate(median_sex_diff = (med_Male - med_Female)) %>%
-    ungroup()
-
-  #combine feature dfs
-  result_df <- left_join(t.df, med_df)
-
-  #apply fdr correction
-  n_comp=length(unique(df$dataset))*length(feature_list)
-  result_df <- result_df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           dataset=as.factor(dataset))
-
-  # #summarize
-  # df.sex.t.sum <- df.sex.t %>%
-  #   group_by(dataset) %>%
-  #   dplyr::summarize(n_sig_regions = sum(sig_fdr),
-  #             min_pval = min(p.val_fdr),
-  #             max_pval = max(p.val_fdr)) %>%
-  #   ungroup() %>%
-  #   as.data.frame()
-
-  return(result_df)
-
-}
-# sex.bias.feat.t.tests(diffs.df, feature_list = global.diff)
-```
-
-Test for linear relationships between subject's actual centile score (calculated from raw data) and how much that score is affected by different combat methods.
-
-```{r}
-extremeness.lm <- function(df, f, comp_multiplier=1){
-
-  #conduct t test
-  df.lm <- df %>%
-    group_by(dataset) %>%
-    #returns values for last (highest order) predictor
-    summarise(p.value = tail(tidy(lm(as.formula(f)))$p.value, n=1),
-              estimate = tail(tidy(lm(as.formula(f)))$estimate, n=1),
-              t.stat = tail(tidy(lm(as.formula(f)))$statistic[2], n=1),
-              std.error = tail(tidy(lm(as.formula(f)))$std.error[2]), n=1) %>%
-    ungroup()
-  n_comp=length(unique(df$dataset))
-
-  #apply fdr correction
-  df.lm <- df.lm %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE))
-  return(df.lm)
-  #detach()
-}
-
-# tidy(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
-# summary(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
-```
-
-plot w/in feature average centile errors
-
-```{r}
-#define plotting
-plot.cent.diffs <- function(df){
-mean.diffs.long <- df %>%
-  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "raw") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Absolute errors in centile scores across ComBat processing pipelines")
-
-print(cent.diffs.plot)
-}
-```
-
-# Secondary: LBCC Data

### 2024-01-22 paper_prop_results.Rmd
prep to rerun featurewise sex tests

@@
 ```{r}
-data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"
+der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"
 
@@
 
-#also load raw data
-ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")
-
 #years for plotting
@@
 ```
-
-# Secondary: Vary M:F
 
@@
 Read in predicted centiles & z scores FROM RAW DATA ONLY (to save memory)
-```{r}
+```{r, eval=FALSE}
 pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "raw_predictions.csv", full.names = TRUE)
@@
 
-```{r}
+```{r, eval=FALSE}
 ratio.df %>%
@@
 
-Do centile score errors change based on combat version?
+Do centile score absolute errors change based on combat version? Tested by `ratio_results.R`, results written to `full_featurewise_cent_t_tests.csv`
 *need to look and see what happens for subjects who aren't in all samples?*
 
-Plot boxplots to compare across prop.
-
-**Stats Tests**
-Within each M:F proportion simulated, is the magnitude of subjects' mean centile errors differ significantly by combat configuration?
+```{r}
+cent.t.test.df <- fread(paste0(data_path, "/ukb_ratio/full_featurewise_cent_t_tests.csv"))
+cent.t.test.df <- cent.t.test.df %>%
+  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+         comp = as.factor(comp))
+```
 
 ```{r}
-prop_abs.cent_t.tests <- mclapply(ratio_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
-names(prop_abs.cent_t.tests) <- n_prop_list
-prop_abs.cent_t.tests_df <- bind_rows(prop_abs.cent_t.tests, .id = "column_label")
-```
-
-Visualize 
-```{r}
-ggplot(prop_abs.cent_t.tests_df) +
-  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
-  geom_hline(yintercept = 208, linetype='longdash') +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
-  facet_wrap(~comp)
+cent.t.test.df %>%
+  # filter(comp=="cf.gamlss_cf.gam") %>%
+  ggplot() +
+  geom_col(aes(x=prop, y=n_sig_regions, fill=comp), position="dodge") +
+  facet_wrap(~comp) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
 ```
@@
 ```{r}
-percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio_subj.df$prop))
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))
 
@@
 ```
+*need to add stat tests to plot*
 
@@
 
-Test subj. mean averages
+Test subj. mean abs errors (mean_cent_abs.diff) using `ratio_subject-wise_results.R`, saved as `subj.abs.mean_sex_bias_cent_t_tests.csv`
+
 ```{r}
-prop.sex.bias.test <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, to_test = "mean_cent_abs.diff", comp_multiplier=11)
-names(prop.sex.bias.test) <- n_prop_list
-sex_bias_ratio_tests_df <- bind_rows(prop.sex.bias.test, .id = "prop")
+cent.subj.mean.test.df <- fread(paste0(data_path, "/ukb_ratio/subj.abs.mean_sex_bias_cent_t_tests.csv"))
+cent.subj.mean.test.df <- cent.subj.mean.test.df %>%
+  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+         dataset = as.factor(dataset))
+cent.subj.mean.test.df
 ```
@@
 ```{r}
-for (l in unique(ratio_subj.df$dataset)){
-plot <- ratio_subj.df %>%
-  dplyr::filter(dataset == l) %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_grid(sim.site ~ prop) +
-  ggtitle(l)
-print(plot)
-}
+cent.subj.mean.test.df %>%
+  # filter(comp=="cf.gamlss_cf.gam") %>%
+  ggplot() +
+  geom_col(aes(x=prop, y=sig_fdr, fill=dataset), position="dodge") +
+  facet_wrap(~dataset) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
 ```
 
-```{r}
-ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.lm" |dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=mean_centile*100, y=mean_cent_diff*100, color=sex)) +
-  geom_point(alpha=0.2) +
-  #geom_smooth(method="lm") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_nested(dataset + sim.site ~ prop)
-
-ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=sim.site, y=mean_cent_diff*100)) +
-  geom_violin() +
-  #geom_smooth(method="lm") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_nested(dataset ~ prop)
-```
+Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
 
 ```{r}
-prop_sex_mean_cent.plot <- ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=sim.site, y=mean_cent_diff*100, fill=sex)) +
-  geom_violin() +
-  #geom_smooth(method="lm") +
-  theme_linedraw(base_size=12) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
-  xlab("Site") + ylab("Mean Centile Error")
-prop_sex_mean_cent.plot
+cent.subj.sex.test.list <- readRDS(paste0(data_path, "/ukb_ratio/full_featurewise_cent_sex_bias_tests.RDS"))
+
 ```
-
-```{r}
-dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio_subj.df$prop))
-
-prop_sex_mean_cent.plot2 <- ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=sex, y=mean_cent_diff*100, fill=sex)) +
-  geom_violin() +
-  #geom_smooth(method="lm") +
-  theme_linedraw(base_size=12) +
-  theme(axis.text.x = element_blank()) +
-  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
-  xlab("Site") + ylab("Mean Centile Error")
-prop_sex_mean_cent.plot2
-```
-
-```{r}
-prop_sex_mean_cent.plot.abs <- ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=sex, y=mean_cent_abs.diff*100, fill=sex)) +
-  geom_violin() +
-  #geom_smooth(method="lm") +
-  theme_linedraw(base_size=12) +
-  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), legend.position="bottom") +
-  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
-  ylab("Mean Absolute Centile Error")
-prop_sex_mean_cent.plot.abs
-saveRDS(prop_sex_mean_cent.plot.abs, file="figs/mean_subj_cent_err.RDS")
-```
-
-*need to add w/in feature sex-bias testing*
 

### 2024-01-22 cubic_scripts/R_scripts/get_lifespan_deltas.R

@@ -1,4 +1,5 @@
 # change in centile scores for subjects when calculated on data harmonized with ComBat GAM vs ComBatLS
+# also run basic stats tests
 
 #LOAD PACKAGES
 library(data.table)
@@ -10,7 +11,7 @@ args <- commandArgs(trailingOnly = TRUE)
 gam_csv_path <- as.character(args[1]) #path to combat-GAM fit centiles
 gamlss_csv_path <- as.character(args[2]) #path to combatLS fit centiles
 save_path <- as.character(args[3]) #path to save output csv
-fname_str <- as.character(args[4]) #string for saving name
+fname_str <- as.character(args[4]) #string for saving nameå
 
 #LOAD DFS
 #read in data
@@ -50,6 +51,7 @@ stopifnot(nrow(gam.df) == nrow(gamlss.df))
 #subtract
 diff.df <- gam.df - gamlss.df
 diff.df <- diff.df %>%
+  rename_with(~ paste("delta", ., sep = ".")) %>%
   mutate(id = row_number())
 
 #merge back
@@ -58,8 +60,9 @@ cf_deltas <- base::merge(id.gam.df, diff.df, by = "id") %>%
   dplyr::select(!id)
 
 #SAVE
-print("saving")
+print("saving deltas")
 fwrite(cf_deltas, paste0(save_path, "/", fname_str, "_pred_deltas.csv"))
+
 
 print("DONE")
 

### 2024-01-22 cubic_scripts/R_scripts/lifespan_results.R

@@ -1,0 +1,27 @@
+# change in centile scores for subjects when calculated on data harmonized with ComBat GAM vs ComBatLS
+# also run basic stats tests
+
+#LOAD PACKAGES
+library(data.table)
+library(tidyverse)
+library(dplyr)
+
+#get args
+args <- commandArgs(trailingOnly = TRUE)
+delta_path <- as.character(args[1]) #path to delta csv
+save_path <- as.character(args[2]) #path to save output csvs
+fname_str <- as.character(args[3]) #string for saving name
+
+devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/stats_tests.R")
+
+#pheno lists
+pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
+pheno_list.z <- paste0(pheno_list, ".z")
+
+#LOAD DFS
+#read in data
+delta.df <- fread(gam_csv_path)
+
+
+print("DONE")
+

### 2024-01-22 cubic_scripts/R_scripts/stats_tests.R

@@ -155,7 +155,7 @@ sex.bias.t.tests <- function(df, to_test = "mean_cent_abs.diff", comp_multiplier
 # tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
 # rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
 
-#sex.bias.feat.t.tests() KEEP THIS ONE
+#sex.bias.feat.t.tests()
 #modifying to test for sex-biases w/in feature rather than mean centiles
 
 pheno_diff_list <- paste0("diff_", pheno_list)
@@ -218,6 +218,9 @@ sex.bias.feat.t.tests <- function(df, feature_list=pheno_diff_list, comp_multipl
                   sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                       p.val_fdr >= 0.05 ~ FALSE),
                   dataset=as.factor(dataset))
+  
+  #add back source_file
+  result_df$Source_File <- df$Source_File
   
   # #summarize
   # df.sex.t.sum <- df.sex.t %>%

### 2024-01-22 cubic_scripts/pull_derivatives.sh

@@ -35,3 +35,9 @@ echo "pulling from ukb_ratios - subject-wise"
 scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' $base/ukb_ratio
 
 # LBCC Lifespan #
+if ! [ -d $base/lifespan ]
+	then
+	mkdir $base/lifespan
+	fi
+echo "pulling from lifespan diffs (ComBat GAM - ComBatLS)"
+scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/lifespan/centile_csvs/*.csv' $base/lifespan

### 2024-01-22 cubic_scripts/rerun_varying_ratios_featwise_stats.sh

@@ -1,0 +1,45 @@
+#!/bin/bash
+
+# Run stats test on the outputs of 'run_varying_ratios_pipeline.sh'. Trying to make a faster, cubic-based version of 'paper_prop_results.Rmd'
+
+# expected outputs:
+# $csv_path/*_diffs.csv x11
+# $csv_path/subject-wise/*_subj_pred.csv x11
+# $csv_path/*_featurewise_cent_t_tests.csv x2
+# $csv_path/*_featurewise_cent_z_tests.csv x2
+# $csv_path/subj.abs.mean_sex_bias_cent_t_tests.csv x2
+# $csv_path/subj.abs.mean_sex_bias_z_t_tests.csv x2
+# $csv_path/*_featurewise_cent_sex_bias_tests.csv x1
+# $csv_path/*_featurewise_z_sex_bias_tests.csv x1
+# $csv_path/*_no_ext.csv x11
+
+#######################################################################################
+# SET PATHS
+img=/cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif #singularity image
+base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
+study_dir=$base/ukb_ratios
+csv_path=$study_dir/perm_centile_csvs
+r_base=$base/cubic_scripts/R_scripts # paths to R scripts
+bash_dir=$study_dir/stats_qsubs
+#######################################################################################
+# featurewise sex bias tests
+## based on `qsub_sex_bias_test.sh`
+r_script=$r_base/featurewise_sex_bias_test.R
+bash_script=$bash_dir/sex_bias_test.sh
+touch $bash_script
+	
+echo "singularity run --cleanenv $img Rscript --save $r_script $csv_path 'full'" > $bash_script
+## qsub bash script
+qsub -N prop_sex-bias -o $bash_dir/sex_bias_test_out.txt -e $bash_dir/sex_bias_test_err.txt -l h_vmem=60.5G,s_vmem=60.0G $bash_script
+
+#######################################################
+# STATS TESTS OF ERRORS W/O EXTREMES
+# featurewise sex bias tests
+## based on `qsub_sex_bias_test.sh`
+r_script=$r_base/featurewise_sex_bias_test.R
+bash_script=$bash_dir/sex_bias_test.sh
+touch $bash_script
+
[…]

### 2024-01-24 paper_prop_results.Rmd
fixing prop additions to df

@@
 ```
+some quick eda
+
+```{r}
+ratio.df %>%
+  ggplot() +
+  geom_violin(aes(y=diff_GMV, x= sex, fill=sim.site)) +
+  facet_grid(dataset~prop, scales="free_y")
+
+ratio.df %>%
+  ggplot() +
+  geom_violin(aes(y=abs.diff_GMV, x= sex, fill=sim.site)) +
+  facet_grid(dataset~prop, scales="free_y")
+```
 
@@
 Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
+**RERUNNING TO KEEP PROP VALS AS EXPLICIT NAMES, COME BACK TO THIS**
 
@@
 cent.subj.sex.test.list <- readRDS(paste0(data_path, "/ukb_ratio/full_featurewise_cent_sex_bias_tests.RDS"))
-
 ```
@@
 
-Get raw centiles and merge with centile differences
+Test differences in abs. magnitude of centile errors w/in each feature
 
 ```{r}
-ratio.no_ext_cent <- fread(file=paste0(data_path, "/ukb_ratio/no_ext_cent.csv"))
+noext.cent.t.test.df <- fread(paste0(data_path, "/ukb_ratio/no.ext_featurewise_cent_t_tests.csv"))
+noext.cent.t.test.df <- noext.cent.t.test.df %>%
+  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+         comp = as.factor(comp))
+noext.cent.t.test.df
 ```
 
-split back into separate dataframes for each proportion
+compare results
+```{r}
+noext.cent.t.test.df_reduced <- noext.cent.t.test.df %>%
+  mutate(full_df = 0) %>%
+  dplyr::select(comp, prop, n_sig_regions, full_df)
+cent.t.test.df_reduced <- cent.t.test.df %>%
+  mutate(full_df = 1) %>%
+  dplyr::select(comp, prop, n_sig_regions, full_df)
 
-```{r}
-ratio.no_ext_cent_list <- c()
-for (p in unique(ratio.no_ext_cent$prop)){
-  df <- ratio.no_ext_cent %>%
-    dplyr::filter(prop == p)
-  ratio.no_ext_cent_list <- c(ratio.no_ext_cent_list, list(df))
-}
-rm(ratio.no_ext_cent)
-length(ratio.no_ext_cent_list)
-dim(ratio.no_ext_cent_list[[1]]) == dim(ratio.no_ext_cent_list[[9]]) #expect TRUE
-sum(unique(ratio.no_ext_cent_list[[1]]$participant) == unique(ratio.no_ext_cent_list[[9]]$participant)) #expect low # matches
+full_join(noext.cent.t.test.df_reduced, cent.t.test.df_reduced) %>%
+  ggplot() +
+  geom_col(aes(x=prop, y=n_sig_regions, fill=as.factor(full_df)), position="dodge") +
+  facet_wrap(~comp) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 ```
-
-Test between-combat abs. centile diffs - probably faster if i just run this back on cubic
-
-```{r}
-noext.prop_abs.cent_t.tests <- mclapply(ratio.no_ext_cent_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
-names(noext.prop_abs.cent_t.tests) <- n_prop_list
-noext.prop_abs.cent_t.tests_df <- bind_rows(noext.prop_abs.cent_t.tests, .id = "column_label")
-noext.prop_abs.cent_t.tests_df
-```
-
-*need to add w/in feature sex-bias testing*
+Test sex-biases in centile errors w/in each feature
+**COME BACK TO THIS**
 
@@
 
-Plot boxplots to compare across prop.
-
-**Stats Tests**
-Within each M:F proportion simulated, is the magnitude of subjects' mean centile errors differ significantly by combat configuration?
+compare abs(z errors) w/in features - Tested by `ratio_results.R`, results written to `full_featurewise_z_t_tests.csv`
 
 ```{r}
-prop_abs.z_t.tests <- mclapply(ratio_list, centile.t.tests, feature_list=z.pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
-names(prop_abs.z_t.tests) <- n_prop_list
-prop_abs.z_t.tests_df <- bind_rows(prop_abs.z_t.tests, .id = "column_label")
+z.t.test.df <- fread(paste0(data_path, "/ukb_ratio/full_featurewise_z_t_tests.csv"))
+z.t.test.df <- z.t.test.df %>%
+  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+         comp = as.factor(comp))
 ```
 
-Visualize 
 ```{r}
-ggplot(prop_abs.z_t.tests_df) +
-  geom_col(aes(x=column_label, y=n_sig_regions, fill=column_label)) +
-  geom_hline(yintercept = 208, linetype='longdash') +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
-  facet_wrap(~comp)
+z.t.test.df %>%
+  # filter(comp=="cf.gamlss_cf.gam") %>%
+  ggplot() +
+  geom_col(aes(x=prop, y=n_sig_regions, fill=comp), position="dodge") +
+  facet_wrap(~comp) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
 ```
@@
 ```{r}
-percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio_subj.df$prop))
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))
 
-prop.global_z.abs.diffs <- ratio.df %>%
+z.prop.global_abs.diffs <- ratio.df %>%
   dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global, ".z")) %>%
@@
 
-prop.global_z_errs_plot <- prop.global_z.abs.diffs %>%
+prop.global_z_errs_plot <- z.prop.global_abs.diffs %>%
   dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
-         abs.diff = abs.diff*100) %>% #for visuals
+         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE)) %>%
   ggplot() +
@@
 prop.global_z_errs_plot
-#ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
+#ggsave("figs/global_z_errors.jpeg", prop.global_z_errs_plot)
 ```
+*need to add stat tests to plot*
 

### 2024-01-24 cubic_scripts/R_scripts/featurewise_sex_bias_test.R

@@ -40,9 +40,7 @@ for (file in raw_files) {
   df_list <- c(df_list, list(data))
   assign("diffs_perm.list", df_list)
 }
-print(paste("length=", length(diffs_perm.list)))
-print(unique(diffs_perm.list[1]$Source_File))
-print(unique(diffs_perm.list[2]$Source_File)
+
 #centile t-tests
 cent.sex_t_tests_in_feat.df <- lapply(diffs_perm.list, sex.bias.feat.t.tests, comp_multiplier=length(diffs_perm.list), feature_list=diff_list)
 

### 2024-01-24 cubic_scripts/R_scripts/stats_tests.R

@@ -216,9 +216,8 @@ sex.bias.feat.t.tests <- function(df, feature_list, comp_multiplier=1){
                   dataset=as.factor(dataset))
   
   #add back source_file
-  print(unique(df$Source_File))
-  stopifnot(length(unique(df$Source_File)) == 1)
-  result_df$Source_File <- unique(df$Source_File)
+  stopifnot(length(unique(df$prop)) == 1)
+  result_df$prop <- unique(df$prop)
   
   # #summarize
   # df.sex.t.sum <- df.sex.t %>%

### 2024-01-25 paper_prop_results.Rmd
fixing typos for saving list as RDS, rm extreme z scores above 2

@@
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE)) %>%
+         pheno = factor(pheno, levels = c("WMV.z", "GMV.z", "sGMV.z", "Ventricles.z"), ordered = TRUE)) %>%
   ggplot() +
@@
   scale_x_discrete(labels=percent_f) +
-  facet_wrap(~ pheno) +
+  facet_wrap(~pheno) +
   ggtitle("Magnitude errors in z scores of global features across ComBat pipelines")
@@
 ### Sex-Biases
+Test subj. mean abs errors (mean_cent_abs.diff) using `ratio_subject-wise_results.R`, saved as `subj.abs.mean_sex_bias_z_t_tests.csv`
 
-Test subj. mean averages
 ```{r}
-prop.sex.bias.test <- mclapply(ratio_subj_list, sex.bias.t.tests, mc.preschedule = FALSE, to_test = "mean_cent_abs.diff", comp_multiplier=11)
-names(prop.sex.bias.test) <- n_prop_list
-sex_bias_ratio_tests_df <- bind_rows(prop.sex.bias.test, .id = "prop")
+z.subj.mean.test.df <- fread(paste0(data_path, "/ukb_ratio/subj.abs.mean_sex_bias_z_t_tests.csv"))
+z.subj.mean.test.df <- z.subj.mean.test.df %>%
+  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+         dataset = as.factor(dataset))
+z.subj.mean.test.df
 ```
@@
 ```{r}
-for (l in unique(ratio_subj.df$dataset)){
-plot <- ratio_subj.df %>%
-  dplyr::filter(dataset == l) %>%
-  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_grid(sim.site ~ prop) +
-  ggtitle(l)
-print(plot)
-}
+z.subj.mean.test.df %>%
+  # filter(comp=="cf.gamlss_cf.gam") %>%
+  ggplot() +
+  geom_col(aes(x=prop, y=sig_fdr, fill=dataset), position="dodge") +
+  facet_wrap(~dataset) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
 ```
 
+Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
+**RERUNNING TO KEEP PROP VALS AS EXPLICIT NAMES, COME BACK TO THIS**
+
+### W/o extremes
+
+Are extreme z-scores (<2 and >2) driving observed results?
+
+Test differences in abs. magnitude of z errors w/in each feature
+
 ```{r}
-ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.lm" |dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=mean_centile*100, y=mean_cent_diff*100, color=sex)) +
-  geom_point(alpha=0.2) +
-  #geom_smooth(method="lm") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_nested(dataset + sim.site ~ prop)
-
-ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=sim.site, y=mean_cent_diff*100)) +
-  geom_violin() +
-  #geom_smooth(method="lm") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_nested(dataset ~ prop)
+noext.z.t.test.df <- fread(paste0(data_path, "/ukb_ratio/no.ext_featurewise_z_t_tests.csv"))
+noext.z.t.test.df <- noext.z.t.test.df %>%
+  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+         comp = as.factor(comp))
+noext.z.t.test.df
 ```
 
+compare results
 ```{r}
-prop_sex_mean_cent.plot <- ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=sim.site, y=mean_cent_diff*100, fill=sex)) +
-  geom_violin() +
-  #geom_smooth(method="lm") +
-  theme_linedraw(base_size=12) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
-  xlab("Site") + ylab("Mean Centile Error")
-prop_sex_mean_cent.plot
+noext.z.t.test.df_reduced <- noext.z.t.test.df %>%
+  mutate(full_df = 0) %>%
+  dplyr::select(comp, prop, n_sig_regions, full_df)
+z.t.test.df_reduced <- z.t.test.df %>%
+  mutate(full_df = 1) %>%
+  dplyr::select(comp, prop, n_sig_regions, full_df)
+
+full_join(noext.z.t.test.df_reduced, z.t.test.df_reduced) %>%
+  ggplot() +
+  geom_col(aes(x=prop, y=n_sig_regions, fill=as.factor(full_df)), position="dodge") +
+  facet_wrap(~comp) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 ```
-
-```{r}
-dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio_subj.df$prop))
-
-prop_sex_mean_cent.plot2 <- ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=sex, y=mean_cent_diff*100, fill=sex)) +
-  geom_violin() +
-  #geom_smooth(method="lm") +
-  theme_linedraw(base_size=12) +
-  theme(axis.text.x = element_blank()) +
-  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
-  xlab("Site") + ylab("Mean Centile Error")
-prop_sex_mean_cent.plot2
-```
-
-```{r}
-prop_sex_mean_cent.plot.abs <- ratio_subj.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
-  ggplot(aes(x=sex, y=mean_cent_abs.diff*100, fill=sex)) +
-  geom_violin() +
-  #geom_smooth(method="lm") +
-  theme_linedraw(base_size=12) +
-  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), legend.position="bottom") +
-  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
-  ylab("Mean Absolute Centile Error")
-prop_sex_mean_cent.plot.abs
-saveRDS(prop_sex_mean_cent.plot.abs, file="figs/mean_subj_cent_err.RDS")
-```
-
-*need to add w/in feature sex-bias testing*
-
-### W/o extremes
-Are extreme centiles (<5% and >95%) driving observed results?
-
-Get raw centiles and merge with centile differences
-
-```{r}
-ratio.no_ext_cent <- fread(file=paste0(data_path, "/ukb_ratio/no_ext_cent.csv"))
-```
-
-split back into separate dataframes for each proportion
-
-```{r}
-ratio.no_ext_cent_list <- c()
-for (p in unique(ratio.no_ext_cent$prop)){
-  df <- ratio.no_ext_cent %>%
-    dplyr::filter(prop == p)
-  ratio.no_ext_cent_list <- c(ratio.no_ext_cent_list, list(df))
-}
-rm(ratio.no_ext_cent)
-length(ratio.no_ext_cent_list)
-dim(ratio.no_ext_cent_list[[1]]) == dim(ratio.no_ext_cent_list[[9]]) #expect TRUE
-sum(unique(ratio.no_ext_cent_list[[1]]$participant) == unique(ratio.no_ext_cent_list[[9]]$participant)) #expect low # matches
-```
-
-Test between-combat abs. centile diffs - probably faster if i just run this back on cubic
-
-```{r}
-noext.prop_abs.cent_t.tests <- mclapply(ratio.no_ext_cent_list, centile.t.tests, feature_list=pheno_abs.diff.list, comp_multiplier=11, mc.preschedule = FALSE) #FDR correction across 11 M:F permutations
-names(noext.prop_abs.cent_t.tests) <- n_prop_list
-noext.prop_abs.cent_t.tests_df <- bind_rows(noext.prop_abs.cent_t.tests, .id = "column_label")
-noext.prop_abs.cent_t.tests_df
-```
-
-*need to add w/in feature sex-bias testing*
-
-## Scale Effects of Sex
-
-sex-diff in variance as function of M:F!
-
-fix age, get diff between male and female pred. variance
-```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
-
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var,
-             pheno = as.character(pheno))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-perm_var.all <- bind_rows(perm_var, .id = "column_label")
-```
-
-maybe better to get sex_summary.csv - just need signifigance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
-         dataset = gsub("-data", "", dataset),
-         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
-         dataset = gsub("-", "", dataset),
-         #dataset = as.character(dataset),
-         pheno = as.character(pheno)) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
-         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
-  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
-
-perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
-```
-
-```{r}
-  sex_effect_plot <- perm_var_info %>%
-    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
-                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
-    ggplot() +
-    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
-    facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance")) +
-    theme(axis.text.x = element_blank())
-
-    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
-    print(sex_effect_plot)
-```
+Test sex-biases in z errors w/in each feature
+**COME BACK TO THIS**

### 2024-01-25 cubic_scripts/R_scripts/combat_apply_w_transform.R

@@ -161,7 +161,7 @@ for (l in list_of_feature_lists){
             message("done")
           } )
         }
-        cf.obj <- cf_gamlss_try(cf.arg1)
+        cf.obj <- cf_gamlss_try(cf.args)
       } else {
         cf.obj <- eval(parse(text = paste("comfam(pheno.df, batch, covar.df,", cf.args, ")")))
       }

### 2024-01-25 cubic_scripts/R_scripts/featurewise_sex_bias_test.R

@@ -45,12 +45,12 @@ for (file in raw_files) {
 cent.sex_t_tests_in_feat.df <- lapply(diffs_perm.list, sex.bias.feat.t.tests, comp_multiplier=length(diffs_perm.list), feature_list=diff_list)
 
 #save
-fwrite(cent.sex_t_tests_in_feat.df, file=paste0(data_path, "/", d.type, "_featurewise_cent_sex_bias_tests.csv"))
+saveRDS(cent.sex_t_tests_in_feat.df, file=paste0(data_path, "/", d.type, "_featurewise_cent_sex_bias_tests.RDS"))
 
 #z-score t-tests
 z.sex_t_tests_in_feat.df <- lapply(diffs_perm.list, sex.bias.feat.t.tests, comp_multiplier=length(diffs_perm.list), feature_list=paste0(diff_list,".z"))
 
 #save
-fwrite(z.sex_t_tests_in_feat.df, file=paste0(data_path, "/", d.type, "_featurewise_z_sex_bias_tests.csv"))
+saveRDS(z.sex_t_tests_in_feat.df, file=paste0(data_path, "/", d.type, "_featurewise_z_sex_bias_tests.RDS"))
 
 print("DONE!")

### 2024-01-25 cubic_scripts/R_scripts/remove_extremes_single.R

@@ -60,7 +60,7 @@ rm_ext_z <- function(og.data, feat_list, col_select){
       df <- og.data %>%
         dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, prop, all_of(list)) %>% #get correct cols
         group_by(participant) %>% 
-        dplyr::filter(!any((get(pheno) < -2 | get(pheno) > 1) & dataset == "raw")) %>%
+        dplyr::filter(!any((get(pheno) < -2 | get(pheno) > 2) & dataset == "raw")) %>%
         ungroup() %>%
         dplyr::filter(dataset != "raw") #now drop raw vals
     } else if (col_select == "perm"){

### 2024-01-25 cubic_scripts/run_permutation_pipeline.sh
correcting args for running pipelines

@@ -8,13 +8,16 @@ img=/cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif #singularity ima
 base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
 og_data=$base/data/ukb_CN_data_agefilt.csv #path to original data for which site assignments should be permuted
 save_data_path=$base/data/ukb_permute
-pheno_path=$base/pheno_lists #path to .txt files listing phenotypes (global & regional)
+
 # paths to R scripts
-permute_script=$base/R_scripts/permute_sites.R
-cf_script=$base/R_scripts/combat_apply.R
-mod_script=$base/R_scripts/fit_perm_basic_mod.R
+r_base=$base/cubic_scripts/R_scripts
+permute_script=$r_base/permute_sites.R
+cf_script=$r_base/combat_apply.R
+mod_script=$r_base/fit_perm_basic_mod.R
+centile_script=$r_base/fit_centiles.R
+cent_subj_parse=$r_base/summarise_cent_subj-wise.R 
 #######################################################################################
-cd $base #to source functions correctly
+cd $base/cubic_scripts #to source functions correctly
 #######################################################################################
 # GET ARGS
 if [ $# -lt 1 ]
@@ -72,6 +75,32 @@ if ! [ -d $gamlss_dir ]
 	mkdir $gamlss_dir
 	fi
 
+#centile qsubs
+cent_bash_dir=$study_dir/centile_qsubs
+if ! [ -d $cent_bash_dir ]
+	then
+	mkdir $cent_bash_dir
+	else
+    rm -f $cent_bash_dir/*.*
+	fi
+
+#centile output csvs
+cent_save_dir=$study_dir/perm_centile_csvs
+if ! [ -d $cent_save_dir ]
+	then
+	mkdir $cent_save_dir
+	else
+    rm -f $cent_save_dir/*.*
+	fi
+
+# ... with subdir for subject-level outputs
+cent_subj_dir=$cent_save_dir/subject-wise
+if ! [ -d $cent_subj_dir ]
+	then
+	mkdir $cent_subj_dir
+	else
+    rm -f $cent_subj_dir/*.*
+	fi
 #######################################################################################
 #######################################################################################
 # SUBMIT PERMUTAITON JOBS
@@ -252,4 +281,97 @@ do
     sleep 60    # wait for 1min before detecting again
 done
 
[…]

### 2024-01-25 cubic_scripts/run_permutation_stats.sh

@@ -53,7 +53,7 @@ if ! [ -d $csv_path/subject-wise ]
 	mkdir $csv_path/subject-wise
 	fi
 # sub jobs
-for n_prop in $(seq -f "%02g" 1 11) #10 sims
+for n_prop in $(seq -f "%03g" 1 10) #10 sims
 do
 	echo "Prepping perm-$n_prop"
 	#write bash script
@@ -124,7 +124,7 @@ qsub -N perm_sex-bias -o $bash_dir/sex_bias_test_out.txt -e $bash_dir/sex_bias_t
 ## based on `qsub_rm_extremes.sh`
 r_script=$r_base/remove_extremes_single.R
 # sub jobs
-for n_prop in $(seq -f "%02g" 1 11) #11 sims
+for n_prop in $(seq -f "%03g" 1 10) #10 sims
 do
 	echo "Prepping prop-$n_prop"
 	#write bash script

### 2024-01-28 paper_main_results.Rmd
plots

@@
 options("mc.cores")
-
+```
+```{r}
 source("./cubic_scripts/R_scripts/stats_tests.R")
+# devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/stats_tests.R")
 ```
@@
 Little plot
-```{r, eval=FALSE}
+```{r}
 #get means w/in feature 
@@
       label = sub("_[^_]*_", "_", pheno))
-
+```
+
+```{r, eval=FALSE}
 dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
@@
 
+```{r, eval=FALSE}
+for (cf in c("cf.gam", "cf.gamlss")) {
+  for (pheno in vol_list_global){
+  val <- paste0("diff_", pheno)
+  med.df <- diffs.df %>%
+    dplyr::filter(dataset == cf) %>%
+    group_by(sex) %>%
+    summarize_at(val, median, na.rm=FALSE)
+
+ p <- diffs.df %>%
+    dplyr::filter(dataset == cf) %>%
+    ggplot(aes(x=.[[val]], fill = sex, color=sex)) +
+      geom_histogram(alpha=0.5, position="identity") +
+    ggtitle(paste(val, cf)) +
+    geom_vline(aes(xintercept = med.df[med.df$sex=="Female", ][[val]])) +
+    geom_vline(aes(xintercept = med.df[med.df$sex=="Male", ][[val]]))
+  print(p)
+  }
+}
+```
+
 not normally distributed, use nonparametric t test of ranks with welch correction for different variances.
@@
 ```{r, message=FALSE}
-sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list)
+sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list, ID_col="File_ID")
 
@@
     as.data.frame()
+```
+
+quick plot for median diffs
+```{r}
+sex_t_tests_in_feat.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+sex_t_tests_in_feat.df %>%
+  group_by(dataset) %>%
+  summarize(min_diff = min(median_sex_diff)*100, max_diff=max(median_sex_diff)*100)
+```
+
+```{r}
+plt.df <- diffs.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", pheno_list)) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         diff = diff*100,
+         pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)))
+```
+```{r}
+custom_strip <- strip_themed(
+     text_x = elem_list_text(angle = 90, color="black"),
+     background_x=element_blank()
+)
+
+big_violin_plt <- plt.df %>% 
+  ggplot() +
+  geom_violin(aes(x=pheno, y=diff, fill=sex, color=sex)) +
+  theme_linedraw(base_size = 11) +
+  labs(y="Centile Error", x = "Brain Feature") +
+  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines")) +
+  facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
+  ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
+
+ggsave("figs/big_sex_violin_plt.jpeg", big_violin_plt, width=12, height=6, units="in")
+```
+*need to add stats tests*
+
+```{r}
+ diffs.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", vol_list_global)) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>% #drop refA levels
+  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         diff = diff*100) %>%
+  ggplot(aes(x=pheno, y=diff, fill=sex)) +
+  geom_violin(trim = FALSE)+
+  theme_linedraw(base_size = 11) +
+  labs(y="Centile Error", x = "Brain Feature") +
+  facet_wrap(~dataset)
+```
+
+```{r}
+diffs.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", pheno_list)) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         diff = diff*100,
+         pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
+  theme_linedraw(base_size = 11) +
+  # scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  # scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Centile Error", x = "Brain Feature") +
+  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
+  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 45)) +
+  ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
 ```
@@
 
-```{r, warning=FALSE}
+```{r, message=FALSE}
 raw_centiles <- combined_preds %>%

### 2024-01-29 paper_main_results.Rmd
troubleshooting featurewise sex bias tests

@@
 ```{r, message=FALSE}
-sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list, ID_col="File_ID")
+sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list)
 
@@
 
-ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", cent.diffs.plot)
+ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", z.diffs.plot)
 ```

### 2024-01-29 paper_prop_results.Rmd
troubleshooting featurewise sex bias tests

@@
 
-#years for plotting
-ukb_cn_df <- ukb_cn_df %>%
-  mutate(age_years = age_days/365.25)
-
 #plotting info from dk atlas
@@
 ```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "_diffs.csv", full.names = TRUE)
+raw_files <- list.files(path = paste0(der_path, "/ukb_ratio"), pattern = "_diffs.csv", full.names = TRUE)
 
@@
 ```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "_subj_pred.csv", full.names = TRUE)
+raw_files <- list.files(path = paste0(der_path, "/ukb_ratio"), pattern = "_subj_pred.csv", full.names = TRUE)
 
@@
 ```{r, eval=FALSE}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "raw_predictions.csv", full.names = TRUE)
+pred.csvs <- list.files(path = paste0(der_path, "/ukb_ratio"), pattern = "raw_predictions.csv", full.names = TRUE)
 
@@
 ```{r, eval=FALSE}
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))
+
+ratio.df %>%
+  mutate(sex_site = paste(sex, sim.site, sep=".")) %>%
+  dplyr::filter(dataset=="cf") %>%
+ggplot() +
+  geom_bar(aes(x=prop, fill=sex_site)) +
+  scale_x_discrete(labels=percent_f) +
+  scale_fill_manual(values=c(Female.Balanced = "#FCC3BF", Male.Balanced = "#ADFDFF", Female.Imbalanced ="#F8766D", Male.Imbalanced="#00BFC4")) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  geom_hline(yintercept = (18800/2), linetype="dashed")+
+  theme_linedraw()
+```
+
+```{r}
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))
+
 ratio.df %>%
@@
 ggplot() +
-  geom_bar(aes(x=prop, fill=sex)) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~ sim.site)
+  geom_bar(aes(x=sim.site, fill=sex)) +
+  #theme_linedraw() +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
+  xlab("Simulated Site") +
+  ylab("Subjects") +
+  facet_wrap(~prop, nrow=1, labeller = as_labeller(percent_f))
 ```
@@
 ```{r}
-cent.t.test.df <- fread(paste0(data_path, "/ukb_ratio/full_featurewise_cent_t_tests.csv"))
+cent.t.test.df <- fread(paste0(der_path, "/ukb_ratio/full_featurewise_cent_t_tests.csv"))
 cent.t.test.df <- cent.t.test.df %>%
@@
 
+Also visualize by sex
+```{r}
+prop.global_cent_errs_plot_bysex <- prop.global_abs.diffs %>%
+  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
+         abs.diff = abs.diff*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
+  theme_linedraw(base_size = 11) +
+  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Study Site") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_x_discrete(labels=percent_f) +
+  facet_grid(sex ~ pheno) +
+  ggtitle("Magnitude errors in centile scores of global features across ComBat pipelines")
+prop.global_cent_errs_plot_bysex
+```
+
 ### Sex-Biases
@@
 ```{r}
-cent.subj.mean.test.df <- fread(paste0(data_path, "/ukb_ratio/subj.abs.mean_sex_bias_cent_t_tests.csv"))
+cent.subj.mean.test.df <- fread(paste0(der_path, "/ukb_ratio/subj.abs.mean_sex_bias_cent_t_tests.csv"))
 cent.subj.mean.test.df <- cent.subj.mean.test.df %>%
@@
 Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
-**RERUNNING TO KEEP PROP VALS AS EXPLICIT NAMES, COME BACK TO THIS**
 
 ```{r}
-cent.subj.sex.test.list <- readRDS(paste0(data_path, "/ukb_ratio/full_featurewise_cent_sex_bias_tests.RDS"))
+cent.subj.sex.test.list <- readRDS(paste0(der_path, "/ukb_ratio/full_featurewise_cent_sex_bias_tests.RDS"))
+cent.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
+```
+
+simple bar-plot to count regions w sigificant sex-biases
+
+```{r}
+cent.subj.sex.test.df %>%
+  ggplot() +
+  geom_bar(aes(x=prop, fill=sig_fdr)) +
+  facet_wrap(~dataset) +
+  scale_x_discrete(labels=percent_f) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+```
+
+visualize magnitude of significant sex-biases
+
+```{r}
+cent.subj.sex.test.df %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
+  facet_wrap(~prop)
+
+#zoom in on combat-GAM vs combat-LS
+cent.subj.sex.test.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_wrap(~prop, labeller = as_labeller(percent_f)) +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+cent.subj.sex.test.df %>%
+  group_by(dataset, prop) %>%
+  summarize(min_diff = min(median_sex_diff)*100, max_diff=max(median_sex_diff)*100)
+```
+
+visualize actual (not ABSOLUTE) errors for global pheonos by sex
+
+```{r}
+ratio.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
+         diff = diff*100) %>% #for visuals
+  ggplot() +
+  geom_violin(aes(x=prop, y=diff, fill=sex, color=sex)) +
+  theme_linedraw(base_size = 11) +
+  # scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  # scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Centile Error", x = "Imbalanced Study Site Composition") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_x_discrete(labels=percent_f) +
+  facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names)) +
+  ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
+```
+
+Plot on brains for 60% F imbalance:
+```{r}
+aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+
+cent.subj.sex.test.df2 <- cent.subj.sex.test.df %>%
+  dplyr::filter(prop == "prop-04") %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    label = sub("_[^_]*_", "_", pheno)) #for plotting
+sex_t_tests_to_plot <- left_join(cent.subj.sex.test.df2, dk.parc, by="label")
+
+# max(sex_t_tests_to_plot$median_sex_diff)*100
+# min(sex_t_tests_to_plot$median_sex_diff)*100
+
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+#var plot
+sex.var.plot.df <- sex_t_tests_to_plot %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         # median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+         #                      TRUE ~ median_sex_diff*100)) %>%
+         median_sex_diff = median_sex_diff*100) %>%
+  group_by(dataset, pheno_cat)
+
+tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="ukb_CN")
+tp$sex_diff <- 1
+
+sex.var.plot <- ggplot(sex.var.plot.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = dk, 
+             position = position_brain(hemi ~ side),
+             aes(fill = median_sex_diff)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-3, 3), name = "Sex Effect") #
+#sex.var.plot
+
+#plot global
+sex.var.plot.global.df <- sex_t_tests_to_plot %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "WM",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA)),
+        # median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+        #                       TRUE ~ median_sex_diff*100)) %>% #drop NS values
+        median_sex_diff = median_sex_diff*100) %>%
+  group_by(dataset)
+
+tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="ukb_CN")
+tp$sex_standard <- 1
+
+sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = median_sex_diff),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-3, 3),  name = "Sex Effect") #,
+# sex.var.plot.global
+
+#Combine plots - need to keep playing around to make the spacing pretty when it saves out
+
+p_sex <- ggarrange(
+  sex.var.plot.global, sex.var.plot,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.1),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+
+p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations when Imbalanced Site has 70% Female", 
+               color = "black", size = 14, hjust=.65))
+p_sexannotate
 ```
@@
 ```{r}
-noext.cent.t.test.df <- fread(paste0(data_path, "/ukb_ratio/no.ext_featurewise_cent_t_tests.csv"))
+noext.cent.t.test.df <- fread(paste0(der_path, "/ukb_ratio/no.ext_featurewise_cent_t_tests.csv"))
 noext.cent.t.test.df <- noext.cent.t.test.df %>%
@@
 ```{r}
-z.t.test.df <- fread(paste0(data_path, "/ukb_ratio/full_featurewise_z_t_tests.csv"))
+z.t.test.df <- fread(paste0(der_path, "/ukb_ratio/full_featurewise_z_t_tests.csv"))
 z.t.test.df <- z.t.test.df %>%
@@
 ```{r}
-z.subj.mean.test.df <- fread(paste0(data_path, "/ukb_ratio/subj.abs.mean_sex_bias_z_t_tests.csv"))
+z.subj.mean.test.df <- fread(paste0(der_path, "/ukb_ratio/subj.abs.mean_sex_bias_z_t_tests.csv"))
 z.subj.mean.test.df <- z.subj.mean.test.df %>%
@@
 ```{r}
-noext.z.t.test.df <- fread(paste0(data_path, "/ukb_ratio/no.ext_featurewise_z_t_tests.csv"))
+noext.z.t.test.df <- fread(paste0(der_path, "/ukb_ratio/no.ext_featurewise_z_t_tests.csv"))
 noext.z.t.test.df <- noext.z.t.test.df %>%

### 2024-01-29 cubic_scripts/R_scripts/stats_tests.R

@@ -7,7 +7,7 @@ library(broom)
 library(purrr)
 library(tidyverse)
 
-devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/ranked_welch_tests.R?token=GHSAT0AAAAAACGAYP4GBRV6DV6ZXXAUMDXMZM5SWKQ")
+devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/ranked_welch_tests.R")
 
 ##########
 ## STATS

### 2024-01-29 cubic_scripts/run_permutation_stats.sh

@@ -169,14 +169,6 @@ done
 echo "launching stats tests on data w extremes removed"
 
 #######################################################
-# RECALC SUBJ-LEVEL STATS
-## using subject-wise (list-wise) deletion, rm all subjects with ANY extreme centiles and recalc w/in subj. means
-## based on `qsub_rm_extremes.sh`
-
-#######################################################
-## CHECK FOR OUTPUTS
-
-#######################################################
 # STATS TESTS OF ERRORS W/O EXTREMES
 
 # featurewise error tests

### 2024-01-29 cubic_scripts/R_scripts/gamlss_helper_funs.R
lifespan centiles w sexMale instead of sex

@@ -505,7 +505,7 @@ get.og.data.centiles.lbcc <- function(gamlss.rds.file, og.data, get.zscores = FA
   pheno <- gamlss.obj$mu.terms[[2]]
   
   newData <- data.frame(log_age=og.data$log_age,
-                        sex=og.data$sex,
+                        sexMale=og.data$sexMale,
                         sex.age=og.data$sex.age,
                         fs_version=og.data$fs_version,
                         TBV=og.data$TBV,

### 2024-01-29 paper_main_results.Rmd
plots

@@
 #plotting info from dk atlas
-dk.parc <- dk$data %>%
+dk.parc <- dk_nomed$data %>%
   as.data.frame() %>%
@@
   distinct()
+
+#write out new dk atlas w/o medial wall
+dk_nomed <- dk
+dk_nomed$data <- dk_nomed$data %>%
+  dplyr::filter(!is.na(region) & region !="corpus callosum")
 ```
@@
   theme(plot.title = element_text(size=12)) +
-  ggtitle("Magnitude errors in centile scores of global features across ComBat pipelines")
+  ggtitle("Magnitude error in centile scores")
 ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
@@
 *need to add stats tests*
-
-```{r}
- diffs.df %>%
+```{r}
+regional_violin_plt <- plt.df %>%
+  dplyr::filter(!pheno %in% vol_list_global) %>%
+  ggplot() +
+  geom_violin(aes(x=pheno, y=diff, fill=sex, color=sex)) +
+  theme_linedraw(base_size = 11) +
+  labs(y="Centile Error", x = "Brain Feature") +
+  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines"), legend.position="bottom") +
+  facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x") +
+  ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
+```
+
+```{r}
+global_violin_sex_plt <- diffs.df %>%
   dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", vol_list_global)) %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>% #drop refA levels
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels | dataset == "cf.lm"
   pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
@@
   ggplot(aes(x=pheno, y=diff, fill=sex)) +
-  geom_violin(trim = FALSE)+
+  geom_rain(likert= TRUE,
+            point.args=list(color=NA, fill=NA),
+ boxplot.args.pos = list(position = ggpp::position_dodgenudge(x = -.07), width = .3),
+ violin.args = list(color=NA, alpha = .7))+
   theme_linedraw(base_size = 11) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom") +
   labs(y="Centile Error", x = "Brain Feature") +
@@
          median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_diff <- 1
+                                     TRUE ~ median_sex_diff*100))
 
 sex.var.plot <- ggplot(sex.var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
              position = position_brain(hemi ~ side),
@@
   #scale_fill_gradient2() +
+  # ggseg::geom_brain(data=med_wall, atlas = dk, 
+  #            position = position_brain(hemi ~ side),
+  #            aes(fill=median_sex_diff)) +
   facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5), name = "Sex Effect") #
-# sex.var.plot
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6)) #
+sex.var.plot
 ```
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork",  limits = c(-6, 6), breaks = c(-5, -2, 0, 2, 5), name = "Sex Effect") #,
+  paletteer::scale_fill_paletteer_c("grDevices::Cork",  limits = c(-6, 6), name = "Sex Effect") #,
 # sex.var.plot.global
@@
   ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))+
+  geom_bar(aes(x=sim.site, fill=sex), position="fill")+
   ggtitle("Distributions of male and female subjects across simulated sites")
@@
 
-```{r}
+```{r, eval=FALSE}
 global.cent.df <- centiles.df %>%
@@
          dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ),
-         age_years = age_days/365.25)
+         age_years = age_days/365.25,
+         label = ifelse(age_days == max(age_days), "Test", NA))
 ```
@@
 adding jitter to age (x-axis) because otherwise frankly it's hard to look at
+centiles = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
 
@@
 dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+
+tmp_labels <- wmv.df %>%
+  dplyr::filter(age_days == max(age_days) & dataset=="ukb-CN-simsites") %>%
+  pivot_longer(cols=starts_with("centile"), values_to = "value", names_to="centile_num", names_prefix = "centile_") %>%
+  mutate(centile = case_when(centile_num == 1 ~ "1%", centile_num==2 ~ "5%", centile_num==3 ~ "10%", 
+                             centile_num == 4 ~ "25%", centile_num==5 ~ "50%", centile_num==6 ~ "75%",
+                             centile_num == 7 ~ "90%", centile_num==8 ~ "95%", centile_num==9 ~ "99%")) %>%
+  dplyr::filter(centile_num %in% c(1, 2, 5, 8, 9))
 
@@
   geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_years, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
+  #geom_line(aes(x=age_years, y=centile_3, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
   geom_line(aes(x=age_years, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_years, y=centile_7, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_7, color=dataset, linetype=dataset)) +
   geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
@@
   scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
+  geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile)), nudge_x = 0.35, size=2) +
   xlab("Age (yrs)") + ylab("Volume (mm3)") +
@@
   theme(plot.title = element_text(size=12)) +
-  labs(title="WMV brain charts")
+  labs(title="White matter volume brain charts")
 wmv_centile_fan
 ```
-
-*need to add labeling of % lines - ask Ayan how he did it*
-also not sure how much i should worry about the super big female outlier
 
@@
                                    sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
-
-ggplot(ven.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$Ventricles), alpha=0.7, width=150, color="lightgray") +
+         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ),
+         age_years = age_days/365.25)
+tmp_labels <- ven.df %>%
+  dplyr::filter(age_years == max(age_years) & dataset=="ukb-CN-simsites") %>%
+  pivot_longer(cols=starts_with("centile"), values_to = "value", names_to="centile_num", names_prefix = "centile_") %>%
+  mutate(centile = case_when(centile_num == 1 ~ "1%", centile_num==2 ~ "5%", centile_num==3 ~ "10%", 
+                             centile_num == 4 ~ "25%", centile_num==5 ~ "50%", centile_num==6 ~ "75%",
+                             centile_num == 7 ~ "90%", centile_num==8 ~ "95%", centile_num==9 ~ "99%")) %>%
+  dplyr::filter(centile_num %in% c(1, 2, 5, 8, 9)) %>%
+  #repel for visuals
+  mutate(value = ifelse(centile_num == 1, value*.75,
+                           value))
+
+ven_plot <- ggplot(ven.df) +
+  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_years, y=ukb_cn_df$Ventricles),  alpha=0.4, width=.2, color="lightgray") +
   #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
-  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_3, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
+  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_7, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
   facet_wrap(~sex) + 
@@
   scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  theme_minimal() +
+  theme_linedraw(base_size = 11) +
+  geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile))) +
   xlab("Age (days)") + ylab("Volume (mm3)") +
-  theme_minimal() +
-  labs(title="Ventricle brain charts")
+  labs(title="Ventricular volume brain charts")
 ```
@@
   ncol=2, nrow=1, 
-  widths = c(1.3, 1),
+  widths = c(1.75, 1),
   #labels=c("A", "B"),
@@
 fig_1_annotate
-ggsave("figs/figure_1.jpeg", fig_1_annotate, width=10, height=5)
+ggsave("figs/figure_1.jpeg", fig_1_annotate, width=11, height=5)
 ```
@@
 ## FIGURE 2
-need to run `paper_prop_results.Rmd` to get panel B
-
-```{r}
-prop_sex_mean_cent.plot.abs <- readRDS(file="figs/mean_subj_cent_err.RDS")
-```
 
@@
 fig_2 <- ggarrange(
-  p_sex, prop_sex_mean_cent.plot.abs,
+  p_sex, global_violin_sex_plt,
   ncol=2, nrow=1, 
-  widths = c(1, 1.2),
+  widths = c(2, 1),
   #labels=c("A", "B"),
@@
 
-fig_2_annotate <- annotate_figure(fig_2, top = text_grob("Figure 2. Sex-biases in centile score errors across combat configurations", color = "black", hjust = 0, x = 0))
+fig_2_annotate <- annotate_figure(fig_2, top = text_grob("Figure 2. Sex-biases in centile score errors across ComBat configurations", color = "black", hjust = 0, x = 0))
 fig_2_annotate
-ggsave("figs/figure_2.jpeg", fig_2_annotate, width=14, height=6)
-```
+ggsave("figs/figure_2.jpeg", fig_2_annotate)
+```
+
+```{r}
+fig_2.opt2 <- ggarrange(
+  p_sex, regional_violin_plt,
+  ncol=2, nrow=1, 
+  widths = c(1, 1.5),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  # align = "hv",
+  common.legend = FALSE,
+  labels="auto"
+  )
+# fig_2
+
+fig_2.opt2_annotate <- annotate_figure(fig_2.opt2, top = text_grob("Figure 2. Sex-biases in centile score errors across ComBat configurations", color = "black", hjust = 0, x = 0))
+fig_2.opt2_annotate
+ggsave("figs/figure_2_option2.jpeg", fig_2.opt2_annotate, width=7, height=4, units="in")
+
+```

### 2024-01-30 paper_main_results.Rmd
finishing fig 1git add cubic_scripts/*git add cubic_scripts/*!

@@
 *TO-DO: SIGNIFICANCE IN PLOT*
+
+plot median on brain
+```{r}
+#get median w/in feature 
+med.diffs.long <- diffs.df %>%
+  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
+  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), median)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="median_centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+med.diffs.long2 <- left_join(med.diffs.long, dk.parc, by="label")
+max(med.diffs.long2$median_centile_abs.diff)*100
+
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+#var plot
+med.diffs.plot.df <- med.diffs.long2 %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_centile_abs.diff = median_centile_abs.diff*100)
+
+med.diffs.plot <- ggplot(med.diffs.plot.df) +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+             aes(fill = median_centile_abs.diff)) +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 7.5)) #
+med.diffs.plot
+```
+
+```{r}
+#plot global
+med.diffs.plot.global.df <- med.diffs.long2 %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_centile_abs.diff = median_centile_abs.diff*100)
+
+med.diffs.plot.global <- ggplot(med.diffs.plot.global.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = median_centile_abs.diff),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 7.5))
+med.diffs.plot.global
+```
+
+Combine plots - need to keep playing around to make the spacing pretty when it saves out
+```{r}
+p_diff <- ggarrange(
+  med.diffs.plot.global, med.diffs.plot,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.1),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+
+p_diff_annotate <- annotate_figure(p_diff)
+p_diff_annotate
+```
 
@@
 wmv_centile_fan
+```
+toy brain chart for figs
+
+```{r}
+ven.df %>%
+  dplyr::filter(sex=="Female") %>%
+  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+ggplot() +
+  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+  # geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_3, color=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_5, color=dataset), linewidth=1.5) +
+  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_7, color=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
+  facet_wrap(~dataset, nrow=1, labeller=labeller(dataset=dataset_names)) + 
+  # scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  xlab("Age (yrs)") + ylab("Brain Feature") +
+  theme_classic(base_size = 11) +
+  theme(plot.title = element_text(size=12), legend.position="none", axis.text.y=element_blank())
 ```
@@
 ```
+
+```{r}
+ven.df %>%
+  dplyr::filter(sex=="Female") %>%
+ggplot() +
+  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+  # geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_3, color=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_5, color=dataset), linewidth=1.5) +
+  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_7, color=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
+  facet_wrap(~dataset, nrow=1, labeller=labeller(dataset_names)) + 
+  # scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  xlab("Age (yrs)") + ylab("Volume (mm3)") +
+  theme_classic(base_size = 11) +
+  theme(plot.title = element_text(size=12), legend.position="none")
+```
+
 Random cortical feature

### 2024-02-02 paper_lbcc_results.Rmd
running lm on cubic

@@
 
-data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/lifespan"
+der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/lifespan"
+
 ```
 
-# QC ComBat Outputs
+#Load
 
-csvs scp'd from cubic
-
-Log-Scaled
+Combat-GAM Centiles
 
 ```{r}
-imp.sites_log.cf_cf.gam <- fread(file=paste0(data_path, "/lifespan-CN-imp-sites_log-cf_cf.gam_data.csv"))
-sum(is.na(imp.sites_log.cf_cf.gam))
-sum(imp.sites_log.cf_cf.gam < 0)
-skimr::skim(imp.sites_log.cf_cf.gam$lh_Vol_posteriorcingulate)
-min(imp.sites_log.cf_cf.gam$lh_Vol_posteriorcingulate)
+gam_centile_df <- fread(file=paste0(der_path, "/cf.gam-batch.study_predictions.csv"))
 ```
 
-```{r}
-imp.sites_log.cf_cf.gamlss <- fread(file=paste0(data_path, "/lifespan-CN-imp-sites_log-cf_cf.gamlss_data.csv"))
-sum(is.na(imp.sites_log.cf_cf.gamlss))
-skimr::skim(imp.sites_log.cf_cf.gam$log_age)
-skimr::skim(imp.sites_log.cf_cf.gam$sex)
-skimr::skim(imp.sites_log.cf_cf.gam$sex.age)
-skimr::skim(imp.sites_log.cf_cf.gam$fs_version)
-skimr::skim(imp.sites_log.cf_cf.gam$Vol_total)
-```
+Combat-LS Centiles
+
+Deltas
 
 ```{r}
-data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"
 
-pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
-vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
-vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
-sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
-ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
-
-pheno_list.z <- paste0(pheno_list, ".z")
-pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)
-z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")
-
-#also load raw data
-ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")
-
-#plotting info from dk atlas
-dk.parc <- dk$data %>%
-  as.data.frame() %>%
-  na.omit() %>%
-  dplyr::select(c(hemi, region, label)) %>%
-  distinct()
 ```

### 2024-02-02 cubic_scripts/R_scripts/lifespan_delta_lms.R

@@ -1,0 +1,115 @@
+
+##LOAD PACKAGES
+library(dplyr)
+library(data.table)
+library(tibble)
+library(broom.mixed)
+library(stats)
+
+#get args
+args <- commandArgs(trailingOnly = TRUE)
+delta_path <- as.character(args[1]) #path to delta csv
+save_path <- as.character(args[2]) #path to delta csv
+
+pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
+
+##READ IN DELTAS
+deltas_df <- fread(delta_path)
+
+#######################################
+# CENTILES
+#######################################
+
+#initialize empty df to store outputs
+tests.df <- data.frame("term" = character(),
+                   "estimate" = double(),
+                   "std.error" = double(),
+                   "statistic" = double(),
+                   "p.value" = double(),
+                   "iv" = character(),
+                   "pheno" = character())
+
+##Loop each Pheno
+for (pheno in pheno_list) {
+  # Test Sex
+  sex.df <- tidy(lm(formula=as.formula(paste0("delta.", pheno, " ~ sexMale")), data=deltas_df)) %>%
+    dplyr::filter(term =="sexMale") %>%
+    mutate(iv = "sexMale")
+  
+  # Test Age
+  age.df <- tidy(lm(formula=as.formula(paste0("delta.", pheno, " ~ age_days")), data=deltas_df)) %>%
+    dplyr::filter(term =="age_days") %>%
+    mutate(iv = "age_days")
+  
+  # Test Sex*Age
+  int.df <- tidy(lm(formula=as.formula(paste0("delta.", pheno, " ~ sexMale*age_days")), data=deltas_df)) %>%
+    dplyr::filter(term =="sexMale:age_days") %>%
+    mutate(iv = "sexMale:age_days")
+  
+  # Compile w/in phenotype
+  stopifnot(dim(sex.df) == dim(age.df) & dim(age.df) == dim(int.df))
+  pheno.df <- rbind(c(list(sex.df), list(age.df), list(int.df)))
+  pheno.df$pheno <- pheno
+  
+  # Add to other phenotypes
+  tests.df <- rbind(tests.df, pheno.df)
+}
+#FDR correct across tests/phenotypes
+result_df <- tests.df %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = nrow(tests.df)), 
+                sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
[…]

### 2024-02-02 cubic_scripts/qsub_lbcc_deltas_stats.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+#see how different predicted centiles are for each subject when ComBat-GAM or ComBatLS are used to harmonize LBCC
+
+#######################################################################################
+# SET PATHS
+img=/cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif #singularity image
+base=/cbica/home/gardnerm/combat-biovar #base path (cubic) - also path that output csvs will be saved to
+r_script=$base/cubic_scripts/R_scripts/lifespan_delta_lms.R #path to .R script
+cent_path=$base/lifespan/centile_csvs #path to centile csvs
+bash_dir=$base/lifespan/centile_qsubs
+
+#search string:
+config="batch.study"
+#######################################################################################
+cd $base/cubic_scripts #to source functions correctly
+#######################################################################################
+bash_script=$bash_dir/${config}_delta_lms.sh
+touch $bash_script
+
+echo "singularity run --cleanenv $img Rscript --save $r_script $cent_path/${config}_pred_deltas.csv $cent_path" > $bash_script
+
+#qsub bash script
+qsub -N $config -o $bash_dir/${config}_delta_lms_out.txt -e $bash_dir/${config}_delta_lms_err.txt $bash_script

### 2024-02-02 cubic_scripts/qsub_lbcc_cent_deltas.sh
troubleshooting

@@ -19,7 +19,7 @@ bash_script=$bash_dir/${config}_deltas.sh
 touch $bash_script
 
 #arg1=gam csv path, arg2=gamlss csv path, arg3=save path, arg4=search string
-echo "singularity run --cleanenv $img Rscript --save $r_script $cent_path/${config}_cf.gam_data_predictions.csv $cent_path/${config}_cf.gamlss_data_predictions.csv $cent_path $config" > $bash_script
+echo "singularity run --cleanenv $img Rscript --save $r_script $cent_path/cf.gam-${config}_predictions.csv $cent_path/cf.gamlss-${config}_predictions.csv $cent_path $config" > $bash_script
 
 #qsub bash script
 qsub -N $config -o $bash_dir/${config}_deltas_out.txt -e $bash_dir/${config}_deltas_err.txt $bash_script

### 2024-02-02 paper_prop_results.Rmd

@@
 
-ratio.df %>%
+sampling.plt <- ratio.df %>%
   dplyr::filter(dataset=="cf") %>%
+  # mutate(sim.site = case_when(sim.site == "Balanced" ~ "A",
+  #                             sim.site == "Imbalanced" ~ "B")) %>%
 ggplot() +
   geom_bar(aes(x=sim.site, fill=sex)) +
-  #theme_linedraw() +
-  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
+  theme_linedraw() +
+  theme(axis.ticks.x = element_blank(), axis.text.x=element_blank(), panel.spacing.x = unit(.1, "lines")) +
   xlab("Simulated Site") +
@@
   facet_wrap(~ pheno) +
-  ggtitle("Magnitude errors in centile scores of global features across ComBat pipelines")
+  ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
 prop.global_cent_errs_plot
@@
   facet_grid(sex ~ pheno) +
-  ggtitle("Magnitude errors in centile scores of global features across ComBat pipelines")
+  ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
 prop.global_cent_errs_plot_bysex
@@
 
-visualize actual (not ABSOLUTE) errors for global pheonos by sex
+visualize actual (not ABSOLUTE) errors for global phenos by sex
 
 ```{r}
-ratio.df %>%
+sex_err_ratio.plt <- ratio.df %>%
   dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
@@
   labs(y="Centile Error", x = "Imbalanced Study Site Composition") +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
   scale_x_discrete(labels=percent_f) +
@@
 **COME BACK TO THIS**
+
+## FIGURE 3
+```{r}
+fig_3 <- ggarrange(ggarrange(sampling.plt, prop.global_cent_errs_plot, 
+                             nrow = 2, labels = c("A", "B"),
+                             heights=c(1,2)), #nest 1st col
+          sex_err_ratio.plt,
+          ncol = 2,
+          # widths=c(1, 1.2),
+          labels = c(NA, "C")                                    
+)
+
+fig_3_annotate <- annotate_figure(fig_3, top = text_grob("Figure 3. Varying M:F ratios across site simulations", color = "black", hjust = 0, x = 0))
+fig_3_annotate
+ggsave("figs/figure_3.jpeg", fig_3_annotate, width=32, height=20, units="cm")
+```

### 2024-02-02 paper_lbcc_results.Rmd
fixing so centiles can't be 0 or 1

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr)
 #Sys.setenv("MC_CORES"=10L)
@@
 gam_centile_df <- fread(file=paste0(der_path, "/cf.gam-batch.study_predictions.csv"))
+dim(gam_centile_df)
 ```
@@
 Combat-LS Centiles
+```{r}
+gamlss_centile_df <- fread(file=paste0(der_path, "/cf.gamlss-batch.study_predictions.csv"))
+dim(gamlss_centile_df)
+```
 
@@
 ```{r}
+deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
+# skimr::skim(deltas_df)
+```
 
+Visualize
+
+```{r}
+ggplot(deltas_df, aes(x=age_days, y=delta.sGMV, color=sex)) +
+  geom_point() +
+  geom_smooth()
+
+tidy(lm(delta.sGMV ~ sexMale*age_days, data=deltas_df))
 ```

### 2024-02-02 cubic_scripts/R_scripts/gamlss_helper_funs.R

@@ -525,6 +525,15 @@ get.og.data.centiles.lbcc <- function(gamlss.rds.file, og.data, get.zscores = FA
   for (i in 1:nrow(og.data)){
     centiles[i] <- eval(call(pfun, og.data[[pheno]][[i]], mu=predModel$mu[[i]], sigma=predModel$sigma[[i]], nu=predModel$nu[[i]]))
     
+    #don't let centile = 1!
+    if (centiles[i] == 1) {
+      centiles[i] <- 0.99999999999999994 #largest number i could get w/o rounding to 1 (trial & error)
+    }
+    #don't let centile = 0!
+    if (centiles[i] == 0) {
+      centiles[i] <- 0.0000000000000000000000001 #25 dec places, should be plenty based on min centile
+    }
+    
   }
   if (get.zscores == FALSE){
     return(centiles)

### 2024-02-03 paper_main_results.Rmd
fixing big sex-differences violin plot and adding sig values

@@
 
+#write out new dk atlas w/o medial wall
+dk_nomed <- dk
+dk_nomed$data <- dk_nomed$data %>%
+  dplyr::filter(!is.na(region) & region !="corpus callosum")
 #plotting info from dk atlas
@@
   distinct()
-
-#write out new dk atlas w/o medial wall
-dk_nomed <- dk
-dk_nomed$data <- dk_nomed$data %>%
-  dplyr::filter(!is.na(region) & region !="corpus callosum")
 ```
@@
 mean.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
+  dplyr::select(sex, dataset, pheno_abs.diff.list) %>%
   dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
@@
 Plot errors for global phenotypes
+
+get stat results for global phenos:
+```{r, message=FALSE}
+full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
+full_cent_res %>%
+  dplyr::filter(pheno %in% vol_list_global) %>%
+  dplyr::filter(!grepl('_refA', comp))
+```
+
+all significant
+
 ```{r}
@@
          abs.diff = abs.diff*100) %>% #for visuals
-  ggplot() +
+  ggplot(aes(pheno, abs.diff)) +
   geom_violin(aes(x=pheno, y=abs.diff, fill=dataset, color=dataset), scale="count") +
@@
   scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #WMV
+geom_signif(y_position = c(29, 30.5, 4), xmin = c(0.7, 0.7, 1), 
+              xmax = c(1, 1.3, 1.3), annotation = c("***","***", "***"),
+              tip_length = 0, vjust = 0.6) +
+  #GMV
+  geom_signif(y_position = c(30, 31.5, 3), xmin = c(1.7, 1.7, 2), 
+              xmax = c(2, 2.3, 2.3), annotation = c("***","***", "***"),
+              tip_length = 0, vjust = 0.6) +
+  #sGMV
+  geom_signif(y_position = c(28, 29.5, 2.8), xmin = c(2.7, 2.7, 3), 
+              xmax = c(3, 3.3, 3.3), annotation = c("***","***", "***"),
+              tip_length = 0, vjust = 0.6) +
+  #Ventricles
+  geom_signif(y_position = c(24, 25.5, 7.5), xmin = c(3.7, 3.7, 4), 
+              xmax = c(4, 4.3, 4.3), annotation = c("***","***", "***"),
+              tip_length = 0, vjust = 0.6) +
   labs(y="Magnitude Centile Error", x = "Phenotype") +
@@
   ggtitle("Magnitude error in centile scores")
-ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
+global_cent_errs_plot
+# ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
 ```
@@
   paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 7.5)) #
-med.diffs.plot
+# med.diffs.plot
 ```
@@
 med.diffs.plot.global <- ggplot(med.diffs.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
   ggseg::geom_brain(atlas = aseg_cerebrum, 
@@
   paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 7.5))
-med.diffs.plot.global
+# med.diffs.plot.global
 ```
@@
   ncol=2, nrow=1, 
-  widths = c(1, 3.1),
+  widths = c(.95, 3.1),
   #labels=c("A", "B"),
@@
 
-p_diff_annotate <- annotate_figure(p_diff)
+p_diff_annotate <- annotate_figure(p_diff, top = text_grob("Magnitude Centile Errors", color = "black", hjust = 0, x = 0.06))
 p_diff_annotate
@@
 plt.df <- diffs.df %>%
-  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", pheno_list)) %>%
+  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", pheno_list)) %>% #select centile errors only
   dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
@@
          diff = diff*100,
-         pheno_cat = as.factor(case_when(
+         pheno_cat = factor(case_when(
     pheno %in% vol_list_global ~ "Global Volume",
@@
     pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_)))
-```
-```{r}
-custom_strip <- strip_themed(
-     text_x = elem_list_text(angle = 90, color="black"),
-     background_x=element_blank()
-)
+    TRUE ~ NA_character_), ordered=TRUE, levels=c("Global Volume", "Regional Volume", "Regional SA", "Regional CT"))) %>%
+    mutate(pheno = factor(pheno, ordered=TRUE, levels=c(vol_list_global, vol_list_regions, sa_list, ct_list)))#sort w/in pheno_cat
+```
+
+stat test re-do w wilcox 
+```{r}
+stat.test <- plt.df %>%
+  group_by(dataset, pheno) %>%
+  wilcox_test(diff ~ sex) %>%
+  adjust_pvalue(method = "fdr") %>%
+  add_significance("p.adj")
+```
+```{r}
+stat.test2 <- stat.test %>%
+  add_xy_position(x = "pheno", dodge = 0.8, fun="max") %>%
+  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*"))
+stat.test2 %>%
+  dplyr::filter(phen_cat=="")
+```
+
+```{r}
+# custom_strip <- strip_themed(
+#      text_x = elem_list_text(angle = 90, color="black"),
+#      background_x=element_blank()
+# )
+
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 
 big_violin_plt <- plt.df %>%
-  ggplot() +
-  geom_violin(aes(x=pheno, y=diff, fill=sex, color=sex)) +
+  ggplot(aes(x=pheno, y=diff)) +
+  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
+  # geom_violin(aes(fill=sex, color=sex)) +
   theme_linedraw(base_size = 11) +
   labs(y="Centile Error", x = "Brain Feature") +
-  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines")) +
-  facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
+  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines"), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 1, tip.length = 0, hide.ns = FALSE, remove.bracket = TRUE) +
+  # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
+  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
+  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 45)) +
   ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
-
+big_violin_plt
 ggsave("figs/big_sex_violin_plt.jpeg", big_violin_plt, width=12, height=6, units="in")
 ```
+
 *need to add stats tests*
@@
   theme_linedraw(base_size = 11) +
+  stat_compare_means(method = "wilcox.test", label="p.signif") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom") +
@@
   facet_wrap(~dataset)
+global_violin_sex_plt
 ```
@@
     TRUE ~ NA_character_))) %>% #for visuals
-  ggplot() +
+  ggplot(aes(y=diff)) +
   geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
@@
   scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
   scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
@@
   theme_linedraw(base_size = 11) +
-  theme(plot.title = element_text(size=12)) +
+  theme(plot.title = element_text(size=12), legend.position="bottom") +
   labs(title="White matter volume brain charts")
@@
 fig_1 <- ggarrange(
-  wmv_centile_fan, global_cent_errs_plot,
+  wmv_centile_fan,
+  ggarrange(global_cent_errs_plot, p_diff_annotate, ncol=1, nrow=2, heights = c(1.2, 2), common.legend = FALSE, labels=c("B", "C")),
   ncol=2, nrow=1, 
-  widths = c(1.75, 1),
-  #labels=c("A", "B"),
+  labels=c("A", NA),
+  widths=c(1.1, 1)
   #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom",
-  labels="auto"
+  # align = "hv"
   )
@@
 fig_1_annotate
-ggsave("figs/figure_1.jpeg", fig_1_annotate, width=11, height=5)
+ggsave("figs/figure_1.jpeg", fig_1_annotate, width=11, height=7, units="in")
 ```

### 2024-02-03 paper_main_results.Rmd
finalizing big violin plot yay

@@
   mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*"))
-stat.test2 %>%
-  dplyr::filter(phen_cat=="")
 ```
@@
 dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+addline_format <- function(x,...){
+    gsub('\\s','\n',x)
+}
 
@@
   theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines"), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 1, tip.length = 0, hide.ns = FALSE, remove.bracket = TRUE) +
+  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 3, tip.length = 0, hide.ns = TRUE, remove.bracket = TRUE) +
   # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
@@
   guides(x = ggh4x::guide_axis_nested(delim = "&")) +
-  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 45)) +
+  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
+  labs(fill = "Sex", color="Sex") +
   ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")

### 2024-02-03 paper_main_results.Rmd
cleaning up unnecesarry plots

@@
   add_xy_position(x = "pheno", dodge = 0.8, fun="max") %>%
-  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*"))
-```
-
-```{r}
-# custom_strip <- strip_themed(
-#      text_x = elem_list_text(angle = 90, color="black"),
-#      background_x=element_blank()
-# )
-
+  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*")) #simplify significance labels
+```
+
+```{r}
 dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
@@
 ggsave("figs/big_sex_violin_plt.jpeg", big_violin_plt, width=12, height=6, units="in")
-```
-
-*need to add stats tests*
-```{r}
-regional_violin_plt <- plt.df %>%
-  dplyr::filter(!pheno %in% vol_list_global) %>%
-  ggplot() +
-  geom_violin(aes(x=pheno, y=diff, fill=sex, color=sex)) +
-  theme_linedraw(base_size = 11) +
-  labs(y="Centile Error", x = "Brain Feature") +
-  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines"), legend.position="bottom") +
-  facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x") +
-  ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
-```
-
-```{r}
-global_violin_sex_plt <- diffs.df %>%
-  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", vol_list_global)) %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels | dataset == "cf.lm"
-  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         diff = diff*100) %>%
-  ggplot(aes(x=pheno, y=diff, fill=sex)) +
-  geom_rain(likert= TRUE,
-            point.args=list(color=NA, fill=NA),
- boxplot.args.pos = list(position = ggpp::position_dodgenudge(x = -.07), width = .3),
- violin.args = list(color=NA, alpha = .7))+
-  theme_linedraw(base_size = 11) +
-  stat_compare_means(method = "wilcox.test", label="p.signif") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom") +
-  labs(y="Centile Error", x = "Brain Feature") +
-  facet_wrap(~dataset)
-global_violin_sex_plt
-```
-
-```{r}
-diffs.df %>%
-  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", pheno_list)) %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
-  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         diff = diff*100,
-         pheno_cat = as.factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_))) %>% #for visuals
-  ggplot(aes(y=diff)) +
-  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
-  theme_linedraw(base_size = 11) +
-  # scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  # scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Centile Error", x = "Brain Feature") +
-  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
-  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
-  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 45)) +
-  ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
 ```

### 2024-02-04 paper_lbcc_results.Rmd
fixing big sex-differences violin plot and adding sig values

@@
 source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
+
+#ATLAS
 aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+#write out new dk atlas w/o medial wall
+dk_nomed <- dk
+dk_nomed$data <- dk_nomed$data %>%
+  dplyr::filter(!is.na(region) & region !="corpus callosum")
+#plotting info from dk atlas
+dk.parc <- dk_nomed$data %>%
+  as.data.frame() %>%
+  na.omit() %>%
+  dplyr::select(c(hemi, region, label)) %>%
+  distinct()
 
+#PHENO LISTS
+pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
+vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
+vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
+sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
+ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
+
+#PATH
 der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/lifespan"
-
 ```
@@
 # skimr::skim(deltas_df)
+cent_deltas_df <- deltas_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
+  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
 ```
 
-Visualize
+#Clean & EDA
+
+Sample
+```{r}
+sample_boxplot <- ggplot(cent_deltas_df, aes(y=reorder(study, -age_years), x=age_years, color=sex)) +
+  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
+  theme_linedraw(base_size = 11) +
+  labs(color="Sex", x = "Age (Years)", y = "Study")
+
+sample_barplot <- ggplot(cent_deltas_df) +
+  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=reorder(study, -age_years), fill=sex)) + 
+  scale_x_continuous(labels = scales::percent) +
+  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
+  theme_linedraw(base_size = 11) +
+  theme(axis.ticks.y = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) +
+  labs(color="Sex", x = "% Study")
+
+ggarrange(sample_boxplot, sample_barplot, 
+          nrow=1, ncol=2, 
+          align = "h", 
+          common.legend = TRUE,
+          widths=c(2, 1))
+```
 
@@
   geom_smooth()
+```
 
-tidy(lm(delta.sGMV ~ sexMale*age_days, data=deltas_df))
+```{r}
+cent_deltas_df_long <- cent_deltas_df %>%
+  pivot_longer(cols=starts_with("delta."), names_to="pheno", names_prefix = "delta.", values_to = "delta")
+
+cent_deltas_df_long %>%
+  ggplot() +
+  geom_histogram(aes(x=delta, color=study, fill=study))
 ```
+
+```{r}
+delta_summary.df <- cent_deltas_df_long %>%
+  group_by(sex, study) %>%
+  summarise(mean_delta = mean(delta),
+            mean_abs.delta = mean(abs(delta)),
+            max_abs.delta=max(abs(delta)),
+            sex_n = n())
+
+abs_delta_summary.df <- cent_deltas_df %>%
+  pivot_longer(cols=starts_with("delta."), names_to="pheno", names_prefix = "delta.", values_to = "delta") %>%
+  mutate(abs.delta = abs(delta)) %>%
+  group_by(study) %>%
+  summarise(study_mean_abs.delta=mean(abs.delta),
+            study_mean_delta=mean(delta),
+            mean_age=mean(age_days),
+            study_max_abs.delta=max(abs.delta),
+            study_n=n())
+```
+
+```{r}
+delta_sum.df <- full_join(delta_summary.df, abs_delta_summary.df) %>%
+  mutate(sex_prop=sex_n/study_n,
+         majority = ifelse(sex_prop > .5, 1, 0))
+
+delta_sum.df_piv <- delta_sum.df %>%
+  pivot_wider(id_cols=c(study, study_n, mean_age, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta, max_abs.delta))
+
+#Merge back female prop
+to_merge <- delta_sum.df_piv %>%
+  dplyr::select(study, sex_prop_Female)
+
+delta_sum.df2 <- full_join(delta_sum.df, to_merge)
+```
+
+plots!
+
+delta ~ sex-imbalance
+
+```{r}
+delta_sum.df2 %>%
+  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=sex)) +
+  geom_count(aes(size=study_n)) +
+  geom_smooth(method="lm")
+
+delta_sum.df2 %>%
+  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=study, shape=sex)) +
+  geom_point() +
+  geom_smooth(aes(color=sex, linetype=sex), method="lm")
+
+delta_sum.df2 %>%
+  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=sex)) +
+  geom_point() +
+  geom_smooth(aes(color=sex, linetype=sex), method="lm")
+
+delta_sum.df %>%
+  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
+  geom_count(aes(size=sex_prop), alpha=0.5) +
+  geom_line(aes(group=study))
+
+delta_sum.df %>%
+  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
+  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
+  geom_line(aes(group=study))
+
+delta_sum.df %>%
+  ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age)) +
+  geom_count(aes(size=sex_n)) +
+  geom_smooth(method="lm") +
+  facet_wrap(~sex)
+
+delta_sum.df_piv %>%
+  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta, color=mean_age)) +
+  geom_count(aes(size=study_n)) +
+  geom_smooth(method="lm")
+```
+
+delta ~ age
+
+```{r}
+delta_sum.df %>%
+  ggplot(aes(x=mean_age, y=mean_abs.delta, color=sex)) +
+  geom_count(aes(size=study_n)) +
+  geom_smooth(method="lm")
+
+delta_sum.df_piv %>%
+  ggplot(aes(x=mean_age, y=(study_mean_abs.delta*100))) +
+  geom_count(aes(size=study_n, color=sex_prop_Female)) +
+  paletteer::scale_color_paletteer_c("grDevices::Oslo", name = "Proportion \nFemale") +
+  geom_smooth(method="lm")
+```
+
+#Across-Feature Mean Results
+
+```{r}
+# age
+summary(lm(study_mean_abs.delta ~ mean_age, data=delta_sum.df_piv))
+summary(lm(study_mean_delta ~ mean_age, data=delta_sum.df_piv))
+```
+
+are studies w younger subjects changing more because there are fewer of them?
+check for association between study size and age
+```{r}
+summary(lm(mean_age ~ study_n, data=delta_sum.df_piv))
+fit <- lm((study_mean_abs.delta*100) ~ mean_age + study_n, data=delta_sum.df_piv)
+summary(fit)
+sjPlot::plot_model(fit, type = "pred", terms = "mean_age")
+
+summary(lm((study_mean_abs.delta*100) ~ mean_age + study_n + sex_prop_Female, data=delta_sum.df_piv))
+```
+
+```{r}
+cent_deltas_df %>%
+  ggplot() +
+  geom_histogram(aes(x=age_days))
+```
+
+```{r}
+# %female
+summary(lm(study_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_piv))
+summary(lm(study_mean_delta ~ sex_prop_Female, data=delta_sum.df_piv))
+```
+
+```{r}
+# age*%female
+summary(lm(study_mean_abs.delta ~ mean_age*sex_prop_Female, data=delta_sum.df_piv))
+summary(lm(study_mean_delta ~ mean_age*sex_prop_Female, data=delta_sum.df_piv))
+```
+
+#Featurewise Results
+
+Linear models
+```{r}
+cent_delta_lms.df <- fread(file=paste0(der_path, "/batch.study_pred_cent_delta_lms.csv"))
+n_unique(cent_delta_lms.df$pheno)
+```
+
+summarise findings
+
+```{r}
+cent_delta_lms.df %>%
+  dplyr::filter(term=="sexMale") %>%
+  dplyr::summarize(n_sig_regions = sum(sig_fdr),
+                   n_regions=nrow(.),
+              mean_est = mean(estimate)) %>%
+    ungroup() %>%
+    as.data.frame()
+```
+
+```{r}
+cent_delta_lms.df %>%
+  dplyr::filter(term=="age_days") %>%
+  dplyr::summarize(n_sig_regions = sum(sig_fdr),
+                   n_regions=nrow(.),
+              mean_est = mean(estimate)) %>%
+    ungroup() %>%
+    as.data.frame()
+```
+```{r}
+cent_delta_lms.df %>%
+  dplyr::filter(term=="sexMale:age_days") %>%
+  dplyr::summarize(n_sig_regions = sum(sig_fdr),
+                   n_regions=nrow(.),
+              mean_est = mean(estimate)) %>%
+    ungroup() %>%
+    as.data.frame()
+```
+
+```{r}
+sig.cent_delta_lms.df <- cent_delta_lms.df %>%
+  dplyr::filter(sig_fdr==TRUE)
+n_unique(sig.cent_delta_lms.df$pheno)
+
+table(sig.cent_delta_lms.df$term, sig.cent_delta_lms.df$pheno)
+
+max(sig.cent_delta_lms.df$estimate)*100
+min(sig.cent_delta_lms.df$estimate)*100
+```
+
+## Plots
+
+```{r}
+deltas_df %>%
+  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
+  geom_point(alpha=0.4) +
+  geom_smooth(method="lm")
+
+deltas_df %>%
+  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
+  geom_histogram(aes(y=after_stat(c(
+      count[group==1]/sum(count[group==1]),
+      count[group==2]/sum(count[group==2])
+    )*100)), alpha=0.4, position="identity", bins=60) +
+  ylab("Percentage")
+
+deltas_df %>%
+  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm")
+
+deltas_df %>%
+  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
+  geom_point(alpha=0.4) +
+  geom_smooth(method="lm")
+```
+
+Plot betas (*100 for centile space) on brain!
+
+```{r}
+cent_delta_lms.df_toplt <- cent_delta_lms.df %>%
+   mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    label = sub("_[^_]*_", "_", pheno)) #for plotting
+```
+
+**Figure out what features didn't converge & remove them from brain atlas** - seems to be happening automatically...?
+
+RERUN MODELS WITH AGE IN YEARS!!!
+
+```{r}
+term_names <- c(sexMale = "Sex", age_days = "Age (Days)", 'sexMale:age_days' = "Sex*Age (Days)")
+
+#Regional
+regional.delta_plt <- cent_delta_lms.df_toplt %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
+                                     TRUE ~ estimate*100)) %>%
+  ggplot() +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+             aes(fill = estimate)) +
+  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names))) + 
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))
+
+#Global
+global.delta_plt <- cent_delta_lms.df_toplt %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
+                                     TRUE ~ estimate*100),
+         tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "WM",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA))) %>%
+  dplyr::select(!any_of(c("label", "region", "hemi"))) %>% #drop cols corresponding to dk regions
+  ggplot() +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             side = c("coronal"),
+             aes(fill = estimate)) +
+  facet_grid(term ~ pheno_cat) + 
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))
+```

### 2024-02-05 paper_lbcc_results.Rmd
figure 4

@@
             mean_age=mean(age_days),
+            mean_age_yrs=mean(age_days)/365.25,
             study_max_abs.delta=max(abs.delta),
@@
 delta_sum.df_piv <- delta_sum.df %>%
-  pivot_wider(id_cols=c(study, study_n, mean_age, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta, max_abs.delta))
+  pivot_wider(id_cols=c(study, study_n, mean_age, mean_age_yrs, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta, max_abs.delta))
 
@@
 delta_sum.df %>%
-  ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age)) +
+  ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age_yrs)) +
   geom_count(aes(size=sex_n)) +
@@
 delta_sum.df_piv %>%
-  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta, color=mean_age)) +
+  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta, color=mean_age_yrs)) +
   geom_count(aes(size=study_n)) +
-  geom_smooth(method="lm")
-```
-
-delta ~ age
-
-```{r}
-delta_sum.df %>%
-  ggplot(aes(x=mean_age, y=mean_abs.delta, color=sex)) +
-  geom_count(aes(size=study_n)) +
-  geom_smooth(method="lm")
-
-delta_sum.df_piv %>%
-  ggplot(aes(x=mean_age, y=(study_mean_abs.delta*100))) +
-  geom_count(aes(size=study_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("grDevices::Oslo", name = "Proportion \nFemale") +
   geom_smooth(method="lm")
@@
 # age
-summary(lm(study_mean_abs.delta ~ mean_age, data=delta_sum.df_piv))
-summary(lm(study_mean_delta ~ mean_age, data=delta_sum.df_piv))
+summary(lm(study_mean_abs.delta ~ mean_age_yrs, data=delta_sum.df_piv))
+summary(lm(study_mean_delta ~ mean_age_yrs, data=delta_sum.df_piv))
+```
+
+delta ~ age
+
+```{r}
+delta_sum.df %>%
+  ggplot(aes(x=mean_age_yrs, y=mean_abs.delta, color=sex)) +
+  geom_count(aes(size=study_n)) +
+  geom_smooth(method="lm")
+
+age_effects.plt <- delta_sum.df_piv %>%
+  ggplot(aes(x=mean_age_yrs, y=(study_mean_abs.delta*100))) +
+  geom_count(aes(size=study_n, color=sex_prop_Female)) +
+  paletteer::scale_color_paletteer_c("grDevices::Oslo", name = "Proportion \nFemale") +
+  geom_smooth(method="lm", color="#060B14FF") +
+  theme_linedraw(base_size = 11) +
+  labs(size="Study N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores")
 ```
@@
 ```{r}
-summary(lm(mean_age ~ study_n, data=delta_sum.df_piv))
-fit <- lm((study_mean_abs.delta*100) ~ mean_age + study_n, data=delta_sum.df_piv)
+summary(lm(mean_age_yrs ~ study_n, data=delta_sum.df_piv))
+fit <- lm((study_mean_abs.delta*100) ~ mean_age_yrs + study_n, data=delta_sum.df_piv)
 summary(fit)
-sjPlot::plot_model(fit, type = "pred", terms = "mean_age")
+sjPlot::plot_model(fit, type = "pred", terms = "mean_age_yrs")
 
-summary(lm((study_mean_abs.delta*100) ~ mean_age + study_n + sex_prop_Female, data=delta_sum.df_piv))
+summary(lm((study_mean_abs.delta*100) ~ mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_piv))
 ```
@@
 
+delta ~ % female
+
 ```{r}
@@
 summary(lm(study_mean_delta ~ sex_prop_Female, data=delta_sum.df_piv))
+summary(lm(study_mean_delta ~ sex_prop_Female + study_n, data=delta_sum.df_piv)) #add study size covariate
+summary(lm(study_mean_delta ~ sex_prop_Female + study_n + mean_age, data=delta_sum.df_piv)) #add study size & age covariates
+```
+
+```{r}
+delta_sum.df %>%
+  ggplot(aes(x=sex, y=(mean_delta*100), color=study, fill=study)) +
+  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
+  geom_line(aes(group=study))
+
+sex_effects.plt <- delta_sum.df_piv %>%
+  ggplot(aes(x=sex_prop_Female, y=(study_mean_delta*100))) +
+  geom_count(aes(size=study_n, color=mean_age_yrs)) +
+  geom_smooth(method="lm", color="#006027FF") +
+  paletteer::scale_color_paletteer_c("grDevices::Greens 2", name = "Mean Age \n(Years)") +
+  theme_linedraw(base_size = 11) +
+  labs(size="Study N", x = "Proportion Female", y = "Mean Difference in Centile Scores")
+
+delta_sum.df %>%
+  ggplot(aes(x=sex, y=(mean_delta*100))) +
+  geom_boxplot()
 ```
@@
 # age*%female
-summary(lm(study_mean_abs.delta ~ mean_age*sex_prop_Female, data=delta_sum.df_piv))
-summary(lm(study_mean_delta ~ mean_age*sex_prop_Female, data=delta_sum.df_piv))
+summary(lm(study_mean_abs.delta ~ mean_age_yrs*sex_prop_Female, data=delta_sum.df_piv))
+summary(lm(study_mean_delta ~ mean_age_yrs*sex_prop_Female, data=delta_sum.df_piv))
 ```
@@
 cent_delta_lms.df %>%
-  dplyr::filter(term=="age_days") %>%
+  dplyr::filter(term=="age_yrs") %>%
   dplyr::summarize(n_sig_regions = sum(sig_fdr),
@@
 ```
+
 ```{r}
 cent_delta_lms.df %>%
-  dplyr::filter(term=="sexMale:age_days") %>%
+  dplyr::filter(term=="sexMale:age_yrs") %>%
   dplyr::summarize(n_sig_regions = sum(sig_fdr),
@@
 ```{r}
-deltas_df %>%
+cent_deltas_df %>%
   ggplot(aes(x=age_days, y=(delta.GMV*100))) +
@@
 
-deltas_df %>%
+cent_deltas_df %>%
   ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
@@
 
-deltas_df %>%
+cent_deltas_df %>%
   ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
@@
 
-deltas_df %>%
+cent_deltas_df %>%
   ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
@@
 
-**Figure out what features didn't converge & remove them from brain atlas** - seems to be happening automatically...?
-
-RERUN MODELS WITH AGE IN YEARS!!!
-
 ```{r}
-term_names <- c(sexMale = "Sex", age_days = "Age (Days)", 'sexMale:age_days' = "Sex*Age (Days)")
+term_names <- c(sexMale = "Sex", age_yrs = "Age", 'sexMale:age_yrs' = "Sex*Age")
 
@@
   facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names))) + 
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
   paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))
@@
              aes(fill = estimate)) +
-  facet_grid(term ~ pheno_cat) + 
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names)), switch = 'y') + 
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
   paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))
 ```
+
+#FIGURE 4
+
+```{r}
+# study demographics
+fig4 <- ggarrange(
+  ggarrange(
+ggarrange(sample_boxplot, sample_barplot,
+          nrow=1, ncol=2, 
+          align = "h", 
+          common.legend = TRUE,
+          legend="bottom",
+          widths=c(2, 1)),
+
+# featurewise effects
+ggarrange(global.delta_plt, regional.delta_plt,
+          nrow=1, ncol=2, 
+          align = "hv", 
+          common.legend = TRUE,
+          legend = "bottom",
+          widths=c(1, 2.5)
+          ),
+nrow=2, ncol=1,
+heights=c(1.5, 1),
+labels=c("A", "C")
+),
+# mean effects
+ggarrange(age_effects.plt,
+sex_effects.plt,
+nrow=2, ncol=1,
+align="hv",
+labels=c("B", "D")),
+common.legend=FALSE,
+nrow=1, ncol=2,
+widths=c(1, 1.3)
+)
+
+fig4_annotate <- annotate_figure(fig4, top = text_grob("Figure 4.
+Differences in centile scores of lifespan consortium data when harmonized with ComBat-GAM and ComBatLS", color = "black", hjust = 0, x = 0))
+fig4_annotate
+ggsave("figs/figure_4.jpeg", fig4_annotate, width=12, height=10, units="in")
+```

### 2024-02-05 eda/lifespan_eda.Rmd
troubleshooting combat
figs

@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+ggplot(cn.df.site) +
+  geom_bar(aes(x=site, fill=sex), position="dodge") + 
+  theme(axis.text.x = element_blank()) +
+  facet_wrap(~study, scales="free")
 ```
@@
 ```
+
+## Combat Troubleshooting
+having some issues with combat on cubic, trying locally.

### 2024-02-05 paper_main_results.Rmd
figs

@@
 
-*TO-DO: SIGNIFICANCE IN PLOT*
-
 plot median on brain
@@
 med.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
+  dplyr::select(sex, dataset, pheno_abs.diff.list) %>%
   dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
@@
   labs(fill = "Sex", color="Sex") +
-  ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
+  ggtitle("Male and female centile errors")
 big_violin_plt
@@
   dplyr::filter(pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
@@
   dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
+  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
@@
 ```
-toy brain chart for figs
-
-```{r}
-ven.df %>%
-  dplyr::filter(sex=="Female") %>%
-  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-ggplot() +
-  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-  # geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
-  # geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_years, y=centile_3, color=dataset)) +
-  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_years, y=centile_5, color=dataset), linewidth=1.5) +
-  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
-  geom_line(aes(x=age_years, y=centile_7, color=dataset)) +
-  # geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
-  # geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
-  facet_wrap(~dataset, nrow=1, labeller=labeller(dataset=dataset_names)) + 
-  # scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
-  xlab("Age (yrs)") + ylab("Brain Feature") +
-  theme_classic(base_size = 11) +
-  theme(plot.title = element_text(size=12), legend.position="none", axis.text.y=element_blank())
-```
-
 Look at other phenotypes
@@
 
+## METHODS FIG
+
+toy brain chart for figs
+
+```{r}
+ven.df %>%
+  dplyr::filter(sex=="Female") %>%
+  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+ggplot() +
+  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
+  # geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_3, color=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_5, color=dataset), linewidth=1.5) +
+  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
+  geom_line(aes(x=age_years, y=centile_7, color=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
+  # geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
+  facet_wrap(~dataset, nrow=1, labeller=labeller(dataset=dataset_names)) + 
+  # scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  xlab("Age (yrs)") + ylab("Brain Feature") +
+  theme_classic(base_size = 11) +
+  theme(plot.title = element_text(size=12), legend.position="none", axis.text.y=element_blank())
+```
+
+toy error distributions
+```{r}
+cf.df <- data.frame("err" = c(rnorm(100, -3, 10)),
+          "df" = c(rep("cf.lm", 100)))
+cf.gam.df <- data.frame("err" = c(rnorm(100, 1, 5)),
+          "df" = c(rep("cf.gam", 100)))
+cf.gamlss.df <- data.frame("err" = c(rnorm(100, 0, 3)),
+          "df" = c(rep("cf.gamlss", 100)))
+
+rbind(cf.df, cf.gam.df, cf.gamlss.df) %>%
+  ggplot() +
+  geom_histogram(aes(x=err, fill=df), color="black", alpha=0.7, position="identity") +
+  scale_fill_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme_classic(base_size=20) +
+  xlab("Centile Error") + 
+  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(), axis.line.y = element_blank()) +
+  scale_x_continuous(labels=c('', '', '', 0, '', '', ''))
+```
+
 ## FIGURE 1
@@
 ```{r}
+blank <- ggplot() + theme_void()
+
 fig_2 <- ggarrange(
-  p_sex, global_violin_sex_plt,
+  ggarrange(blank, 
+            annotate_figure(
+              ggarrange(sex.var.plot.global, sex.var.plot,
   ncol=2, nrow=1, 
-  widths = c(2, 1),
+  widths = c(1, 3.25),
+  hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "right"
+  ), top = text_grob("Differences in males' and females' median centile errors", color = "black", hjust = 0, x = 0)),
+  blank,
+  ncol=3, nrow=1,
+  widths=c(.5, 1, .5)),
+  big_violin_plt,
+  ncol=1, nrow=2, 
   #labels=c("A", "B"),
@@
   # align = "hv",
+  heights=c(1, 1.8),
   common.legend = FALSE,
-  labels="auto"
+  labels=c("A", "B")
   )
@@
 fig_2_annotate
-ggsave("figs/figure_2.jpeg", fig_2_annotate)
-```
-
-```{r}
-fig_2.opt2 <- ggarrange(
-  p_sex, regional_violin_plt,
-  ncol=2, nrow=1, 
-  widths = c(1, 1.5),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  # align = "hv",
-  common.legend = FALSE,
-  labels="auto"
-  )
-# fig_2
-
-fig_2.opt2_annotate <- annotate_figure(fig_2.opt2, top = text_grob("Figure 2. Sex-biases in centile score errors across ComBat configurations", color = "black", hjust = 0, x = 0))
-fig_2.opt2_annotate
-ggsave("figs/figure_2_option2.jpeg", fig_2.opt2_annotate, width=7, height=4, units="in")
-
-```
+ggsave("figs/figure_2.jpeg", fig_2_annotate, width=11.5, height=10, units="in")
+```

### 2024-02-05 paper_prop_results.Rmd
figs

@@
 Plot errors for global phenotypes
+
+trying to add significance testing
+```{r}
+cent.t.test.df %>%
+
+stat.test <- ratio.df(
+  len ~ dose, data = ToothGrowth, group.by = "supp",
+  method = "t.test", ref.group = "0.5"
+)
+```
+
 ```{r}
@@
   ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
-prop.global_cent_errs_plot_bysex
+# prop.global_cent_errs_plot_bysex
 ```

### 2024-02-05 paper_prop_results.Rmd
figured out facetted significance markers for sex-bias plots

@@
 ```{r}
+#trying to add stats test
+global_sex_diffs <- cent.subj.sex.test.df %>%
+  dplyr::filter(pheno %in% vol_list_global & dataset %in% c("cf.gam", "cf.gamlss")) %>%
+  mutate(pheno_prop = paste(pheno, prop, sep="&"))
+
+global_sex_errs.df <- ratio.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff")
+
+wilcox_tbl <- global_sex_errs.df %>%
+  mutate(pheno_prop = paste(pheno, prop, sep="&"),
+         pheno = factor(pheno, levels = c("GMV", "sGMV", "Ventricles", "WMV"), ordered = TRUE),
+         diff=diff*100) %>%
+  arrange(pheno, prop) %>%
+  group_by(dataset, pheno, prop) %>%
+  wilcox_test(diff ~ sex) %>%
+  add_xy_position(x="prop", fun="max")
+
+stat_merge <- full_join(wilcox_tbl, global_sex_diffs)
+stat_merge2 <- stat_merge %>%
+  add_significance("p.val_fdr") %>% #use actual cubic test (nonparametric, full fdr correction) for significance labels
+  mutate(p.val_fdr.signif = ifelse(p.val_fdr.signif == "ns", "", "*")) 
+table(stat_merge2$p.val_fdr.signif, stat_merge2$sig_fdr)
+```
+
+```{r}
 sex_err_ratio.plt <- ratio.df %>%
   dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
   pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
-         diff = diff*100) %>% #for visuals
+         pheno = factor(pheno, levels = c("GMV", "sGMV", "Ventricles", "WMV"), ordered = TRUE),
+         diff = diff*100,
+         pheno_prop = paste(pheno, prop, sep="&")) %>% #for visuals
   ggplot() +
@@
   labs(y="Centile Error", x = "Imbalanced Study Site Composition") +
+  stat_pvalue_manual(stat_merge2, label = "p.val_fdr.signif", label.size = 3, tip.length = 0, hide.ns = TRUE, remove.bracket = TRUE) +
+  scale_x_discrete(labels=percent_f) +
+  # facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2, strip.position = "right") +
+  # guides(x = ggh4x::guide_axis_nested(delim = "&", inv=TRUE)) +
   theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
-  scale_x_discrete(labels=percent_f) +
-  facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names)) +
+  facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names), scales="free_x") +
   ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
+sex_err_ratio.plt
 ```

### 2024-02-05 paper_prop_results.Rmd
sig. figs for abs cent err

@@
 
-trying to add significance testing
+add significance testing
 ```{r}
-cent.t.test.df %>%
+cent.t.test.df_global <- fread(paste0(der_path, "/ukb_ratio/full_featurewise_cent_t_tests_global.csv"))
 
-stat.test <- ratio.df(
-  len ~ dose, data = ToothGrowth, group.by = "supp",
-  method = "t.test", ref.group = "0.5"
-)
+cent.t.test.df_global_filt <- cent.t.test.df_global %>%
+  dplyr::filter(comp %in% c("cf.gamlss_cf.gam", "cf.lm_cf.gam", "cf.lm_cf.gamlss"))
+
+cent.t.test.df_global_filt %>%
+  dplyr::filter(sig_fdr == FALSE)
 ```
+
+```{r}
+prop.global_abs.diffs <- ratio.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global)) %>%
+  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
+  mutate(abs.diff=abs.diff*100) %>%
+  arrange(pheno, prop) %>%
+  as.data.frame()
+
+# mean_comp_tbl <- ggpubr::compare_means(abs.diff ~ dataset, data=prop.global_abs.diffs, group.by = c("pheno", "prop")) 
+# unique(mean_comp_tbl$group1)
+# unique(mean_comp_tbl$group2)
+mean_comp_tbl <- prop.global_abs.diffs %>%
+  group_by(pheno, prop) %>%
+  pairwise_wilcox_test(abs.diff ~ dataset, comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss")),
+                       p.adjust.method = "fdr") %>%
+  add_xy_position(x="prop", fun="max") %>%
+  add_significance("p.adj")
+
+#check ns results
+mean_comp_tbl %>%
+  dplyr::filter(p.adj.signif == "ns")
+
+mean_comp_tbl2 <- mean_comp_tbl %>%
+  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*"))
+mean_comp_tbl2[(mean_comp_tbl2$pheno=="WMV" & mean_comp_tbl2$group1=="cf.gam"), "p.adj.signif"] <- "*" #Overwrite result that doesnt match
+mean_comp_tbl2 %>%
+  dplyr::filter(p.adj.signif == "ns")
+```
+results match! just gotta proceed as is
 
@@
 
-prop.global_abs.diffs <- ratio.df %>%
-  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global)) %>%
-  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")
-
 prop.global_cent_errs_plot <- prop.global_abs.diffs %>%
@@
   mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
-         abs.diff = abs.diff*100) %>% #for visuals
+         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE)) %>%
   ggplot() +
@@
   scale_x_discrete(labels=percent_f) +
+  stat_pvalue_manual(mean_comp_tbl2, label = "p.adj.signif", label.size = 3, tip.length = 0) +
+  # ggsignif::geom_signif(data=mean_comp_tbl2) +
   facet_wrap(~ pheno) +

### 2024-02-07 paper_prop_results.Rmd
yet again formatting significance

@@
   ylab("Subjects") +
+  labs(fill="Sex") +
   facet_wrap(~prop, nrow=1, labeller = as_labeller(percent_f))
@@
 # unique(mean_comp_tbl$group2)
-mean_comp_tbl <- prop.global_abs.diffs %>%
+mean_comp_tbl_calc <- prop.global_abs.diffs %>%
   group_by(pheno, prop) %>%
   pairwise_wilcox_test(abs.diff ~ dataset, comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss")),
-                       p.adjust.method = "fdr") %>%
-  add_xy_position(x="prop", fun="max") %>%
+                       p.adjust.method = "fdr")
+
+mean_comp_tbl <- mean_comp_tbl_calc %>%
+  mutate(comp = factor(paste(group1, group2, sep="_"), ordered=TRUE, levels=c("cf.gamlss_cf.lm", "cf.gam_cf.lm", "cf.gam_cf.gamlss"))) %>%
+  arrange(prop, comp) %>%
+  add_xy_position(x="prop", fun="max", comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss"))) %>%
   add_significance("p.adj")
@@
 
-mean_comp_tbl2 <- mean_comp_tbl %>%
-  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*"))
-mean_comp_tbl2[(mean_comp_tbl2$pheno=="WMV" & mean_comp_tbl2$group1=="cf.gam"), "p.adj.signif"] <- "*" #Overwrite result that doesnt match
-mean_comp_tbl2 %>%
+#get max w/in each pheno, prop & dataset
+max_abs.errs <- prop.global_abs.diffs %>%
+  group_by(pheno, prop, dataset) %>%
+  summarize(max_err = max(abs.diff)) %>%
+  ungroup() %>%
+  pivot_wider(id_cols=c(pheno, prop), names_from=dataset, values_from=max_err, names_prefix="max_")
+
+mean_comp_tbl2 <- full_join(mean_comp_tbl, max_abs.errs)
+
+mean_comp_tbl3 <- mean_comp_tbl2 %>%
+  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*"),
+         xmax=case_when(group1=="cf.gamlss" | group2== "cf.gamlss" ~ (x+.266667),
+                        TRUE ~  x),
+         xmin=case_when(group2=="cf.lm" ~ (x-.266667),
+                        TRUE ~  x),
+         y.position=case_when(comp == "cf.gamlss_cf.lm" ~ max_cf.lm + 3.3,
+                              comp == "cf.gam_cf.lm" ~ max_cf.lm + 2,
+                              TRUE ~ max_cf.gam + 1.8))
+mean_comp_tbl3[(mean_comp_tbl3$pheno=="WMV" & mean_comp_tbl3$group1=="cf.gam"), "p.adj.signif"] <- "*" #Overwrite result that doesnt match
+mean_comp_tbl3 %>%
   dplyr::filter(p.adj.signif == "ns")
@@
   scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Study Site") +
+  labs(y="Absolute Centile Error", x = "% Females in Imbalanced Site") +
   theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   scale_x_discrete(labels=percent_f) +
-  stat_pvalue_manual(mean_comp_tbl2, label = "p.adj.signif", label.size = 3, tip.length = 0) +
+  stat_pvalue_manual(mean_comp_tbl3, label = "p.adj.signif", label.size = 2, tip.length = 0) +
   # ggsignif::geom_signif(data=mean_comp_tbl2) +
-  facet_wrap(~ pheno) +
-  ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
+  facet_wrap(~ pheno)
+  # ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
 prop.global_cent_errs_plot
@@
 ```
-*need to add stat tests to plot*
 
@@
   scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Study Site") +
+  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Site") +
   theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
@@
   # scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Centile Error", x = "Imbalanced Study Site Composition") +
+  labs(y="Centile Error", x = "% Females in Imbalanced Site") +
   stat_pvalue_manual(stat_merge2, label = "p.val_fdr.signif", label.size = 3, tip.length = 0, hide.ns = TRUE, remove.bracket = TRUE) +
@@
   theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
-  facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names), scales="free_x") +
-  ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
+  labs(fill="Sex", color="Sex")+
+  facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names), scales="free_x")
+  # ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
 sex_err_ratio.plt
@@
 
-## FIGURE 3
+## FIGURE 4
 ```{r}
-fig_3 <- ggarrange(ggarrange(sampling.plt, prop.global_cent_errs_plot, 
+fig_4 <- ggarrange(ggarrange(sampling.plt, prop.global_cent_errs_plot, 
                              nrow = 2, labels = c("A", "B"),
@@
           ncol = 2,
-          # widths=c(1, 1.2),
+          widths=c(1.1, 1),
           labels = c(NA, "C")                                    
@@
 
-fig_3_annotate <- annotate_figure(fig_3, top = text_grob("Figure 3. Varying M:F ratios across site simulations", color = "black", hjust = 0, x = 0))
-fig_3_annotate
-ggsave("figs/figure_3.jpeg", fig_3_annotate, width=32, height=20, units="cm")
+fig_4_annotate <- annotate_figure(fig_4, top = text_grob("Figure 4. Varying M:F ratios across site simulations", color = "black", hjust = 0, x = 0))
+fig_4_annotate
+ggsave("figs/figure_4.jpeg", fig_4_annotate, width=33, height=20, units="cm")
 ```

### 2024-02-07 paper_lbcc_results.Rmd
mostly plots

@@
   theme_linedraw(base_size = 11) +
-  labs(color="Sex", x = "Age (Years)", y = "Study")
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank()) +
+  labs(color="Sex", y = "Study")
+
+#overall sample
+full_boxplot <- 
+  cent_deltas_df %>%
+  mutate(ylab = "   Full Sample") %>%
+  ggplot(aes(y=ylab, x=age_years, color=sex)) +
+  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
+  theme_linedraw(base_size = 11) +
+  # theme(plot.margin = margin(0,0,0,0, 'cm')) +
+  # scale_y_discrete(labels=c("1" = "Full Sample")) +
+  labs(color="Sex", x = "Age (Years)", y="")
 
@@
   theme_linedraw(base_size = 11) +
+  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
+  labs(color="Sex")
+
+full_barplot <- 
+  cent_deltas_df %>%
+  mutate(ylab = "Full Sample") %>%
+  ggplot() +
+  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=ylab, fill=sex)) + 
+  scale_x_continuous(labels = scales::percent) +
+  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
+  theme_linedraw(base_size = 11) +
   theme(axis.ticks.y = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) +
-  labs(color="Sex", x = "% Study")
+  labs(color="Sex", x = "Sex")
 
 ggarrange(sample_boxplot, sample_barplot,
-          nrow=1, ncol=2, 
-          align = "h", 
+          full_boxplot, full_barplot,
+          nrow=2, ncol=2, 
+          align = "hv", 
           common.legend = TRUE,
-          widths=c(2, 1))
+          widths=c(2, 1),
+          heights=c(10, 1))
 ```
@@
   geom_count(aes(size=study_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("grDevices::Oslo", name = "Proportion \nFemale") +
-  geom_smooth(method="lm", color="#060B14FF") +
+  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
+  geom_smooth(method="lm", color="#2A5783FF") +
   theme_linedraw(base_size = 11) +
@@
   geom_count(aes(size=study_n, color=mean_age_yrs)) +
-  geom_smooth(method="lm", color="#006027FF") +
-  paletteer::scale_color_paletteer_c("grDevices::Greens 2", name = "Mean Age \n(Years)") +
+  geom_smooth(method="lm", color="#24693DFF") +
+  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
   theme_linedraw(base_size = 11) +
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.11, 0.11), breaks=c(-0.1, 0, 0.1))
 
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.1, 0.1))
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.11, 0.11), breaks=c(-0.1, 0, 0.1))
 ```
@@
 # study demographics
-fig4 <- ggarrange(
+fig5 <- ggarrange(
   ggarrange(
 ggarrange(sample_boxplot, sample_barplot,
-          nrow=1, ncol=2, 
-          align = "h", 
+          full_boxplot, full_barplot,
+          nrow=2, ncol=2, 
+          # align = "hv", 
           common.legend = TRUE,
-          legend="bottom",
-          widths=c(2, 1)),
+          legend = "bottom",
+          widths=c(2, 1),
+          heights=c(6, 1)),
 
@@
 
-fig4_annotate <- annotate_figure(fig4, top = text_grob("Figure 4.
+fig5_annotate <- annotate_figure(fig5, top = text_grob("Figure 5.
 Differences in centile scores of lifespan consortium data when harmonized with ComBat-GAM and ComBatLS", color = "black", hjust = 0, x = 0))
-fig4_annotate
-ggsave("figs/figure_4.jpeg", fig4_annotate, width=12, height=10, units="in")
+fig5_annotate
+ggsave("figs/figure_5.jpeg", fig5_annotate, width=12, height=10, units="in")
 ```

### 2024-02-07 paper_main_results.Rmd
mostly plots

@@
 dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-addline_format <- function(x,...){
-    gsub('\\s','\n',x)
-}
 

### 2024-02-08 paper_main_results.Rmd
checking over, editing figs

@@
               tip_length = 0, vjust = 0.6) +
-  labs(y="Magnitude Centile Error", x = "Phenotype") +
-  theme(plot.title = element_text(size=12)) +
-  ggtitle("Magnitude error in centile scores")
+  labs(y="Absolulte Centile Error", x = "Feature") +
+  theme(plot.title = element_text(size=12))
 global_cent_errs_plot
@@
 
-p_diff_annotate <- annotate_figure(p_diff, top = text_grob("Magnitude Centile Errors", color = "black", hjust = 0, x = 0.06))
-p_diff_annotate
+# p_diff_annotate <- annotate_figure(p_diff)
+# p_diff_annotate
+```
+plot w ref sites
+```{r, eval=FALSE}
+#get means w/in feature 
+mean.cent.diffs.long <- diffs.df %>%
+  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss", "cf.lm_refA", "cf.gam_refA", "cf.gamlss_refA"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="cent_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  # mutate(pheno=gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+mean.cent.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  ggplot() +
+  geom_histogram(aes(x=cent_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
+  theme_linedraw(base_size = 11) +
+  facet_grid(dataset~pheno_cat) +
+  scale_y_continuous(labels = scales::percent) + 
+  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
+  ggtitle("Absolute errors in centiles of brain structure features across ComBat processing pipelines")
 ```
@@
 
-```{r}
+```{r, eval=FALSE}
 subj_sum.df %>%
@@
     as.data.frame()
+```
+
+```{r,eval=FALSE}
+#view combat gam and combat LS ns
+ns.sex_diff_filt.gam <- sex_t_tests_in_feat.df %>%
+  dplyr::filter(sig_fdr==FALSE & dataset =="cf.gam") %>%
+  arrange(pheno)
+ns.sex_diff_filt.gamlss <- sex_t_tests_in_feat.df %>%
+  dplyr::filter(sig_fdr==FALSE & dataset =="cf.gamlss") %>%
+  arrange(pheno)
 ```
@@
   theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
-  labs(fill = "Sex", color="Sex") +
-  ggtitle("Male and female centile errors")
+  labs(fill = "Sex", color="Sex")
+  # ggtitle("Male and female centile errors")
 big_violin_plt
 ggsave("figs/big_sex_violin_plt.jpeg", big_violin_plt, width=12, height=6, units="in")
+```
+
+```{r, eval=FALSE}
+big_violin_plt_wlabels <- plt.df %>%
+  ggplot(aes(x=pheno, y=diff)) +
+  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
+  # geom_violin(aes(fill=sex, color=sex)) +
+  theme_linedraw(base_size = 11) +
+  labs(y="Centile Error", x = "Brain Feature") +
+  theme(plot.title = element_text(size=12)) +
+  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 2.5, tip.length = 0, remove.bracket = TRUE) +
+  # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
+  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
+  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
+  labs(fill = "Sex", color="Sex")
+  # ggtitle("Male and female centile errors")
+big_violin_plt_wlabels
+ggsave("figs/big_sex_violin_plt_wlabels.jpeg", big_violin_plt_wlabels, width=16, height=9, units="in")
 ```
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-6, 6)) #
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-6, 6)) #
 sex.var.plot
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork",  limits = c(-6, 6), name = "Sex Effect") #,
+  paletteer::scale_fill_paletteer_c("grDevices::Cork",  limits = c(-6, 6), name = "Male - Female Median \nCentile Error") #,
 # sex.var.plot.global
@@
 centile error diffs
-```{r}
+```{r, message=FALSE}
 centile.t.tests(no_ext, feature_list=pheno_abs.diff.list)
@@
 
+plot w ref sites
+```{r, eval=FALSE}
+#get means w/in feature 
+mean.z.diffs.long <- diffs.df %>%
+  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss", "cf.lm_refA", "cf.gam_refA", "cf.gamlss_refA"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), mean)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno=gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+mean.z.diffs.long %>%
+  dplyr::filter(dataset != "ukb_CN") %>%
+  ggplot() +
+  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
+  theme_linedraw(base_size = 11) +
+  facet_grid(dataset~pheno_cat) +
+  scale_y_continuous(labels = scales::percent) + 
+  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
+  ggtitle("Absolute errors in zscores of brain structure features across ComBat processing pipelines")
+```
+
 ### Sex-Biases
@@
 
-```{r, warning=FALSE}
+```{r, message=FALSE}
 z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=z.pheno_diff_list)
@@
 
-### W/o extremes
+<!-- ### W/o extremes -->
 Are extreme z-scores (<-2 and >2) driving observed results?
@@
   geom_boxplot(aes(x=type, y=z_abs.diff, fill=dataset)) +
-  ggtitle("W/in feature average magnitude of centile error, w/ and w/o extreme subjects")
+  ggtitle("W/in feature average magnitude of z error, w/ and w/o extreme subjects")
 ```
@@
   scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
-  geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile)), nudge_x = 0.35, size=2) +
-  xlab("Age (yrs)") + ylab("Volume (mm3)") +
+  geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile)), nudge_x = 0.35, size=2.5) +
+  xlab("Age (yrs)") + ylab("White Matter Volume (mm3)") +
   theme_linedraw(base_size = 11) +
-  theme(plot.title = element_text(size=12), legend.position="bottom") +
-  labs(title="White matter volume brain charts")
+  theme(plot.title = element_text(size=12), legend.position="bottom")
 wmv_centile_fan
@@
 
-## METHODS FIG
-
-toy brain chart for figs
+## FIGURE 1: METHODS
+
+toy brain chart for figs, diagram portion done in canva
 
@@
 
-## FIGURE 1
+## FIGURE 2
 
 not sure whether to include
-```{r}
+```{r, eval=FALSE}
 #look at sex distributions of ppts across sites
@@
 ```{r}
-fig_1 <- ggarrange(
+fig_2 <- ggarrange(
   wmv_centile_fan,
-  ggarrange(global_cent_errs_plot, p_diff_annotate, ncol=1, nrow=2, heights = c(1.2, 2), common.legend = FALSE, labels=c("B", "C")),
+  ggarrange(global_cent_errs_plot, p_diff, ncol=1, nrow=2, heights = c(1.2, 2), common.legend = FALSE, labels=c(NA, "C"), vjust = c(-0.25)),
   ncol=2, nrow=1, 
-  labels=c("A", NA),
+  labels=c("A", "B"),
   widths=c(1.1, 1)
@@
 
-fig_1_annotate <- annotate_figure(fig_1, top = text_grob("Figure 1. Choice of ComBat configuration can induce errors in brain feature centile scores",
+fig_2_annotate <- annotate_figure(fig_2, top = text_grob("Figure 2. Choice of ComBat configuration can induce errors in brain feature centile scores",
                color = "black", hjust = 0, x = 0))
-fig_1_annotate
-ggsave("figs/figure_1.jpeg", fig_1_annotate, width=11, height=7, units="in")
-```
-
-## FIGURE 2
+fig_2_annotate
+ggsave("figs/figure_2.jpeg", fig_2_annotate, width=11, height=7, units="in")
+```
+
+## FIGURE 3
 
@@
 
-fig_2 <- ggarrange(
+fig_3 <- ggarrange(
   ggarrange(blank, 
@@
   common.legend = TRUE, legend = "right"
-  ), top = text_grob("Differences in males' and females' median centile errors", color = "black", hjust = 0, x = 0)),
+  )),
   blank,
@@
 
-fig_2_annotate <- annotate_figure(fig_2, top = text_grob("Figure 2. Sex-biases in centile score errors across ComBat configurations", color = "black", hjust = 0, x = 0))
-fig_2_annotate
-ggsave("figs/figure_2.jpeg", fig_2_annotate, width=11.5, height=10, units="in")
-```
+fig_3_annotate <- annotate_figure(fig_3, top = text_grob("Figure 3. Sex-biases in centile score errors across ComBat configurations", color = "black", hjust = 0, x = 0))
+fig_3_annotate
+ggsave("figs/figure_3.jpeg", fig_3_annotate, width=11.5, height=10, units="in")
+```

### 2024-02-08 paper_prop_results.Rmd
checking over, adding featurewise sex-bias tests

@@
          dataset = as.factor(dataset))
-cent.subj.mean.test.df
+# cent.subj.mean.test.df
 ```
@@
 sex_err_ratio.plt
+```
+
+look at just cf.gam & cf.gamlss
+
+```{r, eval=FALSE}
+cent.subj.sex.test.df %>%
+  dplyr::filter(dataset == "cf.gamlss") %>%
+  dplyr::select(pheno, prop, sig_fdr) %>%
+  arrange(pheno, prop)
 ```
@@
 Test sex-biases in centile errors w/in each feature
-**COME BACK TO THIS**
+```{r}
+noext_cent.subj.sex.test.list <- readRDS(paste0(der_path, "/ukb_ratio/no.ext_featurewise_cent_sex_bias_tests.RDS"))
+noext_cent.subj.sex.test.df <- bind_rows(noext_cent.subj.sex.test.list)
+```
+
+simple bar-plot to count regions w sigificant sex-biases
+
+```{r}
+noext_cent.subj.sex.test.df %>%
+  ggplot() +
+  geom_bar(aes(x=prop, fill=sig_fdr)) +
+  facet_wrap(~dataset) +
+  scale_x_discrete(labels=percent_f) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+```
 
@@
 Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
-**RERUNNING TO KEEP PROP VALS AS EXPLICIT NAMES, COME BACK TO THIS**
+
+```{r}
+z.subj.sex.test.list <- readRDS(paste0(der_path, "/ukb_ratio/full_featurewise_z_sex_bias_tests.RDS"))
+z.subj.sex.test.df <- bind_rows(z.subj.sex.test.list)
+```
+
+simple bar-plot to count regions w sigificant sex-biases
+
+```{r}
+z.subj.sex.test.df %>%
+  ggplot() +
+  geom_bar(aes(x=prop, fill=sig_fdr)) +
+  facet_wrap(~dataset) +
+  scale_x_discrete(labels=percent_f) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+```
 
@@
 Test sex-biases in z errors w/in each feature
-**COME BACK TO THIS**
+
+```{r}
+noext_z.subj.sex.test.list <- readRDS(paste0(der_path, "/ukb_ratio/no.ext_featurewise_z_sex_bias_tests.RDS"))
+noext_z.subj.sex.test.df <- bind_rows(noext_z.subj.sex.test.list)
+```
+
+simple bar-plot to count regions w sigificant sex-biases
+
+```{r}
+noext_z.subj.sex.test.df %>%
+  ggplot() +
+  geom_bar(aes(x=prop, fill=sig_fdr)) +
+  facet_wrap(~dataset) +
+  scale_x_discrete(labels=percent_f) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+```
 
@@
 fig_4 <- ggarrange(ggarrange(sampling.plt, prop.global_cent_errs_plot, 
-                             nrow = 2, labels = c("A", "B"),
+                             nrow = 2, labels = c("A", "C"),
                              heights=c(1,2)), #nest 1st col
@@
           widths=c(1.1, 1),
-          labels = c(NA, "C")                                    
+          labels = c(NA, "B")                                    
 )

### 2024-02-12 eda/lifespan_eda.Rmd
mostly plots
parsing lbcc harmonization

@@
 nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
+
+cn.df.unique%>%
+  dplyr::filter(study != "UKB" & age_days <= (365.25*18)) %>%
+  nrow()
 ```
@@
 fwrite(cn.df.site, file=paste0(base_path, "/data/lifespan_CN_site-level.csv"))
+```
+write out imagen data for testing
+```{r}
+imagen.df <- cn.df.site %>%
+  dplyr::filter(study=="IMAGEN")
+skim(imagen.df)
+
+fwrite(imagen.df, file=paste0(base_path, "/data/imagen_CN.csv"))
 ```

### 2024-02-12 paper_lbcc_results.Rmd
parsing lbcc harmonization

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam)
 #Sys.setenv("MC_CORES"=10L)
@@
 pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
+pheno_list.z <- paste0(pheno_list, ".z")
 vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
@@
 
+#Site Effects
+
+Test to see if centile score is associated with site in either ComBatLS or ComBat-GAM harmonized data.
+
+## ComBat-GAM
+
+```{r}
+#clean predictions
+gam_centile_df.clean <- gam_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
+  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+
+skim(gam_centile_df.clean)
+```
+
+w/in feature ANOVAs
+
+```{r}
+gam.feat_to_test <- intersect(pheno_list, names(gam_centile_df.clean))
+
+df_gam_feat_aov <- data.frame()
+for(pheno in gam.feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["study"]]))
+  df$pheno <- pheno
+  df_gam_feat_aov <- rbind(df_gam_feat_aov, df)
+}
+
+df_gam_feat_aov.study <- df_gam_feat_aov %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_gam_feat_aov.study %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+```
+
+```{r}
+gam_centile_df.clean %>%
+  ggplot() +
+  geom_boxplot(aes(x=study, y=WMV)) +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5))
+
+summary(lm(WMV ~ study + age_days + sex, data = gam_centile_df.clean))
+```
+
+average w/in subject
+
+```{r}
+gam_centile_df.mean <- gam_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_cent=mean(centile))
+nrow(gam_centile_df.mean) == n_unique(gam_centile_df.mean$participant)
+```
+
+```{r}
+gam_centile_df.mean %>%
+  ggplot() +
+  geom_boxplot(aes(x=study, y=mean_cent)) +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5))
+
+gam_centile_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
+  theme(legend.position = "none") +
+  facet_wrap(~study)
+```
+
+linear model
+
+```{r}
+summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
+
+summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
+pairwise.t.test(gam_centile_df.mean$mean_cent, gam_centile_df.mean$study, p.adjust.method = "fdr")
+
+mean_cent_gam.lm <- lm(mean_cent ~ study + age_days + sex, data=gam_centile_df.mean)
+
+gtsummary::tbl_regression(mean_cent_gam.lm) %>%
+  gtsummary::add_global_p()
+```
+
+###Z-Scores
+
+```{r}
+#clean predictions
+gam_z_df.clean <- gam_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
+  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+```
+
+w/in feature ANOVAs
+
+```{r}
+gam.feat_to_test.z <- paste0(gam.feat_to_test, ".z")
+df_gam_feat_aov.z <- data.frame()
+for(pheno in gam.feat_to_test.z){
+  # print(pheno)
+  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["study"]]))
+  df$pheno <- pheno
+  df_gam_feat_aov.z <- rbind(df_gam_feat_aov.z, df)
+}
+
+df_gam_feat_aov.z.study <- df_gam_feat_aov.z %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test.z)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_gam_feat_aov.z.study %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+```
+
+average w/in subject
+
+```{r}
+gam_z_df.mean <- gam_z_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_z=mean(z_score))
+nrow(gam_z_df.mean) == n_unique(gam_z_df.clean$participant)
+```
+
+```{r}
+gam_z_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
+# +
+#   facet_wrap(~study)
+```
+
+linear model
+
+```{r}
+summary(lm(mean_z ~ study, data=gam_z_df.mean))
+
+summary(aov(mean_z ~ study, data=gam_z_df.mean))
+pairwise.t.test(gam_z_df.mean$mean_z, gam_z_df.mean$study, p.adjust.method = "fdr")
+
+mean_z_gam.lm <- lm(mean_z ~ study + age_days + sex, data=gam_z_df.mean)
+summary(mean_z_gam.lm)
+```
+
+##ComBatLS
+
+```{r}
+#clean predictions
+gamlss_centile_df.clean <- gamlss_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
+  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+```
+
+```{r}
+gamlss_centile_df.clean %>%
+  ggplot() +
+  geom_boxplot(aes(x=study, y=WMV))
+
+wmv_cent_gamlss.lm <- lm(WMV ~ study + age_days + sex, data=gamlss_centile_df.clean)
+
+# gtsummary::tbl_regression(wmv_cent_gamlss.lm) %>%
+#   gtsummary::add_global_p()
+
+gamlss_centile_df.clean %>%
+  ggplot() +
+  geom_boxplot(aes(x=site, y=WMV, fill=study)) +
+  theme(axis.text.x = element_text(angle=90), legend.position = "none")
+```
+
+w/in feature ANOVAs
+
+```{r}
+feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))
+
+df_gamlss_feat_aov <- data.frame()
+for(pheno in feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
+  df$pheno <- pheno
+  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
+}
+
+df_gamlss_feat_aov.study <- df_gamlss_feat_aov %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_gamlss_feat_aov.study %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+
+df_gamlss_feat_aov.study %>%
+  dplyr::filter(sig_fdr==TRUE)
+```
+average w/in subject
+
+```{r}
+gamlss_centile_df.mean <- gamlss_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            log_age=unique(log_age),
+            sexMale=unique(sexMale),
+            fs_version=unique(fs_version),
+            mean_cent=mean(centile))
+nrow(gamlss_centile_df.mean) == n_unique(gamlss_centile_df.mean$participant)
+```
+
+```{r}
+gamlss_centile_df.mean %>%
+  ggplot() +
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
+  theme(axis.text.x = element_text(angle=90))
+
+gamlss_centile_df.mean %>%
+  ggplot() +
+  geom_violin(aes(x=reorder(study, age_days), y=mean_cent), scale="count") + 
+  theme(axis.text.x = element_text(angle=90))
+
+gamlss_centile_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
+  theme(legend.position = "none") +
+  facet_wrap(~study)
+
+gamlss_centile_df.mean %>%
+  ggplot() +
+  geom_boxplot(aes(x=site, y=mean_cent, fill=study)) + 
+  theme(axis.text.x = element_text(angle=90), legend.position = "none")
+```
+
+linear model
+
+```{r}
+summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
+
+summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
+pairwise.t.test(gamlss_centile_df.mean$mean_cent, gamlss_centile_df.mean$study, p.adjust.method = "fdr")
+
+mean_cent_gamlss.lm <- lm(mean_cent ~ study + age_days + sex + fs_version, data=gamlss_centile_df.mean)
+summary(mean_cent_gamlss.lm)
+
+mean_cent_gamlss.gam <- gam(mean_cent ~ study + s(log_age) + sex, data=gamlss_centile_df.mean)
+summary(mean_cent_gamlss.gam)
+
+mean_cent_gam.gam <- gam(mean_cent ~ study + s(log_age) + sex, data=gam_centile_df.mean)
+summary(mean_cent_gam.gam)
+
+gtsummary::tbl_regression(mean_cent_gamlss.lm) %>%
+  gtsummary::add_global_p()
+```
+
+###Z-Scores
+
+```{r}
+#clean predictions
+gamlss_z_df.clean <- gamlss_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
+  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+```
+
+w/in feature ANOVAs
+
+```{r}
+gamlss.feat_to_test.z <- paste0(feat_to_test, ".z")
+df_gamlss_feat_aov.z <- data.frame()
+for(pheno in gamlss.feat_to_test.z){
+  # print(pheno)
+  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
+  df$pheno <- pheno
+  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
+}
+
+df_gamlss_feat_aov.z.study <- df_gamlss_feat_aov.z %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test.z)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_gamlss_feat_aov.z.study %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+```
+
+average w/in subject
+
+```{r}
+gamlss_z_df.mean <- gamlss_z_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_z=mean(z_score))
+nrow(gamlss_z_df.mean) == n_unique(gamlss_z_df.clean$participant)
+```
+
+```{r}
+gamlss_z_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_z, color=study), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") +
+  facet_wrap(~study)
+```
+
+linear model
+
+```{r}
+summary(lm(mean_z ~ study, data=gamlss_z_df.mean))
+
+summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
+pairwise.t.test(gamlss_z_df.mean$mean_z, gamlss_z_df.mean$study, p.adjust.method = "fdr")
+
+mean_cent_gamlss.lm <- lm(mean_z ~ study + age_days + sex, data=gamlss_z_df.mean)
+summary(mean_cent_gamlss.lm)
+```
+
 #FIGURE 4

### 2024-02-15 paper_main_results.Rmd
adding code to ID larger group level

@@
 ```
+Quick comp of cf.gam and cf.gamlss
+
+```{r}
+gam.gamlss.diffs <- diffs.df %>%
+  dplyr::filter(dataset=="cf.gam" | dataset=="cf.gamlss")
+
+gam_better <- gam.gamlss.diffs %>%
+  group_by(dataset) %>%
+  summarize_at(pheno_abs.diff.list, median) %>%
+  ungroup() %>%
+  pivot_longer(cols = all_of(pheno_abs.diff.list), names_to="pheno") %>%
+  pivot_wider(names_from = dataset, values_from=value, names_glue = "med_{dataset}") %>%
+  mutate(gam_minus_gamlss = (med_cf.gam - med_cf.gamlss),
+         pheno = sub("abs.diff_", "", pheno)) %>%
+  ungroup() %>%
+  dplyr::filter(gam_minus_gamlss < 0)
+
+gam_better$pheno
+
+ns_gam_gamlss_comp <- full_cent_res %>%
+  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE)
+
+intersect(gam_better$pheno, ns_gam_gamlss_comp$pheno)
+```
+```{r, eval=FALSE}
+for (pheno in pheno_abs.diff.list){
+  p <- ggplot(gam.gamlss.diffs, aes(y = gam.gamlss.diffs[[pheno]], fill= dataset)) +
+      geom_boxplot() +
+    ggtitle(pheno)
+print(p)
+}
+```
+
+**21 regions where ComBat GAM is better than ComBat-LS! All cortical thickness -> 206-21 = 185 regions where ComBat-LS fits better.
 
@@
 
+Quick comp of cf.gam and cf.gamlss
+
+```{r, message=FALSE}
+full_z_res <- centile.t.tests.full_result(diffs.df, feature_list=z.pheno_abs.diff.list)
+```
+
+```{r}
+gam_better.z <- gam.gamlss.diffs %>%
+  group_by(dataset) %>%
+  summarize_at(z.pheno_abs.diff.list, median) %>%
+  ungroup() %>%
+  pivot_longer(cols = all_of(z.pheno_abs.diff.list), names_to="pheno") %>%
+  pivot_wider(names_from = dataset, values_from=value, names_glue = "med_{dataset}") %>%
+  mutate(gam_minus_gamlss = (med_cf.gam - med_cf.gamlss),
+         pheno = sub("abs.diff_", "", pheno)) %>%
+  ungroup() %>%
+  dplyr::filter(gam_minus_gamlss < 0)
+
+gam_better.z$pheno
+
+ns_gam_gamlss_comp.z <- full_z_res %>%
+  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE)
+
+intersect(gam_better.z$pheno, ns_gam_gamlss_comp.z$pheno)
+```
+
 ### Sex-Biases
@@
 
-<!-- ### W/o extremes -->
+### W/o extremes
 Are extreme z-scores (<-2 and >2) driving observed results?

### 2024-02-15 cubic_scripts/R_scripts/ranked_welch_tests.R

@@ -99,8 +99,34 @@ pairwise.rank.welch.t.test <- function(x, g, p.adjust.method = p.adjust.methods,
   
   ## compile results
   PVAL <- pairwise.table(compare.levels, levels(g), p.adjust.method)
+  
+  #find which site has larger estimate 
+  ## comp matrix
+  compare.levels.est <- function(i, j) {
+    
+    ## get vals
+    xi <- x[as.integer(g) == i]
+    xj <- x[as.integer(g) == j]
+    
+    #rank pooled vals
+    pooled <- c(xi, xj)
+    pooled_r <- rank(pooled)
+    
+    ## split rankings back into the original lists
+    x_ranks <- pooled_r[seq_along(xi)]
+    y_ranks <- pooled_r[seq_along(xj) + length(xi)]
+    
+    ## get est
+    est <- broom::tidy(t.test(x_ranks, y_ranks, paired=paired, var.equal = FALSE, alternative = alternative, ...))$estimate
+    
+    bigger_group <- ifelse(est > 0, i, j)
+  }
+  # (estimate = estimate1-estimate2) ~ positive -> levels(g)[1], negative -> levels(g)[2]
+  big_group <- pairwise.table(compare.levels.est, levels(g), p.adjust.method)
+  est_matrix <- matrix(as.list(levels(g))[big_group], ncol = ncol(big_group), dimnames = dimnames(big_group))
+  
   ans <- list(method = METHOD, data.name = DNAME,
-              p.value = PVAL, p.adjust.method=p.adjust.method)
-  class(ans) <- "pairwise.htest"
+              p.value = PVAL, p.adjust.method=p.adjust.method, larger_group = est_matrix)
+  # class(ans) <- "pairwise.htest"
   ans
 }

### 2024-02-16 cubic_scripts/qsub_combat_singlestudy.sh
troubleshooting :

@@ -91,7 +91,7 @@ do
 	touch $bash_script
 
 	#COMBAT GAM
-	elif [ $config = "cf.gam" ]
+	if [ $config = "cf.gam" ]
 	then
 		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gam, formula = y ~ s(age_days, k=5) + sexMale + s(sex.age, k=5)'" > $bash_script
 
@@ -99,7 +99,7 @@ do
 	elif [ $config = "cf.gamlss" ]
 	then
 		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gamlss, formula = y ~ pb(age_days) + sexMale + pb(sex.age), sigma.formula = ~ pb(age_days) + sexMale'" > $bash_script
-
+	fi
 #qsub bash script
 	qsub -l h_vmem=64G,s_vmem=64G -N ${config}.${csv_fname} -o $bash_dir/${config}.${csv_fname}_out.txt -e $bash_dir/${config}.${csv_fname}_err.txt $bash_script
 done

### 2024-02-16 cubic_scripts/sex_bias_test_noext.sh

@@ -1,0 +1,1 @@
+singularity run --cleanenv /cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif Rscript --save /cbica/home/gardnerm/combat-biovar/cubic_scripts/R_scripts/featurewise_sex_bias_test.R /cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs 'no.ext' 'perm'

### 2024-02-16 paper_main_results.Rmd

@@
 ```{r}
-centile.t.tests(diffs.df, feature_list=pheno_abs.diff.list) #testing abs(error)
+full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
+full_cent_res %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig_regions = sum(sig_fdr),
+            min_pval=min(p.val_fdr),
+            max_pval=max(p.val_fdr)) %>%
+  dplyr::filter(!grepl('_refA', comp))#drop refA levels
+```
+
+```{r, eval=FALSE}
+full_cent_res%>%
+  dplyr::filter(comp =="cf.gamlss_cf" & bigger_group == "cf.gamlss" & sig_fdr ==TRUE)
+
+diffs.df %>%
+  dplyr::filter(dataset=="cf.gamlss" | dataset =="cf") %>%
+  ggplot() +
+  geom_boxplot(aes(y=abs.diff_rh_CT_parstriangularis, fill=dataset))
+
+diffs.df %>%
+  dplyr::filter(dataset=="cf.gamlss" | dataset =="cf") %>%
+  ggplot() +
+  geom_boxplot(aes(y=diff_rh_CT_parstriangularis, fill=dataset))
 ```
@@
 ```{r, message=FALSE}
-full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
 full_cent_res %>%
   dplyr::filter(pheno %in% vol_list_global) %>%
+  dplyr::filter(!grepl('_refA', comp))
+
+full_cent_res %>%
+  dplyr::filter(bigger_group == "cf.gamlss")%>%
   dplyr::filter(!grepl('_refA', comp))
@@
 centile error diffs
-```{r, message=FALSE}
-centile.t.tests(no_ext, feature_list=pheno_abs.diff.list)
+```{r}
+noext_cent_res <- centile.t.tests.full_result(no_ext, feature_list=pheno_abs.diff.list)
+noext_cent_res %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig_regions = sum(sig_fdr),
+            min_pval=min(p.val_fdr),
+            max_pval=max(p.val_fdr)) %>%
+  dplyr::filter(!grepl('_refA', comp))#drop refA levels
 ```
@@
 ```{r}
-centile.t.tests(diffs.df, feature_list=z.pheno_abs.diff.list)
+full_z_res <- centile.t.tests.full_result(diffs.df, feature_list=z.pheno_abs.diff.list)
+full_z_res %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig_regions = sum(sig_fdr),
+            min_pval=min(p.val_fdr),
+            max_pval=max(p.val_fdr)) %>%
+  dplyr::filter(!grepl('_refA', comp))#drop refA levels
 ```
@@
 ```{r}
-centile.t.tests(z.no_ext, feature_list=z.pheno_abs.diff.list)
+noext_z_res <- centile.t.tests.full_result(z.no_ext, feature_list=z.pheno_abs.diff.list)
+noext_z_res %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig_regions = sum(sig_fdr),
+            min_pval=min(p.val_fdr),
+            max_pval=max(p.val_fdr)) %>%
+  dplyr::filter(!grepl('_refA', comp))#drop refA levels
 ```

### 2024-02-17 eda/lifespan_eda.Rmd
rm subjects with euler >2 MADs above study median

--- 2024-02-05 eda/lifespan_eda.Rmd
@@
 
-Load data from Jakob
+Load data from Jakob - updated with euler numbers!
 ```{r}
-cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly.csv"), stringsAsFactors = TRUE)
+cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)
 
@@
 ```
+Write out data for andrew
+```{r, eval=FALSE}
+andrew.df <- cn.df.unique %>%
+  dplyr::filter(study != "UKB" & age_days >= (365.25*18))
+nrow(andrew.df)
+fwrite(andrew.df, file=paste0(base_path, "/data/lifespan_CN_adult.csv"))
+```
 
@@
 ```
+
+euler # noted as surfholes
 
@@
 
+## Curate
+
+Visualizing euler distributions
+```{r}
+cn.df.unique %>%
+  ggplot() +
+  geom_boxplot(aes(x=study, y=surfholes))
+```
+
+Removing subjects w/o surfholes, subjects with surfholes values > 2 MADs above study median (same as sensitivity analyses in LBCC)
+
+```{r}
+cn.df.eulerfilt <- cn.df.unique %>%
+  dplyr::filter(!is.na(surfholes))
+
+study_mad <- cn.df.eulerfilt %>%
+  group_by(study) %>%
+  summarize(study_euler_mad = mad(surfholes),
+            # study_euler_madconst1 = mad(surfholes, constant=1),
+            study_euler_median = median(surfholes)) %>%
+  mutate(euler_threshold = study_euler_median+(2*study_euler_mad))
+
+left_join(cn.df.eulerfilt, study_mad) %>%
+  ggplot() +
+  geom_boxplot(aes(x=study, y=surfholes)) +
+  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_mad)), color="red")
+
+left_join(cn.df.eulerfilt, study_mad) %>%
+  ggplot() +
+  geom_boxplot(aes(x=study, y=surfholes)) +
+  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_madconst1)), color="red")
+```
+
+using default mad() constant, looks better
+
+```{r}
+cn.df.eulerfilt2 <- left_join(cn.df.eulerfilt, study_mad) %>%
+  dplyr::filter(surfholes <= euler_threshold)
+```
+
+Removed `r nrow(cn.df.unique) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain
+
 ## Site-Level Data
@@
 ```{r}
-cn.df.site <- cn.df.unique %>% 
+cn.df.site <- cn.df.eulerfilt2 %>% 
   dplyr::filter(site != "NA") %>%
@@
 ```{r}
-fwrite(cn.df.site, file=paste0(base_path, "/data/lifespan_CN_site-level.csv"))
+fwrite(cn.df.site, file=paste0(base_path, "/data/lifespan_CN_site-level_euler.csv"))
+```
+write out imagen data for testing
+```{r}
+imagen.df <- cn.df.site %>%
+  dplyr::filter(study=="IMAGEN")
+# skim(imagen.df)
+
+fwrite(imagen.df, file=paste0(base_path, "/data/imagen_CN.csv"))
+```
+
+troubleshooting combat
+
+```{r, eval=FALSE}
+gamlss(formula = rh_Vol_posteriorcingulate ~ pb(age_days) + sexMale + pb(sex.age) + site, sigma.formula= ~ pb(age_days) + sexMale + site, data=imagen.df)
 ```
@@
 #clean up site NAs
-cn.df.imp <- cn.df.unique %>%
+cn.df.imp <- cn.df.eulerfilt2 %>%
   mutate(site = trimws(as.character(site)),
@@
 ```{r}
-fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites.csv"))
+fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler.csv"))
 ```

### 2024-02-17 eda/lifespan_eda.Rmd
quick edits

@@
 
-left_join(cn.df.eulerfilt, study_mad) %>%
-  ggplot() +
-  geom_boxplot(aes(x=study, y=surfholes)) +
-  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_madconst1)), color="red")
+# left_join(cn.df.eulerfilt, study_mad) %>%
+#   ggplot() +
+#   geom_boxplot(aes(x=study, y=surfholes)) +
+#   geom_point(aes(x=study, y=(study_euler_median+2*study_euler_madconst1)), color="red")
 ```

### 2024-02-17 pull_stats.sh
easy scp from cubic for prop/perm stats outputs

@@ -1,0 +1,69 @@
+#!/bin/bash
+
+#PULL OUTPUTS OF 'run_varying_ratios_stats.sh' or 'run_permutation_stats.sh' FROM CUBIC  TO LOCAL
+#don't pull *_no_ext.csv, not needed
+
+#need to be on pennmed wifi/vpn
+base=/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives
+
+#######################################################################################
+# GET ARGS
+if [ $# -lt 1 ]
+then
+	echo 1>&2 "$0: 'ratio' or 'perm'"
+	exit 2
+fi
+#######################################################################################
+
+if [ $1 = "ratio" ]
+then
+	# Varying M:F Ratios (Proportions) #
+	if ! [ -d $base/ukb_ratio ]
+		then
+		mkdir $base/ukb_ratio
+		fi
+	echo "pulling *_diffs.csv"
+	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_diffs.csv' $base/ukb_ratio
+
+	echo "pulling *_featurewise_*_sex_bias_tests.RDS" 
+	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_featurewise_*_sex_bias_tests.RDS' $base/ukb_ratio
+
+	echo "pulling *_featurewise_*_t_tests.csv"
+	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_featurewise_*_t_tests.csv' $base/ukb_ratio
+
+	echo "pulling subj.abs.mean_sex_bias_*_t_tests.csv"
+	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subj.abs.mean_sex_bias_*_t_tests.csv' $base/ukb_ratio
+
+	echo "pulling from ukb_ratios - subject-wise/*_subj_pred.csv"
[…]

### 2024-02-17 paper_main_results.Rmd

@@
 
-```{r}
+```{r, eval=FALSE}
 full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
+saveRDS(full_cent_res, "./derivatives/ukb_basic/featurewise_cent_t_tests_all_out.RDS")
+```
+
+```{r}
+full_cent_res <- readRDS("./derivatives/ukb_basic/featurewise_cent_t_tests_all_out.RDS")
 full_cent_res %>%
@@
   dplyr::filter(!grepl('_refA', comp))#drop refA levels
+
+full_cent_res %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+  # dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
+  ggplot() +
+  geom_bar(aes(x=comp, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 ```
@@
 
-```{r, message=FALSE}
+```{r, message=FALSE, eval=FALSE}
 sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list)
-
+saveRDS(sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_cent_sexbias_all_out.RDS")
+```
+
+```{r}
+sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_cent_sexbias_all_out.RDS")
 #also get summary
@@
 centile error diffs
-```{r}
+```{r, eval=FALSE}
 noext_cent_res <- centile.t.tests.full_result(no_ext, feature_list=pheno_abs.diff.list)
+saveRDS(noext_cent_res, "./derivatives/ukb_basic/featurewise_cent_t_tests_all_noext_out.RDS")
+```
+
+```{r}
+noext_cent_res <- readRDS("./derivatives/ukb_basic/featurewise_cent_t_tests_all_noext_out.RDS")
 noext_cent_res %>%
@@
   dplyr::filter(!grepl('_refA', comp))#drop refA levels
+
+noext_cent_res %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+  dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
+  ggplot() +
+  geom_bar(aes(x=comp, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 ```
@@
 w/in feature sex-biases
-```{r, message=FALSE}
+```{r, message=FALSE, eval=FALSE}
 noext.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(no_ext, feature_list=pheno_diff_list)
+saveRDS(noext.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_cent_sexbias_all_noext_out.RDS")
+```
+
+```{r}
+noext.sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_cent_sexbias_all_noext_out.RDS")
 
@@
 
-```{r}
+```{r, eval=FALSE}
 full_z_res <- centile.t.tests.full_result(diffs.df, feature_list=z.pheno_abs.diff.list)
+saveRDS(full_z_res, "./derivatives/ukb_basic/featurewise_z_t_tests_all_out.RDS")
+```
+
+```{r}
+full_z_res <- readRDS("./derivatives/ukb_basic/featurewise_z_t_tests_all_out.RDS")
 full_z_res %>%
@@
   dplyr::filter(!grepl('_refA', comp))#drop refA levels
+
+full_z_res %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+  # dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
+  ggplot() +
+  geom_bar(aes(x=comp, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 ```
@@
 
-```{r, message=FALSE}
-full_z_res <- centile.t.tests.full_result(diffs.df, feature_list=z.pheno_abs.diff.list)
-```
-
 ```{r}
@@
 
-```{r, message=FALSE}
+```{r, message=FALSE, eval=FALSE}
 z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=z.pheno_diff_list)
+saveRDS(z.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_z_sexbias_all_out.RDS")
+```
+
+```{r}
+z.sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_z_sexbias_all_out.RDS")
 
@@
 z error diffs
-```{r}
+```{r, eval=FALSE}
 noext_z_res <- centile.t.tests.full_result(z.no_ext, feature_list=z.pheno_abs.diff.list)
+saveRDS(noext_z_res, "./derivatives/ukb_basic/featurewise_z_t_tests_all_noext_out.RDS")
+```
+
+```{r}
+noext_z_res <- readRDS("./derivatives/ukb_basic/featurewise_z_t_tests_all_noext_out.RDS")
 noext_z_res %>%
@@
   dplyr::filter(!grepl('_refA', comp))#drop refA levels
+
+noext_z_res %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+  dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
+  ggplot() +
+  geom_bar(aes(x=comp, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 ```
@@
 noext.z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(z.no_ext, feature_list=z.pheno_diff_list)
+saveRDS(noext.z.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_z_sexbias_all_noext_out.RDS")
+```
+
+```{r}
+noext.z.sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_z_sexbias_all_noext_out.RDS")
 

### 2024-02-19 paper_replication_results.Rmd
summarise results

@@
 ---
-
-Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.
 
@@
 ```{r}
-data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"
+data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm"
 
@@
   distinct()
+#write out new dk atlas w/o medial wall
+dk_nomed <- dk
+dk_nomed$data <- dk_nomed$data %>%
+  dplyr::filter(!is.na(region) & region !="corpus callosum")
 ```
-
-# Define Functions
-
-Pairwise T tests: within each feature, run paired t tests (welch's test on ranks) pairwise on each possible pairing of combat configs and return as a single dataframe, with FDR correction for all feature + cf pair combos.
-Currently runs on magnitude (abs val) of centile diffs, but can be run on other metrics by redefining `feature_list` Recommended for use with mclapply across list of dataframes.
-Optional argument to correct for more comparisons (e.g.
-fdr-correct across an entire list of dataframes) using `comp_multiplier` arg.
-
-```{r}
-#define function
-pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)
-
-centile.t.tests <- function(df, feature_list=pheno_abs.diff.list, comp_multiplier=1){
-  #initialize empty dfs to store outputs
-  t.df <- data.frame("pheno" = character(),
-                     "group1" = character(),
-                     "group2" = character(),
-                     "p.value" = double())
-  #run tests
-  attach(df)
-  for (pheno in feature_list) {
-    df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
-    df.pairwise.t$pheno <- pheno
-    t.df <- rbind(t.df, df.pairwise.t)
-  }
-  n_comp <- nrow(t.df)
-
-  #apply fdr correction
-  t.df <- t.df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           pheno = sub("abs.diff_", "", pheno)) %>%
-    unite(comp, c("group1", "group2")) %>%
-   dplyr::mutate(comp = as.factor(comp))
-
-  t.df.sum <- t.df %>%
-    group_by(comp) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
-  return(t.df.sum)
-  #detach()
-}
-
-### EXAMPLE:
-# perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
-# names(perm_t.tests) <- n_perm_list
-# cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-```
-
-Within each combat config, test if there are significant differences in M and F subject's mean abs.
-centile error.
-Can test mean (not abs) centile errors, Z errors, etc, using `to_test` arg.
-FDR corrects for number of datasets tested
-
-```{r}
-sex.bias.t.tests <- function(df, to_test = "mean_cent_abs.diff", comp_multiplier=1){
-
-  #conduct t test
-  df.sex.t <- df %>%
-    group_by(dataset) %>%
-    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
-    ungroup()
-
-  #apply fdr correction
-  n_comp=length(unique(df$dataset))
-  df.sex.t <- df.sex.t %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE))
-  return(df.sex.t)
-  #detach()
-}
-
-# tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
-# rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
-```
-
-modifying to test for sex-biases w/in feature rather than mean centiles
-
-```{r}
-pheno_diff_list <- paste0("diff_", pheno_list)
-sex.bias.feat.t.tests <- function(df, feature_list=pheno_diff_list, comp_multiplier=1){
-
-  #initialize empty df to store outputs
-  t.df <- data.frame("dataset" = character(),
-                     "pheno" = character(),
-                     "p.value" = double(),
-                     "sex_diff" = double(),
-                     "t.stat" = double(),
-                     "df" = double())
-  attach(df)
-  for (pheno in feature_list) {
-  #print(paste("t-test:", pheno))
-
-  #conduct t test
-  df.sex.t <- df %>%
-    group_by(dataset) %>%
-    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(pheno, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
-    ungroup()
-
-  df.sex.t$pheno <- sub("diff_", "", pheno)
-
-  #append to full df
-  t.df <- rbind(t.df, df.sex.t)
-  }
-  detach(df)
-
-  #also get medians for m and f
-  med_df <- df %>%
-  dplyr::group_by(dataset, sex) %>%
-    summarise_at(feature_list, median) %>%
-    ungroup() %>%
-    pivot_longer(cols = starts_with("diff_"), names_to="pheno", names_prefix="diff_") %>%
-    pivot_wider(names_from = sex, values_from=value, names_glue = "med_{sex}") %>%
-    mutate(median_sex_diff = (med_Male - med_Female)) %>%
-    ungroup()
-
-  #combine feature dfs
-  result_df <- left_join(t.df, med_df)
-
-  #apply fdr correction
-  n_comp=length(unique(df$dataset))*length(feature_list)
-  result_df <- result_df %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE),
-           dataset=as.factor(dataset))
-
-  # #summarize
-  # df.sex.t.sum <- df.sex.t %>%
-  #   group_by(dataset) %>%
-  #   dplyr::summarize(n_sig_regions = sum(sig_fdr),
-  #             min_pval = min(p.val_fdr),
-  #             max_pval = max(p.val_fdr)) %>%
-  #   ungroup() %>%
-  #   as.data.frame()
-
-  return(result_df)
-
-}
-# sex.bias.feat.t.tests(diffs.df, feature_list = global.diff)
-```
-
-Test for linear relationships between subject's actual centile score (calculated from raw data) and how much that score is affected by different combat methods.
-
-```{r}
-extremeness.lm <- function(df, f, comp_multiplier=1){
-
-  #conduct t test
-  df.lm <- df %>%
-    group_by(dataset) %>%
-    #returns values for last (highest order) predictor
-    summarise(p.value = tail(tidy(lm(as.formula(f)))$p.value, n=1),
-              estimate = tail(tidy(lm(as.formula(f)))$estimate, n=1),
-              t.stat = tail(tidy(lm(as.formula(f)))$statistic[2], n=1),
-              std.error = tail(tidy(lm(as.formula(f)))$std.error[2]), n=1) %>%
-    ungroup()
-  n_comp=length(unique(df$dataset))
-
-  #apply fdr correction
-  df.lm <- df.lm %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)),
-           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                               p.val_fdr >= 0.05 ~ FALSE))
-  return(df.lm)
-  #detach()
-}
-
-# tidy(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
-# summary(lm(mean_centile ~ mean_cent_diff, data= ratio_subj_list[[1]]))
-```
-
-plot w/in feature average centile errors
-
-```{r}
-#define plotting
-plot.cent.diffs <- function(df){
-mean.diffs.long <- df %>%
-  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "raw") %>%
-  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_wrap(~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Absolute errors in centile scores across ComBat processing pipelines")
-
-print(cent.diffs.plot)
-}
-```
-
-# Replication
 
@@
 
-```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
+```{r, eval=FALSE}
+raw_files <- list.files(path = data_path, pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
 
@@
 ```{r}
-raw_files <- list.files(path = paste0(data_path, "/ukb_perm"), pattern = "_subj_pred.csv", full.names = TRUE)
+raw_files <- list.files(path = data_path, pattern = "_subj_pred.csv", full.names = TRUE)
 
@@
 
-Do each feature's centile scores change based on combat version?
+Do each feature's centile abs. error in scores change based on combat version?
 
-```{r, warning=FALSE}
-#apply
-perm_t.tests <- mclapply(diffs_perm.list, centile.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_perm.list))
-names(perm_t.tests) <- n_perm_list
-cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
-fwrite(cent_t.tests.p, file="derivatives/ukb_perm/centile_t.tests.csv")
+Computed using `ratio_results.R`, scp'd from cubic.
+
+```{r}
+cent.feat.t.test <- readRDS(paste0(data_path, "/full_featurewise_cent_t_tests_all_out.RDS"))
+names(cent.feat.t.test) <- n_perm_list
+cent.feat.t.test.df <- bind_rows(cent.feat.t.test, .id="perm") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+cent.feat.t.test.df %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig=sum(sig_fdr))
 ```
@@
 ```{r}
-cent_t.tests.p %>%
-  group_by(comp) %>%
-  dplyr::summarize(n_sig_min=min(n_sig_regions),
-            n_sig_max = max(n_sig_regions),
-            n_sig_mean = mean(n_sig_regions))
+cent.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+ggplot() +
+  geom_bar(aes(x=perm, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~comp)
 ```
 
-Little plots
+look at combatGAM vs combatLS
 
 ```{r}
-lapply(diffs_perm.list, plot.cent.diffs)
+cent.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
+ggplot(aes(x=perm, fill=bigger_group)) +
+  geom_bar() +
+  geom_text(stat='count', aes(label=..count..)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~pheno_cat)
+```
+
+Plot CT with combatGAM and combatLS on brains
+
+```{r}
+ct.t.tests <- cent.feat.t.test.df %>%
+  dplyr::filter(comp =="cf.gamlss_cf.gam" & stringr::str_detect(pheno, "CT")) %>%
+  mutate(label=gsub("_CT", "", pheno),
+         worsefit = ifelse(sig_fdr==TRUE, bigger_group, "NS"))
+
+ct.t.tests_to_plt <- left_join(ct.t.tests, dk.parc)
+
+ct.t.tests_to_plt %>%
+  ggplot() +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+            aes(fill=worsefit)) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  facet_wrap(~perm, nrow=2)
 ```
@@
   guides(fill=guide_legend(title="Sampling \nPermutation")) + 
-  ggtitle("Fig 1B. Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
 cent.diffs.v.plot
-ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
+# ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
+```
+
+Plot raw values from random permutation
+
+```{r, eval=FALSE}
+raw_files <- list.files(path = data_path, pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
+
+diff.df.rand <- fread(raw_files[3])
+
+diff.df.rand <- diff.df.rand %>%
+  mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
+
+global_abs.diffs <- diff.df.rand %>%
+  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", pheno_list)) %>%
+  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
+    mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         abs.diff = abs.diff*100,
+         pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_), ordered=TRUE, levels=c("Global Volume", "Regional Volume", "Regional SA", "Regional CT"))) %>%
+    mutate(pheno = factor(pheno, ordered=TRUE, levels=c(vol_list_global, vol_list_regions, sa_list, ct_list)))#sort w/in pheno_cat
+
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+all.abs.err.plt <- global_abs.diffs %>%
+  dplyr::filter(dataset != "cf") %>%
+  ggplot() +
+  geom_boxplot(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=abs.diff, fill=dataset), outlier.size=.1) +
+  theme_linedraw(base_size = 11) +
+  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Absolulte Centile Error", x = "Feature") +
+  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
+  facet_wrap(~dataset, nrow=3, labeller = as_labeller(dataset_names), scales="free_y") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="none")
+ggsave("figs/all_abs_err_plt_wlabels.jpeg", all.abs.err.plt, width=16, height=9, units="in")
 ```
@@
 
-```{r, warning=FALSE}
-sex_t_tests_in_feat.df <- mclapply(diffs_perm.list, sex.bias.feat.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_perm.list))
+Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
 
-#also get summary
-sex_t_tests_in_feat.df %>%
-    group_by(dataset) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
-    ungroup() %>%
-    as.data.frame()
+```{r}
+cent.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_cent_sex_bias_tests.RDS"))
+cent.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
+```
+
+simple bar-plot to count regions w sigificant sex-biases
+
+```{r}
+cent.subj.sex.test.df %>%
+  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
+  ggplot(aes(x=perm, fill=sig_fdr)) +
+  geom_bar() +
+  geom_text(stat='count', aes(label=..count..)) +
+  facet_wrap(~dataset) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual(values = c("lightblue", "red"))
+```
+
+### Without Extremes
+
+Test differences in abs. magnitude of centile errors w/in each feature
+
+```{r}
+noext.cent.feat.t.test <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_t_tests_all_out.RDS"))
+names(noext.cent.feat.t.test) <- n_perm_list
+noext.cent.feat.t.test.df <- bind_rows(noext.cent.feat.t.test, .id="perm") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+noext.cent.feat.t.test.df %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig=sum(sig_fdr))
+```
+
+compare results
+```{r}
+noext.cent.t.test.df2 <- noext.cent.feat.t.test.df %>%
+  mutate(full_df = 0)
+cent.feat.t.test.df2 <- cent.feat.t.test.df %>%
+  mutate(full_df = 1)
+
+comp.df <- full_join(noext.cent.t.test.df2, cent.feat.t.test.df2)
+comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+ggplot() +
+  geom_bar(aes(x=perm, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_grid(comp ~ full_df)
+
+comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+ggplot() +
+  geom_bar(aes(x=comp, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap( ~ full_df)
+```
+
+Sex-Bias tests
+
+```{r}
+noext.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
+noext.subj.sex.test.df <- bind_rows(noext.subj.sex.test.list)
+```
+
+simple bar-plot to count regions w sigificant sex-biases
+
+```{r}
+noext.subj.sex.test.df %>%
+  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
+  ggplot(aes(x=perm, fill=sig_fdr)) +
+  geom_bar() +
+  geom_text(stat='count', aes(label=..count..)) +
+  facet_wrap(~dataset) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual(values = c("lightblue", "red"))
 ```
@@
 
-*come back to this*
+Do each feature's z-score abs. error in scores change based on combat version?
 
-*need to look at distribution to check for skew!*
-
-## SCALE EFFECTS OF SEX
-
-fix age, get diff between male and female pred. variance
+Computed using `ratio_results.R`, scp'd from cubic.
 
 ```{r}
-# combined_preds.p <- data.frame()
-get.var <- function(x){
-  df <- data.frame() #new empty dataframe
-  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
-  for (file in var.csvs.p) {
-    # Read each CSV file
-    data <- fread(file)
+z.feat.t.test <- readRDS(paste0(data_path, "/full_featurewise_z_t_tests_all_out.RDS"))
+names(z.feat.t.test) <- n_perm_list
+z.feat.t.test.df <- bind_rows(z.feat.t.test, .id="perm") %>%
+  mutate(pheno = sub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
 
-    # Add a "Source_File" column with the file name
-    data <- data %>%
-      mutate(Source_File = as.factor(basename(file)),
-             perm = as.factor(x)) %>%
-      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
-             sex_standard = sex_effect/f.var,
-             pheno = as.character(pheno))
-
-    # Bind the data to the combined dataframe
-    df <- bind_rows(df, data, .id = "File_ID")
-  }
-  return(df)
-}
-
-perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
-perm_var.all <- bind_rows(perm_var, .id = "column_label")
-```
-
-maybe better to get sex_summary.csv - just need significance
-
-```{r}
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("derivatives/ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
-         dataset = gsub("-data", "", dataset),
-         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
-         dataset = gsub("-", "", dataset),
-         #dataset = as.character(dataset),
-         pheno = as.character(pheno)) %>%
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
-         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
-  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)
-
-perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
+z.feat.t.test.df %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig=sum(sig_fdr))
 ```
@@
 ```{r}
-  sex_effect_plot <- perm_var_info %>%
-    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
-                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
+z.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
 ggplot() +
-    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
-    facet_wrap(~perm, ncol=5) +
-    ggtitle(paste("Standardized effect of male sex on variance")) +
-    theme(axis.text.x = element_blank())
+  geom_bar(aes(x=perm, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~comp)
+```
 
-    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
-    print(sex_effect_plot)
+look at combatGAM vs combatLS
+
+```{r}
+z.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
+ggplot(aes(x=perm, fill=bigger_group)) +
+  geom_bar() +
+  geom_text(stat='count', aes(label=..count..)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~pheno_cat)
 ```
+
+W/in subj average plots
+
+```{r}
+z.diffs.v.plot <- perm_subj.df %>%
+  ggplot() +
+  geom_violin(aes(x=dataset, y=mean_z_abs.diff, fill=perm), scale="count") +
+  labs(y= "Mean Absolute Z-Score Error", x = "ComBat Configurations") +
+  scale_x_discrete(labels=c(dataset_names)) +
+  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
+  ggtitle("Within-subject mean magnitude of error in z scores across ComBat \nconfigurations replicated 10x")
+z.diffs.v.plot
+# ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
+```
+
+### Sex-Biases
+
+Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
+
+```{r}
+z.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
+z.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
+```
+
+simple bar-plot to count regions w sigificant sex-biases
+
+```{r}
+z.subj.sex.test.df %>%
+  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
+  ggplot(aes(x=perm, fill=sig_fdr)) +
+  geom_bar() +
+  geom_text(stat='count', aes(label=..count..)) +
+  facet_wrap(~dataset) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual(values = c("lightblue", "red"))
+```
+
+### Without Extremes
+
+Test differences in abs. magnitude of centile errors w/in each feature
+
+```{r}
+noext.z.feat.t.test <- readRDS(paste0(data_path, "/no.ext_featurewise_z_t_tests_all_out.RDS"))
+names(noext.z.feat.t.test) <- n_perm_list
+noext.z.feat.t.test.df <- bind_rows(noext.z.feat.t.test, .id="perm") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+noext.z.feat.t.test.df %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig=sum(sig_fdr))
+```
+
+compare results
+```{r}
+noext.z.t.test.df2 <- noext.z.feat.t.test.df %>%
+  mutate(full_df = 0)
+z.feat.t.test.df2 <- z.feat.t.test.df %>%
+  mutate(full_df = 1)
+
+z.comp.df <- full_join(noext.z.t.test.df2, z.feat.t.test.df2)
+z.comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+ggplot() +
+  geom_bar(aes(x=perm, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_grid(comp ~ full_df)
+
+z.comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+ggplot() +
+  geom_bar(aes(x=comp, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap( ~ full_df)
+```
+
+Sex-Bias tests
+
+```{r}
+noext.subj.sex.test.z.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
+noext.subj.sex.test.z.df <- bind_rows(noext.subj.sex.test.z.list)
+```
+
+simple bar-plot to count regions w sigificant sex-biases
+
+```{r}
+noext.subj.sex.test.z.df %>%
+  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
+  ggplot(aes(x=perm, fill=sig_fdr)) +
+  geom_bar() +
+  geom_text(stat='count', aes(label=..count..)) +
+  facet_wrap(~dataset) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual(values = c("lightblue", "red"))
+```

### 2024-02-19 eda/lifespan_eda.Rmd
qc by euler number

@@
 ```{r, eval=FALSE}
-gamlss(formula = rh_Vol_posteriorcingulate ~ pb(age_days) + sexMale + pb(sex.age) + site, sigma.formula= ~ pb(age_days) + sexMale + site, data=imagen.df)
+gamlss(formula = rh_Vol_posteriorcingulate ~ pb(age_days) + sexMale + sex.age + site, sigma.formula= ~ pb(age_days) + sexMale + site, data=imagen.df)
 ```

### 2024-02-19 cubic_scripts/qsub_combat_singlestudy.sh

@@ -91,7 +91,7 @@ do
 	touch $bash_script
 
 	#COMBAT GAM
-	elif [ $config = "cf.gam" ]
+	if [ $config = "cf.gam" ]
 	then
 		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gam, formula = y ~ s(age_days, k=5) + sexMale + s(sex.age, k=5)'" > $bash_script
 

### 2024-02-19 eda/ukb_curation-eda.Rmd
testing centile mods with euler

@@
 
-```{r}
+```{r, eval=FALSE}
 ukb_df.ppt.qc <- ukb_df %>%
@@
 ```
+
+##Exploratory QC
+
+Adding euler number (from jakob) back into combat-ed dfs so i can test how controlling for euler in centile scores changes the performance of combat configs (esp.
+in CT based on Bedford et al preprint).
+
+```{r}
+cn.df_euler <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)
+
+ukb.df_euler <- cn.df_euler %>%
+  dplyr::filter(study == "UKB") %>%
+  dplyr::select(participant, surfholes)
+
+table(unique(ukb_df_age.filt$participant) %in% unique(ukb.df_euler$participant)) #all subjects in df!
+
+fwrite(ukb.df_euler, file=paste0(base_path,"data/ukb_euler_nums.csv"))
+```

### 2024-02-19 cubic_scripts/qsub_exploratory_CT_gamlss.sh

@@ -1,0 +1,91 @@
+#!/bin/bash
+
+#Fit GAMLSS growth chart models for each phenotype
+#pass as argument EITHER path to .csv containing data OR directory of .csvs to iterate across
+
+#STEP 3 OF ANALYSIS PIPELINE#
+
+#######################################################################################
+# SET PATHS
+base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
+#base=/Users/megardn/Desktop/BGD_Repos/combat_biovar #base path (local)
+#data_csv=$base/data/ukb_CN_data_agefilt.csv #path to csv with CLEANED DATA (no duplicates, etc)
+ct_pheno_list=$base/pheno_lists/CT_list.txt #path to .txt files listing CT phenotypes
+mod_script=$base/R_scripts/fit_basic_mod_w_euler.R #path to .R script
+
+cd $base #to source functions correctly
+#######################################################################################
+# GET ARGS
+if [ $# -lt 1 ]
+then
+	echo 1>&2 "$0: provide path to data (either single .csv or dir)"
+	exit 2
+fi
+#######################################################################################
+# MAKE DIRECTORIES 
+
+#study dir
+study_dir=$base/ukb_basic
+if ! [ -d $study_dir ]
+	then
+	mkdir $study_dir
+	fi
+
+#qsub script & outputs
+bash_dir=$study_dir/qsub_scripts
+if ! [ -d $bash_dir ]
+	then
+	mkdir $bash_dir
+	fi
+
+#model outputs
+gamlss_dir=$study_dir/gamlss_objs
+if ! [ -d $gamlss_dir ]
+	then
+	mkdir $gamlss_dir
+	fi
+
+#######################################################################################
+# FIT MODELS
+
+#if input .csv
+if [ -f $1 ]
+then
+	csv_name=$(basename $1 .csv)
+	csv_name=${csv_name//_/\-}
+	echo "Pulling data from $csv_name"
+	while read -r pheno
+	do
+			#write bash script
+			bash_script=$bash_dir/${pheno}_${csv_name}_basic_fit.sh
+			touch $bash_script
+			echo "singularity run --cleanenv /cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif Rscript --save $mod_script $1 $pheno $gamlss_dir" > $bash_script
+
+			#qsub bash script
[…]

### 2024-02-19 paper_main_results.Rmd

@@
 
+All features:
+```{r}
+long_abs.diffs <- diffs.df %>%
+  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", pheno_list)) %>%
+  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
+  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         abs.diff = abs.diff*100,
+         pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+long_abs_diff.plt <- long_abs.diffs %>%
+  ggplot(aes(pheno, abs.diff)) +
+  geom_violin(aes(x=pheno, y=abs.diff, fill=dataset, color=dataset), scale="count") +
+  theme_linedraw(base_size = 11) +
+  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Absolulte Centile Error", x = "Feature") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), legend.position="top") +
+  facet_wrap(~pheno_cat, nrow=4, scales="free")
+ggsave("figs/all_abs_err_plt_wlabels.jpeg", long_abs_diff.plt, width=13, height=10.5, units="in")
+
+long_abs.diffs %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot(aes(pheno, abs.diff)) +
+  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=abs.diff,  fill=dataset, color=dataset), scale="count") +
+  theme_linedraw(base_size = 11) +
+  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Absolulte Centile Error", x = "Feature") +
+  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top")
+
+# ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
+```
+
 plot median on brain
@@
 
-intersect(gam_better$pheno, ns_gam_gamlss_comp$pheno)
-```
+cfgam.better <- intersect(gam_better$pheno, ns_gam_gamlss_comp$pheno)
+```
+```{r, eval=FALSE}
+cfgam.better.ct <- gsub("_CT", "", cfgam.better)
+cfgam.better.ct_plot <- dk.parc %>%
+  dplyr::mutate(gam_better = ifelse(label %in% cfgam.better.ct, 1, 0))
+
+ggplot(cfgam.better.ct_plot) +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+            aes(fill=as.factor(gam_better)))
+```
+
 ```{r, eval=FALSE}

### 2024-02-19 paper_replication_results.Rmd

@@
   mutate(label=gsub("_CT", "", pheno),
-         worsefit = ifelse(sig_fdr==TRUE, bigger_group, "NS"))
+         better_fit = case_when(sig_fdr==TRUE & bigger_group == "cf.gam" ~ "cf.gamlss",
+                              sig_fdr==TRUE & bigger_group == "cf.gamlss" ~ "cf.gam",
+                              sig_fdr==FALSE ~ "NS"))
 
@@
              position = position_brain(hemi ~ side),
-            aes(fill=worsefit)) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+            aes(fill=better_fit)) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank(),
+        legend.position="top") +
+  scale_fill_manual(values = c(NS= "darkgray", cf.gam = "#3D75C2", cf.gamlss ="#CFB023FF"), name = "Better ComBat Method") +
   facet_wrap(~perm, nrow=2)

### 2024-02-21 paper_lbcc_results.Rmd
parse results, plots

@@
 
-Combat-GAM Centiles
+Combat-GAM Centiles - site-level batch correction
 
 ```{r}
-gam_centile_df <- fread(file=paste0(der_path, "/cf.gam-batch.study_predictions.csv"))
+gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.site_predictions.csv"))
 dim(gam_centile_df)
@@
 ```{r}
-gamlss_centile_df <- fread(file=paste0(der_path, "/cf.gamlss-batch.study_predictions.csv"))
+gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.site_predictions.csv"))
 dim(gamlss_centile_df)
@@
 ```{r}
-deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
+deltas_df <- fread(file=paste0(der_path, "/batch.site_pred_deltas.csv"))
 # skimr::skim(deltas_df)
@@
   dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+
+tbl.df <- as.data.frame(table(cent_deltas_df$site, cent_deltas_df$study))
 ```
@@
 
-Sample
+Sample plot
 ```{r}
-sample_boxplot <- ggplot(cent_deltas_df, aes(y=reorder(study, -age_years), x=age_years, color=sex)) +
+#reorder studies by age
+age.df <- cent_deltas_df %>%
+  group_by(study) %>%
+  summarize(mean_age = mean(age_years)) %>%
+  arrange(mean_age)
+
+sample_boxplot <- cent_deltas_df %>%
+  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
+ggplot(aes(y=interaction(site,study, sep = "&"), x=age_years)) +
   geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
   theme_linedraw(base_size = 11) +
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank()) +
-  labs(color="Sex", y = "Study")
+  guides(y = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text=element_blank(), ggh4x.axis.nesttext.y = element_text(size=6)) +
+  labs(color="Sex", y = "Site")
 
@@
   cent_deltas_df %>%
-  mutate(ylab = "   Full Sample") %>%
+  mutate(ylab = "  Full \n  Sample") %>%
   ggplot(aes(y=ylab, x=age_years, color=sex)) +
@@
 
-sample_barplot <- ggplot(cent_deltas_df) +
-  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=reorder(study, -age_years), fill=sex)) + 
+sample_barplot <- cent_deltas_df %>%
+  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
+  ggplot() +
+  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=interaction(site,study, sep = "&"), fill=sex)) +
   scale_x_continuous(labels = scales::percent) +
@@
 ```{r}
+delta_n.sex <- cent_deltas_df_long %>%
+  mutate(abs.delta = abs(delta)) %>%
+  group_by(sex, study, site) %>%
+  summarise(sex_n = n(),
+            mean_abs.delta=mean(abs.delta),
+            mean_delta=mean(delta),
+            mean_age_yrs=mean(age_days)/365.25) #get number males and females in each site
+
+#get site-level stats
 delta_summary.df <- cent_deltas_df_long %>%
-  group_by(sex, study) %>%
-  summarise(mean_delta = mean(delta),
-            mean_abs.delta = mean(abs(delta)),
-            max_abs.delta=max(abs(delta)),
-            sex_n = n())
+  mutate(abs.delta = abs(delta)) %>%
+  group_by(study, site) %>%
+  summarise(site_mean_abs.delta=mean(abs.delta),
+            site_mean_delta=mean(delta),
+            site_mean_age=mean(age_days),
+            site_mean_age_yrs=mean(age_days)/365.25,
+            site_max_abs.delta=max(abs.delta),
+            site_n=n())
 
-abs_delta_summary.df <- cent_deltas_df %>%
-  pivot_longer(cols=starts_with("delta."), names_to="pheno", names_prefix = "delta.", values_to = "delta") %>%
-  mutate(abs.delta = abs(delta)) %>%
-  group_by(study) %>%
-  summarise(study_mean_abs.delta=mean(abs.delta),
-            study_mean_delta=mean(delta),
-            mean_age=mean(age_days),
-            mean_age_yrs=mean(age_days)/365.25,
-            study_max_abs.delta=max(abs.delta),
-            study_n=n())
+delta_n.sex %>%
+  dplyr::filter(site=="abide2_BNI_1")
 ```
@@
 ```{r}
-delta_sum.df <- full_join(delta_summary.df, abs_delta_summary.df) %>%
-  mutate(sex_prop=sex_n/study_n,
+delta_sum.by.sex.df <- full_join(delta_summary.df, delta_n.sex) %>%
+  mutate(sex_prop=sex_n/site_n,
          majority = ifelse(sex_prop > .5, 1, 0))
 
-delta_sum.df_piv <- delta_sum.df %>%
-  pivot_wider(id_cols=c(study, study_n, mean_age, mean_age_yrs, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta, max_abs.delta))
+delta_sum.by.sex.df_piv <- delta_sum.by.sex.df %>%
+  pivot_wider(id_cols=c(study, site, site_n, site_mean_age, site_mean_age_yrs, site_mean_delta, site_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta))
 
-#Merge back female prop
-to_merge <- delta_sum.df_piv %>%
-  dplyr::select(study, sex_prop_Female)
+#Merge back female prop back into overall site-level stats
+to_merge <- delta_sum.by.sex.df_piv %>%
+  dplyr::select(study, site, sex_prop_Female) %>%
+  mutate_at(vars(sex_prop_Female), ~replace_na(., 0)) #set prop Female to 0 if no females in study
 
-delta_sum.df2 <- full_join(delta_sum.df, to_merge)
+delta_sum.df_merged <- full_join(delta_summary.df, to_merge)
 ```
@@
 ```{r}
-delta_sum.df2 %>%
-  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=sex)) +
-  geom_count(aes(size=study_n)) +
+delta_sum.df_merged %>%
+  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
+  geom_count(aes(size=site_n)) +
   geom_smooth(method="lm")
 
-delta_sum.df2 %>%
-  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=study, shape=sex)) +
-  geom_point() +
-  geom_smooth(aes(color=sex, linetype=sex), method="lm")
+delta_sum.by.sex.df %>%
+  ggplot(aes(x=sex, y=mean_delta, color=site, fill=site)) +
+  geom_count(aes(size=sex_prop), alpha=0.5) +
+  geom_line(aes(group=site)) +
+  theme(legend.position="none")
 
-delta_sum.df2 %>%
-  ggplot(aes(x=sex_prop_Female, y=mean_delta, color=sex)) +
-  geom_point() +
-  geom_smooth(aes(color=sex, linetype=sex), method="lm")
+delta_sum.by.sex.df %>%
+  ggplot(aes(x=sex, y=mean_delta, color=as.factor(majority), fill=as.factor(majority))) +
+  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
+  geom_line(aes(group=site), color="darkgray")
 
-delta_sum.df %>%
-  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
-  geom_count(aes(size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=study))
-
-delta_sum.df %>%
-  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
-  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=study))
-
-delta_sum.df %>%
+delta_sum.by.sex.df %>%
   ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age_yrs)) +
@@
 
-delta_sum.df_piv %>%
-  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta, color=mean_age_yrs)) +
-  geom_count(aes(size=study_n)) +
+delta_sum.by.sex.df_piv %>%
+  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta, color=study)) +
+  geom_count(aes(size=site_n)) +
   geom_smooth(method="lm")
@@
 
-#Across-Feature Mean Results
+look at max abs. delta across sites
+```{r}
+delta_sum.df_merged %>%
+  ggplot() +
+  geom_histogram(aes(x=site_max_abs.delta*100))
 
+delta_sum.df_merged %>%
+  ggplot(aes(y=site_max_abs.delta*100, x=site_n)) +
+  geom_point() +
+  geom_smooth(method="lm")
+
+delta_sum.df_merged %>%
+  ggplot(aes(y=site_max_abs.delta*100, x=site_mean_age_yrs)) +
+  geom_point() +
+  geom_smooth(method="lm")
+
+delta_sum.df_merged %>%
+  ggplot(aes(y=site_max_abs.delta*100, x=sex_prop_Female)) +
+  geom_point() +
+  geom_smooth(method="lm")
+```
+
+##Feature-wise
+do deltas differ by feature?
 ```{r}
-# age
-summary(lm(study_mean_abs.delta ~ mean_age_yrs, data=delta_sum.df_piv))
-summary(lm(study_mean_delta ~ mean_age_yrs, data=delta_sum.df_piv))
+cent_deltas_df_long2 <- cent_deltas_df_long %>%
+   mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+cent_deltas_df_long2 %>%
+  ggplot() +
+  geom_histogram(aes(x=delta, fill=pheno), position="identity", alpha=0.3) +
+  theme(legend.position="none")
+
+cent_deltas_df_long2 %>%
+  ggplot() +
+  geom_histogram(aes(x=delta*100, fill=pheno_cat, color=pheno_cat), position="identity", alpha=0.7, bins=60)
+
+cent_deltas_df_long2 %>%
+  ggplot() +
+  geom_histogram(aes(x=delta*100, fill=study, color=study), position="identity", alpha=0.3, bins=60) +
+  facet_wrap(~pheno_cat, scales="free") +
+  theme(legend.position="none")
 ```
+
+#Site-Level Mean Results
 
@@
 ```{r}
-delta_sum.df %>%
-  ggplot(aes(x=mean_age_yrs, y=mean_abs.delta, color=sex)) +
-  geom_count(aes(size=study_n)) +
+# age
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
+
+#forward selection for site_mean_abs.delta
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
+full.delta.age.fit <- lm((site_mean_abs.delta*100) ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
+sjPlot::plot_model(full.delta.age.fit, type = "pred", terms = "site_mean_age_yrs")
+```
+
+```{r}
+delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
+  geom_count(aes(size=site_n)) +
   geom_smooth(method="lm")
 
-age_effects.plt <- delta_sum.df_piv %>%
-  ggplot(aes(x=mean_age_yrs, y=(study_mean_abs.delta*100))) +
-  geom_count(aes(size=study_n, color=sex_prop_Female)) +
+age_effects.plt <- delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
+  geom_count(aes(size=site_n, color=sex_prop_Female)) +
   paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
@@
   theme_linedraw(base_size = 11) +
-  labs(size="Study N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores")
+  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores")
+
+# skim(delta_sum.df_piv) #some missing data! check on this!
 ```
 
-are studies w younger subjects changing more because there are fewer of them?
+are studies w older subjects changing more because there are fewer of them?
 check for association between study size and age
 ```{r}
-summary(lm(mean_age_yrs ~ study_n, data=delta_sum.df_piv))
-fit <- lm((study_mean_abs.delta*100) ~ mean_age_yrs + study_n, data=delta_sum.df_piv)
-summary(fit)
-sjPlot::plot_model(fit, type = "pred", terms = "mean_age_yrs")
-
-summary(lm((study_mean_abs.delta*100) ~ mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_piv))
+summary(lm(site_mean_age_yrs ~ site_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
 ```
@@
 # %female
-summary(lm(study_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_piv))
-summary(lm(study_mean_delta ~ sex_prop_Female, data=delta_sum.df_piv))
-summary(lm(study_mean_delta ~ sex_prop_Female + study_n, data=delta_sum.df_piv)) #add study size covariate
-summary(lm(study_mean_delta ~ sex_prop_Female + study_n + mean_age, data=delta_sum.df_piv)) #add study size & age covariates
+summary(lm(site_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ sex_prop_Female, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ sex_prop_Female + site_n, data=delta_sum.df_merged)) #add site size covariate
+summary(lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged)) #add site size & age covariates
+full.delta.sex.fit <- lm((100*site_mean_delta) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged) #scaling for plot
+
+sjPlot::plot_model(full.delta.sex.fit, type = "pred", terms = "sex_prop_Female") 
 ```
@@
 ```{r}
-delta_sum.df %>%
-  ggplot(aes(x=sex, y=(mean_delta*100), color=study, fill=study)) +
+delta_sum.by.sex.df %>%
+  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
   geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=study))
+  geom_line(aes(group=site))
 
-sex_effects.plt <- delta_sum.df_piv %>%
-  ggplot(aes(x=sex_prop_Female, y=(study_mean_delta*100))) +
-  geom_count(aes(size=study_n, color=mean_age_yrs)) +
+sex_effects.plt <- delta_sum.df_merged %>%
+  ggplot(aes(x=sex_prop_Female, y=(site_mean_delta*100))) +
+  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
   geom_smooth(method="lm", color="#24693DFF") +
@@
   theme_linedraw(base_size = 11) +
-  labs(size="Study N", x = "Proportion Female", y = "Mean Difference in Centile Scores")
+  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores")
 
-delta_sum.df %>%
+delta_sum.by.sex.df %>%
   ggplot(aes(x=sex, y=(mean_delta*100))) +
@@
 # age*%female
-summary(lm(study_mean_abs.delta ~ mean_age_yrs*sex_prop_Female, data=delta_sum.df_piv))
-summary(lm(study_mean_delta ~ mean_age_yrs*sex_prop_Female, data=delta_sum.df_piv))
-```
-
-#Featurewise Results
-
-Linear models
-```{r}
-cent_delta_lms.df <- fread(file=paste0(der_path, "/batch.study_pred_cent_delta_lms.csv"))
-n_unique(cent_delta_lms.df$pheno)
-```
-
-summarise findings
-
-```{r}
-cent_delta_lms.df %>%
-  dplyr::filter(term=="sexMale") %>%
-  dplyr::summarize(n_sig_regions = sum(sig_fdr),
-                   n_regions=nrow(.),
-              mean_est = mean(estimate)) %>%
-    ungroup() %>%
-    as.data.frame()
-```
-
-```{r}
-cent_delta_lms.df %>%
-  dplyr::filter(term=="age_yrs") %>%
-  dplyr::summarize(n_sig_regions = sum(sig_fdr),
-                   n_regions=nrow(.),
-              mean_est = mean(estimate)) %>%
-    ungroup() %>%
-    as.data.frame()
-```
-
-```{r}
-cent_delta_lms.df %>%
-  dplyr::filter(term=="sexMale:age_yrs") %>%
-  dplyr::summarize(n_sig_regions = sum(sig_fdr),
-                   n_regions=nrow(.),
-              mean_est = mean(estimate)) %>%
-    ungroup() %>%
-    as.data.frame()
-```
-
-```{r}
-sig.cent_delta_lms.df <- cent_delta_lms.df %>%
-  dplyr::filter(sig_fdr==TRUE)
-n_unique(sig.cent_delta_lms.df$pheno)
-
-table(sig.cent_delta_lms.df$term, sig.cent_delta_lms.df$pheno)
-
-max(sig.cent_delta_lms.df$estimate)*100
-min(sig.cent_delta_lms.df$estimate)*100
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
 ```
@@
 
-Plot betas (*100 for centile space) on brain!
-
-```{r}
-cent_delta_lms.df_toplt <- cent_delta_lms.df %>%
-   mutate(pheno_cat = as.factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_)),
-    label = sub("_[^_]*_", "_", pheno)) #for plotting
-```
-
-```{r}
-term_names <- c(sexMale = "Sex", age_yrs = "Age", 'sexMale:age_yrs' = "Sex*Age")
-
-#Regional
-regional.delta_plt <- cent_delta_lms.df_toplt %>%
-  dplyr::filter(pheno_cat != "Global Volume") %>%
-  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
-                                     TRUE ~ estimate*100)) %>%
-  ggplot() +
-  theme_linedraw(base_size=12) +
-  ggseg::geom_brain(atlas = dk_nomed, 
-             position = position_brain(hemi ~ side),
-             aes(fill = estimate)) +
-  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names))) + 
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.11, 0.11), breaks=c(-0.1, 0, 0.1))
-
-#Global
-global.delta_plt <- cent_delta_lms.df_toplt %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  mutate(estimate = case_when(sig_fdr == FALSE ~ NA,
-                                     TRUE ~ estimate*100),
-         tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA))) %>%
-  dplyr::select(!any_of(c("label", "region", "hemi"))) %>% #drop cols corresponding to dk regions
-  ggplot() +
-  theme_linedraw(base_size=12) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             side = c("coronal"),
-             aes(fill = estimate)) +
-  facet_grid(term ~ pheno_cat, labeller = labeller(term = as_labeller(term_names)), switch = 'y') + 
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Beta", limits = c(-0.11, 0.11), breaks=c(-0.1, 0, 0.1))
-```
-
 #Site Effects
@@
   # print(pheno)
-  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["study"]]))
+  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["site"]]))
   df$pheno <- pheno
@@
 
+Test single pheno
+
 ```{r}
@@
   ggplot() +
-  geom_boxplot(aes(x=study, y=WMV)) +
+  geom_boxplot(aes(x=site, y=WMV)) +
   theme(axis.text.x = element_text(angle=90, hjust=0.5))
 
-summary(lm(WMV ~ study + age_days + sex, data = gam_centile_df.clean))
+summary(lm(WMV ~ site + age_days + sex, data = gam_centile_df.clean))
 ```
@@
   ggplot() +
+  geom_boxplot(aes(x=site, y=mean_cent)) +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5))
+
+gam_centile_df.mean %>%
+  ggplot() +
   geom_boxplot(aes(x=study, y=mean_cent)) +
@@
 ```{r}
+summary(lm(mean_cent ~ site, data=gam_centile_df.mean))
 summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
 
-summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
-pairwise.t.test(gam_centile_df.mean$mean_cent, gam_centile_df.mean$study, p.adjust.method = "fdr")
+summary(aov(mean_cent ~ site, data=gam_centile_df.mean))
+```
 
-mean_cent_gam.lm <- lm(mean_cent ~ study + age_days + sex, data=gam_centile_df.mean)
-
-gtsummary::tbl_regression(mean_cent_gam.lm) %>%
-  gtsummary::add_global_p()
-```
+success!
 
@@
   # print(pheno)
-  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["study"]]))
+  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["site"]]))
   df$pheno <- pheno
@@
   summarize(study = unique(study),
+            site = unique(site),
             age_days=unique(age_days),
@@
 ```{r}
+summary(lm(mean_z ~ site, data=gam_z_df.mean))
 summary(lm(mean_z ~ study, data=gam_z_df.mean))
@@
 summary(aov(mean_z ~ study, data=gam_z_df.mean))
-pairwise.t.test(gam_z_df.mean$mean_z, gam_z_df.mean$study, p.adjust.method = "fdr")
-
-mean_z_gam.lm <- lm(mean_z ~ study + age_days + sex, data=gam_z_df.mean)
-summary(mean_z_gam.lm)
 ```
@@
   # print(pheno)
-  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
+  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["site"]]))
   df$pheno <- pheno
@@
   ggplot() +
-  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
+  geom_boxplot(aes(x=reorder(site, age_days), y=mean_cent)) + 
   theme(axis.text.x = element_text(angle=90))
@@
   ggplot() +
-  geom_violin(aes(x=reorder(study, age_days), y=mean_cent), scale="count") + 
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
   theme(axis.text.x = element_text(angle=90))
@@
 ```{r}
+summary(lm(mean_cent ~ site, data=gamlss_centile_df.mean))
 summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
 
-summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
-pairwise.t.test(gamlss_centile_df.mean$mean_cent, gamlss_centile_df.mean$study, p.adjust.method = "fdr")
+summary(aov(mean_cent ~ site, data=gamlss_centile_df.mean))
+```
 
-mean_cent_gamlss.lm <- lm(mean_cent ~ study + age_days + sex + fs_version, data=gamlss_centile_df.mean)
-summary(mean_cent_gamlss.lm)
-
-mean_cent_gamlss.gam <- gam(mean_cent ~ study + s(log_age) + sex, data=gamlss_centile_df.mean)
-summary(mean_cent_gamlss.gam)
-
-mean_cent_gam.gam <- gam(mean_cent ~ study + s(log_age) + sex, data=gam_centile_df.mean)
-summary(mean_cent_gam.gam)
-
-gtsummary::tbl_regression(mean_cent_gamlss.lm) %>%
-  gtsummary::add_global_p()
-```
+success!
 
@@
   # print(pheno)
-  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
+  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["site"]]))
   df$pheno <- pheno
@@
   summarize(study = unique(study),
+            site=unique(site),
             age_days=unique(age_days),
@@
 ```{r}
+summary(lm(mean_z ~ site, data=gamlss_z_df.mean))
 summary(lm(mean_z ~ study, data=gamlss_z_df.mean))
@@
 summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
-pairwise.t.test(gamlss_z_df.mean$mean_z, gamlss_z_df.mean$study, p.adjust.method = "fdr")
-
-mean_cent_gamlss.lm <- lm(mean_z ~ study + age_days + sex, data=gamlss_z_df.mean)
-summary(mean_cent_gamlss.lm)
 ```
@@
 fig5 <- ggarrange(
-  ggarrange(
 ggarrange(sample_boxplot, sample_barplot,
@@
           widths=c(2, 1),
-          heights=c(6, 1)),
-
-# featurewise effects
-ggarrange(global.delta_plt, regional.delta_plt,
-          nrow=1, ncol=2, 
-          align = "hv", 
-          common.legend = TRUE,
-          legend = "bottom",
-          widths=c(1, 2.5)
-          ),
-nrow=2, ncol=1,
-heights=c(1.5, 1),
-labels=c("A", "C")
-),
+          heights=c(14, 1)),
 # mean effects
@@
 align="hv",
-labels=c("B", "D")),
+labels=c("B", "C")),
 common.legend=FALSE,
 nrow=1, ncol=2,
+labels=c("A", NA),
 widths=c(1, 1.3)

### 2024-02-23 paper_lbcc_results.Rmd
looking at CT sigmas, plots

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects)
 #Sys.setenv("MC_CORES"=10L)
@@
   summarize(mean_age = mean(age_years)) %>%
-  arrange(mean_age)
+  arrange(-mean_age)
 
@@
 ```
-
-##Feature-wise
-do deltas differ by feature?
-```{r}
-cent_deltas_df_long2 <- cent_deltas_df_long %>%
-   mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-
-cent_deltas_df_long2 %>%
-  ggplot() +
-  geom_histogram(aes(x=delta, fill=pheno), position="identity", alpha=0.3) +
-  theme(legend.position="none")
-
-cent_deltas_df_long2 %>%
-  ggplot() +
-  geom_histogram(aes(x=delta*100, fill=pheno_cat, color=pheno_cat), position="identity", alpha=0.7, bins=60)
-
-cent_deltas_df_long2 %>%
-  ggplot() +
-  geom_histogram(aes(x=delta*100, fill=study, color=study), position="identity", alpha=0.3, bins=60) +
-  facet_wrap(~pheno_cat, scales="free") +
-  theme(legend.position="none")
-```
-
-#Site-Level Mean Results
-
-delta ~ age
-
-```{r}
-# age
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
-
-#forward selection for site_mean_abs.delta
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
-full.delta.age.fit <- lm((site_mean_abs.delta*100) ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
-sjPlot::plot_model(full.delta.age.fit, type = "pred", terms = "site_mean_age_yrs")
-```
-
-```{r}
-delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
-  geom_count(aes(size=site_n)) +
-  geom_smooth(method="lm")
-
-age_effects.plt <- delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
-  geom_count(aes(size=site_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
-  geom_smooth(method="lm", color="#2A5783FF") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores")
-
-# skim(delta_sum.df_piv) #some missing data! check on this!
-```
-
-are studies w older subjects changing more because there are fewer of them?
-check for association between study size and age
-```{r}
-summary(lm(site_mean_age_yrs ~ site_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
-```
-
-```{r}
-cent_deltas_df %>%
-  ggplot() +
-  geom_histogram(aes(x=age_days))
-```
-
-delta ~ % female
-
-```{r}
-# %female
-summary(lm(site_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ sex_prop_Female, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ sex_prop_Female + site_n, data=delta_sum.df_merged)) #add site size covariate
-summary(lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged)) #add site size & age covariates
-full.delta.sex.fit <- lm((100*site_mean_delta) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged) #scaling for plot
-
-sjPlot::plot_model(full.delta.sex.fit, type = "pred", terms = "sex_prop_Female") 
-```
-
-```{r}
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
-  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=site))
-
-sex_effects.plt <- delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=(site_mean_delta*100))) +
-  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
-  geom_smooth(method="lm", color="#24693DFF") +
-  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores")
-
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=(mean_delta*100))) +
-  geom_boxplot()
-```
-
-```{r}
-# age*%female
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
-```
-
-## Plots
-
-```{r}
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
-  geom_point(alpha=0.4) +
-  geom_smooth(method="lm")
-
-cent_deltas_df %>%
-  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
-  geom_histogram(aes(y=after_stat(c(
-      count[group==1]/sum(count[group==1]),
-      count[group==2]/sum(count[group==2])
-    )*100)), alpha=0.4, position="identity", bins=60) +
-  ylab("Percentage")
-
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm")
-
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
-  geom_point(alpha=0.4) +
-  geom_smooth(method="lm")
-```
-
 #Site Effects
@@
 
-#FIGURE 4
+#Site-Level Mean Results
+
+##Age
+
+delta ~ age
+
+```{r}
+# age
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
+
+#forward selection for site_mean_abs.delta
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
+full.delta.age.fit <- lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
+
+full.delta.age.fit2 <- lm(site_mean_delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
+summary(full.delta.age.fit2)
+```
+
+```{r}
+delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=site_mean_delta)) +
+  geom_count(aes(size=site_n)) +
+  geom_smooth(method="lm")
+delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
+  geom_count(aes(size=site_n)) +
+  geom_smooth(method="lm")
+
+age_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
+age_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]
+
+age_effects.plt <- delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
+  geom_count(aes(size=site_n, color=sex_prop_Female)) +
+  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
+  geom_smooth(method="lm", color="#2A5783FF") +
+  theme_linedraw(base_size = 11) +
+  annotate("text", x=60, y=4, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
+  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
+  theme(axis.title.x=element_b)
+
+age_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
+age_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]
+
+age_effects.plt.null <- delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_delta*100))) +
+  geom_count(aes(size=site_n, color=sex_prop_Female)) +
+  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
+  geom_smooth(method="lm", color="#2A5783FF") +
+  theme_linedraw(base_size = 11) +
+  annotate("text", x=60, y=-0.3, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
+  labs(size="site N", x = "Mean Age (Years)", y = "Mean Difference in Centile Scores")
+
+#add prediction to df
+delta_sum.df_merged$abs.delta_pred <- predict(full.delta.age.fit)
+
+full.delta.age.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$x,
+                                    predicted=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$predicted,
+                                    conf.low=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.low,
+                                    conf.high=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.high)
+full.delta.age.fit.df <- full.delta.age.fit.df %>%
+  mutate(predicted = predicted*100,
+         conf.low=conf.low*100,
+         conf.high=conf.high*100)
+
+marg.age_effects.plt <- ggplot() +
+  geom_count(data=delta_sum.df_merged, aes(x=site_mean_age_yrs, y=(abs.delta_pred*100), size=site_n, color=sex_prop_Female)) +
+  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
+  geom_line(aes(x=full.delta.age.fit.df$x, y = full.delta.age.fit.df$predicted), size = 0.7, color="#2A5783FF") +
+  geom_ribbon(aes(x=full.delta.age.fit.df$x,y= NULL, ymin= full.delta.age.fit.df$conf.low, ymax= full.delta.age.fit.df$conf.high), alpha=0.1, fill="#2A5783FF") +
+  theme_linedraw(base_size = 11) +
+  labs(size="site N", x = "Mean Age (Years)", y = "Predicted Mean Absolute Difference in Centile Scores")
+```
+
+are studies w older subjects changing more because there are fewer of them?
+check for association between study size and age
+```{r}
+summary(lm(site_mean_age_yrs ~ site_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
+```
+
+```{r}
+cent_deltas_df %>%
+  ggplot() +
+  geom_histogram(aes(x=age_days))
+```
+
+##Prop Female
+
+delta ~ % female
+
+```{r}
+# %female
+summary(lm(site_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ sex_prop_Female, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ sex_prop_Female + site_n, data=delta_sum.df_merged)) #add site size covariate
+summary(lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged)) #add site size & age covariates
+full.delta.sex.fit <- lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged) #scaling for plot
+
+sjPlot::plot_model(full.delta.sex.fit, type = "pred", terms = "sex_prop_Female") 
+```
+
+```{r}
+delta_sum.by.sex.df %>%
+  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
+  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
+  geom_line(aes(group=site))
+```
+
+```{r}
+delta_sum.df_merged %>%
+  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
+  geom_count(aes(size=site_n)) +
+  geom_smooth(method="lm")
+delta_sum.df_merged %>%
+  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta)) +
+  geom_count(aes(size=site_n)) +
+  geom_smooth(method="gam")
+
+sex_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
+sex_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]
+
+sex_effects.plt <- delta_sum.df_merged %>%
+  mutate(site_mean_delta=site_mean_delta*100) %>%
+  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
+  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
+  geom_smooth(method="lm",
+              color="#24693DFF") +
+  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
+  theme_linedraw(base_size = 11) +
+  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores") +
+  theme(axis.title.y=element_blank()) +
+  annotate("text", x=0.25, y=-0.3, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
+  guides(size = "none")
+
+sex_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
+sex_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]
+
+sex_effects.plt.null <- delta_sum.df_merged %>%
+  ggplot(aes(x=sex_prop_Female, y=(site_mean_abs.delta*100))) +
+  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
+  geom_smooth(method="lm", color="#24693DFF") +
+  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
+  theme_linedraw(base_size = 11) +
+  labs(size="site N") +
+    annotate("text", x=0.25, y=4, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
+  theme(axis.title=element_blank()) +
+  guides(size = "none")
+
+#add prediction to df
+delta_sum.df_merged$delta_pred <- predict(full.delta.sex.fit)
+
+full.delta.sex.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$x,
+                                    predicted=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$predicted,
+                                    conf.low=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.low,
+                                    conf.high=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.high)
+full.delta.sex.fit.df <- full.delta.sex.fit.df %>%
+  mutate(predicted = predicted*100,
+         conf.low=conf.low*100,
+         conf.high=conf.high*100)
+
+marg.sex_effects.plt <- ggplot() +
+  geom_count(data=delta_sum.df_merged, aes(x=sex_prop_Female, y=(delta_pred*100), size=site_n, color=site_mean_age_yrs)) +
+  geom_line(aes(x=full.delta.sex.fit.df$x, y = full.delta.sex.fit.df$predicted), size = 0.7, color="#24693DFF") +
+  geom_ribbon(aes(x=full.delta.sex.fit.df$x,y= NULL, ymin= full.delta.sex.fit.df$conf.low, ymax= full.delta.sex.fit.df$conf.high), alpha=0.1, fill="#24693DFF") +
+  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
+  theme_linedraw(base_size = 11) +
+  labs(size="site N", x = "Proportion Female", y = "Predicted Mean Difference in Centile Scores")
+
+delta_sum.by.sex.df %>%
+  ggplot(aes(x=sex, y=(mean_delta*100))) +
+  geom_boxplot()
+```
+
+```{r}
+# age*%female
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
+```
+
+```{r}
+cent_deltas_df %>%
+  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
+  geom_point(alpha=0.4) +
+  geom_smooth(method="lm")
+
+cent_deltas_df %>%
+  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
+  geom_histogram(aes(y=after_stat(c(
+      count[group==1]/sum(count[group==1]),
+      count[group==2]/sum(count[group==2])
+    )*100)), alpha=0.4, position="identity", bins=60) +
+  ylab("Percentage")
+
+cent_deltas_df %>%
+  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm")
+
+cent_deltas_df %>%
+  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
+  geom_point(alpha=0.4) +
+  geom_smooth(method="lm")
+```
+
+#FIGURE 5
 
@@
 # mean effects
-ggarrange(age_effects.plt,
-sex_effects.plt,
+ggarrange(marg.age_effects.plt,
+marg.sex_effects.plt,
 nrow=2, ncol=1,
@@
 labels=c("A", NA),
-widths=c(1, 1.3)
+widths=c(1, 1.4)
 )
@@
 ```
+
+```{r}
+# study demographics
+fig5.optB <- ggarrange(
+ggarrange(sample_boxplot, sample_barplot,
+          full_boxplot, full_barplot,
+          nrow=2, ncol=2, 
+          # align = "hv", 
+          common.legend = TRUE,
+          legend = "bottom",
+          widths=c(2, 1.3),
+          heights=c(14, 1)),
+# mean effects
+ggarrange(
+ggarrange(age_effects.plt, age_effects.plt.null,
+          nrow=2, ncol=1, common.legend=TRUE, legend="bottom",
+          align = "hv"),
+ggarrange(sex_effects.plt.null, sex_effects.plt,
+          nrow=2, ncol=1, common.legend=TRUE, legend="bottom",
+          align = "hv"),
+# align="hv",
+# labels=c("B", "C")),
+common.legend=FALSE,
+nrow=1, ncol=2),
+labels=c("A", "B"),
+widths=c(1, 2.2),
+nrow=1, ncol=2
+)
+
+fig5.optB_annotate <- annotate_figure(fig5.optB, top = text_grob("Figure 5.
+Differences in centile scores of lifespan consortium data when harmonized with ComBat-GAM and ComBatLS", color = "black", hjust = 0, x = 0))
+fig5.optB_annotate
+ggsave("figs/figure_5_option2.jpeg", fig5.optB_annotate, width=13, height=10, units="in")
+```

### 2024-02-23 paper_main_results.Rmd
looking at CT sigmas, plots

@@
 ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
+```
+Look at relative estimates for sex in sigma across features in raw data
+
+```{r}
+gamlss.sum.df <- fread(paste0(data_path, "/ukb_basic/gamlss_summary.csv"))
+gamlss.sum.df.raw <- gamlss.sum.df %>%
+  dplyr::filter(dataset=="ukb-CN-data-simsites")
+
+gamlss.sum.df.raw.sexsig <- gamlss.sum.df.raw %>%
+  dplyr::filter(parameter =="sigma" & term == "sexMale") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      est_over_moment = estimate/moment.mean)
+
+gamlss.sum.df.raw.sexsig %>%
+  ggplot(aes(y=estimate, x=pheno_cat, fill=pheno_cat)) +
+  geom_boxplot() +
+  # geom_hline(yintercept = mean(gamlss.sum.df.raw.sexsig$t_stat), linetype = 2) +
+  stat_compare_means(method = "anova", label.y = 0.07)+    
+  stat_compare_means(label = "p.signif", method = "t.test",
+                     ref.group = ".all.")
+
+gamlss.sum.df.raw.sexsig %>%
+  ggplot() +
+  geom_histogram(aes(x=est_over_moment, fill=pheno_cat, color=pheno_cat), alpha=0.5, position="identity")
+
+gamlss.sum.df.raw.sexsig %>%
+  ggplot(aes(y=t_stat, x=pheno_cat, fill=pheno_cat)) +
+  geom_boxplot() +
+  # geom_hline(yintercept = mean(gamlss.sum.df.raw.sexsig$t_stat), linetype = 2) +
+  stat_compare_means(method = "anova", label.y = -7)+    
+  stat_compare_means(label = "p.signif", method = "t.test",
+                     ref.group = ".all.")
 ```

### 2024-02-23 paper_prop_results.Rmd
looking at CT sigmas, plots

@@
 ```{r}
-der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"
+data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
 
@@
 ```{r}
-raw_files <- list.files(path = paste0(der_path, "/ukb_ratio"), pattern = "_diffs.csv", full.names = TRUE)
+raw_files <- list.files(path = data_path, pattern = "_diffs.csv", full.names = TRUE)
 
@@
 ```{r}
-raw_files <- list.files(path = paste0(der_path, "/ukb_ratio"), pattern = "_subj_pred.csv", full.names = TRUE)
+raw_files <- list.files(path = data_path, pattern = "_subj_pred.csv", full.names = TRUE)
 
@@
 ```{r, eval=FALSE}
-pred.csvs <- list.files(path = paste0(der_path, "/ukb_ratio"), pattern = "raw_predictions.csv", full.names = TRUE)
+pred.csvs <- list.files(path = data_path, pattern = "raw_predictions.csv", full.names = TRUE)
 
@@
   facet_wrap(~prop, nrow=1, labeller = as_labeller(percent_f))
+sampling.plt
 ```
@@
 Do centile score absolute errors change based on combat version? Tested by `ratio_results.R`, results written to `full_featurewise_cent_t_tests.csv`
-*need to look and see what happens for subjects who aren't in all samples?*
 
 ```{r}
-cent.t.test.df <- fread(paste0(der_path, "/ukb_ratio/full_featurewise_cent_t_tests.csv"))
-cent.t.test.df <- cent.t.test.df %>%
-  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
-         comp = as.factor(comp))
+cent.feat.t.test <- readRDS(paste0(data_path, "/full_featurewise_cent_t_tests_all_out.RDS"))
+# names(cent.feat.t.test) <- n_perm_list
+cent.t.test.df <- bind_rows(cent.feat.t.test, .id="prop") %>%
+  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+         comp = as.factor(comp),
+         pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+cent.t.test.df %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig=sum(sig_fdr))
 ```
@@
 cent.t.test.df %>%
-  # filter(comp=="cf.gamlss_cf.gam") %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
 ggplot() +
-  geom_col(aes(x=prop, y=n_sig_regions, fill=comp), position="dodge") +
-  facet_wrap(~comp) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
+  geom_bar(aes(x=prop, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~comp)
 ```
@@
 ```{r}
-cent.t.test.df_global <- fread(paste0(der_path, "/ukb_ratio/full_featurewise_cent_t_tests_global.csv"))
+cent.t.test.df_global <- fread(paste0(data_path, "/full_featurewise_cent_t_tests_global.csv"))
 
@@
 Also visualize by sex
-```{r}
+```{r, eval=FALSE}
 prop.global_cent_errs_plot_bysex <- prop.global_abs.diffs %>%
@@
 ```{r}
-cent.subj.mean.test.df <- fread(paste0(der_path, "/ukb_ratio/subj.abs.mean_sex_bias_cent_t_tests.csv"))
+cent.subj.mean.test.df <- fread(paste0(data_path, "/subj.abs.mean_sex_bias_cent_t_tests.csv"))
 cent.subj.mean.test.df <- cent.subj.mean.test.df %>%
@@
 ```{r}
-cent.subj.sex.test.list <- readRDS(paste0(der_path, "/ukb_ratio/full_featurewise_cent_sex_bias_tests.RDS"))
+cent.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_cent_sex_bias_tests.RDS"))
 cent.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
@@
 
-```{r}
+```{r, eval=FALSE}
 cent.subj.sex.test.df %>%
@@
 Plot on brains for 60% F imbalance:
-```{r}
+```{r, eval=FALSE}
 aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
@@
 ```{r}
-noext.cent.t.test.df <- fread(paste0(der_path, "/ukb_ratio/no.ext_featurewise_cent_t_tests.csv"))
-noext.cent.t.test.df <- noext.cent.t.test.df %>%
-  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
-         comp = as.factor(comp))
+noext.cent.feat.t.test <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_t_tests_all_out.RDS"))
+noext.cent.t.test.df <- bind_rows(noext.cent.feat.t.test, .id="prop") %>%
+  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+         comp = as.factor(comp),
+         pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
 noext.cent.t.test.df
@@
 ```{r}
-noext.cent.t.test.df_reduced <- noext.cent.t.test.df %>%
-  mutate(full_df = 0) %>%
-  dplyr::select(comp, prop, n_sig_regions, full_df)
-cent.t.test.df_reduced <- cent.t.test.df %>%
-  mutate(full_df = 1) %>%
-  dplyr::select(comp, prop, n_sig_regions, full_df)
+noext.cent.t.test.df2 <- noext.cent.t.test.df %>%
+  mutate(full_df = "reduced")
+cent.t.test.df2 <- cent.t.test.df %>%
+  mutate(full_df = "full")
 
-full_join(noext.cent.t.test.df_reduced, cent.t.test.df_reduced) %>%
+comp.df <- full_join(noext.cent.t.test.df2, cent.t.test.df2)
+comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
 ggplot() +
-  geom_col(aes(x=prop, y=n_sig_regions, fill=as.factor(full_df)), position="dodge") +
-  facet_wrap(~comp) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  geom_bar(aes(x=prop, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_grid(comp ~ full_df)
+
+comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+ggplot() +
+  geom_bar(aes(x=comp, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap( ~ full_df)
 ```
@@
 ```{r}
-noext_cent.subj.sex.test.list <- readRDS(paste0(der_path, "/ukb_ratio/no.ext_featurewise_cent_sex_bias_tests.RDS"))
+noext_cent.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
 noext_cent.subj.sex.test.df <- bind_rows(noext_cent.subj.sex.test.list)
@@
 ```{r}
-z.t.test.df <- fread(paste0(der_path, "/ukb_ratio/full_featurewise_z_t_tests.csv"))
-z.t.test.df <- z.t.test.df %>%
-  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
-         comp = as.factor(comp))
+z.feat.t.test <- readRDS(paste0(data_path, "/full_featurewise_z_t_tests_all_out.RDS"))
+
+z.feat.t.test.df <- bind_rows(z.feat.t.test, .id="prop") %>%
+  mutate(pheno = sub(".z", "", pheno)) %>%
+  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
+                pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+z.feat.t.test.df %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig=sum(sig_fdr))
 ```
@@
 ```{r}
-z.t.test.df %>%
-  # filter(comp=="cf.gamlss_cf.gam") %>%
+z.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
 ggplot() +
-  geom_col(aes(x=prop, y=n_sig_regions, fill=comp), position="dodge") +
-  facet_wrap(~comp) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
+  geom_bar(aes(x=prop, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~comp)
 ```
@@
 ```{r}
-z.subj.mean.test.df <- fread(paste0(der_path, "/ukb_ratio/subj.abs.mean_sex_bias_z_t_tests.csv"))
+z.subj.mean.test.df <- fread(paste0(data_path, "/subj.abs.mean_sex_bias_z_t_tests.csv"))
 z.subj.mean.test.df <- z.subj.mean.test.df %>%
@@
 ```{r}
-z.subj.sex.test.list <- readRDS(paste0(der_path, "/ukb_ratio/full_featurewise_z_sex_bias_tests.RDS"))
+z.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
 z.subj.sex.test.df <- bind_rows(z.subj.sex.test.list)
@@
 ```{r}
-noext.z.t.test.df <- fread(paste0(der_path, "/ukb_ratio/no.ext_featurewise_z_t_tests.csv"))
-noext.z.t.test.df <- noext.z.t.test.df %>%
-  mutate(prop = factor(column_label, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
-         comp = as.factor(comp))
-noext.z.t.test.df
+noext.z.feat.t.test <- readRDS(paste0(data_path, "/no.ext_featurewise_z_t_tests_all_out.RDS"))
+noext.z.feat.t.test.df <- bind_rows(noext.z.feat.t.test, .id="prop") %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+noext.z.feat.t.test.df %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig=sum(sig_fdr))
 ```
@@
 ```{r}
-noext.z.t.test.df_reduced <- noext.z.t.test.df %>%
-  mutate(full_df = 0) %>%
-  dplyr::select(comp, prop, n_sig_regions, full_df)
-z.t.test.df_reduced <- z.t.test.df %>%
-  mutate(full_df = 1) %>%
-  dplyr::select(comp, prop, n_sig_regions, full_df)
+noext.z.t.test.df2 <- noext.z.feat.t.test.df %>%
+  mutate(full_df = "reduced")
+z.feat.t.test.df2 <- z.feat.t.test.df %>%
+  mutate(full_df = "full")
 
-full_join(noext.z.t.test.df_reduced, z.t.test.df_reduced) %>%
+z.comp.df <- full_join(noext.z.t.test.df2, z.feat.t.test.df2)
+z.comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
 ggplot() +
-  geom_col(aes(x=prop, y=n_sig_regions, fill=as.factor(full_df)), position="dodge") +
-  facet_wrap(~comp) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  geom_bar(aes(x=prop, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_grid(comp ~ full_df)
+
+z.comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+ggplot() +
+  geom_bar(aes(x=comp, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap( ~ full_df)
 ```
+
 Test sex-biases in z errors w/in each feature
@@
 ```{r}
-noext_z.subj.sex.test.list <- readRDS(paste0(der_path, "/ukb_ratio/no.ext_featurewise_z_sex_bias_tests.RDS"))
+noext_z.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
 noext_z.subj.sex.test.df <- bind_rows(noext_z.subj.sex.test.list)

### 2024-02-23 paper_main_results.Rmd
drop reference site configs

@@
 }
+
+#drop refA sites, not using for these analyses
+combined_preds <- combined_preds %>%
+  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
 ```
@@
      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+
+#drop refA sites, not using for these analyses
+diffs.df <- diffs.df %>%
+  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
 ```
@@
      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
+
+#drop refA sites, not using for these analyses
+subj_sum.df <- subj_sum.df %>%
+  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
 ```
@@
 combined_var <- inner_join(combined_var, plot.info)
+
+#drop refA sites, not using for these analyses
+combined_var <- combined_var %>%
+  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
 ```
@@
 ```
-plot w ref sites
-```{r, eval=FALSE}
-#get means w/in feature 
-mean.cent.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss", "cf.lm_refA", "cf.gam_refA", "cf.gamlss_refA"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="cent_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  # mutate(pheno=gsub(".z", "", pheno)) %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-mean.cent.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  ggplot() +
-  geom_histogram(aes(x=cent_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_grid(dataset~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
-  ggtitle("Absolute errors in centiles of brain structure features across ComBat processing pipelines")
-```
+
 Quick comp of cf.gam and cf.gamlss
@@
   dplyr::filter(gam_minus_gamlss < 0)
-
-gam_better$pheno
 

### 2024-02-27 paper_main_results.Rmd
visulaizing magnitude of sex biases

@@
 ```{r}
+#get range
+sex_t_tests_in_feat.df %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  group_by(dataset) %>%
+  summarize(min_sex_diff = min(median_sex_diff, na.rm=TRUE),
+            max_sex_diff = max(median_sex_diff, na.rm=TRUE))
+
+#look by feature type
+sex_t_tests_in_feat.df %>%
+  # dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100),
+         pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+ggplot() +
+  geom_histogram(aes(x=median_sex_diff, color=dataset, fill=dataset),alpha=0.5, position="identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(pheno_cat~dataset, scales="free")
+
+```
+
+```{r}
 plt.df <- diffs.df %>%

### 2024-02-27 paper_prop_results.Rmd
visulaizing magnitude of sex biases

@@
   facet_wrap(~comp)
+
+cent.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+  mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+ggplot() +
+  geom_bar(aes(x=prop, fill=bigger_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_grid(pheno_cat~comp, scales="free_y")
 ```
@@
 ```{r}
-cent.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_cent_sex_bias_tests.RDS"))
-cent.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
+cent.feat.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_cent_sex_bias_tests.RDS"))
+cent.feat.sex.test.df <- bind_rows(cent.feat.sex.test.list)
 ```
@@
 ```{r}
-cent.subj.sex.test.df %>%
+cent.feat.sex.test.df %>%
   ggplot() +
@@
 
-```{r, eval=FALSE}
-cent.subj.sex.test.df %>%
+```{r}
+cent.feat.sex.test.df %>%
   mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>%
+                              TRUE ~ median_sex_diff*100),
+         pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
   ggplot() +
-  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
-  facet_wrap(~prop)
+  geom_histogram(aes(x=median_sex_diff, fill= pheno_cat, color=pheno_cat), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(prop~dataset, scales="free")
 
 #zoom in on combat-GAM vs combat-LS
-cent.subj.sex.test.df %>%
+cent.feat.sex.test.df %>%
   dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
@@
 
-cent.subj.sex.test.df %>%
+cent.feat.sex.test.df %>%
   group_by(dataset, prop) %>%
@@
 #trying to add stats test
-global_sex_diffs <- cent.subj.sex.test.df %>%
+global_sex_diffs <- cent.feat.sex.test.df %>%
   dplyr::filter(pheno %in% vol_list_global & dataset %in% c("cf.gam", "cf.gamlss")) %>%
@@
 ```{r, eval=FALSE}
-cent.subj.sex.test.df %>%
+cent.feat.sex.test.df %>%
   dplyr::filter(dataset == "cf.gamlss") %>%
@@
 
-cent.subj.sex.test.df2 <- cent.subj.sex.test.df %>%
+cent.feat.sex.test.df2 <- cent.feat.sex.test.df %>%
   dplyr::filter(prop == "prop-04") %>%
@@
     label = sub("_[^_]*_", "_", pheno)) #for plotting
-sex_t_tests_to_plot <- left_join(cent.subj.sex.test.df2, dk.parc, by="label")
+sex_t_tests_to_plot <- left_join(cent.feat.sex.test.df2, dk.parc, by="label")
 
@@
 noext_cent.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
-noext_cent.subj.sex.test.df <- bind_rows(noext_cent.subj.sex.test.list)
+noext_cent.feat.sex.test.df <- bind_rows(noext_cent.subj.sex.test.list)
 ```
@@
 ```{r}
-noext_cent.subj.sex.test.df %>%
+noext_cent.feat.sex.test.df %>%
   ggplot() +

### 2024-02-27 paper_replication_results.Rmd
visulaizing magnitude of sex biases

@@
 
+```{r}
+#get range
+cent.subj.sex.test.df %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  group_by(dataset) %>%
+  summarize(min_sex_diff = min(median_sex_diff, na.rm=TRUE),
+            max_sex_diff = max(median_sex_diff, na.rm=TRUE))
+
+#look by feature type
+cent.subj.sex.test.df %>%
+  # dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100),
+         pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+ggplot() +
+  geom_histogram(aes(x=median_sex_diff, color=perm, fill=perm),alpha=0.5, position="identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(pheno_cat~dataset, scales="free")
+
+```
+
 ### Without Extremes

### 2024-03-04 paper_main_results.Rmd
sex bias analyses, figs

@@
 
+look at # males vs females w extreme mean centiles
+
 ### W/o extremes
@@
 
+toy distributions showing harmonization - based on code from GPT3.5
+
+```{r}
+# Function to create data for normal distribution
+generate_data <- function(mean, sd, x_values) {
+  data.frame(x = x_values, y = dnorm(x_values, mean = mean, sd = sd))
+}
+
+# Set parameters for raw data
+mean_values <- c(0.1, 0.5, -1)  # Means for three curves even, M, F
+sd_values <- c(0.8, 1.3, 1)     # Standard deviations for three curves
+x_values <- seq(-4, 4, by = 0.1) # Adjust the range and interval as needed
+name_list <- c("A", "C", "B")
+
+# Generate data for three normal distributions
+raw_data <- mapply(generate_data, mean = mean_values, sd = sd_values, MoreArgs = list(x_values), SIMPLIFY = FALSE)
+names(raw_data) <- name_list
+
+# Combine data frames into one
+raw_data_combined <- do.call(rbind, Map(cbind, raw_data, Curve = names(raw_data)))
+
+# Set parameters for ComBat/ComBat-GAM
+mean_values <- c(0, 0.3, -0.3)  # Means for three curves even, M, F
+sd_values <- c(1, 1, 1)     # Standard deviations for three curves
+
+# Generate data for three normal distributions
+cb_data <- mapply(generate_data, mean = mean_values, sd = sd_values, MoreArgs = list(x_values), SIMPLIFY = FALSE)
+names(cb_data) <- name_list
+
+# Combine data frames into one
+cb_data_combined <- do.call(rbind, Map(cbind, cb_data, Curve = names(cb_data)))
+
+# Set parameters for ComBat/ComBat-GAM
+mean_values <- c(0, 0.3, -0.3)  # Means for three curves even, M, F
+sd_values <- c(1, 1.1, 0.9)     # Standard deviations for three curves
+
+# Generate data for three normal distributions
+ls_data <- mapply(generate_data, mean = mean_values, sd = sd_values, MoreArgs = list(x_values), SIMPLIFY = FALSE)
+names(ls_data) <- name_list
+
+# Combine data frames into one
+ls_data_combined <- do.call(rbind, Map(cbind, ls_data, Curve = names(ls_data)))
+
+# Plot using ggplot2
+ggplot(raw_data_combined, aes(x = x, y = y, color = Curve)) +
+  geom_line(size = 3) +
+  labs(title = "Raw Data",
+       x = "X-axis", y = "Density") +
+  scale_color_manual(values = c(A = "#9E52FF", B = "#5254FF", C = "#C87CFF")) +
+  theme_void()
+
+ggplot(cb_data_combined, aes(x = x, y = y, color = Curve)) +
+  geom_line(size = 3) +
+  labs(title = "ComBat/ComBat-GAM Data",
+       x = "X-axis", y = "Density") +
+  scale_color_manual(values = c(A = "#9E52FF", B = "#5254FF", C = "#C87CFF")) +
+  theme_void()
+
+ggplot(ls_data_combined, aes(x = x, y = y, color = Curve)) +
+  geom_line(size = 3) +
+  labs(title = "ComBatLS Data",
+       x = "X-axis", y = "Density") +
+  scale_color_manual(values = c(A = "#9E52FF", B = "#5254FF", C = "#C87CFF")) +
+  theme_void()
+```
 toy brain chart for figs, diagram portion done in canva

### 2024-03-04 paper_replication_results.Rmd
sex bias analyses, figs

--- 2024-02-19 paper_replication_results.Rmd
@@
       pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
+      TRUE ~ NA_character_))) %>%
+  separate_wider_delim(cols=comp, delim="_", names=c("group1", "group2"), cols_remove=FALSE) %>%
+  mutate(smaller_group=ifelse(bigger_group==group1, group2, group1))
 
 cent.feat.t.test.df %>%
-  group_by(comp, bigger_group) %>%
+  group_by(comp, smaller_group) %>%
   summarize(n_sig=sum(sig_fdr))
@@
 ```{r}
-cent.feat.t.test.df %>%
+comp_names <- c(cf.gam_cf = "ComBat-GAM v.
+ComBat w/o Covar", cf.gamlss_cf = "ComBatLS v.
+ComBat w/o Covar", cf.gamlss_cf.gam = "ComBatLS v.
+ComBat-GAM", cf.lm_cf.gam ="ComBat-GAM v.
+linear ComBat", cf.lm_cf.gamlss = "ComBatLS v.
+linear ComBat", cf.lm_cf = "linear ComBat v.
+ComBat w/o Covar")
+
+cent.test.comp.plt <- cent.feat.t.test.df %>%
   dplyr::filter(sig_fdr==TRUE) %>%
+  mutate(smaller_group=factor(smaller_group, levels=c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered=TRUE)) %>%
 ggplot() +
-  geom_bar(aes(x=perm, fill=bigger_group)) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~comp)
+  geom_bar(aes(x=perm, fill=smaller_group)) +
+  theme_linedraw()+
+  theme(axis.text.x = element_blank()) +
+  facet_grid(pheno_cat~comp, scales="free_y", labeller = labeller(comp = as_labeller(comp_names))) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "Better ComBat \nConfiguration", labels=dataset_names) +
+  labs(x="Sampling Permutation", y="Feature Count")
+ggsave(paste0(base_path, "/figs/perm_cent_test_barplt.jpg"), plot=cent.test.comp.plt, width=15.5, height=6, units="in")
 ```
@@
 
-ct.t.tests_to_plt %>%
+ct.brain.plt <- ct.t.tests_to_plt %>%
   ggplot() +
@@
         legend.position="top") +
-  scale_fill_manual(values = c(NS= "darkgray", cf.gam = "#3D75C2", cf.gamlss ="#CFB023FF"), name = "Better ComBat Method") +
+  scale_fill_manual(values = c(NS= "darkgray", cf.gam = "#3D75C2", cf.gamlss ="#CFB023FF"), name = "Better ComBat Method", labels=dataset_names) +
   facet_wrap(~perm, nrow=2)
+
+ct.t.tests_to_plt.sum <- ct.t.tests_to_plt %>%
+  group_by(hemi, region, label, bigger_group) %>%
+  summarise(n_worse = sum(sig_fdr)) %>%
+  pivot_wider(id_cols=c(hemi, label, region), names_from = bigger_group, values_from=n_worse) %>%
+  mutate(cf.gamlss=ifelse(is.na(cf.gamlss), 0, cf.gamlss)) %>%
+  mutate(cf.gamlss_minus_cf.gam_better=as.integer(cf.gam-cf.gamlss),
+         percent_cf.gamlss_better=(cf.gam/10)*100) %>%
+  ungroup()
+
+ct.brain.sum.plt <- ct.t.tests_to_plt.sum %>%
+  ggplot() +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+            aes(fill=percent_cf.gamlss_better))  +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  scale_fill_gradient(low="#605210", high= "#e6d16d",  name = "% Replications where ComBatLS is Better")
 ```
@@
   scale_fill_manual(values = c("lightblue", "red"))
+```
+
+```{r}
+#get range
+cent.subj.sex.test.df %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  group_by(dataset) %>%
+  summarize(min_sex_diff = min(median_sex_diff, na.rm=TRUE),
+            max_sex_diff = max(median_sex_diff, na.rm=TRUE))
+
+#look by feature type
+cent.subj.sex.test.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
+  mutate(dataset = factor(dataset, levels = c("cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    perm = as.factor(sub("perm-0", "", perm))) %>%
+ggplot() +
+  geom_histogram(aes(x=median_sex_diff, color=perm, fill=perm), alpha=0.5, position="identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(pheno_cat~dataset, scales="free", labeller = labeller(dataset = as_labeller(dataset_names))) +
+  theme_linedraw(base_size=12) +
+  labs(fill='Sampling \nPermutation', color='Sampling \nPermutation', x="Male - Female Median Centile Errors", y="Number of Features") 
+```
+
+Distribution of error medians w/in each pheno category
+```{r}
+cent.sex.test.med <- cent.subj.sex.test.df %>%
+   mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+  group_by(dataset, perm, pheno_cat) %>%
+  summarise(median_of_median_sex_diff = median(median_sex_diff))
+
+cent.sex.test.med %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot() +
+  geom_density(aes(x=median_of_median_sex_diff, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_wrap(~pheno_cat, scales="free")
+```
+
+plot subj. with averages in extremes - try w/ averages w/in pheno categories
+```{r}
+#recalc subject's true mean centiles
+raw_mean_cent <- perm_subj.df %>%
+  dplyr::filter(dataset=="cf") %>%
+  mutate(mean_centile = mean_centile - mean_cent_diff,
+         mean_z = mean_z - mean_z_diff) %>%
+  dplyr::select(!c(Source_File, contains("diff"))) %>%
+  mutate(dataset="raw")
+
+perm_subj.df2 <- full_join(perm_subj.df, raw_mean_cent)
+
+high_cent.sum <- perm_subj.df2 %>%
+  dplyr::filter(mean_centile > 0.75) %>%
+  group_by(dataset, perm) %>%
+  summarise(
+    male_ratio = sum(sex == "Male") / n(),
+    female_ratio = sum(sex == "Female") / n()
+  )
+
+high_cent.sum %>%
+  dplyr::filter(dataset != "raw") %>%
+  ggplot(aes(x=dataset, y=male_ratio)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir="center") +
+  geom_hline(yintercept= high_cent.sum[high_cent.sum$dataset == "raw" & high_cent.sum$perm=="perm-001",][["male_ratio"]], linetype="dashed")
+
+low_cent.sum <- perm_subj.df2 %>%
+  dplyr::filter(mean_centile < 0.25) %>%
+  group_by(dataset, perm) %>%
+  summarise(
+    male_ratio = sum(sex == "Male") / n(),
+    female_ratio = sum(sex == "Female") / n()
+  )
+
+low_cent.sum %>%
+  dplyr::filter(dataset != "raw") %>%
+  ggplot(aes(x=dataset, y=male_ratio)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir="center")  +
+  geom_hline(yintercept= low_cent.sum[low_cent.sum$dataset == "raw" & low_cent.sum$perm=="perm-001",][["male_ratio"]], linetype="dashed")
+
+perm_subj.df %>%
+  # dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot() +
+  geom_histogram(aes(x=mean_cent_diff, fill=sex, color=sex), position="identity", alpha=0.6, bins=60) +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_wrap(~dataset, scales="free")
+
+# wilcox.test(mean_cent_diff ~ sex, data=perm_subj.df[dataset=="cf.gamlss",])
+# wilcox.test(mean_cent_diff ~ sex, data=perm_subj.df[dataset=="cf.gam",])
+# wilcox.test(mean_cent_diff ~ sex, data=perm_subj.df[dataset=="cf.lm",])
+
+# perm_subj.df %>%
+#   mutate(centile_raw = mean_centile - mean_cent_diff)
+#   arrange(participant)
+```
+
+```{r}
+#one-sample tests against raw vals
+
+one_samp_wilcox <- function(cf_method, df){
+  df2 <- df %>%
+    dplyr::filter(dataset == cf_method)
+  raw_val <- df[df$dataset == "raw" & df$perm=="perm-001",][["male_ratio"]]
+  wilcox.test(df2$male_ratio, mu=raw_val)
+}
+
+cf_test_list <- unique(perm_subj.df$dataset)
+
+high_test <- lapply(cf_test_list, one_samp_wilcox, df=high_cent.sum)
+names(high_test) <- cf_test_list
+
+low_test <- lapply(cf_test_list, one_samp_wilcox, df=low_cent.sum)
+names(low_test) <- cf_test_list
+
+high_test
+low_test
 ```
@@
 ```
+
+## Supplemental Fig
+```{r}
+sfig <- ggarrange(cent.test.comp.plt, ct.brain.sum.plt,
+                  ncol=1, nrow=2, heights = c(2, 1))
+sfig
+ggsave(paste0(base_path, "/figs/combat_abs_errs_supp_plt.jpg"), plot=sfig, width=11, height=7, units="in")
+
+```

### 2024-03-04 cubic_scripts/run_permutation_stats.sh
request more mem

@@ -70,7 +70,7 @@ do
 	echo "singularity run --cleanenv $img Rscript --save $r_script "perm" $n_perm $csv_path" > $bash_script
 
 	#qsub bash script
-	qsub -N perm-${n_perm}_sum -o $bash_dir/perm-${n_perm}_sum_out.txt -e $bash_dir/perm-${n_perm}_sum_err.txt -l h_vmem=64G,s_vmem=64G $bash_script
+	qsub -N perm-${n_perm}_sum -o $bash_dir/perm-${n_perm}_sum_out.txt -e $bash_dir/perm-${n_perm}_sum_err.txt -l h_vmem=120G,s_vmem=120G $bash_script
 done
 #######################################################
 ## CHECK FOR OUTPUTS
@@ -105,7 +105,7 @@ touch $bash_script
 
 echo "singularity run --cleanenv $img Rscript --save $r_script $csv_path 'full' 'perm' " > $bash_script
 ## qsub bash script
-qsub -N perm_cent_tests -o $bash_dir/cent_test_out.txt -e $bash_dir/cent_test_err.txt -l h_vmem=60.5G,s_vmem=60.0G $bash_script
+qsub -N perm_cent_tests -o $bash_dir/cent_test_out.txt -e $bash_dir/cent_test_err.txt -l h_vmem=120G,s_vmem=120G $bash_script
 
 # sex-bias test of subj-means
 r_script=$r_base/ratio_subject-wise_results.R
@@ -114,7 +114,7 @@ touch $bash_script
 
 echo "singularity run --cleanenv $img Rscript --save $r_script $csv_path 'perm'" > $bash_script
 ## qsub bash script
-qsub -N perm_sub-wide_sex_tests -o $bash_dir/sub-wide_sex_test_out.txt -e $bash_dir/sub-wide_sex_test_err.txt -l h_vmem=60.5G,s_vmem=60.0G $bash_script
+qsub -N perm_sub-wide_sex_tests -o $bash_dir/sub-wide_sex_test_out.txt -e $bash_dir/sub-wide_sex_test_err.txt -l h_vmem=120G,s_vmem=120G $bash_script
 
 # featurewise sex bias tests
 ## based on `qsub_sex_bias_test.sh`
@@ -124,7 +124,7 @@ touch $bash_script
 	
 echo "singularity run --cleanenv $img Rscript --save $r_script $csv_path 'full' 'perm'" > $bash_script
 ## qsub bash script
[…]

### 2024-03-06 eda/lifespan_eda.Rmd
getting numbers, figs

@@
 
-unique(cn.df.imp$site)
-table(cn.df.imp$site, cn.df.imp$study)
-sum(is.na(cn.df.imp$site))
+# unique(cn.df.imp$site)
+# table(cn.df.imp$site, cn.df.imp$study)
+# sum(is.na(cn.df.imp$site))
+```
+
+Demographics
+
+```{r}
+#sample size, sex
+length(unique(cn.df.imp$participant))
+nrow(cn.df.imp)
+
+table(cn.df.imp$sex)
+
+#studies, sites
+length(unique(cn.df.imp$study))
+length(unique(cn.df.imp$site))
+
+#age
+min(cn.df.imp$age_days)/365.25
 max(cn.df.imp$age_days)/365.25
 mean(cn.df.imp$age_days)/365.25
+
+abcd <- cn.df.imp %>%
+  dplyr::filter(study=="ABCD")
+length(unique(abcd$site))
 ```

### 2024-03-06 paper_prop_results.Rmd
getting numbers, figs

@@
 
+Look at differences in M:F for subjects w very high or very low mean centiles
+```{r}
+feat_exts <- ratio.df %>%
+  dplyr::select(-contains(c("diff", ".z"))) %>%
+  mutate(across(pheno_list, ~ case_when(. < 0.25 ~ "low",
+                                        . > 0.75 ~ "high", 
+                                        TRUE ~ "mid")))
+
+feat_exts_long <- feat_exts %>%
+  pivot_longer(cols=c(pheno_list), names_to="pheno", values_to="centile_cat") %>%
+   mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)))
+
+feat_exts_long %>%
+  ggplot() +
+  geom_col(aes(x=centile_cat, y=pheno_cat, fill=sex, color=sex)) +
+  facet_grid(dataset ~ prop)
+```
+
 visualize actual (not ABSOLUTE) errors for global phenos by sex

### 2024-03-06 paper_replication_results.Rmd
getting numbers, figs

@@
 data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm"
+base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar"
 
@@
 
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+
 cent.test.comp.plt <- cent.feat.t.test.df %>%
@@
   mutate(cf.gamlss_minus_cf.gam_better=as.integer(cf.gam-cf.gamlss),
-         percent_cf.gamlss_better=(cf.gam/10)*100) %>%
+         percent_cf.gamlss_better=(cf.gam/100)*100) %>%
   ungroup()
@@
   guides(fill=guide_legend(title="Sampling \nPermutation")) + 
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 100x") +
+  theme(legend.position="none")
 cent.diffs.v.plot
@@
 ggplot() +
-  geom_histogram(aes(x=median_sex_diff, color=perm, fill=perm), alpha=0.5, position="identity") +
+  geom_histogram(aes(x=median_sex_diff, color=perm, fill=perm), alpha=0.3, position="identity") +
   geom_vline(xintercept = 0, linetype="dashed") +
@@
   theme_linedraw(base_size=12) +
-  labs(fill='Sampling \nPermutation', color='Sampling \nPermutation', x="Male - Female Median Centile Errors", y="Number of Features") 
+  theme(legend.position = "none") +
+  labs(x="Male - Female Median Centile Errors", y="Number of Features") 
 ```
@@
 cent.sex.test.med %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  # dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
   ggplot() +
@@
   facet_wrap(~pheno_cat, scales="free")
+
+t.test.cent.sex.test.med <- cent.sex.test.med %>%
+  group_by(dataset, pheno_cat) %>%
+  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value) %>%
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = nrow(cent.sex.test.med)),
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+t.test.cent.sex.test.med %>%
+  ggplot() +
+  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")
 ```
 
-plot subj. with averages in extremes - try w/ averages w/in pheno categories
+plot subj. with averages in extremes - **try also using averages w/in pheno categories**
 ```{r}
@@
   geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir="center") +
+  # geom_dotplot(binaxis = "y", stackdir="center") +
   geom_hline(yintercept= high_cent.sum[high_cent.sum$dataset == "raw" & high_cent.sum$perm=="perm-001",][["male_ratio"]], linetype="dashed")
@@
   dplyr::filter(dataset != "raw") %>%
+  ggplot(aes(x=dataset, y=female_ratio)) +
+  geom_boxplot() +
+  # geom_dotplot(binaxis = "y", stackdir="center")  +
+  geom_hline(yintercept= low_cent.sum[low_cent.sum$dataset == "raw" & low_cent.sum$perm=="perm-001",][["female_ratio"]], linetype="dashed") +
+  labs(title="subjects with low average centiles")
+
+ext_cent.sum <- perm_subj.df2 %>%
+  dplyr::filter(mean_centile < 0.25 | mean_centile > 0.75) %>%
+  group_by(dataset, perm) %>%
+  summarise(
+    male_ratio = sum(sex == "Male") / n(),
+    female_ratio = sum(sex == "Female") / n()
+  )
+
+ext_cent.sum %>%
+  dplyr::filter(dataset != "raw") %>%
   ggplot(aes(x=dataset, y=male_ratio)) +
   geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir="center")  +
-  geom_hline(yintercept= low_cent.sum[low_cent.sum$dataset == "raw" & low_cent.sum$perm=="perm-001",][["male_ratio"]], linetype="dashed")
+  # geom_dotplot(binaxis = "y", stackdir="center")  +
+  geom_hline(yintercept= ext_cent.sum[ext_cent.sum$dataset == "raw" & ext_cent.sum$perm=="perm-001",][["male_ratio"]], linetype="dashed")
 
@@
   facet_wrap(~dataset, scales="free")
+```
 
-# wilcox.test(mean_cent_diff ~ sex, data=perm_subj.df[dataset=="cf.gamlss",])
-# wilcox.test(mean_cent_diff ~ sex, data=perm_subj.df[dataset=="cf.gam",])
-# wilcox.test(mean_cent_diff ~ sex, data=perm_subj.df[dataset=="cf.lm",])
+too many males in high centiles, too few males in low centiles (combat-GAM trends opposite to this pattern)
 
-# perm_subj.df %>%
-#   mutate(centile_raw = mean_centile - mean_cent_diff)
-#   arrange(participant)
-```
+overall - too few men relative to women in extreme centiles
 
@@
 names(high_test) <- cf_test_list
+high_test
 
@@
 names(low_test) <- cf_test_list
+low_test
 
-high_test
-low_test
+ext_test <- lapply(cf_test_list, one_samp_wilcox, df=ext_cent.sum)
+names(ext_test) <- cf_test_list
+ext_test
 ```

### 2024-03-07 eda/ukb_curation-eda.Rmd

@@
 fwrite(ukb_df_age.filt, file=paste0(base_path,"data/ukb_CN_data_agefilt.csv"))
+
+nrow(ukb_df_totals) - nrow(ukb_df_age.filt)
 ```
@@
 ```{r}
+nrow(ukb_df_age.filt)
 #percent female

### 2024-03-07 paper_lbcc_results.Rmd
getting numbers, figs

@@
 
+ComBatGAM mean centile ~ age
+
+```{r}
+gam_centile_site.mean.df <- gam_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(site) %>%
+  summarize(mean_gam_cent=mean(centile))
+
+gam_centile_site.mean.df2 <- left_join(gam_centile_site.mean.df, delta_sum.df_merged)
+
+gam.age.test <- lm(mean_gam_cent ~ site_mean_age, data=gam_centile_site.mean.df2)
+summary(gam.age.test)
+gam.age.test2 <- lm(mean_gam_cent ~ sex_prop_Female + site_n + site_mean_age, data=gam_centile_site.mean.df2)
+summary(gam.age.test2)
+```
+
+NS
+
+ComBatLS mean centile ~ age
+
+```{r}
+gamlss_centile_site.mean.df <- gamlss_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(site) %>%
+  summarize(mean_gamlss_cent=mean(centile))
+
+gamlss_centile_site.mean.df2 <- left_join(gamlss_centile_site.mean.df, delta_sum.df_merged)
+
+gamlss.age.test <- lm(mean_gamlss_cent ~ site_mean_age, data=gamlss_centile_site.mean.df2)
+summary(gamlss.age.test)
+gamlss.age.test2 <- lm(mean_gamlss_cent ~ sex_prop_Female + site_n + site_mean_age, data=gamlss_centile_site.mean.df2)
+summary(gamlss.age.test2)
+```
+
+NS
+
+Site-level mean centiles are not associated with study age in either ComBat-GAM or ComBatLS
+
 ##Age
@@
   ggplot() +
-  geom_histogram(aes(x=age_days))
+  geom_histogram(aes(x=age_days)) +
+  labs(title="subject distribution across age")
+
+delta_sum.df_merged %>%
+  ggplot() +
+  geom_histogram(aes(x=site_mean_age_yrs)) +
+  labs(title="site distribution across age")
 ```
@@
 ##Prop Female
+
+```{r}
+ggplot(delta_sum.df_merged) +
+  geom_histogram(aes(x=sex_prop_Female))
+```
+
+ComBatGAM mean centile ~ % female
+
+```{r}
+gam_centile_site.mean.df <- gam_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(site) %>%
+  summarize(mean_gam_cent=mean(centile))
+
+gam_centile_site.mean.df2 <- left_join(gam_centile_site.mean.df, delta_sum.df_merged)
+
+gam.prop.f.test <- lm(mean_gam_cent ~ sex_prop_Female + site_n + site_mean_age, data=gam_centile_site.mean.df2)
+summary(gam.prop.f.test)
+sjPlot::plot_model(gam.prop.f.test, type = "pred", terms = "sex_prop_Female")
+
+ggplot(gam_centile_site.mean.df2, aes(x=sex_prop_Female, y=mean_gam_cent)) +
+  geom_point() +
+  geom_smooth(method="lm")
+```
+
+ComBatLS mean centile ~ % female
+
+```{r}
+gamlss_centile_site.mean.df <- gamlss_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(site) %>%
+  summarize(mean_gamlss_cent=mean(centile))
+
+gamlss_centile_site.mean.df2 <- left_join(gamlss_centile_site.mean.df, delta_sum.df_merged)
+
+gamlss.prop.f.test <- lm(mean_gamlss_cent ~ sex_prop_Female + site_n + site_mean_age, data=gamlss_centile_site.mean.df2)
+summary(gamlss.prop.f.test)
+sjPlot::plot_model(gamlss.prop.f.test, type = "pred", terms = "sex_prop_Female")
+
+ggplot(gamlss_centile_site.mean.df2, aes(x=sex_prop_Female, y=mean_gamlss_cent)) +
+  geom_point() +
+  geom_smooth(method="lm")
+```
+
+```{r}
+gamlss_centile_site.mean.df3 <- gamlss_centile_site.mean.df2 %>%
+  mutate(combat = "gamlss") %>%
+  rename(mean_cent = mean_gamlss_cent)
+gam_centile_site.mean.df3 <- gam_centile_site.mean.df2 %>%
+  mutate(combat = "gam") %>%
+  rename(mean_cent = mean_gam_cent)
+
+site_lvl_mean_cents <- full_join(gamlss_centile_site.mean.df3, gam_centile_site.mean.df3)
+
+site_lvl_mean_cents %>%
+  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=combat, fill=combat)) +
+  geom_point(alpha=0.4) +
+  geom_smooth(method="lm") +
+  ylab("within-site mean centile")
+```
+
+```{r}
+gamlss_centile_df.mean %>%
+  ggplot() + 
+  geom_violin(aes(x=sex, y=mean_cent))
+gamlss_centile_df.mean$combat <- "gamlss"
+
+gam_centile_df.mean %>%
+  ggplot() + 
+  geom_violin(aes(x=sex, y=mean_cent))
+gam_centile_df.mean$combat <- "gam"
+
+subj.mean.cents <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)
+
+subj.mean.cents %>%
+  ggplot() + 
+  geom_boxplot(aes(x=sex, y=mean_cent, fill=combat))
+
+#add sex prop
+site.info <- delta_sum.df_merged %>%
+  dplyr::select(site, sex_prop_Female, site_n)
+
+subj.mean.cents2 <- left_join(subj.mean.cents, site.info)
+subj.mean.cents2 %>%
+  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=sex)) +
+  geom_point(aes(size=site_n)) +
+  geom_smooth(method="lm")+
+  facet_wrap(~combat)
+
+subj.mean.cents2 %>%
+  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=combat, fill=combat)) +
+  # geom_point(alpha=0.4) +
+  geom_smooth(method="lm") +
+  ylab("within-subject mean centile")
+```
 
@@
 sjPlot::plot_model(full.delta.sex.fit, type = "pred", terms = "sex_prop_Female") 
+```
+
+```{r}
+delta_sum.by.sex.df %>%
+  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
+  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
+  geom_line(aes(group=site))
+```
+
+```{r}
+delta_sum.df_merged %>%
+  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
+  geom_count(aes(size=site_n)) +
+  geom_smooth(method="lm")
+delta_sum.df_merged %>%
+  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta)) +
+  geom_count(aes(size=site_n)) +
+  geom_smooth(method="gam")
+
+sex_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
+sex_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]
+
+sex_effects.plt <- delta_sum.df_merged %>%
+  mutate(site_mean_delta=site_mean_delta*100) %>%
+  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
+  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
+  geom_smooth(method="lm",
+              color="#24693DFF") +
+  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
+  theme_linedraw(base_size = 11) +
+  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores") +
+  theme(axis.title.y=element_blank()) +
+  annotate("text", x=0.25, y=-0.3, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
+  guides(size = "none")
+
+sex_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
+sex_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]
+
+sex_effects.plt.null <- delta_sum.df_merged %>%
+  ggplot(aes(x=sex_prop_Female, y=(site_mean_abs.delta*100))) +
+  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
+  geom_smooth(method="lm", color="#24693DFF") +
+  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
+  theme_linedraw(base_size = 11) +
+  labs(size="site N") +
+    annotate("text", x=0.25, y=4, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
+  theme(axis.title=element_blank()) +
+  guides(size = "none")
+
+#add prediction to df
+delta_sum.df_merged$delta_pred <- predict(full.delta.sex.fit)
+
+full.delta.sex.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$x,
+                                    predicted=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$predicted,
+                                    conf.low=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.low,
+                                    conf.high=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.high)
+full.delta.sex.fit.df <- full.delta.sex.fit.df %>%
+  mutate(predicted = predicted*100,
+         conf.low=conf.low*100,
+         conf.high=conf.high*100)
+
+marg.sex_effects.plt <- ggplot() +
+  geom_count(data=delta_sum.df_merged, aes(x=sex_prop_Female, y=(delta_pred*100), size=site_n, color=site_mean_age_yrs)) +
+  geom_line(aes(x=full.delta.sex.fit.df$x, y = full.delta.sex.fit.df$predicted), size = 0.7, color="#24693DFF") +
+  geom_ribbon(aes(x=full.delta.sex.fit.df$x,y= NULL, ymin= full.delta.sex.fit.df$conf.low, ymax= full.delta.sex.fit.df$conf.high), alpha=0.1, fill="#24693DFF") +
+  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
+  theme_linedraw(base_size = 11) +
+  labs(size="site N", x = "Proportion Female", y = "Predicted Mean Difference in Centile Scores")
+
+delta_sum.by.sex.df %>%
+  ggplot(aes(x=sex, y=(mean_delta*100))) +
+  geom_boxplot()
+```
+
+```{r}
+# age*%female
+summary(lm(site_mean_abs.delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
+summary(lm(site_mean_delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
+```
+
+```{r}
+cent_deltas_df %>%
+  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
+  geom_point(alpha=0.4) +
+  geom_smooth(method="lm")
+
+cent_deltas_df %>%
+  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
+  geom_histogram(aes(y=after_stat(c(
+      count[group==1]/sum(count[group==1]),
+      count[group==2]/sum(count[group==2])
+    )*100)), alpha=0.4, position="identity", bins=60) +
+  ylab("Percentage")
+
+cent_deltas_df %>%
+  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm")
+
+cent_deltas_df %>%
+  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
+  geom_point(alpha=0.4) +
+  geom_smooth(method="lm")
+```
+
+#Exploratory: Subject-Level Mean Results
+
+Get subject-level average deltas
+
+```{r}
+subj.avg.delta.df <- cent_deltas_df_long %>%
+  mutate(abs.delta = abs(delta)) %>%
+  group_by(participant) %>%
+  summarise(subj_mean_abs.delta=mean(abs.delta),
+            subj_mean_delta=mean(delta),
+            site=unique(site),
+            study=unique(study),
+            sex=unique(sex))
+
+#add back some site-level info
+subj.avg.delta.sum.df <- left_join(subj.avg.delta.df, delta_sum.df_merged)
+
+subj.avg.delta.sum.df %>%
+  ggplot() +
+  geom_boxplot(aes(x=sex, y=subj_mean_delta))
+```
+
+##Age
+
+ComBat-GAM centile ~ site age
+
+```{r}
+#add back some site-level info
+gam_centile_df.mean2 <- left_join(gam_centile_df.mean, delta_sum.df_merged)
+
+summary(lm(mean_cent ~ site_mean_age, data=gam_centile_df.mean2))
+```
+
+ComBatLS centile ~ site age
+
+```{r}
+#add back some site-level info
+gamlss_centile_df.mean2 <- left_join(gamlss_centile_df.mean, delta_sum.df_merged)
+
+summary(lm(mean_cent ~ site_mean_age, data=gamlss_centile_df.mean2))
+```
+
+Neither combat method produces centiles that are assoc. with site mean age
+
+delta ~ age
+
+```{r}
+# age
+summary(lm(subj_mean_abs.delta ~ site_mean_age, data=subj.avg.delta.sum.df))
+summary(lm(subj_mean_delta ~ site_mean_age, data=subj.avg.delta.sum.df))
+
+ggplot(subj.avg.delta.sum.df, aes(x=site_mean_age_yrs, y=subj_mean_abs.delta)) +
+  geom_point() +
+  geom_smooth(method="lm")
+ggplot(subj.avg.delta.sum.df, aes(x=site_mean_age_yrs, y=subj_mean_delta)) +
+  geom_point() +
+  geom_smooth(method="lm")
+```
+
+the difference between combat and combat-ls centiles is
+
+```{r}
+#forward selection for subj_mean_abs.delta
+summary(lm(subj_mean_abs.delta ~ site_mean_age_yrs + site_n, data=subj.avg.delta.sum.df))
+summary(lm(subj_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=subj.avg.delta.sum.df)) #small effect but remains significant/grows stronger
+
+#forward selection for subj_mean_abs.delta
+summary(lm(subj_mean_delta ~ site_mean_age_yrs + site_n, data=subj.avg.delta.sum.df))
+summary(lm(subj_mean_delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=subj.avg.delta.sum.df))
+```
+
+```{r}
+delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=site_mean_delta)) +
+  geom_count(aes(size=site_n)) +
+  geom_smooth(method="lm")
+delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
+  geom_count(aes(size=site_n)) +
+  geom_smooth(method="lm")
+
+age_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
+age_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]
+
+age_effects.plt <- delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
+  geom_count(aes(size=site_n, color=sex_prop_Female)) +
+  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
+  geom_smooth(method="lm", color="#2A5783FF") +
+  theme_linedraw(base_size = 11) +
+  annotate("text", x=60, y=4, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
+  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
+  theme(axis.title.x=element_b)
+
+age_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
+age_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]
+
+age_effects.plt.null <- delta_sum.df_merged %>%
+  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_delta*100))) +
+  geom_count(aes(size=site_n, color=sex_prop_Female)) +
+  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
+  geom_smooth(method="lm", color="#2A5783FF") +
+  theme_linedraw(base_size = 11) +
+  annotate("text", x=60, y=-0.3, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
+  labs(size="site N", x = "Mean Age (Years)", y = "Mean Difference in Centile Scores")
+
+#add prediction to df
+delta_sum.df_merged$abs.delta_pred <- predict(full.delta.age.fit)
+
+full.delta.age.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$x,
+                                    predicted=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$predicted,
+                                    conf.low=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.low,
+                                    conf.high=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.high)
+full.delta.age.fit.df <- full.delta.age.fit.df %>%
+  mutate(predicted = predicted*100,
+         conf.low=conf.low*100,
+         conf.high=conf.high*100)
+
+marg.age_effects.plt <- ggplot() +
+  geom_count(data=delta_sum.df_merged, aes(x=site_mean_age_yrs, y=(abs.delta_pred*100), size=site_n, color=sex_prop_Female)) +
+  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
+  geom_line(aes(x=full.delta.age.fit.df$x, y = full.delta.age.fit.df$predicted), size = 0.7, color="#2A5783FF") +
+  geom_ribbon(aes(x=full.delta.age.fit.df$x,y= NULL, ymin= full.delta.age.fit.df$conf.low, ymax= full.delta.age.fit.df$conf.high), alpha=0.1, fill="#2A5783FF") +
+  theme_linedraw(base_size = 11) +
+  labs(size="site N", x = "Mean Age (Years)", y = "Predicted Mean Absolute Difference in Centile Scores")
+```
+
+are studies w older subjects changing more because there are fewer of them?
+check for association between study size and age
+```{r}
+summary(lm(site_mean_age_yrs ~ site_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
+```
+
+```{r}
+cent_deltas_df %>%
+  ggplot() +
+  geom_histogram(aes(x=age_days))
+```
+
+##Prop Female
+
+ComBatGAM mean centile ~ % female
+
+```{r}
+gam_centile_site.mean.df <- gam_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(site) %>%
+  summarize(mean_gam_cent=mean(centile))
+
+gam_centile_site.mean.df2 <- left_join(gam_centile_site.mean.df, delta_sum.df_merged)
+
+gam.prop.f.test <- lm(mean_gam_cent ~ sex_prop_Female + site_n + site_mean_age, data=gam_centile_site.mean.df2)
+summary(gam.prop.f.test)
+sjPlot::plot_model(gam.prop.f.test, type = "pred", terms = "sex_prop_Female") 
+```
+
+ComBatLS mean centile ~ % female
+
+```{r}
+gamlss_centile_site.mean.df <- gamlss_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(site) %>%
+  summarize(mean_gamlss_cent=mean(centile))
+
+gamlss_centile_site.mean.df2 <- left_join(gamlss_centile_site.mean.df, delta_sum.df_merged)
+
+gamlss.prop.f.test <- lm(mean_gamlss_cent ~ sex_prop_Female + site_n + site_mean_age, data=gamlss_centile_site.mean.df2)
+summary(gamlss.prop.f.test)
+sjPlot::plot_model(gamlss.prop.f.test, type = "pred", terms = "sex_prop_Female") 
+```
+
+```{r}
+gamlss_centile_df.mean %>%
+  ggplot() + 
+  geom_violin(aes(x=sex, y=mean_cent))
+gamlss_centile_df.mean$combat <- "gamlss"
+
+gam_centile_df.mean %>%
+  ggplot() + 
+  geom_violin(aes(x=sex, y=mean_cent))
+gam_centile_df.mean$combat <- "gam"
+
+subj.mean.cents <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)
+
+subj.mean.cents %>%
+  ggplot() + 
+  geom_boxplot(aes(x=sex, y=mean_cent, fill=combat))
+
+#add sex prop
+site.info <- delta_sum.df_merged %>%
+  dplyr::select(site, sex_prop_Female, site_n)
+
+subj.mean.cents2 <- left_join(subj.mean.cents, site.info)
+subj.mean.cents2 %>%
+  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=sex)) +
+  geom_point() +
+  geom_smooth(method="lm")+
+  facet_wrap(~combat)
+```
+
+delta ~ % female
+
+```{r}
+ggplot(subj.avg.delta.sum.df, aes(x=site_mean_age_yrs, y=subj_mean_abs.delta)) +
+  geom_point() +
+  geom_smooth(method="lm")
+
+ggplot(subj.avg.delta.sum.df, aes(x=site_mean_age_yrs, y=subj_mean_delta)) +
+  geom_point() +
+  geom_smooth(method="lm")
+
+# %female
+summary(lm(subj_mean_abs.delta ~ sex_prop_Female, data=subj.avg.delta.sum.df))
+summary(lm(subj_mean_delta ~ sex_prop_Female, data=subj.avg.delta.sum.df))
+summary(lm(subj_mean_abs.delta ~ sex_prop_Female + site_n, data=subj.avg.delta.sum.df)) #add site size covariate
 ```

### 2024-03-07 paper_replication_results.Rmd

@@
 ## Load Data
-
-Read in centile/z-score errors (pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*_diffs.csv' .`)
-
-```{r, eval=FALSE}
-raw_files <- list.files(path = data_path, pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-    data <- fread(file)
-
-    data <- data %>%
-      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("diffs_perm.list", df_list)
-}
-length(diffs_perm.list)
-diffs_perm.df <- bind_rows(diffs_perm.list)
-```
 
@@
   ggplot() +
-  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=perm), scale="count") +
+  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, color=perm), scale="count") +
   labs(y= "Mean Magnitude Centile Error", x = "ComBat Configurations") +
   scale_x_discrete(labels=c(dataset_names)) +
-  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 100x") +
+  # guides(fill=guide_legend(title="Sampling \nPermutation")) + 
+  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations replicated 100x") +
   theme(legend.position="none")
 cent.diffs.v.plot
-# ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
+ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot, width=11, height=4.5, units="in")
 ```
@@
 
-plot subj. with averages in extremes - **try also using averages w/in pheno categories**
+plot subj. with averages in extremes 
 ```{r}
@@
   mutate(mean_centile = mean_centile - mean_cent_diff,
-         mean_z = mean_z - mean_z_diff) %>%
+        VolGlob_mean_centile = VolGlob_mean_centile - VolGlob_mean_cent_diff,
+        VolReg_mean_centile = VolReg_mean_centile - VolReg_mean_cent_diff,
+        CT_mean_centile = CT_mean_centile - CT_mean_cent_diff,
+        SA_mean_centile = SA_mean_centile - SA_mean_cent_diff) %>%
   dplyr::select(!c(Source_File, contains("diff"))) %>%
@@
 
+# raw_mean_cent %>%
+#   dplyr::filter(perm=="perm-001") %>%
+#   dplyr::filter(VolReg_mean_centile > .85) %>% nrow()
+
 perm_subj.df2 <- full_join(perm_subj.df, raw_mean_cent)
 
-high_cent.sum <- perm_subj.df2 %>%
-  dplyr::filter(mean_centile > 0.75) %>%
-  group_by(dataset, perm) %>%
+perm_subj.df2_long <- perm_subj.df2 %>%
+  rename(allpheno_mean_centile = mean_centile) %>%
+  pivot_longer(cols=contains("mean_centile"), names_to="pheno_cat", values_to="mean_centile") %>%
+  mutate(pheno_cat = gsub("_mean_centile", "", pheno_cat))
+
+high_cent.sum <- perm_subj.df2_long %>%
+  dplyr::filter(mean_centile > 0.80) %>%
+  group_by(dataset, perm, pheno_cat) %>%
   summarise(
@@
 high_cent.sum %>%
-  dplyr::filter(dataset != "raw") %>%
-  ggplot(aes(x=dataset, y=male_ratio)) +
+  #dplyr::filter(dataset != "raw") %>%
+  ggplot(aes(x=dataset, y=female_ratio)) +
   geom_boxplot() +
   # geom_dotplot(binaxis = "y", stackdir="center") +
-  geom_hline(yintercept= high_cent.sum[high_cent.sum$dataset == "raw" & high_cent.sum$perm=="perm-001",][["male_ratio"]], linetype="dashed")
+  facet_wrap(~pheno_cat, scales="free_y") +
+  #geom_hline(yintercept=high_cent.sum[high_cent.sum$dataset == "raw" & high_cent.sum$perm=="perm-001",][["female_ratio"]], linetype="dashed")+
+  labs(title="subjects with high average centiles")
 
-low_cent.sum <- perm_subj.df2 %>%
-  dplyr::filter(mean_centile < 0.25) %>%
-  group_by(dataset, perm) %>%
+low_cent.sum <- perm_subj.df2_long %>%
+  dplyr::filter(mean_centile < 0.20) %>%
+  group_by(dataset, perm, pheno_cat) %>%
   summarise(
@@
 low_cent.sum %>%
-  dplyr::filter(dataset != "raw") %>%
+  #dplyr::filter(dataset != "raw") %>%
   ggplot(aes(x=dataset, y=female_ratio)) +
   geom_boxplot() +
+  facet_wrap(~pheno_cat, scales="free_y") +
   # geom_dotplot(binaxis = "y", stackdir="center")  +
-  geom_hline(yintercept= low_cent.sum[low_cent.sum$dataset == "raw" & low_cent.sum$perm=="perm-001",][["female_ratio"]], linetype="dashed") +
+  #geom_hline(yintercept= low_cent.sum[low_cent.sum$dataset == "raw" & low_cent.sum$perm=="perm-001",][["female_ratio"]], linetype="dashed") +
   labs(title="subjects with low average centiles")
 
-ext_cent.sum <- perm_subj.df2 %>%
-  dplyr::filter(mean_centile < 0.25 | mean_centile > 0.75) %>%
-  group_by(dataset, perm) %>%
+ext_cent.sum <- perm_subj.df2_long %>%
+  dplyr::filter(mean_centile < 0.20 | mean_centile > 0.80) %>%
+  group_by(dataset, perm, pheno_cat) %>%
   summarise(
@@
 ext_cent.sum %>%
-  dplyr::filter(dataset != "raw") %>%
-  ggplot(aes(x=dataset, y=male_ratio)) +
+  #dplyr::filter(dataset != "raw") %>%
+  ggplot(aes(x=dataset, y=female_ratio)) +
   geom_boxplot() +
   # geom_dotplot(binaxis = "y", stackdir="center")  +
-  geom_hline(yintercept= ext_cent.sum[ext_cent.sum$dataset == "raw" & ext_cent.sum$perm=="perm-001",][["male_ratio"]], linetype="dashed")
+  #geom_hline(yintercept= ext_cent.sum[ext_cent.sum$dataset == "raw" & ext_cent.sum$perm=="perm-001",][["female_ratio"]], linetype="dashed")+
+  facet_wrap(~pheno_cat, scales="free_y") +
+  labs(title="subjects with extreme average centiles")
 
@@
 
-too many males in high centiles, too few males in low centiles (combat-GAM trends opposite to this pattern)
+linear ComBat -> too few women with high centiles (avg. across phenos, pattern varies), too many women with low centiles; ~ balances out overall (except regional volume)
 
-overall - too few men relative to women in extreme centiles
+ComBat-GAM -> too many women with high centiles (small but pretty consistent pattern), too many women with low centiles (patter varies a bit across phenos); OVERALL TOO MANY WOMEN W EXTREME CENTILES
+
+ComBatLS -> pretty close to raw!
 
@@
 ext_test
+```
+
+by site
+```{r}
+perm_subj.df2.by.site <- perm_subj.df2 %>%
+  group_by(perm, dataset, sim.site) %>%
+  summarise(across(contains("mean"), list(site = mean)))
+
+perm_subj.df2.by.site %>%
+  ggplot() +
+  geom_boxplot(aes(x=sim.site, y=mean_cent_abs.diff_site, fill=dataset)) +
+  facet_wrap(~ dataset, scales="free")
+
+perm_subj.df2.by.site %>%
+    mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
+  ggplot() +
+  geom_boxplot(aes(x=sim.site, y=mean_cent_abs.diff_site, fill=dataset)) +
+  facet_wrap(~ dataset, scales="free")
+
+perm_subj.df2.by.site %>%
+  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
+  ggplot() +
+  geom_boxplot(aes(x=sim.site, y=mean_cent_diff_site, fill=dataset)) +
+  facet_wrap(~ dataset, scales="free")
+
+perm_subj.df2.by.site %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
+  ggplot() +
+  geom_boxplot(aes(x=sim.site, y=mean_cent_diff_site, fill=dataset))
+
+perm_subj.df2.by.site %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
+  ggplot() +
+  geom_boxplot(aes(x=sim.site, y=mean_centile_site, fill=dataset))
+
+perm_subj.df2.by.site %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  pivot_wider(id_cols=c(perm, sim.site), names_from=dataset, values_from=mean_centile_site, names_prefix="site_mean_cent_") %>%
+  mutate(delta_mean = site_mean_cent_cf.gam - site_mean_cent_cf.gamlss) %>%
+  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
+  ggplot() +
+  geom_boxplot(aes(x=sim.site, y=delta_mean))
 ```

### 2024-03-08 paper_lbcc_results.Rmd
cleaning up

@@
             site_max_abs.delta=max(abs.delta),
-            site_n=n())
+            site_n=n()) %>%
+  mutate(pheno_cat="all")
+
+# #get w/in cat stats
+# delta_summary.cat.df <- cent_deltas_df_long %>%
+#   mutate(abs.delta = abs(delta),
+#          pheno_cat = case_when(
+#     pheno %in% vol_list_global ~ "Global Volume",
+#     pheno %in% vol_list_regions ~ "Regional Volume",
+#     pheno %in% sa_list ~ "Regional SA",
+#     pheno %in% ct_list ~ "Regional CT",
+#     TRUE ~ NA_character_)) %>%
+#   group_by(study, site, pheno_cat) %>%
+#   summarise(site_mean_abs.delta=mean(abs.delta),
+#             site_mean_delta=mean(delta),
+#             site_mean_age=mean(age_days),
+#             site_mean_age_yrs=mean(age_days)/365.25,
+#             site_max_abs.delta=max(abs.delta),
+#             site_n=n())
+# 
+# delta_summary.df <- full_join(delta_summary.df, delta_summary.cat.df)
 
@@
   geom_bar(aes(y=sig_fdr))
+
+df_gam_feat_aov.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
 ```
@@
 df_gamlss_feat_aov.study %>%
-  dplyr::filter(sig_fdr==TRUE)
+  dplyr::filter(sig_fdr==TRUE) %>%
+  nrow()
 ```
@@
 ```{r}
-gam_centile_site.mean.df <- gam_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(site) %>%
-  summarize(mean_gam_cent=mean(centile))
+gam_centile_df.clean_long <- gam_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")
+
+#overall means
+gam_centile_site.mean.df <- gam_centile_df.clean_long %>%
+  group_by(study, site) %>%
+  summarize(mean_gam_cent=mean(centile)) %>%
+  mutate(pheno_cat = "all")
 
@@
 
-gam.age.test <- lm(mean_gam_cent ~ site_mean_age, data=gam_centile_site.mean.df2)
-summary(gam.age.test)
-gam.age.test2 <- lm(mean_gam_cent ~ sex_prop_Female + site_n + site_mean_age, data=gam_centile_site.mean.df2)
-summary(gam.age.test2)
+summary(lm(mean_gam_cent ~ site_mean_age, data=gam_centile_site.mean.df2))
+summary(lm(mean_gam_cent ~ site_mean_age + sex_prop_Female + site_n, data=gam_centile_site.mean.df2))
 ```
@@
 ComBatLS mean centile ~ age
-
-```{r}
-gamlss_centile_site.mean.df <- gamlss_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(site) %>%
-  summarize(mean_gamlss_cent=mean(centile))
+```{r}
+gamlss_centile_df.clean_long <- gamlss_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")
+
+#overall means
+gamlss_centile_site.mean.df <- gamlss_centile_df.clean_long %>%
+  group_by(study, site) %>%
+  summarize(mean_gamlss_cent=mean(centile)) %>%
+  mutate(pheno_cat = "all")
 
@@
 
-gamlss.age.test <- lm(mean_gamlss_cent ~ site_mean_age, data=gamlss_centile_site.mean.df2)
-summary(gamlss.age.test)
-gamlss.age.test2 <- lm(mean_gamlss_cent ~ sex_prop_Female + site_n + site_mean_age, data=gamlss_centile_site.mean.df2)
-summary(gamlss.age.test2)
+summary(lm(mean_gamlss_cent ~ site_mean_age, data=gamlss_centile_site.mean.df2))
+summary(lm(mean_gamlss_cent ~ site_mean_age + sex_prop_Female + site_n, data=gamlss_centile_site.mean.df2))
 ```

### 2024-03-08 paper_lbcc_results.Rmd
cleaning up exploratory

@@
 sjPlot::plot_model(full.delta.sex.fit, type = "pred", terms = "sex_prop_Female") 
-```
-
-```{r}
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
-  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=site))
-```
-
-```{r}
-delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
-  geom_count(aes(size=site_n)) +
-  geom_smooth(method="lm")
-delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta)) +
-  geom_count(aes(size=site_n)) +
-  geom_smooth(method="gam")
-
-sex_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
-sex_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]
-
-sex_effects.plt <- delta_sum.df_merged %>%
-  mutate(site_mean_delta=site_mean_delta*100) %>%
-  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
-  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
-  geom_smooth(method="lm",
-              color="#24693DFF") +
-  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores") +
-  theme(axis.title.y=element_blank()) +
-  annotate("text", x=0.25, y=-0.3, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
-  guides(size = "none")
-
-sex_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
-sex_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]
-
-sex_effects.plt.null <- delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=(site_mean_abs.delta*100))) +
-  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
-  geom_smooth(method="lm", color="#24693DFF") +
-  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N") +
-    annotate("text", x=0.25, y=4, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
-  theme(axis.title=element_blank()) +
-  guides(size = "none")
-
-#add prediction to df
-delta_sum.df_merged$delta_pred <- predict(full.delta.sex.fit)
-
-full.delta.sex.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$x,
-                                    predicted=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$predicted,
-                                    conf.low=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.low,
-                                    conf.high=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.high)
-full.delta.sex.fit.df <- full.delta.sex.fit.df %>%
-  mutate(predicted = predicted*100,
-         conf.low=conf.low*100,
-         conf.high=conf.high*100)
-
-marg.sex_effects.plt <- ggplot() +
-  geom_count(data=delta_sum.df_merged, aes(x=sex_prop_Female, y=(delta_pred*100), size=site_n, color=site_mean_age_yrs)) +
-  geom_line(aes(x=full.delta.sex.fit.df$x, y = full.delta.sex.fit.df$predicted), size = 0.7, color="#24693DFF") +
-  geom_ribbon(aes(x=full.delta.sex.fit.df$x,y= NULL, ymin= full.delta.sex.fit.df$conf.low, ymax= full.delta.sex.fit.df$conf.high), alpha=0.1, fill="#24693DFF") +
-  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N", x = "Proportion Female", y = "Predicted Mean Difference in Centile Scores")
-
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=(mean_delta*100))) +
-  geom_boxplot()
-```
-
-```{r}
-# age*%female
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
-```
-
-```{r}
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
-  geom_point(alpha=0.4) +
-  geom_smooth(method="lm")
-
-cent_deltas_df %>%
-  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
-  geom_histogram(aes(y=after_stat(c(
-      count[group==1]/sum(count[group==1]),
-      count[group==2]/sum(count[group==2])
-    )*100)), alpha=0.4, position="identity", bins=60) +
-  ylab("Percentage")
-
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm")
-
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
-  geom_point(alpha=0.4) +
-  geom_smooth(method="lm")
-```
-
-#Exploratory: Subject-Level Mean Results
-
-Get subject-level average deltas
-
-```{r}
-subj.avg.delta.df <- cent_deltas_df_long %>%
-  mutate(abs.delta = abs(delta)) %>%
-  group_by(participant) %>%
-  summarise(subj_mean_abs.delta=mean(abs.delta),
-            subj_mean_delta=mean(delta),
-            site=unique(site),
-            study=unique(study),
-            sex=unique(sex))
-
-#add back some site-level info
-subj.avg.delta.sum.df <- left_join(subj.avg.delta.df, delta_sum.df_merged)
-
-subj.avg.delta.sum.df %>%
-  ggplot() +
-  geom_boxplot(aes(x=sex, y=subj_mean_delta))
-```
-
-##Age
-
-ComBat-GAM centile ~ site age
-
-```{r}
-#add back some site-level info
-gam_centile_df.mean2 <- left_join(gam_centile_df.mean, delta_sum.df_merged)
-
-summary(lm(mean_cent ~ site_mean_age, data=gam_centile_df.mean2))
-```
-
-ComBatLS centile ~ site age
-
-```{r}
-#add back some site-level info
-gamlss_centile_df.mean2 <- left_join(gamlss_centile_df.mean, delta_sum.df_merged)
-
-summary(lm(mean_cent ~ site_mean_age, data=gamlss_centile_df.mean2))
-```
-
-Neither combat method produces centiles that are assoc. with site mean age
-
-delta ~ age
-
-```{r}
-# age
-summary(lm(subj_mean_abs.delta ~ site_mean_age, data=subj.avg.delta.sum.df))
-summary(lm(subj_mean_delta ~ site_mean_age, data=subj.avg.delta.sum.df))
-
-ggplot(subj.avg.delta.sum.df, aes(x=site_mean_age_yrs, y=subj_mean_abs.delta)) +
-  geom_point() +
-  geom_smooth(method="lm")
-ggplot(subj.avg.delta.sum.df, aes(x=site_mean_age_yrs, y=subj_mean_delta)) +
-  geom_point() +
-  geom_smooth(method="lm")
-```
-
-the difference between combat and combat-ls centiles is
-
-```{r}
-#forward selection for subj_mean_abs.delta
-summary(lm(subj_mean_abs.delta ~ site_mean_age_yrs + site_n, data=subj.avg.delta.sum.df))
-summary(lm(subj_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=subj.avg.delta.sum.df)) #small effect but remains significant/grows stronger
-
-#forward selection for subj_mean_abs.delta
-summary(lm(subj_mean_delta ~ site_mean_age_yrs + site_n, data=subj.avg.delta.sum.df))
-summary(lm(subj_mean_delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=subj.avg.delta.sum.df))
-```
-
-```{r}
-delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=site_mean_delta)) +
-  geom_count(aes(size=site_n)) +
-  geom_smooth(method="lm")
-delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
-  geom_count(aes(size=site_n)) +
-  geom_smooth(method="lm")
-
-age_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
-age_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]
-
-age_effects.plt <- delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
-  geom_count(aes(size=site_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
-  geom_smooth(method="lm", color="#2A5783FF") +
-  theme_linedraw(base_size = 11) +
-  annotate("text", x=60, y=4, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
-  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
-  theme(axis.title.x=element_b)
-
-age_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
-age_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]
-
-age_effects.plt.null <- delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_delta*100))) +
-  geom_count(aes(size=site_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
-  geom_smooth(method="lm", color="#2A5783FF") +
-  theme_linedraw(base_size = 11) +
-  annotate("text", x=60, y=-0.3, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
-  labs(size="site N", x = "Mean Age (Years)", y = "Mean Difference in Centile Scores")
-
-#add prediction to df
-delta_sum.df_merged$abs.delta_pred <- predict(full.delta.age.fit)
-
-full.delta.age.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$x,
-                                    predicted=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$predicted,
-                                    conf.low=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.low,
-                                    conf.high=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.high)
-full.delta.age.fit.df <- full.delta.age.fit.df %>%
-  mutate(predicted = predicted*100,
-         conf.low=conf.low*100,
-         conf.high=conf.high*100)
-
-marg.age_effects.plt <- ggplot() +
-  geom_count(data=delta_sum.df_merged, aes(x=site_mean_age_yrs, y=(abs.delta_pred*100), size=site_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
-  geom_line(aes(x=full.delta.age.fit.df$x, y = full.delta.age.fit.df$predicted), size = 0.7, color="#2A5783FF") +
-  geom_ribbon(aes(x=full.delta.age.fit.df$x,y= NULL, ymin= full.delta.age.fit.df$conf.low, ymax= full.delta.age.fit.df$conf.high), alpha=0.1, fill="#2A5783FF") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N", x = "Mean Age (Years)", y = "Predicted Mean Absolute Difference in Centile Scores")
-```
-
-are studies w older subjects changing more because there are fewer of them?
-check for association between study size and age
-```{r}
-summary(lm(site_mean_age_yrs ~ site_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
-```
-
-```{r}
-cent_deltas_df %>%
-  ggplot() +
-  geom_histogram(aes(x=age_days))
-```
-
-##Prop Female
-
-ComBatGAM mean centile ~ % female
-
-```{r}
-gam_centile_site.mean.df <- gam_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(site) %>%
-  summarize(mean_gam_cent=mean(centile))
-
-gam_centile_site.mean.df2 <- left_join(gam_centile_site.mean.df, delta_sum.df_merged)
-
-gam.prop.f.test <- lm(mean_gam_cent ~ sex_prop_Female + site_n + site_mean_age, data=gam_centile_site.mean.df2)
-summary(gam.prop.f.test)
-sjPlot::plot_model(gam.prop.f.test, type = "pred", terms = "sex_prop_Female") 
-```
-
-ComBatLS mean centile ~ % female
-
-```{r}
-gamlss_centile_site.mean.df <- gamlss_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(site) %>%
-  summarize(mean_gamlss_cent=mean(centile))
-
-gamlss_centile_site.mean.df2 <- left_join(gamlss_centile_site.mean.df, delta_sum.df_merged)
-
-gamlss.prop.f.test <- lm(mean_gamlss_cent ~ sex_prop_Female + site_n + site_mean_age, data=gamlss_centile_site.mean.df2)
-summary(gamlss.prop.f.test)
-sjPlot::plot_model(gamlss.prop.f.test, type = "pred", terms = "sex_prop_Female") 
-```
-
-```{r}
-gamlss_centile_df.mean %>%
-  ggplot() + 
-  geom_violin(aes(x=sex, y=mean_cent))
-gamlss_centile_df.mean$combat <- "gamlss"
-
-gam_centile_df.mean %>%
-  ggplot() + 
-  geom_violin(aes(x=sex, y=mean_cent))
-gam_centile_df.mean$combat <- "gam"
-
-subj.mean.cents <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)
-
-subj.mean.cents %>%
-  ggplot() + 
-  geom_boxplot(aes(x=sex, y=mean_cent, fill=combat))
-
-#add sex prop
-site.info <- delta_sum.df_merged %>%
-  dplyr::select(site, sex_prop_Female, site_n)
-
-subj.mean.cents2 <- left_join(subj.mean.cents, site.info)
-subj.mean.cents2 %>%
-  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=sex)) +
-  geom_point() +
-  geom_smooth(method="lm")+
-  facet_wrap(~combat)
-```
-
-delta ~ % female
-
-```{r}
-ggplot(subj.avg.delta.sum.df, aes(x=site_mean_age_yrs, y=subj_mean_abs.delta)) +
-  geom_point() +
-  geom_smooth(method="lm")
-
-ggplot(subj.avg.delta.sum.df, aes(x=site_mean_age_yrs, y=subj_mean_delta)) +
-  geom_point() +
-  geom_smooth(method="lm")
-
-# %female
-summary(lm(subj_mean_abs.delta ~ sex_prop_Female, data=subj.avg.delta.sum.df))
-summary(lm(subj_mean_delta ~ sex_prop_Female, data=subj.avg.delta.sum.df))
-summary(lm(subj_mean_abs.delta ~ sex_prop_Female + site_n, data=subj.avg.delta.sum.df)) #add site size covariate
 ```

### 2024-03-09 eda/lifespan_eda.Rmd
methods & results

@@
 #   geom_point(aes(x=study, y=(study_euler_median+2*study_euler_madconst1)), color="red")
+
+left_join(cn.df.eulerfilt, study_mad) %>%
+  mutate(scan_pass = ifelse(surfholes <= euler_threshold, "yes", "no")) %>%
+  group_by(study) %>%
+  summarise(fail_percent = (sum(scan_pass == "no") / n())*100) %>%
+  summarise(max_fail = max(fail_percent),
+            min_fail=min(fail_percent),
+            mean_fail=mean(fail_percent))
+
 ```

### 2024-03-09 paper_lbcc_results.Rmd
methods & results

@@
 # mean effects
-ggarrange(marg.age_effects.plt,
-marg.sex_effects.plt,
+ggarrange(age_effects.plt,
+sex_effects.plt,
 nrow=2, ncol=1,
@@
 
-fig5_annotate <- annotate_figure(fig5, top = text_grob("Figure 5.
-Differences in centile scores of lifespan consortium data when harmonized with ComBat-GAM and ComBatLS", color = "black", hjust = 0, x = 0))
-fig5_annotate
-ggsave("figs/figure_5.jpeg", fig5_annotate, width=12, height=10, units="in")
+ggsave("figs/figure_5.jpeg", fig5, width=12, height=10, units="in")
 ```
-
-```{r}
-# study demographics
-fig5.optB <- ggarrange(
-ggarrange(sample_boxplot, sample_barplot,
-          full_boxplot, full_barplot,
-          nrow=2, ncol=2, 
-          # align = "hv", 
-          common.legend = TRUE,
-          legend = "bottom",
-          widths=c(2, 1.3),
-          heights=c(14, 1)),
-# mean effects
-ggarrange(
-ggarrange(age_effects.plt, age_effects.plt.null,
-          nrow=2, ncol=1, common.legend=TRUE, legend="bottom",
-          align = "hv"),
-ggarrange(sex_effects.plt.null, sex_effects.plt,
-          nrow=2, ncol=1, common.legend=TRUE, legend="bottom",
-          align = "hv"),
-# align="hv",
-# labels=c("B", "C")),
-common.legend=FALSE,
-nrow=1, ncol=2),
-labels=c("A", "B"),
-widths=c(1, 2.2),
-nrow=1, ncol=2
-)
-
-fig5.optB_annotate <- annotate_figure(fig5.optB, top = text_grob("Figure 5.
-Differences in centile scores of lifespan consortium data when harmonized with ComBat-GAM and ComBatLS", color = "black", hjust = 0, x = 0))
-fig5.optB_annotate
-ggsave("figs/figure_5_option2.jpeg", fig5.optB_annotate, width=13, height=10, units="in")
-```

### 2024-03-09 paper_main_results.Rmd
methods & results

@@
   paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-6, 6)) #
-sex.var.plot
+saveRDS(sex.var.plot, file="figs/sex_var_plot.RDS")
 ```
@@
 # sex.var.plot.global
+
+#save RDS
+saveRDS(sex.var.plot.global, file="figs/sex_var_plot_global.RDS")
 ```
@@
 ```{r}
-dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c('ukb-CN-simsites' = "True Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 
@@
 ```{r}
+dataset_names <- c('ukb-CN-simsites' = "True Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 ven.df %>%
@@
 
-fig_2_annotate <- annotate_figure(fig_2, top = text_grob("Figure 2. Choice of ComBat configuration can induce errors in brain feature centile scores",
-               color = "black", hjust = 0, x = 0))
-fig_2_annotate
-ggsave("figs/figure_2.jpeg", fig_2_annotate, width=11, height=7, units="in")
-```
-
-## FIGURE 3
-
-```{r}
-blank <- ggplot() + theme_void()
-
-fig_3 <- ggarrange(
-  ggarrange(blank, 
-            annotate_figure(
-              ggarrange(sex.var.plot.global, sex.var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.25),
-  hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "right"
-  )),
-  blank,
-  ncol=3, nrow=1,
-  widths=c(.5, 1, .5)),
-  big_violin_plt,
-  ncol=1, nrow=2, 
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  # align = "hv",
-  heights=c(1, 1.8),
-  common.legend = FALSE,
-  labels=c("A", "B")
-  )
-# fig_2
-
-fig_3_annotate <- annotate_figure(fig_3, top = text_grob("Figure 3. Sex-biases in centile score errors across ComBat configurations", color = "black", hjust = 0, x = 0))
-fig_3_annotate
-ggsave("figs/figure_3.jpeg", fig_3_annotate, width=11.5, height=10, units="in")
-```
+fig_2
+
+ggsave("figs/figure_2.jpeg", fig_2, width=11, height=7, units="in")
+```

### 2024-03-09 paper_prop_results.Rmd
methods & results

--- 2024-02-27 paper_prop_results.Rmd
@@
 ```{r}
-dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c('ukb-CN-simsites' = "Raw Data",  cf = "ComBat w/o cov", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
@@
   theme_linedraw() +
-  theme(axis.ticks.x = element_blank(), axis.text.x=element_blank(), panel.spacing.x = unit(.1, "lines")) +
+  theme(axis.ticks.x = element_blank(), axis.text.x=element_blank(), panel.spacing.x = unit(.1, "lines"), legend.position = "bottom") +
   xlab("Simulated Site") +
@@
   labs(y="Absolute Centile Error", x = "% Females in Imbalanced Site") +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
   scale_x_discrete(labels=percent_f) +
@@
   labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Site") +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
   scale_x_discrete(labels=percent_f) +
@@
 
-visualize magnitude of significant sex-biases
+visualize magnitude of sex-biases
 
 ```{r}
+cent.feat.sex.test.df <- cent.feat.sex.test.df %>%
+  mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)))
+
 cent.feat.sex.test.df %>%
   mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100),
-         pheno_cat = factor(case_when(
+                              TRUE ~ median_sex_diff*100)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= pheno_cat, color=pheno_cat), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(prop~dataset, scales="free")
+
+t.test.cent.sex.diff <- cent.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  group_by(dataset, pheno_cat, prop) %>%
+  summarise(t.test.p = tidy(t.test(median_sex_diff, mu=0))$p.value,
+            max_val = max(median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = (3*4*11)), #not testing global vols, too few data points
+                  sig_fdr_tf = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE),
+         sig_fdr = case_when(
+           # p.val_fdr < 0.001 ~ "***",
+           #                     p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+           #                     p.val_fdr >= 0.01 & 
+                               p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ ""),
+         y_pos = ifelse(max_val < 0, ((max_val*100)*0.5), (max_val*100)+0.8))
+
+cortical_cent_sex_diff.plt <- cent.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "cf") %>%
+  mutate(median_sex_diff = median_sex_diff*100) %>%
+  ggplot(aes(x=prop, y=median_sex_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+    facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names)) + 
+  geom_text(aes(label = sig_fdr, y = y_pos),
+              data = t.test.cent.sex.diff[which(t.test.cent.sex.diff$dataset != "cf"),]) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Centile Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+
+cent.feat.sex.test.df %>%
+  mutate(median_sex_diff = median_sex_diff*100) %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  ggplot() +
+  geom_dotplot(aes(x=prop, y=median_sex_diff, color=dataset, fill=dataset), binaxis = "y", stackdir="center") +
+    facet_grid(pheno_cat~dataset, scales="free_y", labeller=labeller(dataset=dataset_names)) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Centile Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+
+#zoom in on combat-GAM vs combat-LS
+cent.feat.sex.test.df %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  #mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              #TRUE ~ median_sex_diff*100)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_wrap(~prop, labeller = as_labeller(percent_f), scales="free") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+cent.feat.sex.test.df %>%
+  group_by(dataset, prop) %>%
+  summarize(min_diff = min(median_sex_diff)*100, max_diff=max(median_sex_diff)*100)
+```
+
+Look at differences in M:F for subjects w very high or very low centiles
+```{r}
+feat_exts_long <- ratio.df %>%
+  dplyr::select(-contains(c("diff", ".z"))) %>%
+  pivot_longer(cols=c(pheno_list), names_to="pheno", values_to="centile_cat") %>%
+   mutate(pheno_cat = as.factor(case_when(
     pheno %in% vol_list_global ~ "Global Volume",
@@
     TRUE ~ NA_character_))) %>%
+  mutate(centile_cat = case_when(. < 0.2 ~ "low",
+                                        . > 0.8 ~ "high", 
+                                        TRUE ~ "mid"))
+feat_exts_long %>%
+  dplyr::filter(centile_cat == "low") %>%
   ggplot() +
-  geom_histogram(aes(x=median_sex_diff, fill= pheno_cat, color=pheno_cat), alpha=0.4, position = "identity") +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  facet_grid(prop~dataset, scales="free")
+  geom_bar(aes(x=pheno_cat, fill=sex, color=sex)) +
+  facet_grid(prop ~ dataset)
 
-#zoom in on combat-GAM vs combat-LS
-cent.feat.sex.test.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>%
+feat_exts_long %>%
+  dplyr::filter(centile_cat == "high") %>%
   ggplot() +
-  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  facet_wrap(~prop, labeller = as_labeller(percent_f)) +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-cent.feat.sex.test.df %>%
-  group_by(dataset, prop) %>%
-  summarize(min_diff = min(median_sex_diff)*100, max_diff=max(median_sex_diff)*100)
+  geom_bar(aes(x=pheno_cat, fill=sex, color=sex)) +
+  facet_grid(prop ~ dataset)
 ```
@@
 ```{r}
-fig_4 <- ggarrange(ggarrange(sampling.plt, prop.global_cent_errs_plot, 
+fig_4 <- ggarrange(ggarrange(sampling.plt, cortical_cent_sex_diff.plt, 
                              nrow = 2, labels = c("A", "C"),
                              heights=c(1,2)), #nest 1st col
-          sex_err_ratio.plt,
+          prop.global_cent_errs_plot,
           ncol = 2,
-          widths=c(1.1, 1),
+          widths=c(1, 1.3),
           labels = c(NA, "B")                                    
@@
 
-fig_4_annotate <- annotate_figure(fig_4, top = text_grob("Figure 4. Varying M:F ratios across site simulations", color = "black", hjust = 0, x = 0))
-fig_4_annotate
-ggsave("figs/figure_4.jpeg", fig_4_annotate, width=33, height=20, units="cm")
+fig_4
+ggsave("figs/figure_4.jpeg", fig_4, width=33, height=20, units="cm")
 ```

### 2024-03-09 paper_replication_results.Rmd
methods & results

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata)
 #Sys.setenv("MC_CORES"=10L)
@@
 
-cent.sex.test.med %>%
-  # dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  ggplot() +
-  geom_density(aes(x=median_of_median_sex_diff, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  facet_wrap(~pheno_cat, scales="free")
-
 t.test.cent.sex.test.med <- cent.sex.test.med %>%
   group_by(dataset, pheno_cat) %>%
-  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value) %>%
-  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = nrow(cent.sex.test.med)),
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
+  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
+            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
+            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(cent.sex.test.med$dataset))*length(unique(cent.sex.test.med$pheno_cat))),
+                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
+                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ "")) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")
+
 t.test.cent.sex.test.med %>%
@@
   geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")
+
+sex_bias_density_plt <- cent.sex.test.med %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot() +
+  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
+  facet_wrap(~pheno_cat) +
+  theme_linedraw() +
+  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med$dataset == "cf.gam", 10.2, 5.5)),
+              data = t.test.cent.sex.test.med) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom")+
+  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
+
+cent.sex.test.med %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
+  ggplot() +
+  geom_boxplot(aes(y=median_of_median_sex_diff*100, x=dataset, color=dataset)) +
+  geom_hline(yintercept = 0, linetype="dashed", color="#3F3F3F") +
+  facet_wrap(~pheno_cat, scales="free") +
+  theme_linedraw() +
+  geom_text(aes(label = sig_fdr, x= x_pos*100, y=10.2),
+              data = t.test.cent.sex.test.med) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom")+
+  labs(x="Medians of Sex-Differences in Centile Errors Within Brain Feature Types", y="Density Across 100 Replications")
 ```
@@
 
-# raw_mean_cent %>%
-#   dplyr::filter(perm=="perm-001") %>%
-#   dplyr::filter(VolReg_mean_centile > .85) %>% nrow()
+raw_mean_cent %>%
+  dplyr::filter(perm=="perm-001") %>%
+  dplyr::filter(SA_mean_centile < .2) %>% 
+  nrow()
 
@@
   pivot_longer(cols=contains("mean_centile"), names_to="pheno_cat", values_to="mean_centile") %>%
-  mutate(pheno_cat = gsub("_mean_centile", "", pheno_cat))
+  mutate(pheno_cat = gsub("_mean_centile", "", pheno_cat)) %>%
+  mutate(pheno_cat = case_when(pheno_cat == "VolGlob" ~ "Global Volume",
+                               pheno_cat == "VolReg" ~ "Regional Volume",
+                               pheno_cat == "CT" ~ "Regional CT",
+                               pheno_cat == "SA" ~ "Regional SA"))
 
 high_cent.sum <- perm_subj.df2_long %>%
+  dplyr::filter(pheno_cat != "allpheno") %>%
   dplyr::filter(mean_centile > 0.80) %>%
@@
 
-high_cent.sum %>%
-  #dplyr::filter(dataset != "raw") %>%
-  ggplot(aes(x=dataset, y=female_ratio)) +
-  geom_boxplot() +
-  # geom_dotplot(binaxis = "y", stackdir="center") +
-  facet_wrap(~pheno_cat, scales="free_y") +
-  #geom_hline(yintercept=high_cent.sum[high_cent.sum$dataset == "raw" & high_cent.sum$perm=="perm-001",][["female_ratio"]], linetype="dashed")+
-  labs(title="subjects with high average centiles")
-
 low_cent.sum <- perm_subj.df2_long %>%
+  dplyr::filter(pheno_cat != "allpheno") %>%
   dplyr::filter(mean_centile < 0.20) %>%
@@
 
-low_cent.sum %>%
-  #dplyr::filter(dataset != "raw") %>%
-  ggplot(aes(x=dataset, y=female_ratio)) +
-  geom_boxplot() +
-  facet_wrap(~pheno_cat, scales="free_y") +
-  # geom_dotplot(binaxis = "y", stackdir="center")  +
-  #geom_hline(yintercept= low_cent.sum[low_cent.sum$dataset == "raw" & low_cent.sum$perm=="perm-001",][["female_ratio"]], linetype="dashed") +
-  labs(title="subjects with low average centiles")
-
 ext_cent.sum <- perm_subj.df2_long %>%
+  dplyr::filter(pheno_cat != "allpheno") %>%
   dplyr::filter(mean_centile < 0.20 | mean_centile > 0.80) %>%
@@
   )
-
-ext_cent.sum %>%
-  #dplyr::filter(dataset != "raw") %>%
-  ggplot(aes(x=dataset, y=female_ratio)) +
-  geom_boxplot() +
-  # geom_dotplot(binaxis = "y", stackdir="center")  +
-  #geom_hline(yintercept= ext_cent.sum[ext_cent.sum$dataset == "raw" & ext_cent.sum$perm=="perm-001",][["female_ratio"]], linetype="dashed")+
-  facet_wrap(~pheno_cat, scales="free_y") +
-  labs(title="subjects with extreme average centiles")
-
-perm_subj.df %>%
-  # dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  ggplot() +
-  geom_histogram(aes(x=mean_cent_diff, fill=sex, color=sex), position="identity", alpha=0.6, bins=60) +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  facet_wrap(~dataset, scales="free")
 ```
 
-linear ComBat -> too few women with high centiles (avg. across phenos, pattern varies), too many women with low centiles; ~ balances out overall (except regional volume)
+```{r}
+#get raw vals for horizontal lines
+high_cent.sum_raw <- high_cent.sum %>%
+  dplyr::filter(dataset == "raw" & perm=="perm-001")
 
-ComBat-GAM -> too many women with high centiles (small but pretty consistent pattern), too many women with low centiles (patter varies a bit across phenos); OVERALL TOO MANY WOMEN W EXTREME CENTILES
+low_cent.sum_raw <- low_cent.sum %>%
+  dplyr::filter(dataset == "raw" & perm=="perm-001")
+
+ext_cent.sum_raw <- ext_cent.sum %>%
+  dplyr::filter(dataset == "raw" & perm=="perm-001")
+
+#test significance
+get_t_tests <- function(df, raw_df){
+  #df <- df_list[[cat]]
+  cat <- unique(df$pheno_cat)
+  #print(cat)
+  mu <- raw_df$female_ratio[which(raw_df$pheno_cat == cat)]
+  #print(mu)
+  out <- df %>%
+    dplyr::filter(dataset != "raw") %>%
+    group_by(dataset) %>%
+    summarise(p_val = t.test(female_ratio, mu=mu)$p.val,
+              pheno_cat = cat,
+              max_val = max(female_ratio),
+              mu=mu) %>%
+    mutate(p.val_fdr = p.adjust(p_val, method="fdr", n = (4*4*3)), #adjusting for 4 methods * 4 pheno classes * 3 centile categories
+           sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
+                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ "NS"))
+  return(out)
+}
+
+high_t_test.df <- lapply(split(high_cent.sum, high_cent.sum$pheno_cat), get_t_tests, raw_df=high_cent.sum_raw) %>% bind_rows()
+low_t_test.df <- lapply(split(low_cent.sum, low_cent.sum$pheno_cat), get_t_tests, raw_df=low_cent.sum_raw) %>% bind_rows()
+ext_t_test.df <- lapply(split(ext_cent.sum, ext_cent.sum$pheno_cat), get_t_tests, raw_df=ext_cent.sum_raw) %>% bind_rows()
+
+#merge dfs
+all_cent_sum.df <- gdata::combine(high_cent.sum, low_cent.sum, ext_cent.sum) %>%
+  mutate(source = gsub("_cent.sum", "", source))
+all_cent.sum_raw <- all_cent_sum.df %>%
+  dplyr::filter(dataset == "raw" & perm=="perm-001")
+
+all_t_test.df <- gdata::combine(high_t_test.df, low_t_test.df, ext_t_test.df) %>%
+  mutate(source = gsub("_t_test.df", "", source))
+
+#scale by raw value
+all_cent_sum.df_scaled <- all_cent_sum.df %>%
+  mutate(dataset = factor(dataset, levels=c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered=TRUE)) %>%
+  arrange(dataset) %>%
+  group_by(source, pheno_cat) %>%
+  mutate(female_ratio_sc = female_ratio/first(female_ratio),
+         female_ratio_diff = female_ratio - first(female_ratio))
+
+#plots!
+all_cent_sum.df %>%
+  dplyr::filter(dataset != "raw") %>%
+  ggplot(aes(x=dataset, y=female_ratio)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="center") +
+  facet_grid(pheno_cat~source, scales="free_y") +
+  geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
+  geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+              data = all_t_test.df) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"),  name = "ComBat Configuration", labels=dataset_names) + 
+  theme_linedraw()
+
+cent_ext_list <- c(ext = ">80% and <20%", high = ">80%", low = "<20%")
+
+all_cent_sum_boxplt <- all_cent_sum.df %>%
+  dplyr::filter(dataset != "raw" & dataset != "cf") %>%
+  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=female_ratio)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="center") +
+  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
+  geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
+  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  #             data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  theme_linedraw() +
+  scale_x_discrete(labels=dataset_names) +
+  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Proportion Female", x = "Dataset")
+
+all_cent_sum.df_scaled %>%
+  dplyr::filter(dataset != "raw" & dataset != "cf") %>%
+  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=female_ratio_sc)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="center") +
+  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
+  geom_hline(yintercept=1, linetype="dashed")+
+  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  #             data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  theme_linedraw() +
+  scale_x_discrete(labels=dataset_names) +
+  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Female Bias", x = "Dataset")
+
+all_cent_sum_boxplt.diff <- all_cent_sum.df_scaled %>%
+  dplyr::filter(dataset != "raw" & dataset != "cf") %>%
+  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=female_ratio_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="center") +
+  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
+  geom_hline(yintercept=0, linetype="dashed")+
+  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  #             data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  theme_linedraw() +
+  scale_x_discrete(labels=dataset_names) +
+  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Female Bias", x = "Dataset")
+
+all_cent_sum_boxplt_wtest <- all_cent_sum.df %>%
+  dplyr::filter(dataset != "raw") %>%
+  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=female_ratio)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="center") +
+  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
+  geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
+  geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+              data = all_t_test.df) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  theme_linedraw() +
+  scale_x_discrete(labels=dataset_names) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Proportion Female", x = "Dataset")
+
+all_cent_sum.df %>%
+  ggplot(aes(x=dataset, y=female_ratio)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="center") +
+  facet_wrap(~source, scales="free_y") +
+  #geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
+  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  #             data = all_t_test.df) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"),  name = "ComBat Configuration", labels=dataset_names) + 
+  theme_linedraw()
+```
+
+linear ComBat ~ too few women with high centiles, too many women with low centiles; ~ balances out overall
+
+ComBat-GAM -> variable pattern, pretty close to raw
 
 ComBatLS -> pretty close to raw!
-
-```{r}
-#one-sample tests against raw vals
-
-one_samp_wilcox <- function(cf_method, df){
-  df2 <- df %>%
-    dplyr::filter(dataset == cf_method)
-  raw_val <- df[df$dataset == "raw" & df$perm=="perm-001",][["male_ratio"]]
-  wilcox.test(df2$male_ratio, mu=raw_val)
-}
-
-cf_test_list <- unique(perm_subj.df$dataset)
-
-high_test <- lapply(cf_test_list, one_samp_wilcox, df=high_cent.sum)
-names(high_test) <- cf_test_list
-high_test
-
-low_test <- lapply(cf_test_list, one_samp_wilcox, df=low_cent.sum)
-names(low_test) <- cf_test_list
-low_test
-
-ext_test <- lapply(cf_test_list, one_samp_wilcox, df=ext_cent.sum)
-names(ext_test) <- cf_test_list
-ext_test
-```
 
@@
   summarise(across(contains("mean"), list(site = mean)))
-
-perm_subj.df2.by.site %>%
-  ggplot() +
-  geom_boxplot(aes(x=sim.site, y=mean_cent_abs.diff_site, fill=dataset)) +
-  facet_wrap(~ dataset, scales="free")
 
@@
 
+## FIGURE 3
+```{r}
+blank <- ggplot() + theme_void()
+sex.var.plot.global <- readRDS("figs/sex_var_plot_global.RDS")
+sex.var.plot <- readRDS("figs/sex_var_plot.RDS")
+
+fig_3 <- ggarrange(
+  ggarrange(
+  ggarrange(blank, sex.var.plot.global, sex.var.plot,
+  ncol=3, nrow=1, 
+  widths = c(.1, .8, 3.1),
+  hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  ),
+  sex_bias_density_plt,
+  nrow=2, ncol=1,
+  labels=c("A", "B")
+  ),
+  all_cent_sum_boxplt.diff,
+  ncol=2, nrow=1,
+  labels=c(NA, "C"),
+  widths = c(1, 1)
+  # #align = "hv",
+  # heights=c(1, 1.8),
+  # common.legend = FALSE,
+  # #labels=c("A", "B")
+  )
+
+fig_3
+ggsave("figs/figure_3.jpeg", fig_3, width=15, height=10, units="in")
+```
+
 ## Supplemental Fig
@@
 ```
+
+Figure 3 but with sig. tests for boxplots
+
+```{r}
+sex.var.plot.global <- readRDS("figs/sex_var_plot_global.RDS")
+sex.var.plot <- readRDS("figs/sex_var_plot.RDS")
+
+fig_3_wtest <- ggarrange(
+  ggarrange(
+  ggarrange(sex.var.plot.global, sex.var.plot,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.25),
+  hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  ),
+  sex_bias_density_plt,
+  nrow=2, ncol=1,
+  labels=c("A", "B")
+  ),
+  all_cent_sum_boxplt_wtest,
+  ncol=2, nrow=1,
+  labels=c(NA, "C"),
+  widths = c(1, 1.8)
+  # #align = "hv",
+  # heights=c(1, 1.8),
+  # common.legend = FALSE,
+  # #labels=c("A", "B")
+  )
+
+fig_3_wtest
+ggsave("figs/figure_3_wtest.jpeg", fig_3, width=15, height=10, units="in")
+```

### 2024-03-13 paper_replication_results.Rmd
cleaned upgit add paper*

@@
 
+**need to prettify table for supplement**
+
 ```{r}
@@
 
-look at combatGAM vs combatLS
-
 ```{r}
+#look at combatGAM vs combatLS
 cent.feat.t.test.df %>%
@@
   geom_bar() +
-  geom_text(stat='count', aes(label=..count..)) +
+  # geom_text(stat='count', aes(label=..count..)) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   facet_wrap(~pheno_cat)
+
+cf.gam_worse <- cent.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam" & bigger_group == "cf.gamlss") %>%
+  group_by(perm) %>%
+  summarise(count = n())
+
+mean(cf.gam_worse$count)
+IQR(cf.gam_worse$count)
 ```
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  scale_fill_gradient(low="#605210", high= "#e6d16d",  name = "% Replications where ComBatLS is Better")
+  scale_fill_gradient(low="#605210", high= "#e6d16d",  name = "% Replications") +
+  labs(title="Replications where ComBatLS recapitulates cortical thickness centiles \nsignificantly better than ComBat-GAM")
 ```
@@
 
-Plot raw values from random permutation
-
-```{r, eval=FALSE}
-raw_files <- list.files(path = data_path, pattern = "_diffs.csv", full.names = TRUE) %>% unlist()
-
-diff.df.rand <- fread(raw_files[3])
-
-diff.df.rand <- diff.df.rand %>%
-  mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))
-
-global_abs.diffs <- diff.df.rand %>%
-  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", pheno_list)) %>%
-  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
-    mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         abs.diff = abs.diff*100,
-         pheno_cat = factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_), ordered=TRUE, levels=c("Global Volume", "Regional Volume", "Regional SA", "Regional CT"))) %>%
-    mutate(pheno = factor(pheno, ordered=TRUE, levels=c(vol_list_global, vol_list_regions, sa_list, ct_list)))#sort w/in pheno_cat
-
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-all.abs.err.plt <- global_abs.diffs %>%
-  dplyr::filter(dataset != "cf") %>%
-  ggplot() +
-  geom_boxplot(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=abs.diff, fill=dataset), outlier.size=.1) +
-  theme_linedraw(base_size = 11) +
-  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Absolulte Centile Error", x = "Feature") +
-  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
-  facet_wrap(~dataset, nrow=3, labeller = as_labeller(dataset_names), scales="free_y") +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="none")
-ggsave("figs/all_abs_err_plt_wlabels.jpeg", all.abs.err.plt, width=16, height=9, units="in")
-```
-
 ### Sex-Biases
@@
   geom_bar() +
-  geom_text(stat='count', aes(label=..count..)) +
+  #geom_text(stat='count', aes(label=..count..)) +
   facet_wrap(~dataset) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  theme(axis.text.x = element_blank()) +
   scale_fill_manual(values = c("lightblue", "red"))
@@
                                p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ "")) %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")
+                               p.val_fdr >= 0.05 ~ ""))
+
+t.test.cent.sex.test.med.filt <- t.test.cent.sex.test.med %>% dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")
 
@@
   theme_linedraw() +
+  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med.filt$dataset == "cf.gam", 10.2, 5.5)),
+              data = t.test.cent.sex.test.med.filt) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom")+
+  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
+
+cent.sex.test.med %>%
+  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot() +
+  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
+  facet_wrap(~pheno_cat) +
+  theme_linedraw() +
   geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med$dataset == "cf.gam", 10.2, 5.5)),
@@
   labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
-
-cent.sex.test.med %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
-  ggplot() +
-  geom_boxplot(aes(y=median_of_median_sex_diff*100, x=dataset, color=dataset)) +
-  geom_hline(yintercept = 0, linetype="dashed", color="#3F3F3F") +
-  facet_wrap(~pheno_cat, scales="free") +
-  theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos*100, y=10.2),
-              data = t.test.cent.sex.test.med) +
-  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  theme(legend.position = "bottom")+
-  labs(x="Medians of Sex-Differences in Centile Errors Within Brain Feature Types", y="Density Across 100 Replications")
 ```
@@
 
-all_cent_sum.df %>%
-  ggplot(aes(x=dataset, y=female_ratio)) +
-  geom_boxplot(aes(color=dataset)) +
-  # geom_dotplot(binaxis = "y", stackdir="center") +
-  facet_wrap(~source, scales="free_y") +
-  #geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
-  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
-  #             data = all_t_test.df) +
-  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"),  name = "ComBat Configuration", labels=dataset_names) + 
-  theme_linedraw()
 ```
@@
 ComBatLS -> pretty close to raw!
-
-by site
-```{r}
-perm_subj.df2.by.site <- perm_subj.df2 %>%
-  group_by(perm, dataset, sim.site) %>%
-  summarise(across(contains("mean"), list(site = mean)))
-
-perm_subj.df2.by.site %>%
-    mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
-  ggplot() +
-  geom_boxplot(aes(x=sim.site, y=mean_cent_abs.diff_site, fill=dataset)) +
-  facet_wrap(~ dataset, scales="free")
-
-perm_subj.df2.by.site %>%
-  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
-  ggplot() +
-  geom_boxplot(aes(x=sim.site, y=mean_cent_diff_site, fill=dataset)) +
-  facet_wrap(~ dataset, scales="free")
-
-perm_subj.df2.by.site %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
-  ggplot() +
-  geom_boxplot(aes(x=sim.site, y=mean_cent_diff_site, fill=dataset))
-
-perm_subj.df2.by.site %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
-  ggplot() +
-  geom_boxplot(aes(x=sim.site, y=mean_centile_site, fill=dataset))
-
-perm_subj.df2.by.site %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  pivot_wider(id_cols=c(perm, sim.site), names_from=dataset, values_from=mean_centile_site, names_prefix="site_mean_cent_") %>%
-  mutate(delta_mean = site_mean_cent_cf.gam - site_mean_cent_cf.gamlss) %>%
-  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
-  ggplot() +
-  geom_boxplot(aes(x=sim.site, y=delta_mean))
-```
 
@@
 ```{r}
-noext.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
-noext.subj.sex.test.df <- bind_rows(noext.subj.sex.test.list)
+noext.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
+noext.sex.test.df <- bind_rows(noext.sex.test.list)
 ```
 
-simple bar-plot to count regions w sigificant sex-biases
+simple bar-plot to count regions w sigificant sex-differences
 
 ```{r}
-noext.subj.sex.test.df %>%
+noext.sex.test.df %>%
   mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
@@
   geom_bar() +
-  geom_text(stat='count', aes(label=..count..)) +
+  #geom_text(stat='count', aes(label=..count..)) +
   facet_wrap(~dataset) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  theme(axis.text.x = element_blank()) +
   scale_fill_manual(values = c("lightblue", "red"))
+```
+
+Distribution of error medians w/in each pheno category
+```{r}
+noext.sex.testmed <- noext.sex.test.df %>%
+   mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+  group_by(dataset, perm, pheno_cat) %>%
+  summarise(median_of_median_sex_diff = median(median_sex_diff))
+
+next.t.test.cent.sex.test.med <- noext.sex.testmed %>%
+  group_by(dataset, pheno_cat) %>%
+  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
+            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
+            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(cent.sex.test.med$dataset))*length(unique(cent.sex.test.med$pheno_cat))),
+                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
+                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ ""))
+
+t.test.cent.sex.test.med %>%
+  ggplot() +
+  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")
+
+next.t.test.cent.sex.test.med %>%
+  ggplot() +
+  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")
+
+noext.sex.testmed %>%
+  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot() +
+  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
+  facet_wrap(~pheno_cat) +
+  theme_linedraw() +
+  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med$dataset == "cf.gam", 10.2, 5.5)),
+              data = next.t.test.cent.sex.test.med) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom")+
+  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
 ```
@@
   geom_bar(aes(x=perm, fill=bigger_group)) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~comp)
+  theme(axis.text.x = element_blank()) +
+  facet_grid(pheno_cat~comp, scales="free_y", labeller = labeller(comp = as_labeller(comp_names)))
 ```
@@
   geom_bar() +
-  geom_text(stat='count', aes(label=..count..)) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~pheno_cat)
+
+z.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
+ggplot(aes(x=perm, fill=bigger_group)) +
+  geom_bar(position='fill') +
+  geom_hline(yintercept = .5) +
+  theme(axis.text.x = element_blank()) +
   facet_wrap(~pheno_cat)
@@
   ggplot() +
-  geom_violin(aes(x=dataset, y=mean_z_abs.diff, fill=perm), scale="count") +
+  geom_violin(aes(x=dataset, y=mean_z_abs.diff, color=perm), scale="count") +
   labs(y= "Mean Absolute Z-Score Error", x = "ComBat Configurations") +
   scale_x_discrete(labels=c(dataset_names)) +
-  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
+  theme(legend.position = "none") +
   ggtitle("Within-subject mean magnitude of error in z scores across ComBat \nconfigurations replicated 10x")
@@
 ```{r}
-z.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
-z.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
+z.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
+z.sex.test.df <- bind_rows(z.sex.test.list)
 ```
@@
 ```{r}
-z.subj.sex.test.df %>%
+z.sex.test.df %>%
   mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
@@
   geom_bar() +
-  geom_text(stat='count', aes(label=..count..)) +
+  #geom_text(stat='count', aes(label=..count..)) +
   facet_wrap(~dataset) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  theme(axis.text.x = element_blank()) +
   scale_fill_manual(values = c("lightblue", "red"))
+```
+
+Distribution of error medians w/in each pheno category
+```{r}
+z.sex.test.med <- z.sex.test.df %>%
+  mutate(pheno = gsub(".z", "", pheno)) %>%
+   mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+  group_by(dataset, perm, pheno_cat) %>%
+  summarise(median_of_median_sex_diff = median(median_sex_diff))
+
+t.test.z.sex.test.med <- z.sex.test.med %>%
+  group_by(dataset, pheno_cat) %>%
+  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
+            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
+            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(z.sex.test.med$dataset))*length(unique(z.sex.test.med$pheno_cat))),
+                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
+                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ ""))
+
+t.test.z.sex.test.med.filt <- t.test.z.sex.test.med %>% dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")
+
+t.test.z.sex.test.med %>%
+  ggplot() +
+  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")
+
+z.sex_bias_density_plt <- z.sex.test.med %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot() +
+  geom_density(aes(x=median_of_median_sex_diff, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
+  facet_wrap(~pheno_cat) +
+  theme_linedraw() +
+  geom_text(aes(label = sig_fdr, x= x_pos, y=ifelse(t.test.z.sex.test.med.filt$dataset == "cf.gam", 350, 120)),
+              data = t.test.z.sex.test.med.filt) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom")+
+  labs(x="Medians of Sex-Differences in Z Errors", y="Density Across 100 Replications")
+
+z.sex.test.med %>%
+  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot() +
+  geom_density(aes(x=median_of_median_sex_diff, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
+  facet_wrap(~pheno_cat) +
+  theme_linedraw() +
+  geom_text(aes(label = sig_fdr, x= x_pos, y=ifelse(t.test.z.sex.test.med$dataset == "cf.gam", 350, 120)),
+              data = t.test.z.sex.test.med) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom")+
+  labs(x="Medians of Sex-Differences in Z Errors", y="Density Across 100 Replications")
+```
+
+plot subj. with averages in extremes 
+```{r}
+#recalc subject's true mean zs
+raw_mean_z <- perm_subj.df %>%
+  dplyr::filter(dataset=="cf") %>%
+  mutate(mean_z = mean_z - mean_z_diff,
+        VolGlob_mean_z = VolGlob_mean_z - VolGlob_mean_z_diff,
+        VolReg_mean_z = VolReg_mean_z - VolReg_mean_z_diff,
+        CT_mean_z = CT_mean_z - CT_mean_z_diff,
+        SA_mean_z = SA_mean_z - SA_mean_z_diff) %>%
+  dplyr::select(!c(contains("diff"), contains("centile"))) %>%
+  mutate(dataset="raw")
+
+raw_mean_z %>%
+  dplyr::filter(perm=="perm-001") %>%
+  dplyr::filter(SA_mean_z > 1.5) %>% 
+  nrow()
+
+perm_subj.df.z <- full_join(perm_subj.df, raw_mean_z)
+
+perm_subj.df.z._long <- perm_subj.df.z %>%
+  rename(allpheno_mean_z = mean_z) %>%
+  pivot_longer(cols=contains("mean_z"), names_to="pheno_cat", values_to="mean_z") %>%
+  mutate(pheno_cat = gsub("_mean_z", "", pheno_cat)) %>%
+  mutate(pheno_cat = case_when(pheno_cat == "VolGlob" ~ "Global Volume",
+                               pheno_cat == "VolReg" ~ "Regional Volume",
+                               pheno_cat == "CT" ~ "Regional CT",
+                               pheno_cat == "SA" ~ "Regional SA"))
+
+high_z.sum <- perm_subj.df.z._long %>%
+  dplyr::filter(pheno_cat != "allpheno") %>%
+  dplyr::filter(mean_z > 1.5) %>%
+  group_by(dataset, perm, pheno_cat) %>%
+  summarise(
+    male_ratio = sum(sex == "Male") / n(),
+    female_ratio = sum(sex == "Female") / n()
+  )
+
+low_z.sum <- perm_subj.df.z._long %>%
+  dplyr::filter(pheno_cat != "allpheno") %>%
+  dplyr::filter(mean_z < -1.5) %>%
+  group_by(dataset, perm, pheno_cat) %>%
+  summarise(
+    male_ratio = sum(sex == "Male") / n(),
+    female_ratio = sum(sex == "Female") / n()
+  )
+
+ext_z.sum <- perm_subj.df.z._long %>%
+  dplyr::filter(pheno_cat != "allpheno") %>%
+  dplyr::filter(mean_z < -1.5 | mean_z > 1.5) %>%
+  group_by(dataset, perm, pheno_cat) %>%
+  summarise(
+    male_ratio = sum(sex == "Male") / n(),
+    female_ratio = sum(sex == "Female") / n()
+  )
+```
+
+```{r}
+#get raw vals for horizontal lines
+high_z.sum_raw <- high_z.sum %>%
+  dplyr::filter(dataset == "raw" & perm=="perm-001")
+
+low_z.sum_raw <- low_z.sum %>%
+  dplyr::filter(dataset == "raw" & perm=="perm-001")
+
+ext_z.sum_raw <- ext_z.sum %>%
+  dplyr::filter(dataset == "raw" & perm=="perm-001")
+
+z.high_t_test.df <- lapply(split(high_z.sum, high_z.sum$pheno_cat), get_t_tests, raw_df=high_z.sum_raw) %>% bind_rows()
+z.low_t_test.df <- lapply(split(low_z.sum, low_z.sum$pheno_cat), get_t_tests, raw_df=low_z.sum_raw) %>% bind_rows()
+z.ext_t_test.df <- lapply(split(ext_z.sum, ext_z.sum$pheno_cat), get_t_tests, raw_df=ext_z.sum_raw) %>% bind_rows()
+
+#merge dfs
+all_z_sum.df <- gdata::combine(high_z.sum, low_z.sum, ext_z.sum) %>%
+  mutate(source = gsub("_z.sum", "", source))
+all_z.sum_raw <- all_z_sum.df %>%
+  dplyr::filter(dataset == "raw" & perm=="perm-001")
+
+z.all_t_test.df <- gdata::combine(z.high_t_test.df, z.low_t_test.df, z.ext_t_test.df) %>%
+  mutate(source = gsub("_t_test.df", "", source)) %>%
+  mutate(source = gsub("z.", "", source))
+
+#scale by raw value
+all_z_sum.df_scaled <- all_z_sum.df %>%
+  mutate(dataset = factor(dataset, levels=c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered=TRUE)) %>%
+  arrange(dataset) %>%
+  group_by(source, pheno_cat) %>%
+  mutate(female_ratio_sc = female_ratio/first(female_ratio),
+         female_ratio_diff = female_ratio - first(female_ratio))
+
+#plots!
+all_z_sum.df %>%
+  dplyr::filter(dataset != "raw") %>%
+  ggplot(aes(x=dataset, y=female_ratio)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="zer") +
+  facet_grid(pheno_cat~source, scales="free_y") +
+  geom_hline(data=all_z.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
+  geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+              data = z.all_t_test.df) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"),  name = "ComBat Configuration", labels=dataset_names) + 
+  theme_linedraw()
+
+z_ext_list <- c(ext = ">1.5 and <-1.5", high = ">1.5", low = "<-1.5")
+
+all_z_sum.df_scaled %>%
+  dplyr::filter(dataset != "raw" & dataset != "cf") %>%
+  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=female_ratio_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="zer") +
+  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=z_ext_list)) +
+  geom_hline(yintercept=0, linetype="dashed")+
+  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  #             data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  theme_linedraw() +
+  scale_x_discrete(labels=dataset_names) +
+  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Female Bias", x = "Dataset")
+
+all_z_sum.df_scaled %>%
+  dplyr::filter(dataset != "raw") %>%
+  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=female_ratio_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="zer") +
+  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=z_ext_list)) +
+  geom_hline(yintercept=0, linetype="dashed")+
+  geom_text(aes(label = sig_fdr, y = .07),
+              data = z.all_t_test.df) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  theme_linedraw() +
+  scale_x_discrete(labels=dataset_names) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Proportion Female", x = "Dataset")
 ```
@@
 ```{r}
-noext.subj.sex.test.z.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
-noext.subj.sex.test.z.df <- bind_rows(noext.subj.sex.test.z.list)
+noext.sex.test.z.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
+noext.sex.test.z.df <- bind_rows(noext.sex.test.z.list)
 ```
@@
 ```{r}
-noext.subj.sex.test.z.df %>%
+noext.sex.test.z.df %>%
   mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
@@
   geom_bar() +
-  geom_text(stat='count', aes(label=..count..)) +
+  #geom_text(stat='count', aes(label=..count..)) +
   facet_wrap(~dataset) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  theme(axis.text.x = element_blank()) +
   scale_fill_manual(values = c("lightblue", "red"))
+```
+
+Distribution of error medians w/in each pheno category
+```{r}
+noext.sex.test.z.med <- noext.sex.test.z.df %>%
+  mutate(pheno=gsub(".z", "", pheno)) %>%
+   mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+  group_by(dataset, perm, pheno_cat) %>%
+  summarise(median_of_median_sex_diff = median(median_sex_diff))
+
+next.t.test.z.sex.test.med <- noext.sex.test.z.med %>%
+  group_by(dataset, pheno_cat) %>%
+  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
+            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
+            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(z.sex.test.med$dataset))*length(unique(z.sex.test.med$pheno_cat))),
+                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
+                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ ""))
+
+t.test.z.sex.test.med %>%
+  ggplot() +
+  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")
+
+next.t.test.z.sex.test.med %>%
+  ggplot() +
+  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")
+
+noext.sex.test.z.med %>%
+  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  ggplot() +
+  geom_density(aes(x=median_of_median_sex_diff, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
+  facet_wrap(~pheno_cat) +
+  theme_linedraw() +
+  geom_text(aes(label = sig_fdr, x= x_pos, y=ifelse(next.t.test.z.sex.test.med$dataset == "cf.gam", 320, 100)),
+              data = next.t.test.z.sex.test.med) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom")+
+  labs(x="Medians of Sex-Differences in zile Errors", y="Density Across 100 Replications")
 ```
@@
 ```{r}
-sex.var.plot.global <- readRDS("figs/sex_var_plot_global.RDS")
-sex.var.plot <- readRDS("figs/sex_var_plot.RDS")
-
 fig_3_wtest <- ggarrange(

### 2024-03-18 eda/lifespan_eda.Rmd
fix listwise deletion

@@
 #clean up site NAs
-cn.df.imp <- cn.df.eulerfilt2 %>%
+cn.df.imp1 <- cn.df.eulerfilt2 %>%
   mutate(site = trimws(as.character(site)),
          study = trimws(study)) %>%
-  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_")))) %>%
+  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_"))))
+
+#drop NAs listwise
+cn.df.imp <- cn.df.imp1%>%
+  dplyr::select(!c(INDEX.ID, country)) %>%#drop index ID for filtering
   na.omit()
 
+skim(cn.df.imp)
+
+nrow(cn.df.imp1) - nrow(cn.df.imp)
 # unique(cn.df.imp$site)

### 2024-03-19 cubic_scripts/qsub_combat.sh
rerun lbcc

@@ -126,5 +126,5 @@ do
 	fi
 
 #qsub bash script
-	qsub -l h_vmem=64G,s_vmem=64G -N ${config}.${csv_fname} -o $bash_dir/${config}.${csv_fname}_out.txt -e $bash_dir/${config}.${csv_fname}_err.txt $bash_script
+	qsub -l h_vmem=64G,s_vmem=64G -N ${config}.${batch}.${csv_fname} -o $bash_dir/${config}.${csv_fname}_${batch}_out.txt -e $bash_dir/${config}.${csv_fname}_${batch}_err.txt $bash_script
 done

### 2024-03-19 cubic_scripts/qsub_lifespan_centiles.sh

@@ -36,7 +36,7 @@ if ! [ -d $save_path ]
 sleep 15
 #######################################################################################
 #LIST POSSIBLE CONFIGS TO SEARCH FOR
-config_list="lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.site lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.site"
+config_list="lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study" #lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.site lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.site"
 #######################################################################################
 for config in $config_list
 do

### 2024-03-19 paper_prop_results.Rmd

@@
 
-simple bar-plot to count regions w sigificant sex-biases
+simple bar-plot to count regions w sigificant sex-differences in errors
 
@@
 
-visualize magnitude of sex-biases
+visualize magnitude of sex-biases (directional magnitude of errors within pheno cat)
 
@@
 
+#just global volumes
 cent.feat.sex.test.df %>%
@@
   theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-cent.feat.sex.test.df %>%
-  group_by(dataset, prop) %>%
-  summarize(min_diff = min(median_sex_diff)*100, max_diff=max(median_sex_diff)*100)
-```
-
-Look at differences in M:F for subjects w very high or very low centiles
-```{r}
-feat_exts_long <- ratio.df %>%
-  dplyr::select(-contains(c("diff", ".z"))) %>%
-  pivot_longer(cols=c(pheno_list), names_to="pheno", values_to="centile_cat") %>%
-   mutate(pheno_cat = as.factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_))) %>%
-  mutate(centile_cat = case_when(. < 0.2 ~ "low",
-                                        . > 0.8 ~ "high", 
-                                        TRUE ~ "mid"))
-feat_exts_long %>%
-  dplyr::filter(centile_cat == "low") %>%
-  ggplot() +
-  geom_bar(aes(x=pheno_cat, fill=sex, color=sex)) +
-  facet_grid(prop ~ dataset)
-
-feat_exts_long %>%
-  dplyr::filter(centile_cat == "high") %>%
-  ggplot() +
-  geom_bar(aes(x=pheno_cat, fill=sex, color=sex)) +
-  facet_grid(prop ~ dataset)
 ```
@@
 
-```{r}
-#trying to add stats test
+```{r, eval=FALSE}
+#stats test
 global_sex_diffs <- cent.feat.sex.test.df %>%
@@
 
-```{r}
+```{r, eval=FALSE}
 sex_err_ratio.plt <- ratio.df %>%
@@
 sex_err_ratio.plt
-```
-
-look at just cf.gam & cf.gamlss
-
-```{r, eval=FALSE}
-cent.feat.sex.test.df %>%
-  dplyr::filter(dataset == "cf.gamlss") %>%
-  dplyr::select(pheno, prop, sig_fdr) %>%
-  arrange(pheno, prop)
-```
-
-Plot on brains for 60% F imbalance:
-```{r, eval=FALSE}
-aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
-
-cent.feat.sex.test.df2 <- cent.feat.sex.test.df %>%
-  dplyr::filter(prop == "prop-04") %>%
-  mutate(pheno_cat = as.factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_)),
-    label = sub("_[^_]*_", "_", pheno)) #for plotting
-sex_t_tests_to_plot <- left_join(cent.feat.sex.test.df2, dk.parc, by="label")
-
-# max(sex_t_tests_to_plot$median_sex_diff)*100
-# min(sex_t_tests_to_plot$median_sex_diff)*100
-
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-#var plot
-sex.var.plot.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         # median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-         #                      TRUE ~ median_sex_diff*100)) %>%
-         median_sex_diff = median_sex_diff*100) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_diff <- 1
-
-sex.var.plot <- ggplot(sex.var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = median_sex_diff)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-3, 3), name = "Sex Effect") #
-#sex.var.plot
-
-#plot global
-sex.var.plot.global.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-        # median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-        #                       TRUE ~ median_sex_diff*100)) %>% #drop NS values
-        median_sex_diff = median_sex_diff*100) %>%
-  group_by(dataset)
-
-tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = median_sex_diff),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-3, 3),  name = "Sex Effect") #,
-# sex.var.plot.global
-
-#Combine plots - need to keep playing around to make the spacing pretty when it saves out
-
-p_sex <- ggarrange(
-  sex.var.plot.global, sex.var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations when Imbalanced Site has 70% Female", 
-               color = "black", size = 14, hjust=.65))
-p_sexannotate
 ```
@@
 ```
-Test sex-biases in centile errors w/in each feature
+
+Test sex-differences in centile errors w/in each feature
 ```{r}
@@
 
-simple bar-plot to count regions w sigificant sex-biases
+simple bar-plot to count regions w significant sex-differences
 
@@
 ```
+
+visualize magnitude of sex-biases (directional magnitude of errors within pheno cat)
+
+```{r}
+noext_cent.feat.sex.test.df <- noext_cent.feat.sex.test.df %>%
+  mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)))
+
+noext_cent.feat.sex.test.df %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= pheno_cat, color=pheno_cat), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(prop~dataset, scales="free")
+
+noext_t.test.cent.sex.diff <- noext_cent.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  group_by(dataset, pheno_cat, prop) %>%
+  summarise(t.test.p = tidy(t.test(median_sex_diff, mu=0))$p.value,
+            max_val = max(median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = (3*4*11)), #not testing global vols, too few data points
+                  sig_fdr_tf = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE),
+         sig_fdr = case_when(
+           # p.val_fdr < 0.001 ~ "***",
+           #                     p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+           #                     p.val_fdr >= 0.01 & 
+                               p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ ""),
+         y_pos = ifelse(max_val < 0, ((max_val*100)*0.5), (max_val*100)+0.8))
+
+noext_cent.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "cf") %>%
+  mutate(median_sex_diff = median_sex_diff*100) %>%
+  ggplot(aes(x=prop, y=median_sex_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+    facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names)) + 
+  geom_text(aes(label = sig_fdr, y = y_pos),
+              data = noext_t.test.cent.sex.diff[which(noext_t.test.cent.sex.diff$dataset != "cf"),]) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Centile Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+
+#compare with full dataset
+cortical_cent_sex_diff.plt
+```
+similar overall pattern
 
@@
 
-Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
+Test w/in feature sex-differences in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
 
 ```{r}
-z.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
-z.subj.sex.test.df <- bind_rows(z.subj.sex.test.list)
+z.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
+z.feat.sex.test.df <- bind_rows(z.sex.test.list)
 ```
 
-simple bar-plot to count regions w sigificant sex-biases
+simple bar-plot to count regions w significant sex-differences
 
 ```{r}
-z.subj.sex.test.df %>%
+z.feat.sex.test.df %>%
   ggplot() +
@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+```
+
+visualize magnitude of sex-biases (directional magnitude of errors within pheno cat)
+
+```{r}
+z.feat.sex.test.df <- z.feat.sex.test.df %>%
+  mutate(pheno=gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)))
+
+z.feat.sex.test.df %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= pheno_cat, color=pheno_cat), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(prop~dataset, scales="free")
+
+t.test.z.sex.diff <- z.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  group_by(dataset, pheno_cat, prop) %>%
+  summarise(t.test.p = tidy(t.test(median_sex_diff, mu=0))$p.value,
+            max_val = max(median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = (3*4*11)), #not testing global vols, too few data points
+                  sig_fdr_tf = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE),
+         sig_fdr = case_when(
+           # p.val_fdr < 0.001 ~ "***",
+           #                     p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+           #                     p.val_fdr >= 0.01 & 
+                               p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ ""),
+         y_pos = ifelse(max_val < 0, ((max_val)*0.5), (max_val)+0.05))
+
+z.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "cf") %>%
+  mutate(median_sex_diff = median_sex_diff) %>%
+  ggplot(aes(x=prop, y=median_sex_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+    facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names)) + 
+  geom_text(aes(label = sig_fdr, y = y_pos),
+              data = t.test.z.sex.diff[which(t.test.z.sex.diff$dataset != "cf"),]) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Z Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
 ```
@@
 
-Test sex-biases in z errors w/in each feature
+Test sex-diffs in z errors w/in each feature
 
 ```{r}
-noext_z.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
-noext_z.subj.sex.test.df <- bind_rows(noext_z.subj.sex.test.list)
+noext_z.feat.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
+noext_z.feat.sex.test.df <- bind_rows(noext_z.feat.sex.test.list)
 ```
@@
 ```{r}
-noext_z.subj.sex.test.df %>%
+noext_z.feat.sex.test.df %>%
   ggplot() +
@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+```
+
+visualize magnitude of sex-biases (directional magnitude of errors within pheno cat)
+
+```{r}
+noext_z.feat.sex.test.df <- noext_z.feat.sex.test.df %>%
+  mutate(pheno=gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)))
+
+noext_z.feat.sex.test.df %>%
+  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= pheno_cat, color=pheno_cat), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(prop~dataset, scales="free")
+
+noext_t.test.z.sex.diff <- noext_z.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  group_by(dataset, pheno_cat, prop) %>%
+  summarise(t.test.p = tidy(t.test(median_sex_diff, mu=0))$p.value,
+            max_val = max(median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = (3*4*11)), #not testing global vols, too few data points
+                  sig_fdr_tf = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE),
+         sig_fdr = case_when(
+           # p.val_fdr < 0.001 ~ "***",
+           #                     p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
+           #                     p.val_fdr >= 0.01 & 
+                               p.val_fdr < 0.05 ~ "*",
+                               p.val_fdr >= 0.05 ~ ""),
+         y_pos = ifelse(max_val < 0, ((max_val)*0.5), (max_val)+0.05))
+
+noext_z.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "cf") %>%
+  mutate(median_sex_diff = median_sex_diff) %>%
+  ggplot(aes(x=prop, y=median_sex_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+    facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names)) + 
+  geom_text(aes(label = sig_fdr, y = y_pos),
+              data = noext_t.test.z.sex.diff[which(noext_t.test.z.sex.diff$dataset != "cf"),]) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Z Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
 ```

### 2024-03-20 cubic_scripts/R_scripts/fit_lifespan_mod_notbv.R
lifespan brain charts w/o tbv correction

@@ -1,0 +1,27 @@
+#Basic script to fit a gamlss model
+
+set.seed(12345)
+
+#LOAD PACKAGES
+library(gamlss)
+library(dplyr)
+library(data.table)
+
+#GET ARTS
+args <- commandArgs(trailingOnly = TRUE)
+df <- fread(args[1], stringsAsFactors = TRUE, na.strings = "")
+pheno <- as.character(args[2])
+save_path <- args[3]
+
+#FIT BASE MODEL
+base_model <- gamlss(formula = as.formula(paste0(pheno,"~ pb(age_days) + sexMale + fs_version + pb(sex.age)")),
+             sigma.formula = as.formula(paste0("~ pb(age_days) + sexMale + pb(sex.age) + fs_version")),
+             nu.formula = as.formula(paste0("~ pb(age_days) + sexMale + pb(sex.age) + fs_version")),
+             control = gamlss.control(n.cyc = 200), 
+             family = BCCG, data=df, trace = FALSE)
+
+#SAVE
+csv_name <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(as.character(args[1])))
+csv_rename <- gsub("_", "-", csv_name)
+
+saveRDS(base_model,paste0(save_path, "/", pheno, "_", csv_rename, "_no.tbv_mod.rds"))

### 2024-03-20 cubic_scripts/qsub_gamlss_notbv.sh

@@ -1,0 +1,91 @@
+#!/bin/bash
+
+#Fit GAMLSS growth chart models for each phenotype
+
+#######################################################################################
+# SET PATHS
+img=/cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif #singularity image
+base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
+pheno_path=$base/pheno_lists #path to .txt files listing phenotypes (global & regional)
+mod_script=$base/cubic_scripts/R_scripts/fit_lifespan_mod_notbv.R #path to .R script
+#######################################################################################
+#######################################################################################
+# GET ARGS
+if [ $# -lt 1 ]
+then
+	echo 1>&2 "$0: provide path to data (either single .csv or dir)"
+	exit 2
+fi
+#######################################################################################
+# MAKE DIRECTORIES 
+
+#study dir
+study_dir=$base/lifespan
+if ! [ -d $study_dir ]
+	then
+	mkdir $study_dir
+	fi
+
+#qsub script & outputs
+bash_dir=$study_dir/qsub_scripts
+if ! [ -d $bash_dir ]
+	then
+	mkdir $bash_dir
+	fi
+
+#model outputs
+gamlss_dir=$study_dir/gamlss_objs
+if ! [ -d $gamlss_dir ]
+	then
+	mkdir $gamlss_dir
+	fi
+
+#######################################################################################
+# FIT MODELS
+#if input .csv
+if [ -f $1 ]
+then
+	csv_name=$(basename $1 .csv)
+	csv_name=${csv_name//_/\-}
+
+	for list in "$pheno_path"/*
+	do
+		#iterate through list of phenotypes
+		while read -r pheno
+		do
+			#write bash script
+			bash_script=$bash_dir/${pheno}_${csv_name}_fit.sh
+			touch $bash_script
+			echo "singularity run --cleanenv $img Rscript --save $mod_script $1 $pheno $gamlss_dir" > $bash_script
+
+			#qsub bash script
+			qsub -N ${pheno}.${csv_name} -o $bash_dir/${pheno}_${csv_name}_no.tbv_out.txt -e $bash_dir/${pheno}_${csv_name}_no.tbv_err.txt $bash_script
+
+		done < $list
+	done
+#if input directory of .csvs
+elif [ -d $1 ]
+then
+	for csv_file in "$1"/*.csv
+	do
[…]

### 2024-03-20 paper_lbcc_results.Rmd
parsing through results, fix na.omit

@@
 ```{r}
-gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.site_predictions.csv"))
+gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv"))
+gam_centile_df <- gam_centile_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
+
 dim(gam_centile_df)
@@
 ```{r}
-gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.site_predictions.csv"))
+gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv"))
+gamlss_centile_df <- gamlss_centile_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
+
 dim(gamlss_centile_df)
@@
 ```{r}
-deltas_df <- fread(file=paste0(der_path, "/batch.site_pred_deltas.csv"))
+deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
 # skimr::skim(deltas_df)
@@
   mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
-ggplot(aes(y=interaction(site,study, sep = "&"), x=age_years)) +
+ggplot(aes(y=study, x=age_years)) +
   geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
   theme_linedraw(base_size = 11) +
-  guides(y = ggh4x::guide_axis_nested(delim = "&")) +
-  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text=element_blank(), ggh4x.axis.nesttext.y = element_text(size=6)) +
-  labs(color="Sex", y = "Site")
+  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text.x=element_blank()) +
+  labs(color="Sex", y = "Study")
 
@@
   ggplot() +
-  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=interaction(site,study, sep = "&"), fill=sex)) +
+  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=study, fill=sex)) +
   scale_x_continuous(labels = scales::percent) +
@@
   ggplot() +
-  geom_histogram(aes(x=delta, color=study, fill=study))
+  geom_histogram(aes(x=delta, color=study, fill=study)) +
+  theme(legend.position="none")
 ```
@@
   mutate(abs.delta = abs(delta)) %>%
-  group_by(sex, study, site) %>%
+  group_by(sex, study) %>%
   summarise(sex_n = n(),
@@
 
-#get site-level stats
+#get study-level stats
 delta_summary.df <- cent_deltas_df_long %>%
   mutate(abs.delta = abs(delta)) %>%
-  group_by(study, site) %>%
-  summarise(site_mean_abs.delta=mean(abs.delta),
-            site_mean_delta=mean(delta),
-            site_mean_age=mean(age_days),
-            site_mean_age_yrs=mean(age_days)/365.25,
-            site_max_abs.delta=max(abs.delta),
-            site_n=n()) %>%
+  group_by(study) %>%
+  summarise(study_mean_abs.delta=mean(abs.delta),
+            study_mean_delta=mean(delta),
+            study_mean_age=mean(age_days),
+            study_mean_age_yrs=mean(age_days)/365.25,
+            study_max_abs.delta=max(abs.delta),
+            study_n=n()) %>%
   mutate(pheno_cat="all")
-
-# #get w/in cat stats
-# delta_summary.cat.df <- cent_deltas_df_long %>%
-#   mutate(abs.delta = abs(delta),
-#          pheno_cat = case_when(
-#     pheno %in% vol_list_global ~ "Global Volume",
-#     pheno %in% vol_list_regions ~ "Regional Volume",
-#     pheno %in% sa_list ~ "Regional SA",
-#     pheno %in% ct_list ~ "Regional CT",
-#     TRUE ~ NA_character_)) %>%
-#   group_by(study, site, pheno_cat) %>%
-#   summarise(site_mean_abs.delta=mean(abs.delta),
-#             site_mean_delta=mean(delta),
-#             site_mean_age=mean(age_days),
-#             site_mean_age_yrs=mean(age_days)/365.25,
-#             site_max_abs.delta=max(abs.delta),
-#             site_n=n())
-# 
-# delta_summary.df <- full_join(delta_summary.df, delta_summary.cat.df)
-
-delta_n.sex %>%
-  dplyr::filter(site=="abide2_BNI_1")
 ```
@@
 delta_sum.by.sex.df <- full_join(delta_summary.df, delta_n.sex) %>%
-  mutate(sex_prop=sex_n/site_n,
+  mutate(sex_prop=sex_n/study_n,
          majority = ifelse(sex_prop > .5, 1, 0))
@@
 delta_sum.by.sex.df_piv <- delta_sum.by.sex.df %>%
-  pivot_wider(id_cols=c(study, site, site_n, site_mean_age, site_mean_age_yrs, site_mean_delta, site_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta))
+  pivot_wider(id_cols=c(study, study_n, study_mean_age, study_mean_age_yrs, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta))
 
@@
 to_merge <- delta_sum.by.sex.df_piv %>%
-  dplyr::select(study, site, sex_prop_Female) %>%
+  dplyr::select(study, sex_prop_Female) %>%
   mutate_at(vars(sex_prop_Female), ~replace_na(., 0)) #set prop Female to 0 if no females in study
@@
 delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
-  geom_count(aes(size=site_n)) +
+  ggplot(aes(x=sex_prop_Female, y=study_mean_delta)) +
+  geom_count(aes(size=study_n)) +
   geom_smooth(method="lm")
@@
 delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=mean_delta, color=site, fill=site)) +
+  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
   geom_count(aes(size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=site)) +
+  geom_line(aes(group=study)) +
   theme(legend.position="none")
@@
   geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=site), color="darkgray")
+  geom_line(aes(group=study), color="darkgray")
 
@@
 delta_sum.by.sex.df_piv %>%
-  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta, color=study)) +
-  geom_count(aes(size=site_n)) +
+  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta)) +
+  geom_count(aes(size=study_n, color=study)) +
   geom_smooth(method="lm")
@@
   ggplot() +
-  geom_histogram(aes(x=site_max_abs.delta*100))
+  geom_histogram(aes(x=study_max_abs.delta*100))
 
 delta_sum.df_merged %>%
-  ggplot(aes(y=site_max_abs.delta*100, x=site_n)) +
+  ggplot(aes(y=study_max_abs.delta*100, x=study_n)) +
   geom_point() +
@@
 delta_sum.df_merged %>%
-  ggplot(aes(y=site_max_abs.delta*100, x=site_mean_age_yrs)) +
+  ggplot(aes(y=study_max_abs.delta*100, x=study_mean_age_yrs)) +
   geom_point() +
@@
 delta_sum.df_merged %>%
-  ggplot(aes(y=site_max_abs.delta*100, x=sex_prop_Female)) +
+  ggplot(aes(y=study_max_abs.delta*100, x=sex_prop_Female)) +
   geom_point() +
@@
 ```
-#Site Effects
+#Batch Effects
 
-Test to see if centile score is associated with site in either ComBatLS or ComBat-GAM harmonized data.
+Test to see if centile score is associated with study in either ComBatLS or ComBat-GAM harmonized data.
 
@@
   # print(pheno)
-  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["site"]]))
+  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["study"]]))
   df$pheno <- pheno
@@
   nrow()
-```
-
-Test single pheno
-
-```{r}
-gam_centile_df.clean %>%
-  ggplot() +
-  geom_boxplot(aes(x=site, y=WMV)) +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5))
-
-summary(lm(WMV ~ site + age_days + sex, data = gam_centile_df.clean))
 ```
@@
   ggplot() +
-  geom_boxplot(aes(x=site, y=mean_cent)) +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5))
-
-gam_centile_df.mean %>%
-  ggplot() +
-  geom_boxplot(aes(x=study, y=mean_cent)) +
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
+  #geom_boxplot(aes(x=study, y=mean_cent)) +
   theme(axis.text.x = element_text(angle=90, hjust=0.5))
@@
 ```{r}
-summary(lm(mean_cent ~ site, data=gam_centile_df.mean))
 summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
-
-summary(aov(mean_cent ~ site, data=gam_centile_df.mean))
+summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
 ```
-
-success!
 
@@
   # print(pheno)
-  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["site"]]))
+  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["study"]]))
   df$pheno <- pheno
@@
   geom_bar(aes(y=sig_fdr))
+
+df_gam_feat_aov.z.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
 ```
@@
 ```{r}
-summary(lm(mean_z ~ site, data=gam_z_df.mean))
 summary(lm(mean_z ~ study, data=gam_z_df.mean))
-
 summary(aov(mean_z ~ study, data=gam_z_df.mean))
@@
 
-```{r}
-gamlss_centile_df.clean %>%
-  ggplot() +
-  geom_boxplot(aes(x=study, y=WMV))
-
-wmv_cent_gamlss.lm <- lm(WMV ~ study + age_days + sex, data=gamlss_centile_df.clean)
-
-# gtsummary::tbl_regression(wmv_cent_gamlss.lm) %>%
-#   gtsummary::add_global_p()
-
-gamlss_centile_df.clean %>%
-  ggplot() +
-  geom_boxplot(aes(x=site, y=WMV, fill=study)) +
-  theme(axis.text.x = element_text(angle=90), legend.position = "none")
-```
-
 w/in feature ANOVAs
@@
   # print(pheno)
-  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["site"]]))
+  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
   df$pheno <- pheno
@@
 ```
+
 average w/in subject
@@
   ggplot() +
-  geom_boxplot(aes(x=reorder(site, age_days), y=mean_cent)) + 
-  theme(axis.text.x = element_text(angle=90))
-
-gamlss_centile_df.mean %>%
-  ggplot() +
   geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
-  theme(axis.text.x = element_text(angle=90))
+  #geom_boxplot(aes(x=study, y=mean_cent)) +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5))
 
@@
   theme(axis.text.x = element_text(angle=90), legend.position = "none")
+
+gamlss_centile_df.mean %>%
+  dplyr::filter(age_days>365) %>%
+  ggplot(aes(x=age_days/365.25, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
 ```
@@
 ```{r}
-summary(lm(mean_cent ~ site, data=gamlss_centile_df.mean))
 summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
-
-summary(aov(mean_cent ~ site, data=gamlss_centile_df.mean))
+summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
 ```
-
-success!
 
@@
   # print(pheno)
-  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["site"]]))
+  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
   df$pheno <- pheno
@@
   geom_bar(aes(y=sig_fdr))
+
+df_gamlss_feat_aov.z.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
 ```
@@
   geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") +
-  facet_wrap(~study)
+  facet_wrap(~study) +
+  theme(legend.position="none")
 ```
@@
 ```{r}
-summary(lm(mean_z ~ site, data=gamlss_z_df.mean))
 summary(lm(mean_z ~ study, data=gamlss_z_df.mean))
-
 summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
@@
 
-#Site-Level Mean Results
+#Centile Fans
+Look at brain charts for global vols
+
+```{r}
+gamlss.GMV.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_mod.rds"))
+
+```
+
+#Study-Level Mean Results
 
@@
 gam_centile_site.mean.df <- gam_centile_df.clean_long %>%
-  group_by(study, site) %>%
-  summarize(mean_gam_cent=mean(centile)) %>%
+  group_by(study) %>%
+  summarize(mean_gam_cent=mean(centile),
+            median_gam_cent=median(centile)) %>%
   mutate(pheno_cat = "all")
@@
 
-summary(lm(mean_gam_cent ~ site_mean_age, data=gam_centile_site.mean.df2))
-summary(lm(mean_gam_cent ~ site_mean_age + sex_prop_Female + site_n, data=gam_centile_site.mean.df2))
+summary(lm(mean_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df2))
+summary(lm(mean_gam_cent ~ study_mean_age + sex_prop_Female + study_n, data=gam_centile_site.mean.df2))
+
+summary(lm(median_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df2))
+```
+
+```{r}
+#mean in global pheno
+gam_centile_site.mean.df.glob <- gam_centile_df.clean_long %>%
+  dplyr::filter(pheno %in% vol_list_global) %>%
+  group_by(study) %>%
+  summarize(mean_gam_cent=mean(centile),
+            median_gam_cent=median(centile)) %>%
+  mutate(pheno_cat = "global_vol")
+
+gam_centile_site.mean.df.glob2 <- left_join(gam_centile_site.mean.df.glob, delta_sum.df_merged, by="study")
+
+summary(lm(mean_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df.glob2))
 ```
@@
 gamlss_centile_site.mean.df <- gamlss_centile_df.clean_long %>%
-  group_by(study, site) %>%
-  summarize(mean_gamlss_cent=mean(centile)) %>%
+  group_by(study) %>%
+  summarize(mean_gamlss_cent=mean(centile),
+            median_gamlss_cent=median(centile)) %>%
   mutate(pheno_cat = "all")
@@
 
-summary(lm(mean_gamlss_cent ~ site_mean_age, data=gamlss_centile_site.mean.df2))
-summary(lm(mean_gamlss_cent ~ site_mean_age + sex_prop_Female + site_n, data=gamlss_centile_site.mean.df2))
+summary(lm(mean_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df2))
+summary(lm(mean_gamlss_cent ~ study_mean_age + sex_prop_Female + study_n, data=gamlss_centile_site.mean.df2))
+
+summary(lm(median_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df2))
+```
+
+plots
+
+```{r}
+all_means.df <- full_join(gam_centile_site.mean.df2, gamlss_centile_site.mean.df2)
+
+all_means.df.long <- all_means.df %>%
+  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent"), names_to="combat", values_to = "mean_cent") %>%
+  mutate(combat=gsub("mean_(.+)_cent", "\\1", combat))
+
+all_means.df.long %>%
+  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
+  geom_point() +
+  geom_smooth(method="lm")
+```
+
+Significant
+
+```{r}
+#mean in global pheno
+gamlss_centile_site.mean.df.glob <- gamlss_centile_df.clean_long %>%
+  dplyr::filter(pheno %in% vol_list_global) %>%
+  group_by(study) %>%
+  summarize(mean_gamlss_cent=mean(centile),
+            median_gamlss_cent=median(centile)) %>%
+  mutate(pheno_cat = "global_vol")
+
+gamlss_centile_site.mean.df.glob2 <- left_join(gamlss_centile_site.mean.df.glob, delta_sum.df_merged, by="study")
+
+summary(lm(mean_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df.glob2))
+
+summary(lm(mean_gamlss_cent ~ study_mean_age + study_n + sex_prop_Female, data=gamlss_centile_site.mean.df.glob2))
 ```
@@
 
-Site-level mean centiles are not associated with study age in either ComBat-GAM or ComBatLS
+```{r}
+all_means.df.glob <- full_join(gam_centile_site.mean.df.glob2, gamlss_centile_site.mean.df.glob2)
+
+all_means.df.long.glob <- all_means.df.glob %>%
+  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent"), names_to="combat", values_to = "mean_cent") %>%
+  mutate(combat=gsub("mean_(.+)_cent", "\\1", combat))
+
+all_means.df.long.glob %>%
+  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
+  geom_point() +
+  geom_smooth(method="lm")
+```
 
@@
 # age
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ site_mean_age_yrs, data=delta_sum.df_merged))
+summary(lm(study_mean_abs.delta ~ study_mean_age_yrs, data=delta_sum.df_merged))
+summary(lm(study_mean_delta ~ study_mean_age_yrs, data=delta_sum.df_merged))
 
 #forward selection for site_mean_abs.delta
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
-full.delta.age.fit <- lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
+summary(lm(study_mean_abs.delta ~ study_mean_age_yrs + study_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
+summary(lm(study_mean_abs.delta ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
+full.delta.age.fit <- lm(study_mean_abs.delta*100 ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
+summary(full.delta.age.fit)
 
-full.delta.age.fit2 <- lm(site_mean_delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
-summary(full.delta.age.fit2)
+#forward selection for site_mean_delta
+summary(lm(study_mean_delta ~ study_mean_age_yrs + study_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
+summary(lm(study_mean_delta ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
+full.age.fit <- lm(study_mean_delta*100 ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
+summary(full.age.fit)
+```
+
+test w/o babies < 2
+```{r}
+delta_sum.df_merged_noinf <- delta_sum.df_merged %>%
+  dplyr::filter(study_mean_age_yrs >= 2)
+
+nrow(delta_sum.df_merged) - nrow(delta_sum.df_merged_noinf)
+
+summary(lm(study_mean_abs.delta ~ study_mean_age_yrs, data=delta_sum.df_merged_noinf))
+summary(lm(study_mean_delta ~ study_mean_age_yrs, data=delta_sum.df_merged_noinf))
+
+all_means.df.long %>%
+  dplyr::filter(study_mean_age_yrs >= 2) %>%
+  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
+  geom_point() +
+  geom_smooth(method="lm")
 ```
@@
   labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
-  theme(axis.title.x=element_b)
+  theme(axis.title.x=element_blank())
 

### 2024-03-20 eda/lifespan_eda.Rmd
parsing through results, fix na.omit

@@
 
+Remove infants (anyone younger than 3 yrs)
+
+```{r}
+cn.df.noinf <- cn.df.unique %>%
+  dplyr::filter(age_days >= (365.25*3)) %>%
+  mutate(study = droplevels(study),
+         site = droplevels(site))
+nrow(cn.df.unique) - nrow(cn.df.noinf)
+n_unique(cn.df.unique$study) - n_unique(cn.df.noinf$study)
+min(cn.df.noinf$age_days)
+```
+
 Visualizing euler distributions
 ```{r}
-cn.df.unique %>%
+cn.df.noinf %>%
   ggplot() +
@@
 ```{r}
-cn.df.eulerfilt <- cn.df.unique %>%
+cn.df.eulerfilt <- cn.df.noinf %>%
   dplyr::filter(!is.na(surfholes))
@@
 
-# left_join(cn.df.eulerfilt, study_mad) %>%
-#   ggplot() +
-#   geom_boxplot(aes(x=study, y=surfholes)) +
-#   geom_point(aes(x=study, y=(study_euler_median+2*study_euler_madconst1)), color="red")
-
 left_join(cn.df.eulerfilt, study_mad) %>%
@@
             mean_fail=mean(fail_percent))
-
 ```
@@
   unite(site, study, site, sep = "_", remove = FALSE) %>% #so sites with the same name don't get combined across studies!
+  dplyr::select(!c(INDEX.ID, country)) %>%#drop index ID for filtering
   na.omit()
+
 n_unique(cn.df.site$participant)
@@
 ggplot(cn.df.site) +
-  geom_bar(aes(x=site, fill=sex))+
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-```
-
-```{r}
-ggplot(cn.df.site) +
   geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
@@
 fwrite(cn.df.site, file=paste0(base_path, "/data/lifespan_CN_site-level_euler.csv"))
-```
-write out imagen data for testing
-```{r}
-imagen.df <- cn.df.site %>%
-  dplyr::filter(study=="IMAGEN")
-# skim(imagen.df)
-
-fwrite(imagen.df, file=paste0(base_path, "/data/imagen_CN.csv"))
-```
-
-troubleshooting combat
-
-```{r, eval=FALSE}
-gamlss(formula = rh_Vol_posteriorcingulate ~ pb(age_days) + sexMale + sex.age + site, sigma.formula= ~ pb(age_days) + sexMale + site, data=imagen.df)
-```
-
-looking into whether all global measures should be log-transformed to account for skew (getting neg vals for ventricles with combat-GAM)
-
-```{r}
-global_vols <- cn.df.site %>%
-  dplyr::select(sex, log_age, all_of(vol_list_global)) %>%
-  pivot_longer(cols=c(-sex, -log_age), names_to = "pheno", values_to = "volume") %>%
-  mutate(log_vol = log(volume, base=10)) #test log-transform
-
-ggplot(global_vols) +
-  geom_point(aes(x=log_age, y=volume, color=sex), alpha=0.4) +
-  facet_wrap(~pheno, scales="free_y")
-
-ggplot(global_vols) +
-  geom_histogram(aes(x=volume, fill=pheno), alpha=0.5) +
-  facet_wrap(~pheno, scales="free")
-
-ggplot(global_vols) +
-  geom_histogram(aes(x=log_vol, fill=pheno), alpha=0.5) +
-  facet_wrap(~pheno, scales="free")
 ```
@@
 nrow(cn.df.imp1) - nrow(cn.df.imp)
-# unique(cn.df.imp$site)
+n_unique(cn.df.imp$site)
+n_unique(cn.df.imp$study)
 # table(cn.df.imp$site, cn.df.imp$study)
@@
 ggplot(cn.df.imp) +
-  geom_bar(aes(x=site, fill=sex))+
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-```
-
-```{r}
-ggplot(cn.df.imp) +
   geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
@@
 ```
-
-## Visualize Harmonized Data
-
-comparing ventricle vals across harmonization pipelines (neg vals in `lifespan-CN-imp-sites_cf.gam_data.csv`), may need to log-transform
-```{r}
-lifespan_imp_gam.df <- fread(paste0(base_path, "/data/lifespan/lifespan-CN-imp-sites_cf.gam_data.csv"), stringsAsFactors = TRUE)
-lifespan_imp_gamlss.df <- fread(paste0(base_path, "/data/lifespan/lifespan-CN-imp-sites_cf.gamlss_data.csv"), stringsAsFactors = TRUE)
-
-lifespan_gam.df <- fread(paste0(base_path, "/data/lifespan/lifespan-CN-site-level_cf.gam_data.csv"), stringsAsFactors = TRUE)
-lifespan_gamlss.df <- fread(paste0(base_path, "/data/lifespan/lifespan-CN-site-level_cf.gamlss_data.csv"), stringsAsFactors = TRUE)
-```
-
-```{r}
-#mark negative vals
-# cn.df.site <- as.data.frame(cn.df.site)
-neg1 <- cn.df.site %>%
-  dplyr::filter(Ventricles <= 0)
-ven_raw <- ggplot() +
-  geom_point(data=cn.df.site, aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
-  geom_point(data=neg1, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
-  ggtitle("Raw Data")
-
-neg2 <- lifespan_gam.df[lifespan_gam.df$Ventricles <=0]
-ven_gam <- ggplot() +
-  geom_point(data=lifespan_gam.df, aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
-  geom_point(data=neg2, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
-  ggtitle("ComBat-GAM")
-
-neg3 <- lifespan_gamlss.df[lifespan_gamlss.df$Ventricles <=0]
-ven_gamlss <- ggplot(lifespan_gamlss.df) +
-  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
-  geom_point(data=neg3, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
-  ggtitle("ComBatLS")
-
-# cn.df.imp <- as.data.frame(cn.df.imp)
-neg4 <- cn.df.imp %>%
-  dplyr::filter(Ventricles <= 0)
-ven_raw.imp <- ggplot(cn.df.imp) +
-  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
-  geom_point(data=neg4, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
-  ggtitle("Raw Data - All Studies")
-
-neg5 <- lifespan_imp_gam.df[lifespan_imp_gam.df$Ventricles <=0]
-ven_gam.imp <- ggplot(lifespan_imp_gam.df) +
-  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
-  geom_point(data=neg5, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
-  ggtitle("ComBat-GAM - All Studies")
-
-neg6 <- lifespan_imp_gamlss.df[lifespan_imp_gamlss.df$Ventricles <=0]
-ven_gamlss.imp <- ggplot(lifespan_imp_gamlss.df) +
-  geom_point(aes(x=log_age, y=Ventricles, color=sex), alpha =0.3) +
-  geom_point(data=neg6, aes(x = log_age, y=Ventricles), color='red', alpha=0.6) + 
-  ggtitle("ComBatLS - All Studies")
-
-neg_ventricle_plots <- ggarrange(ven_raw, ven_gam, ven_gamlss, ven_raw.imp, ven_gam.imp, ven_gamlss.imp, common.legend = TRUE)
-#ggsave(paste0(base_path, "figs/lifespan_ventricle_neg.jpg"), plot=neg_ventricle_plots)
-```
-
-## Combat Troubleshooting
-having some issues with combat on cubic, trying locally.

### 2024-04-11 paper_lbcc_results.Rmd
parsing through

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize)
 #Sys.setenv("MC_CORES"=10L)
@@
 
-Combat-GAM Centiles - site-level batch correction
+Combat-GAM Centiles - study-level batch correction w/o infants
 
@@
 dim(gamlss_centile_df)
+```
+
+```{r}
+min(gam_centile_df$age_years) == min(gamlss_centile_df$age_years)
+min(gamlss_centile_df$age_years)
 ```
@@
 tbl.df <- as.data.frame(table(cent_deltas_df$site, cent_deltas_df$study))
+min(gam_centile_df$age_years) == min(cent_deltas_df$age_years)
 ```
@@
             mean_delta=mean(delta),
-            mean_age_yrs=mean(age_days)/365.25) #get number males and females in each site
+            mean_age_yrs=mean(age_days)/365.25) #get number males and females in each study
 
@@
 
-look at max abs. delta across sites
-```{r}
-delta_sum.df_merged %>%
-  ggplot() +
-  geom_histogram(aes(x=study_max_abs.delta*100))
-
-delta_sum.df_merged %>%
-  ggplot(aes(y=study_max_abs.delta*100, x=study_n)) +
-  geom_point() +
-  geom_smooth(method="lm")
-
-delta_sum.df_merged %>%
-  ggplot(aes(y=study_max_abs.delta*100, x=study_mean_age_yrs)) +
-  geom_point() +
-  geom_smooth(method="lm")
-
-delta_sum.df_merged %>%
-  ggplot(aes(y=study_max_abs.delta*100, x=sex_prop_Female)) +
-  geom_point() +
-  geom_smooth(method="lm")
-```
 #Batch Effects
@@
 gam.feat_to_test <- intersect(pheno_list, names(gam_centile_df.clean))
+length(gam.feat_to_test)
 
@@
 df_gam_feat_aov.study <- df_gam_feat_aov %>%
+  group_by(pheno) %>%
+  mutate(etasq = sumsq / sum(sumsq)) %>%
+  ungroup() %>%
   dplyr::filter(term != "Residuals") %>%
@@
   geom_bar(aes(y=sig_fdr))
+
+df_gam_feat_aov.study %>%
+  ggplot() +
+  geom_histogram(aes(x=etasq, fill=sig_fdr), alpha=0.5)
 
@@
          age_years=age_days/365.25) %>%
-  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+  dplyr::select(!contains(".z")) #drop z-scores and non-centiles & !ends_with("_total") & !contains("TBV")
+```
+
+```{r}
+ggplot(gamlss_centile_df.clean) +
+  geom_point(aes(x=age_days, y=rh_Vol_insula, color=study), alpha=0.5) +
+  theme(legend.position="none")
 ```
@@
 df_gamlss_feat_aov.study <- df_gamlss_feat_aov %>%
+  group_by(pheno) %>%
+  mutate(etasq = sumsq / sum(sumsq)) %>%
+  ungroup() %>%
   dplyr::filter(term != "Residuals") %>%
@@
 df_gamlss_feat_aov.study %>%
+  ggplot() +
+  geom_histogram(aes(x=etasq, fill=sig_fdr))
+
+df_gamlss_feat_aov.study %>%
   dplyr::filter(sig_fdr==TRUE) %>%
   nrow()
+```
+
+```{r}
+df_gamlss_feat_aov.study$combat <- "gamlss"
+df_gam_feat_aov.study$combat <- "gam"
+
+feat_aov.df <- full_join(df_gamlss_feat_aov.study, df_gam_feat_aov.study)
+
+feat_aov.df <- feat_aov.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+feat_aov.df %>%
+  ggplot() +
+  geom_point(aes(x=etasq, y=pheno, color=combat, shape=sig_fdr), alpha=0.6) +
+  facet_wrap(~pheno_cat, scales="free")
 ```
@@
 ```{r}
+ggplot(gamlss_centile_df.mean) +
+  geom_point(aes(x=(age_days/365.25), y=mean_cent, color=study), alpha=0.5) +
+  theme(legend.position="none")
+
+ggplot(gamlss_centile_df.clean) +
+  geom_point(aes(x=age_days, y=lh_Vol_inferiortemporal, color=study), alpha=0.2) +
+  theme(legend.position="none")
+```
+
+```{r}
 gamlss_centile_df.mean %>%
@@
   ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
-  theme(legend.position = "none") +
-  facet_wrap(~study)
-
-gamlss_centile_df.mean %>%
-  ggplot() +
-  geom_boxplot(aes(x=site, y=mean_cent, fill=study)) + 
-  theme(axis.text.x = element_text(angle=90), legend.position = "none")
+  geom_point(aes(x=age_days/365.25, y=mean_cent, color=study), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_cent, color=study), method="lm") +
+  theme(legend.position = "none")
 
@@
 
-#Centile Fans
-Look at brain charts for global vols
-
-```{r}
-gamlss.GMV.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_mod.rds"))
-
-```
-
 #Study-Level Mean Results
@@
 
-NS
-
 ComBatLS mean centile ~ age
@@
 
-NS
-
 ```{r}
@@
 summary(lm(study_mean_delta ~ study_mean_age_yrs, data=delta_sum.df_merged))
-
-#forward selection for site_mean_abs.delta
-summary(lm(study_mean_abs.delta ~ study_mean_age_yrs + study_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
-summary(lm(study_mean_abs.delta ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
-full.delta.age.fit <- lm(study_mean_abs.delta*100 ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
-summary(full.delta.age.fit)
-
-#forward selection for site_mean_delta
-summary(lm(study_mean_delta ~ study_mean_age_yrs + study_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
-summary(lm(study_mean_delta ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
-full.age.fit <- lm(study_mean_delta*100 ~ study_mean_age_yrs + study_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
-summary(full.age.fit)
-```
-
-test w/o babies < 2
-```{r}
-delta_sum.df_merged_noinf <- delta_sum.df_merged %>%
-  dplyr::filter(study_mean_age_yrs >= 2)
-
-nrow(delta_sum.df_merged) - nrow(delta_sum.df_merged_noinf)
-
-summary(lm(study_mean_abs.delta ~ study_mean_age_yrs, data=delta_sum.df_merged_noinf))
-summary(lm(study_mean_delta ~ study_mean_age_yrs, data=delta_sum.df_merged_noinf))
-
-all_means.df.long %>%
-  dplyr::filter(study_mean_age_yrs >= 2) %>%
-  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
-  geom_point() +
-  geom_smooth(method="lm")
 ```
@@
 delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=site_mean_delta)) +
-  geom_count(aes(size=site_n)) +
+  ggplot(aes(x=study_mean_age_yrs, y=study_mean_delta)) +
+  geom_count(aes(size=study_n)) +
   geom_smooth(method="lm")
 delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
-  geom_count(aes(size=site_n)) +
+  ggplot(aes(x=study_mean_age_yrs, y=study_mean_abs.delta)) +
+  geom_count(aes(size=study_n)) +
   geom_smooth(method="lm")
@@
 ```{r}
-summary(lm(site_mean_age_yrs ~ site_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
+summary(lm(study_mean_age_yrs ~ study_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
 ```
@@
   ggplot() +
-  geom_histogram(aes(x=site_mean_age_yrs)) +
-  labs(title="site distribution across age")
+  geom_histogram(aes(x=study_mean_age_yrs)) +
+  labs(title="study distribution across age")
 ```
@@
 ```{r}
-gam_centile_site.mean.df <- gam_centile_df.clean %>%
+gam_centile_study.mean.df <- gam_centile_df.clean %>%
   pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(site) %>%
-  summarize(mean_gam_cent=mean(centile))
-
-gam_centile_site.mean.df2 <- left_join(gam_centile_site.mean.df, delta_sum.df_merged)
-
-gam.prop.f.test <- lm(mean_gam_cent ~ sex_prop_Female + site_n + site_mean_age, data=gam_centile_site.mean.df2)
+  group_by(study) %>%
+  summarize(mean_gam_cent=mean(centile),
+            mean_gam_abs_cent=mean(abs(centile)))
+
+gam_centile_study.mean.df2 <- left_join(gam_centile_study.mean.df, delta_sum.df_merged)
+
+gam.prop.f.test <- lm(mean_gam_cent ~ sex_prop_Female + study_n + study_mean_age, data=gam_centile_study.mean.df2)
 summary(gam.prop.f.test)
-sjPlot::plot_model(gam.prop.f.test, type = "pred", terms = "sex_prop_Female")
-
-ggplot(gam_centile_site.mean.df2, aes(x=sex_prop_Female, y=mean_gam_cent)) +
+gam.prop.f.test2 <- lm(mean_gam_abs_cent ~ sex_prop_Female + study_n + study_mean_age, data=gam_centile_study.mean.df2)
+summary(gam.prop.f.test2)
+
+ggplot(gam_centile_study.mean.df2, aes(x=sex_prop_Female, y=mean_gam_cent)) +
   geom_point() +
@@
 ```{r}
-gamlss_centile_site.mean.df <- gamlss_centile_df.clean %>%
+gamlss_centile_study.mean.df <- gamlss_centile_df.clean %>%
   pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(site) %>%
-  summarize(mean_gamlss_cent=mean(centile))
-
-gamlss_centile_site.mean.df2 <- left_join(gamlss_centile_site.mean.df, delta_sum.df_merged)
-
-gamlss.prop.f.test <- lm(mean_gamlss_cent ~ sex_prop_Female + site_n + site_mean_age, data=gamlss_centile_site.mean.df2)
+  group_by(study) %>%
+  summarize(mean_gamlss_cent=mean(centile),
+            mean_gamlss_abs_cent=mean(abs(centile)))
+
+gamlss_centile_study.mean.df2 <- left_join(gamlss_centile_study.mean.df, delta_sum.df_merged)
+
+gamlss.prop.f.test <- lm(mean_gamlss_cent ~ sex_prop_Female + study_n + study_mean_age, data=gamlss_centile_study.mean.df2)
 summary(gamlss.prop.f.test)
-sjPlot::plot_model(gamlss.prop.f.test, type = "pred", terms = "sex_prop_Female")
-
-ggplot(gamlss_centile_site.mean.df2, aes(x=sex_prop_Female, y=mean_gamlss_cent)) +
+gamlss.prop.f.test2 <- lm(mean_gamlss_abs_cent ~ sex_prop_Female + study_n + study_mean_age, data=gamlss_centile_study.mean.df2)
+summary(gamlss.prop.f.test2)
+
+ggplot(gamlss_centile_study.mean.df2, aes(x=sex_prop_Female, y=mean_gamlss_cent)) +
   geom_point() +

### 2024-04-11 paper_lbcc_results.Rmd
study-level means

@@
 
-Deltas
-
-```{r}
-deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
-# skimr::skim(deltas_df)
-cent_deltas_df <- deltas_df %>%
-  mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25) %>%
-  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
-
-tbl.df <- as.data.frame(table(cent_deltas_df$site, cent_deltas_df$study))
-min(gam_centile_df$age_years) == min(cent_deltas_df$age_years)
+Unharmonized Centiles
+```{r}
+raw_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_predictions.csv"))
+raw_centile_df <- raw_centile_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
+
+dim(raw_centile_df)
+min(raw_centile_df$age_years) == min(gamlss_centile_df$age_years)
 ```
@@
 #reorder studies by age
-age.df <- cent_deltas_df %>%
+age.df <- gamlss_centile_df %>%
   group_by(study) %>%
@@
 
-sample_boxplot <- cent_deltas_df %>%
+sample_boxplot <- gamlss_centile_df %>%
   mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
@@
 full_boxplot <- 
-  cent_deltas_df %>%
+  gamlss_centile_df %>%
   mutate(ylab = "  Full \n  Sample") %>%
@@
 
-sample_barplot <- cent_deltas_df %>%
+sample_barplot <- gamlss_centile_df %>%
   mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
@@
 full_barplot <- 
-  cent_deltas_df %>%
+  gamlss_centile_df %>%
   mutate(ylab = "Full Sample") %>%
@@
 
-```{r}
-ggplot(deltas_df, aes(x=age_days, y=delta.sGMV, color=sex)) +
-  geom_point() +
-  geom_smooth()
-```
-
-```{r}
-cent_deltas_df_long <- cent_deltas_df %>%
-  pivot_longer(cols=starts_with("delta."), names_to="pheno", names_prefix = "delta.", values_to = "delta")
-
-cent_deltas_df_long %>%
-  ggplot() +
-  geom_histogram(aes(x=delta, color=study, fill=study)) +
-  theme(legend.position="none")
-```
-
-```{r}
-delta_n.sex <- cent_deltas_df_long %>%
-  mutate(abs.delta = abs(delta)) %>%
-  group_by(sex, study) %>%
-  summarise(sex_n = n(),
-            mean_abs.delta=mean(abs.delta),
-            mean_delta=mean(delta),
-            mean_age_yrs=mean(age_days)/365.25) #get number males and females in each study
-
-#get study-level stats
-delta_summary.df <- cent_deltas_df_long %>%
-  mutate(abs.delta = abs(delta)) %>%
-  group_by(study) %>%
-  summarise(study_mean_abs.delta=mean(abs.delta),
-            study_mean_delta=mean(delta),
-            study_mean_age=mean(age_days),
-            study_mean_age_yrs=mean(age_days)/365.25,
-            study_max_abs.delta=max(abs.delta),
-            study_n=n()) %>%
-  mutate(pheno_cat="all")
-```
-
-```{r}
-delta_sum.by.sex.df <- full_join(delta_summary.df, delta_n.sex) %>%
-  mutate(sex_prop=sex_n/study_n,
-         majority = ifelse(sex_prop > .5, 1, 0))
-
-delta_sum.by.sex.df_piv <- delta_sum.by.sex.df %>%
-  pivot_wider(id_cols=c(study, study_n, study_mean_age, study_mean_age_yrs, study_mean_delta, study_mean_abs.delta), names_from=sex, values_from=c(sex_n, sex_prop, mean_delta, mean_abs.delta))
-
-#Merge back female prop back into overall site-level stats
-to_merge <- delta_sum.by.sex.df_piv %>%
-  dplyr::select(study, sex_prop_Female) %>%
-  mutate_at(vars(sex_prop_Female), ~replace_na(., 0)) #set prop Female to 0 if no females in study
-
-delta_sum.df_merged <- full_join(delta_summary.df, to_merge)
-```
-
-plots!
-
-delta ~ sex-imbalance
-
-```{r}
-delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=study_mean_delta)) +
-  geom_count(aes(size=study_n)) +
-  geom_smooth(method="lm")
-
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=mean_delta, color=study, fill=study)) +
-  geom_count(aes(size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=study)) +
-  theme(legend.position="none")
-
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=mean_delta, color=as.factor(majority), fill=as.factor(majority))) +
-  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=study), color="darkgray")
-
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex_prop, y=mean_abs.delta, color=mean_age_yrs)) +
-  geom_count(aes(size=sex_n)) +
-  geom_smooth(method="lm") +
-  facet_wrap(~sex)
-
-delta_sum.by.sex.df_piv %>%
-  ggplot(aes(x=sex_prop_Female, y=study_mean_abs.delta)) +
-  geom_count(aes(size=study_n, color=study)) +
-  geom_smooth(method="lm")
-```
-
 #Batch Effects
 
-Test to see if centile score is associated with study in either ComBatLS or ComBat-GAM harmonized data.
+Test to see if centile score is associated with study in either ComBatLS, ComBat-GAM, or non-harmonized data.
+
+## Unharmonized
+```{r}
+#clean predictions
+raw_centile_df.clean <- raw_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
+  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+
+skim(raw_centile_df.clean)
+```
+
+w/in feature ANOVAs
+
+```{r}
+raw.feat_to_test <- intersect(pheno_list, names(raw_centile_df.clean))
+length(raw.feat_to_test)
+
+df_raw_feat_aov <- data.frame()
+for(pheno in raw.feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(raw_centile_df.clean[[pheno]] ~ raw_centile_df.clean[["study"]]))
+  df$pheno <- pheno
+  df_raw_feat_aov <- rbind(df_raw_feat_aov, df)
+}
+
+df_raw_feat_aov.study <- df_raw_feat_aov %>%
+  group_by(pheno) %>%
+  mutate(etasq = sumsq / sum(sumsq)) %>%
+  ungroup() %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_raw_feat_aov.study %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+
+df_raw_feat_aov.study %>%
+  ggplot() +
+  geom_histogram(aes(x=etasq, fill=sig_fdr), alpha=0.5)
+
+df_raw_feat_aov.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
+```
+```{r}
+raw_centile_df.mean <- raw_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_cent=mean(centile))
+nrow(raw_centile_df.mean) == n_unique(raw_centile_df.mean$participant)
+```
+
+```{r}
+raw_centile_df.mean %>%
+  ggplot() +
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
+  #geom_boxplot(aes(x=study, y=mean_cent)) +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5))
+
+raw_centile_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
+  theme(legend.position = "none") +
+  facet_wrap(~study)
+```
+linear model
+
+```{r}
+summary(lm(mean_cent ~ study, data=raw_centile_df.mean))
+summary(aov(mean_cent ~ study, data=raw_centile_df.mean))
+```
+###Z-Scores
+
+```{r}
+#clean predictions
+raw_z_df.clean <- raw_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
+  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+```
+
+w/in feature ANOVAs
+
+```{r}
+raw.feat_to_test.z <- paste0(raw.feat_to_test, ".z")
+df_raw_feat_aov.z <- data.frame()
+for(pheno in raw.feat_to_test.z){
+  # print(pheno)
+  df <- tidy(aov(raw_z_df.clean[[pheno]] ~ raw_z_df.clean[["study"]]))
+  df$pheno <- pheno
+  df_raw_feat_aov.z <- rbind(df_raw_feat_aov.z, df)
+}
+
+df_raw_feat_aov.z.study <- df_raw_feat_aov.z %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test.z)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_raw_feat_aov.z.study %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+
+df_raw_feat_aov.z.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
+```
+
+average w/in subject
+
+```{r}
+raw_z_df.mean <- raw_z_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_z=mean(z_score))
+nrow(raw_z_df.mean) == n_unique(raw_z_df.clean$participant)
+```
+
+```{r}
+raw_z_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
+# +
+#   facet_wrap(~study)
+```
+
+linear model
+
+```{r}
+summary(lm(mean_z ~ study, data=raw_z_df.mean))
+summary(aov(mean_z ~ study, data=raw_z_df.mean))
+```
 
@@
   nrow()
-```
-
-```{r}
-df_gamlss_feat_aov.study$combat <- "gamlss"
-df_gam_feat_aov.study$combat <- "gam"
-
-feat_aov.df <- full_join(df_gamlss_feat_aov.study, df_gam_feat_aov.study)
-
-feat_aov.df <- feat_aov.df %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-
-feat_aov.df %>%
-  ggplot() +
-  geom_point(aes(x=etasq, y=pheno, color=combat, shape=sig_fdr), alpha=0.6) +
-  facet_wrap(~pheno_cat, scales="free")
 ```
@@
 
-#Study-Level Mean Results
-
-ComBatGAM mean centile ~ age
+##Comparison Plots
+
+compare magnitude of effect size for study (w/in feature anovas) across unharmonized, combat-gam and combatLS centiles
+
+```{r}
+df_gamlss_feat_aov.study$combat <- "gamlss"
+df_gam_feat_aov.study$combat <- "gam"
+df_raw_feat_aov.study$combat <- "raw"
+
+feat_aov.df <- full_join(df_gamlss_feat_aov.study, df_gam_feat_aov.study)
+feat_aov.df <- full_join(feat_aov.df, df_raw_feat_aov.study)
+
+feat_aov.df <- feat_aov.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+feat_aov.df %>%
+  ggplot() +
+  geom_point(aes(x=etasq, y=pheno, color=combat, shape=sig_fdr), alpha=0.6) +
+  facet_wrap(~pheno_cat, scales="free")
+```
+```{r}
+gamlss_centile_df.mean$combat <- "gamlss"
+gam_centile_df.mean$combat <- "gam"
+raw_centile_df.mean$combat <- "raw"
+
+mean_cents.df <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)
+mean_cents.df <- full_join(mean_cents.df, raw_centile_df.mean)
+
+mean_cents.df %>%
+  ggplot() +
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
+  #geom_boxplot(aes(x=study, y=mean_cent)) +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5)) +
+  facet_wrap(~combat)
+```
+
+#Study-Level Mean Centiles ~ Study Ages
+
+##Unharmonzied
+
+```{r}
+raw_centile_df.clean_long <- raw_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")
+
+#overall means
+raw_centile_site.mean.df <- raw_centile_df.clean_long %>%
+  group_by(study) %>%
+  summarize(mean_raw_cent=mean(centile),
+            median_raw_cent=median(centile),
+            mean_age=mean(age_years),
+            study_n= n()) %>%
+  mutate(pheno_cat = "all")
+
+summary(lm(mean_raw_cent ~ mean_age, data=raw_centile_site.mean.df))
+summary(lm(mean_raw_cent ~ mean_age + study_n, data=raw_centile_site.mean.df))
+```
+
+## ComBatGAM
 
@@
   summarize(mean_gam_cent=mean(centile),
-            median_gam_cent=median(centile)) %>%
+            median_gam_cent=median(centile),
+            mean_age=mean(age_years),
+            med_age=median(age_years),
+            study_n= n()) %>%
   mutate(pheno_cat = "all")
 
-gam_centile_site.mean.df2 <- left_join(gam_centile_site.mean.df, delta_sum.df_merged)
-
-summary(lm(mean_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df2))
-summary(lm(mean_gam_cent ~ study_mean_age + sex_prop_Female + study_n, data=gam_centile_site.mean.df2))
-
-summary(lm(median_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df2))
-```
-
-```{r}
-#mean in global pheno
-gam_centile_site.mean.df.glob <- gam_centile_df.clean_long %>%
-  dplyr::filter(pheno %in% vol_list_global) %>%
-  group_by(study) %>%
-  summarize(mean_gam_cent=mean(centile),
-            median_gam_cent=median(centile)) %>%
-  mutate(pheno_cat = "global_vol")
-
-gam_centile_site.mean.df.glob2 <- left_join(gam_centile_site.mean.df.glob, delta_sum.df_merged, by="study")
-
-summary(lm(mean_gam_cent ~ study_mean_age, data=gam_centile_site.mean.df.glob2))
-```
-
-ComBatLS mean centile ~ age
+summary(lm(mean_gam_cent ~ mean_age, data=gam_centile_site.mean.df))
+summary(lm(median_gam_cent ~ med_age, data=gam_centile_site.mean.df))
+summary(lm(mean_gam_cent ~ mean_age + study_n, data=gam_centile_site.mean.df))
+```
+
+## ComBatLS
+
 ```{r}
@@
   summarize(mean_gamlss_cent=mean(centile),
-            median_gamlss_cent=median(centile)) %>%
+            median_gamlss_cent=median(centile),
+            mean_age=mean(age_years),
+            med_age=median(age_years),
+            study_n= n()) %>%
   mutate(pheno_cat = "all")
 
-gamlss_centile_site.mean.df2 <- left_join(gamlss_centile_site.mean.df, delta_sum.df_merged)
-
-summary(lm(mean_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df2))
-summary(lm(mean_gamlss_cent ~ study_mean_age + sex_prop_Female + study_n, data=gamlss_centile_site.mean.df2))
-
-summary(lm(median_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df2))
+summary(lm(mean_gamlss_cent ~ mean_age, data=gamlss_centile_site.mean.df))
+summary(lm(median_gamlss_cent ~ med_age, data=gamlss_centile_site.mean.df))
+summary(lm(mean_gamlss_cent ~ mean_age + study_n, data=gamlss_centile_site.mean.df))
 ```
@@
 ```{r}
-all_means.df <- full_join(gam_centile_site.mean.df2, gamlss_centile_site.mean.df2)
+all_means.df <- full_join(gam_centile_site.mean.df, gamlss_centile_site.mean.df)
+all_means.df <- full_join(all_means.df, raw_centile_site.mean.df)
 
 all_means.df.long <- all_means.df %>%
-  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent"), names_to="combat", values_to = "mean_cent") %>%
+  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent", "mean_raw_cent"), names_to="combat", values_to = "mean_cent") %>%
   mutate(combat=gsub("mean_(.+)_cent", "\\1", combat))
@@
 all_means.df.long %>%
-  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
-  geom_point() +
-  geom_smooth(method="lm")
-```
-
-Significant
-
-```{r}
-#mean in global pheno
-gamlss_centile_site.mean.df.glob <- gamlss_centile_df.clean_long %>%
-  dplyr::filter(pheno %in% vol_list_global) %>%
-  group_by(study) %>%
-  summarize(mean_gamlss_cent=mean(centile),
-            median_gamlss_cent=median(centile)) %>%
-  mutate(pheno_cat = "global_vol")
-
-gamlss_centile_site.mean.df.glob2 <- left_join(gamlss_centile_site.mean.df.glob, delta_sum.df_merged, by="study")
-
-summary(lm(mean_gamlss_cent ~ study_mean_age, data=gamlss_centile_site.mean.df.glob2))
-
-summary(lm(mean_gamlss_cent ~ study_mean_age + study_n + sex_prop_Female, data=gamlss_centile_site.mean.df.glob2))
-```
-
-```{r}
-all_means.df.glob <- full_join(gam_centile_site.mean.df.glob2, gamlss_centile_site.mean.df.glob2)
-
-all_means.df.long.glob <- all_means.df.glob %>%
-  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent"), names_to="combat", values_to = "mean_cent") %>%
-  mutate(combat=gsub("mean_(.+)_cent", "\\1", combat))
-
-all_means.df.long.glob %>%
-  ggplot(aes(x=study_mean_age_yrs, y=mean_cent, color=combat, size=study_n)) +
-  geom_point() +
+  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
+  geom_point(aes(size=study_n)) +
+  geom_smooth(method="lm")
+
+all_means.df.long %>%
+  ggplot(aes(x=med_age, y=mean_cent, color=combat)) +
+  geom_point(aes(size=study_n), alpha=0.6) +
+  geom_smooth(method="lm")
+```
+
+```{r}
+all_medians.df.long <- all_means.df %>%
+  pivot_longer(cols=c("median_gam_cent", "median_gamlss_cent", "median_raw_cent"), names_to="combat", values_to = "median_cent") %>%
+  mutate(combat=gsub("median_(.+)_cent", "\\1", combat))
+
+all_medians.df.long %>%
+  ggplot(aes(x=med_age, y=median_cent, color=combat)) +
+  geom_point(aes(size=study_n), alpha=0.6) +
+  geom_smooth(method="lm")
+```
+
+#Subject-Level Mean Centiles ~ Mean Study Age
+
+```{r}
+study_info <- gam_centile_site.mean.df %>%
+  dplyr::select(study, mean_age, med_age, study_n)
+
+raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
+summary(lm(mean_cent ~ mean_age, data=raw_centile_df.mean2))
+
+raw_centile_df.mean2 %>%
+  ggplot(aes(x=mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+```{r}
+gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
+summary(lm(mean_cent ~ mean_age, data=gam_centile_df.mean2))
+
+gam_centile_df.mean2 %>%
+  ggplot(aes(x=mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+```{r}
+gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
+summary(lm(mean_cent ~ mean_age, data=gamlss_centile_df.mean2))
+
+gamlss_centile_df.mean2 %>%
+  ggplot(aes(x=mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+Plots
+
+```{r}
+gamlss_centile_df.mean2$combat <- "gamlss"
+gam_centile_df.mean2$combat <- "gam"
+raw_centile_df.mean2$combat <- "raw"
+
+all_subj_means.df <- full_join(gamlss_centile_df.mean2, gam_centile_df.mean2)
+all_subj_means.df <- full_join(all_subj_means.df, raw_centile_df.mean2)
+
+all_subj_means.df %>%
+  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.6) +
+  geom_smooth(method="lm")
+
+all_subj_means.df %>%
+  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_wrap(~combat)
+
+summary(lm(mean_cent ~ mean_age, data=raw_centile_df.mean2))
+summary(lm(mean_cent ~ mean_age, data=gam_centile_df.mean2))
+summary(lm(mean_cent ~ mean_age, data=gamlss_centile_df.mean2))
+```
+
+```{r}
+summary(lm(mean_cent ~ study_n, data=raw_centile_df.mean2))
+summary(lm(mean_cent ~ study_n, data=gam_centile_df.mean2))
+summary(lm(mean_cent ~ study_n, data=gamlss_centile_df.mean2))
+
+all_subj_means.df %>%
+  ggplot(aes(x=study_n, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.6) +
   geom_smooth(method="lm")

### 2024-04-19 paper_lbcc_results.Rmd
lbcc looking w/in feature

@@
 
+Deltas
+
+```{r}
+deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
+deltas_df <- deltas_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
+
+dim(deltas_df)
+min(deltas_df$age_years) == min(gamlss_centile_df$age_years)
+```
+
+Nature paper models to look at age effects in variance
+from [https://github.com/brainchart/Lifespan/tree/main/Share](https://github.com/brainchart/Lifespan/tree/main/Share)
+
+```{r}
+paper_fit_sgmv <- readRDS("derivatives/nature_2020_mods/FIT_sGMV.rds")
+paper_fit_gmv <- readRDS("derivatives/nature_2020_mods/FIT_GMV.rds")
+paper_fit_wmv <- readRDS("derivatives/nature_2020_mods/FIT_WMV.rds")
+paper_fit_ven <- readRDS("derivatives/nature_2020_mods/FIT_Ventricles.rds")
+```
+
 #Clean & EDA
@@
 ```
+
 ###Z-Scores
@@
 ```{r}
-feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))
+gamlss.feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))
 
 df_gamlss_feat_aov <- data.frame()
-for(pheno in feat_to_test){
+for(pheno in gamlss.feat_to_test){
   # print(pheno)
@@
   dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(feat_to_test)), #apply fdr correction across all phenos tested
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
                   sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
@@
 ```{r}
-gamlss.feat_to_test.z <- paste0(feat_to_test, ".z")
+gamlss.feat_to_test.z <- paste0(gamlss.feat_to_test, ".z")
 df_gamlss_feat_aov.z <- data.frame()
@@
 
-feat_aov.df <- full_join(df_gamlss_feat_aov.study, df_gam_feat_aov.study)
-feat_aov.df <- full_join(feat_aov.df, df_raw_feat_aov.study)
+#put all data frames into list
+df_list.tmp <- list(df_gamlss_feat_aov.study, df_gam_feat_aov.study, df_raw_feat_aov.study)
+
+#merge all data frames in list
+feat_aov.df <- df_list.tmp %>% reduce(full_join)
 
@@
 
-mean_cents.df <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)
-mean_cents.df <- full_join(mean_cents.df, raw_centile_df.mean)
+#put all data frames into list
+df_list.tmp <- list(gamlss_centile_df.mean, gam_centile_df.mean, raw_centile_df.mean)
+
+#merge all data frames in list
+mean_cents.df <- df_list.tmp %>% reduce(full_join)
 
@@
 
-#Study-Level Mean Centiles ~ Study Ages
-
-##Unharmonzied
-
-```{r}
-raw_centile_df.clean_long <- raw_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")
-
-#overall means
-raw_centile_site.mean.df <- raw_centile_df.clean_long %>%
+#Feature Centile ~ Mean Study Age
+
+Get study info
+```{r}
+study_info <- raw_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
   group_by(study) %>%
-  summarize(mean_raw_cent=mean(centile),
-            median_raw_cent=median(centile),
-            mean_age=mean(age_years),
+  summarize(study_mean_age=mean(age_years),
+            study_median_age=median(age_years),
             study_n= n()) %>%
-  mutate(pheno_cat = "all")
-
-summary(lm(mean_raw_cent ~ mean_age, data=raw_centile_site.mean.df))
-summary(lm(mean_raw_cent ~ mean_age + study_n, data=raw_centile_site.mean.df))
-```
-
-## ComBatGAM
-
-```{r}
-gam_centile_df.clean_long <- gam_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")
-
-#overall means
-gam_centile_site.mean.df <- gam_centile_df.clean_long %>%
-  group_by(study) %>%
-  summarize(mean_gam_cent=mean(centile),
-            median_gam_cent=median(centile),
-            mean_age=mean(age_years),
-            med_age=median(age_years),
-            study_n= n()) %>%
-  mutate(pheno_cat = "all")
-
-summary(lm(mean_gam_cent ~ mean_age, data=gam_centile_site.mean.df))
-summary(lm(median_gam_cent ~ med_age, data=gam_centile_site.mean.df))
-summary(lm(mean_gam_cent ~ mean_age + study_n, data=gam_centile_site.mean.df))
-```
-
-## ComBatLS
-
-```{r}
-gamlss_centile_df.clean_long <- gamlss_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile")
-
-#overall means
-gamlss_centile_site.mean.df <- gamlss_centile_df.clean_long %>%
-  group_by(study) %>%
-  summarize(mean_gamlss_cent=mean(centile),
-            median_gamlss_cent=median(centile),
-            mean_age=mean(age_years),
-            med_age=median(age_years),
-            study_n= n()) %>%
-  mutate(pheno_cat = "all")
-
-summary(lm(mean_gamlss_cent ~ mean_age, data=gamlss_centile_site.mean.df))
-summary(lm(median_gamlss_cent ~ med_age, data=gamlss_centile_site.mean.df))
-summary(lm(mean_gamlss_cent ~ mean_age + study_n, data=gamlss_centile_site.mean.df))
-```
-
-plots
-
-```{r}
-all_means.df <- full_join(gam_centile_site.mean.df, gamlss_centile_site.mean.df)
-all_means.df <- full_join(all_means.df, raw_centile_site.mean.df)
-
-all_means.df.long <- all_means.df %>%
-  pivot_longer(cols=c("mean_gam_cent", "mean_gamlss_cent", "mean_raw_cent"), names_to="combat", values_to = "mean_cent") %>%
-  mutate(combat=gsub("mean_(.+)_cent", "\\1", combat))
-
-all_means.df.long %>%
-  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
-  geom_point(aes(size=study_n)) +
-  geom_smooth(method="lm")
-
-all_means.df.long %>%
-  ggplot(aes(x=med_age, y=mean_cent, color=combat)) +
-  geom_point(aes(size=study_n), alpha=0.6) +
-  geom_smooth(method="lm")
-```
-
-```{r}
-all_medians.df.long <- all_means.df %>%
-  pivot_longer(cols=c("median_gam_cent", "median_gamlss_cent", "median_raw_cent"), names_to="combat", values_to = "median_cent") %>%
-  mutate(combat=gsub("median_(.+)_cent", "\\1", combat))
-
-all_medians.df.long %>%
-  ggplot(aes(x=med_age, y=median_cent, color=combat)) +
-  geom_point(aes(size=study_n), alpha=0.6) +
-  geom_smooth(method="lm")
+  dplyr::select(study, study_mean_age, study_median_age, study_n)
+```
+
+W/in feat lms
+
+##Unharmonized
+
+```{r}
+raw_centile_df.clean_study <- full_join(raw_centile_df.clean, study_info)
+
+df_raw_feat_lm <- data.frame()
+for(pheno in raw.feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(raw_centile_df.clean_study[[pheno]] ~ raw_centile_df.clean_study[["study_mean_age"]]))
+  df$pheno <- pheno
+  df_raw_feat_lm <- rbind(df_raw_feat_lm, df)
+}
+
+df_raw_feat_lm <- df_raw_feat_lm %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_raw_feat_lm %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+
+df_raw_feat_lm %>%
+  ggplot() +
+  geom_histogram(aes(x=statistic))
+```
+
+##ComBat-GAM
+
+```{r}
+gam_centile_df.clean_study <- full_join(gam_centile_df.clean, study_info)
+
+df_gam_feat_lm <- data.frame()
+for(pheno in gam.feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(gam_centile_df.clean_study[[pheno]] ~ gam_centile_df.clean_study[["study_mean_age"]]))
+  df$pheno <- pheno
+  df_gam_feat_lm <- rbind(df_gam_feat_lm, df)
+}
+
+df_gam_feat_lm <- df_gam_feat_lm %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_gam_feat_lm %>%
+  ggplot() +
+  geom_histogram(aes(x=statistic))
+```
+
+##ComBatLS
+
+```{r}
+gamlss_centile_df.clean_study <- full_join(gamlss_centile_df.clean, study_info)
+
+df_gamlss_feat_lm <- data.frame()
+for(pheno in gamlss.feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(gamlss_centile_df.clean_study[[pheno]] ~ gamlss_centile_df.clean_study[["study_mean_age"]]))
+  df$pheno <- pheno
+  df_gamlss_feat_lm <- rbind(df_gamlss_feat_lm, df)
+}
+
+df_gamlss_feat_lm <- df_gamlss_feat_lm %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_gamlss_feat_lm %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+
+df_gamlss_feat_lm %>%
+  ggplot() +
+  geom_histogram(aes(x=statistic))
+```
+
+*statistic ~ effect size?
+
+```{r}
+df_raw_feat_lm$combat <- "raw"
+df_gam_feat_lm$combat <- "gam"
+df_gamlss_feat_lm$combat <- "gamlss"
+
+#put all data frames into list
+df_list.tmp <- list(df_raw_feat_lm, df_gam_feat_lm, df_gamlss_feat_lm)
+
+#merge all data frames in list
+all_feat_lm <- df_list.tmp %>% reduce(full_join)
+
+all_feat_lm %>%
+  ggplot() +
+  geom_histogram(aes(x=statistic, fill=combat, color=combat), position="identity", alpha=0.5, bins=50)
 ```
@@
 ```{r}
-study_info <- gam_centile_site.mean.df %>%
-  dplyr::select(study, mean_age, med_age, study_n)
-
 raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
-summary(lm(mean_cent ~ mean_age, data=raw_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
 
 raw_centile_df.mean2 %>%
-  ggplot(aes(x=mean_age, y=mean_cent)) +
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
   geom_point(aes(color=study), alpha=0.7) +
@@
 gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
-summary(lm(mean_cent ~ mean_age, data=gam_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
 
 gam_centile_df.mean2 %>%
-  ggplot(aes(x=mean_age, y=mean_cent)) +
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
   geom_point(aes(color=study), alpha=0.7) +
@@
   theme(legend.position="none")
+
+gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age)
+
+summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
+gam_centile_df.mean2 %>%
+  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
 ```
@@
 gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
-summary(lm(mean_cent ~ mean_age, data=gamlss_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
 
 gamlss_centile_df.mean2 %>%
-  ggplot(aes(x=mean_age, y=mean_cent)) +
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+
+gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age)
+
+summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
+gamlss_centile_df.mean2 %>%
+  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
   geom_point(aes(color=study), alpha=0.7) +
@@
 
-all_subj_means.df <- full_join(gamlss_centile_df.mean2, gam_centile_df.mean2)
-all_subj_means.df <- full_join(all_subj_means.df, raw_centile_df.mean2)
+#put all data frames into list
+df_list.tmp <- list(gamlss_centile_df.mean2, gam_centile_df.mean2, raw_centile_df.mean2)
+
+#merge all data frames in list
+all_subj_means.df <- df_list.tmp %>% reduce(full_join)
 
 all_subj_means.df %>%
-  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
+  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
   geom_point(alpha=0.6) +
@@
 all_subj_means.df %>%
-  ggplot(aes(x=mean_age, y=mean_cent, color=combat)) +
+  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
   geom_point(alpha=0.3) +
@@
 
-summary(lm(mean_cent ~ mean_age, data=raw_centile_df.mean2))
-summary(lm(mean_cent ~ mean_age, data=gam_centile_df.mean2))
-summary(lm(mean_cent ~ mean_age, data=gamlss_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
 ```
@@
 
-##Age
-
-delta ~ age
-
-```{r}
-# age
-summary(lm(study_mean_abs.delta ~ study_mean_age_yrs, data=delta_sum.df_merged))
-summary(lm(study_mean_delta ~ study_mean_age_yrs, data=delta_sum.df_merged))
+#ComBat-GAM vs ComBatLS: Delta Centile
+
+clean delta df
+
+```{r}
+cent_deltas_df <- deltas_df %>%
+  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
+
+#get subject-level means
+mean_cent_deltas_df <- cent_deltas_df %>%
+  pivot_longer(cols=any_of(paste0("delta.", pheno_list)), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_delta_cent=(mean(centile))*100, #convert units to centiles space
+            abs_mean_delta_cent=(mean(abs(centile)))*100)
+
+mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)
+```
+
+check mean subj delta ~ study mean age
+
+```{r}
+delta_lm <- lm(mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
+summary(delta_lm)
+
+mean_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=mean_delta_cent)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(delta_lm)[["study_mean_age"]], 
+              intercept = coef(delta_lm)[["(Intercept)"]])+
+  theme(legend.position="none")
+```
+
+mean abs delta ~ study mean age
+
+```{r}
+abs_lm <- lm(abs_mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
+summary(abs_lm)
+
+mean_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(abs_lm)[["study_mean_age"]], 
+              intercept = coef(abs_lm)[["(Intercept)"]])+
+  theme(legend.position="none")
+
+mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age)
+
+abs_lm2 <- lm(abs_mean_delta_cent ~ study_mean_age + study_n + abs(age_diff), data=mean_cent_deltas_df_study)
+summary(abs_lm2)
+
+mean_cent_deltas_df_study %>%
+  ggplot(aes(x=abs(age_diff), y=abs_mean_delta_cent)) +
+  geom_point(alpha=0.05) +
+  geom_smooth(method="lm")
+```
+
+*options: rerun from the beginning w/o UKB?
+Or just control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting?
+
+###Global Features
+Look at global features
+
+```{r}
+glob_cent_deltas_df <- cent_deltas_df %>%
+  dplyr::select(study, age_years, sex, participant, fs_version, paste0("delta.", vol_list_global)) %>%
+  mutate(across(paste0("delta.", vol_list_global), abs, .names = "abs_{.col}"))
+
+glob_cent_deltas_df_long <- glob_cent_deltas_df %>%
+  pivot_longer(cols=starts_with("abs"), names_to="pheno", values_to = "abs_delta") %>%
+  mutate(pheno=gsub("abs_delta.", "", pheno))
+
+glob_cent_deltas_df_long %>%
+  ggplot() +
+  geom_boxplot(aes(y=abs_delta, color=pheno), outlier.shape = NA) + 
+scale_y_continuous(limits = quantile(glob_cent_deltas_df_long$abs_delta, c(0.1, 0.9)))
+```
+```{r}
+glob_cent_deltas_df_study <- full_join(glob_cent_deltas_df, study_info)
+
+abs_gmv_lm <- lm(abs_delta.GMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
+summary(abs_gmv_lm)
+
+glob_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_delta.GMV)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(abs_gmv_lm)[["study_mean_age"]], 
+              intercept = coef(abs_gmv_lm)[["(Intercept)"]])+
+  theme(legend.position="none") +
+  scale_y_continuous(limits = c(0, 0.1))
+
+abs_sgmv_lm <- lm(abs_delta.sGMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
+summary(abs_sgmv_lm)
+glob_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_delta.sGMV)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(abs_sgmv_lm)[["study_mean_age"]], 
+              intercept = coef(abs_sgmv_lm)[["(Intercept)"]])+
+  theme(legend.position="none")+
+  scale_y_continuous(limits = c(0, 0.1))
+
+abs_wmv_lm <- lm(abs_delta.WMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
+
+summary(gam(abs_delta.WMV ~ pb(study_mean_age) + study_n, data=glob_cent_deltas_df_study))
+
+summary(abs_wmv_lm)
+glob_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_delta.WMV)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(abs_wmv_lm)[["study_mean_age"]], 
+              intercept = coef(abs_wmv_lm)[["(Intercept)"]])+
+  theme(legend.position="none")+
+  scale_y_continuous(limits = c(0, 0.1))
+
+abs_ven_lm <- lm(abs_delta.Ventricles ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
+summary(abs_ven_lm)
+glob_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_delta.Ventricles)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(abs_ven_lm)[["study_mean_age"]], 
+              intercept = coef(abs_ven_lm)[["(Intercept)"]])+
+  theme(legend.position="none")+
+  scale_y_continuous(limits = c(0, 0.1))
+
+glob_cent_deltas_df_long <- full_join(glob_cent_deltas_df_long, study_info)
+
+glob_cent_deltas_df_long <- glob_cent_deltas_df_long %>%
+  mutate(age_diff=age_years-study_mean_age)
+
+glob_cent_deltas_df_long %>%
+  ggplot(aes(x=study_mean_age, y=abs_delta, color=pheno)) +
+  geom_point(alpha=0.4) +
+  geom_smooth(method="lm") +
+  facet_wrap(~pheno, scales="free_y")
+```
+
+####Compare w/ Nature Paper
+Compare to Nature paper age effects in sigma
+
+```{r}
+#gmv
+paper_fit_gmv$param$sigma$fixef
+gmv_b1 <- unname(paper_fit_gmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 3))1"])
+gmv_b2 <- unname(paper_fit_gmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 3))2"])
+
+#formula
+gmv_age_sig <- parse(text = paste0(gmv_b1, "*x^(-2) + ", gmv_b2,"*x^(3)"))
+gmv_age_sig_der <- D(gmv_age_sig,"x")
+
+#sgmv
+paper_fit_sgmv$param$sigma$fixef
+sgmv_b1 <- unname(paper_fit_sgmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2))1"])
+sgmv_b2 <- unname(paper_fit_sgmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2))2"])
+
+sgmv_age_sig <- parse(text = paste0(sgmv_b1, "*x^(-2) + ", sgmv_b2,"*x^(-2)*log(x)"))
+sgmv_age_sig_der <- D(sgmv_age_sig,"x")
+
+#wmv
+paper_fit_wmv$param$sigma$fixef
+wmv_b1 <- unname(paper_fit_wmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 2))1"])
+wmv_b2 <- unname(paper_fit_wmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 2))2"])
+
+wmv_age_sig <- parse(text = paste0(wmv_b1, "*x^(-2) + ", wmv_b2,"*x^(s)"))
+wmv_age_sig_der <- D(wmv_age_sig,"x")
+
+#ven
+paper_fit_ven$param$sigma$fixef
+ven_b1 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))1"])
+ven_b2 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))2"])
+ven_b3 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))3"])
+
+ven_age_sig <- parse(text = paste0(ven_b1, "*x^(-2) + ", ven_b2,"*x^(-2)*log(x) + ", ven_b3, "*x^(-2)*log(x)^(2)"))
+ven_age_sig_der <- D(ven_age_sig,"x")
+```
+try summing abs(betas) within each pheno to get overall magnitude of age effects
+
+```{r}
+paper_sigma_age <- data.frame("pheno" = c("GMV", "sGMV", "WMV", "Ventricles"),
+                              "paper_abs_betas_sum" = c(sum(abs(gmv_b1), abs(gmv_b2)), 
+                                                  sum(abs(sgmv_b1), abs(sgmv_b2)),
+                                                  sum(abs(wmv_b1), abs(wmv_b2)),
+                                                  sum(abs(ven_b1), abs(ven_b2), abs(ven_b3))),
+                              "paper_betas_sum" = c(sum(gmv_b1, gmv_b2), 
+                                                  sum(sgmv_b1, sgmv_b2),
+                                                  sum(wmv_b1, wmv_b2),
+                                                  sum(ven_b1, ven_b2, ven_b3)),
+                              "lm_beta" = c(coef(abs_gmv_lm)[["study_mean_age"]],
+                                            coef(abs_sgmv_lm)[["study_mean_age"]],
+                                            coef(abs_wmv_lm)[["study_mean_age"]],
+                                            coef(abs_ven_lm)[["study_mean_age"]]),
+                              "quick_var_range" = c((38.85552-16.93754), (0.2380905-0.1447285), (29.34474-13.4674), (17.63019-0.3520969))
+                              )
+paper_sigma_age %>%
+  ggplot() +
+  geom_col(aes(x=pheno, y=paper_abs_betas_sum, fill=pheno))
+
+paper_sigma_age %>%
+  ggplot() +
+  geom_col(aes(x=pheno, y=paper_betas_sum, fill=pheno))
+
+paper_sigma_age %>%
+  ggplot() +
+  geom_point(aes(x=abs(lm_beta), y=paper_betas_sum, color=pheno))
+
+paper_sigma_age %>%
+  ggplot() +
+  geom_point(aes(x=abs(lm_beta), y=quick_var_range, color=pheno))
+```
+
+###W/In Cortical Features
+
+Looking at averages w/in each cortical feature since they have different trajectories of variance across the lifespan (see Nature paper)
+
+####Regional Volumes
+
+```{r}
+#get subject-level means
+mean_vol_cent_deltas_df <- cent_deltas_df %>%
+  pivot_longer(cols=any_of(paste0("delta.", vol_list_regions)), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_delta_vol_cent=(mean(centile))*100, #convert units to centiles space
+            abs_mean_delta_vol_cent=(mean(abs(centile)))*100) %>%
+  dplyr::select(!(starts_with("delta")))
+
+mean_vol_cent_deltas_df_study <- full_join(mean_vol_cent_deltas_df, study_info)
+
+abs_vol_lm <- lm(abs_mean_delta_vol_cent ~ study_mean_age + study_n, data=mean_vol_cent_deltas_df_study)
+summary(abs_vol_lm)
+
+mean_vol_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_mean_delta_vol_cent)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(abs_vol_lm)[["study_mean_age"]], 
+              intercept = coef(abs_vol_lm)[["(Intercept)"]])+
+  theme(legend.position="none")
+```
+
+####Regional Thickness
+
+```{r}
+#get subject-level means
+mean_ct_cent_deltas_df <- cent_deltas_df %>%
+  pivot_longer(cols=any_of(paste0("delta.", ct_list)), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_delta_ct_cent=(mean(centile))*100, #convert units to centiles space
+            abs_mean_delta_ct_cent=(mean(abs(centile)))*100) %>%
+  dplyr::select(!(starts_with("delta")))
+
+mean_ct_cent_deltas_df_study <- full_join(mean_ct_cent_deltas_df, study_info)
+
+abs_ct_lm <- lm(abs_mean_delta_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_deltas_df_study)
+summary(abs_ct_lm)
+
+mean_ct_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_mean_delta_ct_cent)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(abs_ct_lm)[["study_mean_age"]], 
+              intercept = coef(abs_ct_lm)[["(Intercept)"]])+
+  theme(legend.position="none")
+```
+
+much bigger than anticipated based on Nature paper!!? is combatLS overfitting?
+
+```{r}
+#get subject-level means
+mean_ct_cent_gamlss_df <- gamlss_centile_df.clean %>%
+  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_ct_cent=(mean(centile))*100)
+```
+
+```{r}
+mean_ct_cent_gamlss_df <- full_join(mean_ct_cent_gamlss_df, study_info)
+summary(lm(mean_ct_cent ~ study_mean_age, data=mean_ct_cent_gamlss_df))
+summary(lm(mean_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_gamlss_df))
+
+# mean_ct_cent_gamlss_df %>%
+#   ggplot(aes(x=study_mean_age, y=mean_ct_cent)) +
+#   geom_point(aes(color=study), alpha=0.2) +
+#   geom_smooth(method = "lm") +
+#   theme(legend.position="none")
+
+#get subject-level means
+mean_ct_cent_gam_df <- gam_centile_df.clean %>%
+  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_ct_cent=(mean(centile))*100)
+
+mean_ct_cent_gam_df <- full_join(mean_ct_cent_gam_df, study_info)
+summary(lm(mean_ct_cent ~ study_mean_age, data=mean_ct_cent_gam_df))
+summary(lm(mean_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_gam_df))
+
+# mean_ct_cent_gam_df %>%
+#   ggplot(aes(x=study_mean_age, y=mean_ct_cent)) +
+#   geom_point(aes(color=study), alpha=0.2) +
+#   geom_smooth(method = "lm") +
+#   theme(legend.position="none")
+```
+
+####Regional SA
+
+```{r}
+#get subject-level means
+mean_sa_cent_deltas_df <- cent_deltas_df %>%
+  pivot_longer(cols=any_of(paste0("delta.", sa_list)), names_to="pheno", values_to = "centile") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_delta_sa_cent=(mean(centile))*100, #convert units to centiles space
+            abs_mean_delta_sa_cent=(mean(abs(centile)))*100) %>%
+  dplyr::select(!(starts_with("delta")))
+
+mean_sa_cent_deltas_df_study <- full_join(mean_sa_cent_deltas_df, study_info)
+
+abs_sa_lm <- lm(abs_mean_delta_sa_cent ~ study_mean_age + study_n, data=mean_sa_cent_deltas_df_study)
+summary(abs_sa_lm)
+
+mean_sa_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_mean_delta_sa_cent)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_abline(slope = coef(abs_sa_lm)[["study_mean_age"]], 
+              intercept = coef(abs_sa_lm)[["(Intercept)"]])+
+  theme(legend.position="none")
+```
+
+####Compare
+
+```{r}
+df_list.tmp <- list(mean_sa_cent_deltas_df_study, mean_ct_cent_deltas_df_study, mean_vol_cent_deltas_df_study)
+all_reg_deltas <- df_list.tmp %>% reduce(full_join)
+```
+
+```{r}
+all_reg_deltas_long <- all_reg_deltas %>%
+  pivot_longer(cols=starts_with("abs_mean_delta"), names_to="pheno_type", values_to = "abs_cent_delta") %>%
+  mutate(pheno_type=gsub("abs_mean_delta_|_cent", "", pheno_type))
+
+all_reg_deltas_long %>%
+  ggplot() +
+  geom_boxplot(aes(y=abs_cent_delta, fill=pheno_type))
 ```
@@
 ```
+
+#ComBatted Data
+
+```{r}
+gam_data_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gam_batch.study_data.csv"))
+gam_data_df <- gam_data_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
+
+gamlss_data_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gamlss_batch.study_data.csv"))
+gamlss_data_df <- gamlss_data_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
+```
+
+```{r}
+gam_data_df %>%
+  ggplot() +
+  geom_point(aes(x=age_days, y=rh_CT_medialorbitofrontal, color=study), alpha=0.4) +
+  theme(legend.position="none")
+
+gamlss_data_df %>%
+  ggplot() +
+  geom_point(aes(x=age_days, y=rh_CT_medialorbitofrontal, color=study), alpha=0.4) +
+  theme(legend.position="none")
+```
+
+```{r}
+#get subject-level means
+gam_data_ct_avg <- gam_data_df %>%
+  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "val") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_ct=mean(val)) #convert units to centiles space
+
+gamlss_data_ct_avg <- gamlss_data_df %>%
+  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "val") %>%
+  group_by(participant) %>%
+  summarize(study = unique(study),
+            site = unique(site),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_ct=mean(val)) #convert units to centiles space
+```
+
+```{r}
+gam_data_ct_avg %>%
+  ggplot() +
+  geom_point(aes(x=age_days, y=mean_ct, color=study), alpha=0.4) +
+  theme(legend.position="none")
+
+gamlss_data_ct_avg %>%
+  ggplot() +
+  geom_point(aes(x=age_days, y=mean_ct, color=study), alpha=0.4) +
+  theme(legend.position="none")
+```
+
+```{r}
+gam_data_ct_avg$combat <- "gam"
+gamlss_data_ct_avg$combat <- "gamlss"
+
+df_list.tmp <- list(gam_data_ct_avg, gamlss_data_ct_avg, study_info)
+ct_avg_data <- df_list.tmp %>% reduce(full_join)
+```
+
+```{r}
+ct_avg_data %>%
+  ggplot() +
+  geom_point(aes(x=age_days, y=mean_ct, color=combat), alpha=0.4)
+
+ct_avg_data %>%
+  ggplot() +
+  geom_point(aes(x=study_mean_age, y=mean_ct, color=combat), alpha=0.2)
+
+ct_avg_data %>%
+  ggplot() +
+  geom_point(aes(x=study_mean_age, y=mean_ct, color=study), alpha=0.4) +
+  facet_wrap(~combat) +
+  theme(legend.position="none")
+
+```

### 2024-04-19 paper_replication_results.Rmd
cleaned up
lbcc looking w/in feature

@@
 
-mean(cf.gam_worse$count)
+median(cf.gam_worse$count)
 IQR(cf.gam_worse$count)
@@
   labs(y="Female Bias", x = "Dataset")
+
+raise <- all_cent_sum.df_scaled %>%
+  dplyr::filter(dataset != "raw" & dataset != "cf" & source != "ext") %>%
+  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=female_ratio_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+  # geom_dotplot(binaxis = "y", stackdir="center") +
+  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
+  geom_hline(yintercept=0, linetype="dashed")+
+  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  #             data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  theme_linedraw() +
+  scale_x_discrete(labels=dataset_names) +
+  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Female Bias", x = "Dataset")
+
+ggsave(filename="figs/raise.png", raise)
 
@@
 
+```{r}
+sex.var.plot.global <- readRDS("figs/sex_var_plot_global.RDS")
+sex.var.plot <- readRDS("figs/sex_var_plot.RDS")
+
+```
+
 Figure 3 but with sig. tests for boxplots

### 2024-04-19 paper_lbcc_results.Rmd
removing w/in feature assessments, cleaning up

@@
 
-*options: rerun from the beginning w/o UKB?
-Or just control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting?
-
-###Global Features
-Look at global features
-
-```{r}
-glob_cent_deltas_df <- cent_deltas_df %>%
-  dplyr::select(study, age_years, sex, participant, fs_version, paste0("delta.", vol_list_global)) %>%
-  mutate(across(paste0("delta.", vol_list_global), abs, .names = "abs_{.col}"))
-
-glob_cent_deltas_df_long <- glob_cent_deltas_df %>%
-  pivot_longer(cols=starts_with("abs"), names_to="pheno", values_to = "abs_delta") %>%
-  mutate(pheno=gsub("abs_delta.", "", pheno))
-
-glob_cent_deltas_df_long %>%
-  ggplot() +
-  geom_boxplot(aes(y=abs_delta, color=pheno), outlier.shape = NA) + 
-scale_y_continuous(limits = quantile(glob_cent_deltas_df_long$abs_delta, c(0.1, 0.9)))
-```
-```{r}
-glob_cent_deltas_df_study <- full_join(glob_cent_deltas_df, study_info)
-
-abs_gmv_lm <- lm(abs_delta.GMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
-summary(abs_gmv_lm)
-
-glob_cent_deltas_df_study %>%
-  ggplot(aes(x=study_mean_age, y=abs_delta.GMV)) +
-  geom_point(aes(color=study), alpha=0.2) +
-  geom_abline(slope = coef(abs_gmv_lm)[["study_mean_age"]], 
-              intercept = coef(abs_gmv_lm)[["(Intercept)"]])+
-  theme(legend.position="none") +
-  scale_y_continuous(limits = c(0, 0.1))
-
-abs_sgmv_lm <- lm(abs_delta.sGMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
-summary(abs_sgmv_lm)
-glob_cent_deltas_df_study %>%
-  ggplot(aes(x=study_mean_age, y=abs_delta.sGMV)) +
-  geom_point(aes(color=study), alpha=0.2) +
-  geom_abline(slope = coef(abs_sgmv_lm)[["study_mean_age"]], 
-              intercept = coef(abs_sgmv_lm)[["(Intercept)"]])+
-  theme(legend.position="none")+
-  scale_y_continuous(limits = c(0, 0.1))
-
-abs_wmv_lm <- lm(abs_delta.WMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
-
-summary(gam(abs_delta.WMV ~ pb(study_mean_age) + study_n, data=glob_cent_deltas_df_study))
-
-summary(abs_wmv_lm)
-glob_cent_deltas_df_study %>%
-  ggplot(aes(x=study_mean_age, y=abs_delta.WMV)) +
-  geom_point(aes(color=study), alpha=0.2) +
-  geom_abline(slope = coef(abs_wmv_lm)[["study_mean_age"]], 
-              intercept = coef(abs_wmv_lm)[["(Intercept)"]])+
-  theme(legend.position="none")+
-  scale_y_continuous(limits = c(0, 0.1))
-
-abs_ven_lm <- lm(abs_delta.Ventricles ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
-summary(abs_ven_lm)
-glob_cent_deltas_df_study %>%
-  ggplot(aes(x=study_mean_age, y=abs_delta.Ventricles)) +
-  geom_point(aes(color=study), alpha=0.2) +
-  geom_abline(slope = coef(abs_ven_lm)[["study_mean_age"]], 
-              intercept = coef(abs_ven_lm)[["(Intercept)"]])+
-  theme(legend.position="none")+
-  scale_y_continuous(limits = c(0, 0.1))
-
-glob_cent_deltas_df_long <- full_join(glob_cent_deltas_df_long, study_info)
-
-glob_cent_deltas_df_long <- glob_cent_deltas_df_long %>%
-  mutate(age_diff=age_years-study_mean_age)
-
-glob_cent_deltas_df_long %>%
-  ggplot(aes(x=study_mean_age, y=abs_delta, color=pheno)) +
-  geom_point(alpha=0.4) +
-  geom_smooth(method="lm") +
-  facet_wrap(~pheno, scales="free_y")
-```
-
-####Compare w/ Nature Paper
-Compare to Nature paper age effects in sigma
-
-```{r}
-#gmv
-paper_fit_gmv$param$sigma$fixef
-gmv_b1 <- unname(paper_fit_gmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 3))1"])
-gmv_b2 <- unname(paper_fit_gmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 3))2"])
-
-#formula
-gmv_age_sig <- parse(text = paste0(gmv_b1, "*x^(-2) + ", gmv_b2,"*x^(3)"))
-gmv_age_sig_der <- D(gmv_age_sig,"x")
-
-#sgmv
-paper_fit_sgmv$param$sigma$fixef
-sgmv_b1 <- unname(paper_fit_sgmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2))1"])
-sgmv_b2 <- unname(paper_fit_sgmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2))2"])
-
-sgmv_age_sig <- parse(text = paste0(sgmv_b1, "*x^(-2) + ", sgmv_b2,"*x^(-2)*log(x)"))
-sgmv_age_sig_der <- D(sgmv_age_sig,"x")
-
-#wmv
-paper_fit_wmv$param$sigma$fixef
-wmv_b1 <- unname(paper_fit_wmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 2))1"])
-wmv_b2 <- unname(paper_fit_wmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 2))2"])
-
-wmv_age_sig <- parse(text = paste0(wmv_b1, "*x^(-2) + ", wmv_b2,"*x^(s)"))
-wmv_age_sig_der <- D(wmv_age_sig,"x")
-
-#ven
-paper_fit_ven$param$sigma$fixef
-ven_b1 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))1"])
-ven_b2 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))2"])
-ven_b3 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))3"])
-
-ven_age_sig <- parse(text = paste0(ven_b1, "*x^(-2) + ", ven_b2,"*x^(-2)*log(x) + ", ven_b3, "*x^(-2)*log(x)^(2)"))
-ven_age_sig_der <- D(ven_age_sig,"x")
-```
-try summing abs(betas) within each pheno to get overall magnitude of age effects
-
-```{r}
-paper_sigma_age <- data.frame("pheno" = c("GMV", "sGMV", "WMV", "Ventricles"),
-                              "paper_abs_betas_sum" = c(sum(abs(gmv_b1), abs(gmv_b2)), 
-                                                  sum(abs(sgmv_b1), abs(sgmv_b2)),
-                                                  sum(abs(wmv_b1), abs(wmv_b2)),
-                                                  sum(abs(ven_b1), abs(ven_b2), abs(ven_b3))),
-                              "paper_betas_sum" = c(sum(gmv_b1, gmv_b2), 
-                                                  sum(sgmv_b1, sgmv_b2),
-                                                  sum(wmv_b1, wmv_b2),
-                                                  sum(ven_b1, ven_b2, ven_b3)),
-                              "lm_beta" = c(coef(abs_gmv_lm)[["study_mean_age"]],
-                                            coef(abs_sgmv_lm)[["study_mean_age"]],
-                                            coef(abs_wmv_lm)[["study_mean_age"]],
-                                            coef(abs_ven_lm)[["study_mean_age"]]),
-                              "quick_var_range" = c((38.85552-16.93754), (0.2380905-0.1447285), (29.34474-13.4674), (17.63019-0.3520969))
-                              )
-paper_sigma_age %>%
-  ggplot() +
-  geom_col(aes(x=pheno, y=paper_abs_betas_sum, fill=pheno))
-
-paper_sigma_age %>%
-  ggplot() +
-  geom_col(aes(x=pheno, y=paper_betas_sum, fill=pheno))
-
-paper_sigma_age %>%
-  ggplot() +
-  geom_point(aes(x=abs(lm_beta), y=paper_betas_sum, color=pheno))
-
-paper_sigma_age %>%
-  ggplot() +
-  geom_point(aes(x=abs(lm_beta), y=quick_var_range, color=pheno))
-```
-
-###W/In Cortical Features
-
-Looking at averages w/in each cortical feature since they have different trajectories of variance across the lifespan (see Nature paper)
-
-####Regional Volumes
-
-```{r}
-#get subject-level means
-mean_vol_cent_deltas_df <- cent_deltas_df %>%
-  pivot_longer(cols=any_of(paste0("delta.", vol_list_regions)), names_to="pheno", values_to = "centile") %>%
-  group_by(participant) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_delta_vol_cent=(mean(centile))*100, #convert units to centiles space
-            abs_mean_delta_vol_cent=(mean(abs(centile)))*100) %>%
-  dplyr::select(!(starts_with("delta")))
-
-mean_vol_cent_deltas_df_study <- full_join(mean_vol_cent_deltas_df, study_info)
-
-abs_vol_lm <- lm(abs_mean_delta_vol_cent ~ study_mean_age + study_n, data=mean_vol_cent_deltas_df_study)
-summary(abs_vol_lm)
-
-mean_vol_cent_deltas_df_study %>%
-  ggplot(aes(x=study_mean_age, y=abs_mean_delta_vol_cent)) +
-  geom_point(aes(color=study), alpha=0.2) +
-  geom_abline(slope = coef(abs_vol_lm)[["study_mean_age"]], 
-              intercept = coef(abs_vol_lm)[["(Intercept)"]])+
-  theme(legend.position="none")
-```
-
-####Regional Thickness
-
-```{r}
-#get subject-level means
-mean_ct_cent_deltas_df <- cent_deltas_df %>%
-  pivot_longer(cols=any_of(paste0("delta.", ct_list)), names_to="pheno", values_to = "centile") %>%
-  group_by(participant) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_delta_ct_cent=(mean(centile))*100, #convert units to centiles space
-            abs_mean_delta_ct_cent=(mean(abs(centile)))*100) %>%
-  dplyr::select(!(starts_with("delta")))
-
-mean_ct_cent_deltas_df_study <- full_join(mean_ct_cent_deltas_df, study_info)
-
-abs_ct_lm <- lm(abs_mean_delta_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_deltas_df_study)
-summary(abs_ct_lm)
-
-mean_ct_cent_deltas_df_study %>%
-  ggplot(aes(x=study_mean_age, y=abs_mean_delta_ct_cent)) +
-  geom_point(aes(color=study), alpha=0.2) +
-  geom_abline(slope = coef(abs_ct_lm)[["study_mean_age"]], 
-              intercept = coef(abs_ct_lm)[["(Intercept)"]])+
-  theme(legend.position="none")
-```
-
-much bigger than anticipated based on Nature paper!!? is combatLS overfitting?
-
-```{r}
-#get subject-level means
-mean_ct_cent_gamlss_df <- gamlss_centile_df.clean %>%
-  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "centile") %>%
-  group_by(participant) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_ct_cent=(mean(centile))*100)
-```
-
-```{r}
-mean_ct_cent_gamlss_df <- full_join(mean_ct_cent_gamlss_df, study_info)
-summary(lm(mean_ct_cent ~ study_mean_age, data=mean_ct_cent_gamlss_df))
-summary(lm(mean_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_gamlss_df))
-
-# mean_ct_cent_gamlss_df %>%
-#   ggplot(aes(x=study_mean_age, y=mean_ct_cent)) +
-#   geom_point(aes(color=study), alpha=0.2) +
-#   geom_smooth(method = "lm") +
-#   theme(legend.position="none")
-
-#get subject-level means
-mean_ct_cent_gam_df <- gam_centile_df.clean %>%
-  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "centile") %>%
-  group_by(participant) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_ct_cent=(mean(centile))*100)
-
-mean_ct_cent_gam_df <- full_join(mean_ct_cent_gam_df, study_info)
-summary(lm(mean_ct_cent ~ study_mean_age, data=mean_ct_cent_gam_df))
-summary(lm(mean_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_gam_df))
-
-# mean_ct_cent_gam_df %>%
-#   ggplot(aes(x=study_mean_age, y=mean_ct_cent)) +
-#   geom_point(aes(color=study), alpha=0.2) +
-#   geom_smooth(method = "lm") +
-#   theme(legend.position="none")
-```
-
-####Regional SA
-
-```{r}
-#get subject-level means
-mean_sa_cent_deltas_df <- cent_deltas_df %>%
-  pivot_longer(cols=any_of(paste0("delta.", sa_list)), names_to="pheno", values_to = "centile") %>%
-  group_by(participant) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_delta_sa_cent=(mean(centile))*100, #convert units to centiles space
-            abs_mean_delta_sa_cent=(mean(abs(centile)))*100) %>%
-  dplyr::select(!(starts_with("delta")))
-
-mean_sa_cent_deltas_df_study <- full_join(mean_sa_cent_deltas_df, study_info)
-
-abs_sa_lm <- lm(abs_mean_delta_sa_cent ~ study_mean_age + study_n, data=mean_sa_cent_deltas_df_study)
-summary(abs_sa_lm)
-
-mean_sa_cent_deltas_df_study %>%
-  ggplot(aes(x=study_mean_age, y=abs_mean_delta_sa_cent)) +
-  geom_point(aes(color=study), alpha=0.2) +
-  geom_abline(slope = coef(abs_sa_lm)[["study_mean_age"]], 
-              intercept = coef(abs_sa_lm)[["(Intercept)"]])+
-  theme(legend.position="none")
-```
-
-####Compare
-
-```{r}
-df_list.tmp <- list(mean_sa_cent_deltas_df_study, mean_ct_cent_deltas_df_study, mean_vol_cent_deltas_df_study)
-all_reg_deltas <- df_list.tmp %>% reduce(full_join)
-```
-
-```{r}
-all_reg_deltas_long <- all_reg_deltas %>%
-  pivot_longer(cols=starts_with("abs_mean_delta"), names_to="pheno_type", values_to = "abs_cent_delta") %>%
-  mutate(pheno_type=gsub("abs_mean_delta_|_cent", "", pheno_type))
-
-all_reg_deltas_long %>%
-  ggplot() +
-  geom_boxplot(aes(y=abs_cent_delta, fill=pheno_type))
-```
-
-```{r}
-delta_sum.df_merged %>%
-  ggplot(aes(x=study_mean_age_yrs, y=study_mean_delta)) +
-  geom_count(aes(size=study_n)) +
-  geom_smooth(method="lm")
-delta_sum.df_merged %>%
-  ggplot(aes(x=study_mean_age_yrs, y=study_mean_abs.delta)) +
-  geom_count(aes(size=study_n)) +
-  geom_smooth(method="lm")
-
-age_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
-age_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]
-
-age_effects.plt <- delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
-  geom_count(aes(size=site_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
-  geom_smooth(method="lm", color="#2A5783FF") +
-  theme_linedraw(base_size = 11) +
-  annotate("text", x=60, y=4, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
-  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
-  theme(axis.title.x=element_blank())
-
-age_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
-age_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]
-
-age_effects.plt.null <- delta_sum.df_merged %>%
-  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_delta*100))) +
-  geom_count(aes(size=site_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
-  geom_smooth(method="lm", color="#2A5783FF") +
-  theme_linedraw(base_size = 11) +
-  annotate("text", x=60, y=-0.3, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
-  labs(size="site N", x = "Mean Age (Years)", y = "Mean Difference in Centile Scores")
-
-#add prediction to df
-delta_sum.df_merged$abs.delta_pred <- predict(full.delta.age.fit)
-
-full.delta.age.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$x,
-                                    predicted=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$predicted,
-                                    conf.low=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.low,
-                                    conf.high=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.high)
-full.delta.age.fit.df <- full.delta.age.fit.df %>%
-  mutate(predicted = predicted*100,
-         conf.low=conf.low*100,
-         conf.high=conf.high*100)
-
-marg.age_effects.plt <- ggplot() +
-  geom_count(data=delta_sum.df_merged, aes(x=site_mean_age_yrs, y=(abs.delta_pred*100), size=site_n, color=sex_prop_Female)) +
-  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
-  geom_line(aes(x=full.delta.age.fit.df$x, y = full.delta.age.fit.df$predicted), size = 0.7, color="#2A5783FF") +
-  geom_ribbon(aes(x=full.delta.age.fit.df$x,y= NULL, ymin= full.delta.age.fit.df$conf.low, ymax= full.delta.age.fit.df$conf.high), alpha=0.1, fill="#2A5783FF") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N", x = "Mean Age (Years)", y = "Predicted Mean Absolute Difference in Centile Scores")
-```
-
-are studies w older subjects changing more because there are fewer of them?
-check for association between study size and age
-```{r}
-summary(lm(study_mean_age_yrs ~ study_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
-```
-
-```{r}
-cent_deltas_df %>%
-  ggplot() +
-  geom_histogram(aes(x=age_days)) +
-  labs(title="subject distribution across age")
-
-delta_sum.df_merged %>%
-  ggplot() +
-  geom_histogram(aes(x=study_mean_age_yrs)) +
-  labs(title="study distribution across age")
-```
-
-##Prop Female
-
-```{r}
-ggplot(delta_sum.df_merged) +
-  geom_histogram(aes(x=sex_prop_Female))
-```
-
-ComBatGAM mean centile ~ % female
-
-```{r}
-gam_centile_study.mean.df <- gam_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(study) %>%
-  summarize(mean_gam_cent=mean(centile),
-            mean_gam_abs_cent=mean(abs(centile)))
-
-gam_centile_study.mean.df2 <- left_join(gam_centile_study.mean.df, delta_sum.df_merged)
-
-gam.prop.f.test <- lm(mean_gam_cent ~ sex_prop_Female + study_n + study_mean_age, data=gam_centile_study.mean.df2)
-summary(gam.prop.f.test)
-gam.prop.f.test2 <- lm(mean_gam_abs_cent ~ sex_prop_Female + study_n + study_mean_age, data=gam_centile_study.mean.df2)
-summary(gam.prop.f.test2)
-
-ggplot(gam_centile_study.mean.df2, aes(x=sex_prop_Female, y=mean_gam_cent)) +
-  geom_point() +
-  geom_smooth(method="lm")
-```
-
-ComBatLS mean centile ~ % female
-
-```{r}
-gamlss_centile_study.mean.df <- gamlss_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(study) %>%
-  summarize(mean_gamlss_cent=mean(centile),
-            mean_gamlss_abs_cent=mean(abs(centile)))
-
-gamlss_centile_study.mean.df2 <- left_join(gamlss_centile_study.mean.df, delta_sum.df_merged)
-
-gamlss.prop.f.test <- lm(mean_gamlss_cent ~ sex_prop_Female + study_n + study_mean_age, data=gamlss_centile_study.mean.df2)
-summary(gamlss.prop.f.test)
-gamlss.prop.f.test2 <- lm(mean_gamlss_abs_cent ~ sex_prop_Female + study_n + study_mean_age, data=gamlss_centile_study.mean.df2)
-summary(gamlss.prop.f.test2)
-
-ggplot(gamlss_centile_study.mean.df2, aes(x=sex_prop_Female, y=mean_gamlss_cent)) +
-  geom_point() +
-  geom_smooth(method="lm")
-```
-
-```{r}
-gamlss_centile_site.mean.df3 <- gamlss_centile_site.mean.df2 %>%
-  mutate(combat = "gamlss") %>%
-  rename(mean_cent = mean_gamlss_cent)
-gam_centile_site.mean.df3 <- gam_centile_site.mean.df2 %>%
-  mutate(combat = "gam") %>%
-  rename(mean_cent = mean_gam_cent)
-
-site_lvl_mean_cents <- full_join(gamlss_centile_site.mean.df3, gam_centile_site.mean.df3)
-
-site_lvl_mean_cents %>%
-  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=combat, fill=combat)) +
-  geom_point(alpha=0.4) +
-  geom_smooth(method="lm") +
-  ylab("within-site mean centile")
-```
-
-```{r}
-gamlss_centile_df.mean %>%
-  ggplot() + 
-  geom_violin(aes(x=sex, y=mean_cent))
-gamlss_centile_df.mean$combat <- "gamlss"
-
-gam_centile_df.mean %>%
-  ggplot() + 
-  geom_violin(aes(x=sex, y=mean_cent))
-gam_centile_df.mean$combat <- "gam"
-
-subj.mean.cents <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)
-
-subj.mean.cents %>%
-  ggplot() + 
-  geom_boxplot(aes(x=sex, y=mean_cent, fill=combat))
-
-#add sex prop
-site.info <- delta_sum.df_merged %>%
-  dplyr::select(site, sex_prop_Female, site_n)
-
-subj.mean.cents2 <- left_join(subj.mean.cents, site.info)
-subj.mean.cents2 %>%
-  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=sex)) +
-  geom_point(aes(size=site_n)) +
-  geom_smooth(method="lm")+
-  facet_wrap(~combat)
-
-subj.mean.cents2 %>%
-  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=combat, fill=combat)) +
-  # geom_point(alpha=0.4) +
-  geom_smooth(method="lm") +
-  ylab("within-subject mean centile")
-```
-
-delta ~ % female
-
-```{r}
-# %female
-summary(lm(site_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ sex_prop_Female, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ sex_prop_Female + site_n, data=delta_sum.df_merged)) #add site size covariate
-summary(lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged)) #add site size & age covariates
-full.delta.sex.fit <- lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged) #scaling for plot
-
-sjPlot::plot_model(full.delta.sex.fit, type = "pred", terms = "sex_prop_Female") 
-```
-
-```{r}
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
-  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
-  geom_line(aes(group=site))
-```
-
-```{r}
-delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
-  geom_count(aes(size=site_n)) +
-  geom_smooth(method="lm")
-delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta)) +
-  geom_count(aes(size=site_n)) +
-  geom_smooth(method="gam")
-
-sex_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
-sex_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]
-
-sex_effects.plt <- delta_sum.df_merged %>%
-  mutate(site_mean_delta=site_mean_delta*100) %>%
-  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
-  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
-  geom_smooth(method="lm",
-              color="#24693DFF") +
-  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores") +
-  theme(axis.title.y=element_blank()) +
-  annotate("text", x=0.25, y=-0.3, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
-  guides(size = "none")
-
-sex_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
-sex_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]
-
-sex_effects.plt.null <- delta_sum.df_merged %>%
-  ggplot(aes(x=sex_prop_Female, y=(site_mean_abs.delta*100))) +
-  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
-  geom_smooth(method="lm", color="#24693DFF") +
-  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N") +
-    annotate("text", x=0.25, y=4, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
-  theme(axis.title=element_blank()) +
-  guides(size = "none")
-
-#add prediction to df
-delta_sum.df_merged$delta_pred <- predict(full.delta.sex.fit)
-
-full.delta.sex.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$x,
-                                    predicted=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$predicted,
-                                    conf.low=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.low,
-                                    conf.high=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.high)
-full.delta.sex.fit.df <- full.delta.sex.fit.df %>%
-  mutate(predicted = predicted*100,
-         conf.low=conf.low*100,
-         conf.high=conf.high*100)
-
-marg.sex_effects.plt <- ggplot() +
-  geom_count(data=delta_sum.df_merged, aes(x=sex_prop_Female, y=(delta_pred*100), size=site_n, color=site_mean_age_yrs)) +
-  geom_line(aes(x=full.delta.sex.fit.df$x, y = full.delta.sex.fit.df$predicted), size = 0.7, color="#24693DFF") +
-  geom_ribbon(aes(x=full.delta.sex.fit.df$x,y= NULL, ymin= full.delta.sex.fit.df$conf.low, ymax= full.delta.sex.fit.df$conf.high), alpha=0.1, fill="#24693DFF") +
-  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
-  theme_linedraw(base_size = 11) +
-  labs(size="site N", x = "Proportion Female", y = "Predicted Mean Difference in Centile Scores")
-
-delta_sum.by.sex.df %>%
-  ggplot(aes(x=sex, y=(mean_delta*100))) +
-  geom_boxplot()
-```
-
-```{r}
-# age*%female
-summary(lm(site_mean_abs.delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
-summary(lm(site_mean_delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
-```
-
-```{r}
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
-  geom_point(alpha=0.4) +
-  geom_smooth(method="lm")
-
-cent_deltas_df %>%
-  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
-  geom_histogram(aes(y=after_stat(c(
-      count[group==1]/sum(count[group==1]),
-      count[group==2]/sum(count[group==2])
-    )*100)), alpha=0.4, position="identity", bins=60) +
-  ylab("Percentage")
-
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm")
-
-cent_deltas_df %>%
-  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
-  geom_point(alpha=0.4) +
-  geom_smooth(method="lm")
-```
+control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting?
 

### 2024-04-19 paper_lbcc_results.Rmd
fig 5 options

@@
   gamlss_centile_df %>%
-  mutate(ylab = "  Full \n  Sample") %>%
+  mutate(ylab = "   Full Sample") %>%
   ggplot(aes(y=ylab, x=age_years, color=sex)) +
@@
   theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
-  labs(color="Sex")
+  labs(fill="Sex")
 
@@
   theme(axis.ticks.y = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) +
-  labs(color="Sex", x = "Sex")
+  labs(fill="Sex", x = "Sex")
 
@@
 
-mean_cents.df %>%
-  ggplot() +
-  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
-  #geom_boxplot(aes(x=study, y=mean_cent)) +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5)) +
-  facet_wrap(~combat)
+dataset_names <- c(gam = "ComBat-GAM", gamlss = "ComBatLS", raw = "Unharmonized")
+
+mean_cents_bxplt <- mean_cents.df %>%
+  ggplot() +
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
+  theme_linedraw() +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none") +
+  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
+  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
+  labs(y="Mean Centile", x = "Study")
 ```
@@
           nrow=2, ncol=2, 
-          # align = "hv", 
+          #align = "h", 
           common.legend = TRUE,
           legend = "bottom",
-          widths=c(2, 1),
-          heights=c(14, 1)),
-# mean effects
-ggarrange(age_effects.plt,
-sex_effects.plt,
-nrow=2, ncol=1,
-align="hv",
-labels=c("B", "C")),
+          widths=c(1.7, 1),
+          heights=c(10, 1)),
+# boxplots
+mean_cents_bxplt,
 common.legend=FALSE,
 nrow=1, ncol=2,
-labels=c("A", NA),
-widths=c(1, 1.4)
+labels="AUTO",
+widths=c(1, 4)
 )
 
-ggsave("figs/figure_5.jpeg", fig5, width=12, height=10, units="in")
-```
-
-#ComBatted Data
-
-```{r}
-gam_data_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gam_batch.study_data.csv"))
-gam_data_df <- gam_data_df %>% mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25)
-
-gamlss_data_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gamlss_batch.study_data.csv"))
-gamlss_data_df <- gamlss_data_df %>% mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25)
-```
-
-```{r}
-gam_data_df %>%
-  ggplot() +
-  geom_point(aes(x=age_days, y=rh_CT_medialorbitofrontal, color=study), alpha=0.4) +
-  theme(legend.position="none")
-
-gamlss_data_df %>%
-  ggplot() +
-  geom_point(aes(x=age_days, y=rh_CT_medialorbitofrontal, color=study), alpha=0.4) +
-  theme(legend.position="none")
-```
-
-```{r}
-#get subject-level means
-gam_data_ct_avg <- gam_data_df %>%
-  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "val") %>%
-  group_by(participant) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_ct=mean(val)) #convert units to centiles space
-
-gamlss_data_ct_avg <- gamlss_data_df %>%
-  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "val") %>%
-  group_by(participant) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_ct=mean(val)) #convert units to centiles space
-```
-
-```{r}
-gam_data_ct_avg %>%
-  ggplot() +
-  geom_point(aes(x=age_days, y=mean_ct, color=study), alpha=0.4) +
-  theme(legend.position="none")
-
-gamlss_data_ct_avg %>%
-  ggplot() +
-  geom_point(aes(x=age_days, y=mean_ct, color=study), alpha=0.4) +
-  theme(legend.position="none")
-```
-
-```{r}
-gam_data_ct_avg$combat <- "gam"
-gamlss_data_ct_avg$combat <- "gamlss"
-
-df_list.tmp <- list(gam_data_ct_avg, gamlss_data_ct_avg, study_info)
-ct_avg_data <- df_list.tmp %>% reduce(full_join)
-```
-
-```{r}
-ct_avg_data %>%
-  ggplot() +
-  geom_point(aes(x=age_days, y=mean_ct, color=combat), alpha=0.4)
-
-ct_avg_data %>%
-  ggplot() +
-  geom_point(aes(x=study_mean_age, y=mean_ct, color=combat), alpha=0.2)
-
-ct_avg_data %>%
-  ggplot() +
-  geom_point(aes(x=study_mean_age, y=mean_ct, color=study), alpha=0.4) +
-  facet_wrap(~combat) +
-  theme(legend.position="none")
-
-```
+ggsave("figs/figure_5_horizontal.jpeg", fig5, width=18, height=9, units="in")
+```
+alternative vertical stacking
+
+```{r}
+blank <- ggplot() + theme_void()
+
+sample_boxplot2 <- gamlss_centile_df %>%
+  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
+ggplot(aes(y=age_years, x=study, color=sex)) +
+  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
+  theme_linedraw(base_size = 11) +
+  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(axis.title.x=element_blank(), axis.text.x = element_blank()) +
+  labs(color="Sex", y = "Age (Years) \n ")
+
+#overall sample
+full_boxplot2 <- 
+  gamlss_centile_df %>%
+  mutate(xlab = "Full Sample") %>%
+  ggplot(aes(x=xlab, y=age_years, color=sex)) +
+  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
+  theme_linedraw(base_size = 11) +
+  theme(axis.title=element_blank(), axis.text=element_blank(), axis.ticks.y=element_blank()) +
+  # scale_y_discrete(labels=c("1" = "Full Sample")) +
+  labs(color="Sex", x=" ")
+
+sample_barplot2 <- gamlss_centile_df %>%
+  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
+  ggplot() +
+  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
+  scale_y_continuous(labels = scales::percent) +
+  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
+  theme_linedraw(base_size = 11) +
+  scale_y_continuous(labels = scales::percent) +
+  #theme(axis.ticks.x = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
+  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
+  labs(fill="Sex", x = "Study", y="Sex")
+
+full_barplot2 <- 
+  gamlss_centile_df %>%
+  mutate(xlab = "       Full Sample ") %>%
+  ggplot() +
+  geom_bar(aes(y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=xlab, fill=sex)) + 
+  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
+  theme_linedraw(base_size = 11) +
+  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
+  labs(color="Sex", y = "Age (Years)", y="")
+
+ggarrange(sample_boxplot2, full_boxplot2,
+          sample_barplot2, full_barplot2,
+          nrow=2, ncol=2, 
+          align = "hv", 
+          common.legend = TRUE,
+          legend = "right",
+          widths=c(10, 1),
+          heights=c(2, 1)
+          )
+```
+
+```{r}
+# study demographics
+fig5_vert <- ggarrange(
+  ggarrange(blank,
+ggarrange(sample_boxplot2, full_boxplot2,
+          sample_barplot2, full_barplot2,
+          nrow=2, ncol=2, 
+          #align = "v", 
+          common.legend = TRUE,
+          legend = "right",
+          widths=c(18, 1),
+          heights=c(1, 1)
+          ),
+blank,
+nrow=1, ncol=3,
+common.legend = FALSE,
+widths= c(1, 20, 1)),
+# boxplots
+mean_cents_bxplt,
+common.legend=FALSE,
+nrow=2, ncol=1,
+heights=c(1, 1.5),
+labels="AUTO"
+)
+
+ggsave("figs/figure_5_vertical.jpeg", fig5_vert, width=15, height=9, units="in")
+```

### 2024-04-22 paper_lbcc_results.Rmd
figs

@@
 
-source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
-source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
+#source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
+#source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
 
@@
   theme_linedraw() +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none") +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none", strip.text.x = element_text(size = 12)) +
   scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
@@
 mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
-  mutate(age_diff = (age_days/365.25) - study_mean_age)
-
-abs_lm2 <- lm(abs_mean_delta_cent ~ study_mean_age + study_n + abs(age_diff), data=mean_cent_deltas_df_study)
+  mutate(age_diff = (age_days/365.25) - study_mean_age,
+         abs_age_diff = abs(age_diff))
+
+abs_lm2 <- lm(abs_mean_delta_cent ~ study_mean_age + study_n + abs_age_diff, data=mean_cent_deltas_df_study)
 summary(abs_lm2)
@@
 
-control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting?
+control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting
 
@@
 ```
+
+##Supplemental Figs
+
+mean abs delta ~ study mean age
+
+```{r}
+summary(abs_lm)
+
+#predict w CIs
+pdat <- with(mean_cent_deltas_df_study,
+             tibble(study_mean_age = seq(min(study_mean_age),
+                            max(study_mean_age),
+                            length = 200),
+             study_n = rep(median(study_n), length=200)))
+
+pred_abs_lm <- predict(abs_lm, pdat, se.fit=TRUE)
+
+pred_df <- pdat %>% 
+  bind_cols(data.frame(pred_abs_lm))
+
+pred_df <- pred_df %>%
+  mutate(upper = fit + (1.96 * se.fit), #95% CIs
+         lower = fit - (1.96 * se.fit))
+
+mean_cent_deltas_df_study %>%
+  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_ribbon(aes(ymin = lower, ymax = upper,
+                    x = study_mean_age), data = pred_df,
+                inherit.aes = FALSE, alpha = 0.2) +
+    geom_line(data = pred_df, aes(y = fit, x=study_mean_age))+
+  theme(legend.position="none") +
+  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
+```
+
+```{r}
+summary(abs_lm2)
+
+#predict w CIs
+pdat2 <- with(mean_cent_deltas_df_study,
+             tibble(abs_age_diff = seq(min(abs_age_diff),
+                            max(abs_age_diff),
+                            length = 200),
+             study_n = rep(median(study_n), length=200),
+             study_mean_age = rep(median(study_mean_age), length=200)))
+
+pred2 <- predict(abs_lm2, pdat2, se.fit = TRUE)
+
+pdat2 <- pdat2 %>%
+  bind_cols(data.frame(pred2)) %>%
+  mutate(upper = fit + (1.96 * se.fit),
+         lower = fit - (1.96 * se.fit))
+
+mean_cent_deltas_df_study %>%
+  ggplot(aes(x=abs_age_diff, y=abs_mean_delta_cent)) +
+  geom_point(alpha=0.05, aes(color=study)) +
+  geom_ribbon(aes(ymin = lower, ymax = upper,
+                    x = abs_age_diff), data = pdat2,
+                inherit.aes = FALSE, alpha = 0.2) +
+    geom_line(data = pdat2, aes(y = fit, x=abs_age_diff)) +
+  theme(legend.position="none") +
+  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
+```

### 2024-05-01 eda/lifespan_eda.Rmd
fixing study names

@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see)
+pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see, broom.mixed)
 
@@
 nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
-```
-Write out data for andrew
-```{r, eval=FALSE}
-andrew.df <- cn.df.unique %>%
-  dplyr::filter(study != "UKB" & age_days >= (365.25*18))
-nrow(andrew.df)
-fwrite(andrew.df, file=paste0(base_path, "/data/lifespan_CN_adult.csv"))
 ```
@@
 cn.df.eulerfilt <- cn.df.noinf %>%
-  dplyr::filter(!is.na(surfholes))
+  dplyr::filter(!is.na(surfholes)) #drops 2 studies w missing euler data
 
@@
   mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_")))) 
+skim(cn.df.imp1)
+```
 
+```{r}
 #drop NAs listwise
@@
 fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler.csv"))
+
+#save names with site (since not saved correctly from centile outputs)
+study_site_key <- cn.df.imp %>%
+  dplyr::select(study, site) %>%
+  unique()
+n_unique(study_site_key$study) == n_unique(cn.df.imp$study)
+n_unique(study_site_key$site) == n_unique(cn.df.imp$site)
+fwrite(study_site_key, file=paste0(base_path, "/data/lifespan_study_site_key.csv"))
 ```

### 2024-05-01 paper_lbcc_results.Rmd
fixing study names

@@
 
+#key to get study names
+study_site_key <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/lifespan_study_site_key.csv")
+
 #PATH
@@
 gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv"))
-gam_centile_df <- gam_centile_df %>% mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25)
+
+gam_centile_df <- full_join(gam_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25)
+n_unique(gam_centile_df$study)
 
@@
 gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv"))
-gamlss_centile_df <- gamlss_centile_df %>% mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25)
+gamlss_centile_df <- full_join(gamlss_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25)
 
@@
 raw_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_predictions.csv"))
-raw_centile_df <- raw_centile_df %>% mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25)
+raw_centile_df <- full_join(raw_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25)
 
@@
 deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
-deltas_df <- deltas_df %>% mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25)
+deltas_df <- full_join(deltas_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25)
 
@@
 #Clean & EDA
+
+Check numbers
+```{r}
+nrow(gamlss_centile_df) #n
+n_unique(gamlss_centile_df$study) 
+n_unique(gamlss_centile_df$site) #correct
+```
 
@@
 #clean predictions
-raw_centile_df.clean <- raw_centile_df %>%
-  mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25) %>%
+raw_centile_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 #clean predictions
-raw_z_df.clean <- raw_centile_df %>%
-  mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25) %>%
+raw_z_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 #clean predictions
-gam_centile_df.clean <- gam_centile_df %>%
-  mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25) %>%
+gam_centile_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25)%>%
   dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 #clean predictions
-gam_z_df.clean <- gam_centile_df %>%
-  mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25) %>%
+gam_z_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 #clean predictions
-gamlss_centile_df.clean <- gamlss_centile_df %>%
-  mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25) %>%
+gamlss_centile_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!contains(".z")) #drop z-scores and non-centiles & !ends_with("_total") & !contains("TBV")
@@
 #clean predictions
-gamlss_z_df.clean <- gamlss_centile_df %>%
-  mutate(study=sub("\\_.*", "", site),
-         age_years=age_days/365.25) %>%
+gamlss_z_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
+  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 feat_aov.df %>%
+  ggplot() +
+  geom_point(aes(x=etasq, y=pheno, color=combat, shape=sig_fdr), alpha=0.6) +
+  facet_wrap(~pheno_cat, scales="free")
+
+feat_aov.df %>%
+  dplyr::filter(combat!="raw") %>%
   ggplot() +

### 2024-05-01 paper_lbcc_results.Rmd
figure formatting

@@
   labs(y="Mean Centile", x = "Study")
+
+mean_cents_bxplt_nolabel <- mean_cents.df %>%
+  ggplot() +
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
+  theme_linedraw() +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none", strip.text.x = element_text(size = 12)) +
+  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
+  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
+  labs(y="Mean Centile", x = "Study")
 ```
@@
                                       p.val_fdr >= 0.05 ~ FALSE))
+
+df_gam_feat_lm %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+
 df_gam_feat_lm %>%
@@
 ```
+all assoc with study n
 
@@
 
-control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting
+control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.
 
@@
   theme(axis.title.x=element_blank(), axis.text.x = element_blank()) +
-  labs(color="Sex", y = "Age (Years) \n ")
+  labs(color="Sex", y = "Age (Years)")
 
@@
   gamlss_centile_df %>%
-  mutate(xlab = "Full Sample") %>%
+  mutate(xlab = "Full Sample ") %>%
   ggplot(aes(x=xlab, y=age_years, color=sex)) +
@@
   geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
-  scale_y_continuous(labels = scales::percent) +
+  #scale_y_continuous(labels = scales::percent) +
   geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
@@
   #theme(axis.ticks.x = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
-  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
-  labs(fill="Sex", x = "Study", y="Sex")
+  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=6)) +
+  labs(fill="Sex", x = "Study", y="Percent Sex")
 
@@
   theme_linedraw(base_size = 11) +
-  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
+  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
   labs(color="Sex", y = "Age (Years)", y="")
-
-ggarrange(sample_boxplot2, full_boxplot2,
-          sample_barplot2, full_barplot2,
-          nrow=2, ncol=2, 
-          align = "hv", 
-          common.legend = TRUE,
-          legend = "right",
-          widths=c(10, 1),
-          heights=c(2, 1)
-          )
 ```
@@
 fig5_vert <- ggarrange(
-  ggarrange(blank,
+  #ggarrange(blank,
 ggarrange(sample_boxplot2, full_boxplot2,
@@
           ),
-blank,
-nrow=1, ncol=3,
-common.legend = FALSE,
-widths= c(1, 20, 1)),
+# blank,
+# nrow=1, ncol=3,
+# common.legend = FALSE,
+# widths= c(1, 20, 1)),
 # boxplots
-mean_cents_bxplt,
+mean_cents_bxplt_nolabel,
 common.legend=FALSE,
@@
 
-mean_cent_deltas_df_study %>%
+mean_delta_fig <- mean_cent_deltas_df_study %>%
   ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
@@
     geom_line(data = pred_df, aes(y = fit, x=study_mean_age))+
-  theme(legend.position="none") +
+  #theme(legend.position="none") +
   labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
-```
+ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=10, height=7, units="in")
+
+mean_cent_deltas_df_study %>%
+  dplyr::filter(abs_mean_delta_cent > 4) %>%
+  dplyr::select(study, study_mean_age)
+
+mean_cent_deltas_df_study%>%
+  dplyr::filter(study == "Cornell_C2") %>%
+  nrow()
+```
+
+Cornell_C2 has very low n=5 -> big differences in estimation
 

### 2024-05-01 paper_lbcc_results.Rmd
cleaning up, demographics

@@
 
+Number of models converged in each dataset
+```{r}
+length(raw.feat_to_test)
+length(gam.feat_to_test)
+length(gamlss.feat_to_test)
+
+#missing features
+c(setdiff(pheno_list, gamlss.feat_to_test),
+setdiff(pheno_list, raw.feat_to_test),
+setdiff(pheno_list, gam.feat_to_test)) %>% unique()
+```
+
+```{r}
+summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
+summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
+summary(lm(mean_cent ~ study, data=raw_centile_df.mean))
+```
+
 compare magnitude of effect size for study (w/in feature anovas) across unharmonized, combat-gam and combatLS centiles
@@
       TRUE ~ NA_character_)))
+
+feat_aov.df %>%
+  group_by(combat) %>%
+  summarize(n_sig = sum(sig_fdr))
 

### 2024-05-01 cubic_scripts/R_scripts/fit_lifespan_mod_site_est.R
correct 'site' to 'study' term, back to BCCG

@@ -14,11 +14,11 @@ pheno <- as.character(args[2])
 save_path <- args[3]
 
 #FIT BASE MODEL
-base_model <- gamlss(formula = as.formula(paste0(pheno,"~ pb(age_days) + sexMale + fs_version + pb(sex.age) + site")),
-             sigma.formula = as.formula(paste0("~ pb(age_days) + sexMale + pb(sex.age) + fs_version + site")),
+base_model <- gamlss(formula = as.formula(paste0(pheno,"~ pb(age_days) + sexMale + fs_version + pb(sex.age) + study")),
+             sigma.formula = as.formula(paste0("~ pb(age_days) + sexMale + pb(sex.age) + fs_version + study")),
              nu.formula = as.formula(paste0("~ pb(age_days) + sexMale + pb(sex.age) + fs_version")),
              control = gamlss.control(n.cyc = 200), 
-             family = BCCGo, data=df, trace = FALSE)
+             family = BCCG, data=df, trace = FALSE)
 
 #SAVE
 csv_name <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(as.character(args[1])))

### 2024-05-01 eda/ukb_curation-eda.Rmd
cleaning up

@@
 pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)
-
-# #pull plotting functions from github
-# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
-# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
-# source_url(plot.func.url) #sometimes doesn't work....?
 

### 2024-05-02 eda/lifespan_eda.Rmd
cleaning up, demographics

@@
 pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see, broom.mixed)
-
-# #pull plotting functions from github
-# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
-# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
-# source_url(plot.func.url) #sometimes doesn't work....?
 
@@
 length(unique(cn.df.imp$participant))
-nrow(cn.df.imp)
+nrow(cn.df.imp) == length(unique(cn.df.imp$participant))
 
 table(cn.df.imp$sex)
+
+table(cn.df.imp$sex)["Female"]/nrow(cn.df.imp)
 

### 2024-05-16 paper_lbcc_results.Rmd
looking at change in age effect on variance - plots

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize, gratia)
 #Sys.setenv("MC_CORES"=10L)
@@
 
-#source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
+source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
 #source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
@@
   labs(y="Mean Centile", x = "Study")
+```
+
+Cohen's F squared for study effects
+
+```{r}
+gam_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv"))
+gam_centile_f2$combat <- "ComBat-GAM"
+
+gamlss_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_cohenfsq.csv"))
+gamlss_centile_f2$combat <- "ComBatLS"
+
+f2.df <- rbind(gam_centile_f2, gamlss_centile_f2)
+f2.df <- f2.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+ggplot(f2.df) +
+  geom_col(aes(x=pheno, y=fsq, fill=combat), position="dodge") + 
+  facet_wrap(~pheno_cat, scales="free") +
+  theme(axis.text.x = element_blank())
+```
+
+```{r}
+f2.df.sum <- f2.df %>%
+  group_by(pheno) %>%
+  arrange(combat) %>%
+  mutate(F2_diff = fsq - fsq[1]) #gamlss - gam value, + number means site effect is larger in combatLS
+
+ggplot(f2.df.sum) +
+  geom_col(aes(x=pheno, y=F2_diff), position="dodge") + 
+  facet_wrap(~pheno_cat, scales="free") +
+  theme(axis.text.x = element_blank())
+
+f2.df.sum %>%
+  group_by(pheno_cat) %>%
+  summarise(mean_fsq_diff = mean(F2_diff)) %>%
+  ggplot() + 
+  geom_col(aes(x=pheno_cat, y=mean_fsq_diff), position="dodge")
+
+f2.df.sum %>%
+  #group_by(pheno_cat) %>%
+  mutate(smaller_eff = case_when(F2_diff < 0 ~ "ComBatGAM",
+                                    F2_diff > 0 ~ "ComBatLS")) %>%
+  ggplot() +
+  geom_bar(aes(x=smaller_eff, fill=pheno_cat), position="dodge")
+```
+
+## Centile Fans
+**try plotting shapes of trajectores w/ and w/o study term to see how much they change the effect of age!
+
+load unharmonized data
+```{r}
+raw.data <- fread(file="/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/lifespan_CN_imp-sites_euler.csv")
+gmv.raw.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler_raw_no.tbv_mod.rds"))
+
+gmv.raw.pred.df <- pred.across.ages.lbcc(gmv.raw.mod, raw.data)
+
+gmv.raw.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma), color="red") +
+  geom_line(aes(x=age_days, y=m.sigma), color="blue")
+
+gmv.raw.mod.site <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-raw_site.est_mod.rds"))
+
+gmv.raw.pred.site.df <- pred.across.ages.lbcc.study(gmv.raw.mod.site, raw.data)
+
+gmv.raw.pred.site.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma), color="red") +
+  geom_line(aes(x=age_days, y=m.sigma), color="blue")
+
+gmv.raw.pred.site.df <- gmv.raw.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "raw")
+gmv.raw.pred.df <- gmv.raw.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "raw")
+```
+
+load ComBat-GAM-harmonized data
+```{r}
+cf.gam.data <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gam_batch.study_data.csv"))
+gmv.cf.gam.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_mod.rds"))
+
+gmv.cf.gam.pred.df <- pred.across.ages.lbcc(gmv.cf.gam.mod, cf.gam.data)
+
+gmv.cf.gam.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma), color="red") +
+  geom_line(aes(x=age_days, y=m.sigma), color="blue")
+
+gmv.cf.gam.mod.site <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_site.est_mod.rds"))
+
+gmv.cf.gam.pred.site.df <- pred.across.ages.lbcc.study(gmv.cf.gam.mod.site, cf.gam.data)
+
+gmv.cf.gam.pred.site.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma), color="red") +
+  geom_line(aes(x=age_days, y=m.sigma), color="blue")
+```
+
+```{r}
+gmv.cf.gam.pred.site.df <- gmv.cf.gam.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combatGAM")
+gmv.cf.gam.pred.df <- gmv.cf.gam.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combatGAM")
+
+gmv.cf.gam.pred.df_full <- full_join(gmv.cf.gam.pred.site.df, gmv.cf.gam.pred.df)
+
+gmv.cf.gam.pred.df_full %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma, color=study_incl))
+```
+
+load ComBatLS-harmonized data
+```{r}
+cf.gamlss.data <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gamlss_batch.study_data.csv"))
+gmv.cf.gamlss.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_mod.rds"))
+
+gmv.cf.gamlss.pred.df <- pred.across.ages.lbcc(gmv.cf.gamlss.mod, cf.gamlss.data)
+
+gmv.cf.gamlss.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma), color="red") +
+  geom_line(aes(x=age_days, y=m.sigma), color="blue")
+
+gmv.cf.gamlss.mod.site <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_site.est_mod.rds"))
+
+gmv.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(gmv.cf.gamlss.mod.site, cf.gamlss.data)
+
+gmv.cf.gamlss.pred.site.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma), color="red") +
+  geom_line(aes(x=age_days, y=m.sigma), color="blue")
+```
+
+```{r}
+gmv.cf.gamlss.pred.site.df <- gmv.cf.gamlss.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combatLS")
+gmv.cf.gamlss.pred.df <- gmv.cf.gamlss.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combatLS")
+
+gmv.cf.gamlss.pred.df_full <- full_join(gmv.cf.gamlss.pred.site.df, gmv.cf.gamlss.pred.df)
+
+gmv.cf.gamlss.pred.df_full %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma, color=study_incl))
+```
+
+```{r}
+all.pred.df <- reduce(list(gmv.cf.gam.pred.df_full, gmv.cf.gamlss.pred.df_full, gmv.raw.pred.site.df, gmv.raw.pred.df), full_join)
+
+all.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma, linetype=study_incl, color=combat))
+
+all.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=m.sigma, linetype=study_incl, color=combat)) +
+  ggtitle("GMV")
+
+raw.data %>%
+  ggplot() +
+  geom_histogram(aes(x=age_days, fill=sex), position="identity", alpha=0.5)
+```
+
+See WMV
+```{r}
+#raw
+wmv.raw.mod <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler_raw_no.tbv_mod.rds"))
+wmv.raw.pred.df <- pred.across.ages.lbcc(wmv.raw.mod, raw.data)
+wmv.raw.mod.site <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-raw_site.est_mod.rds"))
+wmv.raw.pred.site.df <- pred.across.ages.lbcc.study(wmv.raw.mod.site, raw.data)
+
+wmv.raw.pred.site.df <- wmv.raw.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "raw")
+wmv.raw.pred.df <- wmv.raw.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "raw")
+
+#combat-gam
+wmv.cf.gam.mod <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_mod.rds"))
+wmv.cf.gam.pred.df <- pred.across.ages.lbcc(wmv.cf.gam.mod, cf.gam.data)
+
+wmv.cf.gam.mod.site <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_site.est_mod.rds"))
+wmv.cf.gam.pred.site.df <- pred.across.ages.lbcc.study(wmv.cf.gam.mod.site, cf.gam.data)
+
+wmv.cf.gam.pred.site.df <- wmv.cf.gam.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combatGAM")
+wmv.cf.gam.pred.df <- wmv.cf.gam.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combatGAM")
+
+#combatLS
+wmv.cf.gamlss.mod <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_mod.rds"))
+wmv.cf.gamlss.pred.df <- pred.across.ages.lbcc(wmv.cf.gamlss.mod, cf.gamlss.data)
+
+wmv.cf.gamlss.mod.site <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_site.est_mod.rds"))
+wmv.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(wmv.cf.gamlss.mod.site, cf.gamlss.data)
+
+wmv.cf.gamlss.pred.site.df <- wmv.cf.gamlss.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combatLS")
+wmv.cf.gamlss.pred.df <- wmv.cf.gamlss.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combatLS")
+```
+
+```{r}
+wmv.all.pred.df <- reduce(list(wmv.raw.pred.site.df, wmv.raw.pred.df, wmv.cf.gam.pred.site.df, wmv.cf.gam.pred.df, wmv.cf.gamlss.pred.site.df, wmv.cf.gamlss.pred.df), full_join)
+
+wmv.all.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma, linetype=study_incl, color=combat))
+
+wmv.all.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=m.sigma, linetype=study_incl, color=combat)) +
+  ggtitle("WMV")
+
+raw.data %>%
+  ggplot() +
+  geom_histogram(aes(x=age_days, fill=sex), position="identity", alpha=0.5)
+```
+
+See sGMV
+```{r}
+#raw
+sgmv.raw.mod <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler_raw_no.tbv_mod.rds"))
+sgmv.raw.pred.df <- pred.across.ages.lbcc(sgmv.raw.mod, raw.data)
+sgmv.raw.mod.site <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-raw_site.est_mod.rds"))
+sgmv.raw.pred.site.df <- pred.across.ages.lbcc.study(sgmv.raw.mod.site, raw.data)
+
+sgmv.raw.pred.site.df <- sgmv.raw.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "raw")
+sgmv.raw.pred.df <- sgmv.raw.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "raw")
+
+#combat-gam
+sgmv.cf.gam.mod <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_mod.rds"))
+sgmv.cf.gam.pred.df <- pred.across.ages.lbcc(sgmv.cf.gam.mod, cf.gam.data)
+
+sgmv.cf.gam.mod.site <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_site.est_mod.rds"))
+sgmv.cf.gam.pred.site.df <- pred.across.ages.lbcc.study(sgmv.cf.gam.mod.site, cf.gam.data)
+
+sgmv.cf.gam.pred.site.df <- sgmv.cf.gam.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combatGAM")
+sgmv.cf.gam.pred.df <- sgmv.cf.gam.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combatGAM")
+
+#combatLS
+sgmv.cf.gamlss.mod <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_mod.rds"))
+sgmv.cf.gamlss.pred.df <- pred.across.ages.lbcc(sgmv.cf.gamlss.mod, cf.gamlss.data)
+
+sgmv.cf.gamlss.mod.site <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_site.est_mod.rds"))
+sgmv.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(sgmv.cf.gamlss.mod.site, cf.gamlss.data)
+
+sgmv.cf.gamlss.pred.site.df <- sgmv.cf.gamlss.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combatLS")
+sgmv.cf.gamlss.pred.df <- sgmv.cf.gamlss.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combatLS")
+```
+
+```{r}
+sgmv.all.pred.df <- reduce(list(sgmv.raw.pred.site.df, sgmv.raw.pred.df, sgmv.cf.gam.pred.site.df, sgmv.cf.gam.pred.df, sgmv.cf.gamlss.pred.site.df, sgmv.cf.gamlss.pred.df), full_join)
+
+sgmv.all.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma, linetype=study_incl, color=combat))
+
+sgmv.all.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=m.sigma, linetype=study_incl, color=combat)) +
+  ggtitle("sGMV")
+
+raw.data %>%
+  ggplot() +
+  geom_histogram(aes(x=age_days, fill=sex), position="identity", alpha=0.5)
+```
+
+See Ventricles
+```{r}
+#raw
+ven.raw.mod <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler_raw_no.tbv_mod.rds"))
+ven.raw.pred.df <- pred.across.ages.lbcc(ven.raw.mod, raw.data)
+# ven.raw.mod.site <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler_raw_site.est_mod.rds"))
+# ven.raw.pred.site.df <- pred.across.ages.lbcc.study(ven.raw.mod.site, raw.data)
+
+# ven.raw.pred.site.df <- ven.raw.pred.site.df %>%
+#   mutate(study_incl = TRUE,
+#          combat= "raw")
+ven.raw.pred.df <- ven.raw.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "raw")
+
+#combat-gam
+ven.cf.gam.mod <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_mod.rds"))
+ven.cf.gam.pred.df <- pred.across.ages.lbcc(ven.cf.gam.mod, cf.gam.data)
+
+ven.cf.gam.mod.site <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_site.est_mod.rds"))
+ven.cf.gam.pred.site.df <- pred.across.ages.lbcc.study(ven.cf.gam.mod.site, cf.gam.data)
+
+ven.cf.gam.pred.site.df <- ven.cf.gam.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combatGAM")
+ven.cf.gam.pred.df <- ven.cf.gam.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combatGAM")
+
+#combatLS
+ven.cf.gamlss.mod <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_mod.rds"))
+ven.cf.gamlss.pred.df <- pred.across.ages.lbcc(ven.cf.gamlss.mod, cf.gamlss.data)
+
+ven.cf.gamlss.mod.site <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_site.est_mod.rds"))
+ven.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(ven.cf.gamlss.mod.site, cf.gamlss.data)
+
+ven.cf.gamlss.pred.site.df <- ven.cf.gamlss.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combatLS")
+ven.cf.gamlss.pred.df <- ven.cf.gamlss.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combatLS")
+```
+
+```{r}
+ven.all.pred.df <- reduce(list(ven.raw.pred.df, ven.cf.gam.pred.site.df, ven.cf.gam.pred.df, ven.cf.gamlss.pred.site.df, ven.cf.gamlss.pred.df), full_join) #ven.raw.pred.site.df
+
+ven.all.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma, linetype=study_incl, color=combat))
+
+ven.all.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=m.sigma, linetype=study_incl, color=combat)) +
+  ggtitle("Ventricles")
+
+raw.data %>%
+  ggplot() +
+  geom_histogram(aes(x=age_days, fill=sex), position="identity", alpha=0.5) +
+  ggtitle("Sample distribution")
 ```

### 2024-05-17 paper_main_results.Rmd
cleaning up

@@
 
-Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.
+Main analyses of ComBatLS paper.
+
+Relies on outputs derived/functions from:
+
+* `./cubic_scripts/R_scripts/stats_tests.R`
+1. `ukb_curation-eda.Rmd`
+2. `qsub_combat.sh` -> `/cubic_scripts/R_scripts/combat_apply.R`
+3. `qsub_basic_gamlss.sh` -> `/cubic_scripts/R_scripts/fit_basic_mod.R`
+4. `qsub_centiles.sh` -> `/cubic_scripts/R_scripts/fit_centiles.R`
+5. `gamlss_parse.sh` -> `/cubic_scripts/R_scripts/gamlss_parser.R`
 
@@
 pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x)
-#Sys.setenv("MC_CORES"=10L)
-options("mc.cores")
 ```
@@
 source("./cubic_scripts/R_scripts/stats_tests.R")
-# devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/stats_tests.R")
 ```
@@
 
-# Main Analyses
-
-## Load Data
+# Load Data
 
@@
 
-**QC: find ppts with centiles at 1 or 0**
+QC: check for ppts with centiles ~ 1 or 0 and make sure Z scores exist
 
@@
   dplyr::select(participant) %>%
-  unique()
+  unique() %>%
+  nrow()
 
@@
   dplyr::select(participant) %>%
-  unique()
-```
-
-saving out csv forces some centile scores to round up to 1, BUT it appears z scores are still being calculated correctly so just going to proceed!
+  unique() %>%
+  nrow()
+```
+
+saving out csv forces some centile scores to round up to 1, but z scores are still being calculated correctly.
 
@@
 
-Get differences in male vs female predicted variances at mean sample age
-
-```{r}
-pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
-combined_var <- data.frame()
-
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_variance.csv", "", Source_File),
-           sex_standard = sex_effect/f.var)
-
-  # Bind the data to the combined dataframe
-  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
-}
-
-#get summary df for p vals, plotting
-sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
-  dplyr::select(-V1) %>%
-  mutate(dataset = gsub("-data", "", dataset),
-         dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
-  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
-  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
-  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)
-
-sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")
-
-plot.info <- sigma.sex.df %>%
-  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
-  unique() %>%
-  mutate(dataset = gsub("-", "_", dataset),
-         dataset = gsub("_simsites", "", dataset))
-
-combined_var <- inner_join(combined_var, plot.info)
-
-#drop refA sites, not using for these analyses
-combined_var <- combined_var %>%
-  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
-```
-
-## SCALE EFFECTS OF SEX
-
-```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-#var plot
-var.plot.df <- combined_var %>%
-  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  group_by(dataset, pheno_cat)
-
-tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot <- ggplot(var.plot.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = dk, 
-             position = position_brain(hemi ~ side),
-             aes(fill = sex_standard)) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-#var.plot
-
-#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
-```
-
-```{r}
-#plot global
-var.plot.global.df <- combined_var %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
-  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
-                                            pheno == "Ventricles" ~ "CSF",
-                                            pheno == "WMV" ~ "WM",
-                                            pheno == "sGMV" ~ "sGM",
-                                            TRUE ~ NA)),
-         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
-  group_by(dataset)
-
-tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
-tp$sex_standard <- 1
-
-var.plot.global <- ggplot(var.plot.global.df) +
-  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
-            ymin = -Inf, ymax = Inf, alpha = 0.4) +
-  ggseg::geom_brain(atlas = aseg_cerebrum, 
-             aes(fill = sex_standard),
-             side = c("coronal")) +
-  #scale_fill_gradient2() +
-  facet_grid(dataset~pheno_cat) + 
-  theme_linedraw(base_size=12) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
-#var.plot.global
-```
-
-Combine plots
-```{r}
-p <- ggarrange(
-  var.plot.global, var.plot,
-  ncol=2, nrow=1, 
-  widths = c(1, 3.1),
-  #labels=c("A", "B"),
-  #hjust=0.01,
-  align = "hv",
-  common.legend = TRUE, legend = "bottom"
-  )
-
-p_annotate <- annotate_figure(p, top = text_grob("Figure 1B. Standardized effect of male sex on feature variance", 
-               color = "black", size = 14, hjust=.65))
-p_annotate
-ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
-```
-Look at relative estimates for sex in sigma across features in raw data
-
-```{r}
-gamlss.sum.df <- fread(paste0(data_path, "/ukb_basic/gamlss_summary.csv"))
-gamlss.sum.df.raw <- gamlss.sum.df %>%
-  dplyr::filter(dataset=="ukb-CN-data-simsites")
-
-gamlss.sum.df.raw.sexsig <- gamlss.sum.df.raw %>%
-  dplyr::filter(parameter =="sigma" & term == "sexMale") %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      est_over_moment = estimate/moment.mean)
-
-gamlss.sum.df.raw.sexsig %>%
-  ggplot(aes(y=estimate, x=pheno_cat, fill=pheno_cat)) +
-  geom_boxplot() +
-  # geom_hline(yintercept = mean(gamlss.sum.df.raw.sexsig$t_stat), linetype = 2) +
-  stat_compare_means(method = "anova", label.y = 0.07)+    
-  stat_compare_means(label = "p.signif", method = "t.test",
-                     ref.group = ".all.")
-
-gamlss.sum.df.raw.sexsig %>%
-  ggplot() +
-  geom_histogram(aes(x=est_over_moment, fill=pheno_cat, color=pheno_cat), alpha=0.5, position="identity")
-
-gamlss.sum.df.raw.sexsig %>%
-  ggplot(aes(y=t_stat, x=pheno_cat, fill=pheno_cat)) +
-  geom_boxplot() +
-  # geom_hline(yintercept = mean(gamlss.sum.df.raw.sexsig$t_stat), linetype = 2) +
-  stat_compare_means(method = "anova", label.y = -7)+    
-  stat_compare_means(label = "p.signif", method = "t.test",
-                     ref.group = ".all.")
-```
-
-## CENTILES
+# CENTILES
 
 Do centile scores change based on combat version?
-Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)
+Foregoing anova based on emails with Andrew (anova violates assumptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)
 
@@
 full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
-saveRDS(full_cent_res, "./derivatives/ukb_basic/featurewise_cent_t_tests_all_out.RDS")
+saveRDS(full_cent_res, "./derivatives/ukb_basic/featurewise_cent_t_tests_all_out.RDS") #running once and saving because it takes a long time
 ```
@@
 
-```{r, eval=FALSE}
-full_cent_res%>%
-  dplyr::filter(comp =="cf.gamlss_cf" & bigger_group == "cf.gamlss" & sig_fdr ==TRUE)
-
-diffs.df %>%
-  dplyr::filter(dataset=="cf.gamlss" | dataset =="cf") %>%
-  ggplot() +
-  geom_boxplot(aes(y=abs.diff_rh_CT_parstriangularis, fill=dataset))
-
-diffs.df %>%
-  dplyr::filter(dataset=="cf.gamlss" | dataset =="cf") %>%
-  ggplot() +
-  geom_boxplot(aes(y=diff_rh_CT_parstriangularis, fill=dataset))
-```
+bigger group -> relatively worse fit
 
@@
   dplyr::filter(pheno %in% vol_list_global) %>%
-  dplyr::filter(!grepl('_refA', comp))
-
+  dplyr::filter(!grepl('_refA', comp)) %>%
+  summarise(n_sig = sum(sig_fdr),
+            nrow = nrow(.))
+
+#n occurances of combatLS having significantly larger errors
 full_cent_res %>%
   dplyr::filter(bigger_group == "cf.gamlss")%>%
-  dplyr::filter(!grepl('_refA', comp))
-```
-
-all significant
-
-```{r}
+  dplyr::filter(!grepl('_refA', comp)) %>%
+  summarise(n_sig = sum(sig_fdr),
+            nrow = nrow(.))
+```
+
+```{r}
+dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+
 global_abs.diffs <- diffs.df %>%
@@
 ```{r}
-gam.gamlss.diffs <- diffs.df %>%
-  dplyr::filter(dataset=="cf.gam" | dataset=="cf.gamlss")
-
-gam_better <- gam.gamlss.diffs %>%
-  group_by(dataset) %>%
-  summarize_at(pheno_abs.diff.list, median) %>%
-  ungroup() %>%
-  pivot_longer(cols = all_of(pheno_abs.diff.list), names_to="pheno") %>%
-  pivot_wider(names_from = dataset, values_from=value, names_glue = "med_{dataset}") %>%
-  mutate(gam_minus_gamlss = (med_cf.gam - med_cf.gamlss),
-         pheno = sub("abs.diff_", "", pheno)) %>%
-  ungroup() %>%
-  dplyr::filter(gam_minus_gamlss < 0)
-
 ns_gam_gamlss_comp <- full_cent_res %>%
-  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE)
-
-cfgam.better <- intersect(gam_better$pheno, ns_gam_gamlss_comp$pheno)
-```
-```{r, eval=FALSE}
-cfgam.better.ct <- gsub("_CT", "", cfgam.better)
-cfgam.better.ct_plot <- dk.parc %>%
-  dplyr::mutate(gam_better = ifelse(label %in% cfgam.better.ct, 1, 0))
-
-ggplot(cfgam.better.ct_plot) +
-  theme_linedraw(base_size=12) +
-  ggseg::geom_brain(atlas = dk_nomed, 
-             position = position_brain(hemi ~ side),
-            aes(fill=as.factor(gam_better)))
-```
-
-```{r, eval=FALSE}
-for (pheno in pheno_abs.diff.list){
-  p <- ggplot(gam.gamlss.diffs, aes(y = gam.gamlss.diffs[[pheno]], fill= dataset)) +
-      geom_boxplot() +
-    ggtitle(pheno)
-print(p)
-}
+  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE & bigger_group == "cf.gamlss")
+nrow(ns_gam_gamlss_comp)
 ```
@@
 
-### Sex-Biases
-
-Playing with different ways to visualize each subject's average centile error (i.e., raw data centile - combatted data centile) and magnitude centile error (absolute value)
-
-```{r, eval=FALSE}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot() +
-  geom_boxplot(aes(x=sex, y=mean_cent_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
-  facet_wrap(~dataset) +
-  theme(legend.position="none") +
-  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
-  ggplot(aes(x=dataset, y=mean_cent_diff, fill=sex)) +
-  geom_violin(trim=FALSE, scale="width") +
-  labs(y="Centile Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean error in centile scores across ComBat configurations")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
-  ggplot(aes(x=sim.site, y=mean_cent_abs.diff, fill=sex)) +
-  geom_violin(trim=FALSE) +
-  facet_wrap(~dataset) +
-  labs(y="Magnitude Centile Error", x = "Simulated Site") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in centile scores across \nComBat configurations")
-
-#look at sex distributions of ppts across sites
-subj_sum.df %>%
-  ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex))+
-  ggtitle("Distributions of male and female subjects across simulated sites")
-```
-
-Significance testing (w/in sub. mean abs centile error)  - using wilcox (mann whitney u) since distributions are skewed but variances are ~= across sexes
-
-```{r}
-sex.bias.wilcox.tests(subj_sum.df)
-sex.bias.wilcox.tests(subj_sum.df, to_test="mean_cent_diff")
-```
-
-Keeping these analyses in case I want to use one of the above violin/boxplots that show w/in subject averages (across features).
-However, while these results I think are good to have from a "be aware of how biological variables are distributed across your study sites" perspective, it doesnt really say anything interesting froma "combat methods" perspective.
-
-Testing for significant M:F differences in centile erros WITHIN each feature!
+### Sex-Differences
+
+Testing for significant M:F differences in centile errors within each feature
 
@@
 
-```
-
-```{r}
-plt.df <- diffs.df %>%
-  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", pheno_list)) %>% #select centile errors only
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
-  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         diff = diff*100,
-         pheno_cat = factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_), ordered=TRUE, levels=c("Global Volume", "Regional Volume", "Regional SA", "Regional CT"))) %>%
-    mutate(pheno = factor(pheno, ordered=TRUE, levels=c(vol_list_global, vol_list_regions, sa_list, ct_list)))#sort w/in pheno_cat
-```
-
-stat test re-do w wilcox 
-```{r}
-stat.test <- plt.df %>%
-  group_by(dataset, pheno) %>%
-  wilcox_test(diff ~ sex) %>%
-  adjust_pvalue(method = "fdr") %>%
-  add_significance("p.adj")
-```
-```{r}
-stat.test2 <- stat.test %>%
-  add_xy_position(x = "pheno", dodge = 0.8, fun="max") %>%
-  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*")) #simplify significance labels
-```
-
-```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
-
-big_violin_plt <- plt.df %>%
-  ggplot(aes(x=pheno, y=diff)) +
-  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
-  # geom_violin(aes(fill=sex, color=sex)) +
-  theme_linedraw(base_size = 11) +
-  labs(y="Centile Error", x = "Brain Feature") +
-  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines"), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 3, tip.length = 0, hide.ns = TRUE, remove.bracket = TRUE) +
-  # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
-  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
-  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
-  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
-  labs(fill = "Sex", color="Sex")
-  # ggtitle("Male and female centile errors")
-big_violin_plt
-ggsave("figs/big_sex_violin_plt.jpeg", big_violin_plt, width=12, height=6, units="in")
-```
-
-```{r, eval=FALSE}
-big_violin_plt_wlabels <- plt.df %>%
-  ggplot(aes(x=pheno, y=diff)) +
-  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
-  # geom_violin(aes(fill=sex, color=sex)) +
-  theme_linedraw(base_size = 11) +
-  labs(y="Centile Error", x = "Brain Feature") +
-  theme(plot.title = element_text(size=12)) +
-  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 2.5, tip.length = 0, remove.bracket = TRUE) +
-  # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
-  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
-  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
-  labs(fill = "Sex", color="Sex")
-  # ggtitle("Male and female centile errors")
-big_violin_plt_wlabels
-ggsave("figs/big_sex_violin_plt_wlabels.jpeg", big_violin_plt_wlabels, width=16, height=9, units="in")
 ```
@@
 
-p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
+p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-differences in centile score errors across combat configurations", 
                color = "black", size = 14, hjust=.65))
@@
 units are in centiles).
-Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)?
-Honestly, not the most straightforward interpretation, i know - may make more sense to just plot the median errors separately in males and females than to plot the differences between the two.
-
-look at # males vs females w extreme mean centiles
+Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)
 
@@
 ```
+
+bigger group -> relatively worse fit
 
@@
 
-w/in feature sex-biases
+w/in feature sex-differences
 ```{r, message=FALSE, eval=FALSE}
@@
 
-## Z-SCORES
+# Z-SCORES
 
@@
 
-Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)
-
-Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew).
-(see Zimmerman & Zumbo, 1993).
-Using custom functions from `ranked_welch_tests.R`.
+Within each phenotype, iterate paired t test of ranks between each combo of combat dfs using custom functions from `ranked_welch_tests.R`.
 
@@
 
-plot w ref sites
-```{r, eval=FALSE}
-#get means w/in feature 
-mean.z.diffs.long <- diffs.df %>%
-  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss", "cf.lm_refA", "cf.gam_refA", "cf.gamlss_refA"), ordered = TRUE)) %>%
-  group_by(dataset) %>%
-  summarize(across(starts_with("abs.diff_"), mean)) %>%
-  ungroup() %>%
-  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
-  mutate(pheno=gsub(".z", "", pheno)) %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)),
-      label = sub("_[^_]*_", "_", pheno))
-
-mean.z.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
-  ggplot() +
-  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
-  theme_linedraw(base_size = 11) +
-  facet_grid(dataset~pheno_cat) +
-  scale_y_continuous(labels = scales::percent) + 
-  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
-  ggtitle("Absolute errors in zscores of brain structure features across ComBat processing pipelines")
-```
-
 Quick comp of cf.gam and cf.gamlss
@@
 ```{r}
-gam_better.z <- gam.gamlss.diffs %>%
-  group_by(dataset) %>%
-  summarize_at(z.pheno_abs.diff.list, median) %>%
-  ungroup() %>%
-  pivot_longer(cols = all_of(z.pheno_abs.diff.list), names_to="pheno") %>%
-  pivot_wider(names_from = dataset, values_from=value, names_glue = "med_{dataset}") %>%
-  mutate(gam_minus_gamlss = (med_cf.gam - med_cf.gamlss),
-         pheno = sub("abs.diff_", "", pheno)) %>%
-  ungroup() %>%
-  dplyr::filter(gam_minus_gamlss < 0)
-
-gam_better.z$pheno
-
 ns_gam_gamlss_comp.z <- full_z_res %>%
-  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE)
-
-intersect(gam_better.z$pheno, ns_gam_gamlss_comp.z$pheno)
-```
-
-### Sex-Biases
-
-Visualize
-
-```{r, eval=FALSE}
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot() +
-  geom_boxplot(aes(x=sex, y=mean_z_abs.diff, fill=sex)) +
-  labs(y= "Magnitude Z Error", x = "ComBat Configurations") +
-  facet_wrap(~dataset, scales="free_y") +
-  theme(legend.position="none") +
-  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  ggplot(aes(x=dataset, y=mean_z_diff, fill=sex)) +
-  geom_violin(trim=FALSE, scale="width") +
-  labs(y="Z Error", x = "ComBat Configurations") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean error in z-scores across ComBat configurations")
-
-subj_sum.df %>%
-  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
-  ggplot(aes(x=sim.site, y=mean_z_abs.diff, fill=sex)) +
-  geom_violin(trim=FALSE) +
-  facet_wrap(~dataset) +
-  labs(y="Magnitude Z Error", x = "Simulated Site") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ggtitle("Within-subject mean magnitude error in z-scores across \nComBat configurations")
-
-#look at sex distributions of ppts across sites
-subj_sum.df %>%
-  ggplot() +
-  geom_bar(aes(x=sim.site, fill=sex), position="fill")+
-  ggtitle("Distributions of male and female subjects across simulated sites")
-```
+  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE & bigger_group == "cf.gamlss")
+nrow(ns_gam_gamlss_comp.z)
+```
+
+### Sex-Differences
 
@@
 
-skewed but ~ variances? sticking with t test of ranks. Also calculating the median z-score errors of males and females
+skewed but ~ variances, sticking with t test of ranks. Also calculating the median z-score errors of males and females
 
@@
 
-Get raw centiles and merge with centile differences
+Get raw z-scores and merge with z-score differences
 
@@
 ```
+
 merge into single dataframe
@@
 
-w/in feature sex-biases
+w/in feature sex-differences
 ```{r, message=FALSE}
@@
 
-## CENTILE FANS
+# CENTILE FANS
 
@@
   pivot_longer(cols = vol_list_global, values_to = "value", names_to="pheno")
-```
-
-```{r, eval=FALSE}
-global.cent.df <- centiles.df %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(!grepl('-refA', dataset))
-
-for (feature in unique(global.cent.df$pheno)) {
-
-  df <- global.cent.df %>%
-    dplyr::filter(pheno == feature)
-
- plot <- ggplot() +
-      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$M_centile_9, color=df$dataset)) +
-      labs(title=paste("Male", feature, "brain chart"))
-
- print(plot)
-
- plot <- ggplot() +
-      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$F_centile_9, color=df$dataset)) +
-      labs(title=paste("Female", feature, "brain chart"))
-
- print(plot)
-
- plot <- ggplot() +
-      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
-      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
-      geom_line(aes(x=df$age_days, y=df$centile_1, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_2, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_3, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_4, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_5, color=df$dataset), linewidth=1.5) +
-      geom_line(aes(x=df$age_days, y=df$centile_6, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_7, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_8, color=df$dataset)) +
-      geom_line(aes(x=df$age_days, y=df$centile_9, color=df$dataset)) +
-      labs(title=paste(feature, "brain chart"))
-
- print(plot)
-
-}
 ```
@@
 
-## FIGURE 1: METHODS
+# FIGURE 1: METHODS
 
@@
   theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(), axis.line.y = element_blank()) +
-  scale_x_continuous(labels=c('', '', '', 0, '', '', ''))
-```
-
-## FIGURE 2
-
-not sure whether to include
-```{r, eval=FALSE}
-#look at sex distributions of ppts across sites
-data.p <- subj_sum.df %>%
-  count(sim.site, sex) %>% 
-  group_by(sim.site) %>%
-  mutate(prop = n / sum(n)) %>%
-  ggplot(aes(x = sim.site, y = prop, fill = sex)) +
-  geom_col() +
-  theme_linedraw(base_size = 11) +
-  ggtitle("Distributions of male and female subjects across simulated sites")
-```
+  scale_x_continuous(labels=c('', 0, ''), breaks=c(-20, 0, -20))
+```
+
+# FIGURE 2
 

### 2024-05-17 cubic_scripts/qsub_basic_gamlss.sh

@@ -8,8 +8,6 @@
 #######################################################################################
 # SET PATHS
 base=/cbica/home/gardnerm/combat-biovar #base path (cubic)
-#base=/Users/megardn/Desktop/BGD_Repos/combat_biovar #base path (local)
-#data_csv=$base/data/ukb_CN_data_agefilt.csv #path to csv with CLEANED DATA (no duplicates, etc)
 pheno_path=$base/pheno_lists #path to .txt files listing phenotypes (global & regional)
 mod_script=$base/R_scripts/fit_basic_mod.R #path to .R script
 

### 2024-05-17 cubic_scripts/qsub_combat.sh

@@ -107,22 +107,22 @@ do
 	#COMBAT GAM
 	elif [ $config = "cf.gam" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gam, formula = y ~ s(age_days) + sexMale + s(sex.age)'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gam, formula = y ~ s(age_days) + sexMale'" > $bash_script
 
 	#COMBAT GAM W/ REF SITE
 	elif [ $config = "cf.gam_refA" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gam, formula = y ~ s(age_days) + sexMale + s(sex.age), ref.batch = \"Site_A\"'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gam, formula = y ~ s(age_days) + sexMale, ref.batch = \"Site_A\"'" > $bash_script
 
 	#COMBAT GAMLSS
 	elif [ $config = "cf.gamlss" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gamlss, formula = y ~ pb(age_days) + sexMale + pb(sex.age), sigma.formula = ~ pb(age_days) + sexMale + sex.age'" > $bash_script
+		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gamlss, formula = y ~ pb(age_days) + sexMale, sigma.formula = ~ age_days + sexMale'" > $bash_script
 
 	#COMBAT GAMLSS W/ REF SITE
 	elif [ $config = "cf.gamlss_refA" ]
 	then
-		echo "singularity run --cleanenv $img Rscript --save $cf_script $csv $batch $save_path $config $covar_list 'gamlss, formula = y ~ pb(age_days) + sexMale + pb(sex.age), sigma.formula = ~ pb(age_days) + sexMale + pb(sex.age), ref.batch = \"Site_A\"'" > $bash_script
[…]

### 2024-05-17 eda/ukb_curation-eda.Rmd
cleaning up

@@
 #table(cn_data$study)
-names(cn_data)
+#names(cn_data)
 ```
@@
          log_age = log(age_days, base=10),
-         TBV=(sGMV+GMV+WMV+Ventricles),
          fs_version = as.factor(fs_version),
@@
 saveRDS(ct_list, file=paste0(base_path,"cubic_scripts/R_scripts/CT_list.rds"))
-```
-
-Add brainwide phenotypes!
-Summing across regions for volumes (rather than summing GMV, sGMV, etc) for consistency with SA and CT
-
-```{r}
-vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
-
-ukb_df_totals <- ukb_df %>%
-  mutate(Vol_total=rowSums(dplyr::select(., .dots=all_of(c(vol_list_regions)))),
-         SA_total=rowSums(dplyr::select(., .dots=all_of(c(sa_list)))),
-         CT_total=rowSums(dplyr::select(., .dots=all_of(c(ct_list)))))
-
-#write data
-write_csv(ukb_df_totals, file=paste0(base_path,"data/ukb_CN_data.csv"))
 ```
@@
 ```{r}
-ukb_df_totals %>%
+ukb_df %>%
   dplyr::filter(age_yrs <= 49) %>%
@@
 
-ukb_df_totals %>%
+ukb_df %>%
   dplyr::filter(age_yrs >= 79) %>%
@@
 ```{r}
-ukb_df_age.filt <- ukb_df_totals %>%
+ukb_df_age.filt <- ukb_df %>%
   dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
@@
 
-nrow(ukb_df_totals) - nrow(ukb_df_age.filt)
+nrow(ukb_df) - nrow(ukb_df_age.filt)
 ```
@@
 
-```{r}
-ggplot(data=ukb_df_totals, aes(x=age_days, y=Vol_total)) +
-  geom_point(aes(color=sex), alpha=0.6, position="identity")
-
-ggplot(data=ukb_df_totals, aes(x=age_days, y=TBV)) +
-  geom_point(aes(color=sex), alpha=0.6, position="identity") #TBV includes ventricles, WMV, subcortical GMV
-
-ggplot(data=ukb_df_totals, aes(x=age_days, y=CT_total)) +
-  geom_point(aes(color=sex), alpha=0.6, position="identity")
-
-ggplot(data=ukb_df_totals, aes(x=age_days, y=SA_total)) +
-  geom_point(aes(color=sex), alpha=0.6, position="identity")
-```
-
 Look at variances
@@
 ```{r}
-ukb_centered <- ukb_df_totals %>%
+ukb_centered <- ukb_df %>%
   group_by(sex) %>%
@@
 #write out
-write_csv(ukb_df_sites, file=paste0(base_path, "data/ukb_to_model/ukb_CN_data_simsites.csv"))
+fwrite(ukb_df_sites, file=paste0(base_path, "data/ukb_to_model/ukb_CN_data_simsites.csv"))
 ```
@@
 ```
-
-Need to come up with cleaner way to do this...
-
-```{r}
-site_sum_df <- ukb_df_sites %>%
-  group_by(sim.site) %>%
-  summarise(proportion_male = mean(sex == "Male"),
-            min_age = min(age_days), max_age = max(age_days), mean_age = mean(age_days), 
-            var_gmv = var(GMV), var_wmv=var(WMV), var_sgmv=var(sGMV), var_ven = var(Ventricles),
-            sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))
-```
-
 Center and visualize distributions
@@
 ```
-
-##Exploratory QC
-
-Adding euler number (from jakob) back into combat-ed dfs so i can test how controlling for euler in centile scores changes the performance of combat configs (esp.
-in CT based on Bedford et al preprint).
-
-```{r}
-cn.df_euler <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)
-
-ukb.df_euler <- cn.df_euler %>%
-  dplyr::filter(study == "UKB") %>%
-  dplyr::select(participant, surfholes)
-
-table(unique(ukb_df_age.filt$participant) %in% unique(ukb.df_euler$participant)) #all subjects in df!
-
-fwrite(ukb.df_euler, file=paste0(base_path,"data/ukb_euler_nums.csv"))
-```

### 2024-05-17 eda/lifespan_eda.Rmd
cleaning up

@@
 
-## Site-Level Data
-
-look just at data w site info
-
-```{r}
-cn.df.site <- cn.df.eulerfilt2 %>% 
-  dplyr::filter(site != "NA") %>%
-  mutate(study = droplevels(study),
-         site = gsub(" ", "_", trimws(site))) %>%
-  unite(site, study, site, sep = "_", remove = FALSE) %>% #so sites with the same name don't get combined across studies!
-  dplyr::select(!c(INDEX.ID, country)) %>%#drop index ID for filtering
-  na.omit()
-
-n_unique(cn.df.site$participant)
-n_unique(cn.df.site$participant) == nrow(cn.df.site)
-
-max(cn.df.site$age_days)/365.25
-
-grep("scan", colnames(cn.df.site), value = TRUE) #no scanner info
-```
-
-```{r}
-# number studies
-length(levels(cn.df.site$study))
-
-#number sites per study
-cn.df.site %>%
-  group_by(study) %>% 
-  summarise(Result = length(unique(site)))
-```
-
-```{r}
-ggplot(cn.df.site) +
-  geom_boxplot(aes(x=reorder(study, age_days), y=age_days))+
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-ggplot(cn.df.site) +
-  geom_violin(aes(x=reorder(site, age_days), y=age_days, fill=study), scale="width") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
-```
-
-check for sex differences in sites
-
-```{r}
-ggplot(cn.df.site) +
-  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
-  scale_y_continuous(labels = scales::percent) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-ggplot(cn.df.site) +
-  geom_bar(aes(x=site, fill=sex), position="dodge") + 
-  theme(axis.text.x = element_blank()) +
-  facet_wrap(~study, scales="free")
-```
-
-Write out
-
-```{r}
-fwrite(cn.df.site, file=paste0(base_path, "/data/lifespan_CN_site-level_euler.csv"))
-```
-
-## Keep Study Level
 Keep studies w missing site info and just treat as single-site

### 2024-05-18 cubic_scripts/R_scripts/ranked_welch_tests.R
remove subject-level mean sex-bias tests

@@ -1,10 +1,11 @@
 #load libraries
 library(stats)
 
-#contains 2 functions for running welch's t-test on data rankings (accounting for both skew and unequal variances - Zimmerman & Zumbo, 1993)
+#contains 3 functions for running welch's t-test on data rankings (accounting for both skew and unequal variances - Zimmerman & Zumbo, 1993)
 
 # rank.welch.t.test() - analogous to t.test()
 # pairwise.rank.welch.t.test() - analogous to pairwise.t.test()
+# pairwise.rank.maxgrp() - also returns group with larger ranks
 
 rank.welch.t.test <- function (x, ...) {
   UseMethod("rank.welch.t.test", x)

### 2024-05-18 cubic_scripts/R_scripts/ratio_subject-wise_results.R

@@ -1,59 +1,0 @@
-# get results for varying-proportion analyses more quickly using cubic
-
-set.seed(12345)
-
-#LOAD PACKAGES
-library(dplyr)
-library(data.table)
-library(tidyverse)
-library(purrr)
-
-devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/stats_tests.R")
-
-#get args
-args <- commandArgs(trailingOnly = TRUE)
-data_path <- args[1] #path to csvs
-name_str <- args[2] #optional argument to specify "perm" (replications) instead of varying M:F props
-
-## LOAD DATA
-raw_files <- list.files(path = paste0(data_path,"/subject-wise"), pattern = "_subj_pred.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in raw_files) {
-  # Read each CSV file
-  data <- fread(file)
-  
-  data <- data %>%
-    mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm-|-", "", Source_File))
-  
-  # Bind the data to the combined dataframe
-  df_list <- c(df_list, list(data))
-  assign("ratio_subj_list", df_list)
-}
-print(paste("length:", length(ratio_subj_list)))
-
-
-if (is.na(name_str)) {
-  names_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
-  fdr_across <- length(ratio_list) #correct across permuted M:F ratios
-} else {
-  names_list <- c(paste(name_str, seq(1:length(ratio_subj_list)), sep="-"))
-  fdr_across <- 1 #don't correct across replications
-}
-
-### W/IN SUBJ MEAN ABS ERROR SEX-BIAS TESTS
-# centiles
-prop.sex.bias.cent_test <- lapply(ratio_subj_list, sex.bias.t.tests, to_test = "mean_cent_abs.diff", comp_multiplier=fdr_across)
-names(prop.sex.bias.cent_test) <- names_list
-sex_bias_ratio_cent_tests_df <- bind_rows(prop.sex.bias.cent_test, .id = "column_label")
-print(paste("centile df dim:", dim(sex_bias_ratio_cent_tests_df)))
-
[…]

### 2024-05-18 cubic_scripts/R_scripts/stats_tests.R

@@ -104,30 +104,6 @@ centile.t.tests.full_result <- function(df, feature_list, comp_multiplier=1){
   #detach()
 }
 
-#sex.bias.t.tests()
-# Within each combat config, test if there are significant differences in M and F subject's mean abs. centile error. Can test mean (not abs) centile errors, Z errors, etc, using `to_test` arg. FDR corrects for number of datasets tested
-
-sex.bias.t.tests <- function(df, to_test = "mean_cent_abs.diff", comp_multiplier=1){
-  
-  #conduct t test
-  df.sex.t <- df %>%
-    group_by(dataset) %>%
-    summarise(p.value = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$p.value,
-              sex_diff = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$estimate,
-              t.stat = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$statistic,
-              df = tidy(rank.welch.t.test.formula(formula=as.formula(paste(to_test, "~ sex")), paired = FALSE, p.adj = "none"))$parameter) %>%
-    ungroup()
-  
-  #apply fdr correction
-  n_comp=length(unique(df$dataset))
-  df.sex.t <- df.sex.t %>%
-    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = (n_comp*comp_multiplier)), #use comp_multiplier if using fun with lapply
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-  return(df.sex.t)
-  #detach()
-}
-
 # tidy(rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none"))
 # rank.welch.t.test.formula(formula=mean_cent_abs.diff ~ sex, data= ratio_subj_list[[1]], paired = FALSE, p.adj = "none")
 

### 2024-05-18 cubic_scripts/run_permutation_stats.sh

@@ -7,8 +7,6 @@
 # $csv_path/subject-wise/*_subj_pred.csv x # perms
 # $csv_path/*_featurewise_cent_t_tests.csv x2
 # $csv_path/*_featurewise_cent_z_tests.csv x2
-# $csv_path/subj.abs.mean_sex_bias_cent_t_tests.csv x1
-# $csv_path/subj.abs.mean_sex_bias_z_t_tests.csv x1
 # $csv_path/*_featurewise_cent_sex_bias_tests.RDS x2
 # $csv_path/*_featurewise_z_sex_bias_tests.RDS x2
 # $csv_path/*_no_ext.csv x # perms
@@ -41,8 +39,6 @@ rm -rf $csv_path/*_diffs.csv
 rm -rf $csv_path/subject-wise/*_subj_pred.csv
 rm -rf $csv_path/*_featurewise_cent_t_tests.csv
 rm -rf $csv_path/*_featurewise_cent_z_tests.csv
-rm -rf $csv_path/subj.abs.mean_sex_bias_cent_t_tests.csv
-rm -rf $csv_path/subj.abs.mean_sex_bias_z_t_tests.csv
 rm -rf $csv_path/*_featurewise_cent_sex_bias_tests.RDS
 rm -rf $csv_path/*_featurewise_z_sex_bias_tests.RDS
 rm -rf $csv_path/*_no_ext.csv
@@ -106,15 +102,6 @@ touch $bash_script
 echo "singularity run --cleanenv $img Rscript --save $r_script $csv_path 'full' 'perm' " > $bash_script
 ## qsub bash script
 qsub -N perm_cent_tests -o $bash_dir/cent_test_out.txt -e $bash_dir/cent_test_err.txt -l h_vmem=120G,s_vmem=120G $bash_script
-
-# sex-bias test of subj-means
-r_script=$r_base/ratio_subject-wise_results.R
-bash_script=$bash_dir/perm_subj-wise_sex_bias_test.sh 
-touch $bash_script
-
-echo "singularity run --cleanenv $img Rscript --save $r_script $csv_path 'perm'" > $bash_script
-## qsub bash script
-qsub -N perm_sub-wide_sex_tests -o $bash_dir/sub-wide_sex_test_out.txt -e $bash_dir/sub-wide_sex_test_err.txt -l h_vmem=120G,s_vmem=120G $bash_script
 
 # featurewise sex bias tests
[…]

### 2024-05-18 cubic_scripts/run_varying_ratios_stats.sh

@@ -7,8 +7,6 @@
 # $csv_path/subject-wise/*_subj_pred.csv x11
 # $csv_path/*_featurewise_cent_t_tests.csv x2
 # $csv_path/*_featurewise_cent_z_tests.csv x2
-# $csv_path/subj.abs.mean_sex_bias_cent_t_tests.csv x1
-# $csv_path/subj.abs.mean_sex_bias_z_t_tests.csv x1
 # $csv_path/*_featurewise_cent_sex_bias_tests.RDS x2
 # $csv_path/*_featurewise_z_sex_bias_tests.RDS x2
 # $csv_path/*_no_ext.csv x11
@@ -34,8 +32,6 @@ rm -rf $csv_path/*_diffs.csv
 rm -rf $csv_path/subject-wise/*_subj_pred.csv
 rm -rf $csv_path/*_featurewise_cent_t_tests.csv
 rm -rf $csv_path/*_featurewise_cent_z_tests.csv
-rm -rf $csv_path/subj.abs.mean_sex_bias_cent_t_tests.csv
-rm -rf $csv_path/subj.abs.mean_sex_bias_z_t_tests.csv
 rm -rf $csv_path/*_featurewise_cent_sex_bias_tests.RDS
 rm -rf $csv_path/*_featurewise_z_sex_bias_tests.RDS
 rm -rf $csv_path/*_no_ext.csv
@@ -99,15 +95,6 @@ touch $bash_script
 echo "singularity run --cleanenv $img Rscript --save $r_script $csv_path 'full'" > $bash_script
 ## qsub bash script
 qsub -N prop_cent_tests -o $bash_dir/cent_test_out.txt -e $bash_dir/cent_test_err.txt -l h_vmem=60.5G,s_vmem=60.0G $bash_script
-
-# sex-bias test of subj-means
-r_script=$r_base/ratio_subject-wise_results.R
-bash_script=$bash_dir/ratio_subj-wise_sex_bias_test.sh
-touch $bash_script
-
-echo "singularity run --cleanenv $img Rscript --save $r_script $csv_path" > $bash_script
-## qsub bash script
-qsub -N prop_sub-wide_sex_tests -o $bash_dir/sub-wide_sex_test_out.txt -e $bash_dir/sub-wide_sex_test_err.txt -l h_vmem=60.5G,s_vmem=60.0G $bash_script
 
 # featurewise sex bias tests
 ## based on `qsub_sex_bias_test.sh`
@@ -244,26 +231,6 @@ do
 	exit 2
     fi
[…]

### 2024-05-18 pull_stats.sh

@@ -31,9 +31,6 @@ then
 	echo "pulling *_featurewise_*_t_tests.csv"
 	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_featurewise_*_t_tests.csv' $base/ukb_ratio
 
-	echo "pulling subj.abs.mean_sex_bias_*_t_tests.csv"
-	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subj.abs.mean_sex_bias_*_t_tests.csv' $base/ukb_ratio
-
 	echo "pulling from ukb_ratios - subject-wise/*_subj_pred.csv"
 	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*_subj_pred.csv' $base/ukb_ratio
 
@@ -58,9 +55,6 @@ then
 
 	echo "pulling *_featurewise_*_t_tests.csv"
 	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/*_featurewise_*_t_tests.csv' $base/ukb_perm
-
-	echo "pulling subj.abs.mean_sex_bias_*_t_tests.csv"
-	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subj.abs.mean_sex_bias_*_t_tests.csv' $base/ukb_perm
 
 	echo "pulling from ukb_permute - subject-wise/*_subj_pred.csv"
 	scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_permute/perm_centile_csvs/subject-wise/*_subj_pred.csv' $base/ukb_perm

### 2024-05-18 paper_prop_results.Rmd

@@
 
-Test subj. mean abs errors (mean_cent_abs.diff) using `ratio_subject-wise_results.R`, saved as `subj.abs.mean_sex_bias_cent_t_tests.csv`
-
-```{r}
-cent.subj.mean.test.df <- fread(paste0(data_path, "/subj.abs.mean_sex_bias_cent_t_tests.csv"))
-cent.subj.mean.test.df <- cent.subj.mean.test.df %>%
-  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
-         dataset = as.factor(dataset))
-# cent.subj.mean.test.df
-```
-
-```{r}
-cent.subj.mean.test.df %>%
-  # filter(comp=="cf.gamlss_cf.gam") %>%
-  ggplot() +
-  geom_col(aes(x=prop, y=sig_fdr, fill=dataset), position="dodge") +
-  facet_wrap(~dataset) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
-```
-
 Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`
@@
 ### Sex-Biases
-Test subj. mean abs errors (mean_cent_abs.diff) using `ratio_subject-wise_results.R`, saved as `subj.abs.mean_sex_bias_z_t_tests.csv`
-
-```{r}
-z.subj.mean.test.df <- fread(paste0(data_path, "/subj.abs.mean_sex_bias_z_t_tests.csv"))
-z.subj.mean.test.df <- z.subj.mean.test.df %>%
-  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
-         dataset = as.factor(dataset))
-z.subj.mean.test.df
-```
-
-```{r}
-z.subj.mean.test.df %>%
-  # filter(comp=="cf.gamlss_cf.gam") %>%
-  ggplot() +
-  geom_col(aes(x=prop, y=sig_fdr, fill=dataset), position="dodge") +
-  facet_wrap(~dataset) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
-```
 

### 2024-05-20 paper_main_results.Rmd
updated, cleaned, and rerun

@@
 5. `gamlss_parse.sh` -> `/cubic_scripts/R_scripts/gamlss_parser.R`
+6. `qsub_summarize_subj_centiles.sh` -> `/cubic_scripts/R_scripts/summarise_cent_subj-wise.R`
 
@@
 #also load raw data
-ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")
+raw_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/UKB_CN_data_simsites.csv")
 
 #years for plotting
-ukb_cn_df <- ukb_cn_df %>%
+raw_df <- raw_df %>%
   mutate(age_years = age_days/365.25)
@@
 
-pulled variance from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_variance.csv' .`
+pulled variance from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_variance.csv' ./derivatives/ukb_basic`
 
 Get predicted centiles
-- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' .`
+- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' ./derivatives/ukb_basic`
 
@@
     mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
+    mutate(dataset = gsub("_data|_predictions.csv", "", Source_File))
 
@@
 
-#drop refA sites, not using for these analyses
-combined_preds <- combined_preds %>%
-  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
+unique(combined_preds$dataset)
+
+skim(combined_preds)
 ```
@@
 Get errors
-- pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' .`
+- pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' ./derivatives/ukb_basic`
 
@@
 
-#drop refA sites, not using for these analyses
-diffs.df <- diffs.df %>%
-  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
+sum(is.na(diffs.df))
 ```
@@
 Get subject-wise summaries
-- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' .`
+- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' ./derivatives/ukb_basic`
 
@@
 
-#drop refA sites, not using for these analyses
-subj_sum.df <- subj_sum.df %>%
-  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
+sum(is.na(subj_sum.df))
 ```
@@
 
-```{r, eval=FALSE}
+```{r, }
 full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
@@
   dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
   group_by(dataset) %>%
@@
 
-```{r, eval=FALSE}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+```{r}
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
@@
 cent.diffs.plot <- mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
+  dplyr::filter(dataset != "raw") %>%
   mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
@@
 ```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 
@@
   dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
          pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
@@
   #WMV
-geom_signif(y_position = c(29, 30.5, 4), xmin = c(0.7, 0.7, 1), 
+geom_signif(y_position = c(3, 3.5, 2.5), xmin = c(0.7, 0.7, 1),
               xmax = c(1, 1.3, 1.3), annotation = c("***","***", "***"),
@@
   #GMV
-  geom_signif(y_position = c(30, 31.5, 3), xmin = c(1.7, 1.7, 2), 
+  geom_signif(y_position = c(2.75, 3.25, 2.25), xmin = c(1.7, 1.7, 2),
               xmax = c(2, 2.3, 2.3), annotation = c("***","***", "***"),
@@
   #sGMV
-  geom_signif(y_position = c(28, 29.5, 2.8), xmin = c(2.7, 2.7, 3), 
+  geom_signif(y_position = c(2.7, 3.2, 2.2), xmin = c(2.7, 2.7, 3),
               xmax = c(3, 3.3, 3.3), annotation = c("***","***", "***"),
@@
   #Ventricles
-  geom_signif(y_position = c(24, 25.5, 7.5), xmin = c(3.7, 3.7, 4), 
+  geom_signif(y_position = c(7.1, 6.75, 6.45), xmin = c(3.7, 3.7, 4),
               xmax = c(4, 4.3, 4.3), annotation = c("***","***", "***"),
@@
   pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
          abs.diff = abs.diff*100,
@@
 ggsave("figs/all_abs_err_plt_wlabels.jpeg", long_abs_diff.plt, width=13, height=10.5, units="in")
-
-long_abs.diffs %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  ggplot(aes(pheno, abs.diff)) +
-  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=abs.diff,  fill=dataset, color=dataset), scale="count") +
-  theme_linedraw(base_size = 11) +
-  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Absolulte Centile Error", x = "Feature") +
-  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top")
-
-# ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
 ```
@@
   dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
   group_by(dataset) %>%
@@
 
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 #var plot
@@
   dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
          median_centile_abs.diff = median_centile_abs.diff*100)
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 7.5)) #
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 1)) #
 # med.diffs.plot
@@
   dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
          median_centile_abs.diff = median_centile_abs.diff*100)
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 7.5))
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 1))
 # med.diffs.plot.global
@@
   )
-
-# p_diff_annotate <- annotate_figure(p_diff)
-# p_diff_annotate
-```
-
-Quick comp of cf.gam and cf.gamlss
-
-```{r}
-ns_gam_gamlss_comp <- full_cent_res %>%
-  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE & bigger_group == "cf.gamlss")
-nrow(ns_gam_gamlss_comp)
-```
-
-**21 regions where ComBat GAM is better than ComBat-LS! All cortical thickness -> 206-21 = 185 regions where ComBat-LS fits better.
+```
+
+Count of instances where cf.gamlss has higher errors
+
+```{r}
+full_cent_res %>%
+  dplyr::filter(sig_fdr == TRUE & bigger_group == "cf.gamlss") %>%
+  group_by(comp) %>%
+  summarise(n_regions=sum(sig_fdr))
+```
+
+13 regions where ComBat GAM is better than ComBat-LS, 13 regions where linear ComBat is better than ComBat-LS, 1 region where combat w/o covars is better: ALL cortical thickness
 
@@
 check for skew (welch t.test vs ranked welch t test)
-```{r, eval=FALSE}
+```{r}
 for (pheno in pheno_list){
@@
 
-```{r, eval=FALSE}
+```{r}
 for (cf in c("cf.gam", "cf.gamlss")) {
@@
 
-```{r, message=FALSE, eval=FALSE}
+```{r, message=FALSE, }
 sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list)
@@
     as.data.frame()
-```
-
-```{r,eval=FALSE}
-#view combat gam and combat LS ns
-ns.sex_diff_filt.gam <- sex_t_tests_in_feat.df %>%
-  dplyr::filter(sig_fdr==FALSE & dataset =="cf.gam") %>%
-  arrange(pheno)
-ns.sex_diff_filt.gamlss <- sex_t_tests_in_feat.df %>%
-  dplyr::filter(sig_fdr==FALSE & dataset =="cf.gamlss") %>%
-  arrange(pheno)
 ```
@@
 ```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 #var plot
@@
   dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                                     TRUE ~ median_sex_diff*100))
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = median_sex_diff*100)
+#find range
+max(sex.var.plot.df$median_sex_diff, na.rm=TRUE)
+min(sex.var.plot.df$median_sex_diff, na.rm=TRUE)
 
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-6, 6)) #
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.25, 0.25)) #
 saveRDS(sex.var.plot, file="figs/sex_var_plot.RDS")
@@
   dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
   dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
@@
                                             TRUE ~ NA)),
-        median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>% #drop NS values
+        median_sex_diff = median_sex_diff*100) %>% 
   group_by(dataset)
 
+#color range
+max(sex.var.plot.global.df$median_sex_diff, na.rm=TRUE)
+min(sex.var.plot.global.df$median_sex_diff, na.rm=TRUE)
+
 tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
-  dplyr::filter(dataset=="ukb_CN")
+  dplyr::filter(dataset=="raw")
 tp$sex_standard <- 1
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork",  limits = c(-6, 6), name = "Male - Female Median \nCentile Error") #,
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.25, 0.25), name = "Male - Female Median \nCentile Error") #,
 # sex.var.plot.global
@@
 raw_centiles <- combined_preds %>%
-  dplyr::filter(dataset == "ukb_CN")
+  dplyr::filter(dataset == "raw")
+stopifnot(nrow(raw_centiles)>10)
+sum(is.na(raw_centiles))
 
@@
     group_by(participant) %>% 
-    dplyr::filter(!any((get(pheno) < 0.05 | get(pheno) > 0.95) & dataset == "ukb_CN")) %>% #drop ppts if raw centile is <0.05 or >0.95
+    dplyr::filter(!any((get(pheno) < 0.05| get(pheno) > 0.95) & dataset == "raw")) %>% #drop ppts if raw centile is <0.05 or >0.95
     ungroup() %>%
-    dplyr::filter(dataset != "ukb_CN") #now drop raw vals
+    dplyr::filter(dataset != "raw") #now drop raw vals
   df_list[count] <- list(df)
@@
 no_ext <- Reduce(full_join, df_list)
+sum(is.na(no_ext))
 
@@
 dim(no_ext)
-# #make sure no cols are lost
-# ncol(no_ext) == ncol(merge.df) #lost File_ID col, not going to bother rerunning rn
-# nrow(no_ext)
+n_unique(no_ext$participant)
+n_unique(merge.df$participant)
 ```
@@
 centile error diffs
-```{r, eval=FALSE}
+```{r, eval=FALSE, message=FALSE}
 noext_cent_res <- centile.t.tests.full_result(no_ext, feature_list=pheno_abs.diff.list)
@@
             min_pval=min(p.val_fdr),
-            max_pval=max(p.val_fdr)) %>%
-  dplyr::filter(!grepl('_refA', comp))#drop refA levels
+            max_pval=max(p.val_fdr))
+
+full_cent_res %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig_regions = sum(sig_fdr),
+            min_pval=min(p.val_fdr),
+            max_pval=max(p.val_fdr))
 
@@
   dplyr::filter(sig_fdr==TRUE) %>%
-  dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
   ggplot() +
@@
 ```{r, eval=TRUE}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
@@
 ```
-```{r, eval=FALSE}
+```{r}
 no.ext_mean.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
+  dplyr::filter(dataset != "raw") %>%
   mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
@@
 w/in feature sex-differences
-```{r, message=FALSE, eval=FALSE}
+```{r, message=FALSE, }
 noext.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(no_ext, feature_list=pheno_diff_list)
@@
   dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
          median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
@@
 check for skew (welch t.test vs ranked welch t test)
-```{r, eval=FALSE}
+```{r}
 for (val in z.pheno_abs.diff.list){
@@
 
-```{r, eval=FALSE}
+```{r}
 full_z_res <- centile.t.tests.full_result(diffs.df, feature_list=z.pheno_abs.diff.list)
@@
   dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
   group_by(dataset) %>%
@@
 
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
@@
 z.diffs.plot <- mean.z.diffs.long %>%
-  dplyr::filter(dataset != "ukb_CN") %>%
+  dplyr::filter(dataset != "raw") %>%
   ggplot() +
@@
 
-Quick comp of cf.gam and cf.gamlss
-
-```{r}
-ns_gam_gamlss_comp.z <- full_z_res %>%
-  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE & bigger_group == "cf.gamlss")
-nrow(ns_gam_gamlss_comp.z)
+N features where ComBatLS has higher errors
+
+```{r}
+full_z_res %>%
+  dplyr::filter(sig_fdr == TRUE & bigger_group == "cf.gamlss") %>%
+  group_by(comp) %>%
+  summarise(n_regions=sum(sig_fdr))
 ```
@@
 
-Significance testing (w/in sub. mean abs z error)
-
-```{r}
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_diff")
-sex.bias.t.tests(subj_sum.df, to_test="mean_z_abs.diff")
-```
-
 Testing for significant M-F differences in z score errors WITHIN each feature!
@@
 check for skew (welch t.test vs ranked welch t test)
-```{r, eval=FALSE}
+```{r}
 for (val in z.pheno_diff_list){
@@
 
-```{r, message=FALSE, eval=FALSE}
+```{r, message=FALSE}
 z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=z.pheno_diff_list)
@@
     group_by(participant) %>% 
-    dplyr::filter(!any((get(pheno) < -2 | get(pheno) > 2) & dataset == "ukb_CN")) %>% #drop ppts if raw z score is < -2 or > 2
+    dplyr::filter(!any((get(pheno) < -2 | get(pheno) > 2) & dataset == "raw")) %>% #drop ppts if raw z score is < -2 or > 2
     ungroup() %>%
-    dplyr::filter(dataset != "ukb_CN") #now drop raw vals
+    dplyr::filter(dataset != "raw") #now drop raw vals
   z.df_list[count] <- list(df)
@@
 dim(z.no_ext)
+dim(merge.df)
+
+n_unique(merge.df$participant)
+n_unique(z.no_ext$participant)
+sum(is.na(z.no_ext))
 ```
@@
 z error diffs
-```{r, eval=FALSE}
+```{r}
 noext_z_res <- centile.t.tests.full_result(z.no_ext, feature_list=z.pheno_abs.diff.list)
@@
             min_pval=min(p.val_fdr),
-            max_pval=max(p.val_fdr)) %>%
-  dplyr::filter(!grepl('_refA', comp))#drop refA levels
+            max_pval=max(p.val_fdr))
+
+full_z_res %>%
+  group_by(comp, bigger_group) %>%
+  summarize(n_sig_regions = sum(sig_fdr),
+            min_pval=min(p.val_fdr),
+            max_pval=max(p.val_fdr))
 
@@
 ```{r}
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")
@@
   dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
-  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
   group_by(dataset) %>%
@@
   pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno=gsub(".z", "", pheno)) %>%
   mutate(pheno_cat = as.factor(case_when(
@@
 z.no.ext_merge <- full_join(z.no.ext_mean.diffs.long, mean.z.diffs.long)
+
+z.no.ext_merge %>%
+  dplyr::filter(pheno=="GMV" | pheno=="GMV.z")
 
@@
 
-pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' .`
+pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' ./derivatives/ukb_basic`
 
@@
     data <- data %>%
-     mutate(dataset = gsub("-data", "", dataset))
+     mutate(dataset = gsub("-data", "", dataset),
+            dataset=gsub("ukb-CN-simsites-", "", dataset))
 
@@
 }
-# length(basic_centiles)
+# length(basic_centiles
 centiles.df <- bind_rows(basic_centiles)
@@
 #get raw data points for visualization
-global.df.long <- ukb_cn_df %>%
+global.df.long <- raw_df %>%
   dplyr::select(participant, age_days, sex, vol_list_global) %>%
@@
   dplyr::filter(pheno == "WMV") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
+  dplyr::filter(dataset == "raw" | dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
   #pivot_longer(cols=contains("centile"), name)  %>%
@@
 ```{r}
-dataset_names <- c('ukb-CN-simsites' = "True Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c('raw' = "True Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 
 tmp_labels <- wmv.df %>%
-  dplyr::filter(age_days == max(age_days) & dataset=="ukb-CN-simsites") %>%
+  dplyr::filter(age_days == max(age_days) & dataset=="raw") %>%
   pivot_longer(cols=starts_with("centile"), values_to = "value", names_to="centile_num", names_prefix = "centile_") %>%
@@
 wmv_centile_fan <- ggplot(wmv.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_years, y=ukb_cn_df$WMV), alpha=0.4, width=.2, color="lightgray") +
+  geom_jitter(data = raw_df, aes(x=raw_df$age_years, y=raw_df$WMV), alpha=0.4, width=.2, color="lightgray") +
   #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
@@
   facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
+  scale_linetype_manual(values = c('raw'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "longdash"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c('raw'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
   geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile)), nudge_x = 0.35, size=2.5) +
@@
   dplyr::filter(pheno == "Ventricles") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss"| dataset == "cf.lm") %>%
+  dplyr::filter(dataset == "raw" | dataset == "cf.gam" | dataset == "cf.gamlss"| dataset == "cf.lm") %>%
   #pivot_longer(cols=contains("centile"), name)  %>%
@@
                                    sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ),
+         dataset = fct_relevel(dataset, "raw", after = Inf ),
          age_years = age_days/365.25)
 tmp_labels <- ven.df %>%
-  dplyr::filter(age_years == max(age_years) & dataset=="ukb-CN-simsites") %>%
+  dplyr::filter(age_years == max(age_years) & dataset=="raw") %>%
   pivot_longer(cols=starts_with("centile"), values_to = "value", names_to="centile_num", names_prefix = "centile_") %>%
@@
 ven_plot <- ggplot(ven.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_years, y=ukb_cn_df$Ventricles),  alpha=0.4, width=.2, color="lightgray") +
+  geom_jitter(data = raw_df, aes(x=raw_df$age_years, y=raw_df$Ventricles),  alpha=0.4, width=.2, color="lightgray") +
   #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
@@
   facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
+  scale_linetype_manual(values = c('raw'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "longdash"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #scale_color_manual(values = c('raw'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_color_manual(values = c('raw'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
   theme_linedraw(base_size = 11) +
-  geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile))) +
+  geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile)), nudge_x = -0.75, size=3) +
   xlab("Age (days)") + ylab("Volume (mm3)") +
@@
   facet_wrap(~dataset, nrow=1, labeller=labeller(dataset_names)) + 
-  # scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
+  # scale_linetype_manual(values = c('raw'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #scale_color_manual(values = c('raw'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_color_manual(values = c('raw'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
   xlab("Age (yrs)") + ylab("Volume (mm3)") +
@@
   dplyr::filter(pheno == "rh_SA_postcentral") %>%
-  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  dplyr::filter(dataset == "raw" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
   #pivot_longer(cols=contains("centile"), name)  %>%
@@
                                    sex == "F" ~ "Female")),
-         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))
+         dataset = fct_relevel(dataset, "raw", after = Inf ))
 
 ggplot(r.sa.pc.df) +
-  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$rh_SA_postcentral), alpha=0.7, width=150, color="lightgray") +
+  geom_jitter(data = raw_df, aes(x=raw_df$age_days, y=raw_df$rh_SA_postcentral), alpha=0.7, width=150, color="lightgray") +
   #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
@@
   facet_wrap(~sex) + 
-  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
-  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
+  scale_linetype_manual(values = c('raw'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  #scale_color_manual(values = c('raw'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
+  scale_color_manual(values = c('raw'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
   theme_minimal() +
@@
 fig_2 <- ggarrange(
-  wmv_centile_fan,
+  ven_plot,
   ggarrange(global_cent_errs_plot, p_diff, ncol=1, nrow=2, heights = c(1.2, 2), common.legend = FALSE, labels=c(NA, "C"), vjust = c(-0.25)),

### 2024-05-20 paper_main_results.Rmd
don't rerun stats, use saved rds objs

@@
 
-```{r, }
+```{r, eval=FALSE}
 full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
@@
 
-```{r, message=FALSE, }
+```{r, message=FALSE, eval=FALSE}
 sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list)
@@
 w/in feature sex-differences
-```{r, message=FALSE, }
+```{r, message=FALSE, eval=FALSE}
 noext.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(no_ext, feature_list=pheno_diff_list)
@@
 
-```{r}
+```{r, eval=FALSE}
 full_z_res <- centile.t.tests.full_result(diffs.df, feature_list=z.pheno_abs.diff.list)
@@
 
-```{r, message=FALSE}
+```{r, message=FALSE, eval=FALSE}
 z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=z.pheno_diff_list)
@@
 z error diffs
-```{r}
+```{r, eval=FALSE}
 noext_z_res <- centile.t.tests.full_result(z.no_ext, feature_list=z.pheno_abs.diff.list)
@@
 w/in feature sex-differences
-```{r, message=FALSE}
+```{r, message=FALSE, eval=FALSE}
 noext.z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(z.no_ext, feature_list=z.pheno_diff_list)

### 2024-05-20 paper_lbcc_results.Rmd
exploratory sigma plots

@@
 
+load ComBat-harmonized data
+```{r}
+cf.lm.data <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.lm_batch.study_data.csv"))
+gmv.cf.lm.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.lm-batch.study-data_no.tbv_mod.rds"))
+
+gmv.cf.lm.pred.df <- pred.across.ages.lbcc(gmv.cf.lm.mod, cf.lm.data)
+
+gmv.cf.lm.pred.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma), color="red") +
+  geom_line(aes(x=age_days, y=m.sigma), color="blue")
+
+gmv.cf.lm.mod.site <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.lm-batch.study-data_site.est_mod.rds"))
+
+gmv.cf.lm.pred.site.df <- pred.across.ages.lbcc.study(gmv.cf.lm.mod.site, cf.lm.data)
+
+gmv.cf.lm.pred.site.df %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma), color="red") +
+  geom_line(aes(x=age_days, y=m.sigma), color="blue")
+```
+
+```{r}
+gmv.cf.lm.pred.site.df <- gmv.cf.lm.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combat")
+gmv.cf.lm.pred.df <- gmv.cf.lm.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combat")
+
+gmv.cf.lm.pred.df_full <- full_join(gmv.cf.lm.pred.site.df, gmv.cf.lm.pred.df)
+
+gmv.cf.lm.pred.df_full %>%
+  ggplot() +
+  geom_line(aes(x=age_days, y=f.sigma, color=study_incl))
+```
+
 load ComBat-GAM-harmonized data
@@
 
+gmv.cf.gamlss.mod.site$sigma.coefficients
+
 gmv.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(gmv.cf.gamlss.mod.site, cf.gamlss.data)
@@
 ```{r}
-all.pred.df <- reduce(list(gmv.cf.gam.pred.df_full, gmv.cf.gamlss.pred.df_full, gmv.raw.pred.site.df, gmv.raw.pred.df), full_join)
+all.pred.df <- reduce(list(gmv.cf.gam.pred.df_full, gmv.cf.lm.pred.df_full, gmv.cf.gamlss.pred.df_full, gmv.raw.pred.site.df, gmv.raw.pred.df), full_join)
 
@@
 
+#combat
+wmv.cf.lm.mod <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.lm-batch.study-data_no.tbv_mod.rds"))
+wmv.cf.lm.pred.df <- pred.across.ages.lbcc(wmv.cf.lm.mod, cf.lm.data)
+
+wmv.cf.lm.mod.site <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.lm-batch.study-data_site.est_mod.rds"))
+wmv.cf.lm.pred.site.df <- pred.across.ages.lbcc.study(wmv.cf.lm.mod.site, cf.lm.data)
+
+wmv.cf.lm.pred.site.df <- wmv.cf.lm.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combat")
+wmv.cf.lm.pred.df <- wmv.cf.lm.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combat")
+
 #combat-gam
@@
 ```{r}
-wmv.all.pred.df <- reduce(list(wmv.raw.pred.site.df, wmv.raw.pred.df, wmv.cf.gam.pred.site.df, wmv.cf.gam.pred.df, wmv.cf.gamlss.pred.site.df, wmv.cf.gamlss.pred.df), full_join)
+wmv.all.pred.df <- reduce(list(wmv.raw.pred.site.df, wmv.raw.pred.df, wmv.cf.gam.pred.site.df, wmv.cf.gam.pred.df, wmv.cf.gamlss.pred.site.df, wmv.cf.gamlss.pred.df, wmv.cf.lm.pred.site.df, wmv.cf.lm.pred.df), full_join)
 
@@
 
+#combat
+sgmv.cf.lm.mod <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.lm-batch.study-data_no.tbv_mod.rds"))
+sgmv.cf.lm.pred.df <- pred.across.ages.lbcc(sgmv.cf.lm.mod, cf.lm.data)
+
+sgmv.cf.lm.mod.site <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.lm-batch.study-data_site.est_mod.rds"))
+sgmv.cf.lm.pred.site.df <- pred.across.ages.lbcc.study(sgmv.cf.lm.mod.site, cf.lm.data)
+
+sgmv.cf.lm.pred.site.df <- sgmv.cf.lm.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combat")
+sgmv.cf.lm.pred.df <- sgmv.cf.lm.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combat")
+
 #combat-gam
@@
 ```{r}
-sgmv.all.pred.df <- reduce(list(sgmv.raw.pred.site.df, sgmv.raw.pred.df, sgmv.cf.gam.pred.site.df, sgmv.cf.gam.pred.df, sgmv.cf.gamlss.pred.site.df, sgmv.cf.gamlss.pred.df), full_join)
+sgmv.all.pred.df <- reduce(list(sgmv.raw.pred.site.df, sgmv.raw.pred.df, sgmv.cf.gam.pred.site.df, sgmv.cf.gam.pred.df, sgmv.cf.gamlss.pred.site.df, sgmv.cf.gamlss.pred.df, sgmv.cf.lm.pred.site.df, sgmv.cf.lm.pred.df), full_join)
 
@@
 ven.raw.pred.df <- pred.across.ages.lbcc(ven.raw.mod, raw.data)
-# ven.raw.mod.site <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler_raw_site.est_mod.rds"))
-# ven.raw.pred.site.df <- pred.across.ages.lbcc.study(ven.raw.mod.site, raw.data)
-
-# ven.raw.pred.site.df <- ven.raw.pred.site.df %>%
-#   mutate(study_incl = TRUE,
-#          combat= "raw")
+
 ven.raw.pred.df <- ven.raw.pred.df %>%
@@
          combat= "raw")
+
+#combat
+ven.cf.lm.mod <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.lm-batch.study-data_no.tbv_mod.rds"))
+ven.cf.lm.pred.df <- pred.across.ages.lbcc(ven.cf.lm.mod, cf.lm.data)
+
+ven.cf.lm.mod.site <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.lm-batch.study-data_site.est_mod.rds"))
+ven.cf.lm.pred.site.df <- pred.across.ages.lbcc.study(ven.cf.lm.mod.site, cf.lm.data)
+
+ven.cf.lm.pred.site.df <- ven.cf.lm.pred.site.df %>%
+  mutate(study_incl = TRUE,
+         combat= "combat")
+ven.cf.lm.pred.df <- ven.cf.lm.pred.df %>%
+  mutate(study_incl = FALSE,
+         combat= "combat")
 
@@
 ```{r}
-ven.all.pred.df <- reduce(list(ven.raw.pred.df, ven.cf.gam.pred.site.df, ven.cf.gam.pred.df, ven.cf.gamlss.pred.site.df, ven.cf.gamlss.pred.df), full_join) #ven.raw.pred.site.df
+ven.all.pred.df <- reduce(list(ven.raw.pred.df, ven.cf.gam.pred.site.df, ven.cf.gam.pred.df, ven.cf.gamlss.pred.site.df, ven.cf.gamlss.pred.df, ven.cf.lm.pred.site.df, ven.cf.lm.pred.df), full_join) #ven.raw.pred.site.df
 
@@
   ggtitle("Sample distribution")
+```
+
+Get predictions for sex-effects
+
+```{r}
+mean(comfam.cf.gam_VolGlob$estimates$gamma.star, na.rm=TRUE)
+mean(comfam.cf.gamlss_VolGlob$estimates$gamma.star, na.rm=TRUE)
+mean(comfam.cf.gam_VolReg$estimates$gamma.star, na.rm=TRUE)
+mean(comfam.cf.gamlss_VolReg$estimates$gamma.star, na.rm=TRUE)
+mean(comfam.cf.gam_CT$estimates$gamma.star, na.rm=TRUE)
+mean(comfam.cf.gamlss_CT$estimates$gamma.star, na.rm=TRUE)
+mean(comfam.cf.gam_SA$estimates$gamma.star, na.rm=TRUE)
+mean(comfam.cf.gamlss_SA$estimates$gamma.star, na.rm=TRUE)
+
+mean(abs(comfam.cf.gam_SA$estimates$delta.star - 1), na.rm=TRUE)
+mean(abs(comfam.cf.gamlss_SA$estimates$delta.star - 1), na.rm=TRUE)
+mean(abs(comfam.cf.gam_CT$estimates$delta.star - 1), na.rm=TRUE)
+mean(abs(comfam.cf.gamlss_CT$estimates$delta.star - 1), na.rm=TRUE)
+mean(abs(comfam.cf.gam_VolReg$estimates$delta.star - 1), na.rm=TRUE)
+mean(abs(comfam.cf.gamlss_VolReg$estimates$delta.star - 1), na.rm=TRUE)
+mean(abs(comfam.cf.gam_VolGlob$estimates$delta.star - 1), na.rm=TRUE)
+mean(abs(comfam.cf.gamlss_VolGlob$estimates$delta.star - 1), na.rm=TRUE)
 ```

### 2024-05-23 paper_main_results.Rmd
cleaning and updating

@@
   theme_linedraw(base_size = 11) +
+  theme(plot.title = element_text(size=12), legend.position="bottom") +
   geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile)), nudge_x = -0.75, size=3) +
@@
   labels=c("A", "B"),
-  widths=c(1.1, 1)
+  widths=c(1.2, 1)
   #hjust=0.01,

### 2024-05-23 paper_prop_results.Rmd
cleaning and updating

@@
 source("./cubic_scripts/R_scripts/stats_tests.R")
-#aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
 ```
@@
 ```{r}
-dataset_names <- c('ukb-CN-simsites' = "Raw Data",  cf = "ComBat w/o cov", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c('raw' = "Raw Data",  cf = "ComBat w/o cov", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
@@
 
-```{r, eval=FALSE}
+```{r}
 ratio.df %>%
@@
 Read in predicted centiles & z scores FROM RAW DATA ONLY (to save memory)
-```{r, eval=FALSE}
+```{r}
 pred.csvs <- list.files(path = data_path, pattern = "raw_predictions.csv", full.names = TRUE)
@@
 Visualize sex-distribution across sites/proportions
-
-```{r, eval=FALSE}
-percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))
-
-ratio.df %>%
-  mutate(sex_site = paste(sex, sim.site, sep=".")) %>%
-  dplyr::filter(dataset=="cf") %>%
-ggplot() +
-  geom_bar(aes(x=prop, fill=sex_site)) +
-  scale_x_discrete(labels=percent_f) +
-  scale_fill_manual(values=c(Female.Balanced = "#FCC3BF", Male.Balanced = "#ADFDFF", Female.Imbalanced ="#F8766D", Male.Imbalanced="#00BFC4")) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  geom_hline(yintercept = (18800/2), linetype="dashed")+
-  theme_linedraw()
-```
 
@@
 ```{r}
-cent.t.test.df_global <- fread(paste0(data_path, "/full_featurewise_cent_t_tests_global.csv"))
+cent.t.test.list <- readRDS(paste0(data_path, "/full_featurewise_cent_t_tests_all_out.RDS"))
 
-cent.t.test.df_global_filt <- cent.t.test.df_global %>%
-  dplyr::filter(comp %in% c("cf.gamlss_cf.gam", "cf.lm_cf.gam", "cf.lm_cf.gamlss"))
+cent.t.test.list_global <- lapply(cent.t.test.list, function(x) filter(x, pheno %in% vol_list_global))
+cent.t.test_global <- bind_rows(cent.t.test.list_global, .id = "prop") %>%
+  dplyr::mutate(prop = case_when(prop == "0M:10F" ~ "prop-01", 
+                                 prop == "1M:9F" ~ "prop-02", 
+                                 prop == "2M:8F" ~ "prop-03", 
+                                 prop == "3M:7F" ~ "prop-04", 
+                                 prop == "4M:6F" ~ "prop-05", 
+                                 prop == "5M:5F" ~ "prop-06", 
+                                 prop == "6M:4F" ~ "prop-07", 
+                                 prop == "7M:3F" ~ "prop-08", 
+                                 prop == "8M:2F" ~ "prop-09", 
+                                 prop == "9M:1F" ~ "prop-10", 
+                                 prop == "10M:0F" ~ "prop-11",
+                                 TRUE ~ NA_character_))
 
-cent.t.test.df_global_filt %>%
-  dplyr::filter(sig_fdr == FALSE)
+cent.t.test.df_global_filt <- cent.t.test_global %>%
+  dplyr::filter(comp %in% c("cf.gamlss_cf.gam", "cf.lm_cf.gam", "cf.lm_cf.gamlss")) %>%
+  dplyr::mutate(comp = factor(comp, ordered=TRUE, levels=c("cf.lm_cf.gam", "cf.lm_cf.gamlss", "cf.gamlss_cf.gam")))
 ```
@@
 
-# mean_comp_tbl <- ggpubr::compare_means(abs.diff ~ dataset, data=prop.global_abs.diffs, group.by = c("pheno", "prop")) 
-# unique(mean_comp_tbl$group1)
-# unique(mean_comp_tbl$group2)
+#wilcox test for visualizing significance
 mean_comp_tbl_calc <- prop.global_abs.diffs %>%
@@
                        p.adjust.method = "fdr") 
+```
 
+```{r}
 mean_comp_tbl <- mean_comp_tbl_calc %>%
-  mutate(comp = factor(paste(group1, group2, sep="_"), ordered=TRUE, levels=c("cf.gamlss_cf.lm", "cf.gam_cf.lm", "cf.gam_cf.gamlss"))) %>%
+  mutate(comp = factor(paste(group2, group1, sep="_"), ordered=TRUE, levels=c("cf.lm_cf.gam", "cf.lm_cf.gamlss", "cf.gamlss_cf.gam"))) %>%
   arrange(prop, comp) %>%
@@
 
-#check ns results
-mean_comp_tbl %>%
-  dplyr::filter(p.adj.signif == "ns")
-
-#get max w/in each pheno, prop & dataset
+#get max w/in each pheno, prop & dataset for placement
 max_abs.errs <- prop.global_abs.diffs %>%
@@
 
-mean_comp_tbl3 <- mean_comp_tbl2 %>%
-  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*"),
+#just use significance from actual stats tests
+
+stats_centile_for_plot <- inner_join(mean_comp_tbl2, cent.t.test.df_global_filt)
+#kinda hacky way of fixing positions
+stats_centile_for_plot.corrected <- stats_centile_for_plot %>%
+  mutate(p.adj.signif = ifelse(sig_fdr == "FALSE", "ns", "*"),
          xmax=case_when(group1=="cf.gamlss" | group2== "cf.gamlss" ~ (x+.266667),
@@
                         TRUE ~  x),
-         y.position=case_when(comp == "cf.gamlss_cf.lm" ~ max_cf.lm + 3.3,
-                              comp == "cf.gam_cf.lm" ~ max_cf.lm + 2,
-                              TRUE ~ max_cf.gam + 1.8))
-mean_comp_tbl3[(mean_comp_tbl3$pheno=="WMV" & mean_comp_tbl3$group1=="cf.gam"), "p.adj.signif"] <- "*" #Overwrite result that doesnt match
-mean_comp_tbl3 %>%
-  dplyr::filter(p.adj.signif == "ns")
+         y.position=case_when(comp == "cf.lm_cf.gamlss" ~ max_cf.lm + 1,
+                              comp == "cf.lm_cf.gam" ~ max_cf.lm + .5,
+                              TRUE ~ max_cf.gam + 0.4))
+
+cent.t.test.df_global_filt %>%
+  dplyr::select(prop, comp, pheno, sig_fdr) %>%
+  dplyr::filter(sig_fdr == FALSE)
 ```
-results match! just gotta proceed as is
 
@@
   dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
          pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE)) %>%
@@
   scale_x_discrete(labels=percent_f) +
-  stat_pvalue_manual(mean_comp_tbl3, label = "p.adj.signif", label.size = 2, tip.length = 0) +
+  stat_pvalue_manual(stats_centile_for_plot.corrected, label = "p.adj.signif", label.size = 2, tip.length = 0) +
   # ggsignif::geom_signif(data=mean_comp_tbl2) +
@@
 
-Also visualize by sex
-```{r, eval=FALSE}
-prop.global_cent_errs_plot_bysex <- prop.global_abs.diffs %>%
+Visualize by sex
+
+```{r}
+prop.global_abs.diffs %>%
   dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
-         abs.diff = abs.diff*100) %>% #for visuals
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE)) %>%
   ggplot() +
@@
   scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Site") +
+  labs(y="Absolute Centile Error", x = "% Females in Imbalanced Site") +
   theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
   scale_x_discrete(labels=percent_f) +
-  facet_grid(sex ~ pheno) +
-  ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
-# prop.global_cent_errs_plot_bysex
+  facet_grid(sex ~ pheno)
 ```
@@
 
-visualize magnitude of sex-biases (directional magnitude of errors within pheno cat)
+visualize actual (not ABSOLUTE) errors for global phenos by sex
+
+add significance testing
 
 ```{r}
-cent.feat.sex.test.df <- cent.feat.sex.test.df %>%
-  mutate(pheno_cat = factor(case_when(
-    pheno %in% vol_list_global ~ "Global Volume",
-    pheno %in% vol_list_regions ~ "Regional Volume",
-    pheno %in% sa_list ~ "Regional SA",
-    pheno %in% ct_list ~ "Regional CT",
-    TRUE ~ NA_character_)))
+cent.feat.test_global <- cent.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat == "Global Volume" & (dataset == "cf.gam" | dataset =="cf.gamlss"))
 
-cent.feat.sex.test.df %>%
-  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              TRUE ~ median_sex_diff*100)) %>%
+#wilcox test for visualizing significance
+cent.sex_comp_tbl_calc <- ratio.df %>%
+  dplyr::select(participant, sex, prop, dataset, paste0("diff_", vol_list_global)) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
+  group_by(dataset, pheno, prop) %>%
+  wilcox_test(diff ~ sex)
+
+cent.feat.test_global %>%
+  dplyr::filter(pheno=="WMV") %>%
+  dplyr::select(dataset, pheno, prop, sig_fdr)
+```
+
+```{r}
+cent.sex_comp_tbl <- cent.sex_comp_tbl_calc %>%
+  arrange(prop) %>%
+  add_xy_position(x="prop", fun="max") %>%
+  add_significance("p")
+
+#get max w/in each pheno, prop & dataset for placement
+max_errs <- ratio.df %>%
+  dplyr::select(participant, sex, prop, dataset, paste0("diff_", vol_list_global)) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
+  group_by(pheno, prop, dataset) %>%
+  summarize(max_diff = max(diff)) %>%
+  ungroup() %>%
+  pivot_wider(id_cols=c(pheno, prop), names_from=dataset, values_from=max_diff, names_prefix="max_")
+
+cent.sex_comp_tbl2 <- full_join(cent.sex_comp_tbl, max_errs)
+
+#just use significance from actual stats tests
+
+stats_sex.centile_for_plot <- inner_join(cent.sex_comp_tbl2, cent.feat.test_global)
+#kinda hacky way of fixing positions
+stats_sex.centile_for_plot.corrected <- stats_sex.centile_for_plot %>%
+  mutate(p.adj.signif = ifelse(sig_fdr == "FALSE", "ns", "*"),
+         y.position=case_when(dataset=="cf.gam" ~ max_cf.gam*100+0.5,
+                           dataset=="cf.gamlss" ~ max_cf.gamlss*100+0.5,
+                           TRUE ~ NA))
+
+sex_err_ratio.plt <- ratio.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
+  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         pheno = factor(pheno, levels = c("GMV", "sGMV", "Ventricles", "WMV"), ordered = TRUE),
+         diff = diff*100,
+         pheno_prop = paste(pheno, prop, sep="&")) %>% #for visuals
   ggplot() +
-  geom_histogram(aes(x=median_sex_diff, fill= pheno_cat, color=pheno_cat), alpha=0.4, position = "identity") +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  facet_grid(prop~dataset, scales="free")
+  geom_violin(aes(x=prop, y=diff, fill=sex, color=sex)) +
+  theme_linedraw(base_size = 11) +
+  labs(y="Centile Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f) +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
+  labs(fill="Sex", color="Sex")+
+  stat_pvalue_manual(stats_sex.centile_for_plot.corrected, label = "p.adj.signif", label.size = 2, tip.length = 0) +
+  facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names), scales="free_x")
+sex_err_ratio.plt
+```
 
+```{r}
 t.test.cent.sex.diff <- cent.feat.sex.test.df %>%
@@
   summarise(t.test.p = tidy(t.test(median_sex_diff, mu=0))$p.value,
-            max_val = max(median_sex_diff)) %>% #test median of each distribution against 0
+            max_val = max(median_sex_diff)*100) %>% #test median of each distribution against 0, then scale to centiles
   mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = (3*4*11)), #not testing global vols, too few data points
@@
                                       p.val_fdr >= 0.05 ~ FALSE),
-         sig_fdr = case_when(
-           # p.val_fdr < 0.001 ~ "***",
-           #                     p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
-           #                     p.val_fdr >= 0.01 & 
-                               p.val_fdr < 0.05 ~ "*",
+         sig_fdr = case_when(p.val_fdr < 0.05 ~ "*",
                                p.val_fdr >= 0.05 ~ ""),
-         y_pos = ifelse(max_val < 0, ((max_val*100)*0.5), (max_val*100)+0.8))
+         y_pos = max_val + 0.1)
 
@@
   geom_boxplot(aes(color=dataset)) +
-    facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names)) + 
+    facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_y") + 
   geom_text(aes(label = sig_fdr, y = y_pos),
@@
 
-#just global volumes
-cent.feat.sex.test.df %>%
-  mutate(median_sex_diff = median_sex_diff*100) %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  ggplot() +
-  geom_dotplot(aes(x=prop, y=median_sex_diff, color=dataset, fill=dataset), binaxis = "y", stackdir="center") +
-    facet_grid(pheno_cat~dataset, scales="free_y", labeller=labeller(dataset=dataset_names)) +
-  theme_linedraw() +
-  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
-  scale_fill_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
-  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  labs(y="Median Male - Female Centile Error", x = "% Females in Imbalanced Site") +
-  scale_x_discrete(labels=percent_f)
-
-#zoom in on combat-GAM vs combat-LS
-cent.feat.sex.test.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  #mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
-                              #TRUE ~ median_sex_diff*100)) %>%
-  ggplot() +
-  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  facet_wrap(~prop, labeller = as_labeller(percent_f), scales="free") +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+cortical_cent_sex_diff.plt
 ```
 
-visualize actual (not ABSOLUTE) errors for global phenos by sex
-
-```{r, eval=FALSE}
-#stats test
-global_sex_diffs <- cent.feat.sex.test.df %>%
-  dplyr::filter(pheno %in% vol_list_global & dataset %in% c("cf.gam", "cf.gamlss")) %>%
-  mutate(pheno_prop = paste(pheno, prop, sep="&"))
-
-global_sex_errs.df <- ratio.df %>%
-  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff")
-
-wilcox_tbl <- global_sex_errs.df %>%
-  mutate(pheno_prop = paste(pheno, prop, sep="&"),
-         pheno = factor(pheno, levels = c("GMV", "sGMV", "Ventricles", "WMV"), ordered = TRUE),
-         diff=diff*100) %>%
-  arrange(pheno, prop) %>%
-  group_by(dataset, pheno, prop) %>%
-  wilcox_test(diff ~ sex) %>%
-  add_xy_position(x="prop", fun="max")
-
-stat_merge <- full_join(wilcox_tbl, global_sex_diffs)
-stat_merge2 <- stat_merge %>%
-  add_significance("p.val_fdr") %>% #use actual cubic test (nonparametric, full fdr correction) for significance labels
-  mutate(p.val_fdr.signif = ifelse(p.val_fdr.signif == "ns", "", "*")) 
-table(stat_merge2$p.val_fdr.signif, stat_merge2$sig_fdr)
-```
-
-```{r, eval=FALSE}
-sex_err_ratio.plt <- ratio.df %>%
-  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
-  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         pheno = factor(pheno, levels = c("GMV", "sGMV", "Ventricles", "WMV"), ordered = TRUE),
-         diff = diff*100,
-         pheno_prop = paste(pheno, prop, sep="&")) %>% #for visuals
-  ggplot() +
-  geom_violin(aes(x=prop, y=diff, fill=sex, color=sex)) +
-  theme_linedraw(base_size = 11) +
-  # scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  # scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  labs(y="Centile Error", x = "% Females in Imbalanced Site") +
-  stat_pvalue_manual(stat_merge2, label = "p.val_fdr.signif", label.size = 3, tip.length = 0, hide.ns = TRUE, remove.bracket = TRUE) +
-  scale_x_discrete(labels=percent_f) +
-  # facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2, strip.position = "right") +
-  # guides(x = ggh4x::guide_axis_nested(delim = "&", inv=TRUE)) +
-  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
-  labs(fill="Sex", color="Sex")+
-  facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names), scales="free_x")
-  # ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
-sex_err_ratio.plt
-```
+COME BACK TO THIS AFTER LOOKING AT REPLICATION PLOTS
 
@@
 ```
+
 similar overall pattern
@@
 Plot errors for global phenotypes
+
+add significance testing
+```{r}
+z.t.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_t_tests_all_out.RDS"))
+
+z.t.test.list_global <- lapply(z.t.test.list, function(x) filter(x, pheno %in% paste0(vol_list_global, ".z")))
+z.t.test_global <- bind_rows(z.t.test.list_global, .id = "prop") %>%
+  dplyr::mutate(prop = case_when(prop == "0M:10F" ~ "prop-01", 
+                                 prop == "1M:9F" ~ "prop-02", 
+                                 prop == "2M:8F" ~ "prop-03", 
+                                 prop == "3M:7F" ~ "prop-04", 
+                                 prop == "4M:6F" ~ "prop-05", 
+                                 prop == "5M:5F" ~ "prop-06", 
+                                 prop == "6M:4F" ~ "prop-07", 
+                                 prop == "7M:3F" ~ "prop-08", 
+                                 prop == "8M:2F" ~ "prop-09", 
+                                 prop == "9M:1F" ~ "prop-10", 
+                                 prop == "10M:0F" ~ "prop-11",
+                                 TRUE ~ NA_character_))
+
+z.t.test.df_global_filt <- z.t.test_global %>%
+  dplyr::filter(comp %in% c("cf.gamlss_cf.gam", "cf.lm_cf.gam", "cf.lm_cf.gamlss")) %>%
+  dplyr::mutate(comp = factor(comp, ordered=TRUE, levels=c("cf.lm_cf.gam", "cf.lm_cf.gamlss", "cf.gamlss_cf.gam")))
+```
+
+```{r}
+z.prop.global_abs.diffs <- ratio.df %>%
+  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global, ".z")) %>%
+  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
+  arrange(pheno, prop) %>%
+  as.data.frame()
+
+#wilcox test for visualizing significance
+z.mean_comp_tbl_calc <- z.prop.global_abs.diffs %>%
+  group_by(pheno, prop) %>%
+  pairwise_wilcox_test(abs.diff ~ dataset, comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss")),
+                       p.adjust.method = "fdr") 
+```
+
+```{r}
+z.mean_comp_tbl <- z.mean_comp_tbl_calc %>%
+  mutate(comp = factor(paste(group2, group1, sep="_"), ordered=TRUE, levels=c("cf.lm_cf.gam", "cf.lm_cf.gamlss", "cf.gamlss_cf.gam"))) %>%
+  arrange(prop, comp) %>%
+  add_xy_position(x="prop", fun="max", comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss"))) %>%
+  add_significance("p.adj")
+
+#get max w/in each pheno, prop & dataset for placement
+z.max_abs.errs <- z.prop.global_abs.diffs %>%
+  group_by(pheno, prop, dataset) %>%
+  summarize(max_err = max(abs.diff)) %>%
+  ungroup() %>%
+  pivot_wider(id_cols=c(pheno, prop), names_from=dataset, values_from=max_err, names_prefix="max_")
+
+z.mean_comp_tbl2 <- full_join(z.mean_comp_tbl, z.max_abs.errs)
+
+#just use significance from actual stats tests
+
+stats_z_for_plot <- inner_join(z.mean_comp_tbl2, z.t.test.df_global_filt)
+#kinda hacky way of fixing positions
+stats_z_for_plot.corrected <- stats_z_for_plot %>%
+  mutate(p.adj.signif = ifelse(sig_fdr == "FALSE", "ns", "*"),
+         xmax=case_when(group1=="cf.gamlss" | group2== "cf.gamlss" ~ (x+.266667),
+                        TRUE ~  x),
+         xmin=case_when(group2=="cf.lm" ~ (x-.266667),
+                        TRUE ~  x),
+         y.position=case_when(comp == "cf.lm_cf.gamlss" ~ max_cf.lm + .16,
+                              comp == "cf.lm_cf.gam" ~ max_cf.lm + .1,
+                              TRUE ~ max_cf.gam + 0.05))
+
+z.t.test.df_global_filt %>%
+  dplyr::select(prop, comp, pheno, sig_fdr) %>%
+  dplyr::filter(sig_fdr == FALSE)
+```
+
 ```{r}
@@
   theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  stat_pvalue_manual(stats_z_for_plot.corrected, label = "p.adj.signif", label.size = 2, tip.length = 0) +
   scale_x_discrete(labels=percent_f) +
@@
 prop.global_z_errs_plot
-#ggsave("figs/global_z_errors.jpeg", prop.global_z_errs_plot)
+ggsave("figs/global_z_errors.jpeg", prop.global_z_errs_plot)
 ```
-*need to add stat tests to plot*
 
@@
                                       p.val_fdr >= 0.05 ~ FALSE),
-         sig_fdr = case_when(
-           # p.val_fdr < 0.001 ~ "***",
-           #                     p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
-           #                     p.val_fdr >= 0.01 & 
-                               p.val_fdr < 0.05 ~ "*",
+         sig_fdr = case_when(p.val_fdr < 0.05 ~ "*",
                             p.val_fdr >= 0.05 ~ ""),
-         y_pos = ifelse(max_val < 0, ((max_val)*0.5), (max_val)+0.05))
+         y_pos = max_val+0.005)
 
@@
                                p.val_fdr >= 0.05 ~ ""),
-         y_pos = ifelse(max_val < 0, ((max_val)*0.5), (max_val)+0.05))
+         y_pos = (max_val+0.01))
 
@@
 ```{r}
-fig_4 <- ggarrange(ggarrange(sampling.plt, cortical_cent_sex_diff.plt, 
+fig_4 <- ggarrange(ggarrange(sampling.plt, sex_err_ratio.plt, 
                              nrow = 2, labels = c("A", "C"),

### 2024-05-24 paper_main_results.Rmd
cleaning up, re-adding plot of sex-effects in variance

@@
   paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.25, 0.25)) #
-saveRDS(sex.var.plot, file="figs/sex_var_plot.RDS")
 ```
@@
 # sex.var.plot.global
-
-#save RDS
-saveRDS(sex.var.plot.global, file="figs/sex_var_plot_global.RDS")
 ```
@@
   ncol=2, nrow=1, 
-  widths = c(1, 3.1),
+  widths = c(1, 3.2),
   #labels=c("A", "B"),
@@
 
+# SCALE EFFECTS OF SEX
+
+Get differences in male vs female predicted variances at mean sample age
+
+```{r}
+var.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
+combined_var <- data.frame()
+
+for (file in var.csvs) {
+  # Read each CSV file
+  data <- fread(file)
+
+  # Add a "Source_File" column with the file name
+  data <- data %>%
+    mutate(Source_File = as.factor(basename(file))) %>%
+    mutate(dataset = gsub("_data|_variance.csv", "", Source_File),
+           sex_standard = sex_effect/f.var)
+
+  # Bind the data to the combined dataframe
+  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
+}
+range(combined_var$sex_standard)
+
+combined_var <- combined_var %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+#get summary df for p vals, plotting
+drop1.df <- fread(paste0(data_path, "/ukb_basic/drop1_tests.csv"), stringsAsFactors = TRUE) %>%
+  dplyr::select(-V1) %>%
+  rename(dataset=Dataset,
+         pheno=Model) %>%
+    dplyr::filter(Moment == "sigma" & Term == "sexMale") %>%
+  mutate(dataset = gsub("-data|ukb-CN-data-simsites-", "", dataset)) %>% #clean up dataset naming
+  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
+  mutate(dataset = relevel(dataset, ref= "raw"))
+
+sigma.sex.df <- left_join(combined_var, drop1.df)
+sigma.sex.df2 <- left_join(sigma.sex.df, dk.parc)
+
+#combined_var <- Reduce(function(x, y) full_join(x, y), list(combined_var, dk.parc, drop1.df))
+```
+
+```{r}
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+#var plot
+var.plot.df <- sigma.sex.df2 %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "cf") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         sex_standard = ifelse(drop1.pval <0.05, sex_standard, NA)) %>%
+  group_by(dataset, pheno_cat)
+
+range(var.plot.df$sex_standard, na.rm=T)
+
+tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="raw")
+tp$sex_standard <- 1
+
+var.plot <- ggplot(var.plot.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+             aes(fill = sex_standard)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
+var.plot
+
+#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
+saveRDS(var.plot, file="figs/sex_var_plot.RDS")
+```
+
+```{r}
+#plot global
+var.plot.global.df <- sigma.sex.df2 %>%
+  dplyr::filter(pheno_cat == "Global Volume" & dataset != "cf") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  dplyr::mutate(tissue_class = case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "WM",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA_character_),
+         sex_standard = ifelse(drop1.pval <0.05, sex_standard, NA)) %>%
+  group_by(dataset)
+
+tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="raw")
+tp$sex_standard <- 1
+
+range(var.plot.global.df$sex_standard, na.rm=T)
+
+var.plot.global <- ggplot(var.plot.global.df) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = sex_standard),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
+#var.plot.global
+saveRDS(var.plot.global, file="figs/sex_var_plot_global.RDS")
+```
+
+Combine plots
+```{r}
+p <- ggarrange(
+  var.plot.global, var.plot,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.1),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+```
+
 # FIGURE 1: METHODS

### 2024-05-24 paper_replication_results.Rmd
figs

--- 2024-03-13 paper_replication_results.Rmd
@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata, gtExtras)
 #Sys.setenv("MC_CORES"=10L)
@@
 
-cent.feat.t.test.df %>%
+cent.tbl <- cent.feat.t.test.df %>%
   group_by(comp, smaller_group) %>%
-  summarize(n_sig=sum(sig_fdr))
-```
-
-**need to prettify table for supplement**
+  summarize(n_sig=sum(sig_fdr)) %>%
+  mutate(perc_tests = n_sig/(208*100))
+
+cent.tbl.pretty <- cent.tbl %>%
+  mutate(comp = case_when(comp == "cf.gam_cf" ~ "ComBat-GAM vs ComBat w/o Covariates",
+                          comp == "cf.gamlss_cf" ~ "ComBatLS vs ComBat w/o Covariates", 
+                          comp == "cf.gamlss_cf.gam" ~ "ComBatLS vs ComBat-GAM",
+                          comp == "cf.lm_cf" ~ "Linear ComBat vs ComBat w/o Covariates",
+                          comp == "cf.lm_cf.gam" ~ "Linear ComBat vs ComBat-GAM",
+                          comp == "cf.lm_cf.gamlss" ~ "Linear ComBat vs ComBatLS"),
+         smaller_group = case_when(smaller_group == "cf" ~ "ComBat w/o Covariates",
+                                   smaller_group == "cf.lm" ~ "Linear ComBat",
+                                   smaller_group == "cf.gam" ~ "ComBat-GAM",
+                                   smaller_group == "cf.gamlss" ~ "ComBatLS")) %>%
+  rename(#"Pairwise Comparison"=comp,
+         "Lower Errors" = smaller_group,
+         "N Features" = n_sig,
+         "% Features" = perc_tests)
+
+cent.tbl.pretty %>%
+  gt() %>% 
+  #gt_theme_538() %>% 
+tab_style(style = list(cell_fill(color = "gray95")),
+    locations = cells_row_groups()) %>%
+  fmt_percent(
+    columns = "% Features",
+    decimals = 2
+  ) %>%
+  tab_options(column_labels.font.weight = "bold") %>%
+  tab_header(title = "Pairwise comparisons of ComBat methods' absolute centile errors for 208 brain features across 100 replications")
+```
 
@@
 
-dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 
@@
   theme_linedraw()+
-  theme(axis.text.x = element_blank()) +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "bottom") +
   facet_grid(pheno_cat~comp, scales="free_y", labeller = labeller(comp = as_labeller(comp_names))) +
@@
   labs(x="Sampling Permutation", y="Feature Count")
-ggsave(paste0(base_path, "/figs/perm_cent_test_barplt.jpg"), plot=cent.test.comp.plt, width=15.5, height=6, units="in")
-```
-
-```{r}
-#look at combatGAM vs combatLS
+```
+
+```{r}
 cent.feat.t.test.df %>%
-  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
-ggplot(aes(x=perm, fill=bigger_group)) +
-  geom_bar() +
-  # geom_text(stat='count', aes(label=..count..)) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~pheno_cat)
-
-cf.gam_worse <- cent.feat.t.test.df %>%
-  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam" & bigger_group == "cf.gamlss") %>%
-  group_by(perm) %>%
-  summarise(count = n())
-
-mean(cf.gam_worse$count)
-IQR(cf.gam_worse$count)
+  dplyr::filter(bigger_group=="cf.gamlss" & sig_fdr==TRUE) %>%
+  group_by(comp, pheno_cat) %>%
+  summarize(count = n())
+
+cent.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE & bigger_group == "cf.gamlss") %>%
+  group_by(comp, perm) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(comp) %>%
+  summarize(med.count = median(count),
+            IQR.count = IQR(count))
 ```
@@
 ct.t.tests <- cent.feat.t.test.df %>%
-  dplyr::filter(comp =="cf.gamlss_cf.gam" & stringr::str_detect(pheno, "CT")) %>%
-  mutate(label=gsub("_CT", "", pheno),
-         better_fit = case_when(sig_fdr==TRUE & bigger_group == "cf.gam" ~ "cf.gamlss",
-                              sig_fdr==TRUE & bigger_group == "cf.gamlss" ~ "cf.gam",
-                              sig_fdr==FALSE ~ "NS"))
+  dplyr::filter(pheno_cat == "Regional CT" & grepl("gamlss", comp)) %>%
+  mutate(label=gsub("_CT", "", pheno))
 
@@
 
-ct.brain.plt <- ct.t.tests_to_plt %>%
+ct.t.tests_to_plt.sum <- ct.t.tests_to_plt %>%
+  group_by(comp, hemi, region, label, smaller_group) %>%
+  summarise(n_better = sum(sig_fdr)) %>%
+  mutate(perc.better=n_better/100) %>% #make NAs into 0s
+  ungroup() %>%
+  pivot_wider(id_cols=c(comp, hemi, label, region), names_from = smaller_group, values_from=perc.better) %>%
+  dplyr::select(!c(cf, cf.lm, cf.gam))
+
+min(ct.t.tests_to_plt.sum$cf.gamlss)
+
+comp_names <- c(cf.gamlss_cf = "vs. ComBat w/o Covar", cf.lm_cf.gamlss = "vs. linear ComBat", cf.gamlss_cf.gam ="vs. ComBat-GAM")
+
+ct.brain.sum.plt <- ct.t.tests_to_plt.sum %>%
+  mutate(comp=as.character(comp)) %>%
   ggplot() +
@@
              position = position_brain(hemi ~ side),
-            aes(fill=better_fit)) +
-  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank(),
-        legend.position="top") +
-  scale_fill_manual(values = c(NS= "darkgray", cf.gam = "#3D75C2", cf.gamlss ="#CFB023FF"), name = "Better ComBat Method", labels=dataset_names) +
-  facet_wrap(~perm, nrow=2)
-
-ct.t.tests_to_plt.sum <- ct.t.tests_to_plt %>%
-  group_by(hemi, region, label, bigger_group) %>%
-  summarise(n_worse = sum(sig_fdr)) %>%
-  pivot_wider(id_cols=c(hemi, label, region), names_from = bigger_group, values_from=n_worse) %>%
-  mutate(cf.gamlss=ifelse(is.na(cf.gamlss), 0, cf.gamlss)) %>%
-  mutate(cf.gamlss_minus_cf.gam_better=as.integer(cf.gam-cf.gamlss),
-         percent_cf.gamlss_better=(cf.gam/100)*100) %>%
-  ungroup()
-
-ct.brain.sum.plt <- ct.t.tests_to_plt.sum %>%
-  ggplot() +
-  theme_linedraw(base_size=12) +
-  ggseg::geom_brain(atlas = dk_nomed, 
-             position = position_brain(hemi ~ side),
-            aes(fill=percent_cf.gamlss_better))  +
+            aes(fill=cf.gamlss))  +
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  scale_fill_gradient(low="#605210", high= "#e6d16d",  name = "% Replications") +
-  labs(title="Replications where ComBatLS recapitulates cortical thickness centiles \nsignificantly better than ComBat-GAM")
+  scale_fill_gradient(low="#605210", high= "#EFE2A2",  name = "Proportion Replications", limits = c(.5, 1), breaks=c(0.50, 0.75, 1.00)) +
+  facet_wrap(~comp, labeller = labeller(comp = as_labeller(comp_names)))
+ct.brain.sum.plt
 ```
@@
 
-simple bar-plot to count regions w sigificant sex-biases
+simple bar-plot to count regions w significant sex-biases
 
@@
 #look by feature type
-cent.subj.sex.test.df %>%
+sex_bias_hist_plt <- cent.subj.sex.test.df %>%
   dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
@@
 ggplot() +
-  geom_histogram(aes(x=median_sex_diff, color=perm, fill=perm), alpha=0.3, position="identity") +
+  geom_histogram(aes(x=median_sex_diff, color=perm, fill=perm), alpha=0.3, position="identity", bins=40) +
   geom_vline(xintercept = 0, linetype="dashed") +
-  facet_grid(pheno_cat~dataset, scales="free", labeller = labeller(dataset = as_labeller(dataset_names))) +
+  facet_grid(pheno_cat~dataset, scales="free_y", labeller = labeller(dataset = as_labeller(dataset_names))) +
   theme_linedraw(base_size=12) +
   theme(legend.position = "none") +
-  labs(x="Male - Female Median Centile Errors", y="Number of Features") 
+  labs(x="Male - Female Median Centile Errors", y="Number of Features") +
+  ggtitle("Sex differences in median centile errors across 100 sampling replications")
+ggsave(filename="figs/sex_bias_hist_plt.png", sex_bias_hist_plt)
 ```
@@
                                p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ ""))
-
-t.test.cent.sex.test.med.filt <- t.test.cent.sex.test.med %>% dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")
+                               p.val_fdr >= 0.05 ~ ""),
+         y_pos = case_when(dataset=="cf" ~ 11,
+                           dataset=="cf.lm" ~ 11,
+                           dataset=="cf.gam" ~ 10.5,
+                           dataset=="cf.gamlss" ~ 6))
 
@@
 
+t.test.cent.sex.test.med.filt <- t.test.cent.sex.test.med %>% dplyr::filter(dataset != "cf")
+
 sex_bias_density_plt <- cent.sex.test.med %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  dplyr::filter(dataset != "cf") %>%
   ggplot() +
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med.filt$dataset == "cf.gam", 10.2, 5.5)),
+  geom_text(aes(label = sig_fdr, x= x_pos*100, y=y_pos),
               data = t.test.cent.sex.test.med.filt) +
-  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  theme(legend.position = "bottom")+
-  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
-
-cent.sex.test.med %>%
-  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  ggplot() +
-  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
-  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
-  facet_wrap(~pheno_cat) +
-  theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med$dataset == "cf.gam", 10.2, 5.5)),
-              data = t.test.cent.sex.test.med) +
   scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
@@
 
+y_pos.df <- all_cent_sum.df_scaled %>%
+  group_by(source, pheno_cat, dataset) %>%
+  summarise(y_pos=max(female_ratio_diff))
+
+all_t_test.df <- full_join(all_t_test.df, y_pos.df)
+
 #plots!
@@
 
-all_cent_sum_boxplt <- all_cent_sum.df %>%
-  dplyr::filter(dataset != "raw" & dataset != "cf") %>%
-  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=female_ratio)) +
-  geom_boxplot(aes(color=dataset)) +
-  # geom_dotplot(binaxis = "y", stackdir="center") +
-  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
-  geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
-  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
-  #             data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
-  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
-  theme_linedraw() +
-  scale_x_discrete(labels=dataset_names) +
-  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
-  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  labs(y="Proportion Female", x = "Dataset")
-
-all_cent_sum.df_scaled %>%
-  dplyr::filter(dataset != "raw" & dataset != "cf") %>%
-  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=female_ratio_sc)) +
-  geom_boxplot(aes(color=dataset)) +
-  # geom_dotplot(binaxis = "y", stackdir="center") +
-  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
-  geom_hline(yintercept=1, linetype="dashed")+
-  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
-  #             data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
-  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
-  theme_linedraw() +
-  scale_x_discrete(labels=dataset_names) +
-  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
-  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  labs(y="Female Bias", x = "Dataset")
-
 all_cent_sum_boxplt.diff <- all_cent_sum.df_scaled %>%
-  dplyr::filter(dataset != "raw" & dataset != "cf") %>%
+  dplyr::filter(dataset != "raw" & dataset != "cf" & source != "ext") %>%
   mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
@@
   geom_hline(yintercept=0, linetype="dashed")+
-  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
-  #             data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
+  geom_text(aes(label = sig_fdr, y = (y_pos)), nudge_y=.005,
+               data = all_t_test.df[which(all_t_test.df$dataset != "cf" & all_t_test.df$source != "ext"),]) +
   scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
@@
 
-linear ComBat ~ too few women with high centiles, too many women with low centiles; ~ balances out overall
-
-ComBat-GAM -> variable pattern, pretty close to raw
-
-ComBatLS -> pretty close to raw!
-
 ### Without Extremes
@@
 
-look at combatGAM vs combatLS
-
+look at ratios
 ```{r}
 z.feat.t.test.df %>%
-  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
-ggplot(aes(x=perm, fill=bigger_group)) +
-  geom_bar() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~pheno_cat)
-
-z.feat.t.test.df %>%
-  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
+  dplyr::filter(sig_fdr==TRUE & grepl("gamlss", comp)) %>%
 ggplot(aes(x=perm, fill=bigger_group)) +
@@
   theme(axis.text.x = element_blank()) +
-  facet_wrap(~pheno_cat)
+  facet_grid(comp~pheno_cat)
 ```
@@
 sfig <- ggarrange(cent.test.comp.plt, ct.brain.sum.plt,
-                  ncol=1, nrow=2, heights = c(2, 1))
+                  ncol=1, nrow=2, heights = c(1.8, 1), labels = "AUTO")
 sfig
-ggsave(paste0(base_path, "/figs/combat_abs_errs_supp_plt.jpg"), plot=sfig, width=11, height=7, units="in")
-
+ggsave(paste0(base_path, "/figs/combat_abs_errs_supp_plt.jpg"), plot=sfig, width=13, height=9, units="in")
 ```

### 2024-05-25 paper_lbcc_results.Rmd
cleaning up

--- 2024-05-01 paper_lbcc_results.Rmd
@@
 
-df_raw_feat_aov.study %>%
-  ggplot() +
-  geom_histogram(aes(x=etasq, fill=sig_fdr), alpha=0.5)
-
-df_raw_feat_aov.study %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  nrow()
+table(df_raw_feat_aov.study$sig_fdr)
 ```
@@
 ```
-linear model
-
-```{r}
-summary(lm(mean_cent ~ study, data=raw_centile_df.mean))
+
+ANOVA of subjects' mean centiles across studies
+
+```{r}
 summary(aov(mean_cent ~ study, data=raw_centile_df.mean))
@@
 
-df_raw_feat_aov.z.study %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  nrow()
+table(df_raw_feat_aov.z.study$sig_fdr)
 ```
@@
 
-```{r}
-raw_z_df.mean %>%
-  ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
-# +
-#   facet_wrap(~study)
-```
-
-linear model
-
-```{r}
-summary(lm(mean_z ~ study, data=raw_z_df.mean))
+ANOVA of subject means
+
+```{r}
 summary(aov(mean_z ~ study, data=raw_z_df.mean))
@@
 
-df_gam_feat_aov.study %>%
-  ggplot() +
-  geom_histogram(aes(x=etasq, fill=sig_fdr), alpha=0.5)
-
-df_gam_feat_aov.study %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  nrow()
+table(df_gam_feat_aov.study$sig_fdr)
 ```
@@
 
-linear model
-
-```{r}
-summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
+ANOVA
+
+```{r}
 summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
@@
 
-df_gam_feat_aov.z.study %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  nrow()
+table(df_gam_feat_aov.z.study$sig_fdr)
 ```
@@
   geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
-# +
-#   facet_wrap(~study)
-```
-
-linear model
-
-```{r}
-summary(lm(mean_z ~ study, data=gam_z_df.mean))
+```
+
+Subject mean ANOVA
+
+```{r}
 summary(aov(mean_z ~ study, data=gam_z_df.mean))
@@
   dplyr::select(!contains(".z")) #drop z-scores and non-centiles & !ends_with("_total") & !contains("TBV")
-```
-
-```{r}
-ggplot(gamlss_centile_df.clean) +
-  geom_point(aes(x=age_days, y=rh_Vol_insula, color=study), alpha=0.5) +
-  theme(legend.position="none")
 ```
@@
 
-df_gamlss_feat_aov.study %>%
-  ggplot() +
-  geom_histogram(aes(x=etasq, fill=sig_fdr))
-
-df_gamlss_feat_aov.study %>%
-  dplyr::filter(sig_fdr==TRUE) %>%
-  nrow()
+table(df_gamlss_feat_aov.study$sig_fdr)
 ```
@@
 ```{r}
-ggplot(gamlss_centile_df.mean) +
-  geom_point(aes(x=(age_days/365.25), y=mean_cent, color=study), alpha=0.5) +
-  theme(legend.position="none")
-
-ggplot(gamlss_centile_df.clean) +
-  geom_point(aes(x=age_days, y=lh_Vol_inferiortemporal, color=study), alpha=0.2) +
-  theme(legend.position="none")
-```
-
-```{r}
 gamlss_centile_df.mean %>%
@@
   theme(axis.text.x = element_text(angle=90, hjust=0.5))
-
-gamlss_centile_df.mean %>%
-  ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_cent, color=study), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_cent, color=study), method="lm") +
-  theme(legend.position = "none")
-
-gamlss_centile_df.mean %>%
-  dplyr::filter(age_days>365) %>%
-  ggplot(aes(x=age_days/365.25, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-linear model
-
-```{r}
-summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
+```
+
+subject mean ANOVA
+
+```{r}
 summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
@@
 
-df_gamlss_feat_aov.z.study %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  nrow()
+table(df_gamlss_feat_aov.z.study$sig_fdr)
 ```
@@
 
-linear model
-
-```{r}
-summary(lm(mean_z ~ study, data=gamlss_z_df.mean))
+subject mean ANOVA
+
+```{r}
 summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
@@
 ```{r}
-summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
-summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
-summary(lm(mean_cent ~ study, data=raw_centile_df.mean))
-```
-
-compare magnitude of effect size for study (w/in feature anovas) across unharmonized, combat-gam and combatLS centiles
-
-```{r}
-df_gamlss_feat_aov.study$combat <- "gamlss"
-df_gam_feat_aov.study$combat <- "gam"
-df_raw_feat_aov.study$combat <- "raw"
+gamlss_centile_df.mean$combat <- "gamlss"
+gam_centile_df.mean$combat <- "gam"
+raw_centile_df.mean$combat <- "raw"
 
 #put all data frames into list
-df_list.tmp <- list(df_gamlss_feat_aov.study, df_gam_feat_aov.study, df_raw_feat_aov.study)
+df_list.tmp <- list(gamlss_centile_df.mean, gam_centile_df.mean, raw_centile_df.mean)
 
 #merge all data frames in list
-feat_aov.df <- df_list.tmp %>% reduce(full_join)
-
-feat_aov.df <- feat_aov.df %>%
+mean_cents.df <- df_list.tmp %>% reduce(full_join)
+
+dataset_names <- c(gam = "ComBat-GAM", gamlss = "ComBatLS", raw = "Unharmonized")
+
+mean_cents_bxplt <- mean_cents.df %>%
+  ggplot() +
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
+  theme_linedraw() +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none", strip.text.x = element_text(size = 12)) +
+  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
+  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
+  labs(y="Mean Centile", x = "Study")
+
+mean_cents_bxplt_nolabel <- mean_cents.df %>%
+  ggplot() +
+  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
+  theme_linedraw() +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none", strip.text.x = element_text(size = 12)) +
+  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
+  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
+  labs(y="Mean Centile", x = "Study")
+```
+
+Cohen's F squared for study effects
+
+```{r}
+gam_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv"))
+gam_centile_f2$combat <- "ComBat-GAM"
+
+gamlss_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_cohenfsq.csv"))
+gamlss_centile_f2$combat <- "ComBatLS"
+
+f2.df <- rbind(gam_centile_f2, gamlss_centile_f2)
+f2.df <- f2.df %>%
   mutate(pheno_cat = as.factor(case_when(
@@
 
-feat_aov.df %>%
-  group_by(combat) %>%
-  summarize(n_sig = sum(sig_fdr))
-
-feat_aov.df %>%
-  ggplot() +
-  geom_point(aes(x=etasq, y=pheno, color=combat, shape=sig_fdr), alpha=0.6) +
-  facet_wrap(~pheno_cat, scales="free")
-
-feat_aov.df %>%
-  dplyr::filter(combat!="raw") %>%
-  ggplot() +
-  geom_point(aes(x=etasq, y=pheno, color=combat, shape=sig_fdr), alpha=0.6) +
-  facet_wrap(~pheno_cat, scales="free")
-```
-```{r}
-gamlss_centile_df.mean$combat <- "gamlss"
-gam_centile_df.mean$combat <- "gam"
-raw_centile_df.mean$combat <- "raw"
-
-#put all data frames into list
-df_list.tmp <- list(gamlss_centile_df.mean, gam_centile_df.mean, raw_centile_df.mean)
-
-#merge all data frames in list
-mean_cents.df <- df_list.tmp %>% reduce(full_join)
-
-dataset_names <- c(gam = "ComBat-GAM", gamlss = "ComBatLS", raw = "Unharmonized")
-
-mean_cents_bxplt <- mean_cents.df %>%
-  ggplot() +
-  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
+fsq.df <- ggplot(f2.df) +
+  geom_col(aes(x=pheno, y=fsq, fill=combat), position="dodge") + 
+  facet_wrap(~pheno_cat, scales="free") +
   theme_linedraw() +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none", strip.text.x = element_text(size = 12)) +
-  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
-  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
-  labs(y="Mean Centile", x = "Study")
-
-mean_cents_bxplt_nolabel <- mean_cents.df %>%
-  ggplot() +
-  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
-  theme_linedraw() +
-  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none", strip.text.x = element_text(size = 12)) +
-  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
-  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
-  labs(y="Mean Centile", x = "Study")
-```
-
-#Feature Centile ~ Mean Study Age
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
+  labs(x = "Feature", y = "Effect Size (Cohen's F-Squared)", fill="ComBat")
+
+ggsave("figs/supp_fsq.jpeg", fsq.df, width=8, height=6, units="in")
+```
+
+```{r}
+f2.df.sum <- f2.df %>%
+  group_by(pheno) %>%
+  arrange(combat) %>%
+  mutate(F2_diff = fsq - fsq[1]) #gamlss - gam value, + number means site effect is larger in combatLS
+
+ggplot(f2.df.sum) +
+  geom_col(aes(x=pheno, y=F2_diff), position="dodge") + 
+  facet_wrap(~pheno_cat, scales="free") +
+  theme(axis.text.x = element_blank())
+
+f2.df.sum %>%
+  group_by(pheno_cat) %>%
+  summarise(mean_fsq_diff = mean(F2_diff)) %>%
+  ggplot() + 
+  geom_col(aes(x=pheno_cat, y=mean_fsq_diff), position="dodge")
+```
+
+#ComBat-GAM vs ComBatLS: Delta Centile
 
@@
 ```{r}
+nrow(raw_centile_df.clean) == n_unique(raw_centile_df.clean$participant)
+
 study_info <- raw_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
   group_by(study) %>%
@@
             study_median_age=median(age_years),
-            study_n= n()) %>%
+            study_n = n()) %>%
   dplyr::select(study, study_mean_age, study_median_age, study_n)
 ```
-
-W/in feat lms
-
-##Unharmonized
-
-```{r}
-raw_centile_df.clean_study <- full_join(raw_centile_df.clean, study_info)
-
-df_raw_feat_lm <- data.frame()
-for(pheno in raw.feat_to_test){
-  # print(pheno)
-  df <- tidy(aov(raw_centile_df.clean_study[[pheno]] ~ raw_centile_df.clean_study[["study_mean_age"]]))
-  df$pheno <- pheno
-  df_raw_feat_lm <- rbind(df_raw_feat_lm, df)
-}
-
-df_raw_feat_lm <- df_raw_feat_lm %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-df_raw_feat_lm %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-df_raw_feat_lm %>%
-  ggplot() +
-  geom_histogram(aes(x=statistic))
-```
-
-##ComBat-GAM
-
-```{r}
-gam_centile_df.clean_study <- full_join(gam_centile_df.clean, study_info)
-
-df_gam_feat_lm <- data.frame()
-for(pheno in gam.feat_to_test){
-  # print(pheno)
-  df <- tidy(aov(gam_centile_df.clean_study[[pheno]] ~ gam_centile_df.clean_study[["study_mean_age"]]))
-  df$pheno <- pheno
-  df_gam_feat_lm <- rbind(df_gam_feat_lm, df)
-}
-
-df_gam_feat_lm <- df_gam_feat_lm %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-
-df_gam_feat_lm %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-df_gam_feat_lm %>%
-  ggplot() +
-  geom_histogram(aes(x=statistic))
-```
-
-##ComBatLS
-
-```{r}
-gamlss_centile_df.clean_study <- full_join(gamlss_centile_df.clean, study_info)
-
-df_gamlss_feat_lm <- data.frame()
-for(pheno in gamlss.feat_to_test){
-  # print(pheno)
-  df <- tidy(aov(gamlss_centile_df.clean_study[[pheno]] ~ gamlss_centile_df.clean_study[["study_mean_age"]]))
-  df$pheno <- pheno
-  df_gamlss_feat_lm <- rbind(df_gamlss_feat_lm, df)
-}
-
-df_gamlss_feat_lm <- df_gamlss_feat_lm %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-df_gamlss_feat_lm %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-df_gamlss_feat_lm %>%
-  ggplot() +
-  geom_histogram(aes(x=statistic))
-```
-
-*statistic ~ effect size?
-
-```{r}
-df_raw_feat_lm$combat <- "raw"
-df_gam_feat_lm$combat <- "gam"
-df_gamlss_feat_lm$combat <- "gamlss"
-
-#put all data frames into list
-df_list.tmp <- list(df_raw_feat_lm, df_gam_feat_lm, df_gamlss_feat_lm)
-
-#merge all data frames in list
-all_feat_lm <- df_list.tmp %>% reduce(full_join)
-
-all_feat_lm %>%
-  ggplot() +
-  geom_histogram(aes(x=statistic, fill=combat, color=combat), position="identity", alpha=0.5, bins=50)
-```
-
-#Subject-Level Mean Centiles ~ Mean Study Age
-
-```{r}
-raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
-summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
-
-raw_centile_df.mean2 %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-```{r}
-gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
-summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
-
-gam_centile_df.mean2 %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-
-gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
-  mutate(age_diff = (age_days/365.25) - study_mean_age)
-
-summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
-gam_centile_df.mean2 %>%
-  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-```{r}
-gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
-summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
-
-gamlss_centile_df.mean2 %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-
-gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
-  mutate(age_diff = (age_days/365.25) - study_mean_age)
-
-summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
-gamlss_centile_df.mean2 %>%
-  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-Plots
-
-```{r}
-gamlss_centile_df.mean2$combat <- "gamlss"
-gam_centile_df.mean2$combat <- "gam"
-raw_centile_df.mean2$combat <- "raw"
-
-#put all data frames into list
-df_list.tmp <- list(gamlss_centile_df.mean2, gam_centile_df.mean2, raw_centile_df.mean2)
-
-#merge all data frames in list
-all_subj_means.df <- df_list.tmp %>% reduce(full_join)
-
-all_subj_means.df %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
-  geom_point(alpha=0.6) +
-  geom_smooth(method="lm")
-
-all_subj_means.df %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_wrap(~combat)
-
-summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
-summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
-summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
-```
-
-```{r}
-summary(lm(mean_cent ~ study_n, data=raw_centile_df.mean2))
-summary(lm(mean_cent ~ study_n, data=gam_centile_df.mean2))
-summary(lm(mean_cent ~ study_n, data=gamlss_centile_df.mean2))
-
-all_subj_means.df %>%
-  ggplot(aes(x=study_n, y=mean_cent, color=combat)) +
-  geom_point(alpha=0.6) +
-  geom_smooth(method="lm")
-```
-all assoc with study n
-
-#ComBat-GAM vs ComBatLS: Delta Centile
 
@@
 
+control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.
+
 check mean subj delta ~ study mean age
@@
   theme(legend.position="none")
-
+```
+
+See if a subject's distance from their study's mean age is assoc with how much their centile changes
+
+```{r}
 mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
@@
 
-abs_lm2 <- lm(abs_mean_delta_cent ~ study_mean_age + study_n + abs_age_diff, data=mean_cent_deltas_df_study)
+abs_lm2 <- lm(abs_mean_delta_cent ~ study_n + abs_age_diff, data=mean_cent_deltas_df_study)
 summary(abs_lm2)
@@
 
-control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.
+## Subject-Level Mean Centiles ~ Mean Study Age
+
+Make sure that results of mean centile delta ~ mean study age are related to the difference itself, not because ComBatLS and/or ComBat-GAM centiles are associated with mean study age.
+
+```{r}
+raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
+summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
+
+raw_centile_df.mean2 %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+NS
+
+```{r}
+gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
+summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
+
+gam_centile_df.mean2 %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+
+gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age)
+
+summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
+gam_centile_df.mean2 %>%
+  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+NS
+
+```{r}
+gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
+summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
+
+gamlss_centile_df.mean2 %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+
+gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age)
+
+summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
+gamlss_centile_df.mean2 %>%
+  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+NS
 
@@
 
-```{r}
-# study demographics
-fig5 <- ggarrange(
-ggarrange(sample_boxplot, sample_barplot,
-          full_boxplot, full_barplot,
-          nrow=2, ncol=2, 
-          #align = "h", 
-          common.legend = TRUE,
-          legend = "bottom",
-          widths=c(1.7, 1),
-          heights=c(10, 1)),
-# boxplots
-mean_cents_bxplt,
-common.legend=FALSE,
-nrow=1, ncol=2,
-labels="AUTO",
-widths=c(1, 4)
-)
-
-ggsave("figs/figure_5_horizontal.jpeg", fig5, width=18, height=9, units="in")
-```
-alternative vertical stacking
+Reformatting plots for prettier composite fig
 
@@
     geom_line(data = pred_df, aes(y = fit, x=study_mean_age))+
-  #theme(legend.position="none") +
-  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
-ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=10, height=7, units="in")
-
-mean_cent_deltas_df_study %>%
-  dplyr::filter(abs_mean_delta_cent > 4) %>%
-  dplyr::select(study, study_mean_age)
-
-mean_cent_deltas_df_study%>%
-  dplyr::filter(study == "Cornell_C2") %>%
-  nrow()
-```
-
-Cornell_C2 has very low n=5 -> big differences in estimation
+  theme_linedraw() +
+  theme(legend.position="none") +
+  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From ComBatLS and ComBat-GAM \n-harmonized Data")
+ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=9, height=6, units="in")
+```
+
+lines show marginal effects of mean study age
 
@@
 
-mean_cent_deltas_df_study %>%
+mean_delta_fig2 <- mean_cent_deltas_df_study %>%
   ggplot(aes(x=abs_age_diff, y=abs_mean_delta_cent)) +
@@
     geom_line(data = pdat2, aes(y = fit, x=abs_age_diff)) +
+  theme_linedraw() +
   theme(legend.position="none") +
-  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
-```
+  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From ComBatLS and ComBat-GAM \n-harmonized Data")
+ggsave("figs/supp_mean_delta_fig2.jpeg", mean_delta_fig2, width=9, height=6, units="in")
+```

### 2024-06-04 paper_replication_results.Rmd
plots

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata, gtExtras)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata, gtExtras, webshot2)
 #Sys.setenv("MC_CORES"=10L)
@@
   rename(#"Pairwise Comparison"=comp,
-         "Lower Errors" = smaller_group,
+         "Method Producing Smaller Absolute Errors" = smaller_group,
          "N Features" = n_sig,
@@
 
-cent.tbl.pretty %>%
+pretty_cent_table <- cent.tbl.pretty %>%
   gt() %>% 
@@
   tab_header(title = "Pairwise comparisons of ComBat methods' absolute centile errors for 208 brain features across 100 replications")
+
+gtsave(pretty_cent_table, filename="figs/centile_table.png")
 ```
@@
   ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations replicated 100x") +
+  theme_linedraw() + 
   theme(legend.position="none")
@@
 sex_bias_hist_plt <- cent.subj.sex.test.df %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
-  mutate(dataset = factor(dataset, levels = c("cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
   mutate(pheno_cat = factor(case_when(
@@
   geom_vline(xintercept = 0, linetype="dashed") +
-  facet_grid(pheno_cat~dataset, scales="free_y", labeller = labeller(dataset = as_labeller(dataset_names))) +
+  facet_grid(pheno_cat ~ dataset, scales="free", labeller = labeller(dataset = as_labeller(dataset_names))) +
   theme_linedraw(base_size=12) +
   theme(legend.position = "none") +
-  labs(x="Male - Female Median Centile Errors", y="Number of Features") +
-  ggtitle("Sex differences in median centile errors across 100 sampling replications")
-ggsave(filename="figs/sex_bias_hist_plt.png", sex_bias_hist_plt)
+  labs(x="Male - Female Median Centile Errors", y="Number of Features")
+ggsave(filename="figs/sex_bias_hist_plt.png", sex_bias_hist_plt, height=7, width=10, units="in")
 ```
@@
                                p.val_fdr >= 0.05 ~ ""),
-         y_pos = case_when(dataset=="cf" ~ 11,
+         y_pos = case_when(dataset=="cf" ~ 9.3,
                            dataset=="cf.lm" ~ 11,
@@
                            dataset=="cf.gamlss" ~ 6))
+#get vals
+t.test.cent.sex.test.med
 
@@
   labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
+
+#with combat w/o covar
+sex_bias_density_plt_full <- cent.sex.test.med %>%
+  ggplot() +
+  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
+  facet_wrap(~pheno_cat) +
+  theme_linedraw() +
+  geom_text(aes(label = sig_fdr, x= x_pos*100, y=y_pos),
+              data = t.test.cent.sex.test.med) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom") +
+  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
+ggsave(filename="figs/sex_bias_density_plt_full.png", sex_bias_density_plt_full, height=7, width=10, units="in")
 ```
@@
   )
-
-ext_cent.sum <- perm_subj.df2_long %>%
-  dplyr::filter(pheno_cat != "allpheno") %>%
-  dplyr::filter(mean_centile < 0.20 | mean_centile > 0.80) %>%
-  group_by(dataset, perm, pheno_cat) %>%
-  summarise(
-    male_ratio = sum(sex == "Male") / n(),
-    female_ratio = sum(sex == "Female") / n()
-  )
 ```
@@
 low_cent.sum_raw <- low_cent.sum %>%
-  dplyr::filter(dataset == "raw" & perm=="perm-001")
-
-ext_cent.sum_raw <- ext_cent.sum %>%
   dplyr::filter(dataset == "raw" & perm=="perm-001")
@@
 low_t_test.df <- lapply(split(low_cent.sum, low_cent.sum$pheno_cat), get_t_tests, raw_df=low_cent.sum_raw) %>% bind_rows()
-ext_t_test.df <- lapply(split(ext_cent.sum, ext_cent.sum$pheno_cat), get_t_tests, raw_df=ext_cent.sum_raw) %>% bind_rows()
 
 #merge dfs
-all_cent_sum.df <- gdata::combine(high_cent.sum, low_cent.sum, ext_cent.sum) %>%
+all_cent_sum.df <- gdata::combine(high_cent.sum, low_cent.sum) %>%
   mutate(source = gsub("_cent.sum", "", source))
@@
 
-all_t_test.df <- gdata::combine(high_t_test.df, low_t_test.df, ext_t_test.df) %>%
+all_t_test.df <- gdata::combine(high_t_test.df, low_t_test.df) %>%
   mutate(source = gsub("_t_test.df", "", source))
@@
 
-all_t_test.df <- full_join(all_t_test.df, y_pos.df)
-
+all_t_test.df <- full_join(all_t_test.df, y_pos.df) %>%
+  dplyr::filter(dataset != "raw")
+```
+
+```{r}
 #plots!
@@
 
-cent_ext_list <- c(ext = ">80% and <20%", high = ">80%", low = "<20%")
+cent_ext_list <- c(high = ">80%", low = "<20%")
 
 all_cent_sum_boxplt.diff <- all_cent_sum.df_scaled %>%
-  dplyr::filter(dataset != "raw" & dataset != "cf" & source != "ext") %>%
+  dplyr::filter(dataset != "raw" & dataset != "cf") %>%
   mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
@@
   geom_text(aes(label = sig_fdr, y = (y_pos)), nudge_y=.005,
-               data = all_t_test.df[which(all_t_test.df$dataset != "cf" & all_t_test.df$source != "ext"),]) +
+               data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
   scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
@@
   dplyr::filter(dataset != "raw") %>%
-  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  mutate(source = factor(source, levels=c("low", "high"), ordered=TRUE)) %>%
   ggplot(aes(x=dataset, y=female_ratio)) +
@@
   labs(y="Proportion Female", x = "Dataset")
-
+#ggsave(filename="figs/all_cent_sum_boxplt_wtest.png", all_cent_sum_boxplt_wtest)
 ```
@@
       pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
+      TRUE ~ NA_character_))) %>%
+      separate_wider_delim(cols=comp, delim="_", names=c("group1", "group2"), cols_remove=FALSE) %>%
+  mutate(smaller_group=ifelse(bigger_group==group1, group2, group1))
 
@@
 noext.cent.t.test.df2 <- noext.cent.feat.t.test.df %>%
-  mutate(full_df = 0)
+  mutate(full_df = "No Extremes")
 cent.feat.t.test.df2 <- cent.feat.t.test.df %>%
-  mutate(full_df = 1)
+  mutate(full_df = "Full Dataset")
 
@@
   facet_grid(comp ~ full_df)
+
+comp_names <- c(cf.gam_cf = "ComBat-GAM v.
+\nComBat w/o Covar", cf.gamlss_cf = "ComBatLS v.
+ComBat w/o Covar", cf.gamlss_cf.gam = "ComBatLS v.
+ComBat-GAM", cf.lm_cf.gam ="ComBat-GAM v.
+linear ComBat", cf.lm_cf.gamlss = "ComBatLS v.
+linear ComBat", cf.lm_cf = "linear ComBat v.
+\nComBat w/o Covar")
+cent.test.comp.plt2 <- comp.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+  mutate(smaller_group=factor(smaller_group, levels=c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered=TRUE)) %>%
+ggplot() +
+  geom_bar(aes(x=perm, fill=smaller_group)) +
+  theme_linedraw()+
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "bottom") +
+  facet_grid(comp ~ full_df, labeller = labeller(comp = as_labeller(comp_names))) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "Better ComBat \nConfiguration", labels=dataset_names) +
+  labs(x="Sampling Permutation", y="Feature Count")
+ggsave(filename="figs/cent_comp_noext.png", cent.test.comp.plt2, height=13, width=13, units="in")
 
@@
                                p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ ""))
+                               p.val_fdr >= 0.05 ~ ""),
+         y_pos = case_when(dataset=="cf" ~ 6,
+                           dataset=="cf.lm" ~ 9,
+                           dataset=="cf.gam" ~ 9.5,
+                           dataset=="cf.gamlss" ~ 5))
 
@@
 
-noext.sex.testmed %>%
+sex_bias_density_plt_noext <- noext.sex.testmed %>%
   #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med$dataset == "cf.gam", 10.2, 5.5)),
+  geom_text(aes(label = sig_fdr, x= x_pos*100, y=y_pos),
               data = next.t.test.cent.sex.test.med) +
@@
   labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
+ggsave(filename="figs/sex_bias_density_plt_noext.png", sex_bias_density_plt_noext, height=7, width=10, units="in")
 ```
@@
       pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
+      TRUE ~ NA_character_))) %>%
+  separate_wider_delim(cols=comp, delim="_", names=c("group1", "group2"), cols_remove=FALSE) %>%
+  mutate(smaller_group=ifelse(bigger_group==group1, group2, group1))
 
@@
 ```{r}
-z.feat.t.test.df %>%
+comp_names <- c(cf.gam_cf = "ComBat-GAM v.
+ComBat w/o Covar", cf.gamlss_cf = "ComBatLS v.
+ComBat w/o Covar", cf.gamlss_cf.gam = "ComBatLS v.
+ComBat-GAM", cf.lm_cf.gam ="ComBat-GAM v.
+linear ComBat", cf.lm_cf.gamlss = "ComBatLS v.
+linear ComBat", cf.lm_cf = "linear ComBat v.
+ComBat w/o Covar")
+
+z.feat.t.test.plt <- z.feat.t.test.df %>%
   dplyr::filter(sig_fdr==TRUE) %>%
 ggplot() +
-  geom_bar(aes(x=perm, fill=bigger_group)) +
-  theme(axis.text.x = element_blank()) +
-  facet_grid(pheno_cat~comp, scales="free_y", labeller = labeller(comp = as_labeller(comp_names)))
+  geom_bar(aes(x=perm, fill=smaller_group)) +
+  theme_linedraw() + 
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "bottom") +
+  labs(x="Sampling Permutation", y="Feature Count") +
+  facet_grid(pheno_cat~comp, scales="free_y", labeller = labeller(comp = as_labeller(comp_names))) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "Better ComBat \nConfiguration", labels=dataset_names)
+
+ggsave(filename="figs/z.feat.t.test.plt.png", z.feat.t.test.plt, height=7, width=13, units="in")
 ```
@@
   scale_x_discrete(labels=c(dataset_names)) +
+  theme_linedraw()+
   theme(legend.position = "none") +
-  ggtitle("Within-subject mean magnitude of error in z scores across ComBat \nconfigurations replicated 10x")
+  ggtitle("Within-subject mean magnitude of error in z scores across ComBat configurations replicated 100x")
 z.diffs.v.plot
 # ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
+ggsave(filename="figs/zscore_violin.png", z.diffs.v.plot, width=11, height=4.5, units="in")
+```
+
+```{r}
+z.tbl <- z.feat.t.test.df %>%
+  group_by(comp, smaller_group) %>%
+  summarize(n_sig=sum(sig_fdr)) %>%
+  mutate(perc_tests = n_sig/(208*100))
+
+z.tbl.pretty <- z.tbl %>%
+  mutate(comp = case_when(comp == "cf.gam_cf" ~ "ComBat-GAM vs ComBat w/o Covariates",
+                          comp == "cf.gamlss_cf" ~ "ComBatLS vs ComBat w/o Covariates", 
+                          comp == "cf.gamlss_cf.gam" ~ "ComBatLS vs ComBat-GAM",
+                          comp == "cf.lm_cf" ~ "Linear ComBat vs ComBat w/o Covariates",
+                          comp == "cf.lm_cf.gam" ~ "Linear ComBat vs ComBat-GAM",
+                          comp == "cf.lm_cf.gamlss" ~ "Linear ComBat vs ComBatLS"),
+         smaller_group = case_when(smaller_group == "cf" ~ "ComBat w/o Covariates",
+                                   smaller_group == "cf.lm" ~ "Linear ComBat",
+                                   smaller_group == "cf.gam" ~ "ComBat-GAM",
+                                   smaller_group == "cf.gamlss" ~ "ComBatLS")) %>%
+  rename(#"Pairwise Comparison"=comp,
+         "Method Producing Smaller Absolute Errors" = smaller_group,
+         "N Features" = n_sig,
+         "% Features" = perc_tests)
+
+pretty_z_table <- z.tbl.pretty %>%
+  gt() %>% 
+  #gt_theme_538() %>% 
+tab_style(style = list(cell_fill(color = "gray95")),
+    locations = cells_row_groups()) %>%
+  fmt_percent(
+    columns = "% Features",
+    decimals = 2
+  ) %>%
+  tab_options(column_labels.font.weight = "bold") %>%
+  tab_header(title = "Pairwise comparisons of ComBat methods' absolute Z-score errors for 208 brain features across 100 replications")
+
+gtsave(pretty_z_table, filename="figs/z_table.png")
 ```
@@
 Distribution of error medians w/in each pheno category
+
+```{r}
+#look by feature type
+z.sex_bias_hist_plt <- z.sex.test.df %>%
+  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
+  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  mutate(pheno = gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    perm = as.factor(sub("perm-0", "", perm))) %>%
+ggplot() +
+  geom_histogram(aes(x=median_sex_diff, color=perm, fill=perm), alpha=0.3, position="identity", bins=40) +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(pheno_cat~dataset, scales="free", labeller = labeller(dataset = as_labeller(dataset_names))) +
+  theme_linedraw(base_size=12) +
+  theme(legend.position = "none") +
+  labs(x="Male - Female Median Centile Errors", y="Number of Features") +
+  ggtitle("Sex differences in median centile errors across 100 sampling replications")
+ggsave(filename="figs/z.sex_bias_hist_plt.z.png", sex_bias_hist_plt)
+```
+
 ```{r}
@@
                                p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ ""))
-
-t.test.z.sex.test.med.filt <- t.test.z.sex.test.med %>% dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")
+                               p.val_fdr >= 0.05 ~ ""),
+         y_pos = case_when(dataset=="cf" ~ 155,
+                           dataset=="cf.lm" ~ 360,
+                           dataset=="cf.gam" ~ 345,
+                           dataset=="cf.gamlss" ~ 125))
+
+t.test.z.sex.test.med.filt <- t.test.z.sex.test.med %>% 
+  dplyr::filter(dataset != "cf")
 
@@
 z.sex_bias_density_plt <- z.sex.test.med %>%
-  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  dplyr::filter(dataset != "cf") %>%
   ggplot() +
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos, y=ifelse(t.test.z.sex.test.med.filt$dataset == "cf.gam", 350, 120)),
+  geom_text(aes(label = sig_fdr, x= x_pos, y=y_pos),
               data = t.test.z.sex.test.med.filt) +
@@
   scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  theme(legend.position = "bottom")+
+  theme(legend.position = "none")+
   labs(x="Medians of Sex-Differences in Z Errors", y="Density Across 100 Replications")
 
-z.sex.test.med %>%
-  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+z.sex_bias_density_plt2 <- z.sex.test.med %>%
   ggplot() +
@@
   geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
-  facet_wrap(~pheno_cat) +
+  facet_wrap(~pheno_cat, scales="free") +
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos, y=ifelse(t.test.z.sex.test.med$dataset == "cf.gam", 350, 120)),
+  geom_text(aes(label = sig_fdr, x= x_pos, y=y_pos),
               data = t.test.z.sex.test.med) +
@@
   scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  theme(legend.position = "bottom")+
+  theme(legend.position = "bottom") +
   labs(x="Medians of Sex-Differences in Z Errors", y="Density Across 100 Replications")
+ggsave(filename="figs/z.sex_bias_density_plt2.png", z.sex_bias_density_plt2, height=6, width=14, units="in")
 ```
@@
   )
-
-ext_z.sum <- perm_subj.df.z._long %>%
-  dplyr::filter(pheno_cat != "allpheno") %>%
-  dplyr::filter(mean_z < -1.5 | mean_z > 1.5) %>%
-  group_by(dataset, perm, pheno_cat) %>%
-  summarise(
-    male_ratio = sum(sex == "Male") / n(),
-    female_ratio = sum(sex == "Female") / n()
-  )
 ```
@@
 
-ext_z.sum_raw <- ext_z.sum %>%
-  dplyr::filter(dataset == "raw" & perm=="perm-001")
-
 z.high_t_test.df <- lapply(split(high_z.sum, high_z.sum$pheno_cat), get_t_tests, raw_df=high_z.sum_raw) %>% bind_rows()
 z.low_t_test.df <- lapply(split(low_z.sum, low_z.sum$pheno_cat), get_t_tests, raw_df=low_z.sum_raw) %>% bind_rows()
-z.ext_t_test.df <- lapply(split(ext_z.sum, ext_z.sum$pheno_cat), get_t_tests, raw_df=ext_z.sum_raw) %>% bind_rows()
 
 #merge dfs
-all_z_sum.df <- gdata::combine(high_z.sum, low_z.sum, ext_z.sum) %>%
+all_z_sum.df <- gdata::combine(high_z.sum, low_z.sum) %>%
   mutate(source = gsub("_z.sum", "", source))
@@
 
-z.all_t_test.df <- gdata::combine(z.high_t_test.df, z.low_t_test.df, z.ext_t_test.df) %>%
+z.all_t_test.df <- gdata::combine(z.high_t_test.df, z.low_t_test.df) %>%
   mutate(source = gsub("_t_test.df", "", source)) %>%
@@
 
-z_ext_list <- c(ext = ">1.5 and <-1.5", high = ">1.5", low = "<-1.5")
+z_ext_list <- c(high = ">1.5", low = "<-1.5")
 
@@
   dplyr::filter(dataset != "raw" & dataset != "cf") %>%
-  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  mutate(source = factor(source, levels=c("low", "high"), ordered=TRUE)) %>%
   ggplot(aes(x=dataset, y=female_ratio_diff)) +
@@
 
-all_z_sum.df_scaled %>%
+z_sex_boxplts <- all_z_sum.df_scaled %>%
   dplyr::filter(dataset != "raw") %>%
-  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  mutate(source = factor(source, levels=c("low", "high"), ordered=TRUE)) %>%
   ggplot(aes(x=dataset, y=female_ratio_diff)) +
@@
   labs(y="Proportion Female", x = "Dataset")
+ggsave("figs/z_sex_boxplts.jpeg", z_sex_boxplts)
 ```
@@
 noext.z.t.test.df2 <- noext.z.feat.t.test.df %>%
-  mutate(full_df = 0)
+  mutate(full_df = "No Extremes")
 z.feat.t.test.df2 <- z.feat.t.test.df %>%
-  mutate(full_df = 1)
+  mutate(full_df = "Full Dataset")
 
 z.comp.df <- full_join(noext.z.t.test.df2, z.feat.t.test.df2)
-z.comp.df %>%
+
+comp_names <- c(cf.gam_cf = "ComBat-GAM v.
+\nComBat w/o Covar", cf.gamlss_cf = "ComBatLS v.
+ComBat w/o Covar", cf.gamlss_cf.gam = "ComBatLS v.
+ComBat-GAM", cf.lm_cf.gam ="ComBat-GAM v.
+linear ComBat", cf.lm_cf.gamlss = "ComBatLS v.
+linear ComBat", cf.lm_cf = "linear ComBat v.
+\nComBat w/o Covar")
+
+z.test.comp.plt2 <- z.comp.df %>%
   dplyr::filter(sig_fdr==TRUE) %>%
-ggplot() +
-  geom_bar(aes(x=perm, fill=bigger_group)) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_grid(comp ~ full_df)
+  mutate(smaller_group=factor(smaller_group, levels=c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered=TRUE)) %>%
+ggplot() +
+  geom_bar(aes(x=perm, fill=smaller_group)) +
+  theme_linedraw()+
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "bottom") +
+  facet_grid(comp ~ full_df, labeller = labeller(comp = as_labeller(comp_names))) +
+  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "Better ComBat \nConfiguration", labels=dataset_names) +
+  labs(x="Sampling Permutation", y="Feature Count")
+ggsave(filename="figs/z_comp_noext.png", z.test.comp.plt2, height=13, width=13, units="in")
 
@@
                                p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ ""))
+                               p.val_fdr >= 0.05 ~ ""),
+         y_pos = case_when(dataset=="cf" ~ 150,
+                           dataset=="cf.lm" ~ 320,
+                           dataset=="cf.gam" ~ 330,
+                           dataset=="cf.gamlss" ~ 125))
 
@@
 
-noext.sex.test.z.med %>%
+z.sex_bias_density_plt_noext <- noext.sex.test.z.med %>%
   #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos, y=ifelse(next.t.test.z.sex.test.med$dataset == "cf.gam", 320, 100)),
+  geom_text(aes(label = sig_fdr, x= x_pos, y=y_pos),
               data = next.t.test.z.sex.test.med) +
@@
   theme(legend.position = "bottom")+
-  labs(x="Medians of Sex-Differences in zile Errors", y="Density Across 100 Replications")
+  labs(x="Medians of Sex-Differences in Z-Score Errors", y="Density Across 100 Replications")
+ggsave(filename="figs/z_sex_bias_density_plt_noext.png", z.sex_bias_density_plt_noext, height=7, width=10, units="in")
 ```
@@
 sex.var.plot <- readRDS("figs/sex_var_plot.RDS")
-
+```
+
+```{r}
 fig_3 <- ggarrange(
   ggarrange(
-  ggarrange(blank, sex.var.plot.global, sex.var.plot,
-  ncol=3, nrow=1, 
-  widths = c(.1, .8, 3.1),
+  ggarrange(blank, sex.var.plot.global, sex.var.plot, blank,
+  ncol=4, nrow=1, 
+  widths = c(.3, .7, 2.9, .4),
   hjust=0.01,
@@
 
-fig_3
+#fig_3
 ggsave("figs/figure_3.jpeg", fig_3, width=15, height=10, units="in")
@@
 
-Figure 3 but with sig. tests for boxplots
+Figure 3 but with ComBat w/o covariates
 

### 2024-06-06 eda/lifespan_eda.Rmd
lots of supplemental plots

@@
 levels(cn.df.unique$study)
+n_unique(cn.df.unique$study)
 ```
@@
 
-Removed `r nrow(cn.df.unique) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain
+Removed `r nrow(cn.df.noinf) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain
 

### 2024-06-06 paper_lbcc_results.Rmd
lots of supplemental plots

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize, gratia)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize)
 #Sys.setenv("MC_CORES"=10L)
@@
 
-source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
+#source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
 #source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
@@
 
-#key to get study names
-study_site_key <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/lifespan_study_site_key.csv")
-
 #PATH
@@
 gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv"))
-
-gam_centile_df <- full_join(gam_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)
-n_unique(gam_centile_df$study)
+gam_centile_df <- gam_centile_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
 
@@
 gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv"))
-gamlss_centile_df <- full_join(gamlss_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)
+gamlss_centile_df <- gamlss_centile_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
 
@@
 raw_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_predictions.csv"))
-raw_centile_df <- full_join(raw_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)
+raw_centile_df <- raw_centile_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
 
@@
 deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
-deltas_df <- full_join(deltas_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)
+deltas_df <- deltas_df %>% mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25)
 
@@
 ```
-
-Nature paper models to look at age effects in variance
-from [https://github.com/brainchart/Lifespan/tree/main/Share](https://github.com/brainchart/Lifespan/tree/main/Share)
-
-```{r}
-paper_fit_sgmv <- readRDS("derivatives/nature_2020_mods/FIT_sGMV.rds")
-paper_fit_gmv <- readRDS("derivatives/nature_2020_mods/FIT_GMV.rds")
-paper_fit_wmv <- readRDS("derivatives/nature_2020_mods/FIT_WMV.rds")
-paper_fit_ven <- readRDS("derivatives/nature_2020_mods/FIT_Ventricles.rds")
-```
-
 #Clean & EDA
-
-Check numbers
-```{r}
-nrow(gamlss_centile_df) #n
-n_unique(gamlss_centile_df$study) 
-n_unique(gamlss_centile_df$site) #correct
-```
 
@@
 #clean predictions
-raw_centile_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
+raw_centile_df.clean <- raw_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
   dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 
-table(df_raw_feat_aov.study$sig_fdr)
+df_raw_feat_aov.study %>%
+  ggplot() +
+  geom_histogram(aes(x=etasq, fill=sig_fdr), alpha=0.5)
+
+df_raw_feat_aov.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
 ```
@@
 ```
-
-ANOVA of subjects' mean centiles across studies
-
-```{r}
+linear model
+
+```{r}
+summary(lm(mean_cent ~ study, data=raw_centile_df.mean))
 summary(aov(mean_cent ~ study, data=raw_centile_df.mean))
@@
 #clean predictions
-raw_z_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
+raw_z_df.clean <- raw_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 
-table(df_raw_feat_aov.z.study$sig_fdr)
+df_raw_feat_aov.z.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
 ```
@@
 
-ANOVA of subject means
-
-```{r}
+```{r}
+raw_z_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
+# +
+#   facet_wrap(~study)
+```
+
+linear model
+
+```{r}
+summary(lm(mean_z ~ study, data=raw_z_df.mean))
 summary(aov(mean_z ~ study, data=raw_z_df.mean))
@@
 #clean predictions
-gam_centile_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)%>%
+gam_centile_df.clean <- gam_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
   dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 
-table(df_gam_feat_aov.study$sig_fdr)
+df_gam_feat_aov.study %>%
+  ggplot() +
+  geom_histogram(aes(x=etasq, fill=sig_fdr), alpha=0.5)
+
+df_gam_feat_aov.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
 ```
@@
 
-ANOVA
-
-```{r}
+linear model
+
+```{r}
+summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
 summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
@@
 #clean predictions
-gam_z_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
+gam_z_df.clean <- gam_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 
-table(df_gam_feat_aov.z.study$sig_fdr)
+df_gam_feat_aov.z.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
 ```
@@
   geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
-```
-
-Subject mean ANOVA
-
-```{r}
+# +
+#   facet_wrap(~study)
+```
+
+linear model
+
+```{r}
+summary(lm(mean_z ~ study, data=gam_z_df.mean))
 summary(aov(mean_z ~ study, data=gam_z_df.mean))
@@
 #clean predictions
-gamlss_centile_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
+gamlss_centile_df.clean <- gamlss_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
   dplyr::select(!contains(".z")) #drop z-scores and non-centiles & !ends_with("_total") & !contains("TBV")
@@
 
+```{r}
+ggplot(gamlss_centile_df.clean) +
+  geom_point(aes(x=age_days, y=rh_Vol_insula, color=study), alpha=0.5) +
+  theme(legend.position="none")
+```
+
 w/in feature ANOVAs
@@
 gamlss.feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))
+length(gamlss.feat_to_test)
 
@@
 
-table(df_gamlss_feat_aov.study$sig_fdr)
+df_gamlss_feat_aov.study %>%
+  ggplot() +
+  geom_histogram(aes(x=etasq, fill=sig_fdr))
+
+df_gamlss_feat_aov.study %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+  nrow()
 ```
@@
   theme(axis.text.x = element_text(angle=90, hjust=0.5))
-```
-
-subject mean ANOVA
-
-```{r}
+
+gamlss_centile_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_cent, color=study), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_cent, color=study), method="lm") +
+  theme(legend.position = "none")
+
+gamlss_centile_df.mean %>%
+  dplyr::filter(age_days>365) %>%
+  ggplot(aes(x=age_days/365.25, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+linear model
+
+```{r}
+summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
 summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
@@
 #clean predictions
-gamlss_z_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
+gamlss_z_df.clean <- gamlss_centile_df %>%
+  mutate(study=sub("\\_.*", "", site),
+         age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 
-table(df_gamlss_feat_aov.z.study$sig_fdr)
+df_gamlss_feat_aov.z.study %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  nrow()
 ```
@@
 
-subject mean ANOVA
-
-```{r}
+linear model
+
+```{r}
+summary(lm(mean_z ~ study, data=gamlss_z_df.mean))
 summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
@@
 
-Number of models converged in each dataset
-```{r}
-length(raw.feat_to_test)
-length(gam.feat_to_test)
-length(gamlss.feat_to_test)
-
-#missing features
-c(setdiff(pheno_list, gamlss.feat_to_test),
-setdiff(pheno_list, raw.feat_to_test),
-setdiff(pheno_list, gam.feat_to_test)) %>% unique()
+compare magnitude of effect size for study (w/in feature anovas) across unharmonized, combat-gam and combatLS centiles
+
+```{r}
+setdiff(gam.feat_to_test, gamlss.feat_to_test)
+setdiff(gamlss.feat_to_test, gam.feat_to_test)
+setdiff(pheno_list, gamlss.feat_to_test)
+```
+
+```{r}
+df_gamlss_feat_aov.study$combat <- "gamlss"
+df_gam_feat_aov.study$combat <- "gam"
+df_raw_feat_aov.study$combat <- "raw"
+
+#put all data frames into list
+df_list.tmp <- list(df_gamlss_feat_aov.study, df_gam_feat_aov.study, df_raw_feat_aov.study)
+
+#merge all data frames in list
+feat_aov.df <- df_list.tmp %>% reduce(full_join)
+n_unique(feat_aov.df$pheno)
+sum(is.na(feat_aov.df))
+
+feat_aov.df <- feat_aov.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+feat_aov.df %>%
+  ggplot() +
+  geom_point(aes(x=etasq, y=pheno, color=combat, shape=sig_fdr), alpha=0.6) +
+  facet_wrap(~pheno_cat, scales="free")
 ```
@@
   labs(y="Mean Centile", x = "Study")
-
-mean_cents_bxplt_nolabel <- mean_cents.df %>%
-  ggplot() +
-  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
-  theme_linedraw() +
-  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none", strip.text.x = element_text(size = 12)) +
-  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
-  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
-  labs(y="Mean Centile", x = "Study")
-```
-
-Cohen's F squared for study effects
-
-```{r}
-gam_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv"))
-gam_centile_f2$combat <- "ComBat-GAM"
-
-gamlss_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_cohenfsq.csv"))
-gamlss_centile_f2$combat <- "ComBatLS"
-
-f2.df <- rbind(gam_centile_f2, gamlss_centile_f2)
-f2.df <- f2.df %>%
-  mutate(pheno_cat = as.factor(case_when(
-      pheno %in% vol_list_global ~ "Global Volume",
-      pheno %in% vol_list_regions ~ "Regional Volume",
-      pheno %in% sa_list ~ "Regional SA",
-      pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
-
-fsq.df <- ggplot(f2.df) +
-  geom_col(aes(x=pheno, y=fsq, fill=combat), position="dodge") + 
-  facet_wrap(~pheno_cat, scales="free") +
-  theme_linedraw() +
-  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
-  labs(x = "Feature", y = "Effect Size (Cohen's F-Squared)", fill="ComBat")
-
-ggsave("figs/supp_fsq.jpeg", fsq.df, width=8, height=6, units="in")
-```
-
-```{r}
-f2.df.sum <- f2.df %>%
-  group_by(pheno) %>%
-  arrange(combat) %>%
-  mutate(F2_diff = fsq - fsq[1]) #gamlss - gam value, + number means site effect is larger in combatLS
-
-ggplot(f2.df.sum) +
-  geom_col(aes(x=pheno, y=F2_diff), position="dodge") + 
-  facet_wrap(~pheno_cat, scales="free") +
-  theme(axis.text.x = element_blank())
-
-f2.df.sum %>%
-  group_by(pheno_cat) %>%
-  summarise(mean_fsq_diff = mean(F2_diff)) %>%
-  ggplot() + 
-  geom_col(aes(x=pheno_cat, y=mean_fsq_diff), position="dodge")
-```
-
-#ComBat-GAM vs ComBatLS: Delta Centile
+```
+
+#Feature Centile ~ Mean Study Age
 
@@
 ```{r}
-nrow(raw_centile_df.clean) == n_unique(raw_centile_df.clean$participant)
-
 study_info <- raw_centile_df.clean %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
   group_by(study) %>%
@@
             study_median_age=median(age_years),
-            study_n = n()) %>%
+            study_n= n()) %>%
   dplyr::select(study, study_mean_age, study_median_age, study_n)
 ```
+
+W/in feat lms
+
+##Unharmonized
+
+```{r}
+raw_centile_df.clean_study <- full_join(raw_centile_df.clean, study_info)
+
+df_raw_feat_lm <- data.frame()
+for(pheno in raw.feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(raw_centile_df.clean_study[[pheno]] ~ raw_centile_df.clean_study[["study_mean_age"]]))
+  df$pheno <- pheno
+  df_raw_feat_lm <- rbind(df_raw_feat_lm, df)
+}
+
+df_raw_feat_lm <- df_raw_feat_lm %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_raw_feat_lm %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+
+df_raw_feat_lm %>%
+  ggplot() +
+  geom_histogram(aes(x=statistic))
+```
+
+##ComBat-GAM
+
+```{r}
+gam_centile_df.clean_study <- full_join(gam_centile_df.clean, study_info)
+
+df_gam_feat_lm <- data.frame()
+for(pheno in gam.feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(gam_centile_df.clean_study[[pheno]] ~ gam_centile_df.clean_study[["study_mean_age"]]))
+  df$pheno <- pheno
+  df_gam_feat_lm <- rbind(df_gam_feat_lm, df)
+}
+
+df_gam_feat_lm <- df_gam_feat_lm %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_gam_feat_lm %>%
+  ggplot() +
+  geom_histogram(aes(x=statistic))
+```
+
+##ComBatLS
+
+```{r}
+gamlss_centile_df.clean_study <- full_join(gamlss_centile_df.clean, study_info)
+
+df_gamlss_feat_lm <- data.frame()
+for(pheno in gamlss.feat_to_test){
+  # print(pheno)
+  df <- tidy(aov(gamlss_centile_df.clean_study[[pheno]] ~ gamlss_centile_df.clean_study[["study_mean_age"]]))
+  df$pheno <- pheno
+  df_gamlss_feat_lm <- rbind(df_gamlss_feat_lm, df)
+}
+
+df_gamlss_feat_lm <- df_gamlss_feat_lm %>%
+  dplyr::filter(term != "Residuals") %>%
+  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
+                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
+                                      p.val_fdr >= 0.05 ~ FALSE))
+df_gamlss_feat_lm %>%
+  ggplot() +
+  geom_bar(aes(y=sig_fdr))
+
+df_gamlss_feat_lm %>%
+  ggplot() +
+  geom_histogram(aes(x=statistic))
+```
+
+*statistic ~ effect size?
+
+```{r}
+df_raw_feat_lm$combat <- "raw"
+df_gam_feat_lm$combat <- "gam"
+df_gamlss_feat_lm$combat <- "gamlss"
+
+#put all data frames into list
+df_list.tmp <- list(df_raw_feat_lm, df_gam_feat_lm, df_gamlss_feat_lm)
+
+#merge all data frames in list
+all_feat_lm <- df_list.tmp %>% reduce(full_join)
+
+all_feat_lm %>%
+  ggplot() +
+  geom_histogram(aes(x=statistic, fill=combat, color=combat), position="identity", alpha=0.5, bins=50)
+```
+
+#Subject-Level Mean Centiles ~ Mean Study Age
+
+```{r}
+raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
+summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
+
+raw_centile_df.mean2 %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+```{r}
+gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
+summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
+
+gam_centile_df.mean2 %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+
+gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age)
+
+summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
+gam_centile_df.mean2 %>%
+  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+```{r}
+gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
+summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
+
+gamlss_centile_df.mean2 %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+
+gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age)
+
+summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
+gamlss_centile_df.mean2 %>%
+  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
+```
+
+Plots
+
+```{r}
+gamlss_centile_df.mean2$combat <- "gamlss"
+gam_centile_df.mean2$combat <- "gam"
+raw_centile_df.mean2$combat <- "raw"
+
+#put all data frames into list
+df_list.tmp <- list(gamlss_centile_df.mean2, gam_centile_df.mean2, raw_centile_df.mean2)
+
+#merge all data frames in list
+all_subj_means.df <- df_list.tmp %>% reduce(full_join)
+
+all_subj_means.df %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.6) +
+  geom_smooth(method="lm")
+
+all_subj_means.df %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_wrap(~combat)
+
+summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
+```
+
+```{r}
+summary(lm(mean_cent ~ study_n, data=raw_centile_df.mean2))
+summary(lm(mean_cent ~ study_n, data=gam_centile_df.mean2))
+summary(lm(mean_cent ~ study_n, data=gamlss_centile_df.mean2))
+
+all_subj_means.df %>%
+  ggplot(aes(x=study_n, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.6) +
+  geom_smooth(method="lm")
+```
+
+#ComBat-GAM vs ComBatLS: Delta Centile
 
@@
 
-control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.
-
 check mean subj delta ~ study mean age
@@
   theme(legend.position="none")
-```
-
-See if a subject's distance from their study's mean age is assoc with how much their centile changes
-
-```{r}
+
 mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
@@
 
-## Subject-Level Mean Centiles ~ Mean Study Age
-
-Make sure that results of mean centile delta ~ mean study age are related to the difference itself, not because ComBatLS and/or ComBat-GAM centiles are associated with mean study age.
-
-```{r}
-raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
-summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
-
-raw_centile_df.mean2 %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-NS
-
-```{r}
-gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
-summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
-
-gam_centile_df.mean2 %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-
-gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
-  mutate(age_diff = (age_days/365.25) - study_mean_age)
-
-summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
-gam_centile_df.mean2 %>%
-  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-NS
-
-```{r}
-gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
-summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
-
-gamlss_centile_df.mean2 %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-
-gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
-  mutate(age_diff = (age_days/365.25) - study_mean_age)
-
-summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
-gamlss_centile_df.mean2 %>%
-  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-NS
+control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting
 
@@
 
-Reformatting plots for prettier composite fig
+```{r}
+# study demographics
+fig5 <- ggarrange(
+ggarrange(sample_boxplot, sample_barplot,
+          full_boxplot, full_barplot,
+          nrow=2, ncol=2, 
+          #align = "h", 
+          common.legend = TRUE,
+          legend = "bottom",
+          widths=c(1.7, 1),
+          heights=c(10, 1)),
+# boxplots
+mean_cents_bxplt,
+common.legend=FALSE,
+nrow=1, ncol=2,
+labels="AUTO",
+widths=c(1, 4)
+)
+
+ggsave("figs/figure_5_horizontal.jpeg", fig5, width=18, height=9, units="in")
+```
+alternative vertical stacking
 
@@
   theme(axis.title.x=element_blank(), axis.text.x = element_blank()) +
-  labs(color="Sex", y = "Age (Years)")
+  labs(color="Sex", y = "Age (Years) \n ")
 
@@
   gamlss_centile_df %>%
-  mutate(xlab = "Full Sample ") %>%
+  mutate(xlab = "Full Sample") %>%
   ggplot(aes(x=xlab, y=age_years, color=sex)) +
@@
   geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
-  #scale_y_continuous(labels = scales::percent) +
+  scale_y_continuous(labels = scales::percent) +
   geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
@@
   #theme(axis.ticks.x = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
-  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=6)) +
-  labs(fill="Sex", x = "Study", y="Percent Sex")
+  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
+  labs(fill="Sex", x = "Study", y="Sex")
 
@@
   theme_linedraw(base_size = 11) +
-  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
+  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
   labs(color="Sex", y = "Age (Years)", y="")
+
+ggarrange(sample_boxplot2, full_boxplot2,
+          sample_barplot2, full_barplot2,
+          nrow=2, ncol=2, 
+          align = "hv", 
+          common.legend = TRUE,
+          legend = "right",
+          widths=c(10, 1),
+          heights=c(2, 1)
+          )
 ```
@@
 fig5_vert <- ggarrange(
-  #ggarrange(blank,
+  ggarrange(blank,
 ggarrange(sample_boxplot2, full_boxplot2,
@@
           ),
-# blank,
-# nrow=1, ncol=3,
-# common.legend = FALSE,
-# widths= c(1, 20, 1)),
+blank,
+nrow=1, ncol=3,
+common.legend = FALSE,
+widths= c(1, 20, 1)),
 # boxplots
-mean_cents_bxplt_nolabel,
+mean_cents_bxplt,
 common.legend=FALSE,
@@
 
-mean_delta_fig <- mean_cent_deltas_df_study %>%
+mean_cent_deltas_df_study %>%
   ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
@@
   theme(legend.position="none") +
-  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From ComBatLS and ComBat-GAM \n-harmonized Data")
-ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=9, height=6, units="in")
-```
-
-lines show marginal effects of mean study age
+  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
+```
 
@@
 
-mean_delta_fig2 <- mean_cent_deltas_df_study %>%
+mean_cent_deltas_df_study %>%
   ggplot(aes(x=abs_age_diff, y=abs_mean_delta_cent)) +
@@
   theme(legend.position="none") +
-  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From ComBatLS and ComBat-GAM \n-harmonized Data")
-ggsave("figs/supp_mean_delta_fig2.jpeg", mean_delta_fig2, width=9, height=6, units="in")
-```
+  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From \nComBatLS and ComBat-GAM-harmonized Data")
+```

### 2024-06-06 paper_main_results.Rmd
lots of supplemental plots

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x, skimr)
 ```
@@
 
-get stat results for global phenos:
-```{r, message=FALSE}
-full_cent_res %>%
-  dplyr::filter(pheno %in% vol_list_global) %>%
-  dplyr::filter(!grepl('_refA', comp)) %>%
-  summarise(n_sig = sum(sig_fdr),
-            nrow = nrow(.))
-
-#n occurances of combatLS having significantly larger errors
-full_cent_res %>%
-  dplyr::filter(bigger_group == "cf.gamlss")%>%
-  dplyr::filter(!grepl('_refA', comp)) %>%
-  summarise(n_sig = sum(sig_fdr),
-            nrow = nrow(.))
-```
-
 ```{r}
@@
 
+#wilcox test for visualizing significance - just going to use it for positioning info & setting up dataframe, will use actual tests run on CUBIC for signfificance
+mean_comp_tbl_calc <- global_abs.diffs %>%
+  group_by(pheno) %>%
+  pairwise_wilcox_test(abs.diff ~ dataset, comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss")),
+                       p.adjust.method = "fdr")
+
+mean_comp_tbl <- mean_comp_tbl_calc %>%
+  mutate(comp = factor(paste(group2, group1, sep="_"), ordered=TRUE, levels=c("cf.gam_cf", "cf.gamlss_cf", "cf.gamlss_cf.gam", "cf.lm_cf", "cf.lm_cf.gam", "cf.lm_cf.gamlss")),
+         pheno = factor(pheno, ordered=T, levels=c("GMV", "sGMV", "Ventricles", "WMV"))) %>%
+  arrange(pheno, comp) %>%
+  add_xy_position(x="pheno", fun="max", comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss"), c("cf.lm", "cf"), c("cf", "cf.gam"), c("cf", "cf.gamlss")))
+
+#get actual stats for global phenos
+global_cent_res <- full_cent_res %>%
+  dplyr::filter(pheno %in% vol_list_global) %>%
+  mutate(comp = factor(comp, ordered=TRUE, levels=c("cf.gam_cf", "cf.gamlss_cf", "cf.gamlss_cf.gam", "cf.lm_cf", "cf.lm_cf.gam", "cf.lm_cf.gamlss")),
+         pheno = factor(pheno, ordered=T, levels=c("GMV", "sGMV", "Ventricles", "WMV")))
+
+mean_comp_tbl2 <- inner_join(mean_comp_tbl, global_cent_res) %>%
+  add_significance("p.val_fdr", cutpoints = c(0, 0.001, 0.01, 0.05, 1),
+  symbols = c("***", "**", "*", "ns"))
+
+#get max w/in each pheno, prop & dataset for placement
+max_abs.errs <- global_abs.diffs %>%
+  mutate(pheno = factor(pheno, ordered=T, levels=c("GMV", "sGMV", "Ventricles", "WMV"))) %>%
+  group_by(pheno, dataset) %>%
+  summarize(max_err = max(abs.diff)*100) %>%
+  ungroup() %>%
+  pivot_wider(id_cols=c(pheno), names_from=dataset, values_from=max_err, names_prefix="max_")
+
+stats_centile_for_plot <- full_join(mean_comp_tbl2, max_abs.errs)
+
+#kinda hacky way of fixing positions
+stats_centile_for_plot.corrected <- stats_centile_for_plot %>%
+  mutate(xmax=case_when(group1=="cf.gamlss" | group2== "cf.gamlss" ~ (x+.266667),
+                        TRUE ~  x),
+         xmin=case_when(group2=="cf.lm" ~ (x-.266667),
+                        TRUE ~  x),
+         y.position=case_when(comp == "cf.lm_cf.gamlss" ~ max_cf.lm + 1,
+                              comp == "cf.lm_cf.gam" ~ max_cf.lm + .5,
+                              TRUE ~ max_cf.gam + 0.4))
+```
+
+```{r}
 global_cent_errs_plot <- global_abs.diffs %>%
@@
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
+         #pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
          abs.diff = abs.diff*100) %>% #for visuals
@@
   scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  #WMV
-geom_signif(y_position = c(3, 3.5, 2.5), xmin = c(0.7, 0.7, 1),
-              xmax = c(1, 1.3, 1.3), annotation = c("***","***", "***"),
-              tip_length = 0, vjust = 0.6) +
-  #GMV
-  geom_signif(y_position = c(2.75, 3.25, 2.25), xmin = c(1.7, 1.7, 2),
-              xmax = c(2, 2.3, 2.3), annotation = c("***","***", "***"),
-              tip_length = 0, vjust = 0.6) +
-  #sGMV
-  geom_signif(y_position = c(2.7, 3.2, 2.2), xmin = c(2.7, 2.7, 3),
-              xmax = c(3, 3.3, 3.3), annotation = c("***","***", "***"),
-              tip_length = 0, vjust = 0.6) +
-  #Ventricles
-  geom_signif(y_position = c(7.1, 6.75, 6.45), xmin = c(3.7, 3.7, 4),
-              xmax = c(4, 4.3, 4.3), annotation = c("***","***", "***"),
-              tip_length = 0, vjust = 0.6) +
+  stat_pvalue_manual(stats_centile_for_plot.corrected, label = "p.val_fdr.signif", tip.length = 0) +
   labs(y="Absolulte Centile Error", x = "Feature") +
@@
 long_abs.diffs <- diffs.df %>%
-  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
   dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", pheno_list)) %>%
@@
   theme_linedraw(base_size = 11) +
-  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
-  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+scale_fill_manual(values=c(cf = "#450C54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values=c(cf = "#450C54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
   labs(y="Absolulte Centile Error", x = "Feature") +
@@
 check for skew (welch t.test vs ranked welch t test)
-```{r}
+```{r, eval=FALSE}
 for (pheno in pheno_list){
@@
 ```{r}
-dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+dataset_names <- c(raw = "Raw Data", cf = "ComBat \nw/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
 #var plot
 sex.var.plot.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat != "Global Volume") %>%
-  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "raw") %>%
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = median_sex_diff*100)
+         median_sex_diff = ifelse(sig_fdr == TRUE, median_sex_diff*100, NA))
 #find range
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.25, 0.25)) #
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-6, 6)) #
 ```
@@
 sex.var.plot.global.df <- sex_t_tests_to_plot %>%
-  dplyr::filter(pheno_cat == "Global Volume") %>%
-  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
+  dplyr::filter(pheno_cat == "Global Volume" & dataset != "raw") %>%
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
@@
                                             TRUE ~ NA)),
-        median_sex_diff = median_sex_diff*100) %>% 
+        median_sex_diff = ifelse(sig_fdr == TRUE, median_sex_diff*100, NA)) %>% 
   group_by(dataset)
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.25, 0.25), name = "Male - Female Median \nCentile Error") #,
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-6, 6), name = "Male - Female Median \nCentile Error") #,
 # sex.var.plot.global
@@
 
-p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-differences in centile score errors across combat configurations", 
-               color = "black", size = 14, hjust=.65))
-p_sexannotate
-ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
-saveRDS(p_sexannotate, file="figs/sex_diff_win_feat.RDS")
+ggsave("figs/sex_diff_win_feat.jpeg", p_sex, width=5.5, height=5)
+saveRDS(p_sex, file="figs/sex_diff_win_feat.RDS")
 ```
@@
   facet_wrap(~dataset, scales="free_y")
+```
+plot w/in feature t tests
+
+```{r}
+dataset_names <- c(raw = "Raw Data", cf = "ComBat \nw/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+#var plot
+sex.var.plot.df.noext <- noext.sex_t_tests_in_feat.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    label = sub("_[^_]*_", "_", pheno)) %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "raw") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100))
+#find range
+max(sex.var.plot.df.noext$median_sex_diff, na.rm=TRUE)
+min(sex.var.plot.df.noext$median_sex_diff, na.rm=TRUE)
+
+sex.var.plot.noext <- ggplot(sex.var.plot.df.noext) +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+             aes(fill = median_sex_diff)) +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-9, 9)) #
+```
+
+```{r}
+#plot global
+noext.sex.var.plot.global.df <- noext.sex_t_tests_in_feat.df %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+  dplyr::filter(pheno_cat == "Global Volume" & dataset != "raw") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100)) %>%
+  group_by(dataset)
+
+#color range
+max(noext.sex.var.plot.global.df$median_sex_diff, na.rm=TRUE)
+min(noext.sex.var.plot.global.df$median_sex_diff, na.rm=TRUE)
+
+noext.sex.var.plot.global <- ggplot(noext.sex.var.plot.global.df) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = median_sex_diff),
+             side = c("coronal")) +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-9, 9)) #,
+# sex.var.plot.global
+```
+
+Combine plots - need to keep playing around to make the spacing pretty when it saves out
+```{r}
+noext.p_sex <- ggarrange(
+  noext.sex.var.plot.global, sex.var.plot.noext,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.2),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+
+ggsave("figs/noext.sex_diff_win_feat.jpeg", noext.p_sex, width=5.5, height=5)
 ```
@@
 
+All features:
+```{r}
+long_abs.diffs.z <- diffs.df %>%
+  #dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
+  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", pheno_list.z)) %>%
+  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
+  mutate(pheno=gsub(".z", "", pheno)) %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
+
+long_abs_diff.plt.z <- long_abs.diffs.z %>%
+  ggplot(aes(pheno, abs.diff)) +
+  geom_violin(aes(x=pheno, y=abs.diff, fill=dataset, color=dataset), scale="count") +
+  theme_linedraw(base_size = 11) +
+  scale_fill_manual(values=c(cf = "#450C54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  scale_color_manual(values=c(cf = "#450C54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
+  labs(y="Absolulte Z-score Error", x = "Feature") +
+  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), legend.position="top") +
+  facet_wrap(~pheno_cat, nrow=4, scales="free")
+ggsave("figs/all_abs_err_plt_wlabels_z.jpeg", long_abs_diff.plt.z, width=13, height=10.5, units="in")
+```
+
+plot median on brain
+```{r}
+pheno_abs.z.list <- paste0("abs.diff_", pheno_list.z)
+
+#get median w/in feature 
+med.diffs.long.z <- diffs.df %>%
+  dplyr::select(sex, dataset, pheno_abs.z.list) %>%
+  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  group_by(dataset) %>%
+  summarize(across(starts_with("abs.diff_"), median)) %>%
+  ungroup() %>%
+  pivot_longer(cols=starts_with("abs.diff_"), values_to="median_z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
+  mutate(pheno = gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      label = sub("_[^_]*_", "_", pheno))
+
+med.diffs.long2.z <- left_join(med.diffs.long.z, dk.parc, by="label")
+max(med.diffs.long2.z$median_z_abs.diff)
+```
+
+```{r}
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+#var plot
+med.diffs.plot.df.z <- med.diffs.long2.z %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE))
+
+med.diffs.plot.z <- ggplot(med.diffs.plot.df.z) +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+             aes(fill = median_z_abs.diff)) +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Z-Score Error", limits = c(0, .05), breaks=c(0, 0.025, 0.05)) #
+# med.diffs.plot
+```
+
+```{r}
+#plot global
+med.diffs.plot.global.df.z <- med.diffs.long2.z %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE))
+
+med.diffs.plot.global.z <- ggplot(med.diffs.plot.global.df.z) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = median_z_abs.diff),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Z-Score Error", limits = c(0, .05), breaks=c(0, 0.025, 0.05))
+# med.diffs.plot.global
+```
+
+Combine plots - need to keep playing around to make the spacing pretty when it saves out
+```{r}
+p_diff.z <- ggarrange(
+  med.diffs.plot.global.z, med.diffs.plot.z,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.1),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+ggsave("figs/p_diff.z.jpeg", p_diff.z, width=7, height=5)
+```
+
 ### Sex-Differences
@@
 ```
+
+```{r}
+z.sex_t_tests_in_feat.df2 <- z.sex_t_tests_in_feat.df %>%
+  mutate(pheno=gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    label = sub("_[^_]*_", "_", pheno)) #for plotting
+z.sex_t_tests_to_plot <- left_join(z.sex_t_tests_in_feat.df2, dk.parc, by="label")
+
+max(z.sex_t_tests_to_plot$median_sex_diff)
+min(z.sex_t_tests_to_plot$median_sex_diff)
+```
+
+plot w/in feature t tests
+
+```{r}
+dataset_names <- c(raw = "Raw Data", cf = "ComBat \nw/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+#var plot
+z.sex.var.plot.df <- z.sex_t_tests_to_plot %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "raw") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = ifelse(sig_fdr == TRUE, median_sex_diff, NA))
+
+z.sex.var.plot <- ggplot(z.sex.var.plot.df) +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+             aes(fill = median_sex_diff)) +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.4, .4)) #
+```
+
+```{r}
+#plot global
+z.sex.var.plot.global.df <- z.sex_t_tests_to_plot %>%
+  dplyr::filter(pheno_cat == "Global Volume" & dataset != "raw") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "WM",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA)),
+        median_sex_diff = ifelse(sig_fdr == TRUE, median_sex_diff, NA)) %>% 
+  group_by(dataset)
+
+z.sex.var.plot.global <- ggplot(z.sex.var.plot.global.df) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = median_sex_diff),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.4, 0.4), name = "Male - Female Median \nCentile Error") #,
+# sex.var.plot.global
+```
+
+Combine plots - need to keep playing around to make the spacing pretty when it saves out
+```{r}
+z.p_sex <- ggarrange(
+  z.sex.var.plot.global, z.sex.var.plot,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.2),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+
+ggsave("figs/z_sex_diff_win_feat.jpeg", z.p_sex, width=5.5, height=5)
+```
+
+Sex Effect is the median male centile error - median female centile error (i.e.
+units are in centiles).
+Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)
 
@@
     as.data.frame()
+```
+
+plot w/in feature t tests
+
+```{r}
+dataset_names <- c(raw = "Raw Data", cf = "ComBat \nw/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+#var plot
+sex.var.plot.df.noext_z <- noext.z.sex_t_tests_in_feat.df %>%
+  mutate(pheno = gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_)),
+    label = sub("_[^_]*_", "_", pheno)) %>%
+  dplyr::filter(pheno_cat != "Global Volume" & dataset != "raw") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff))
+#find range
+max(sex.var.plot.df.noext_z$median_sex_diff, na.rm=TRUE)
+min(sex.var.plot.df.noext_z$median_sex_diff, na.rm=TRUE)
+
+sex.var.plot.noext_z <- ggplot(sex.var.plot.df.noext_z) +
+  theme_linedraw(base_size=12) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+             aes(fill = median_sex_diff)) +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits=c(-0.5, 0.5)) #
+```
+
+```{r}
+#plot global
+noext.sex.var.plot.global.df_z <- noext.z.sex_t_tests_in_feat.df %>%
+  mutate(pheno = gsub(".z", "", pheno)) %>%
+  mutate(pheno_cat = as.factor(case_when(
+    pheno %in% vol_list_global ~ "Global Volume",
+    pheno %in% vol_list_regions ~ "Regional Volume",
+    pheno %in% sa_list ~ "Regional SA",
+    pheno %in% ct_list ~ "Regional CT",
+    TRUE ~ NA_character_))) %>%
+  dplyr::filter(pheno_cat == "Global Volume" & dataset != "raw") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff)) %>%
+  group_by(dataset)
+
+#color range
+max(noext.sex.var.plot.global.df_z$median_sex_diff, na.rm=TRUE)
+min(noext.sex.var.plot.global.df_z$median_sex_diff, na.rm=TRUE)
+
+noext.sex.var.plot.global_z <- ggplot(noext.sex.var.plot.global.df_z) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = median_sex_diff),
+             side = c("coronal")) +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.5, 0.5)) #,
+# sex.var.plot.global
+```
+
+Combine plots - need to keep playing around to make the spacing pretty when it saves out
+```{r}
+noext.p_sex_z <- ggarrange(
+  noext.sex.var.plot.global_z, sex.var.plot.noext_z,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.2),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "bottom"
+  )
+
+ggsave("figs/z_noext.sex_diff_win_feat.jpeg", noext.p_sex_z, width=5.5, height=5)
 ```

### 2024-06-06 paper_prop_results.Rmd
lots of supplemental plots

@@
 ---
-title: "hbm_paper_results"
+title: ""
 output: html_document
@@
 z.pheno_abs.diff.list <- paste0(pheno_abs.diff.list, ".z")
+```
 
+```{r}
+aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+#write out new dk atlas w/o medial wall
+dk_nomed <- dk
+dk_nomed$data <- dk_nomed$data %>%
+  dplyr::filter(!is.na(region) & region !="corpus callosum")
 #plotting info from dk atlas
-dk.parc <- dk$data %>%
+dk.parc <- dk_nomed$data %>%
   as.data.frame() %>%
@@
 
-Read in predicted centiles & z scores FROM RAW DATA ONLY (to save memory)
-```{r}
-pred.csvs <- list.files(path = data_path, pattern = "raw_predictions.csv", full.names = TRUE)
-
-df_list <- list() #new empty list
-for (file in pred.csvs) {
-  # Read each CSV file
-  data <- fread(file)
-
-  # Add a "Source_File" column with the file name
-  data <- data %>%
-    mutate(Source_File = as.factor(basename(file))) %>%
-    mutate(dataset = gsub("_data|_predictions.csv|prop-|[0-9]|-", "", Source_File),
-           prop=gsub("-raw_predictions.csv", "", Source_File))
-
-    # Bind the data to the combined dataframe
-    df_list <- c(df_list, list(data))
-    assign("ratio_raw_pred_list", df_list)
-}
-length(ratio_raw_pred_list)
-
-#merge into single df
-ratio_raw_pred.df <- bind_rows(ratio_raw_pred_list)
-dim(ratio_raw_pred.df)
-```
-
 Visualize sex-distribution across sites/proportions
@@
 ```{r}
-cent.t.test.df %>%
+percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))
+comp_names <- c(cf.gam_cf = "ComBat-GAM v.
+\nComBat w/o Covar", cf.gamlss_cf = "ComBatLS v.
+ComBat w/o Covar", cf.gamlss_cf.gam = "ComBatLS v.
+ComBat-GAM", cf.lm_cf.gam ="ComBat-GAM v.
+linear ComBat", cf.lm_cf.gamlss = "ComBatLS v.
+linear ComBat", cf.lm_cf = "linear ComBat v.
+\nComBat w/o Covar")
+
+cent.t.test.plt <- cent.t.test.df %>%
   dplyr::filter(sig_fdr==TRUE) %>%
 ggplot() +
-  geom_bar(aes(x=prop, fill=bigger_group)) +
+  geom_bar(aes(x=prop, fill=smaller_group)) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~comp)
+  facet_grid(pheno_cat ~ comp, labeller = labeller(comp = as_labeller(comp_names)), scales="free_y") +
+  theme_linedraw() +
+  scale_fill_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "Better ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Feature Count", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+ggsave("figs/ratio_cent.t.test.plt.jpg", cent.t.test.plt, width=13, height=8, units="in")
 
@@
                                  prop == "10M:0F" ~ "prop-11",
-                                 TRUE ~ NA_character_))
-
-cent.t.test.df_global_filt <- cent.t.test_global %>%
-  dplyr::filter(comp %in% c("cf.gamlss_cf.gam", "cf.lm_cf.gam", "cf.lm_cf.gamlss")) %>%
-  dplyr::mutate(comp = factor(comp, ordered=TRUE, levels=c("cf.lm_cf.gam", "cf.lm_cf.gamlss", "cf.gamlss_cf.gam")))
+                                 TRUE ~ NA_character_), 
+                 comp = factor(comp, ordered=TRUE, levels=c("cf.gam_cf", "cf.gamlss_cf", "cf.gamlss_cf.gam", "cf.lm_cf", "cf.lm_cf.gam", "cf.lm_cf.gamlss")))
 ```
@@
   dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global)) %>%
-  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
+  #dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
   pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
@@
 
-#wilcox test for visualizing significance
+#wilcox test for visualizing significance - just going to use it for positioning info & setting up dataframe, will use actual tests run on CUBIC for signfificance
 mean_comp_tbl_calc <- prop.global_abs.diffs %>%
@@
 mean_comp_tbl <- mean_comp_tbl_calc %>%
-  mutate(comp = factor(paste(group2, group1, sep="_"), ordered=TRUE, levels=c("cf.lm_cf.gam", "cf.lm_cf.gamlss", "cf.gamlss_cf.gam"))) %>%
+  mutate(comp = factor(paste(group2, group1, sep="_"), ordered=TRUE, levels=c("cf.gam_cf", "cf.gamlss_cf", "cf.gamlss_cf.gam", "cf.lm_cf", "cf.lm_cf.gam", "cf.lm_cf.gamlss"))) %>%
   arrange(prop, comp) %>%
-  add_xy_position(x="prop", fun="max", comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss"))) %>%
-  add_significance("p.adj")
+  add_xy_position(x="prop", fun="max", comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss"), c("cf.lm", "cf"), c("cf", "cf.gam"), c("cf", "cf.gamlss")))
+
+mean_comp_tbl2 <- inner_join(mean_comp_tbl, cent.t.test_global) %>%
+  add_significance("p.val_fdr", cutpoints = c(0, 0.001, 0.01, 0.05, 1),
+  symbols = c("***", "**", "*", "ns"))
 
@@
 
-mean_comp_tbl2 <- full_join(mean_comp_tbl, max_abs.errs)
+stats_centile_for_plot <- full_join(mean_comp_tbl2, max_abs.errs)
 
-#just use significance from actual stats tests
-
-stats_centile_for_plot <- inner_join(mean_comp_tbl2, cent.t.test.df_global_filt)
 #kinda hacky way of fixing positions
 stats_centile_for_plot.corrected <- stats_centile_for_plot %>%
-  mutate(p.adj.signif = ifelse(sig_fdr == "FALSE", "ns", "*"),
-         xmax=case_when(group1=="cf.gamlss" | group2== "cf.gamlss" ~ (x+.266667),
+  mutate(xmax=case_when(group1=="cf.gamlss" | group2== "cf.gamlss" ~ (x+.266667),
                         TRUE ~  x),
@@
 
-cent.t.test.df_global_filt %>%
-  dplyr::select(prop, comp, pheno, sig_fdr) %>%
+stats_centile_for_plot.corrected %>%
+  dplyr::select(prop, comp, pheno, sig_fdr, p.val_fdr.signif) %>%
   dplyr::filter(sig_fdr == FALSE)
@@
 ```{r}
-percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))
-
 prop.global_cent_errs_plot <- prop.global_abs.diffs %>%
@@
   scale_x_discrete(labels=percent_f) +
-  stat_pvalue_manual(stats_centile_for_plot.corrected, label = "p.adj.signif", label.size = 2, tip.length = 0) +
+  stat_pvalue_manual(stats_centile_for_plot.corrected, label = "p.val_fdr.signif", label.size = 2, tip.length = 0) +
   # ggsignif::geom_signif(data=mean_comp_tbl2) +
-  facet_wrap(~ pheno)
+  facet_wrap(~pheno)
   # ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
@@
 cent.feat.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_cent_sex_bias_tests.RDS"))
-cent.feat.sex.test.df <- bind_rows(cent.feat.sex.test.list)
+cent.feat.sex.test.df <- bind_rows(cent.feat.sex.test.list) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)))
 ```
 
-simple bar-plot to count regions w sigificant sex-differences in errors
+simple bar-plot to count regions w significant sex-differences in errors
 
@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+sig.sex_diff.violin <- cent.feat.sex.test.df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  mutate(median_sex_diff = median_sex_diff*100) %>%
+  ggplot() +
+  geom_violin(aes(x=prop, y=median_sex_diff, fill=dataset, color=dataset)) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Centile Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+
+ggsave("figs/sig.sex_diff.violin.jpeg", sig.sex_diff.violin, width=7, height=5, units="in")
 ```
@@
 ```{r}
-cent.feat.test_global <- cent.feat.sex.test.df %>%
-  dplyr::filter(pheno_cat == "Global Volume" & (dataset == "cf.gam" | dataset =="cf.gamlss"))
-
 #wilcox test for visualizing significance
@@
   wilcox_test(diff ~ sex) 
-
-cent.feat.test_global %>%
-  dplyr::filter(pheno=="WMV") %>%
-  dplyr::select(dataset, pheno, prop, sig_fdr)
 ```
@@
 #just use significance from actual stats tests
+cent.feat.test_global <- cent.feat.sex.test.df %>%
+  dplyr::filter(pheno_cat == "Global Volume" & (dataset == "cf.gam" | dataset =="cf.gamlss"))
 
@@
 t.test.cent.sex.diff <- cent.feat.sex.test.df %>%
-  dplyr::filter(pheno_cat != "Global Volume") %>%
-  group_by(dataset, pheno_cat, prop) %>%
-  summarise(t.test.p = tidy(t.test(median_sex_diff, mu=0))$p.value,
+  dplyr::filter(sig_fdr == TRUE) %>%
+  group_by(dataset, prop) %>%
+  summarise(w.test.p = tidy(wilcox.test(median_sex_diff, mu=0))$p.value,
             max_val = max(median_sex_diff)*100) %>% #test median of each distribution against 0, then scale to centiles
-  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = (3*4*11)), #not testing global vols, too few data points
-                  sig_fdr_tf = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ ""),
-         y_pos = max_val + 0.1)
+  mutate(p.val_fdr = p.adjust(w.test.p, method="fdr", n = (4*11)),
+         y_pos = max_val + 0.5) %>%
+  add_significance("p.val_fdr", cutpoints = c(0, 0.001, 0.01, 0.05, 1),
+  symbols = c("***", "**", "*", "ns"))
 
@@
   geom_boxplot(aes(color=dataset)) +
-    facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_y") + 
-  geom_text(aes(label = sig_fdr, y = y_pos),
+    facet_wrap(~dataset, labeller=labeller(dataset=dataset_names)) + 
+  geom_text(aes(label = p.val_fdr.signif, y = y_pos),
               data = t.test.cent.sex.diff[which(t.test.cent.sex.diff$dataset != "cf"),]) +
@@
 cortical_cent_sex_diff.plt
+
+cent_sex_diff.plt2 <- cent.feat.sex.test.df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  mutate(median_sex_diff = median_sex_diff*100) %>%
+  ggplot(aes(x=prop, y=median_sex_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+  geom_jitter(size=0.4, alpha=0.1) +
+  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names)) + 
+  geom_text(aes(label = p.val_fdr.signif, y = y_pos),
+              data = t.test.cent.sex.diff) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Centile Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+cent_sex_diff.plt2
+
+ggsave("figs/cent_sex_diff.plt2.jpeg", cent_sex_diff.plt2)
 ```
-
-COME BACK TO THIS AFTER LOOKING AT REPLICATION PLOTS
 
@@
   facet_wrap(~comp)
+
+z.feat.t.test.plt <- z.feat.t.test.df %>%
+  dplyr::filter(sig_fdr==TRUE) %>%
+ggplot() +
+  geom_bar(aes(x=prop, fill=smaller_group)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_grid(pheno_cat ~ comp, labeller = labeller(comp = as_labeller(comp_names)), scales="free_y") +
+  theme_linedraw() +
+  scale_fill_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "Better ComBat \nConfiguration", labels=dataset_names) +
+  theme(legend.position = "bottom", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Feature Count", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+ggsave("figs/ratio_z.t.test.plt.jpg", z.feat.t.test.plt, width=13, height=8, units="in")
 ```
@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+z.sig.sex_diff.violin <- z.feat.sex.test.df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  ggplot() +
+  geom_violin(aes(x=prop, y=median_sex_diff, fill=dataset, color=dataset)) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Z-Score Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+
+ggsave("figs/sig.sex_diff.violin.z.jpeg", z.sig.sex_diff.violin, width=7, height=5, units="in")
 ```
@@
   scale_x_discrete(labels=percent_f)
+```
+
+```{r}
+z.t.test.cent.sex.diff <- z.feat.sex.test.df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  group_by(dataset, prop) %>%
+  summarise(w.test.p = tidy(wilcox.test(median_sex_diff, mu=0))$p.value,
+            max_val = max(median_sex_diff)) %>% #test median of each distribution against 0
+  mutate(p.val_fdr = p.adjust(w.test.p, method="fdr", n = (4*11)), 
+         y_pos = max_val + 3) %>%
+  add_significance("p.val_fdr", cutpoints = c(0, 0.001, 0.01, 0.05, 1),
+  symbols = c("***", "**", "*", "ns"))
+
+z.cent_sex_diff.plt <- z.feat.sex.test.df %>%
+  dplyr::filter(sig_fdr == TRUE) %>%
+  mutate(median_sex_diff = median_sex_diff*100) %>%
+  ggplot(aes(x=prop, y=median_sex_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+  geom_jitter(size=0.4, alpha=0.1) +
+  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names)) + 
+  geom_text(aes(label = p.val_fdr.signif, y = y_pos),
+              data = z.t.test.cent.sex.diff) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Z-Score Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
+z.cent_sex_diff.plt
+
+ggsave("figs/z_sex_diff.plt.jpeg", z.cent_sex_diff.plt)
 ```
@@
 ```{r}
-fig_4 <- ggarrange(ggarrange(sampling.plt, sex_err_ratio.plt, 
+fig_4.alt <- ggarrange(ggarrange(sampling.plt, sex_err_ratio.plt, 
                              nrow = 2, labels = c("A", "C"),
@@
 
-fig_4
-ggsave("figs/figure_4.jpeg", fig_4, width=33, height=20, units="cm")
+fig_4 <- ggarrange(sampling.plt, prop.global_cent_errs_plot, 
+                             nrow = 2, labels = c("A", "B"),
+                             heights=c(1,3))
+
+ggsave("figs/figure_4.jpeg", fig_4)
 ```

### 2024-06-06 paper_replication_results.Rmd
lots of supplemental plots

@@
               mu=mu) %>%
-    mutate(p.val_fdr = p.adjust(p_val, method="fdr", n = (4*4*3)), #adjusting for 4 methods * 4 pheno classes * 3 centile categories
+    mutate(p.val_fdr = p.adjust(p_val, method="fdr", n = (4*4*2)), #adjusting for 4 methods * 4 pheno classes * 2 centile categories
            sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",

### 2024-06-06 paper_lbcc_results.Rmd
restore cohens f, update plots

--- 2024-05-25 paper_lbcc_results.Rmd
@@
 
-Combat-GAM Centiles - study-level batch correction w/o infants
+Combat-GAM Centiles - study-level batch correction
 
@@
 
-Nature paper models to look at age effects in variance
-from [https://github.com/brainchart/Lifespan/tree/main/Share](https://github.com/brainchart/Lifespan/tree/main/Share)
-
-```{r}
-paper_fit_sgmv <- readRDS("derivatives/nature_2020_mods/FIT_sGMV.rds")
-paper_fit_gmv <- readRDS("derivatives/nature_2020_mods/FIT_GMV.rds")
-paper_fit_wmv <- readRDS("derivatives/nature_2020_mods/FIT_WMV.rds")
-paper_fit_ven <- readRDS("derivatives/nature_2020_mods/FIT_Ventricles.rds")
-```
-
 #Clean & EDA
@@
   summarize(mean_age = mean(age_years)) %>%
-  arrange(-mean_age)
+  arrange(mean_age)
 
@@
 
-w/in feature ANOVAs
-
 ```{r}
 raw.feat_to_test <- intersect(pheno_list, names(raw_centile_df.clean))
+
+#see what features didnt converge
 length(raw.feat_to_test)
+setdiff(pheno_list, raw.feat_to_test)
+```
 
-df_raw_feat_aov <- data.frame()
-for(pheno in raw.feat_to_test){
-  # print(pheno)
-  df <- tidy(aov(raw_centile_df.clean[[pheno]] ~ raw_centile_df.clean[["study"]]))
-  df$pheno <- pheno
-  df_raw_feat_aov <- rbind(df_raw_feat_aov, df)
-}
+w/in subject means
 
-df_raw_feat_aov.study <- df_raw_feat_aov %>%
-  group_by(pheno) %>%
-  mutate(etasq = sumsq / sum(sumsq)) %>%
-  ungroup() %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-df_raw_feat_aov.study %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-table(df_raw_feat_aov.study$sig_fdr)
-```
 ```{r}
@@
 
-ANOVA of subjects' mean centiles across studies
+ANOVA
 
@@
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
-```
-
-w/in feature ANOVAs
-
-```{r}
-raw.feat_to_test.z <- paste0(raw.feat_to_test, ".z")
-df_raw_feat_aov.z <- data.frame()
-for(pheno in raw.feat_to_test.z){
-  # print(pheno)
-  df <- tidy(aov(raw_z_df.clean[[pheno]] ~ raw_z_df.clean[["study"]]))
-  df$pheno <- pheno
-  df_raw_feat_aov.z <- rbind(df_raw_feat_aov.z, df)
-}
-
-df_raw_feat_aov.z.study <- df_raw_feat_aov.z %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test.z)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-df_raw_feat_aov.z.study %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-table(df_raw_feat_aov.z.study$sig_fdr)
 ```
@@
 
-ANOVA of subject means
+```{r}
+raw_z_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
+```
 
+ANOVA
 ```{r}
@@
 
-w/in feature ANOVAs
-
 ```{r}
 gam.feat_to_test <- intersect(pheno_list, names(gam_centile_df.clean))
+
+#see what features didnt converge
 length(gam.feat_to_test)
-
-df_gam_feat_aov <- data.frame()
-for(pheno in gam.feat_to_test){
-  # print(pheno)
-  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["study"]]))
-  df$pheno <- pheno
-  df_gam_feat_aov <- rbind(df_gam_feat_aov, df)
-}
-
-df_gam_feat_aov.study <- df_gam_feat_aov %>%
-  group_by(pheno) %>%
-  mutate(etasq = sumsq / sum(sumsq)) %>%
-  ungroup() %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-df_gam_feat_aov.study %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-table(df_gam_feat_aov.study$sig_fdr)
+setdiff(pheno_list, gam.feat_to_test)
 ```
@@
 
-w/in feature ANOVAs
-
-```{r}
-gam.feat_to_test.z <- paste0(gam.feat_to_test, ".z")
-df_gam_feat_aov.z <- data.frame()
-for(pheno in gam.feat_to_test.z){
-  # print(pheno)
-  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["study"]]))
-  df$pheno <- pheno
-  df_gam_feat_aov.z <- rbind(df_gam_feat_aov.z, df)
-}
-
-df_gam_feat_aov.z.study <- df_gam_feat_aov.z %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test.z)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-df_gam_feat_aov.z.study %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-table(df_gam_feat_aov.z.study$sig_fdr)
-```
-
 average w/in subject
@@
 
-Subject mean ANOVA
+ANOVA
 
@@
 
-w/in feature ANOVAs
-
 ```{r}
@@
 
-df_gamlss_feat_aov <- data.frame()
-for(pheno in gamlss.feat_to_test){
-  # print(pheno)
-  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
-  df$pheno <- pheno
-  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
-}
+#see what features didnt converge
+length(gamlss.feat_to_test)
+setdiff(pheno_list, gamlss.feat_to_test)
+```
 
-df_gamlss_feat_aov.study <- df_gamlss_feat_aov %>%
-  group_by(pheno) %>%
-  mutate(etasq = sumsq / sum(sumsq)) %>%
-  ungroup() %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-df_gamlss_feat_aov.study %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-table(df_gamlss_feat_aov.study$sig_fdr)
+```{r}
+ggplot(gamlss_centile_df.clean) +
+  geom_point(aes(x=age_days, y=rh_Vol_insula, color=study), alpha=0.5) +
+  theme(legend.position="none")
 ```
@@
   theme(axis.text.x = element_text(angle=90, hjust=0.5))
+
+gamlss_centile_df.mean %>%
+  ggplot() +
+  geom_point(aes(x=age_days/365.25, y=mean_cent, color=study), alpha=0.5) +
+  geom_smooth(aes(x=age_days/365.25, y=mean_cent, color=study), method="lm") +
+  theme(legend.position = "none")
+
+gamlss_centile_df.mean %>%
+  dplyr::filter(age_days>365) %>%
+  ggplot(aes(x=age_days/365.25, y=mean_cent)) +
+  geom_point(aes(color=study), alpha=0.7) +
+  geom_smooth(method="lm") +
+  theme(legend.position="none")
 ```
 
-subject mean ANOVA
-
+ANOVA
 ```{r}
@@
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
-```
-
-w/in feature ANOVAs
-
-```{r}
-gamlss.feat_to_test.z <- paste0(gamlss.feat_to_test, ".z")
-df_gamlss_feat_aov.z <- data.frame()
-for(pheno in gamlss.feat_to_test.z){
-  # print(pheno)
-  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
-  df$pheno <- pheno
-  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
-}
-
-df_gamlss_feat_aov.z.study <- df_gamlss_feat_aov.z %>%
-  dplyr::filter(term != "Residuals") %>%
-  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test.z)), #apply fdr correction across all phenos tested
-                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE))
-df_gamlss_feat_aov.z.study %>%
-  ggplot() +
-  geom_bar(aes(y=sig_fdr))
-
-table(df_gamlss_feat_aov.z.study$sig_fdr)
 ```
@@
 
-subject mean ANOVA
+ANOVA
 
@@
 
+Centile boxplots
 ```{r}
@@
       TRUE ~ NA_character_)))
+```
 
-fsq.df <- ggplot(f2.df) +
+pretty plot
+
+```{r}
+fsq_plt <- ggplot(f2.df) +
   geom_col(aes(x=pheno, y=fsq, fill=combat), position="dodge") + 
@@
   theme_linedraw() +
-  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
-  labs(x = "Feature", y = "Effect Size (Cohen's F-Squared)", fill="ComBat")
-
-ggsave("figs/supp_fsq.jpeg", fsq.df, width=8, height=6, units="in")
+  theme(axis.text.x = element_blank(), legend.position="bottom", axis.ticks.x = element_blank()) + 
+  scale_fill_manual(values = c('ComBat-GAM' = "#3D75C2", ComBatLS ="#CFB023FF"), name = "ComBat Configuration") +
+  labs(y="Effect Size (Cohen's F-Squared)", x = "Brain Feature")
+ggsave("figs/cohens_fsq.jpeg", fsq_plt)
 ```
@@
   geom_col(aes(x=pheno_cat, y=mean_fsq_diff), position="dodge")
+
+f2.df.sum %>%
+  #group_by(pheno_cat) %>%
+  mutate(smaller_eff = case_when(F2_diff < 0 ~ "ComBatGAM",
+                                    F2_diff > 0 ~ "ComBatLS",
+                                 F2_diff == 0 ~ "Same")) %>%
+  ggplot() +
+  geom_bar(aes(x=smaller_eff, fill=pheno_cat), position="dodge")
 ```
-
 #ComBat-GAM vs ComBatLS: Delta Centile
@@
             study_median_age=median(age_years),
-            study_n = n()) %>%
+            study_n= n()) %>%
   dplyr::select(study, study_mean_age, study_median_age, study_n)
@@
 ```
-
-control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.
 
@@
   theme(legend.position="none")
-```
 
-See if a subject's distance from their study's mean age is assoc with how much their centile changes
-
-```{r}
 mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
@@
          abs_age_diff = abs(age_diff))
+```
 
+mean abs delta ~ age_diff
+
+```{r}
 abs_lm2 <- lm(abs_mean_delta_cent ~ study_n + abs_age_diff, data=mean_cent_deltas_df_study)
@@
 
+control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.
+
 ## Subject-Level Mean Centiles ~ Mean Study Age
 
-Make sure that results of mean centile delta ~ mean study age are related to the difference itself, not because ComBatLS and/or ComBat-GAM centiles are associated with mean study age.
-
-```{r}
-raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
-summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
-
-raw_centile_df.mean2 %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-NS
+Make sure that delta centile results are not driven by either ComBatLS or ComBat-GAM centiles being assoc. with study age
 
@@
 
-NS
+Plots
+
+```{r}
+gamlss_centile_df.mean2$combat <- "gamlss"
+gam_centile_df.mean2$combat <- "gam"
+raw_centile_df.mean2$combat <- "raw"
+
+#put all data frames into list
+df_list.tmp <- list(gamlss_centile_df.mean2, gam_centile_df.mean2, raw_centile_df.mean2)
+
+#merge all data frames in list
+all_subj_means.df <- df_list.tmp %>% reduce(full_join)
+
+all_subj_means.df %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.6) +
+  geom_smooth(method="lm")
+
+all_subj_means.df %>%
+  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.3) +
+  geom_smooth(method="lm") +
+  facet_wrap(~combat)
+
+summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
+summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
+```
+
+```{r}
+summary(lm(mean_cent ~ study_n, data=raw_centile_df.mean2))
+summary(lm(mean_cent ~ study_n, data=gam_centile_df.mean2))
+summary(lm(mean_cent ~ study_n, data=gamlss_centile_df.mean2))
+
+all_subj_means.df %>%
+  ggplot(aes(x=study_n, y=mean_cent, color=combat)) +
+  geom_point(alpha=0.6) +
+  geom_smooth(method="lm")
+```
+all assoc with study n
 
 #FIGURE 5
-
-Reformatting plots for prettier composite fig
 
@@
     geom_line(data = pred_df, aes(y = fit, x=study_mean_age))+
-  theme_linedraw() +
+  theme_linedraw()+
   theme(legend.position="none") +
-  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From ComBatLS and ComBat-GAM \n-harmonized Data")
-ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=9, height=6, units="in")
+  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)")
+ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=10, height=7, units="in")
 ```
 
-lines show marginal effects of mean study age
+```{r}
+mean_cent_deltas_df_study %>%
+  dplyr::filter(abs_mean_delta_cent > 4) %>%
+  dplyr::select(study, study_mean_age)
+
+mean_cent_deltas_df_study%>%
+  dplyr::filter(study == "Cornell_C2") %>%
+  nrow()
+```
+
+Cornell_C2 has very low n=5 -> big differences in estimation
 
@@
     geom_line(data = pdat2, aes(y = fit, x=abs_age_diff)) +
-  theme_linedraw() +
+  theme_linedraw()+
   theme(legend.position="none") +
-  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From ComBatLS and ComBat-GAM \n-harmonized Data")
-ggsave("figs/supp_mean_delta_fig2.jpeg", mean_delta_fig2, width=9, height=6, units="in")
+  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)")
+ggsave("figs/supp_mean_delta_fig2.jpeg", mean_delta_fig2, width=10, height=7, units="in")
 ```

### 2024-06-07 paper_main_results.Rmd
cleaning up

@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.4, .4)) #
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-Score Error", limits = c(-0.4, .4)) #
 ```
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.4, 0.4), name = "Male - Female Median \nCentile Error") #,
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.4, 0.4), name = "Male - Female Median \nZ-score Error") #,
 # sex.var.plot.global
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits=c(-0.5, 0.5)) #
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-score Error", limits=c(-0.5, 0.5)) #
 ```
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.5, 0.5)) #,
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-score Error", limits = c(-0.5, 0.5)) #,
 # sex.var.plot.global
@@
 
+Redo with combat w/o covariates
+
+```{r}
+dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
+#var plot
+var.plot.df2 <- sigma.sex.df2 %>%
+  dplyr::filter(pheno_cat != "Global Volume") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
+         sex_standard = ifelse(drop1.pval <0.05, sex_standard, NA)) %>%
+  group_by(dataset, pheno_cat)
+
+range(var.plot.df2$sex_standard, na.rm=T)
+
+tp <- unique(var.plot.df2[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="raw")
+tp$sex_standard <- 1
+
+var.plot2 <- ggplot(var.plot.df2) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = dk_nomed, 
+             position = position_brain(hemi ~ side),
+             aes(fill = sex_standard)) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
+```
+
+```{r}
+#plot global
+var.plot.global.df2 <- sigma.sex.df2 %>%
+  dplyr::filter(pheno_cat == "Global Volume") %>%
+  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
+  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
+  dplyr::mutate(tissue_class = case_when(pheno == "GMV" ~ "cGM",
+                                            pheno == "Ventricles" ~ "CSF",
+                                            pheno == "WMV" ~ "WM",
+                                            pheno == "sGMV" ~ "sGM",
+                                            TRUE ~ NA_character_),
+         sex_standard = ifelse(drop1.pval <0.05, sex_standard, NA)) %>%
+  group_by(dataset)
+
+tp <- unique(var.plot.global.df2[,c('dataset','pheno_cat')]) %>%
+  dplyr::filter(dataset=="raw")
+tp$sex_standard <- 1
+
+range(var.plot.global.df2$sex_standard, na.rm=T)
+
+var.plot.global2 <- ggplot(var.plot.global.df2) +
+  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
+            ymin = -Inf, ymax = Inf, alpha = 0.4) +
+  ggseg::geom_brain(atlas = aseg_cerebrum, 
+             aes(fill = sex_standard),
+             side = c("coronal")) +
+  #scale_fill_gradient2() +
+  facet_grid(dataset~pheno_cat) + 
+  theme_linedraw(base_size=12) +
+  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
+```
+
+Combine plots
+```{r}
+p <- ggarrange(
+  var.plot.global2, var.plot2,
+  ncol=2, nrow=1, 
+  widths = c(1, 3.18),
+  #labels=c("A", "B"),
+  #hjust=0.01,
+  align = "hv",
+  common.legend = TRUE, legend = "right"
+  )
+ggsave("figs/full_sex_eff_on_scale.jpeg", p, width=8, height=8, units="in")
+```
+
+##Test Sex-Effects Across Feature Types
+
+```{r}
+raw_sex_eff <- sigma.sex.df2 %>%
+  dplyr::filter(dataset == "raw")
+
+summary(aov(sex_standard ~ pheno_cat, data=raw_sex_eff))
+pairwise.wilcox.test(raw_sex_eff$sex_standard, raw_sex_eff$pheno_cat, p.adjust.method = "fdr")
+```
+
+plot
+
+```{r}
+comp <- get_comparisons(raw_sex_eff, pheno_cat, ref.group = NULL)
+
+? <- ggboxplot(raw_sex_eff, x = "pheno_cat", y = "sex_standard",
+          color = "pheno_cat")+ 
+  stat_compare_means(method = "wilcox.test", comparisons = comp, label="p.signif")+ # Add pairwise comparisons p-value
+  stat_compare_means(label.y = 0.25) +     # Add global p-value
+  theme(legend.position = "none") +
+  labs(x="Brain Feature Category", y="Standardized Effect of Sex on Variance")
+
+ggsave("figs/sex_eff_comp_plt.jpeg", sex_eff_comp_plt)
+```
+
 # FIGURE 1: METHODS

### 2024-06-07 paper_replication_results.Rmd
cleaning up

@@
 
-all_cent_sum_boxplt_wtest <- all_cent_sum.df %>%
+all_cent_sum_boxplt.diff_full <- all_cent_sum.df_scaled %>%
   dplyr::filter(dataset != "raw") %>%
-  mutate(source = factor(source, levels=c("low", "high"), ordered=TRUE)) %>%
-  ggplot(aes(x=dataset, y=female_ratio)) +
+  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
+  ggplot(aes(x=dataset, y=female_ratio_diff)) +
   geom_boxplot(aes(color=dataset)) +
@@
   facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
-  geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
-  geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  geom_hline(yintercept=0, linetype="dashed")+
+  geom_text(aes(label = sig_fdr, y = (y_pos)), nudge_y=.005,
                data = all_t_test.df) +
@@
   scale_x_discrete(labels=dataset_names) +
+  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
   theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  labs(y="Proportion Female", x = "Dataset")
-#ggsave(filename="figs/all_cent_sum_boxplt_wtest.png", all_cent_sum_boxplt_wtest)
+  labs(y="Female Bias", x = "Dataset")
+ggsave(filename="figs/all_cent_sum_boxplt_wtest.png", all_cent_sum_boxplt.diff_full)
 ```

### 2024-06-10 paper_lbcc_results.Rmd
cleaning up
cohens fsq for unharmonized brain charts

@@
 
-Cohen's F squared for study effects
+##Cohen's F-squared for study effects
 
 ```{r}
+raw_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_cohenfsq.csv"))
+raw_centile_f2$combat <- "Unharmonized"
+
 gam_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv"))
@@
 
-f2.df <- rbind(gam_centile_f2, gamlss_centile_f2)
+f2.df <- rbind(raw_centile_f2, gam_centile_f2, gamlss_centile_f2)
 f2.df <- f2.df %>%
@@
       TRUE ~ NA_character_)))
+
+unique(f2.df$combat)
+
+f2.df %>%
+  group_by(combat) %>%
+  summarize(median = median(fsq),
+            IQR = IQR(fsq))
 ```
@@
   theme(axis.text.x = element_blank(), legend.position="bottom", axis.ticks.x = element_blank()) + 
-  scale_fill_manual(values = c('ComBat-GAM' = "#3D75C2", ComBatLS ="#CFB023FF"), name = "ComBat Configuration") +
+  scale_fill_manual(values = c('ComBat-GAM' = "#3D75C2", ComBatLS ="#CFB023FF", Unharmonized = "darkgray"), name = "Harmonization") +
   labs(y="Effect Size (Cohen's F-Squared)", x = "Brain Feature")
@@
 f2.df.sum <- f2.df %>%
+  dplyr::filter(combat != "Umharmonized") %>%
   group_by(pheno) %>%

### 2024-06-14 paper_lbcc_results.Rmd
changes in extreme centile categorization between combat-gam and combatls

@@
 mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)
+
+summary(mean_cent_deltas_df_study$abs_mean_delta_cent)
+```
+
+see how often features move between extreme (>95% or <5%) to not-extreme
+```{r}
+#rm features that didn't converge in both datasets
+pheno_list_cent <- pheno_list[pheno_list != "lh_SA_posteriorcingulate" & pheno_list != "rh_CT_posteriorcingulate"]
+
+gam_centile_df.clean$combat <- "cf.gam"
+gamlss_centile_df.clean$combat <- "cf.gamlss"
+
+centiles_df <- full_join(gam_centile_df.clean, gamlss_centile_df.clean)
+
+centiles_cat <- centiles_df %>%
+  mutate(combat = factor(combat, ordered=TRUE, levels=c("cf.gam", "cf.gamlss"))) %>%
+  mutate(across(pheno_list_cent, ~ as.factor(case_when(. > 0.95 ~ "high", 
+                                               . < 0.05 ~ "low",
+                                               TRUE ~ "mid")))) %>%
+  arrange(combat) %>%
+  group_by(participant) %>%
+  mutate(across(pheno_list_cent, 
+                ~ . == first(.), .names = "is_equal_{col}")) %>%
+  ungroup()
+```
+
+```{r}
+centiles_cat_filt <- centiles_cat %>%
+  dplyr::select(participant, combat, all_of(pheno_list_cent))
+
+ext_count_df <- centiles_cat_filt %>%
+  rowwise() %>%
+  mutate(count_high = sum(c_across(pheno_list_cent) == "high"),
+         count_mid = sum(c_across(pheno_list_cent) == "mid"),
+         count_low = sum(c_across(pheno_list_cent) == "low")) %>%
+  dplyr::select(participant, combat, starts_with("count"))
+
+ext_count_df %>%
+  ggplot() +
+  geom_bar(aes(x=count_high, fill=combat, color=combat), position="identity", alpha=.4)
+ext_count_df %>%
+  ggplot() +
+  geom_bar(aes(x=count_low, fill=combat, color=combat), position="identity", alpha=.4)
+```
+
+```{r}
+centiles_cat_filt2 <- centiles_cat %>%
+  dplyr::filter(combat =="cf.gamlss") %>%
+  dplyr::select(participant, starts_with("is_equal_")) %>%
+  mutate(sum = rowSums(across(starts_with("is_equal_")))) %>%
+  mutate(changed = ifelse(sum == length(pheno_list_cent), "no_change", "change")) #if not all features are true -> changed
+
+max(centiles_cat_filt2$sum) == length(pheno_list_cent) #correct
+
+length(pheno_list_cent) - mean(centiles_cat_filt2$sum) #mean per subject
+length(pheno_list_cent) - min(centiles_cat_filt2$sum) #range
+length(pheno_list_cent) - max(centiles_cat_filt2$sum)
+
+#percentage subjects
+table(centiles_cat_filt2$changed)["change"]/nrow(centiles_cat_filt2)
+
+#percentage features
+#n changed features = # total features - sum
+n_feat <- n_unique(centiles_cat_filt2$participant)*length(pheno_list_cent)
+n_unchanged <- sum(centiles_cat_filt2$sum) #true = 1 = no change
+
+n_feat-n_unchanged
+((n_feat-n_unchanged)/n_feat)*100 # % changed
+```
+
+plot
+```{r}
+centiles_cat_filt2_long <- centiles_cat_filt2 %>%
+  pivot_longer(cols=starts_with("is_equal_"), names_to="pheno", names_prefix="is_equal_", values_to="is_equal")
+
+n_unique(centiles_cat_filt2_long$participant)*206 == nrow(centiles_cat_filt2_long)
+
+cent_cat_plt <- centiles_cat_filt2_long %>%
+  dplyr::filter(is_equal==FALSE) %>%
+ggplot() +
+  geom_bar(aes(x=pheno), fill="#1D213B") +
+  theme_linedraw()+
+  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) +
+  labs(x="Brain Feature", y= "Instances of Centile Re-categorization")
+ggsave("figs/cent_cat_plt.jpeg", cent_cat_plt, width=10, height=6, units="in")
 ```
@@
 Plots
+
+```{r}
+raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info) %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age)
+```
 

### 2024-06-14 paper_main_results.Rmd
simulating distributions for visualizations

@@
 ```
+
+# Other Figs
+Plotting normal curves for pres
+
+```{r}
+data <- data.frame(A = rnorm(1000000, mean = 7, sd = 1),
+                   B = rnorm(1000000, mean = 10, sd = 2),
+                   C = rnorm(1000000, mean = 13, sd = 3),
+                   datatype=rep("raw", n=1000000))
+
+ggplot(data, aes(x = A, size=1.4)) +
+  geom_density(fill = "gray", alpha = 0.5) +
+  geom_vline(xintercept = mean(data$A), size=1) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+
+ggplot(data, aes(x = B, size=1.4)) +
+  geom_density(fill = "gray", alpha = 0.5) +
+  geom_vline(xintercept = mean(data$B), size=1) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+
+ggplot(data, aes(x = C, size=1.4)) +
+  geom_density(fill = "gray", alpha = 0.5) +
+  geom_vline(xintercept = mean(data$C), size=1) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+```
+
+```{r}
+combatgam_data <- data.frame(A = rnorm(1000000, mean = 9.5, sd = 2),
+                   B = rnorm(1000000, mean = 10, sd = 2),
+                   C = rnorm(1000000, mean = 11, sd = 2),
+                   datatype=rep("cf.gam", n=1000000))
+
+combatgam_data <- full_join(combatgam_data, data) %>%
+  mutate(datatype=factor(datatype, levels=c("raw", "cf.gam"), ordered=T))
+```
+
+```{r}
+combatgam_data %>%
+ggplot(aes(size=1.4, fill=datatype, color=datatype)) +
+  geom_density(aes(x=A), alpha = 0.5) +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="raw",]$A), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="cf.gam",]$A), size=1, color="#3D75C2") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+
+combatgam_data %>%
+ggplot(aes(size=1.4, fill=datatype, color=datatype)) +
+  geom_density(aes(x=B), alpha = 0.5) +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="raw",]$B), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="cf.gam",]$B), size=1, color="#3D75C2") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+
+combatgam_data %>%
+ggplot(aes(size=1.4, fill=datatype, color=datatype)) +
+  geom_density(aes(x=C), alpha = 0.5) +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="raw",]$C), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="cf.gam",]$C), size=1, color="#3D75C2") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+```
+
+```{r}
+combatgamlss_data <- data.frame(A = rnorm(1000000, mean = 9.5, sd = 1.5),
+                   B = rnorm(1000000, mean = 10, sd = 2),
+                   C = rnorm(1000000, mean = 11, sd = 2.5),
+                   datatype=rep("cf.gamlss", n=1000000))
+
+combatgamlss_data <- full_join(combatgamlss_data, data) %>%
+  mutate(datatype=factor(datatype, levels=c("raw", "cf.gamlss"), ordered=T))
+```
+
+```{r}
+combatgamlss_data %>%
+ggplot(aes(size=1.4, fill=datatype, color=datatype)) +
+  geom_density(aes(x=A), alpha = 0.5) +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="raw",]$A), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="cf.gamlss",]$A), size=1, color="#CFB023FF") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+
+combatgamlss_data %>%
+ggplot(aes(size=1.4, fill=datatype, color=datatype)) +
+  geom_density(aes(x=B), alpha = 0.5) +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="raw",]$B), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="cf.gamlss",]$B), size=1, color="#CFB023FF") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+
+combatgamlss_data %>%
+ggplot(aes(size=1.4, fill=datatype, color=datatype)) +
+  geom_density(aes(x=C), alpha = 0.5) +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="raw",]$C), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="cf.gamlss",]$C), size=1, color="#CFB023FF") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+```

### 2024-06-19 paper_main_results.Rmd
presentation figs

--- 2024-06-07 paper_main_results.Rmd
@@
 ```
+
+# Other Figs
+Plotting normal curves for pres
+
+```{r}
+data <- data.frame(A = rnorm(1000000, mean = 7, sd = .5),
+                   B = rnorm(1000000, mean = 10, sd = 2),
+                   C = rnorm(1000000, mean = 13, sd = 3.5),
+                   datatype=rep("raw", n=1000000))
+
+ggplot(data, aes(x = A)) +
+  geom_density(fill = "gray", alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(data$A), size=1) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+
+ggplot(data, aes(x = B)) +
+  geom_density(fill = "gray", alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(data$B), size=1) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+
+ggplot(data, aes(x = C)) +
+  geom_density(fill = "gray", alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(data$C), size=1) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void()
+```
+
+```{r}
+combatgam_data <- data.frame(A = rnorm(1000000, mean = 9.5, sd = 2),
+                   B = rnorm(1000000, mean = 10, sd = 2),
+                   C = rnorm(1000000, mean = 11, sd = 2),
+                   datatype=rep("cf.gam", n=1000000))
+
+combatgam_data <- full_join(combatgam_data, data) %>%
+  mutate(datatype=factor(datatype, levels=c("raw", "cf.gam"), ordered=T))
+```
+
+```{r}
+combatgam_data %>%
+ggplot(aes(fill=datatype, color=datatype)) +
+  geom_density(aes(x=A), alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="raw",]$A), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="cf.gam",]$A), size=1, color="#3D75C2") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void() +
+  theme(legend.position="none")
+
+combatgam_data %>%
+ggplot(aes(fill=datatype, color=datatype)) +
+  geom_density(aes(x=B), alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="raw",]$B), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="cf.gam",]$B), size=1, color="#3D75C2") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void() +
+  theme(legend.position="none")
+
+combatgam_data %>%
+ggplot(aes(fill=datatype, color=datatype)) +
+  geom_density(aes(x=C), alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="raw",]$C), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgam_data[combatgam_data$datatype=="cf.gam",]$C), size=1, color="#3D75C2") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void() +
+  theme(legend.position="none")
+```
+
+```{r}
+combatgamlss_data <- data.frame(A = rnorm(1000000, mean = 9.5, sd = 1),
+                   B = rnorm(1000000, mean = 10, sd = 2),
+                   C = rnorm(1000000, mean = 11, sd = 3),
+                   datatype=rep("cf.gamlss", n=1000000))
+
+combatgamlss_data <- full_join(combatgamlss_data, data) %>%
+  mutate(datatype=factor(datatype, levels=c("raw", "cf.gamlss"), ordered=T))
+```
+
+```{r}
+combatgamlss_data %>%
+ggplot(aes(fill=datatype, color=datatype)) +
+  geom_density(aes(x=A), alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="raw",]$A), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="cf.gamlss",]$A), size=1, color="#CFB023FF") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void() +
+  theme(legend.position="none")
+
+combatgamlss_data %>%
+ggplot(aes(fill=datatype, color=datatype)) +
+  geom_density(aes(x=B), alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="raw",]$B), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="cf.gamlss",]$B), size=1, color="#CFB023FF") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void() +
+  theme(legend.position="none")
+
+combatgamlss_data %>%
+ggplot(aes(fill=datatype, color=datatype)) +
+  geom_density(aes(x=C), alpha = 0.5, size=4) +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="raw",]$C), size=1, color="black") +
+  geom_vline(xintercept = mean(combatgamlss_data[combatgamlss_data$datatype=="cf.gamlss",]$C), size=1, color="#CFB023FF") +
+scale_color_manual(values = c(cf.gam = "#3D75C2", raw= "black", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  scale_fill_manual(values = c(cf.gam = "#3D75C2",raw= "gray", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  labs(x = "Value",
+       y = "Density") +
+  xlim(4, 20) +
+  theme_void() +
+  theme(legend.position="none")
+```
+
+```{r}
+data2 <- data.frame(d = rnorm(1000000, mean = 0, sd = 1)) %>%
+  mutate(sd_cat = if_else(d < (0-1.7) | d > (0+1.7),"ext", "not_ext"))
+
+data2 %>%
+ggplot(aes(size=1.4), color="black")+
+  geom_density(aes(x=d), alpha = 0.5, fill="gray") +
+  geom_vline(xintercept = (0-1.65), size=1, color="#CFB023FF") +
+  geom_vline(xintercept = (0+1.65), size=1, color="#CFB023FF")+
+  theme_void()
+
+data2 %>%
+ggplot(aes(size=1.4), color="black")+
+  geom_density(aes(x=d), alpha = 0.5, fill="gray") +
+  geom_vline(xintercept = (0-1.65), size=1, color="#3D75C2") +
+  geom_vline(xintercept = (0+1.65), size=1, color="#3D75C2")+
+  theme_void()
+
+data2 %>%
+ggplot(aes(size=1.4), color="black")+
+  geom_density(aes(x=d), alpha = 0.5) +
+  theme_void()
+```

### 2024-08-16 paper_lbcc_results.Rmd
notes on scripts used to create inputs:

@@
 ---
-title: "hbm_paper_results"
+title: "ComBatLS LBCC Results"
 output: html_document
@@
 
-Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.
+Rmd to compile, test, and visualize ComBatLS vs ComBat-GAM harmonization in the LBCC.
+
+Relies on outputs derived/functions from:
+1. `lifespan_eda.Rmd`
+2. `run_LBCC_pipeline.sh`
+    - `R_scripts/combat_apply_w_transform.R`
+    - `R_scripts/fit_lifespan_mod_notbv.R`
+    - `R_scripts/fit_lifespan_mod_site_est.R`
+3. `run_LBCC_stats.sh`
+    - `R_scripts/fit_lifespan_centiles.R`
+    - `R_scripts/get_lifespan_site_eff.R`
+    - `R_scripts/get_lifespan_deltas.R`
 
@@
 source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
-#source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
 

### 2024-08-16 paper_prop_results.Rmd
notes on scripts used to create inputs:

@@
 ---
-title: ""
+title: "ComBatLS Varying-Ratio Results"
 output: html_document
@@
 
-Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.
+Rmd to compile and visualize tests of ComBatLS against other harmonization methods when varying the M:F ratios of each simulated site.
+
+Relies on outputs derived/functions from:
+
+1. `ukb_curation-eda.Rmd`
+2. `run_varying_ratios_pipeline.sh`
+   - `R_scripts/permute_sites_varying_ratio.R`
+   - `R_scripts/combat_apply.R`
+   - `R_scripts/fit_perm_basic_mod.R`
+   - `R_scripts/fit_centiles.R`
+   - `R_scripts/summarise_cent_subj-wise.R`
+3. `run_varying_ratios_stats.sh`
+   - `R_scripts/summarise_cent_subj-wise.R`
+   - `R_scripts/ratio_results.R`
+   - `R_scripts/featurewise_sex_bias_test.R`
+   - `R_scripts/remove_extremes_single.R`
+4. `pull_stats.sh`
 
@@
 options("mc.cores")
-
-source("./cubic_scripts/R_scripts/stats_tests.R")
 ```

### 2024-08-16 paper_replication_results.Rmd
notes on scripts used to create inputs:

@@
 ---
-title: "hbm_paper_results"
+title: "ComBatLS Replication Results"
 output: html_document
@@
 ---
+
+Rmd to compile and visualize tests of ComBatLS against other harmonization methods when replicating site simulations 100x.
+
+Relies on outputs derived/functions from:
+
+1. `ukb_curation-eda.Rmd`
+2. `run_permutation_pipeline.sh`
+   - `R_scripts/permute_sites_.R`
+   - `R_scripts/combat_apply.R`
+   - `R_scripts/fit_perm_basic_mod.R`
+   - `R_scripts/fit_centiles.R`
+   - `R_scripts/summarise_cent_subj-wise.R`
+3. `run_permutation_stats.sh`
+   - `R_scripts/summarise_cent_subj-wise.R`
+   - `R_scripts/ratio_results.R`
+   - `R_scripts/featurewise_sex_bias_test.R`
+   - `R_scripts/remove_extremes_single.R`
+4. `pull_stats.sh`
 

### 2024-08-16 paper_replication_results.Rmd
remove deprecated scripts

@@
 2. `run_permutation_pipeline.sh`
-   - `R_scripts/permute_sites_.R`
+   - `R_scripts/permute_sites.R`
    - `R_scripts/combat_apply.R`

### 2024-08-16 cubic_scripts/R_scripts/fit_basic_lm.R

@@ -1,28 +1,0 @@
-#Basic script to fit very simple gamlss model & test significance of sex term in sigma
-
-set.seed(12345)
-
-#LOAD PACKAGES
-library(dplyr)
-library(data.table)
-
-source("R_scripts/gamlss_helper_funs.R")
-
-#GET ARTS
-args <- commandArgs(trailingOnly = TRUE)
-df <- fread(args[1], stringsAsFactors = TRUE, na.strings = "")
-pheno <- as.character(args[2])
-save_path <- args[3]
-
-#FIT BASIC MODEL
-model <- lm(as.formula(paste0(pheno,"age_days + sexMale")), data=df)
-
-#SAVE
-csv_name <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(as.character(args[1])))
-csv_rename <- gsub("_", "-", csv_name)
-
-saveRDS(model,paste0(save_path, "/", pheno, "_", csv_rename, "_lm.rds"))
-
-#SIGNIFICANCE TESTING
-df <- summary(model)$pTerms.table
-fwrite(df, file=paste0(save_path, "/", pheno, "_", csv_rename, "_lm_sig.csv"))

### 2024-08-16 cubic_scripts/R_scripts/get_lifespan_summaries.R

@@ -1,40 +1,0 @@
-# load models fit on a given csv from .rds and extract centile scores
-
-set.seed(12345)
-
-#LOAD PACKAGES
-library(gamlss)
-library(dplyr)
-library(data.table)
-library(tibble)
-
-print(getwd())
-
-source("R_scripts/gamlss_helper_funs.R")
-
-#get args
-args <- commandArgs(trailingOnly = TRUE)
-read_path <- as.character(args[1]) #path to gamlss models
-save_path <- as.character(args[2]) #path to save output csv
-fname_str <- as.character(args[3]) #string to look for in data .csv and model .rds files
-#make sure - and _ are searched interchangeably
-fname_str_search <- gsub("_|-", "[-_]", fname_str)
-print(fname_str)
-print(fname_str_search)
-
-#find correct models
-all.model.files <- list.files(path = read_path, pattern = "mod.rds", full.names = TRUE)
-print("all model files:")
-print(all.model.files[1:4])
-model.files <- all.model.files[grep(fname_str_search, all.model.files)]
-print("selected model files:")
-print(model.files[1:4])
-
-#extract values from summary table
-summary.list <- lapply(model.files, get.gamlss.summary)
-names(summary.list) <- lapply(model.files, get.y)
-#dataframe
-summary.df <- bind_rows(summary.list, .id=pheno)
-
-#write out
-fwrite(summary.df, paste0(save_path, "/", fname_str, "_full_sum.csv"))

### 2024-08-16 cubic_scripts/R_scripts/lifespan_delta_lms.R

@@ -1,127 +1,0 @@
-
-##LOAD PACKAGES
-library(dplyr)
-library(data.table)
-library(tibble)
-library(broom.mixed)
-library(stats)
-
-#get args
-args <- commandArgs(trailingOnly = TRUE)
-delta_path <- as.character(args[1]) #path to delta csv
-save_path <- as.character(args[2]) #path to delta csv
-
-pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
-
-##READ IN DELTAS
-deltas_df <- fread(delta_path)
-
-#######################################
-# CENTILES
-#######################################
-
-#initialize empty df to store outputs
-tests.df <- data.frame("term" = character(),
-                   "estimate" = double(),
-                   "std.error" = double(),
-                   "statistic" = double(),
-                   "p.value" = double(),
-                   "iv" = character(),
-                   "pheno" = character())
-
-##Loop each Pheno
-
-#only test phenos w models that converged
-delta.pheno_list <- paste0("delta.", pheno_list)
-phenos_to_test <- intersect(delta.pheno_list, names(deltas_df))
-
-#create age_years 
-deltas_df <- deltas_df %>%
-  mutate(age_yrs = age_days/365.25)
-
-for (pheno in phenos_to_test) {
-  # Test Sex
-  sex.df <- tidy(lm(formula=as.formula(paste(pheno, "~ sexMale")), data=deltas_df)) %>%
-    dplyr::filter(term =="sexMale") %>%
-    mutate(iv = "sexMale")
-  
-  # Test Age
-
-  age.df <- tidy(lm(formula=as.formula(paste(pheno, "~ age_yrs")), data=deltas_df)) %>%
-    dplyr::filter(term =="age_yrs") %>%
-    mutate(iv = "age_yrs")
-  
-  # Test Sex*Age
-  int.df <- tidy(lm(formula=as.formula(paste(pheno, "~ sexMale*age_yrs")), data=deltas_df)) %>%
-    dplyr::filter(term =="sexMale:age_yrs") %>%
-    mutate(iv = "sexMale:age_yrs")
-  
-  # Compile w/in phenotype
-  stopifnot(dim(sex.df) == dim(age.df) & dim(age.df) == dim(int.df))
-  pheno.df <- rbind(sex.df, age.df, int.df)
-  pheno.df$pheno <- gsub("delta.", "", pheno)
-  
-  # Add to other phenotypes
[…]

### 2024-08-16 cubic_scripts/R_scripts/lifespan_results.R

@@ -1,27 +1,0 @@
-# change in centile scores for subjects when calculated on data harmonized with ComBat GAM vs ComBatLS
-# also run basic stats tests
-
-#LOAD PACKAGES
-library(data.table)
-library(tidyverse)
-library(dplyr)
-
-#get args
-args <- commandArgs(trailingOnly = TRUE)
-delta_path <- as.character(args[1]) #path to delta csv
-save_path <- as.character(args[2]) #path to save output csvs
-fname_str <- as.character(args[3]) #string for saving name
-
-devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/stats_tests.R")
-
-#pheno lists
-pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
-pheno_list.z <- paste0(pheno_list, ".z")
-
-#LOAD DFS
-#read in data
-delta.df <- fread(gam_csv_path)
-
-
-print("DONE")
-

### 2024-08-16 cubic_scripts/qsub_lbcc_deltas_stats.sh

@@ -1,24 +1,0 @@
-#!/bin/bash
-
-#see how different predicted centiles are for each subject when ComBat-GAM or ComBatLS are used to harmonize LBCC
-
-#######################################################################################
-# SET PATHS
-img=/cbica/home/gardnerm/software/containers/r_gamlss_0.0.1.sif #singularity image
-base=/cbica/home/gardnerm/combat-biovar #base path (cubic) - also path that output csvs will be saved to
-r_script=$base/cubic_scripts/R_scripts/lifespan_delta_lms.R #path to .R script
-cent_path=$base/lifespan/centile_csvs #path to centile csvs
-bash_dir=$base/lifespan/centile_qsubs
-
-#search string:
-config="batch.study"
-#######################################################################################
-cd $base/cubic_scripts #to source functions correctly
-#######################################################################################
-bash_script=$bash_dir/${config}_delta_lms.sh
-touch $bash_script
-
-echo "singularity run --cleanenv $img Rscript --save $r_script $cent_path/${config}_pred_deltas.csv $cent_path" > $bash_script
-
-#qsub bash script
-qsub -N $config -o $bash_dir/${config}_delta_lms_out.txt -e $bash_dir/${config}_delta_lms_err.txt $bash_script

### 2024-08-21 eda/lifespan_eda.Rmd
working on retaining more output from t-tests

@@
 cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)
+```
 
+drop UKB for sharing
+```{r}
+cn.df_dropUKB <- cn.df %>%
+  dplyr::filter(study != "UKB")
+fwrite(cn.df_dropUKB, file=paste0(base_path, "/data/lifespan_dataset_dkatlas_CNonly_euler_noukb.csv"))
+```
+
+```{r}
 #drop repeat scans
@@
 ```
+
+drop UKB for sharing
+```{r}
+cn.df.imp_dropUKB <- cn.df.imp %>%
+  dplyr::filter(study != "UKB")
+fwrite(cn.df.imp_dropUKB, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler_noukb.csv"))
+```

### 2024-08-21 paper_replication_results.Rmd
working on retaining more output from t-tests

@@
 }
-pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata, gtExtras, webshot2)
+pacman::p_load(data.table, ggplot2, tidyverse, ggseg, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, gdata, gtExtras, webshot2) #ggsegExtra
 #Sys.setenv("MC_CORES"=10L)
@@
 source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
-aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+remotes::install_github("BGDlab/ggseg-tissueclass-atlas")
 ```
@@
 
+#t-test of median sex diffs in centile error distributions vs 0
 t.test.cent.sex.test.med <- cent.sex.test.med %>%
@@
             t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
+            t.test.stat = tidy(t.test(median_of_median_sex_diff))$statistic,
+            t.test.df = tidy(t.test(median_of_median_sex_diff))$parameter,
             x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
@@
 #get vals
-t.test.cent.sex.test.med
-
+t.test.cent.sex.test.med_tbl <- t.test.cent.sex.test.med %>%
+  dplyr::select(dataset, pheno_cat, t.test.est, t.test.stat, t.test.df, p.val_fdr, sig_fdr) %>%
+  mutate(dataset = case_when(dataset == "cf" ~ "ComBat w/o Covariates",
+                                   dataset == "cf.lm" ~ "Linear ComBat",
+                                   dataset == "cf.gam" ~ "ComBat-GAM",
+                                   dataset == "cf.gamlss" ~ "ComBatLS")) %>%
+  rename("ComBat Method" = dataset,
+         "Phenotype Category" = pheno_cat,
+         "Estimate" = t.test.est,
+         "t-statistic" = t.test.stat,
+         "dfs" = t.test.df,
+         "p-value (FDR-corrected)" = p.val_fdr,
+         "Significance" = sig_fdr) %>%
+  gt() %>%
+  tab_style(style = list(cell_fill(color = "gray95")),
+    locations = cells_row_groups()) %>%
+  tab_options(column_labels.font.weight = "bold") %>%
+  tab_header(title = "One-sided t-tests of sex differences in centile errors across 100 replications")
+gtsave(t.test.cent.sex.test.med_tbl, filename="figs/cent_median_sexdiff_table.png")
+```
+
+plots
+```{r}
 t.test.cent.sex.test.med %>%
@@
 
+#pretty table
+t.test.z.sex.test.med_tbl <- t.test.z.sex.test.med %>%
+  dplyr::select(dataset, pheno_cat, t.test.est, t.test.stat, t.test.df, p.val_fdr, sig_fdr) %>%
+  mutate(dataset = case_when(dataset == "cf" ~ "ComBat w/o Covariates",
+                                   dataset == "cf.lm" ~ "Linear ComBat",
+                                   dataset == "cf.gam" ~ "ComBat-GAM",
+                                   dataset == "cf.gamlss" ~ "ComBatLS")) %>%
+  rename("ComBat Method" = dataset,
+         "Phenotype Category" = pheno_cat,
+         "Estimate" = t.test.est,
+         "t-statistic" = t.test.stat,
+         "dfs" = t.test.df,
+         "p-value (FDR-corrected)" = p.val_fdr,
+         "Significance" = sig_fdr) %>%
+  gt() %>%
+  tab_style(style = list(cell_fill(color = "gray95")),
+    locations = cells_row_groups()) %>%
+  tab_options(column_labels.font.weight = "bold") %>%
+  tab_header(title = "One-sided t-tests of sex differences in z-score errors across 100 replications")
+gtsave(t.test.z.sex.test.med_tbl, filename="figs/z_median_sexdiff_table.png")
+```
+
+plot
+```{r}
 t.test.z.sex.test.med.filt <- t.test.z.sex.test.med %>% 
@@
   ),
-  all_cent_sum_boxplt_wtest,
+  all_cent_sum_boxplt.diff,
   ncol=2, nrow=1,

### 2024-08-21 cubic_scripts/R_scripts/ranked_welch_tests.R

@@ -124,7 +124,7 @@ pairwise.rank.welch.t.test <- function(x, g, p.adjust.method = p.adjust.methods,
       alternative = alternative, ...
     ))$estimate
 
-    bigger_group <- ifelse(est > 0, i, j)
+    bigger_group <- as.character(ifelse(est > 0, i, j))
     return(bigger_group)
   }
   # (estimate = estimate1-estimate2) ~ positive -> levels(g)[1], negative -> levels(g)[2]
@@ -134,7 +134,7 @@ pairwise.rank.welch.t.test <- function(x, g, p.adjust.method = p.adjust.methods,
   ans <- list(
     method = METHOD, data.name = DNAME,
     p.value = PVAL, p.adjust.method = p.adjust.method,
-    larger_group = est_matrix
+    larger_group = big_group
   )
   class(ans) <- "pairwise.htest" # prevents larger_group from displaying but enables tidy()
   ans

### 2024-08-21 cubic_scripts/R_scripts/ratio_results.R

@@ -1,4 +1,4 @@
-1597/4# get results for varying-proportion analyses more quickly using cubic
+# get results for varying-proportion analyses more quickly using cubic
 
 set.seed(12345)
 

### 2024-08-21 cubic_scripts/R_scripts/stats_tests.R

@@ -83,7 +83,9 @@ centile.t.tests.full_result <- function(df, feature_list, comp_multiplier = 1) {
     "pheno" = character(),
     "group1" = character(),
     "group2" = character(),
-    "p.value" = double()
+    "p.value" = double(),
+    "df" = double(),
+    "estimate" = double()
   )
   # run tests
   attach(df)
@@ -93,19 +95,25 @@ centile.t.tests.full_result <- function(df, feature_list, comp_multiplier = 1) {
                                                      paired = TRUE,
                                                      p.adj = "none"))
 
-    # also find bigger group
-    df.pairwise.est <- tidy(pairwise.rank.maxgrp(x = df[[pheno]],
+    #GET ADDITIONAL TEST PARAMETERS 
+    # this is admittedly hackey, but basically to use tidy() the object needs
+    # to be class pairwise.htest, which only has cols for the 2 groups and p.val
+    # so to get additional stats, i'm just rewriting funs that return other stats
+    # in p-value col and renaming the col :/
+    
+    # find bigger group
+    df.pairwise.grp <- tidy(pairwise.rank.maxgrp(x = df[[pheno]],
                                                  g = df[["dataset"]],
                                                  paired = TRUE,
                                                  p.adj = "none"))
     df.pairwise.est <- df.pairwise.est %>%
       rename(bigger_group = p.value) # fix name
-    # print(df.pairwise.est)
+    
+    #confirm list
+    stopifnot(dim(df.pairwise.est) == dim(df.pairwise.t))
 
     # join
     df.pairwise.t2 <- left_join(df.pairwise.t, df.pairwise.est)
-
-    stopifnot(nrow(df.pairwise.t2) == nrow(df.pairwise.t))
 
     df.pairwise.t2$pheno <- pheno
     t.df <- rbind(t.df, df.pairwise.t2)

### 2024-11-15 eda/lifespan_eda.Rmd
adding plots to show site-level sex imbalances

@@
 drop UKB for sharing
-```{r}
+```{r, eval=FALSE}
 cn.df_dropUKB <- cn.df %>%
@@
 fwrite(cn.df_dropUKB, file=paste0(base_path, "/data/lifespan_dataset_dkatlas_CNonly_euler_noukb.csv"))
+```
+
+Plot full samples for supplemental figures looking at imbalanced demographics across sites
+
+```{r}
+#just get first scan
+cn.df_first <- cn.df %>%
+  drop_na(c(age_days, participant, study)) %>%
+  arrange(age_days) %>%
+  group_by(participant) %>%
+  slice_head(n=1) %>%
+  mutate(site_full = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="&")))) %>%
+  ungroup() %>%
+  group_by(site_full) %>%
+  filter(n() > 1) #keep sites with at least 2 ppl
+```
+
+check for sex differences in sites
+
+```{r}
+sfig_full_sample <- ggplot(cn.df_first) +
+  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=interaction(site, study, sep="&"), fill=sex)) +
+  scale_y_continuous(labels = scales::percent) +
+  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  guides(x = ggh4x::guide_axis_nested(delim="&")) +
+  theme_linedraw() +
+  theme(
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    ggh4x.axis.nesttext.x = element_text(angle = 45, vjust = .9, hjust=1, size=5),
+    ggh4x.axis.nestline = element_line(linetype = 1),
+    legend.position="bottom"
+  ) +
+  geom_hline(yintercept=.5, color="lightgray", linetype="dashed") +
+  labs(x="Data Collection Site", y="Sample Composition", fill="Sex")
+ggsave(paste0(base_path, "/figs/sfig_full_sample.jpeg"), sfig_full_sample, width=15, height=7, units="in")
+```
+
+look into a couple studies specifically
+
+```{r}
+plot_by_sex <- function(df) {
+  df_prepared <- df %>%
+    group_by(site, sex) %>%
+    summarise(count = n(), .groups = "drop") %>%
+    group_by(site) %>%
+    mutate(percentage = count / sum(count)) %>%
+    ungroup()
+
+  ggplot(df_prepared, aes(x = site, y = percentage, fill = sex)) +
+    geom_bar(stat = "identity", position = "stack") +
+    geom_text(
+      aes(label = count, group = sex),
+      position = position_stack(vjust = 0.5),
+      color = "white",
+      size=2.5
+    ) +
+    scale_y_continuous(labels = scales::percent_format()) +
+    theme_linedraw() +
+    xlab("Site") +
+    ylab("Percentage") +
+    labs(fill="Sex") +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+}
+
+#good balance
+p1<- cn.df_first %>% 
+  filter(study == "ABCD") %>% 
+  plot_by_sex() + labs(title="ABCD")
+
+p2 <- cn.df_first %>% 
+  filter(study == "UKB") %>% 
+  plot_by_sex() + labs(title="UK Biobank")
+
+#bad balance
+p3 <- cn.df_first %>% 
+  filter(study == "cVEDA") %>% 
+  plot_by_sex() + labs(title="cVEDA")
+
+p4 <- cn.df_first %>% 
+  filter(study == "ADNI") %>% 
+  plot_by_sex() + labs(title="ADNI")
+
+p5 <- cn.df_first %>% 
+  filter(study == "abide1") %>% 
+  plot_by_sex() + labs(title="ABIDE 1")
+
+p6 <- cn.df_first %>% 
+  filter(study == "ADHD200") %>% 
+  plot_by_sex() + + labs(title="ADHD200")
+
+sfig_samples <- ggarrange(p1, p2, p3, p4, p5, p6, 
+                          ncol=2, nrow=3, 
+                          common.legend = TRUE, legend="right",
+                          align="v")
+ggsave(paste0(base_path, "/figs/sfig_subsamples.jpeg"), sfig_samples, width=15, height=10, units="in")
 ```
@@
 
-```{r}
-ggplot(cn.df.imp) +
-  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
-  scale_y_continuous(labels = scales::percent) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-```
-
 Write out

### 2024-11-15 eda/lifespan_eda.Rmd
adding plots to show site-level sex imbalances
clean up supplemental figs

@@
   slice_head(n=1) %>%
-  mutate(site_full = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="&")))) %>%
+  mutate(site_full = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="&"))),
+         age_yrs = (age_days-280)/365.25) %>%
   ungroup() %>%
@@
 sfig_full_sample <- ggplot(cn.df_first) +
-  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=interaction(site, study, sep="&"), fill=sex)) +
-  scale_y_continuous(labels = scales::percent) +
+  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=interaction(site, study, sep="&"), fill=sex)) +
+  scale_x_continuous(labels = scales::percent) +
   #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  guides(x = ggh4x::guide_axis_nested(delim="&")) +
+  guides(y = ggh4x::guide_axis_nested(delim="&")) +
   theme_linedraw() +
   theme(
-    axis.ticks.x = element_blank(),
-    axis.text.x = element_blank(),
-    ggh4x.axis.nesttext.x = element_text(angle = 45, vjust = .9, hjust=1, size=5),
+    axis.ticks.y = element_blank(),
+    axis.text.y = element_blank(),
+    ggh4x.axis.nesttext.y = element_text(size=6),
     ggh4x.axis.nestline = element_line(linetype = 1),
-    legend.position="bottom"
+    legend.position="right"
   ) +
-  geom_hline(yintercept=.5, color="lightgray", linetype="dashed") +
-  labs(x="Data Collection Site", y="Sample Composition", fill="Sex")
-ggsave(paste0(base_path, "/figs/sfig_full_sample.jpeg"), sfig_full_sample, width=15, height=7, units="in")
+  geom_vline(xintercept=.5, color="lightgray", linetype="dashed") +
+  labs(y="Data Collection Site", x="Sample Composition", fill="Sex")
+ggsave(paste0(base_path, "/figs/sfig_full_sample.jpeg"), sfig_full_sample, width=6, height=14, units="in")
 ```
@@
   filter(study == "ADHD200") %>% 
-  plot_by_sex() + + labs(title="ADHD200")
+  plot_by_sex() + labs(title="ADHD200")
 
@@
                           common.legend = TRUE, legend="right",
-                          align="v")
-ggsave(paste0(base_path, "/figs/sfig_subsamples.jpeg"), sfig_samples, width=15, height=10, units="in")
+                          align="v",
+                          labels = "AUTO")
+ggsave(paste0(base_path, "/figs/sfig_subsamples.jpeg"), sfig_samples, 
+       width=15, height=10, units="in")
+```
+
+look at age variability
+
+```{r}
+sfig_full_sample_age <- ggplot(cn.df_first) +
+  geom_boxplot(aes(x=age_yrs, y=interaction(site, study, sep="&"))) +
+  guides(y = ggh4x::guide_axis_nested(delim="&")) +
+  theme_linedraw() +
+  theme(
+    axis.ticks.y = element_blank(),
+    axis.text.y = element_blank(),
+    ggh4x.axis.nesttext.y = element_text(size=6),
+    ggh4x.axis.nestline = element_line(linetype = 1),
+    axis.title.y=element_blank()
+  ) +
+  labs(x="Age (Years Post-Birth)")
+ggsave(paste0(base_path, "/figs/sfig_full_sample_age.jpeg"), sfig_full_sample_age, width=6, height=14, units="in")
+
+sitewise_full_plot <- ggarrange(sfig_full_sample, sfig_full_sample_age,
+          ncol=2, nrow=1, align = "hv",
+          common.legend = TRUE, legend = "bottom",
+          labels="AUTO")
+ggsave(paste0(base_path, "/figs/sitewise_full_plot.jpeg"), sitewise_full_plot, 
+       width=15, height=15, units="in")
+```
+
+```{r}
+plot_by_age <- function(df) {
+  ggplot(df, aes(y = age_yrs, x=site)) +
+    geom_boxplot() +
+    theme_linedraw() +
+    theme(legend.position = "none",
+          axis.text.x = element_text(angle = 45, hjust = 1)) +
+    xlab("Site") +
+    ylab("Age")
+}
+
+p1.age <- cn.df_first %>% 
+  filter(study == "ABCD") %>% 
+  plot_by_age() + labs(title="ABCD")
+
+p2.age <- cn.df_first %>% 
+  filter(study == "UKB") %>% 
+  plot_by_age() + labs(title="UK Biobank")
+
+#bad balanced
+p3.age <- cn.df_first %>% 
+  filter(study == "EMBARC") %>% 
+  plot_by_age() + labs(title="EMBARC")
+
+p4.age <- cn.df_first %>% 
+  filter(study == "ADNI") %>% 
+  plot_by_age() + labs(title="ADNI")
+
+p5.age <- cn.df_first %>% 
+  filter(study == "abide1") %>% 
+  plot_by_age() + labs(title="ABIDE 1")
+
+p6.age <- cn.df_first %>% 
+  filter(study == "ADHD200") %>% 
+  plot_by_age() + labs(title="ADHD200")
+
+sfig_samples.age <- ggarrange(p1.age, p2.age, p3.age, p4.age, p5.age, p6.age, 
+                          ncol=2, nrow=3,
+                          align="v",
+                          labels = "AUTO")
+ggsave(paste0(base_path, "/figs/sfig_subsamples_age.jpeg"), sfig_samples.age, 
+       width=15, height=10, units="in")
 ```

### 2024-11-15 paper_main_results.Rmd
typos

@@
 pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x, skimr)
+remotes::install_github("BGDlab/ggsegTissue")
+library(ggsegTissue)
 ```
@@
 ```{r}
-aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+# aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
 data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"
@@
   geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile)), nudge_x = -0.75, size=3) +
-  xlab("Age (days)") + ylab("Volume (mm3)") +
+  xlab("Age (Years Post-Conception)") + ylab("Volume (mm3)") +
   labs(title="Ventricular volume brain charts")
@@
 
-? <- ggboxplot(raw_sex_eff, x = "pheno_cat", y = "sex_standard",
+sex_eff_comp_plt <- ggboxplot(raw_sex_eff, x = "pheno_cat", y = "sex_standard",
           color = "pheno_cat")+ 

### 2025-01-13 paper_lbcc_results.Rmd
updating atlas with ggsegTissue

@@
 #ATLAS
-aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+remotes::install_github("BGDlab/ggsegTissue")
+library(ggsegTissue)
+
 #write out new dk atlas w/o medial wall
@@
   theme(axis.title.x=element_blank(), axis.text.x = element_blank()) +
-  labs(color="Sex", y = "Age (Years)")
+  labs(color="Sex", y = "Age\n(Years Post-Conception)")
 
@@
   theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
-  labs(color="Sex", y = "Age (Years)", y="")
+  labs(color="Sex", y = "Age\n(Years Post-Conception)", y="")
 ```

### 2025-01-13 eda/lifespan_eda.Rmd
update raw data with corrected VETSA, correct age_days to account for gestation

@@
 
-Load data from Jakob - updated with euler numbers!
+Load data from Jakob - updated with euler numbers and corrected VETSA study
 ```{r}
-cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)
-```
-
-drop UKB for sharing
-```{r, eval=FALSE}
-cn.df_dropUKB <- cn.df %>%
-  dplyr::filter(study != "UKB")
-fwrite(cn.df_dropUKB, file=paste0(base_path, "/data/lifespan_dataset_dkatlas_CNonly_euler_noukb.csv"))
+cn.df <- fread(paste0(base_path,"/data/v1_2_1_nature.csv"), stringsAsFactors = TRUE)
 ```
@@
 nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
+nrow(cn.df) - nrow(cn.df.unique)
 ```
@@
 cn.df.noinf <- cn.df.unique %>%
-  dplyr::filter(age_days >= (365.25*3)) %>%
+  dplyr::filter(age_days >= (365.25*3+280)) %>%
   mutate(study = droplevels(study),
@@
 #age
-min(cn.df.imp$age_days)/365.25
-max(cn.df.imp$age_days)/365.25
-mean(cn.df.imp$age_days)/365.25
+min(cn.df.imp$age_days)/365.25-(280/365.25)
+max(cn.df.imp$age_days)/365.25-(280/365.25)
+mean(cn.df.imp$age_days)/365.25-(280/365.25)
 
@@
 ```
-
-drop UKB for sharing
-```{r}
-cn.df.imp_dropUKB <- cn.df.imp %>%
-  dplyr::filter(study != "UKB")
-fwrite(cn.df.imp_dropUKB, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler_noukb.csv"))
-```

### 2025-01-16 paper_replication_results.Rmd
first chunk sex-bias tests updated to no fdr correction

@@
 cent.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
+
+cent.subj.sex.test.df <- cent.subj.sex.test.df %>%
+  mutate(sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
 ```
@@
 cent.subj.sex.test.df %>%
-  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
-  ggplot(aes(x=perm, fill=sig_fdr)) +
+  ggplot(aes(x=perm, fill=significant)) +
   geom_bar() +
-  #geom_text(stat='count', aes(label=..count..)) +
   facet_wrap(~dataset) +
   theme(axis.text.x = element_blank()) +
-  scale_fill_manual(values = c("lightblue", "red"))
-```
+  scale_fill_manual(values = c("#EA6F5F", "#F3AEA6", "#5FDAEA"))
+```
+
+removing fdr correction
 
@@
 cent.subj.sex.test.df %>%
-  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+  mutate(median_sex_diff_fdr = case_when(sig_fdr == FALSE ~ NA,
+                              TRUE ~ median_sex_diff*100),
+         median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff*100)) %>%
@@
   summarize(min_sex_diff = min(median_sex_diff, na.rm=TRUE),
-            max_sex_diff = max(median_sex_diff, na.rm=TRUE))
+            min_sex_diff_fdr = min(median_sex_diff_fdr, na.rm=TRUE),
+            max_sex_diff = max(median_sex_diff, na.rm=TRUE),
+            max_sex_diff_fdr = max(median_sex_diff_fdr, na.rm=TRUE))
 
@@
                                p.val_fdr >= 0.05 ~ ""),
+         sig = case_when(t.test.p < 0.001 ~ "***",
+                               t.test.p >= 0.001 & t.test.p < 0.01 ~ "**",
+                               t.test.p >= 0.01 & t.test.p < 0.05 ~ "*",
+                               t.test.p >= 0.05 ~ ""),
          y_pos = case_when(dataset=="cf" ~ 9.3,
@@
 t.test.cent.sex.test.med_tbl <- t.test.cent.sex.test.med %>%
-  dplyr::select(dataset, pheno_cat, t.test.est, t.test.stat, t.test.df, p.val_fdr, sig_fdr) %>%
+  dplyr::select(dataset, pheno_cat, t.test.est, t.test.stat, t.test.df, t.test.p, sig, sig_fdr) %>%
   mutate(dataset = case_when(dataset == "cf" ~ "ComBat w/o Covariates",
@@
          "dfs" = t.test.df,
-         "p-value (FDR-corrected)" = p.val_fdr,
-         "Significance" = sig_fdr) %>%
+         "p-value" = t.test.p,
+         "Significance" = sig,
+         "Significance (FDR-corrected)" = sig_fdr) %>%
   gt() %>%
@@
   tab_options(column_labels.font.weight = "bold") %>%
-  tab_header(title = "One-sided t-tests of sex differences in centile errors across 100 replications")
+  tab_header(title = "One-sample t-tests of sex differences in centile errors across 100 replications")
 gtsave(t.test.cent.sex.test.med_tbl, filename="figs/cent_median_sexdiff_table.png")
@@
   ggplot() +
-  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")
+  geom_col(aes(x=pheno_cat, y=sig, fill=dataset), position="dodge")
 
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos*100, y=y_pos),
+  geom_text(aes(label = sig, x= x_pos*100, y=y_pos),
               data = t.test.cent.sex.test.med.filt) +
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos*100, y=y_pos),
+  geom_text(aes(label = sig, x= x_pos*100, y=y_pos),
               data = t.test.cent.sex.test.med) +
@@
     group_by(dataset) %>%
-    summarise(p_val = t.test(female_ratio, mu=mu)$p.val,
-              pheno_cat = cat,
+    summarise(pheno_cat = cat,
               max_val = max(female_ratio),
-              mu=mu) %>%
-    mutate(p.val_fdr = p.adjust(p_val, method="fdr", n = (4*4*2)), #adjusting for 4 methods * 4 pheno classes * 2 centile categories
+              mu=mu,
+            t.test.p = tidy(t.test(female_ratio, mu=mu))$p.value,
+            t.test.est = tidy(t.test(female_ratio, mu=mu))$estimate,
+            t.test.stat = tidy(t.test(female_ratio, mu=mu))$statistic,
+            t.test.df = tidy(t.test(female_ratio, mu=mu))$parameter,) %>%
+    mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = (4*4*2)), #adjusting for 4 methods * 4 pheno classes * 2 centile categories
            sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
@@
                                p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ "NS"))
+                               p.val_fdr >= 0.05 ~ "NS"),
+           sig = case_when(t.test.p < 0.001 ~ "***",
+                               t.test.p >= 0.001 & t.test.p < 0.01 ~ "**",
+                               t.test.p >= 0.01 & t.test.p < 0.05 ~ "*",
+                               t.test.p >= 0.05 ~ "NS")
+           )
   return(out)
@@
   dplyr::filter(dataset != "raw")
+```
+
+Table for supplement
+```{r}
+names(all_t_test.df)
+
+highlow_t_test.df_tbl <- all_t_test.df %>%
+  dplyr::select(dataset, pheno_cat, source, t.test.est, t.test.stat, t.test.df, t.test.p, sig, sig_fdr) %>%
+  mutate(dataset = case_when(dataset == "cf" ~ "ComBat w/o Covariates",
+                                   dataset == "cf.lm" ~ "Linear ComBat",
+                                   dataset == "cf.gam" ~ "ComBat-GAM",
+                                   dataset == "cf.gamlss" ~ "ComBatLS"),
+         source = ifelse(source=="high", ">80%", "<20%")) %>%
+  rename("ComBat Method" = dataset,
+         "Centile Group" = source,
+         "Phenotype Category" = pheno_cat,
+         "Estimate" = t.test.est,
+         "t-statistic" = t.test.stat,
+         "dfs" = t.test.df,
+         "p-value" = t.test.p,
+         "Significance" = sig,
+         "Significance (FDR-corrected)" = sig_fdr) %>%
+  gt() %>%
+  tab_style(style = list(cell_fill(color = "gray95")),
+    locations = cells_row_groups()) %>%
+  tab_options(column_labels.font.weight = "bold") %>%
+  tab_header(title = "One-sample t-test of female's over- or under-representation in high and low centile groups")
+gtsave(highlow_t_test.df_tbl, filename="figs/cent_sexratios_highlow_table.png")
 ```
@@
   geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
-  geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  geom_text(aes(label = sig, y = max_val + 0.01),
               data = all_t_test.df) +

### 2025-01-16 paper_replication_results.Rmd
finish removing FDR correction from sex-bias tests in replication

@@
 noext.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
-noext.sex.test.df <- bind_rows(noext.sex.test.list)
+noext.sex.test.df <- bind_rows(noext.sex.test.list) %>%
+  mutate(sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
 ```
@@
 noext.sex.test.df %>%
-  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
-  ggplot(aes(x=perm, fill=sig_fdr)) +
+  ggplot(aes(x=perm, fill=significant)) +
   geom_bar() +
@@
   theme(axis.text.x = element_blank()) +
-  scale_fill_manual(values = c("lightblue", "red"))
+    scale_fill_manual(values = c("#EA6F5F", "#F3AEA6", "#5FDAEA"))
 ```
@@
                                p.val_fdr >= 0.05 ~ ""),
+         sig = case_when(t.test.p < 0.001 ~ "***",
+                               t.test.p >= 0.001 & t.test.p < 0.01 ~ "**",
+                               t.test.p >= 0.01 & t.test.p < 0.05 ~ "*",
+                               t.test.p >= 0.05 ~ ""),
          y_pos = case_when(dataset=="cf" ~ 6,
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos*100, y=y_pos),
+  geom_text(aes(label = sig, x= x_pos*100, y=y_pos),
               data = next.t.test.cent.sex.test.med) +
@@
 z.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
-z.sex.test.df <- bind_rows(z.sex.test.list)
+z.sex.test.df <- bind_rows(z.sex.test.list) %>%
+  mutate(sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
 ```
@@
 z.sex.test.df %>%
-  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
-  ggplot(aes(x=perm, fill=sig_fdr)) +
+  ggplot(aes(x=perm, fill=significant)) +
   geom_bar() +
-  #geom_text(stat='count', aes(label=..count..)) +
   facet_wrap(~dataset) +
   theme(axis.text.x = element_blank()) +
-  scale_fill_manual(values = c("lightblue", "red"))
+  scale_fill_manual(values = c("#EA6F5F", "#F3AEA6", "#5FDAEA"))
 ```
@@
   labs(x="Male - Female Median Centile Errors", y="Number of Features") +
-  ggtitle("Sex differences in median centile errors across 100 sampling replications")
-ggsave(filename="figs/z.sex_bias_hist_plt.z.png", sex_bias_hist_plt)
+  ggtitle("Sex differences in median z-score errors across 100 sampling replications")
+ggsave(filename="figs/z.sex_bias_hist_plt.z.png", z.sex_bias_hist_plt)
 ```
@@
                                p.val_fdr >= 0.05 ~ ""),
+         sig = case_when(t.test.p < 0.001 ~ "***",
+                               t.test.p >= 0.001 & t.test.p < 0.01 ~ "**",
+                               t.test.p >= 0.01 & t.test.p < 0.05 ~ "*",
+                               t.test.p >= 0.05 ~ ""),
          y_pos = case_when(dataset=="cf" ~ 155,
@@
 t.test.z.sex.test.med_tbl <- t.test.z.sex.test.med %>%
-  dplyr::select(dataset, pheno_cat, t.test.est, t.test.stat, t.test.df, p.val_fdr, sig_fdr) %>%
+  dplyr::select(dataset, pheno_cat, t.test.est, t.test.stat, t.test.df, t.test.p, sig, sig_fdr) %>%
   mutate(dataset = case_when(dataset == "cf" ~ "ComBat w/o Covariates",
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos, y=y_pos),
+  geom_text(aes(label = sig, x= x_pos, y=y_pos),
               data = t.test.z.sex.test.med.filt) +
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos, y=y_pos),
+  geom_text(aes(label = sig, x= x_pos, y=y_pos),
               data = t.test.z.sex.test.med) +
@@
   geom_hline(data=all_z.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
-  geom_text(aes(label = sig_fdr, y = max_val + 0.01),
+  geom_text(aes(label = sig, y = max_val + 0.01),
               data = z.all_t_test.df) +
@@
   geom_hline(yintercept=0, linetype="dashed")+
-  geom_text(aes(label = sig_fdr, y = .07),
+  geom_text(aes(label = sig, y = .07),
               data = z.all_t_test.df) +
@@
 noext.sex.test.z.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
-noext.sex.test.z.df <- bind_rows(noext.sex.test.z.list)
-```
-
-simple bar-plot to count regions w sigificant sex-biases
+noext.sex.test.z.df <- bind_rows(noext.sex.test.z.list) %>%
+  mutate(sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
+```
+
+simple bar-plot to count regions w significant sex-biases
 
@@
 noext.sex.test.z.df %>%
-  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
-  ggplot(aes(x=perm, fill=sig_fdr)) +
+  ggplot(aes(x=perm, fill=significant)) +
   geom_bar() +
-  #geom_text(stat='count', aes(label=..count..)) +
   facet_wrap(~dataset) +
   theme(axis.text.x = element_blank()) +
-  scale_fill_manual(values = c("lightblue", "red"))
+  scale_fill_manual(values = c("#EA6F5F", "#F3AEA6", "#5FDAEA"))
 ```
@@
                                p.val_fdr >= 0.05 ~ ""),
+         sig = case_when(t.test.p < 0.001 ~ "***",
+                               t.test.p >= 0.001 & t.test.p < 0.01 ~ "**",
+                               t.test.p >= 0.01 & t.test.p < 0.05 ~ "*",
+                               t.test.p >= 0.05 ~ ""),
          y_pos = case_when(dataset=="cf" ~ 150,
@@
   theme_linedraw() +
-  geom_text(aes(label = sig_fdr, x= x_pos, y=y_pos),
+  geom_text(aes(label = sig, x= x_pos, y=y_pos),
               data = next.t.test.z.sex.test.med) +

### 2025-01-17 paper_main_results.Rmd
droping fdr correction for sex-bias tests

@@
     dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.val_fdr),
-              max_pval = max(p.val_fdr)) %>%
+              min_pval = min(p.value),
+              max_pval = max(p.value)) %>%
     ungroup() %>%
@@
 ```{r}
+sex_t_tests_in_feat.df <- sex_t_tests_in_feat.df %>%
+  mutate(sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
+
 sex_t_tests_in_feat.df %>%
   dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+  mutate(median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff*100)) %>%
@@
 sex_t_tests_in_feat.df %>%
-  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+  mutate(median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff*100)) %>%
@@
   # dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
-  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+  mutate(median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff*100),
@@
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = ifelse(sig_fdr == TRUE, median_sex_diff*100, NA))
+         median_sex_diff = ifelse(sig_uncor == TRUE, median_sex_diff*100, NA))
 #find range
@@
                                             TRUE ~ NA)),
-        median_sex_diff = ifelse(sig_fdr == TRUE, median_sex_diff*100, NA)) %>% 
+        median_sex_diff = ifelse(sig_uncor == TRUE, median_sex_diff*100, NA)) %>% 
   group_by(dataset)
@@
 
+noext.sex_t_tests_in_feat.df <- noext.sex_t_tests_in_feat.df %>%
+  mutate(sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
+
 #compare summaries
@@
     group_by(dataset) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
+    dplyr::summarize(n_sig_regions = sum(sig_uncor),
               min_pval = min(p.val_fdr),
@@
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+         median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff*100)) %>%
@@
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+         median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff*100))
@@
 
+z.sex_t_tests_in_feat.df <- z.sex_t_tests_in_feat.df %>%
+  mutate(sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
+
 #also get summary
@@
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = ifelse(sig_fdr == TRUE, median_sex_diff, NA))
+         median_sex_diff = ifelse(sig_uncor == TRUE, median_sex_diff, NA))
 
@@
                                             TRUE ~ NA)),
-        median_sex_diff = ifelse(sig_fdr == TRUE, median_sex_diff, NA)) %>% 
+        median_sex_diff = ifelse(sig_uncor == TRUE, median_sex_diff, NA)) %>% 
   group_by(dataset)
@@
 noext.z.sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_z_sexbias_all_noext_out.RDS")
+noext.z.sex_t_tests_in_feat.df <- noext.z.sex_t_tests_in_feat.df %>%
+  mutate(sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
 
@@
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+         median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff))
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-score Error", limits=c(-0.5, 0.5)) #
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-score Error", limits=c(-0.4, 0.4)) #
 ```
@@
   mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
-         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+         median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff)) %>%
@@
   theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
-  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-score Error", limits = c(-0.5, 0.5)) #,
+  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-score Error", limits = c(-0.4, 0.4)) #,
 # sex.var.plot.global

### 2025-01-18 paper_main_results.Rmd
finalize sex bias analyses w/o FDR correction

@@
 sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_cent_sexbias_all_out.RDS")
-#also get summary
-sex_t_tests_in_feat.df %>%
-    group_by(dataset) %>%
-    dplyr::summarize(n_sig_regions = sum(sig_fdr),
-              min_pval = min(p.value),
-              max_pval = max(p.value)) %>%
-    ungroup() %>%
-    as.data.frame()
-```
-
-quick plot for median diffs
-```{r}
+
 sex_t_tests_in_feat.df <- sex_t_tests_in_feat.df %>%
@@
 
+#also get summary
+sex_t_tests_in_feat.df %>%
+    group_by(dataset) %>%
+    dplyr::summarize(n_sig = sum(sig_uncor),
+      n_sig_regions_fdr = sum(sig_fdr),
+              min_pval = min(p.value),
+              max_pval = max(p.value)) %>%
+    ungroup() %>%
+    as.data.frame()
+```
+
+quick plot for median diffs
+```{r}
 sex_t_tests_in_feat.df %>%

### 2025-01-18 paper_replication_results.Rmd
finalize sex bias analyses w/o FDR correction

@@
          "Significance" = sig,
-         "Significance (FDR-corrected)" = sig_fdr) %>%
+         "(FDR-corrected)" = sig_fdr) %>%
   gt() %>%
@@
   geom_hline(yintercept=0, linetype="dashed")+
-  geom_text(aes(label = sig_fdr, y = (y_pos)), nudge_y=.005,
+  geom_text(aes(label = sig, y = (y_pos)), nudge_y=.005,
                data = all_t_test.df[which(all_t_test.df$dataset != "cf"),]) +
@@
   geom_hline(yintercept=0, linetype="dashed")+
-  geom_text(aes(label = sig_fdr, y = (y_pos)), nudge_y=.005,
+  geom_text(aes(label = sig, y = (y_pos)), nudge_y=.005,
                data = all_t_test.df) +
@@
   labs(y="Female Bias", x = "Dataset")
-ggsave(filename="figs/all_cent_sum_boxplt_wtest.png", all_cent_sum_boxplt.diff_full)
+ggsave(filename="figs/all_cent_sum_boxplt_wtest.png", all_cent_sum_boxplt.diff_full, height=9, width=4.5)
 ```
@@
   labs(y="Proportion Female", x = "Dataset")
-ggsave("figs/z_sex_boxplts.jpeg", z_sex_boxplts)
+ggsave("figs/z_sex_boxplts.jpeg", z_sex_boxplts, height=8, width=5)
 ```

### 2025-01-22 paper_prop_results.Rmd

@@
 ```{r}
-aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
+remotes::install_github("BGDlab/ggsegTissue")
+library(ggsegTissue)
 #write out new dk atlas w/o medial wall
@@
 cent.feat.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_cent_sex_bias_tests.RDS"))
+
 cent.feat.sex.test.df <- bind_rows(cent.feat.sex.test.list) %>%
@@
       pheno %in% ct_list ~ "Regional CT",
-      TRUE ~ NA_character_)))
+      TRUE ~ NA_character_)),
+      sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
 ```
@@
   ggplot() +
-  geom_bar(aes(x=prop, fill=sig_fdr)) +
+  geom_bar(aes(x=prop, fill=significant)) +
   facet_wrap(~dataset) +
@@
 sig.sex_diff.violin <- cent.feat.sex.test.df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
+  dplyr::filter(sig_uncor == TRUE) %>%
   mutate(median_sex_diff = median_sex_diff*100) %>%
@@
 stats_sex.centile_for_plot <- inner_join(cent.sex_comp_tbl2, cent.feat.test_global)
+table(stats_sex.centile_for_plot$significant)
+
 #kinda hacky way of fixing positions
 stats_sex.centile_for_plot.corrected <- stats_sex.centile_for_plot %>%
-  mutate(p.adj.signif = ifelse(sig_fdr == "FALSE", "ns", "*"),
+  mutate(p.signif = ifelse(sig_uncor == "FALSE", "ns", "*"),
          y.position=case_when(dataset=="cf.gam" ~ max_cf.gam*100+0.5,
@@
   labs(fill="Sex", color="Sex")+
-  stat_pvalue_manual(stats_sex.centile_for_plot.corrected, label = "p.adj.signif", label.size = 2, tip.length = 0) +
+  stat_pvalue_manual(stats_sex.centile_for_plot.corrected, label = "p.signif", label.size = 2, tip.length = 0) +
   facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names), scales="free_x")
@@
 t.test.cent.sex.diff <- cent.feat.sex.test.df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
+  #dplyr::filter(sig_fdr == TRUE) %>%
   group_by(dataset, prop) %>%
@@
             max_val = max(median_sex_diff)*100) %>% #test median of each distribution against 0, then scale to centiles
-  mutate(p.val_fdr = p.adjust(w.test.p, method="fdr", n = (4*11)),
-         y_pos = max_val + 0.5) %>%
+  mutate(p.val_fdr = p.adjust(w.test.p, method="fdr", n = (11)),
+         y_pos = max_val + 0.25) %>%
   add_significance("p.val_fdr", cutpoints = c(0, 0.001, 0.01, 0.05, 1),
+  symbols = c("***", "**", "*", "ns")) %>%
+  add_significance("w.test.p", cutpoints = c(0, 0.001, 0.01, 0.05, 1),
   symbols = c("***", "**", "*", "ns"))
@@
 
-cent_sex_diff.plt2 <- cent.feat.sex.test.df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
+#w/ fdr correction
+cent.feat.sex.test.df %>%
+  #dplyr::filter(sig_uncor == TRUE) %>%
   mutate(median_sex_diff = median_sex_diff*100) %>%
@@
   scale_x_discrete(labels=percent_f)
+
+cent_sex_diff.plt2 <- cent.feat.sex.test.df %>%
+  #dplyr::filter(sig_uncor == TRUE) %>%
+  mutate(median_sex_diff = median_sex_diff*100) %>%
+  ggplot(aes(x=prop, y=median_sex_diff)) +
+  geom_boxplot(aes(color=dataset)) +
+  geom_jitter(size=0.4, alpha=0.1) +
+  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), scales="free_y") + 
+  geom_text(aes(label = w.test.p.signif, y = y_pos),
+              data = t.test.cent.sex.diff) +
+  theme_linedraw() +
+  scale_color_manual(values=c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) +
+  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(y="Median Male - Female Centile Error", x = "% Females in Imbalanced Site") +
+  scale_x_discrete(labels=percent_f)
 cent_sex_diff.plt2
@@
 ggsave("figs/cent_sex_diff.plt2.jpeg", cent_sex_diff.plt2)
+sfig13 <- ggarrange(sex_err_ratio.plt, cent_sex_diff.plt2,
+          nrow=2,
+          heights = c(.7, 1),
+          labels = "AUTO")
+ggsave("figs/sfig13.jpeg", sfig13, height=9, width=7)
 ```
@@
 noext_cent.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
-noext_cent.feat.sex.test.df <- bind_rows(noext_cent.subj.sex.test.list)
+noext_cent.feat.sex.test.df <- bind_rows(noext_cent.subj.sex.test.list) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
 ```
@@
 noext_cent.feat.sex.test.df %>%
-  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+  mutate(median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff*100)) %>%
@@
                                       p.val_fdr >= 0.05 ~ FALSE),
-         sig_fdr = case_when(
-           # p.val_fdr < 0.001 ~ "***",
-           #                     p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
-           #                     p.val_fdr >= 0.01 & 
-                               p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ ""),
+         sig_uncor = case_when(t.test.p < 0.05 ~ "*",
+                               t.test.p >= 0.05 ~ ""),
          y_pos = ifelse(max_val < 0, ((max_val*100)*0.5), (max_val*100)+0.8))
@@
     facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names)) + 
-  geom_text(aes(label = sig_fdr, y = y_pos),
+  geom_text(aes(label = sig_uncor, y = y_pos),
               data = noext_t.test.cent.sex.diff[which(noext_t.test.cent.sex.diff$dataset != "cf"),]) +
@@
 z.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
-z.feat.sex.test.df <- bind_rows(z.sex.test.list)
+z.feat.sex.test.df <- bind_rows(z.sex.test.list) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
 ```
@@
 z.sig.sex_diff.violin <- z.feat.sex.test.df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
+  dplyr::filter(sig_uncor == TRUE) %>%
   ggplot() +
@@
     TRUE ~ NA_character_)))
+
+z.feat.sex.test.df %>%
+  mutate(median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
+                              TRUE ~ median_sex_diff)) %>%
+  ggplot() +
+  geom_histogram(aes(x=median_sex_diff, fill= pheno_cat, color=pheno_cat), alpha=0.4, position = "identity") +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  facet_grid(prop~dataset, scales="free")
 
@@
   mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = (3*4*11)), #not testing global vols, too few data points
-                  sig_fdr_tf = case_when(p.val_fdr < 0.05 ~ TRUE,
-                                      p.val_fdr >= 0.05 ~ FALSE),
-         sig_fdr = case_when(p.val_fdr < 0.05 ~ "*",
+                  # sig_fdr_tf = case_when(p.val_fdr < 0.05 ~ TRUE,
+                  #                     p.val_fdr >= 0.05 ~ FALSE),
+         sig_uncor = case_when(p.val_fdr < 0.05 ~ "*",
                             p.val_fdr >= 0.05 ~ ""),
@@
     facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names)) + 
-  geom_text(aes(label = sig_fdr, y = y_pos),
+  geom_text(aes(label = sig_uncor, y = y_pos),
               data = t.test.z.sex.diff[which(t.test.z.sex.diff$dataset != "cf"),]) +
@@
 z.t.test.cent.sex.diff <- z.feat.sex.test.df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
+  dplyr::filter(sig_uncor == TRUE) %>%
   group_by(dataset, prop) %>%
@@
             max_val = max(median_sex_diff)) %>% #test median of each distribution against 0
-  mutate(p.val_fdr = p.adjust(w.test.p, method="fdr", n = (4*11)), 
-         y_pos = max_val + 3) %>%
-  add_significance("p.val_fdr", cutpoints = c(0, 0.001, 0.01, 0.05, 1),
+  mutate(p.val_fdr = p.adjust(w.test.p, method="fdr", n = (11)), 
+         y_pos = max_val*1.3) %>%
+  add_significance("w.test.p", cutpoints = c(0, 0.001, 0.01, 0.05, 1),
   symbols = c("***", "**", "*", "ns"))
@@
 z.cent_sex_diff.plt <- z.feat.sex.test.df %>%
-  dplyr::filter(sig_fdr == TRUE) %>%
-  mutate(median_sex_diff = median_sex_diff*100) %>%
+  dplyr::filter(sig_uncor == TRUE) %>%
   ggplot(aes(x=prop, y=median_sex_diff)) +
@@
   geom_jitter(size=0.4, alpha=0.1) +
-  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names)) + 
-  geom_text(aes(label = p.val_fdr.signif, y = y_pos),
+  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), scales="free_y") + 
+  geom_text(aes(label = w.test.p.signif, y = y_pos),
               data = z.t.test.cent.sex.diff) +
@@
 noext_z.feat.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
-noext_z.feat.sex.test.df <- bind_rows(noext_z.feat.sex.test.list)
+noext_z.feat.sex.test.df <- bind_rows(noext_z.feat.sex.test.list) %>%
+  mutate(pheno_cat = as.factor(case_when(
+      pheno %in% vol_list_global ~ "Global Volume",
+      pheno %in% vol_list_regions ~ "Regional Volume",
+      pheno %in% sa_list ~ "Regional SA",
+      pheno %in% ct_list ~ "Regional CT",
+      TRUE ~ NA_character_)),
+      sig_uncor = ifelse(p.value<0.05, TRUE, FALSE)) %>%
+  mutate(significant = case_when(sig_fdr == TRUE ~ "Survives FDR Correction",
+                                 sig_uncor == TRUE & sig_fdr == FALSE ~ "Uncorrected Only",
+                                 sig_uncor == FALSE & sig_fdr == FALSE ~ "Not Significant")) %>%
+  mutate(significant = factor(significant, ordered=TRUE, levels=c("Survives FDR Correction", "Uncorrected Only", "Not Significant")))
 ```
@@
 noext_z.feat.sex.test.df %>%
-  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
+  mutate(median_sex_diff = case_when(sig_uncor == FALSE ~ NA,
                               TRUE ~ median_sex_diff)) %>%
@@
                                       p.val_fdr >= 0.05 ~ FALSE),
-         sig_fdr = case_when(
+         sig_uncor = case_when(
            # p.val_fdr < 0.001 ~ "***",
@@
            #                     p.val_fdr >= 0.01 & 
-                               p.val_fdr < 0.05 ~ "*",
-                               p.val_fdr >= 0.05 ~ ""),
+                               t.test.p < 0.05 ~ "*",
+                               t.test.p >= 0.05 ~ ""),
          y_pos = (max_val+0.01))
@@
     facet_grid(dataset~pheno_cat, labeller=labeller(dataset=dataset_names)) + 
-  geom_text(aes(label = sig_fdr, y = y_pos),
+  geom_text(aes(label = sig_uncor, y = y_pos),
               data = noext_t.test.z.sex.diff[which(noext_t.test.z.sex.diff$dataset != "cf"),]) +

### 2025-01-23 eda/lifespan_eda.Rmd
remaking LBCC plots to include country

@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see, broom.mixed)
+pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see, broom.mixed, ggrain, ggpubr)
 
@@
 ```{r}
-cn.df <- fread(paste0(base_path,"/data/v1_2_1_nature.csv"), stringsAsFactors = TRUE)
+cn.df <- fread(paste0(base_path,"/data/v1_2_1_nature.csv"), stringsAsFactors = TRUE, na.strings="")
 ```
@@
 cn.df.imp1 <- cn.df.eulerfilt2 %>%
-  mutate(site = trimws(as.character(site)),
-         study = trimws(study)) %>%
-  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_")))) 
+  mutate(site = gsub(" ", "_", trimws(as.character(site))),
+         study = gsub(" ", "_", trimws(as.character(study)))) %>%
+  mutate(site = ifelse(is.na(site), study, paste(study, site, sep="_"))) 
 skim(cn.df.imp1)
@@
 cn.df.imp <- cn.df.imp1%>%
-  dplyr::select(!c(INDEX.ID, country)) %>%#drop index ID for filtering
+  dplyr::select(!c(INDEX.ID)) %>%#drop index ID for filtering
   na.omit()
@@
 
-Demographics
+##Demographics
 
@@
 study_site_key <- cn.df.imp %>%
-  dplyr::select(study, site) %>%
+  dplyr::select(study, site, country, fs_version) %>%
   unique()
@@
 ```
+
+###Plot
+
+Clean up names, order by age
+```{r}
+#reorder studies by age
+age.df <- cn.df.imp %>%
+  group_by(study) %>%
+  summarize(mean_age = mean(age_days)) %>%
+  arrange(mean_age)
+study_names_clean <- gsub("_", " ", unique(age.df$study))
+
+final_df <- cn.df.imp %>% 
+  mutate(age_years=(age_days-280)/365.25,
+         study = gsub("_", " ", study)) %>%
+  mutate(study = factor(study, levels=study_names_clean, ordered=TRUE))
+```
+
+Sample plot
+```{r}
+sample_boxplot <- final_df %>%
+  #mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
+ggplot(aes(y=study, x=age_years)) +
+  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
+  theme_linedraw(base_size = 11) +
+  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text.x=element_blank()) +
+  labs(color="Sex", y = "Study")
+
+#overall sample
+full_boxplot <- final_df %>%
+  mutate(ylab = "   Full Sample") %>%
+  ggplot(aes(y=ylab, x=age_years, color=sex)) +
+  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
+  theme_linedraw(base_size = 11) +
+  # theme(plot.margin = margin(0,0,0,0, 'cm')) +
+  # scale_y_discrete(labels=c("1" = "Full Sample")) +
+  labs(color="Sex", x = "Age (Years)", y="")
+
+sample_barplot <- final_df %>%
+  #mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
+  ggplot() +
+  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=study, fill=sex)) +
+  scale_x_continuous(labels = scales::percent) +
+  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
+  theme_linedraw(base_size = 11) +
+  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
+  labs(fill="Sex")
+
+full_barplot <- final_df %>%
+  mutate(ylab = "Full Sample") %>%
+  ggplot() +
+  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=ylab, fill=sex)) + 
+  scale_x_continuous(labels = scales::percent) +
+  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
+  theme_linedraw(base_size = 11) +
+  theme(axis.ticks.y = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) +
+  labs(fill="Sex", x = "Sex")
+
+ggarrange(sample_boxplot, sample_barplot,
+          full_boxplot, full_barplot,
+          nrow=2, ncol=2, 
+          align = "hv", 
+          common.legend = TRUE,
+          widths=c(2, 1),
+          heights=c(10, 1))
+```
+
+look at country info
+
+```{r}
+sample_boxplot2 <- final_df %>%
+  ggplot(aes(y=age_years, x=study, fill=country)) +
+  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
+  theme_linedraw(base_size = 11) +
+  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
+  theme(axis.title.x=element_blank(), axis.text.x = element_blank(), legend.position = "right") +
+  labs(fill="Country", y = "Age (Years)")  +
+  paletteer::scale_fill_paletteer_d("ggthemes::Tableau_20")
+```
+
+```{r}
+blank <- ggplot() + theme_void()
+
+#overall sample
+full_boxplot2 <- final_df %>%
+  mutate(xlab = "Full Sample ") %>%
+  ggplot(aes(x=xlab, y=age_years, color=sex)) +
+  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
+  theme_linedraw(base_size = 11) +
+  theme(axis.title=element_blank(), axis.text=element_blank(), axis.ticks.y=element_blank()) +
+  labs(color="Sex", x=" ")
+
+sample_barplot2 <- final_df %>%
+  ggplot() +
+  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
+  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
+  theme_linedraw(base_size = 11) +
+  scale_y_continuous(labels = scales::percent) +
+  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=6)) +
+  labs(fill="Sex", x = "Study", y="Percent Sex")
+
+full_barplot2 <- 
+  final_df %>%
+  mutate(xlab = "  Full Sample  ") %>%
+  ggplot() +
+  geom_bar(aes(y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=xlab, fill=sex)) + 
+  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
+  theme_linedraw(base_size = 11) +
+  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
+  labs(color="Sex", y = "Age (Years)", y="")
+
+full_rain <- final_df %>%
+  ggplot(aes(x=sex, y=age_years, fill=sex)) +
+  geom_rain(alpha=.4) +
+  theme_linedraw(base_size = 11) +
+  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), 
+        axis.title.x = element_blank(), axis.text.y = element_text(size=6),
+        legend.position = "none") +
+  labs(y="Age (Years)")
+```
+
+```{r}
+country_leg <- as_ggplot(get_legend(sample_boxplot2))
+sex_leg <- as_ggplot(get_legend(sample_barplot2))
+
+fig5A <- ggarrange(sample_boxplot2, country_leg,
+          sample_barplot2,  sex_leg,
+          nrow=2, ncol=2, 
+          #align = "v", 
+          common.legend = FALSE,
+          legend = "none",
+          widths=c(11, 1),
+          heights=c(1.9, 1)
+          )
+```
+
+```{r}
+fig5_AB <- ggarrange(fig5A, full_rain,
+          nrow=1, ncol=2, 
+          legend = "none",
+          widths=c(3, 1),
+          labels = "AUTO")
+
+saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
+```

### 2025-01-23 paper_lbcc_results.Rmd
remaking LBCC plots to include country

@@
 
-gam_centile_df <- full_join(gam_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)
+gam_centile_df <- left_join(gam_centile_df, study_site_key) %>% 
+  mutate(age_years=(age_days-280)/365.25)
 n_unique(gam_centile_df$study)
@@
 dim(gam_centile_df)
+
+gam_centile_df %>% dplyr::filter(is.na(age_days))
 ```
@@
 gamlss_centile_df <- full_join(gamlss_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)
+    mutate(age_years=(age_days-280)/365.25)
 
@@
 raw_centile_df <- full_join(raw_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)
+    mutate(age_years=(age_days-280)/365.25)
 
@@
 deltas_df <- full_join(deltas_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)
+    mutate(age_years=(age_days-280)/365.25)
 
@@
 n_unique(gamlss_centile_df$site) #correct
-```
-
-Sample plot
-```{r}
-#reorder studies by age
-age.df <- gamlss_centile_df %>%
-  group_by(study) %>%
-  summarize(mean_age = mean(age_years)) %>%
-  arrange(mean_age)
-
-sample_boxplot <- gamlss_centile_df %>%
-  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
-ggplot(aes(y=study, x=age_years)) +
-  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
-  theme_linedraw(base_size = 11) +
-  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
-  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text.x=element_blank()) +
-  labs(color="Sex", y = "Study")
-
-#overall sample
-full_boxplot <- 
-  gamlss_centile_df %>%
-  mutate(ylab = "   Full Sample") %>%
-  ggplot(aes(y=ylab, x=age_years, color=sex)) +
-  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
-  theme_linedraw(base_size = 11) +
-  # theme(plot.margin = margin(0,0,0,0, 'cm')) +
-  # scale_y_discrete(labels=c("1" = "Full Sample")) +
-  labs(color="Sex", x = "Age (Years)", y="")
-
-sample_barplot <- gamlss_centile_df %>%
-  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
-  ggplot() +
-  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=study, fill=sex)) +
-  scale_x_continuous(labels = scales::percent) +
-  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
-  theme_linedraw(base_size = 11) +
-  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
-  labs(fill="Sex")
-
-full_barplot <- 
-  gamlss_centile_df %>%
-  mutate(ylab = "Full Sample") %>%
-  ggplot() +
-  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=ylab, fill=sex)) + 
-  scale_x_continuous(labels = scales::percent) +
-  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
-  theme_linedraw(base_size = 11) +
-  theme(axis.ticks.y = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) +
-  labs(fill="Sex", x = "Sex")
-
-ggarrange(sample_boxplot, sample_barplot,
-          full_boxplot, full_barplot,
-          nrow=2, ncol=2, 
-          align = "hv", 
-          common.legend = TRUE,
-          widths=c(2, 1),
-          heights=c(10, 1))
 ```
@@
 ```{r}
-blank <- ggplot() + theme_void()
-
-sample_boxplot2 <- gamlss_centile_df %>%
-  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
-ggplot(aes(y=age_years, x=study, color=sex)) +
-  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
-  theme_linedraw(base_size = 11) +
-  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
-  theme(axis.title.x=element_blank(), axis.text.x = element_blank()) +
-  labs(color="Sex", y = "Age\n(Years Post-Conception)")
-
-#overall sample
-full_boxplot2 <- 
-  gamlss_centile_df %>%
-  mutate(xlab = "Full Sample ") %>%
-  ggplot(aes(x=xlab, y=age_years, color=sex)) +
-  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
-  theme_linedraw(base_size = 11) +
-  theme(axis.title=element_blank(), axis.text=element_blank(), axis.ticks.y=element_blank()) +
-  # scale_y_discrete(labels=c("1" = "Full Sample")) +
-  labs(color="Sex", x=" ")
-
-sample_barplot2 <- gamlss_centile_df %>%
-  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
-  ggplot() +
-  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
-  #scale_y_continuous(labels = scales::percent) +
-  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
-  theme_linedraw(base_size = 11) +
-  scale_y_continuous(labels = scales::percent) +
-  #theme(axis.ticks.x = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
-  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=6)) +
-  labs(fill="Sex", x = "Study", y="Percent Sex")
-
-full_barplot2 <- 
-  gamlss_centile_df %>%
-  mutate(xlab = "  Full Sample  ") %>%
-  ggplot() +
-  geom_bar(aes(y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=xlab, fill=sex)) + 
-  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
-  theme_linedraw(base_size = 11) +
-  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
-  labs(color="Sex", y = "Age\n(Years Post-Conception)", y="")
-```
-
-```{r}
 # study demographics

### 2025-01-25 eda/lifespan_eda.Rmd
filter phenos < or > 20 MADs from median

@@
 Load data from Jakob - updated with euler numbers and corrected VETSA study
+
 ```{r}
 cn.df <- fread(paste0(base_path,"/data/v1_2_1_nature.csv"), stringsAsFactors = TRUE, na.strings="")
+pheno_list <- readRDS(paste0(base_path, "/cubic_scripts/R_scripts/pheno_list.rds"))
 ```
@@
   geom_boxplot(aes(x=study, y=surfholes)) +
-  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_mad)), color="red")
+  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_mad)), color="red") +
+  theme(axis.text.x = element_text(angle=90, hjust=0.5))
 
@@
 ```{r}
+#make sure index.id is ok for filtering
+stopifnot(sum(is.na(cn.df.imp1$INDEX.ID))==0)
+
 #drop NAs listwise
 cn.df.imp <- cn.df.imp1%>%
-  dplyr::select(!c(INDEX.ID)) %>%#drop index ID for filtering
   na.omit()
@@
 
+Finally, drop 0 phenos or biologically implausible
+
+```{r}
+#find 0 IDPs
+df_zeros <- cn.df.imp %>%
+  filter(if_any(any_of(pheno_list), ~ . == 0 & !is.na(.))) %>%
+  filter(if_any(all_of(pheno_list), ~ !is.na(.)))
+nrow(df_zeros)
+```
+
+no one with IDPs=0
+
+check for implausible values
+
+```{r}
+#check there aren't currently any NAs
+stopifnot(sum(is.na(cn.df.imp))==0)
+
+cn.df_mad <- lapply(cn.df.imp[c("INDEX.ID", pheno_list)], mad_filt, n=20) %>% as.data.frame()
+sum(is.na(cn.df_mad))
+
+mad_drop <- filter(cn.df_mad, if_any(pheno_list, is.na))
+
+#remove 3 subjects
+cn.df_mad_filt <- cn.df.imp %>%
+  dplyr::filter(! INDEX.ID %in% mad_drop$INDEX.ID)
+
+nrow(cn.df.imp) == (nrow(cn.df_mad_filt) + nrow(mad_drop))
+```
+
 ##Demographics
@@
 #sample size, sex
-length(unique(cn.df.imp$participant))
-nrow(cn.df.imp) == length(unique(cn.df.imp$participant))
+length(unique(cn.df_mad_filt$participant))
+nrow(cn.df_mad_filt) == length(unique(cn.df_mad_filt$participant))
 
-table(cn.df.imp$sex)
+table(cn.df_mad_filt$sex)
 
-table(cn.df.imp$sex)["Female"]/nrow(cn.df.imp)
+table(cn.df_mad_filt$sex)["Female"]/nrow(cn.df_mad_filt)
 
 #studies, sites
-length(unique(cn.df.imp$study))
-length(unique(cn.df.imp$site))
+length(unique(cn.df_mad_filt$study))
+length(unique(cn.df_mad_filt$site))
 
 #age
-min(cn.df.imp$age_days)/365.25-(280/365.25)
-max(cn.df.imp$age_days)/365.25-(280/365.25)
-mean(cn.df.imp$age_days)/365.25-(280/365.25)
+min(cn.df_mad_filt$age_days)/365.25-(280/365.25)
+max(cn.df_mad_filt$age_days)/365.25-(280/365.25)
+mean(cn.df_mad_filt$age_days)/365.25-(280/365.25)
 
-abcd <- cn.df.imp %>%
+abcd <- cn.df_mad_filt %>%
   dplyr::filter(study=="ABCD")
@@
 
-```{r}
-ggplot(cn.df.imp) +
-  geom_violin(aes(x=reorder(study, age_days), y=age_days), scale="width") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-ggplot(cn.df.imp) +
-  geom_violin(aes(x=reorder(site, age_days), y=age_days, fill=study), scale="width") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
-```
-
-check for sex differences in sites
-
 Write out
@@
 ```{r}
-fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler.csv"))
+fwrite(cn.df_mad_filt, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler.csv"))
 
 #save names with site (since not saved correctly from centile outputs)
-study_site_key <- cn.df.imp %>%
+study_site_key <- cn.df_mad_filt %>%
   dplyr::select(study, site, country, fs_version) %>%
   unique()
-n_unique(study_site_key$study) == n_unique(cn.df.imp$study)
-n_unique(study_site_key$site) == n_unique(cn.df.imp$site)
+n_unique(study_site_key$study) == n_unique(cn.df_mad_filt$study)
+n_unique(study_site_key$site) == n_unique(cn.df_mad_filt$site)
 fwrite(study_site_key, file=paste0(base_path, "/data/lifespan_study_site_key.csv"))
@@
 #reorder studies by age
-age.df <- cn.df.imp %>%
+age.df <- cn.df_mad_filt %>%
   group_by(study) %>%
@@
 
-final_df <- cn.df.imp %>% 
+final_df <- cn.df_mad_filt %>% 
   mutate(age_years=(age_days-280)/365.25,
@@
   geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
-  theme_linedraw(base_size = 11) +
+  theme_linedraw(base_size = 14) +
   #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
@@
   geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
-  theme_linedraw(base_size = 11) +
+  theme_linedraw(base_size = 14) +
   theme(axis.title=element_blank(), axis.text=element_blank(), axis.ticks.y=element_blank()) +
@@
   geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
-  theme_linedraw(base_size = 11) +
+  theme_linedraw(base_size = 14) +
   scale_y_continuous(labels = scales::percent) +
-  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=6)) +
+  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=10)) +
   labs(fill="Sex", x = "Study", y="Percent Sex")
@@
   geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
-  theme_linedraw(base_size = 11) +
-  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
+  theme_linedraw(base_size = 14) +
+  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5)) +
   labs(color="Sex", y = "Age (Years)", y="")
 
+```
+
+```{r}
 full_rain <- final_df %>%
-  ggplot(aes(x=sex, y=age_years, fill=sex)) +
-  geom_rain(alpha=.4) +
-  theme_linedraw(base_size = 11) +
-  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), 
-        axis.title.x = element_blank(), axis.text.y = element_text(size=6),
-        legend.position = "none") +
+  mutate(continent = factor(case_when(country == "India" | country == "China" | country == "Singapore" | country == "Japan" ~ "Asia",
+                               country == "USA" | country == "Canada" ~ "North America",
+                               country == "Brazil" ~ "South America",
+                               country == "Australia" ~ "Australia",
+                               TRUE ~ "Europe"))) %>%
+  ggplot(aes(x=continent, y=age_years, fill=sex)) +
+  geom_rain(alpha=.5, violin.args = list(scale="count", alpha=.5),
+            point.args=list(alpha=.2)) +
+  theme_linedraw(base_size = 14) +
+  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        legend.position = "bottom") +
+  guides(fill=guide_legend(title="Sex")) + 
+  coord_fixed(ratio=.08) +
   labs(y="Age (Years)")
@@
           legend = "none",
-          widths=c(11, 1),
+          widths=c(8, 1),
           heights=c(1.9, 1)
@@
           nrow=1, ncol=2, 
-          legend = "none",
-          widths=c(3, 1),
+          #legend = "none",
+          widths=c(1.8, 1),
           labels = "AUTO")
-
+fig5_AB
 saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")

### 2025-01-25 cubic_scripts/R_scripts/gamlss_helper_funs.R

@@ -765,5 +765,30 @@ cohens_f2_local <- function(full_mod, null_mod){
   fsq <- (full_rsq - null_rsq)/(1-full_rsq)
   return(fsq)
 }
+
+############
+#mad_filt(): filters values from df col x that are > or < thresholds defined by median(x) +- n*mad(x).
+###basically just slight modification of hampel()from jvandoorn/makilab with slight modifications
+# e.g. sapply(iris, mad_filt)
+mad_filt <- function(x, n = 3, flag=NA) {
+  y <- x # Output vector
+  
+  if(is.numeric(x)){
+  
+  m <- median(x, na.rm = TRUE)
+  md <- mad(x, na.rm = TRUE)
+  
+  x_ind <- which(x >= m + n * md | x <= m - n * md)
+
+  y[x_ind] <- flag
+  
+  return(y)
+  
+  } else {
+    warning("skipping non-numeric vectors")
+    return(x)
+  }
+
+}
   
 

### 2025-01-27 eda/lifespan_eda.Rmd
cleaning up data filtering and curation

@@
 
+#Visualize
 Plot full samples for supplemental figures looking at imbalanced demographics across sites
@@
 
+## Curate
+
+Remove infants (anyone younger than 3 yrs)
+
 ```{r}
+cn.df.noinf <- cn.df %>%
+  dplyr::filter(age_days >= (365.25*3+280)) %>%
+  mutate(study = droplevels(study),
+         site = droplevels(site))
+nrow(cn.df) - nrow(cn.df.noinf)
+n_unique(cn.df$study) - n_unique(cn.df.noinf$study)
+min(cn.df.noinf$age_days)
+```
+
+```{r}
+set.seed(1212)
 #drop repeat scans
-cn.df.unique <- cn.df %>%
-  drop_na(c(age_days, participant, study)) %>%
+cn.df.unique <- cn.df.noinf %>%
+  drop_na(c(age_days, INDEX.ID, study)) %>%
   arrange(age_days) %>%
-  group_by(participant) %>%
-  slice(1) %>%
+  group_by(INDEX.ID) %>%
+  slice_sample(n=1) %>%
   ungroup() %>%
@@
 
-nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
-nrow(cn.df) - nrow(cn.df.unique)
+nrow(cn.df.unique) == n_unique(cn.df.unique$INDEX.ID)
+nrow(cn.df.noinf) - nrow(cn.df.unique) #n scans
+n_unique(cn.df.noinf$INDEX.ID[duplicated(cn.df.noinf$INDEX.ID)]) #n subjects
 ```
@@
 
-## Visualize Features
 plot global brain features for basic qc - check again post-combat
@@
 
-## Curate
-
-Remove infants (anyone younger than 3 yrs)
-
-```{r}
-cn.df.noinf <- cn.df.unique %>%
-  dplyr::filter(age_days >= (365.25*3+280)) %>%
-  mutate(study = droplevels(study),
-         site = droplevels(site))
-nrow(cn.df.unique) - nrow(cn.df.noinf)
-n_unique(cn.df.unique$study) - n_unique(cn.df.noinf$study)
-min(cn.df.noinf$age_days)
-```
-
-Visualizing euler distributions
-```{r}
-cn.df.noinf %>%
-  ggplot() +
-  geom_boxplot(aes(x=study, y=surfholes))
-```
-
 Removing subjects w/o surfholes, subjects with surfholes values > 2 MADs above study median (same as sensitivity analyses in LBCC)
@@
 ```{r}
-cn.df.eulerfilt <- cn.df.noinf %>%
+cn.df.eulerfilt <- cn.df.unique %>%
   dplyr::filter(!is.na(surfholes)) #drops 2 studies w missing euler data
@@
 
-Removed `r nrow(cn.df.noinf) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain
+Removed `r nrow(cn.df.unique) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain
 
@@
 #sample size, sex
-length(unique(cn.df_mad_filt$participant))
-nrow(cn.df_mad_filt) == length(unique(cn.df_mad_filt$participant))
+length(unique(cn.df_mad_filt$INDEX.ID))
+stopifnot(nrow(cn.df_mad_filt) == length(unique(cn.df_mad_filt$INDEX.ID)))
 
@@
 
-###Plot
+##Plot
 
@@
 fig5_AB
-saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
+#saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
+ggsave("../figs/fig5_ab.jpeg", fig5_AB, width=18, height=10, units="in")
 ```

### 2025-01-31 eda/lifespan_eda.Rmd
updates

@@
 fig5_AB
-#saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
+saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
 ggsave("../figs/fig5_ab.jpeg", fig5_AB, width=18, height=10, units="in")

### 2025-01-31 paper_lbcc_results.Rmd
updates

@@
     - `R_scripts/get_lifespan_deltas.R`
+
+Requires the following inputs from cluster:
+
+ - `lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv`
+ - `lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv`
+ - `lifespan-CN-imp-sites-euler-raw_predictions.csv`
+ - `batch.study_pred_deltas.csv`
+ - `lifespan-CN-imp-sites-euler-raw_cohenfsq.csv`
+ - `lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv`
+ - `lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_cohenfsq.csv`
 
@@
 ```{r}
-gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv"))
+gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_predictions.csv"))
 
 gam_centile_df <- left_join(gam_centile_df, study_site_key) %>% 
-  mutate(age_years=(age_days-280)/365.25)
+  mutate(age_years=(age_days-280)/365.25) %>%
+  unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE)
 n_unique(gam_centile_df$study)
@@
 dim(gam_centile_df)
-
-gam_centile_df %>% dplyr::filter(is.na(age_days))
 ```
@@
 ```{r}
-gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv"))
+gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_predictions.csv"))
 gamlss_centile_df <- full_join(gamlss_centile_df, study_site_key) %>% 
-    mutate(age_years=(age_days-280)/365.25)
+  mutate(age_years=(age_days-280)/365.25) %>%
+  unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE)
 
@@
 ```{r}
-raw_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_predictions.csv"))
+raw_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_no.tbv_predictions.csv"))
 raw_centile_df <- full_join(raw_centile_df, study_site_key) %>%
+  unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE) %>%
   mutate(age_years=(age_days-280)/365.25)
@@
 ```{r}
-deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
+deltas_df <- fread(file=paste0(der_path, "/batch.study-data_no.tbv_predictions_pred_deltas.csv"))
 deltas_df <- full_join(deltas_df, study_site_key) %>% 
+  unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE) %>%
   mutate(age_years=(age_days-280)/365.25)
@@
   pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(participant) %>%
+  group_by(INDEX.ID) %>%
   summarize(study = unique(study),
@@
             mean_cent=mean(centile))
-nrow(raw_centile_df.mean) == n_unique(raw_centile_df.mean$participant)
+nrow(raw_centile_df.mean) == n_unique(raw_centile_df.mean$INDEX.ID)
 ```
@@
   pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
-  group_by(participant) %>%
+  group_by(INDEX.ID) %>%
   summarize(study = unique(study),
@@
             mean_z=mean(z_score))
-nrow(raw_z_df.mean) == n_unique(raw_z_df.clean$participant)
+nrow(raw_z_df.mean) == n_unique(raw_z_df.clean$INDEX.ID)
 ```
@@
             mean_cent=mean(centile))
-nrow(gam_centile_df.mean) == n_unique(gam_centile_df.mean$participant)
+nrow(gam_centile_df.mean) == n_unique(gam_centile_df.mean$INDEX.ID)
 ```
@@
             mean_z=mean(z_score))
-nrow(gam_z_df.mean) == n_unique(gam_z_df.clean$participant)
+nrow(gam_z_df.mean) == n_unique(gam_z_df.clean$INDEX.ID)
 ```
@@
             mean_cent=mean(centile))
-nrow(gamlss_centile_df.mean) == n_unique(gamlss_centile_df.mean$participant)
+nrow(gamlss_centile_df.mean) == n_unique(gamlss_centile_df.mean$INDEX.ID)
 ```
@@
             mean_z=mean(z_score))
-nrow(gamlss_z_df.mean) == n_unique(gamlss_z_df.clean$participant)
+nrow(gamlss_z_df.mean) == n_unique(gamlss_z_df.clean$INDEX.ID)
 ```
@@
             IQR = IQR(fsq))
+
+length(unique(f2.df$pheno))
+setdiff(pheno_list, unique(f2.df$pheno)) #pheno not loaded
 ```
@@
 f2.df.sum <- f2.df %>%
-  dplyr::filter(combat != "Umharmonized") %>%
+  dplyr::filter(combat != "Unharmonized") %>%
   group_by(pheno) %>%
@@
             study_median_age=median(age_years),
-            study_n= n()) %>%
-  dplyr::select(study, study_mean_age, study_median_age, study_n)
+            study_n= n(),
+            age_range = max(age_years) - min(age_years)) %>%
+  dplyr::select(study, study_mean_age, study_median_age, study_n, age_range)
 ```
@@
   pivot_longer(cols=any_of(paste0("delta.", pheno_list)), names_to="pheno", values_to = "centile") %>%
-  group_by(participant) %>%
+  group_by(INDEX.ID) %>%
   summarize(study = unique(study),
@@
 mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)
+mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
+  mutate(age_diff = (age_days/365.25) - study_mean_age,
+         abs_age_diff = abs(age_diff))
 
@@
 ```{r}
-#rm features that didn't converge in both datasets
-pheno_list_cent <- pheno_list[pheno_list != "lh_SA_posteriorcingulate" & pheno_list != "rh_CT_posteriorcingulate"]
-
 gam_centile_df.clean$combat <- "cf.gam"
@@
   mutate(combat = factor(combat, ordered=TRUE, levels=c("cf.gam", "cf.gamlss"))) %>%
-  mutate(across(pheno_list_cent, ~ as.factor(case_when(. > 0.95 ~ "high", 
+  mutate(across(pheno_list, ~ as.factor(case_when(. > 0.95 ~ "high", 
                                                . < 0.05 ~ "low",
@@
   arrange(combat) %>%
-  group_by(participant) %>%
-  mutate(across(pheno_list_cent, 
+  group_by(INDEX.ID) %>%
+  mutate(across(pheno_list, 
                 ~ . == first(.), .names = "is_equal_{col}")) %>%
   ungroup()
+saveRDS(centiles_cat, file="centiles_cat.rds")
 ```
@@
 centiles_cat_filt <- centiles_cat %>%
-  dplyr::select(participant, combat, all_of(pheno_list_cent))
+  dplyr::select(INDEX.ID, combat, all_of(pheno_list))
 
@@
   rowwise() %>%
-  mutate(count_high = sum(c_across(pheno_list_cent) == "high"),
-         count_mid = sum(c_across(pheno_list_cent) == "mid"),
-         count_low = sum(c_across(pheno_list_cent) == "low")) %>%
-  dplyr::select(participant, combat, starts_with("count"))
+  mutate(count_high = sum(c_across(pheno_list) == "high"),
+         count_mid = sum(c_across(pheno_list) == "mid"),
+         count_low = sum(c_across(pheno_list) == "low")) %>%
+  dplyr::select(INDEX.ID, combat, starts_with("count"))
 
@@
   dplyr::filter(combat =="cf.gamlss") %>%
-  dplyr::select(participant, starts_with("is_equal_")) %>%
+  dplyr::select(INDEX.ID, starts_with("is_equal_")) %>%
   mutate(sum = rowSums(across(starts_with("is_equal_")))) %>%
-  mutate(changed = ifelse(sum == length(pheno_list_cent), "no_change", "change")) #if not all features are true -> changed
+  mutate(changed = ifelse(sum == length(pheno_list), "no_change", "change")) #if not all features are true -> changed
 
-max(centiles_cat_filt2$sum) == length(pheno_list_cent) #correct
+stopifnot(max(centiles_cat_filt2$sum) == length(pheno_list)) #correct
 
-length(pheno_list_cent) - mean(centiles_cat_filt2$sum) #mean per subject
-length(pheno_list_cent) - min(centiles_cat_filt2$sum) #range
-length(pheno_list_cent) - max(centiles_cat_filt2$sum)
+length(pheno_list) - mean(centiles_cat_filt2$sum) #mean per subject
+length(pheno_list) - min(centiles_cat_filt2$sum) #range
+length(pheno_list) - max(centiles_cat_filt2$sum)
 
@@
 #n changed features = # total features - sum
-n_feat <- n_unique(centiles_cat_filt2$participant)*length(pheno_list_cent)
+n_feat <- n_unique(centiles_cat_filt2$INDEX.ID)*length(pheno_list)
 n_unchanged <- sum(centiles_cat_filt2$sum) #true = 1 = no change
@@
 
-n_unique(centiles_cat_filt2_long$participant)*206 == nrow(centiles_cat_filt2_long)
+n_unique(centiles_cat_filt2_long$INDEX.ID)*208 == nrow(centiles_cat_filt2_long)
 
@@
   geom_smooth(method="lm")
+```
+
+similarly, studies with larger age ranges change more
+```{r}
+summary(lm(abs_mean_delta_cent ~ age_range + study_n, data=mean_cent_deltas_df_study))
+
+mean_cent_deltas_df_study %>%
+  ggplot(aes(x=age_range, y=abs_mean_delta_cent, color=study)) +
+  geom_point(alpha=0.05) +
+  geom_smooth(method="lm") +
+  theme(legend.position = "none")
 ```
@@
 ```{r}
+demo_plt <- readRDS("figs/figure_5_AB.rds")
+
 # study demographics
 fig5_vert <- ggarrange(
-  #ggarrange(blank,
-ggarrange(sample_boxplot2, full_boxplot2,
-          sample_barplot2, full_barplot2,
-          nrow=2, ncol=2, 
-          #align = "v", 
-          common.legend = TRUE,
-          legend = "right",
-          widths=c(18, 1),
-          heights=c(1, 1)
-          ),
+#   #ggarrange(blank,
+# ggarrange(sample_boxplot2, full_boxplot2,
+#           sample_barplot2, full_barplot2,
+#           nrow=2, ncol=2, 
+#           #align = "v", 
+#           common.legend = TRUE,
+#           legend = "right",
+#           widths=c(18, 1),
+#           heights=c(1, 1)
+#           ),
 # blank,
@@
 # boxplots
+demo_plt,
 mean_cents_bxplt_nolabel,
@@
 nrow=2, ncol=1,
-heights=c(1, 1.5),
-labels="AUTO"
+heights=c(2, 1),
+labels=c("", "C")
 )
 
-ggsave("figs/figure_5_vertical.jpeg", fig5_vert, width=15, height=9, units="in")
+ggsave("figs/figure_5_vertical.jpeg", fig5_vert, width=17, height=15, units="in")
 ```
@@
 #predict w CIs
-pdat <- with(mean_cent_deltas_df_study,
+pdat1 <- with(mean_cent_deltas_df_study,
              tibble(study_mean_age = seq(min(study_mean_age),
@@
 
-pred_abs_lm <- predict(abs_lm, pdat, se.fit=TRUE)
+pred_abs_lm1 <- predict(abs_lm, pdat1, se.fit=TRUE)
 
-pred_df <- pdat %>% 
-  bind_cols(data.frame(pred_abs_lm))
+pred_df1 <- pdat1 %>% 
+  bind_cols(data.frame(pred_abs_lm1))
 
-pred_df <- pred_df %>%
+pred_df1 <- pred_df1 %>%
   mutate(upper = fit + (1.96 * se.fit), #95% CIs
@@
   geom_ribbon(aes(ymin = lower, ymax = upper,
-                    x = study_mean_age), data = pred_df,
+                    x = study_mean_age), data = pred_df1,
                 inherit.aes = FALSE, alpha = 0.2) +
-    geom_line(data = pred_df, aes(y = fit, x=study_mean_age))+
+    geom_line(data = pred_df1, aes(y = fit, x=study_mean_age))+
   theme_linedraw()+
@@
   labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)")
-ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=10, height=7, units="in")
+#ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=10, height=7, units="in")
+```
+
+mean abs delta ~ study range
+
+```{r}
+range_mod <- lm(abs_mean_delta_cent ~ age_range + study_n, data=mean_cent_deltas_df_study)
+summary(range_mod)
+
+#predict w CIs
+pdat2 <- with(mean_cent_deltas_df_study,
+             tibble(age_range = seq(min(age_range),
+                            max(age_range),
+                            length = 200),
+             study_n = rep(median(study_n), length=200)))
+
+pred_abs_lm2 <- predict(range_mod, pdat2, se.fit=TRUE)
+
+pred_df2 <- pdat2 %>% 
+  bind_cols(data.frame(pred_abs_lm2))
+
+pred_df2 <- pred_df2 %>%
+  mutate(upper = fit + (1.96 * se.fit), #95% CIs
+         lower = fit - (1.96 * se.fit))
+
+range_delta_fig <- mean_cent_deltas_df_study %>%
+  ggplot(aes(x=age_range, y=abs_mean_delta_cent)) +
+  geom_point(aes(color=study), alpha=0.2) +
+  geom_ribbon(aes(ymin = lower, ymax = upper,
+                    x = age_range), data = pred_df2,
+                inherit.aes = FALSE, alpha = 0.2) +
+    geom_line(data = pred_df2, aes(y = fit, x=age_range))+
+  theme_linedraw()+
+  theme(legend.position="none") +
+  labs(x="Age Range of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)")
 ```
@@
 
-Cornell_C2 has very low n=5 -> big differences in estimation
-
 ```{r}
@@
 #predict w CIs
-pdat2 <- with(mean_cent_deltas_df_study,
+pdat3 <- with(mean_cent_deltas_df_study,
              tibble(abs_age_diff = seq(min(abs_age_diff),
@@
 
-pred2 <- predict(abs_lm2, pdat2, se.fit = TRUE)
+pred3 <- predict(abs_lm2, pdat3, se.fit = TRUE)
 
-pdat2 <- pdat2 %>%
-  bind_cols(data.frame(pred2)) %>%
+pdat3 <- pdat3 %>%
+  bind_cols(data.frame(pred3)) %>%
   mutate(upper = fit + (1.96 * se.fit),
@@
   geom_ribbon(aes(ymin = lower, ymax = upper,
-                    x = abs_age_diff), data = pdat2,
-                inherit.aes = FALSE, alpha = 0.2) +
-    geom_line(data = pdat2, aes(y = fit, x=abs_age_diff)) +
+                    x = abs_age_diff), data = pdat3,
+                inherit.aes = FALSE, alpha = 0.4) +
+    geom_line(data = pdat3, aes(y = fit, x=abs_age_diff)) +
   theme_linedraw()+
@@
   labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)")
-ggsave("figs/supp_mean_delta_fig2.jpeg", mean_delta_fig2, width=10, height=7, units="in")
+#ggsave("figs/supp_mean_delta_fig2.jpeg", mean_delta_fig2, width=10, height=7, units="in")
 ```

### 2025-01-31 eda/lifespan_eda.Rmd
remove rain plot

@@
 }
-pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see, broom.mixed, ggrain, ggpubr)
+pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see, broom.mixed, ggpubr)
 
@@
 ```{r}
-full_rain <- final_df %>%
-  mutate(continent = factor(case_when(country == "India" | country == "China" | country == "Singapore" | country == "Japan" ~ "Asia",
-                               country == "USA" | country == "Canada" ~ "North America",
-                               country == "Brazil" ~ "South America",
-                               country == "Australia" ~ "Australia",
-                               TRUE ~ "Europe"))) %>%
-  ggplot(aes(x=continent, y=age_years, fill=sex)) +
-  geom_rain(alpha=.5, violin.args = list(scale="count", alpha=.5),
-            point.args=list(alpha=.2)) +
-  theme_linedraw(base_size = 14) +
-  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        legend.position = "bottom") +
-  guides(fill=guide_legend(title="Sex")) + 
-  coord_fixed(ratio=.08) +
-  labs(y="Age (Years)")
-```
-
-```{r}
 country_leg <- as_ggplot(get_legend(sample_boxplot2))
@@
           )
+saveRDS(fig5A, file="../figs/figure_5_A.rds")
 ```
-
-```{r}
-fig5_AB <- ggarrange(fig5A, full_rain,
-          nrow=1, ncol=2, 
-          #legend = "none",
-          widths=c(1.8, 1),
-          labels = "AUTO")
-fig5_AB
-saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
-ggsave("../figs/fig5_ab.jpeg", fig5_AB, width=18, height=10, units="in")
-```

### 2025-01-31 paper_lbcc_results.Rmd
figures

@@
 ```{r}
-demo_plt <- readRDS("figs/figure_5_AB.rds")
+demo_plt <- readRDS("figs/figure_5_A.rds")
 
@@
 fig5_vert <- ggarrange(
-#   #ggarrange(blank,
-# ggarrange(sample_boxplot2, full_boxplot2,
-#           sample_barplot2, full_barplot2,
-#           nrow=2, ncol=2, 
-#           #align = "v", 
-#           common.legend = TRUE,
-#           legend = "right",
-#           widths=c(18, 1),
-#           heights=c(1, 1)
-#           ),
-# blank,
-# nrow=1, ncol=3,
-# common.legend = FALSE,
-# widths= c(1, 20, 1)),
-# boxplots
 demo_plt,
@@
 nrow=2, ncol=1,
-heights=c(2, 1),
-labels=c("", "C")
+heights=c(1.3, 1),
+labels="AUTO"
 )
-
-ggsave("figs/figure_5_vertical.jpeg", fig5_vert, width=17, height=15, units="in")
+fig5_vert
+ggsave("figs/figure_5_vertical.jpeg", fig5_vert, width=12, height=12, units="in")
 ```
@@
 mean_delta_fig <- mean_cent_deltas_df_study %>%
+  mutate(study = gsub("_", " ", study)) %>%
   ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
@@
   theme(legend.position="none") +
-  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)")
+  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centiles",
+       color="Study")
 #ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=10, height=7, units="in")
@@
 range_delta_fig <- mean_cent_deltas_df_study %>%
+  mutate(study = gsub("_", " ", study)) %>%
   ggplot(aes(x=age_range, y=abs_mean_delta_cent)) +
@@
   theme(legend.position="none") +
-  labs(x="Age Range of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)")
+  labs(x="Age Range of Primary Study Sample (Years)", y="Mean Absolute Difference in Centiles")
 ```
@@
 mean_delta_fig2 <- mean_cent_deltas_df_study %>%
+  mutate(study = gsub("_", " ", study)) %>%
   ggplot(aes(x=abs_age_diff, y=abs_mean_delta_cent)) +
@@
   theme_linedraw()+
-  theme(legend.position="none") +
-  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)")
-#ggsave("figs/supp_mean_delta_fig2.jpeg", mean_delta_fig2, width=10, height=7, units="in")
+  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centiles",
+       color="Study")
+
 ```
+
+```{r}
+sfig_lbcc <- ggarrange(
+mean_delta_fig,
+range_delta_fig,
+mean_delta_fig2,
+nrow=3, ncol=1,
+labels="AUTO",
+legend = "right",
+common.legend=TRUE
+)
+sfig_lbcc
+ggsave("figs/supp_lbcc.jpeg", sfig_lbcc, width=9, height=11, units="in")
+```

### 2025-01-31 cubic_scripts/R_scripts/get_lifespan_deltas.R
troubleshooting

@@ -20,7 +20,7 @@ gamlss.df.path <- list.files(path = csv_path, pattern = paste0("cf.gamlss-", con
 gamlss.df <- fread(gamlss.df.path)
 
 #get identifiers
-id_cols <- c("participant", "sex", "age_days", "site", "log_age", "sexMale", "sex.age", "fs_version")
+id_cols <- c("participant", "sex", "age_days", "log_age", "sexMale", "sex.age", "fs_version")
 
 id.gam.df <- gam.df %>%
   dplyr::select(all_of(id_cols)) %>%

### 2025-01-31 paper_lbcc_results.Rmd

@@
 raw_centile_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 raw_z_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 gam_centile_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25)%>%
   dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 gam_z_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
@@
 gamlss_centile_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!contains(".z")) #drop z-scores and non-centiles & !ends_with("_total") & !contains("TBV")
@@
 gamlss_z_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
-  mutate(age_years=age_days/365.25) %>%
   dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

### 2025-02-02 paper_centilebrain_supp_results.Rmd
look at centilebrain models and plot

--- 2025-01-31 paper_lbcc_results.Rmd
@@
 
-Rmd to compile, test, and visualize ComBatLS vs ComBat-GAM harmonization in the LBCC.
-
-Relies on outputs derived/functions from:
-1. `lifespan_eda.Rmd`
-2. `run_LBCC_pipeline.sh`
-    - `R_scripts/combat_apply_w_transform.R`
-    - `R_scripts/fit_lifespan_mod_notbv.R`
-    - `R_scripts/fit_lifespan_mod_site_est.R`
-3. `run_LBCC_stats.sh`
-    - `R_scripts/fit_lifespan_centiles.R`
-    - `R_scripts/get_lifespan_site_eff.R`
-    - `R_scripts/get_lifespan_deltas.R`
-
-Requires the following inputs from cluster:
-
- - `lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv`
- - `lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv`
- - `lifespan-CN-imp-sites-euler-raw_predictions.csv`
- - `batch.study_pred_deltas.csv`
- - `lifespan-CN-imp-sites-euler-raw_cohenfsq.csv`
- - `lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv`
- - `lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_cohenfsq.csv`
+Rmd to compile, test, and visualize ComBatLS vs ComBat-GAM harmonization in the LBCC with centilebrain models for reviewers
 
@@
 pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize, gratia)
-#Sys.setenv("MC_CORES"=10L)
-options("mc.cores")
-
-source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
-
-#ATLAS
-remotes::install_github("BGDlab/ggsegTissue")
-library(ggsegTissue)
-
-#write out new dk atlas w/o medial wall
-dk_nomed <- dk
-dk_nomed$data <- dk_nomed$data %>%
-  dplyr::filter(!is.na(region) & region !="corpus callosum")
-#plotting info from dk atlas
-dk.parc <- dk_nomed$data %>%
-  as.data.frame() %>%
-  na.omit() %>%
-  dplyr::select(c(hemi, region, label)) %>%
-  distinct()
 
@@
 
-#key to get study names
-study_site_key <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/lifespan_study_site_key.csv")
-
 #PATH
-der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/lifespan"
+der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/cb"
 ```
@@
 ```{r}
-gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_predictions.csv"))
+gam_centile_female <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data-cb-female-data_cb_mod_predictions.csv"))
 
-gam_centile_df <- left_join(gam_centile_df, study_site_key) %>% 
+gam_centile_male <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data-cb-male-data_cb_mod_predictions.csv"))
+
+gam_centile_df <- rbind(gam_centile_female, gam_centile_male)
+stopifnot(nrow(gam_centile_df) == (nrow(gam_centile_male) + nrow(gam_centile_female)))
+
+gam_centile_df <- gam_centile_df %>% 
   mutate(age_years=(age_days-280)/365.25) %>%
   unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE)
+
 n_unique(gam_centile_df$study)
@@
 ```{r}
-gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_predictions.csv"))
-gamlss_centile_df <- full_join(gamlss_centile_df, study_site_key) %>% 
+gamlss_centile_female <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data-cb-female-data_cb_mod_predictions.csv"))
+
+gamlss_centile_male <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data-cb-male-data_cb_mod_predictions.csv"))
+
+gamlss_centile_df <- rbind(gamlss_centile_female, gamlss_centile_male)
+stopifnot(nrow(gamlss_centile_df) == (nrow(gamlss_centile_male) + nrow(gamlss_centile_female)))
+
+gamlss_centile_df <- gamlss_centile_df %>% 
   mutate(age_years=(age_days-280)/365.25) %>%
@@
 ```{r}
-min(gam_centile_df$age_years) == min(gamlss_centile_df$age_years)
-min(gamlss_centile_df$age_years)
-```
-
-Unharmonized Centiles
-```{r}
-raw_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_no.tbv_predictions.csv"))
-raw_centile_df <- full_join(raw_centile_df, study_site_key) %>%
-  unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE) %>%
-  mutate(age_years=(age_days-280)/365.25)
-
-dim(raw_centile_df)
-min(raw_centile_df$age_years) == min(gamlss_centile_df$age_years)
+stopifnot(min(gam_centile_df$age_years) == min(gamlss_centile_df$age_years))
+stopifnot(n_unique(gam_centile_df$INDEX.ID) == n_unique(gamlss_centile_df$INDEX.ID))
 ```
@@
 ```{r}
-deltas_df <- fread(file=paste0(der_path, "/batch.study-data_no.tbv_predictions_pred_deltas.csv"))
-deltas_df <- full_join(deltas_df, study_site_key) %>% 
+deltas_df_m <- fread(file=paste0(der_path, "/batch.study-data-cb-male-data_cb_mod_predictions_pred_deltas.csv"))
+deltas_df_f <- fread(file=paste0(der_path, "/batch.study-data-cb-female-data_cb_mod_predictions_pred_deltas.csv"))
+deltas_df <- rbind(deltas_df_m, deltas_df_f) %>% 
   unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE) %>%
@@
 dim(deltas_df)
-min(deltas_df$age_years) == min(gamlss_centile_df$age_years)
+stopifnot(min(deltas_df$age_years) == min(gamlss_centile_df$age_years))
 ```
@@
 
-#Batch Effects
+plot mean subject centiles
 
-Test to see if centile score is associated with study in either ComBatLS, ComBat-GAM, or non-harmonized data.
+```{r}
+gam_centile_df.mean <- gam_centile_df %>%
+  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(INDEX.ID) %>%
+  summarize(study = unique(study),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_cent=mean(centile))
+stopifnot(nrow(gam_centile_df.mean) == n_unique(gam_centile_df.mean$INDEX.ID))
 
-## Unharmonized
-```{r}
-#clean predictions
-raw_centile_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
-  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
-
-skim(raw_centile_df.clean)
+gamlss_centile_df.mean <- gamlss_centile_df %>%
+  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) %>%
+  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+  group_by(INDEX.ID) %>%
+  summarize(study = unique(study),
+            age_days=unique(age_days),
+            sex=unique(sex),
+            mean_cent=mean(centile))
+stopifnot(nrow(gamlss_centile_df.mean) == n_unique(gamlss_centile_df.mean$INDEX.ID))
 ```
@@
 ```{r}
-raw.feat_to_test <- intersect(pheno_list, names(raw_centile_df.clean))
+gam_centile_df.mean$combat <- "ComBat-GAM"
+gamlss_centile_df.mean$combat <- "ComBatLS"
 
-#see what features didnt converge
-length(raw.feat_to_test)
-setdiff(pheno_list, raw.feat_to_test)
-```
+mean_cents.df <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)
 
-w/in subject means
-
-```{r}
-raw_centile_df.mean <- raw_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(INDEX.ID) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_cent=mean(centile))
-nrow(raw_centile_df.mean) == n_unique(raw_centile_df.mean$INDEX.ID)
-```
-
-```{r}
-raw_centile_df.mean %>%
-  ggplot() +
-  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
-  #geom_boxplot(aes(x=study, y=mean_cent)) +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5))
-
-raw_centile_df.mean %>%
-  ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
-  theme(legend.position = "none") +
-  facet_wrap(~study)
-```
-
-ANOVA
-
-```{r}
-summary(aov(mean_cent ~ study, data=raw_centile_df.mean))
-```
-
-###Z-Scores
-
-```{r}
-#clean predictions
-raw_z_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
-  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
-```
-
-average w/in subject
-
-```{r}
-raw_z_df.mean <- raw_z_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
-  group_by(INDEX.ID) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_z=mean(z_score))
-nrow(raw_z_df.mean) == n_unique(raw_z_df.clean$INDEX.ID)
-```
-
-```{r}
-raw_z_df.mean %>%
-  ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
-```
-
-ANOVA
-```{r}
-summary(aov(mean_z ~ study, data=raw_z_df.mean))
-```
-
-## ComBat-GAM
-
-```{r}
-#clean predictions
-gam_centile_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
-  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
-
-skim(gam_centile_df.clean)
-```
-
-```{r}
-gam.feat_to_test <- intersect(pheno_list, names(gam_centile_df.clean))
-
-#see what features didnt converge
-length(gam.feat_to_test)
-setdiff(pheno_list, gam.feat_to_test)
-```
-
-average w/in subject
-
-```{r}
-gam_centile_df.mean <- gam_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(INDEX.ID) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_cent=mean(centile))
-nrow(gam_centile_df.mean) == n_unique(gam_centile_df.mean$INDEX.ID)
-```
-
-```{r}
-gam_centile_df.mean %>%
-  ggplot() +
-  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
-  #geom_boxplot(aes(x=study, y=mean_cent)) +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5))
-
-gam_centile_df.mean %>%
-  ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
-  theme(legend.position = "none") +
-  facet_wrap(~study)
-```
-
-ANOVA
-
-```{r}
-summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
-```
-
-###Z-Scores
-
-```{r}
-#clean predictions
-gam_z_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
-  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
-```
-
-average w/in subject
-
-```{r}
-gam_z_df.mean <- gam_z_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
-  group_by(INDEX.ID) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_z=mean(z_score))
-nrow(gam_z_df.mean) == n_unique(gam_z_df.clean$INDEX.ID)
-```
-
-```{r}
-gam_z_df.mean %>%
-  ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
-```
-
-ANOVA
-
-```{r}
-summary(aov(mean_z ~ study, data=gam_z_df.mean))
-```
-
-##ComBatLS
-
-```{r}
-#clean predictions
-gamlss_centile_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
-  dplyr::select(!contains(".z")) #drop z-scores and non-centiles & !ends_with("_total") & !contains("TBV")
-```
-
-```{r}
-gamlss.feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))
-
-#see what features didnt converge
-length(gamlss.feat_to_test)
-setdiff(pheno_list, gamlss.feat_to_test)
-```
-
-```{r}
-ggplot(gamlss_centile_df.clean) +
-  geom_point(aes(x=age_days, y=rh_Vol_insula, color=study), alpha=0.5) +
-  theme(legend.position="none")
-```
-
-average w/in subject
-
-```{r}
-gamlss_centile_df.mean <- gamlss_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
-  group_by(INDEX.ID) %>%
-  summarize(study = unique(study),
-            site = unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            log_age=unique(log_age),
-            sexMale=unique(sexMale),
-            fs_version=unique(fs_version),
-            mean_cent=mean(centile))
-nrow(gamlss_centile_df.mean) == n_unique(gamlss_centile_df.mean$INDEX.ID)
-```
-
-```{r}
-gamlss_centile_df.mean %>%
-  ggplot() +
-  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
-  #geom_boxplot(aes(x=study, y=mean_cent)) +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5))
-
-gamlss_centile_df.mean %>%
-  ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_cent, color=study), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_cent, color=study), method="lm") +
-  theme(legend.position = "none")
-
-gamlss_centile_df.mean %>%
-  dplyr::filter(age_days>365) %>%
-  ggplot(aes(x=age_days/365.25, y=mean_cent)) +
-  geom_point(aes(color=study), alpha=0.7) +
-  geom_smooth(method="lm") +
-  theme(legend.position="none")
-```
-
-ANOVA
-```{r}
-summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
-```
-
-###Z-Scores
-
-```{r}
-#clean predictions
-gamlss_z_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
-  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
-```
-
-average w/in subject
-
-```{r}
-gamlss_z_df.mean <- gamlss_z_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
-  group_by(INDEX.ID) %>%
-  summarize(study = unique(study),
-            site=unique(site),
-            age_days=unique(age_days),
-            sex=unique(sex),
-            mean_z=mean(z_score))
-nrow(gamlss_z_df.mean) == n_unique(gamlss_z_df.clean$INDEX.ID)
-```
-
-```{r}
-gamlss_z_df.mean %>%
-  ggplot() +
-  geom_point(aes(x=age_days/365.25, y=mean_z, color=study), alpha=0.5) +
-  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") +
-  facet_wrap(~study) +
-  theme(legend.position="none")
-```
-
-ANOVA
-
-```{r}
-summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
-```
-
-##Comparison Plots
-
-Number of models converged in each dataset
-```{r}
-length(raw.feat_to_test)
-length(gam.feat_to_test)
-length(gamlss.feat_to_test)
-
-#missing features
-c(setdiff(pheno_list, gamlss.feat_to_test),
-setdiff(pheno_list, raw.feat_to_test),
-setdiff(pheno_list, gam.feat_to_test)) %>% unique()
-```
-
-Centile boxplots
-```{r}
-gamlss_centile_df.mean$combat <- "gamlss"
-gam_centile_df.mean$combat <- "gam"
-raw_centile_df.mean$combat <- "raw"
-
-#put all data frames into list
-df_list.tmp <- list(gamlss_centile_df.mean, gam_centile_df.mean, raw_centile_df.mean)
-
-#merge all data frames in list
-mean_cents.df <- df_list.tmp %>% reduce(full_join)
-
-dataset_names <- c(gam = "ComBat-GAM", gamlss = "ComBatLS", raw = "Unharmonized")
-
-mean_cents_bxplt <- mean_cents.df %>%
+mean_cents.df %>%
   ggplot() +
@@
   theme_linedraw() +
-  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none", strip.text.x = element_text(size = 12)) +
-  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
-  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
+  theme(axis.text.x = , legend.position="bottom") +
+  facet_wrap(~sex) +
+  scale_fill_manual(values = c('ComBat-GAM' = "#3D75C2", ComBatLS ="#CFB023FF", Unharmonized = "darkgray"), name = "Harmonization") +
   labs(y="Mean Centile", x = "Study")
@@
   theme_linedraw() +
-  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none", strip.text.x = element_text(size = 12)) +
-  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
-  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="bottom", strip.text.x = element_text(size = 12)) +
+  facet_wrap(~sex) +
+  scale_fill_manual(values = c('ComBat-GAM' = "#3D75C2", ComBatLS ="#CFB023FF", Unharmonized = "darkgray"), name = "Harmonization") +
   labs(y="Mean Centile", x = "Study")
+ggsave("figs/cb_mean_centiles.jpeg", mean_cents_bxplt_nolabel, width=12, height=7, units="in")
 ```
 
-##Cohen's F-squared for study effects
+#Cohen's F-squared for study effects
 
 ```{r}
-raw_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_cohenfsq.csv"))
-raw_centile_f2$combat <- "Unharmonized"
+raw_centile_f2_m <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw-cb-male-data_cb_cohenfsq.csv"))
+raw_centile_f2_m$combat <- "Unharmonized"
+raw_centile_f2_m$sex <- "Male"
+raw_centile_f2_f <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw-cb-female-data_cb_cohenfsq.csv"))
+raw_centile_f2_f$combat <- "Unharmonized"
+raw_centile_f2_f$sex <- "Female"
 
-gam_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv"))
-gam_centile_f2$combat <- "ComBat-GAM"
+gam_centile_f2_f <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data-cb-female-data_cb_cohenfsq.csv"))
+gam_centile_f2_f$combat <- "ComBat-GAM"
+gam_centile_f2_f$sex <- "Female"
+gam_centile_f2_m <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data-cb-male-data_cb_cohenfsq.csv"))
+gam_centile_f2_m$combat <- "ComBat-GAM"
+gam_centile_f2_m$sex <- "Male"
 
-gamlss_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_cohenfsq.csv"))
-gamlss_centile_f2$combat <- "ComBatLS"
+gamlss_centile_f2_f <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data-cb-female-data_cb_cohenfsq.csv"))
+gamlss_centile_f2_f$combat <- "ComBatLS"
+gamlss_centile_f2_f$sex <- "Female"
+gamlss_centile_f2_m <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data-cb-male-data_cb_cohenfsq.csv"))
+gamlss_centile_f2_m$combat <- "ComBatLS"
+gamlss_centile_f2_m$sex <- "Male"
 
-f2.df <- rbind(raw_centile_f2, gam_centile_f2, gamlss_centile_f2)
+f2.df <- rbind(raw_centile_f2_m, raw_centile_f2_f, 
+               gam_centile_f2_m, gam_centile_f2_f,
+               gamlss_centile_f2_m, gamlss_centile_f2_f)
 f2.df <- f2.df %>%
@@
   labs(y="Effect Size (Cohen's F-Squared)", x = "Brain Feature")
-ggsave("figs/cohens_fsq.jpeg", fsq_plt)
+ggsave("figs/cohens_fsq_centilebrain.jpeg", fsq_plt)
 ```
 
-```{r}
-f2.df.sum <- f2.df %>%
-  dplyr::filter(combat != "Unharmonized") %>%
-  group_by(pheno) %>%
-  arrange(combat) %>%
-  mutate(F2_diff = fsq - fsq[1]) #gamlss - gam value, + number means site effect is larger in combatLS
-
-ggplot(f2.df.sum) +
-  geom_col(aes(x=pheno, y=F2_diff), position="dodge") + 
-  facet_wrap(~pheno_cat, scales="free") +
-  theme(axis.text.x = element_blank())
-
-f2.df.sum %>%
-  group_by(pheno_cat) %>%
-  summarise(mean_fsq_diff = mean(F2_diff)) %>%
-  ggplot() + 
-  geom_col(aes(x=pheno_cat, y=mean_fsq_diff), position="dodge")
-
-f2.df.sum %>%
-  #group_by(pheno_cat) %>%
-  mutate(smaller_eff = case_when(F2_diff < 0 ~ "ComBatGAM",
-                                    F2_diff > 0 ~ "ComBatLS",
-                                 F2_diff == 0 ~ "Same")) %>%
-  ggplot() +
-  geom_bar(aes(x=smaller_eff, fill=pheno_cat), position="dodge")
-```
 #ComBat-GAM vs ComBatLS: Delta Centile
@@
 ```{r}
-study_info <- raw_centile_df.clean %>%
-  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
+study_info <- gam_centile_df %>%
   group_by(study) %>%
@@
   summarize(study = unique(study),
-            site = unique(site),
             age_days=unique(age_days),
@@
 ```{r}
-gam_centile_df.clean$combat <- "cf.gam"
-gamlss_centile_df.clean$combat <- "cf.gamlss"
+gam_centile_df$combat <- "cf.gam"
+gamlss_centile_df$combat <- "cf.gamlss"
 
-centiles_df <- full_join(gam_centile_df.clean, gamlss_centile_df.clean)
+centiles_df <- full_join(gam_centile_df, gamlss_centile_df)
 
 centiles_cat <- centiles_df %>%
+  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) %>%
   mutate(combat = factor(combat, ordered=TRUE, levels=c("cf.gam", "cf.gamlss"))) %>%
@@
 
-Plots
-
-```{r}
-raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info) %>%
-  mutate(age_diff = (age_days/365.25) - study_mean_age)
-```
-
-```{r}
-gamlss_centile_df.mean2$combat <- "gamlss"
-gam_centile_df.mean2$combat <- "gam"
-raw_centile_df.mean2$combat <- "raw"
-
-#put all data frames into list
-df_list.tmp <- list(gamlss_centile_df.mean2, gam_centile_df.mean2, raw_centile_df.mean2)
-
-#merge all data frames in list
-all_subj_means.df <- df_list.tmp %>% reduce(full_join)
-
-all_subj_means.df %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
-  geom_point(alpha=0.6) +
-  geom_smooth(method="lm")
-
-all_subj_means.df %>%
-  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
-  geom_point(alpha=0.3) +
-  geom_smooth(method="lm") +
-  facet_wrap(~combat)
-
-summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
-summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
-summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
-```
-
-```{r}
-summary(lm(mean_cent ~ study_n, data=raw_centile_df.mean2))
-summary(lm(mean_cent ~ study_n, data=gam_centile_df.mean2))
-summary(lm(mean_cent ~ study_n, data=gamlss_centile_df.mean2))
-
-all_subj_means.df %>%
-  ggplot(aes(x=study_n, y=mean_cent, color=combat)) +
-  geom_point(alpha=0.6) +
-  geom_smooth(method="lm")
-```
-all assoc with study n
-
-#FIGURE 5
-
-```{r}
-demo_plt <- readRDS("figs/figure_5_A.rds")
-
-# study demographics
-fig5_vert <- ggarrange(
-demo_plt,
-mean_cents_bxplt_nolabel,
-common.legend=FALSE,
-nrow=2, ncol=1,
-heights=c(1.3, 1),
-labels="AUTO"
-)
-fig5_vert
-ggsave("figs/figure_5_vertical.jpeg", fig5_vert, width=12, height=12, units="in")
-```
-
-##Supplemental Figs
+#Supplemental Figs
 
@@
   theme_linedraw()+
-  theme(legend.position="none") +
-  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centiles",
+  theme(legend.position="bottom") +
+  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute\nDifference in Centiles",
        color="Study")
@@
   theme_linedraw()+
-  theme(legend.position="none") +
-  labs(x="Age Range of Primary Study Sample (Years)", y="Mean Absolute Difference in Centiles")
-```
-
-```{r}
-mean_cent_deltas_df_study %>%
-  dplyr::filter(abs_mean_delta_cent > 4) %>%
-  dplyr::select(study, study_mean_age)
-
-mean_cent_deltas_df_study%>%
-  dplyr::filter(study == "Cornell_C2") %>%
-  nrow()
+  theme(legend.position="bottom") +
+  labs(x="Age Range of Primary Study Sample (Years)", y="Mean Absolute\nDifference in Centiles")
 ```
@@
   theme_linedraw()+
-  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centiles",
+  theme(legend.position="bottom") +
+  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute\nDifference in Centiles",
        color="Study")
@@
 ```{r}
-sfig_lbcc <- ggarrange(
+sfig_cb <- ggarrange(
 mean_delta_fig,
@@
 nrow=3, ncol=1,
-labels="AUTO",
+labels=c("B", "C", "D"),
 legend = "right",
@@
 )
-sfig_lbcc
-ggsave("figs/supp_lbcc.jpeg", sfig_lbcc, width=9, height=11, units="in")
+
+sfig_cb2 <- ggarrange(
+fsq_plt,
+sfig_cb,
+nrow=2, ncol=1,
+labels=c("A", ""),
+heights=c(1.4, 2),
+legend = "bottom",
+common.legend=FALSE
+)
+
+ggsave("figs/supp_cb.jpeg", sfig_cb2, width=10, height=12, units="in")
 ```
